### 2008-11-05 teoria_algoritmos/examen1.tex
commit inicial (cosas de teoria de algoritmos)

\documentclass{article}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage[spanish]{babel}

% Para codigo bonito
\newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}

\newcommand{\algoritmo}[2]
{
{
  \singlespace
  \begin{algorithm}[h]
    \caption{#1}
    \begin{algorithmic}[1]
      #2
    \end{algorithmic}
  \end{algorithm}
}
}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\lstset{mathescape=true}

\begin{document}

% Título

\title{Tarea 1 de Teoría de Algoritmos}
\author{Daniel Izquierdo \\ \#08-86809}
\date{9 de octubre de 2008}

\maketitle

% Pregunta 1
\section{}

\begin{align*}
 & (n+\alpha)^{n+\beta} \\
 = & \\
 & n^{n+\beta} * (1+\frac{\alpha}{n})^{n+\beta} \\
 = & \\
 & n^{n+\beta} * e^{lg (1+\frac{\alpha}{n})^{n+\beta}} \\
 = & \\
 & n^{n+\beta} * e^{(n+\beta) * lg (1+\frac{\alpha}{n})} \\
 = & \\
 & n^{n+\beta} * e^{(n+\beta) * (\frac{\alpha}{n} - \frac{\alpha^2}{2n^2}+O(n^{-3})}) \\
 = & \\
 & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+n*O(n^{-3})+\frac{\alpha\beta}{n}-\frac{\alpha^2\beta}{2n^2}+\beta*O(n^{-3})} \\
 = & \\
 & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+O(n^{-2})+\frac{\alpha\beta}{n}-\frac{\alpha^2\beta}{2n^2}+O(n^{-3})} \\
 = & \\
 & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+O(n^{-2})+\frac{\alpha\beta}{n}-\frac{\alpha^2\beta}{2n^2}} \\
 = & \\
 & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+\frac{\alpha\beta}{n}+O(n^{-2})} \\
 = & \\
 & n^{n+\beta} * e^{\alpha} * e^{-\frac{\alpha^2}{2n}+\frac{\alpha\beta}{n}+O(n^{-2})} \\
 = & \\
 & n^{n+\beta} * e^{\alpha} * (1-\frac{\alpha^2}{2n}+\frac{\alpha\beta}{n}+O(n^{-2}) + O(n^2)) \\
 = & \\
 & n^{n+\beta} * e^{\alpha} * (1+\alpha(\beta-\frac{1}{2}\alpha)n^{-1}+O(n^{2})) \\
\end{align*}

% Pregunta 2
\section{}

La longitud de la representación de un entero en binario o decimal es logarítmica con
respecto a su magnitud, pero en palabras de un computador es la longitud entre el tamaño
de la palabra. Es este factor el que hace la diferencia al usar algoritmos exponenciales.

% Pregunta 3
\section{}

Para el análisis del algoritmo descrito en el libro tomo la suma

\begin{equation}
 C(n-1,k-1) + C(n-1,k)
\end{equation}

como barómetro. En todos los casos excepto los base se realiza esa suma.
La función $t(n)$ a calcular se define de manera recursiva.
En el caso en que $n = 0$, obligatoriamente se mantiene que $k = 0$ ya que
tenemos $0 \leq k \leq n$. El algoritmo retorna una respuesta sin realizar otro
cálculo. En caso contrario, el algoritmo realiza dos llamadas recursivas que
toman --en el peor caso-- tiempo $t(n-1)$. Luego realiza una suma
en tiempo constante. Entonces,

$$
 t(n) = \left\{ \begin{array}{ll}
 0 & \textrm{si $n = 0$} \\
 2*t(n-1)+1 & \textrm{si $n > 0$}
 \end{array} \right.
$$

La solución a esta recurrencia es dada en el ejemplo {\bf 4.7.6} del libro, y es

\begin{equation}
 t(n) = 2^n-1
\end{equation}

Por lo tanto,

\begin{equation}
t(n) = \Theta(2^n)
\end{equation}

% Pregunta 4
\section{}

Decir que $q$ es el cociente de la divisón entera de $A$ entre $B$ y $r$ es el
resto de la división entera de $A$ entre $B$ es equivalente a decir que:

\begin{align*}
 A = q*B + r \wedge 0 \leq r < b
\end{align*}

La última instrucción del programa es una asignación. Tenemos:

\begin{align*}
 & wp.(r \leftarrow A-q*B).(A=q*B+r \wedge 0 \leq r < B) \\
 \equiv & <\text{precondición más débil de la asignación}> \\
 & (A=q*B+r \wedge 0 \leq r < b)[r := A-q*B] \\
 \equiv & <\text{sustitución}> \\
 & A=q*B+(A-q*B) \wedge 0 \leq (A-q*B) < B \\
 \equiv & <\text{aritmética}> \\
 & true \wedge 0 \leq (A-q*B) < B \\
 \equiv & <\text{true es neutro de } \wedge> \\
 & 0 \leq (A-q*B) < B \\
 \equiv & <\text{definición de } \leq \text{ y } <> \\
 & 0 \leq A-q*B \wedge A-q*B < B
\end{align*}

Llamemos $P$ a este resultado. Entonces $P$ es la postcondición que debe ser
satisfecha por el ciclo \texttt{while} para
que el programa sea correcto. Si defino la invariante $Inv$ y la cota $Bound$
del ciclo

\begin{eqnarray*}
 Inv & \equiv & A - q*B \geq 0 \\
 Bound & \equiv & A-q*B \\
\end{eqnarray*}

y tenemos el guardia $B$ definido

\begin{align*}
 B \equiv A-(q+1)*B \geq 0
\end{align*}

entonces debo demostrar lo siguiente, suponiendo que $X$ es alguna constante y
$A$ y $B$ son enteros:

\begin{eqnarray}
 \label{divmod:init} \text{\{} A \geq 0 \wedge B > 0 \text{\}} q \leftarrow 0 \text{\{}Inv\text{\}} & & \\
 \label{divmod:inv} \text{\{} Inv \wedge B \text{\}} q \leftarrow q + 1 \text{\{}Inv\text{\}} & & \\
 \label{divmod:final} Inv \wedge \neg B \Rightarrow P & & \\
 \label{divmod:bounddecr} \text{\{} Inv \wedge X = Bound \text{\}} q \leftarrow q + 1 \text{\{} Bound < X \text{\}} & & \\
 \label{divmod:boundnn} Inv \Rightarrow Bound \geq 0 & & 
\end{eqnarray}

Empiezo con el predicado \ref{divmod:init}:

\begin{align*}
 & wp.(q \leftarrow 0).Inv \\
 \equiv & <\text{precondición más débil de la asignación}> \\
 & Inv[q := 0] \\
 \equiv & <\text{definición de } Inv \text{ y sustitución}> \\
 & A - 0*B \geq 0 \\
 \equiv & <\text{aritmética}> \\
 & A \geq 0 \\
 \Leftarrow & <P \wedge Q \Rightarrow P> \\
 & A \geq 0 \wedge B > 0 \\
 \therefore & <\text{relación entre tripletas de Hoare y precondición más débil}> \\
 & \text{\{} A \geq 0 \wedge B > 0 \text{\}} q \leftarrow 0 \text{\{}Inv\text{\}}
\end{align*}

Luego demuestro el predicado \ref{divmod:inv}:

\begin{align*}
 & wp.(q \leftarrow q + 1).Inv \\
 \equiv & <\text{precondición más débil de la asignación}> \\
 & Inv[q := q + 1] \\
 \equiv & <\text{definición de } Inv \text{ y sustitución}> \\
 & A-(q+1)*B \geq 0 \\
 \equiv & <\text{definición de } B> \\
 & B \\
 \Leftarrow & <P \wedge Q \Rightarrow P> \\
 & Inv \wedge B \\
 \therefore & <\text{relación entre tripletas de Hoare y precondición más débil}> \\
 & \text{\{} Inv \wedge B \text{\}} q \leftarrow q + 1 \text{\{}Inv\text{\}}
\end{align*}

Continúo con el predicado \ref{divmod:final}:

\begin{align*}
 & Inv \wedge \neg B \\
 \equiv & <\text{definiciones de } Inv \text{ y de } B \text{ y aritmética}> \\
 & A - q*B \geq 0 \wedge \neg (A \geq (q+1)*B) \\
 \equiv & <\text{negación de } \geq> \\
 & A - q*B \geq 0 \wedge A < (q+1)*B \\
 \equiv & <\text{aritmética}> \\
 & A - q*B \geq 0 \wedge A - q*B < B \\
 \equiv & <\text{definición de P}> \\
 & P
\end{align*}

Queda comprobar la terminación del algoritmo. Por una parte,
si suponemos $Inv \wedge X = Bound = A-q*B$, entonces

\begin{align*}
 & wp.(q \leftarrow q + 1).(Bound < X) \\
 \equiv & <\text{precondición más débil de la asignación}> \\
 & (Bound < X)[q := q + 1] \\
 \equiv & <\text{definición de Bound y sustitución}> \\
 & A-(q+1)*B < X \\
 \Leftarrow & <\text{aritmética y suposición } X = Bound = A-q*B> \\
 & A-q*B-B < A-q*B \\
 \equiv & <\text{aritmética}> \\
 & 0 < B \\
 \Leftarrow & <\text{suposición } Inv> \\
 & true
\end{align*}

con lo que queda demostrado el predicado \ref{divmod:bounddecr}. El predicado
\ref{divmod:boundnn} sigue trivialmente a partir de las definiciones de $Inv$ y
$Bound$.

% Pregunta 5
\section{}

Llamaré \texttt{convertir} al algoritmo que presento para convertir un arreglo 
en un heap binario. El algoritmo utiliza un procedimiento auxiliar llamado
\texttt{heapify}. Es igual al descrito en el ``Introduction to Algorithms'' de
Cormen y otros. Suponemos que $tamano(A)$ es el tamaño del heap $A$, y
escribimos

\algoritmo{heapify(A,i)}
{
\STATE $l \leftarrow 2*i$
\STATE $r \leftarrow 2*i+1$
\IF {$l \leq tamano(A) \wedge A[l] < A[i]$}
\STATE $maspequeno \leftarrow l$
\ELSE
\STATE $maspequeno \leftarrow i$
\ENDIF
\IF {$r \leq tamano(A) \wedge A[r] < A[maspequeno]$}
\STATE $maspequeno \leftarrow r$
\ENDIF
\IF {$maspequeno \neq i$}
\STATE intercambiar A[i] con A[masgrande]
\STATE heapify(A, masgrande)
\ENDIF
}

\algoritmo{convertir(A)}
{
\STATE $tamano(A) \leftarrow longitud [A]$
\FOR {$i \leftarrow piso(longitud[A]/2)$ hasta $1$}
\STATE heapify(A, i)
\ENDFOR
}

El procedimiento \texttt{heapify} toma un nodo del árbol formado en el arreglo
y tiene como precondición que sus dos hijos sean raíces de un heap. Ejecutar el
procedimiento tiene como resultado que el valor del nodo dado ``descienda''
por el árbol
hasta una posición tal que se preserve la propiedad de heap. Si el valor es
mayor
que el menor de sus hijos, se intercambian sus posiciones y se continúa sobre
la posición del nodo que está descendiendo. Al final, el árbol termina
cumpliendo la propiedad de heap a partir del nodo dado.

Los
elementos en la segunda mitad del arreglo dado representan árboles de un sólo
nodo, por lo que cumplen la propiedad del heap.
\texttt{convertir} trabaja sobre la primera mitad del arreglo de derecha a
izquierda, llamando a \texttt{heapify} sobre cada nodo visto. \texttt{heapify}
funciona porque cada nodo a la derecha del nodo que se está viendo ya es un heap
por llamadas previas a \texttt{heapify}. Al final, hacer \texttt{heapify} sobre
el primer elemento del arreglo asegura que esté organizado como un heap.

Definimos la altura de un nodo de un heap como el número de lados a recorrer
desde ese nodo hasta la hoja del heap más cercana y la altura del heap como la
altura de su raíz. Se puede demostrar que la altura de un heap con $n$ elementos
tiene altura $\lfloor lg n\rfloor$ y que un heap cualquiera tiene como máximo
$\lceil n / 2^{h+1} \rceil$ nodos con altura $h$. Por otra parte,
\texttt{heapify} toma tiempo $O(h)$ para trabajar sobre un árbol de altura $h$.
Además, \texttt{convertir} realiza $n/2 = O(n)$ llamadas a \texttt{heapify}
sobre un arreglo de tamaño $n$.

Por todo esto, el tiempo de ejecución de \texttt{convertir} está acotado por la
siguiente fórmula suponiendo que $n$ es potencia de $2$:

\begin{align*}
 & \sum_{h=0}^{lg n} \frac{n}{2^{h+1}} O(h) \\
 = & <\text{aritmética, propiedades de O(f(n))}> \\
 & O(n \sum_{h=0}^{lg n} \frac{h}{2^{h}})
\end{align*}

Usando la ecuación {\bf A.8} del Cormen, tenemos

\begin{align*}
 \sum_{h=0}^{\infty} \frac{h}{2^h} = 2 \\
\end{align*}

Esto acota el tiempo de ejecución de \texttt{convertir}:

\begin{align*}
 O(n \sum_{h=0}^{lg n} \frac{h}{2^{h}}) &= O(n \sum_{h=0}^{\infty} \frac{h}{2^h}) \\
 &= O(n) \\
\end{align*}

Sabemos que $n$ es uniforme. Además,

\begin{equation}
 \sum_{h=0}^{lg n} \frac{n}{2^{h+1}} O(h) \\
\end{equation}

es no decreciente sobre $n$ porque es una suma de términos positivos cuyo
rango aumenta al aumentar $n$. Entonces, por la regla de la suavidad, el tiempo
de ejecución de \texttt{convertir} es $O(n)$.

% Pregunta 6
\section{}

La estrucura de datos cola tiene dos operaciones: \texttt{encolar} y
\texttt{desencolar}. Para implementar una cola con dos pilas $S_1$ y $S_2$,
añado una operación auxiliar \texttt{mudar}. Para una cola $Q$, voy a llamar a
sus dos pilas $Q.S_1$ y $Q.S_2$. Supongo que nunca se ejecuta
\texttt{desencolar} sobre una cola vacía (lo cual es un error), y defino
entonces:

\begin{lstlisting}
encolar(Q, x) = empilar(Q.$S_1$, x)

desencolar(Q) = si vacia(Q.$S_2$)
                    mudar(Q.$S_1$, Q.$S_2$)
                return desempilar(Q.$S_2$)

mudar(Q.$S_1$, Q.$S_2$) = mientras $\neg$vacia(Q.$S_1$)
                              x $\leftarrow$ desempilar(Q.$S_1$)
                              empilar(Q.$S_2$, x)
\end{lstlisting}

Ahora hago un análisis del costo en tiempo amortizado de las operaciones
\texttt{encolar} y \texttt{desencolar} para verificar que ambas son $O(1)$. Para
ello utilizo el método del potencial. Defino la función potencial $\Phi$ para
una cola $Q$ como

\[\Phi(Q) = 3 * \#(Q.S_1) + \#(Q.S_2)\]

donde

\[\#(S)\]

es el número de elementos en la pila $S$. Es claro que, como el número de
elementos en una pila es no negativo, $\Phi$ es no negativo. Entonces, si $Q_0$
es una pila recién creada y $Q_i$ es la misma pila luego de haberle aplicado
alguna secuencia de operaciones, tenemos

\begin{align*}
\Phi(Q_i) &\geq 0 \\
          &= \Phi(Q_0)
\end{align*}

Así, $\Phi$ cumple esa propiedad necesaria del método del potencial. Ahora
calculo el costo amortizado $\hat{c}_i$ de las operaciones que definí:

\begin{itemize}
 \item \texttt{encolar($Q$, $x$)}:
 Realizamos una operación de inserción en una pila y el potencial de la
 estructura aumenta en 3.
 \begin{align*}
  \hat{c}_{i+1} &= c_{i+1} + \Phi(D_{i+1}) - \Phi(D_i) \\
          &= 1 + (3 + \Phi(D_i)) - \Phi(D_i) \\
          &= 1 + 3 \\
          &= 4
 \end{align*}

 \item \texttt{mudar($Q.S_1$, $Q.S_2$)}:
 Si $Q.S_1$ tiene $k$ elementos, entonces realizamos $k$ operaciones de
 desempilar sobre $Q.S_1$ y $k$ operaciones de empilar sobre $Q.S_2$. Eliminar
 $k$ elementos de $Q.S_1$ hace que el potencial disminuya en $3*k$ y agregarlos
 a $Q.S_2$ hace que aumente en $k$. Entonces,
 \begin{align*}
  \hat{c}_{i+1} &= c_{i+1} + \Phi(D_{i+1}) - \Phi(D_i) \\
                &= k + k + \Phi(D_{i+1}) - \Phi(D_i) \\
                &= 2*k + (k - 3*k + \Phi(D_i)) - \Phi(D_i) \\
                &= 0
 \end{align*}

 \item \texttt{desencolar($Q$)}: Si $Q.S_2$ no tiene elementos, realizamos la
 la operación \texttt{mudar} --cuyo costo amortizado es $0$-- y estamos en el
 caso en que sí tiene. Cuando $Q.S_2$ tiene al menos un elemento, realizamos una
 operación de \texttt{desempilar} y el potencial de la estructura disminuye en
 $1$.
 \begin{align*}
  \hat{c}_{i+1} &= c_{i+1} + \Phi(D_{i+1}) - \Phi(D_i) \\
                &= 1 + (\Phi(D_i) -1) - \Phi(D_i) \\
                &= 0
 \end{align*}

\end{itemize}

En conclusión, tenemos una estructura de datos cola implementada con dos pilas
y cuyas operaciones tienen costo amortizado $O(1)$.

\end{document}

### 2008-11-05 teoria_algoritmos/recuperacion/recuperacion.tex
commit inicial (cosas de teoria de algoritmos)

\documentclass{article}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage[spanish]{babel}

% Para codigo bonito
\newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}

\newcommand{\algoritmo}[2]
{
{
  \singlespace
  \begin{algorithm}[h]
    \caption{#1}
    \begin{algorithmic}[1]
      #2
    \end{algorithmic}
  \end{algorithm}
}
}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\lstset{mathescape=true}

\begin{document}

% Título

\title{Tarea de recuperación de Teoría de Algoritmos}
\author{Daniel Izquierdo \\ \#08-86809}
\date{23 de octubre de 2008}

\maketitle

% Pregunta 1
%\section{}

Para demostrar que $(C,\Gamma)$ es una matroide basta demostrar que

\begin{itemize}
 \item $C$ es un conjunto finito.
 \item $\Gamma$ es un conjunto de subconjuntos de C hereditario. Es decir,

 $$
 (x \in \Gamma \Rightarrow x \subseteq C) \wedge (B \in \Gamma \wedge A \subseteq B \Rightarrow A \in \Gamma)
 $$

 \item Si $A \in \Gamma$, $B \in \Gamma$ y $|A| < |B|$, entonces existe $x \in B - A$ tal que
 $A \bigcup \{x\} \in \Gamma$.
\end{itemize}

Las dos primeras condiciones siguen de la definición de una estructura combinatoria
vista en clases. %Para demostrar que la tercera condición se cumple, voy a suponer al
%absurdo que no se cumple y llegar a una contradicción.

Supongamos al absurdo que la tercera condición no se cumple.
Si se cumple lo descrito en el enunciado, entonces si tenemos una función de costos $c$ tal que

$$
x \in C \Rightarrow c(x) = -1
$$

tendremos al final del algoritmo que $F$ es el conjunto de menor costo en $\Gamma$.
Sean $A, B$ conjuntos tales que $A \in \Gamma$, $B \in \Gamma$ y $|A| < |B|$.
$A \bigcup B$ debe pertenecer
a $\Gamma$ porque $F$ es el conjunto de menor costo en $\Gamma$ y, por lo tanto, contiene
los elementos de $A$ y $B$.
Todos los subconjuntos de $A \bigcup B$ también deben pertenecer a $\Gamma$. Llamemos $D$ al conjunto resultante de
quitar de $A \bigcup B$ todos los elementos de $B$ menos uno (se puede porque $|B| > |A| \geq 0$). Entonces $D$ es subconjunto
de $A \bigcup B$ y pertenece a $\Gamma$. Además, $D = A \bigcup \{x\}$ para algún $x \in B$.

Esto contradice la suposición inicial de que la tercera condición no se cumple. En conclusión, $(C,\Gamma)$ es una 
matroide.

%Como todos los elementos de C
%tienen el mismo costo negativo, la elección del menor elemento de $C^*$ en el algoritmo
%del enunciado es no
%determinística. Luego de $|A|$ iteraciones en el algoritmo, es posible tener que $F = A$
%ya que $A \in \Gamma$ y el algoritmo puede haber elegido únicamente elementos de A hasta ese
%momento. Como $c(A)>c(B)$, el algoritmo no puede haber terminado ya que $A$ no es el elemento
%de menor costo en $\Gamma$ (porque por lo menos $B$ tiene menor costo). Entonces va a agregar
%algún otro elemento de $C^*$.

\end{document}

### 2008-11-05 teoria_algoritmos/examen2/examen2.tex
limpieza del .tex de la tarea 2 de teoria de algoritmos

--- 2008-11-05 teoria_algoritmos/examen1.tex
@@
 
-\title{Tarea 1 de Teoría de Algoritmos}
+\title{Tarea 2 de Teoría de Algoritmos}
 \author{Daniel Izquierdo \\ \#08-86809}
-\date{9 de octubre de 2008}
+\date{6 de noviembre de 2008}
 
@@
 
-\begin{align*}
- & (n+\alpha)^{n+\beta} \\
- = & \\
- & n^{n+\beta} * (1+\frac{\alpha}{n})^{n+\beta} \\
- = & \\
- & n^{n+\beta} * e^{lg (1+\frac{\alpha}{n})^{n+\beta}} \\
- = & \\
- & n^{n+\beta} * e^{(n+\beta) * lg (1+\frac{\alpha}{n})} \\
- = & \\
- & n^{n+\beta} * e^{(n+\beta) * (\frac{\alpha}{n} - \frac{\alpha^2}{2n^2}+O(n^{-3})}) \\
- = & \\
- & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+n*O(n^{-3})+\frac{\alpha\beta}{n}-\frac{\alpha^2\beta}{2n^2}+\beta*O(n^{-3})} \\
- = & \\
- & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+O(n^{-2})+\frac{\alpha\beta}{n}-\frac{\alpha^2\beta}{2n^2}+O(n^{-3})} \\
- = & \\
- & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+O(n^{-2})+\frac{\alpha\beta}{n}-\frac{\alpha^2\beta}{2n^2}} \\
- = & \\
- & n^{n+\beta} * e^{\alpha-\frac{\alpha^2}{2n}+\frac{\alpha\beta}{n}+O(n^{-2})} \\
- = & \\
- & n^{n+\beta} * e^{\alpha} * e^{-\frac{\alpha^2}{2n}+\frac{\alpha\beta}{n}+O(n^{-2})} \\
- = & \\
- & n^{n+\beta} * e^{\alpha} * (1-\frac{\alpha^2}{2n}+\frac{\alpha\beta}{n}+O(n^{-2}) + O(n^2)) \\
- = & \\
- & n^{n+\beta} * e^{\alpha} * (1+\alpha(\beta-\frac{1}{2}\alpha)n^{-1}+O(n^{2})) \\
-\end{align*}
-
-% Pregunta 2
-\section{}
-
-La longitud de la representación de un entero en binario o decimal es logarítmica con
-respecto a su magnitud, pero en palabras de un computador es la longitud entre el tamaño
-de la palabra. Es este factor el que hace la diferencia al usar algoritmos exponenciales.
-
-% Pregunta 3
-\section{}
-
-Para el análisis del algoritmo descrito en el libro tomo la suma
-
-\begin{equation}
- C(n-1,k-1) + C(n-1,k)
-\end{equation}
-
-como barómetro. En todos los casos excepto los base se realiza esa suma.
-La función $t(n)$ a calcular se define de manera recursiva.
-En el caso en que $n = 0$, obligatoriamente se mantiene que $k = 0$ ya que
-tenemos $0 \leq k \leq n$. El algoritmo retorna una respuesta sin realizar otro
-cálculo. En caso contrario, el algoritmo realiza dos llamadas recursivas que
-toman --en el peor caso-- tiempo $t(n-1)$. Luego realiza una suma
-en tiempo constante. Entonces,
-
-$$
- t(n) = \left\{ \begin{array}{ll}
- 0 & \textrm{si $n = 0$} \\
- 2*t(n-1)+1 & \textrm{si $n > 0$}
- \end{array} \right.
-$$
-
-La solución a esta recurrencia es dada en el ejemplo {\bf 4.7.6} del libro, y es
-
-\begin{equation}
- t(n) = 2^n-1
-\end{equation}
-
-Por lo tanto,
-
-\begin{equation}
-t(n) = \Theta(2^n)
-\end{equation}
-
-% Pregunta 4
-\section{}
-
-Decir que $q$ es el cociente de la divisón entera de $A$ entre $B$ y $r$ es el
-resto de la división entera de $A$ entre $B$ es equivalente a decir que:
-
-\begin{align*}
- A = q*B + r \wedge 0 \leq r < b
-\end{align*}
-
-La última instrucción del programa es una asignación. Tenemos:
-
-\begin{align*}
- & wp.(r \leftarrow A-q*B).(A=q*B+r \wedge 0 \leq r < B) \\
- \equiv & <\text{precondición más débil de la asignación}> \\
- & (A=q*B+r \wedge 0 \leq r < b)[r := A-q*B] \\
- \equiv & <\text{sustitución}> \\
- & A=q*B+(A-q*B) \wedge 0 \leq (A-q*B) < B \\
- \equiv & <\text{aritmética}> \\
- & true \wedge 0 \leq (A-q*B) < B \\
- \equiv & <\text{true es neutro de } \wedge> \\
- & 0 \leq (A-q*B) < B \\
- \equiv & <\text{definición de } \leq \text{ y } <> \\
- & 0 \leq A-q*B \wedge A-q*B < B
-\end{align*}
-
-Llamemos $P$ a este resultado. Entonces $P$ es la postcondición que debe ser
-satisfecha por el ciclo \texttt{while} para
-que el programa sea correcto. Si defino la invariante $Inv$ y la cota $Bound$
-del ciclo
-
-\begin{eqnarray*}
- Inv & \equiv & A - q*B \geq 0 \\
- Bound & \equiv & A-q*B \\
-\end{eqnarray*}
-
-y tenemos el guardia $B$ definido
-
-\begin{align*}
- B \equiv A-(q+1)*B \geq 0
-\end{align*}
-
-entonces debo demostrar lo siguiente, suponiendo que $X$ es alguna constante y
-$A$ y $B$ son enteros:
-
-\begin{eqnarray}
- \label{divmod:init} \text{\{} A \geq 0 \wedge B > 0 \text{\}} q \leftarrow 0 \text{\{}Inv\text{\}} & & \\
- \label{divmod:inv} \text{\{} Inv \wedge B \text{\}} q \leftarrow q + 1 \text{\{}Inv\text{\}} & & \\
- \label{divmod:final} Inv \wedge \neg B \Rightarrow P & & \\
- \label{divmod:bounddecr} \text{\{} Inv \wedge X = Bound \text{\}} q \leftarrow q + 1 \text{\{} Bound < X \text{\}} & & \\
- \label{divmod:boundnn} Inv \Rightarrow Bound \geq 0 & & 
-\end{eqnarray}
-
-Empiezo con el predicado \ref{divmod:init}:
-
-\begin{align*}
- & wp.(q \leftarrow 0).Inv \\
- \equiv & <\text{precondición más débil de la asignación}> \\
- & Inv[q := 0] \\
- \equiv & <\text{definición de } Inv \text{ y sustitución}> \\
- & A - 0*B \geq 0 \\
- \equiv & <\text{aritmética}> \\
- & A \geq 0 \\
- \Leftarrow & <P \wedge Q \Rightarrow P> \\
- & A \geq 0 \wedge B > 0 \\
- \therefore & <\text{relación entre tripletas de Hoare y precondición más débil}> \\
- & \text{\{} A \geq 0 \wedge B > 0 \text{\}} q \leftarrow 0 \text{\{}Inv\text{\}}
-\end{align*}
-
-Luego demuestro el predicado \ref{divmod:inv}:
-
-\begin{align*}
- & wp.(q \leftarrow q + 1).Inv \\
- \equiv & <\text{precondición más débil de la asignación}> \\
- & Inv[q := q + 1] \\
- \equiv & <\text{definición de } Inv \text{ y sustitución}> \\
- & A-(q+1)*B \geq 0 \\
- \equiv & <\text{definición de } B> \\
- & B \\
- \Leftarrow & <P \wedge Q \Rightarrow P> \\
- & Inv \wedge B \\
- \therefore & <\text{relación entre tripletas de Hoare y precondición más débil}> \\
- & \text{\{} Inv \wedge B \text{\}} q \leftarrow q + 1 \text{\{}Inv\text{\}}
-\end{align*}
-
-Continúo con el predicado \ref{divmod:final}:
-
-\begin{align*}
- & Inv \wedge \neg B \\
- \equiv & <\text{definiciones de } Inv \text{ y de } B \text{ y aritmética}> \\
- & A - q*B \geq 0 \wedge \neg (A \geq (q+1)*B) \\
- \equiv & <\text{negación de } \geq> \\
- & A - q*B \geq 0 \wedge A < (q+1)*B \\
- \equiv & <\text{aritmética}> \\
- & A - q*B \geq 0 \wedge A - q*B < B \\
- \equiv & <\text{definición de P}> \\
- & P
-\end{align*}
-
-Queda comprobar la terminación del algoritmo. Por una parte,
-si suponemos $Inv \wedge X = Bound = A-q*B$, entonces
-
-\begin{align*}
- & wp.(q \leftarrow q + 1).(Bound < X) \\
- \equiv & <\text{precondición más débil de la asignación}> \\
- & (Bound < X)[q := q + 1] \\
- \equiv & <\text{definición de Bound y sustitución}> \\
- & A-(q+1)*B < X \\
- \Leftarrow & <\text{aritmética y suposición } X = Bound = A-q*B> \\
- & A-q*B-B < A-q*B \\
- \equiv & <\text{aritmética}> \\
- & 0 < B \\
- \Leftarrow & <\text{suposición } Inv> \\
- & true
-\end{align*}
-
-con lo que queda demostrado el predicado \ref{divmod:bounddecr}. El predicado
-\ref{divmod:boundnn} sigue trivialmente a partir de las definiciones de $Inv$ y
-$Bound$.
-
-% Pregunta 5
-\section{}
-
-Llamaré \texttt{convertir} al algoritmo que presento para convertir un arreglo 
-en un heap binario. El algoritmo utiliza un procedimiento auxiliar llamado
-\texttt{heapify}. Es igual al descrito en el ``Introduction to Algorithms'' de
-Cormen y otros. Suponemos que $tamano(A)$ es el tamaño del heap $A$, y
-escribimos
-
-\algoritmo{heapify(A,i)}
-{
-\STATE $l \leftarrow 2*i$
-\STATE $r \leftarrow 2*i+1$
-\IF {$l \leq tamano(A) \wedge A[l] < A[i]$}
-\STATE $maspequeno \leftarrow l$
-\ELSE
-\STATE $maspequeno \leftarrow i$
-\ENDIF
-\IF {$r \leq tamano(A) \wedge A[r] < A[maspequeno]$}
-\STATE $maspequeno \leftarrow r$
-\ENDIF
-\IF {$maspequeno \neq i$}
-\STATE intercambiar A[i] con A[masgrande]
-\STATE heapify(A, masgrande)
-\ENDIF
-}
-
-\algoritmo{convertir(A)}
-{
-\STATE $tamano(A) \leftarrow longitud [A]$
-\FOR {$i \leftarrow piso(longitud[A]/2)$ hasta $1$}
-\STATE heapify(A, i)
-\ENDFOR
-}
-
-El procedimiento \texttt{heapify} toma un nodo del árbol formado en el arreglo
-y tiene como precondición que sus dos hijos sean raíces de un heap. Ejecutar el
-procedimiento tiene como resultado que el valor del nodo dado ``descienda''
-por el árbol
-hasta una posición tal que se preserve la propiedad de heap. Si el valor es
-mayor
-que el menor de sus hijos, se intercambian sus posiciones y se continúa sobre
-la posición del nodo que está descendiendo. Al final, el árbol termina
-cumpliendo la propiedad de heap a partir del nodo dado.
-
-Los
-elementos en la segunda mitad del arreglo dado representan árboles de un sólo
-nodo, por lo que cumplen la propiedad del heap.
-\texttt{convertir} trabaja sobre la primera mitad del arreglo de derecha a
-izquierda, llamando a \texttt{heapify} sobre cada nodo visto. \texttt{heapify}
-funciona porque cada nodo a la derecha del nodo que se está viendo ya es un heap
-por llamadas previas a \texttt{heapify}. Al final, hacer \texttt{heapify} sobre
-el primer elemento del arreglo asegura que esté organizado como un heap.
-
-Definimos la altura de un nodo de un heap como el número de lados a recorrer
-desde ese nodo hasta la hoja del heap más cercana y la altura del heap como la
-altura de su raíz. Se puede demostrar que la altura de un heap con $n$ elementos
-tiene altura $\lfloor lg n\rfloor$ y que un heap cualquiera tiene como máximo
-$\lceil n / 2^{h+1} \rceil$ nodos con altura $h$. Por otra parte,
-\texttt{heapify} toma tiempo $O(h)$ para trabajar sobre un árbol de altura $h$.
-Además, \texttt{convertir} realiza $n/2 = O(n)$ llamadas a \texttt{heapify}
-sobre un arreglo de tamaño $n$.
-
-Por todo esto, el tiempo de ejecución de \texttt{convertir} está acotado por la
-siguiente fórmula suponiendo que $n$ es potencia de $2$:
-
-\begin{align*}
- & \sum_{h=0}^{lg n} \frac{n}{2^{h+1}} O(h) \\
- = & <\text{aritmética, propiedades de O(f(n))}> \\
- & O(n \sum_{h=0}^{lg n} \frac{h}{2^{h}})
-\end{align*}
-
-Usando la ecuación {\bf A.8} del Cormen, tenemos
-
-\begin{align*}
- \sum_{h=0}^{\infty} \frac{h}{2^h} = 2 \\
-\end{align*}
-
-Esto acota el tiempo de ejecución de \texttt{convertir}:
-
-\begin{align*}
- O(n \sum_{h=0}^{lg n} \frac{h}{2^{h}}) &= O(n \sum_{h=0}^{\infty} \frac{h}{2^h}) \\
- &= O(n) \\
-\end{align*}
-
-Sabemos que $n$ es uniforme. Además,
-
-\begin{equation}
- \sum_{h=0}^{lg n} \frac{n}{2^{h+1}} O(h) \\
-\end{equation}
-
-es no decreciente sobre $n$ porque es una suma de términos positivos cuyo
-rango aumenta al aumentar $n$. Entonces, por la regla de la suavidad, el tiempo
-de ejecución de \texttt{convertir} es $O(n)$.
-
-% Pregunta 6
-\section{}
-
-La estrucura de datos cola tiene dos operaciones: \texttt{encolar} y
-\texttt{desencolar}. Para implementar una cola con dos pilas $S_1$ y $S_2$,
-añado una operación auxiliar \texttt{mudar}. Para una cola $Q$, voy a llamar a
-sus dos pilas $Q.S_1$ y $Q.S_2$. Supongo que nunca se ejecuta
-\texttt{desencolar} sobre una cola vacía (lo cual es un error), y defino
-entonces:
-
-\begin{lstlisting}
-encolar(Q, x) = empilar(Q.$S_1$, x)
-
-desencolar(Q) = si vacia(Q.$S_2$)
-                    mudar(Q.$S_1$, Q.$S_2$)
-                return desempilar(Q.$S_2$)
-
-mudar(Q.$S_1$, Q.$S_2$) = mientras $\neg$vacia(Q.$S_1$)
-                              x $\leftarrow$ desempilar(Q.$S_1$)
-                              empilar(Q.$S_2$, x)
-\end{lstlisting}
-
-Ahora hago un análisis del costo en tiempo amortizado de las operaciones
-\texttt{encolar} y \texttt{desencolar} para verificar que ambas son $O(1)$. Para
-ello utilizo el método del potencial. Defino la función potencial $\Phi$ para
-una cola $Q$ como
-
-\[\Phi(Q) = 3 * \#(Q.S_1) + \#(Q.S_2)\]
-
-donde
-
-\[\#(S)\]
-
-es el número de elementos en la pila $S$. Es claro que, como el número de
-elementos en una pila es no negativo, $\Phi$ es no negativo. Entonces, si $Q_0$
-es una pila recién creada y $Q_i$ es la misma pila luego de haberle aplicado
-alguna secuencia de operaciones, tenemos
-
-\begin{align*}
-\Phi(Q_i) &\geq 0 \\
-          &= \Phi(Q_0)
-\end{align*}
-
-Así, $\Phi$ cumple esa propiedad necesaria del método del potencial. Ahora
-calculo el costo amortizado $\hat{c}_i$ de las operaciones que definí:
-
-\begin{itemize}
- \item \texttt{encolar($Q$, $x$)}:
- Realizamos una operación de inserción en una pila y el potencial de la
- estructura aumenta en 3.
- \begin{align*}
-  \hat{c}_{i+1} &= c_{i+1} + \Phi(D_{i+1}) - \Phi(D_i) \\
-          &= 1 + (3 + \Phi(D_i)) - \Phi(D_i) \\
-          &= 1 + 3 \\
-          &= 4
- \end{align*}
-
- \item \texttt{mudar($Q.S_1$, $Q.S_2$)}:
- Si $Q.S_1$ tiene $k$ elementos, entonces realizamos $k$ operaciones de
- desempilar sobre $Q.S_1$ y $k$ operaciones de empilar sobre $Q.S_2$. Eliminar
- $k$ elementos de $Q.S_1$ hace que el potencial disminuya en $3*k$ y agregarlos
- a $Q.S_2$ hace que aumente en $k$. Entonces,
- \begin{align*}
-  \hat{c}_{i+1} &= c_{i+1} + \Phi(D_{i+1}) - \Phi(D_i) \\
-                &= k + k + \Phi(D_{i+1}) - \Phi(D_i) \\
-                &= 2*k + (k - 3*k + \Phi(D_i)) - \Phi(D_i) \\
-                &= 0
- \end{align*}
-
- \item \texttt{desencolar($Q$)}: Si $Q.S_2$ no tiene elementos, realizamos la
- la operación \texttt{mudar} --cuyo costo amortizado es $0$-- y estamos en el
- caso en que sí tiene. Cuando $Q.S_2$ tiene al menos un elemento, realizamos una
- operación de \texttt{desempilar} y el potencial de la estructura disminuye en
- $1$.
- \begin{align*}
-  \hat{c}_{i+1} &= c_{i+1} + \Phi(D_{i+1}) - \Phi(D_i) \\
-                &= 1 + (\Phi(D_i) -1) - \Phi(D_i) \\
-                &= 0
- \end{align*}
-
-\end{itemize}
-
-En conclusión, tenemos una estructura de datos cola implementada con dos pilas
-y cuyas operaciones tienen costo amortizado $O(1)$.
+test
 

### 2008-11-05 teoria_algoritmos/examen2/examen2.tex
añadí una explicación al caso de potencias de dos del algoritmo del ejercicio 3

@@
 
-\lstset{mathescape=true}
+\lstset{mathescape=true,frame=single}
+
+\newcommand{\asig}{\ensuremath{\leftarrow}}
+\renewcommand{\lstlistingname}{Listado} %listings en español
 
 \begin{document}
-
 % Título
@@
 
-test
+% Pregunta 2
+\section{}
+
+En ambos casos, el algoritmo
+devuelve la matriz $partidos$ donde $partidos[d][c]$
+es el competidor contra el cual
+jugará el competidor $c$ en el día $d$. Si
+$partidos[d][c] = 0$, el competidor $c$ no juega el día $d$. Supongo que
+$matriz(n,m)$ devuelve una matriz de tamaño $n x m$ con todas las casillas
+inicializadas en cero. También supongo que tengo los siguientes procedimientos disponibles
+como ayuda:
+
+\begin{lstlisting}
+procedure horizontal(M1[1..n][1..m1]; M2[1..n][1..m2])
+  resultado $\asig matriz(n,m1+m2)$
+
+  for i $\asig 1$ to n do
+    for j $\asig 1$ to $m1$ do
+      resultado[i][j] $\asig M1[i][j]$
+
+    for j $\asig 1$ to $m2$ do
+      resultado[i][m1+j] $\asig M2[i][j]$
+
+  return resultado
+
+procedure vertical(M1[1..n1][1..m1]; M2[1..n2][1..m2])
+  resultado $\asig matriz(n,m1+m2)$
+
+  for j $\asig 1$ to $m$ do
+    for i $\asig 1$ to $n1$ do
+      resultado[i][j] $\asig M1[i][j]$
+
+    for i $\asig 1$ to $n2$ do
+      resultado[n1+i][j] $\asig M2[i][j]$
+
+  return resultado
+\end{lstlisting}
+
+\begin{itemize}
+
+\item Caso $n$ potencia de dos:
+
+No considero los problemas de tamaño $1$ por no tener sentido un torneo con $1$ equipo y de $0$
+días de duración. Para resolver el problema creo el procedimiento auxiliar $tabla$ donde
+$tabla(n, inicio)$ es un torneo válido para los $n$ jugadores $inicio, inicio+1, ..., inicio+n-1$.
+Entonces la solución del problema es igual a $tabla(n, 1)$.
+
+El algoritmo divide un problema de tamaño $n$ en dos sub-problemas
+de tamaño $n/2$, lo cual se puede hacer por ser $n$ potencia de dos. El caso base $n = 2$
+se resuelve en un sólo día con un sólo partido donde los dos competidores juegan entre sí.
+En el caso recursivo, tengo dos sub-problemas $T_1$ y $T_2$, donde $T_1$ es un torneo de $n/2-1$ días
+de duración para los primeros $n/2$ jugadores y $T_2$ para los últimos $n/2$ jugadores. Tras resolver
+recursivamente ambos sub-problemas debo unir los resultados para obtener la solución. Observo que ningún
+jugador que aparece en $T_1$ aparece en $T_2$ y vice-versa. Por lo tanto puedo agregar $n/2-1$ juegos
+asignando juegos entre cada jugador $x$ del primer sub-problema y los contrincantes asignados a $x+n/2$
+en el segundo sub-problema. Eso se logra con la matriz
+
+$$
+\left[
+\begin{array}{cc}
+T_1 & T_2 \\
+T_2 & T_1 \\
+\end{array}
+\right]
+$$
+
+De esta manera me queda un torneo con $n/2-2$ juegos. Se puede ver que sólo faltan los juegos
+entre $x$ y $x+n/2$ con $x$ en $T_1$, ya que $x+n/2$ no aparece como contrincante de sí mismo
+en $T_2$. Entonces puedo agregar en el último día del torneo todos los juegos con esa forma para
+obtener un resultado válido de $n/2-1$ días. El algoritmo se puede observar en el listado
+\ref{alg:ejercicio3}.
+
+\begin{lstlisting}[float,caption={Solución al ejercicio 3},label=alg:ejercicio3]
+procedure torneo(n)
+  return tabla(n, 1)
+
+procedure tabla(n, inicio)
+  if n = 2 do
+    partidos[1][inicio] $\asig inicio+1$
+    partidos[1][inicio+1] $\asig inicio$
+    return partidos
+
+  m1 $\asig tabla(n/2, inicio)$
+  m2 $\asig tabla(n/2, inicio+(n/2))$
+
+  vi $\asig vertical(m1, m2)$
+  vd $\asig vertical(m2, m1)$
+
+  piso $\asig matriz(1,n)$
+
+  for i $\asig 1$ to $n / 2$ do
+    piso[1][i] $\asig i+(n/2)$
+    piso[1][$i+(n/2)$] $\asig i$
+
+  total $\asig horizontal(vi, vd)$
+
+  return $vertical(total, piso)$
+\end{lstlisting}
+
+\item Caso $n > 1$:
+
+\end{itemize}
 

### 2008-11-05 teoria_algoritmos/examen2/examen2.tex
el listing de las funciones auxiliares del ejercicio 3 es flotante y con caption
partes a y b de la pregunta 3

@@
 
-\begin{lstlisting}
+\begin{lstlisting}[float,caption={Procedimientos auxiliares para el ejercicio 3},label=alg:aux-ejercicio3]
 procedure horizontal(M1[1..n][1..m1]; M2[1..n][1..m2])
@@
 
+% Pregunta 3
+\section{}
+
+\renewcommand{\labelenumi}{(\alph{enumi})}
+
+\begin{enumerate}
+ \item El costo promedio en comparaciones a realizar es
+
+$$
+0,1 + (0,25+0,3)*2 + (0,2+0,05+0,05+0,05)*3 = 2,25
+$$
+
+El costo del árbol en la figura 8.8 del libro es menor y, por lo tanto, mejor.
+
+ \item
+
+Para construir un arbol con un conjunto de elementos se elige el elemento con mayor probabilidad
+de ser accedido y se coloca como raíz. Su sub-árbol izquierdo es el resultado de aplicar el algoritmo
+sobre el subconjunto del conjunto original formado por los elementos menores a la raíz. El derecho es el
+resultado
+de aplicar el algoritmo sobre el subconjunto de elementos mayores a la raíz. En caso de que alguno
+de los conjuntos sea vacío, entonces el resultado para ese sub-árbol es una hoja.
+
+ \item
+\end{enumerate}
+
+% Pregunta 4
+\section{}
+
 \end{document}

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
casi lista parte c de pregunta 3 de examen 2

@@
 \usepackage[spanish]{babel}
+\usepackage{qtree}
 
@@
 $$
-0,1 + (0,25+0,3)*2 + (0,2+0,05+0,05+0,05)*3 = 2,25
+0.1 + (0.25+0.3)*2 + (0.2+0.05+0.05+0.05)*3 = 2.25
 $$
@@
  \item
+
+El siguiente árbol es más eficiente en promedio:
+
+El costo promedio en comparaciones a realizar es
+
+$$
+0.25 + (0.2+0.3)*2 + (0.1+0.05)*3 + (0.05+0.05)*4 = 2.1
+$$
+
+Podemos ver que es menor que el costo promedio del árbol en la figura 8.8 del libro.
+Esto refuerza el hecho de que es usual que un problema no pueda ser resuelto con un
+algoritmo voraz (como el utilizado para construir dicho árbol).
+
 \end{enumerate}

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
arbol en la parte c de la pregunta 3, con lo que queda lista
bastante desarrollo de la pregunta 4

@@
 
+\Tree [.12 6 [.34 [.20 18 27 ] 35 ] ]
+
 El costo promedio en comparaciones a realizar es
@@
 
+Supongo que --como aparece en el libro-- la secuencia de entrada está ordenada.
+Voy a usar la definición de $C_{i,j}$ dada en el libro con la nueva condición de que
+$C_{i,j} = 0$ si $j < i$. Supongamos que tenemos una solución $X$
+óptima para la secuencia de elementos entre $i$ y $j$. Sea $k$ el índice del elemento
+raíz de esa solución. El sub-árbol izquierdo contiene los elementos menores que el elemento
+$k$, es decir, los elementos con índices entre $i$ y $k-1$ inclusive. El sub-árbol derecho contiene
+los elementos mayores que el elemento $k$, es decir, los elementos con índices entre $k+1$ y $j$
+inclusive. Sean $K_i$ y $K_d$ los costos de los sub-árboles izquierdo y
+derecho de la solución respectivamente. El costo de la solución $X$ es
+
+$$
+K_i + K_d + \sum_{k=i}^j p_k
+$$
+
+ya que debemos considerar el costo de ambos sub-árboles más el costo de acceder al
+elemento raíz más el costo añadido de incrementar la profundidad de cada nodo en
+cualquiera de los sub-árboles en uno (ya que estamos agregando la raíz a altura mayor).
+Si tenemos que $K_i \neq C_{i,k-1}$, entonces una solución mejor consiste en reemplazar
+el sub-árbol $K_i$ por uno tal que su costo sea igual a $C_{i,k-1}$, el cual es el costo
+mínimo para un sub-árbol con los elementos entre $i$ y $k-1$ inclusive. Esto contradice
+la suposición de que $X$ es una solución óptima.
+
+Por lo tanto tenemos obligatoriamente que el costo de $K_i$ es igual a $C_{i,k-1}$.
+Ocurre de manera análoga con el sub-árbol derecho y un árbol de costo $C_{k+1,j}$.
+Entonces el costo de la solución $X$ debe ser
+
+$$
+C_{i,k-1} + C_{k+1,j} + \sum_{k=i}^j p_k
+$$
+
+Ahora, supongamos al absurdo que existe $m \neq k$ tal que $i \leq m \leq j$ y
+$C_{i,m-1}+C_{m+1,j} < C_{i,k-1}+C_{k+1,j}$. Entonces el árbol que contiene los
+elementos entre $i$ y $j$ inclusive y tiene como raíz el elemento $m$ tiene menor
+costo, ya que
+
+$$
+C_{i,m-1} + C_{m+1,j} + \sum_{k=i}^j p_k < C_{i,k-1} + C_{k+1,j} + \sum_{k=i}^j p_k
+$$
+
+Esto contradice el hecho de que $X$ sea una solución óptima. Con esto, $k$ debe ser tal que
+no exista índice $m$ como el descrito. Así, he probado con el principio de optimalidad
+que
+
+$$
+C_{i,j} = (\text{{\bf min} } k : i \leq k \leq j : C_{i,k-1} + C_{k+1,j}) + \sum_{k=i}^j p_k
+$$
+
+Para obtener el algoritmo de programación dinámica solicitado,
+defino las tablas $S$ y $C$, ambas de tamaño $n$. Usaré la tabla $S$ para pre-calcular las sumatorias
+
+$$
+\sum_{k=i}^j p_k
+$$
+
+para todo $i < j$. Hago esto ya que necesito esos valores para resolver el problema. La tabla $C$ contendrá
+los valores $C_{ij}$.
+
+Para obtener la tabla $S$ utilizo el siguiente algoritmo
+
+\begin{lstlisting}[caption={Cálculo de la tabla S},label=alg:tablaS]
+for i $\asig 1$ to $n$ do
+    for j $\asig i+1$ to $n$ do
+        suma $\asig 0$
+
+        for k $\asig i$ to $j$ do
+            suma $\asig suma + p_k$
+
+        S[i][j] $\asig suma$
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Cálculo de la tabla C},label=alg:tablaC]
+for i $\asig 1$ to $n$ do
+    for j $\asig i+1$ to $n$ do
+        suma $\asig 0$
+
+        for k $\asig i$ to $j$ do
+            suma $\asig suma + p_k$
+
+        S[i][j] $\asig suma$
+\end{lstlisting}
+
+\begin{lstlisting}[caption={Cálculo del árbol},label=alg:arbol]
+for i $\asig 1$ to $n$ do
+    for j $\asig i+1$ to $n$ do
+        suma $\asig 0$
+
+        for k $\asig i$ to $j$ do
+            suma $\asig suma + p_k$
+
+        S[i][j] $\asig suma$
+\end{lstlisting}
+
 \end{document}

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
mas desarrollo de la pregunta 4

@@
 \renewcommand{\labelenumi}{(\alph{enumi})}
-
 \begin{enumerate}
@@
 Voy a usar la definición de $C_{i,j}$ dada en el libro con la nueva condición de que
-$C_{i,j} = 0$ si $j < i$. Supongamos que tenemos una solución $X$
+$C_{i,j} = 0$ si $j < i$ ya que un árbol sin elementos tiene como tiempo promedio de
+búsqueda $0$. Supongamos que tenemos una solución $X$
 óptima para la secuencia de elementos entre $i$ y $j$. Sea $k$ el índice del elemento
@@
 para todo $i < j$. Hago esto ya que necesito esos valores para resolver el problema. La tabla $C$ contendrá
-los valores $C_{ij}$.
+los valores $C_{i,j}$.
 
@@
 for i $\asig 1$ to $n$ do
+    suma $\asig p_i$
+
     for j $\asig i+1$ to $n$ do
-        suma $\asig 0$
-
-        for k $\asig i$ to $j$ do
-            suma $\asig suma + p_k$
-
+        suma $\asig suma + p_j$
         S[i][j] $\asig suma$
@@
 
+En el algoritmo encontramos dos ciclos, uno dentro de otro, ambos realizando $O(n)$ operaciones. Entonces
+el costo en tiempo del algoritmo es $O(n^2)$. Hecho esto utilizaré programación dinámica para calcular
+$C_{1,n}$. Recordemos que los casos para $C$ son
+
+\begin{equation*}
+C_{i,j} = 
+\begin{cases}
+p_i & \text{si $i=j$,} \\
+0 &\text{si $i>j$,} \\
+C_{i,j} = (\text{{\bf min} } k : i \leq k \leq j : C_{i,k-1} + C_{k+1,j}) + \sum_{k=i}^j p_k &\text{si $i<j$.} \\
+\end{cases}
+\end{equation*}
+
+Para empezar relleno en la tabla los casos base ($i \geq j$):
+
+\begin{lstlisting}[caption={Inicialización de casos base de la tabla C},label=alg:tablaC_base]
+for i $\asig 1$ to $n$ do
+    S[i][i] = p_i
+
+for i $asig 1$ to $n$ do
+    for j $asig 1$ to $i-1$ do
+        S[i][j]=0
+\end{lstlisting}
+
+Nos encontramos otra vez con un algoritmo de tiempo $O(n^2)$. Una vez resueltos los
+casos base intento llenar el resto de la tabla. Hay que observar que el valor de $C_{i,j}$
+depende únicamente de $S_{i,j}$ (ya calculado) y los $C_{x,y}$ tales que $x=i$ y $y<j$ o
+$y=j$ y $x<i$. En otras palabras, depende de los elementos que están a la izquierda en su
+fila y de los que están abajo en su columna. Puedo calcular los valores de la diagonal inmediatamente
+a la derecha de la diagonal recién calculada,
+
 \begin{lstlisting}[caption={Cálculo de la tabla C},label=alg:tablaC]
-for i $\asig 1$ to $n$ do
-    for j $\asig i+1$ to $n$ do
-        suma $\asig 0$
 
-        for k $\asig i$ to $j$ do
-            suma $\asig suma + p_k$
-
-        S[i][j] $\asig suma$
 \end{lstlisting}

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
pregunta 4 casi lista

@@
         suma $\asig suma + p_j$
-        S[i][j] $\asig suma$
+        $S_{i,j} \asig suma$
 \end{lstlisting}
@@
 
-Para empezar relleno en la tabla los casos base ($i \geq j$):
+Para empezar relleno en la tabla los casos base ($i \geq j$) e inicializo los demás valores
+con $+\infty$ (porque estoy buscando el mínimo):
 
@@
 for i $\asig 1$ to $n$ do
-    S[i][i] = p_i
+    $S_{i,i} = p_i$
 
-for i $asig 1$ to $n$ do
-    for j $asig 1$ to $i-1$ do
-        S[i][j]=0
+for i $\asig 1$ to $n$ do
+    for j $\asig 1$ to $i-1$ do
+        $S_{i,j} \asig 0$
+
+for i $\asig 1$ to $n$ do
+    for j $\asig i+1$ to $n$ do
+        $S_{i,j} \asig +\infty$
 \end{lstlisting}
 
-Nos encontramos otra vez con un algoritmo de tiempo $O(n^2)$. Una vez resueltos los
+Una vez resueltos los
 casos base intento llenar el resto de la tabla. Hay que observar que el valor de $C_{i,j}$
-depende únicamente de $S_{i,j}$ (ya calculado) y los $C_{x,y}$ tales que $x=i$ y $y<j$ o
-$y=j$ y $x<i$. En otras palabras, depende de los elementos que están a la izquierda en su
-fila y de los que están abajo en su columna. Puedo calcular los valores de la diagonal inmediatamente
-a la derecha de la diagonal recién calculada,
+depende únicamente de $S_{i,j}$ (ya calculado) y los $C_{x,y}$ tales que ($x=i \wedge y<j$) o
+($y=j \wedge x>i$). En otras palabras, depende de los elementos que están a la izquierda en su
+fila y de los que están abajo en su columna si en la tabla tenemos índices crecientes de izquierda
+a derecha y de arriba hacia abajo.
 
-\begin{lstlisting}[caption={Cálculo de la tabla C},label=alg:tablaC]
+Llamemos diagonal $p$ a la diagonal descendiente que comienza en la casilla $C_{i,i+p}$. Entonces
+puedo calcular la diagonal $p$ si y sólo si todas las diagonales $q < p$ están calculadas
+ya que para una casilla
+dada de esa diagonal $C_{i,i+p}$ es necesario y suficiente que estén calculadas todas las casillas
+entre $C_{i,i-1}$ y $C_{i,i+p-1}$ y las casillas entre $C_{i+1,i+p}$ y $C_{i+p+1,i+p}$.
 
+Entonces puedo calcular progresivamente todas las diagonales desde la mitad
+hacia la derecha hasta calcular $C_{1,n}$ en la última diagonal. Esto resulta en
+el siguiente algoritmo:
+
+\begin{lstlisting}[caption={Cálculo de la tabla C},label=alg:completoC]
+for j $\asig 1$ to $n-1$ do /* diagonal j */
+    for i $\asig 1$ to $n$ do /* fila i de la diagonal j */
+        for k $\asig i$ to $i+j$ do
+            actual $\asig$ $C_{i,k-1}+C_{k+1,i+j}+S_{i,i+j}$
+            $C_{i,i+j}$=min($C_{i,i+j}$,actual)
 \end{lstlisting}
 
+El algoritmo contiene tres ciclos anidados que pueden realizar $O(n)$ operaciones. Su
+complejidad es $O(n^3)$. Habiendo hallado el valor para $C_{1,n}$ debo construir un árbol
+de búsqueda óptimo con los valores $C_{i,j}$. Observemos que dados $i, j$
+debe existir $k$ tal que $C_{i,j} - \sum_{k=i}^j p_k = C_{i,k-1} + C_{k+1,j}$.
+Es claro que el costo de un árbol con raíz en $k$ tal que ambos sub-árboles son óptimos es
+igual a $C_{i,j}$. Entonces para construir un árbol óptimo simplemente debemos hallar $k$
+y construir recursivamente el árbol izquierdo (elementos entre $i$ y $k-1$) y el derecho
+(elementos entre $k+1$ y $j$). Suponiendo que $arbol(A,I,D)$ es un árbol cuyo valor es el elemento
+de la secuencia original con índice $A$ y cuyos
+sub-árboles son $I$ y $D$, el siguiente algoritmo recursivo halla el árbol óptimo para
+$i,j$ de la manera descrita:
+
 \begin{lstlisting}[caption={Cálculo del árbol},label=alg:arbol]
-for i $\asig 1$ to $n$ do
-    for j $\asig i+1$ to $n$ do
-        suma $\asig 0$
+procedure optimo(i, j)
+    if i > j do
+        return hoja
+
+    if i = j do
+        return arbol(i, hoja, hoja)
 
     for k $\asig i$ to $j$ do
-            suma $\asig suma + p_k$
-
-        S[i][j] $\asig suma$
+        if $C_{i,j} - S_{i,j} = C_{i,k-1} + C_{k+1,j}$ do
+            return arbol(k, optimo(i,k-1), optimo(k+1,j))
 \end{lstlisting}

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
respuesta 4 parece estar lista

@@
 
+A continuación presento el algoritmo completo para conseguir un árbol de búsqueda óptimo.
+$abo()$ será un árbol de búsqueda óptimo para los elementos de la secuencia con índice
+entre $1$ y $n$ inclusive.
+
+\begin{lstlisting}[float,caption={Cálculo del árbol},label=alg:completo]
+procedure optimo(i, j)
+    if i > j do
+        return hoja
+
+    if i = j do
+        return arbol(i, hoja, hoja)
+
+    for k $\asig i$ to $j$ do
+        if $C_{i,j} - S_{i,j} = C_{i,k-1} + C_{k+1,j}$ do
+            return arbol(k, optimo(i,k-1), optimo(k+1,j))
+
+procedure abo()
+    /* calculo de S */
+    for i $\asig 1$ to $n$ do
+        suma $\asig p_i$
+
+        for j $\asig i+1$ to $n$ do
+            suma $\asig suma + p_j$
+            $S_{i,j} \asig suma$
+
+    /* calculo de C */
+    for i $\asig 1$ to $n$ do
+        $S_{i,i} = p_i$
+
+    for i $\asig 1$ to $n$ do
+        for j $\asig 1$ to $i-1$ do
+            $S_{i,j} \asig 0$
+
+    for i $\asig 1$ to $n$ do
+        for j $\asig i+1$ to $n$ do
+            $S_{i,j} \asig +\infty$
+
+    for j $\asig 1$ to $n-1$ do /* diagonal j */
+        for i $\asig 1$ to $n$ do /* fila i de la diagonal j */
+            for k $\asig i$ to $i+j$ do
+                actual $\asig$ $C_{i,k-1}+C_{k+1,i+j}+S_{i,i+j}$
+                $C_{i,i+j}$=min($C_{i,i+j}$,actual)
+
+    /* calculo del arbol */
+    return optimo(1,n)
+\end{lstlisting}
+
 \end{document}

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
arreglitos de la pregunta 4

@@
 $abo()$ será un árbol de búsqueda óptimo para los elementos de la secuencia con índice
-entre $1$ y $n$ inclusive.
+entre $1$ y $n$ inclusive. Este algoritmo necesita tiempo $O(n^3)$ para correr
+sobre una secuencia de $n$ elementos. Aplicar el algoritmo a la instancia del problema 8.33
+resultó en el siguiente árbol de costo promedio 2.1:
 
-\begin{lstlisting}[float,caption={Cálculo del árbol},label=alg:completo]
+\Tree [.12 6 [.34 [.20 18 27 ] 35 ] ]
+
+\begin{lstlisting}[float,caption={Solución del problema 8.34},label=alg:completo]
 procedure optimo(i, j)

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
agregue varias cosas de la pregunta 2

@@
 \item Caso $n > 1$:
+
+El caso base $n = 2$
+se resuelve en un sólo día con un sólo partido donde los dos competidores juegan entre sí.
+Cuando $n$ es impar agrego un jugador ficticio nuevo y resuelvo el problema
+para los $n+1$ jugadores, lo que me dará una solución de $(n+1)-1 = n$ jugadores
+(porque $n+1$ es par). Considero entonces que los juegos entre un jugador original
+y el jugador ficticio son días libres para dicho jugador original. Me queda
+entonces por resolver el caso en que $n$ es par, lo cual voy a suponer de ahora en
+adelante.
+
+Supongamos que debo organizar el torneo para los $n$ jugadores a partir del jugador $inicio$.
+Si $n/2$ es par, entonces el resultado de los sub-problemas $<inicio,inicio+1,\ldots,inicio+n/2-1>$
+y $<inicio+n/2,inicio+n/2+1,\ldots,inicio+n-1>$ no contiene jugadores con días libres ya que es
+posible organizarlos en $n/2-1$ días. En ese caso, y de manera análoga a la unión de sub-problemas
+hecha en el caso de $n$ potencia de dos, si $T_1$ es solución del primer sub-problema y $T_2$ del segundo,
+puedo organizar $n/2-2$ juegos de la siguiente manera:
+
+$$
+\left[
+\begin{array}{cc}
+T_1 & T_2 \\
+T_2 & T_1 \\
+\end{array}
+\right]
+$$
+
+Al igual que ocurrió con el caso de las potencias de 2, sólo quedan por agregar
+los juegos en los que un jugador $x$ del primer sub-problema
+juega con el jugador $x+n/2$. Esto añade un día de juegos y nos queda el torneo
+completo en $n/2-1$ días.
+
+Resta por ver qué pasa cuando $n/2$ es impar. Si $T_1$ es solución del primer sub-problem
+ y $T_2$ del segundo, tenemos que $T_1$ y $T_2$ van a tener $n/2$ días cada uno, y en cada
+día aparece un jugador libre (porque $n$ es impar).
 

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
algoritmo hablado de la 2 listo

@@
 hecha en el caso de $n$ potencia de dos, si $T_1$ es solución del primer sub-problema y $T_2$ del segundo,
-puedo organizar $n/2-2$ juegos de la siguiente manera:
+puedo organizar $n-2$ juegos de la siguiente manera:
 
@@
 juega con el jugador $x+n/2$. Esto añade un día de juegos y nos queda el torneo
-completo en $n/2-1$ días.
+completo en $n-1$ días.
 
-Resta por ver qué pasa cuando $n/2$ es impar. Si $T_1$ es solución del primer sub-problem
+Resta por ver qué pasa cuando $n/2$ es impar. Si $T_1$ es solución del primer sub-problema
 y $T_2$ del segundo, tenemos que $T_1$ y $T_2$ van a tener $n/2$ días cada uno, y en cada
-día aparece un jugador libre (porque $n$ es impar).
+día aparece un jugador libre (porque $n$ es impar). Puedo entonces organizar $n/2$ días
+de torneo uniendo ambos resultados y poniendo en cada día a jugar los dos jugadores que
+en los sub-problemas tienen ese día libre. Con esto quedan $n/2-1$ días por organizar.
+
+Observemos que los juegos hechos entre jugadores que tenían días libres en los sub-problemas
+son tales que sus participantes son $x$ y $x+n/2$ con $x$ en los jugadores de $T_1$. Esto
+porque al resolver subproblemas asignamos días libres al mismo jugador relativo a $inicio$.
+Salvo por este caso no hay juegos entre jugadores de $T_1$ y de $T_2$.
+Entonces para los $n/2-1$ días que quedan realizo juegos entre los jugadores $x$ y $x+n/2+k$
+con $x$ en $T_1$ y $k > 0$. Esto termina por completar los $n/2-1$ días que faltan.
+
+A continuación el algoritmo desarrollado a partir de este razonamiento:
 

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
mejor definicion de union vertical y horizontal

@@
 
-En ambos casos, el algoritmo
-devuelve la matriz $partidos$ donde $partidos[d][c]$
+En ambos casos, el algoritmo una matriz $T$ donde $T[d][c]$
 es el competidor contra el cual
-jugará el competidor $c$ en el día $d$. Si
-$partidos[d][c] = 0$, el competidor $c$ no juega el día $d$. Supongo que
-$matriz(n,m)$ devuelve una matriz de tamaño $n x m$ con todas las casillas
-inicializadas en cero. También supongo que tengo los siguientes procedimientos disponibles
-como ayuda:
+jugará el competidor $c$ en el día $d$.
+Supongo que
+$matriz(n,m)$ devuelve una matriz de tamaño $n x m$. También supongo que tengo las funciones
+$unir_horizontal$ y $unir_vertical$ tales que si $H_1$ y $H_2$ son matrices con la misma altura
+y $V_1$ y $V_2$ son matrices con el mismo ancho, entonces
+
+$$
+unir\_horizontal(H_1, H_2) = \left[
+\begin{array}{cc}
+H_1 & H_2 \\
+\end{array}
+\right]
+$$
+
+y
+
+$$
+unir\_vertical(V_1, V_2) = \left[
+\begin{array}{c}
+V_1 \\
+V_2 \\
+\end{array}
+\right]
+$$
 
@@
   if n = 2 do
-    partidos[1][inicio] $\asig inicio+1$
-    partidos[1][inicio+1] $\asig inicio$
-    return partidos
+    T = matriz(1,2)
+
+    T[1][inicio] $\asig inicio+1$
+    T[1][inicio+1] $\asig inicio$
+    return T
 

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
corregidos pequenos errores de las funciones auxiliares y eliminada version vieja

@@
 $matriz(n,m)$ devuelve una matriz de tamaño $n x m$. También supongo que tengo las funciones
-$unir_horizontal$ y $unir_vertical$ tales que si $H_1$ y $H_2$ son matrices con la misma altura
+$unir\_horizontal$ y $unir\_vertical$ tales que si $H_1$ y $H_2$ son matrices con la misma altura
 y $V_1$ y $V_2$ son matrices con el mismo ancho, entonces
@@
 $$
-
-\begin{lstlisting}[float,caption={Procedimientos auxiliares para el ejercicio 3},label=alg:aux-ejercicio3]
-procedure horizontal(M1[1..n][1..m1]; M2[1..n][1..m2])
-  resultado $\asig matriz(n,m1+m2)$
-
-  for i $\asig 1$ to n do
-    for j $\asig 1$ to $m1$ do
-      resultado[i][j] $\asig M1[i][j]$
-
-    for j $\asig 1$ to $m2$ do
-      resultado[i][m1+j] $\asig M2[i][j]$
-
-  return resultado
-
-procedure vertical(M1[1..n1][1..m1]; M2[1..n2][1..m2])
-  resultado $\asig matriz(n,m1+m2)$
-
-  for j $\asig 1$ to $m$ do
-    for i $\asig 1$ to $n1$ do
-      resultado[i][j] $\asig M1[i][j]$
-
-    for i $\asig 1$ to $n2$ do
-      resultado[n1+i][j] $\asig M2[i][j]$
-
-  return resultado
-\end{lstlisting}
 
@@
 
-  vi $\asig vertical(m1, m2)$
-  vd $\asig vertical(m2, m1)$
+  vi $\asig unir_vertical(m1, m2)$
+  vd $\asig unir_vertical(m2, m1)$
 
@@
 
-  total $\asig horizontal(vi, vd)$
+  total $\asig unir_horizontal(vi, vd)$
 
-  return $vertical(total, piso)$
+  return $unir_vertical(total, piso)$
 \end{lstlisting}

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
algoritmo del 2 listo aunque chucuto

@@
 obtener un resultado válido de $n/2-1$ días. El algoritmo se puede observar en el listado
-\ref{alg:ejercicio3}.
+\ref{alg:ejercicio3a}.
 
-\begin{lstlisting}[float,caption={Solución al ejercicio 3},label=alg:ejercicio3]
+\begin{lstlisting}[float,caption={Solución al ejercicio 3},label=alg:ejercicio3a]
 procedure torneo(n)
@@
 
-A continuación el algoritmo desarrollado a partir de este razonamiento:
+En el listado \ref{alg:ejercicio3b} se puede observar
+el algoritmo desarrollado a partir de este razonamiento.
+
+\begin{lstlisting}[caption={Solución al ejercicio 3},label=alg:ejercicio3b]
+procedure torneo(n)
+  return tabla(n, 1)
+
+procedure tabla(n, inicio)
+  if n = 2 do
+    T = matriz(1,2)
+
+    T[1][inicio] $\asig inicio+1$
+    T[1][inicio+1] $\asig inicio$
+    return T
+
+  if n es impar do
+    return tabla(n+1, inicio)
+
+  m1 $\asig tabla(n/2, inicio)$
+  m2 $\asig tabla(n/2, inicio+(n/2))$
+
+  if n/2 es par do
+    vi $\asig unir\_vertical(m1, m2)$
+    vd $\asig unir\_vertical(m2, m1)$
+
+    piso $\asig matriz(1,n)$
+
+    for i $\asig 1$ to $n / 2$ do
+      piso[1][i] $\asig i+(n/2)$
+      piso[1][$i+(n/2)$] $\asig i$
+
+    total $\asig unir\_horizontal(vi, vd)$
+
+    return $unir\_vertical(total, piso)$
+
+  techo $\asig matriz(n/2, n)$
+
+  for $jug \asig 1$ to $n/2$
+    for $dia \asig 1$ to $n/2$
+        if m1[dia][jug] = $n/2+1$ do
+          techo[dia][jug] = jug+n/2
+          techo[dia][jug+n/2] = jug
+        else do
+          techo[dia][jug] = m1[dia][jug]
+          techo[dia][jug+n/2] = m2[dia][jug]
+
+  piso $\asig matriz(n/2-1, n)$
+
+  for $x \asig 1$ to $n/2-1$ do
+    for $y \asig 1$ to $n/2$ do
+      if y+(n/2+1) $\leq$ n do
+        piso[x][y] = y+(n/2+1)
+        piso[x][y+(n/2+1)] = y
+      else do
+        piso[x][y] = n/2+1
+        piso[x][n/2+1] = y
+
+  return $unir\_vertical(techo, piso)$
+
+\end{lstlisting}
 

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
fix bug en algoritmo de caso general de ej. 2

@@
 \section{}
+
+\begin{lstlisting}[float,caption={Solución al ejercicio 1},label=alg:ejercicio1]
+procedure buscar(A[1$\ldots$N], X)
+    $i \asig 1$
+    $j \asig 2$
+
+    for $k \asig 1$ to $N$ do
+
+\end{lstlisting}
 
@@
 
-\begin{lstlisting}[float,caption={Solución al ejercicio 3},label=alg:ejercicio3a]
+\begin{lstlisting}[float,caption={Solución al ejercicio 2, caso potencias de dos},label=alg:ejercicio3a]
 procedure torneo(n)
@@
 
-\begin{lstlisting}[caption={Solución al ejercicio 3},label=alg:ejercicio3b]
+\begin{lstlisting}[caption={Solución al ejercicio 2, caso general},label=alg:ejercicio3b]
 procedure torneo(n)
@@
     for $y \asig 1$ to $n/2$ do
-      if y+(n/2+1) $\leq$ n do
-        piso[x][y] = y+(n/2+1)
-        piso[x][y+(n/2+1)] = y
+      if (n/2+y)+x $\leq$ n do
+        piso[x][y] = (n/2+y)+x
+        piso[x][(n/2+y)+x] = y
       else do
-        piso[x][y] = n/2+1
-        piso[x][n/2+1] = y
+        piso[x][y] = y+x
+        piso[x][y+x] = y
 

### 2008-11-06 teoria_algoritmos/examen2/examen2.tex
version final del examen2, bleh

@@
 \begin{document}
+
 % Título
@@
 
-\begin{lstlisting}[float,caption={Solución al ejercicio 1},label=alg:ejercicio1]
+La siguiente es una versión iterativa del algoritmo:
+
+\begin{lstlisting}[caption={Solución al ejercicio 1},label=alg:ejercicio1]
 procedure buscar(A[1$\ldots$N], X)
     $i \asig 1$
-    $j \asig 2$
+    $j \asig 1$
 
-    for $k \asig 1$ to $N$ do
+    while true do
+        if $i \geq N+1$
+            return false
 
+        if $j \geq N+1$
+            return false
+
+        if A[i]+A[j]=X do
+            return (i,j)
+
+        if(A[i]+A[j]>X) do
+            i $\asig$ i+1
+            j $\asig$ j-1
+        else do
+            j $\asig$ j+1
 \end{lstlisting}
+
+Es $O(N)$ porque siempre $i$ se incrementa y $j$ se decrementa o $j$ se incrementa en uno, y ambas
+variables empiezan en $1$ y no crecen más allá de $N+1$. En un arreglo con elementos repetidos
+el algoritmo no funcionaría, porque en el paso en que se hace $j $\asig$ j-1$ habría que
+retroceder posiblemente hasta el principio. No es posible hacer el algoritmo $O(N)$ en un
+arreglo con elementos repetidos.
 

### 2008-11-21 teoria_algoritmos/exposicion/presentacion.tex
anadi los archivos para la exposicion

\documentclass{beamer}

\usepackage{beamerthemesplit}
\usepackage[spanish,activeacute]{babel}
\usepackage[utf8]{inputenc}

\begin{document}
\title{Solución exacta de problemas difíciles: Backtraking para "Satisfiability"}
\author{Daniel Izquierdo}
\date{25 de noviembre de 2008}

\frame{\titlepage}

\section[Contenido]{}
\frame
{
  \begin{itemize}
  \item Marco teórico.
  \item Diseño e implementación.
  \item Resultados.
  \end{itemize}
}

\section{Introducción}

\frame
{
  \frametitle{Introducción}

  Este proyecto consiste del diseño, implementación y evaluación de un
  compilador paralelizante.
}

\section{Preguntas}

\frame
{
  \frametitle{Preguntas}
}

\end{document}

### 2008-11-21 construccion/tarea1.tex
directorio de construccion junto archivos de la tarea 1

\documentclass{article}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage[spanish]{babel}
\usepackage{alltt}

% Para codigo bonito
\newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}

\newcommand{\algoritmo}[2]
{
{
  \singlespace
  \begin{algorithm}[h]
    \caption{#1}
    \begin{algorithmic}[1]
      #2
    \end{algorithmic}
  \end{algorithm}
}
}

\newcommand{\theHalgorithm}{\arabic{algorithm}}

\lstset{language=GCL,mathescape=true,frame=single}

\begin{document}

% Título

\title{Tarea 1 \\ Construcción Formal de Programas}
\author{Daniel Izquierdo \\ \#08-86809}
\date{4 de noviembre de 2008}

\maketitle

% Pregunta 1
\section{Ejercicio 2.0.0}

\begin{itemize}
\item $\{true\} S \{true\}$: el programa $S$ siempre termina.
\item $\{false\} S \{true\}$: no da información sobre la terminación de $S$.
\end{itemize}

% Pregunta 2
\section{Ejercicio 2.2.1}

Debo demostrar $\{true\} x := x + 1 \{true\}$. Tenemos

\begin{align*}
 & \{true\} x := x + 1 \{true\} \\
 \equiv & <\text{equivalencia TH / wp}> \\
 & [ true \Rightarrow wp.(x := x + 1).true ] \\
 \equiv & <true => P \equiv P> \\
 & [ wp.(x := x + 1).true ] \\
 \equiv & <wp.(x := E).Q \equiv Q(x := E)> \\
 & [ true(x := x + 1) ] \\
 \equiv & <\text{sustitución}> \\
 & [ true ] \\
 \equiv & <\text{cuantificación sin variables ligadas}> \\
 & true
\end{align*}

% Pregunta 3
\section{Ejercicio 2.4.1}

\begin{align*}
 & \{P\} \text{{\bf if }} B_0 \rightarrow S_0 ; S | B_1 \rightarrow S_1 ; S \text{ {\bf fi}} \{Q\} \\
 \equiv & <\text{definición de {\bf if}}> \\
 & [ P \Rightarrow B_0 \vee B_1 ] \wedge \{P \wedge B_0\} S_0 ; S \{Q\} \wedge \{P \wedge B_1\} S_1 ; S \{Q\} \\
 \equiv & <\text{equivalencia TH / wp}> \\
 & [ P \Rightarrow B_0 \vee B_1 ] \wedge [P \wedge B_0 \Rightarrow wp.(S_0 ; S).Q] \wedge [P \wedge B_1 \Rightarrow wp.(S_1 ; S).Q] \\
 \equiv & <[wp.(S ; T).Q \equiv wp.S.(wp.T.Q)]> \\
 & [ P \Rightarrow B_0 \vee B_1 ] \wedge [P \wedge B_0 \Rightarrow wp.S_0.(wp.S.Q)] \wedge [P \wedge B_1 \Rightarrow wp.S_1.(wp.S.Q)] \\
 \equiv & <\text{equivalencia TH / wp}> \\
 & [ P \Rightarrow B_0 \vee B_1 ] \wedge \{P \wedge B_0\} S_0 \{wp.S.Q\} \wedge \{P \wedge B_1\} S_1 \{wp.S.Q\} \\
 \equiv & <\text{definición de {\bf if}}> \\
 & \{P\} \text{{\bf if }} B_0 \rightarrow S_0 | B_1 \rightarrow S_1 \text{ {\bf fi}} \{wp.S.Q\} \\
 \equiv & <\text{equivalencia TH / wp}> \\
 & [P \Rightarrow wp.(\text{{\bf if }} B_0 \rightarrow S_0 | B_1 \rightarrow S_1 \text{ {\bf fi}}).(wp.S.Q)] \\
 \equiv & <[wp.(S ; T).Q \equiv wp.S.(wp.T.Q)]> \\
 & [P \Rightarrow wp.(\text{{\bf if }} B_0 \rightarrow S_0 | B_1 \rightarrow S_1 \text{ {\bf fi}} ; S).Q] \\
 \equiv & <\text{equivalencia TH / wp}> \\
 & \{P\} \text{{\bf if }} B_0 \rightarrow S_0 | B_1 \rightarrow S_1 \text{ {\bf fi}} ; S \{Q\}
\end{align*}

% Pregunta 4
\section{Ejercicio 2.6.2}

\begin{align*}
 & \{P\} S \{Q\} \\
 \equiv & <\text{equivalencia TH / wp}> \\
 & [P => wp.S.Q] \\
 \Rightarrow & <\text{introducción del existencial con } y=y> \\
 & [P => (\exists y : y \in \mathbb{Z} : wp.S.Q)] \\
 \Rightarrow & <wp.S.Q \vee wp.S.R \Rightarrow wp.S.(Q \vee R)> \\
 & [P => wp.S.(\exists y : y \in \mathbb{Z} : Q)] \\
 \equiv & <y \text{ no ocurre libre en } (\exists y : y \in \mathbb{Z} : Q)> \\
 & [P => wp.S.(\forall y : y \in \mathbb{Z} : (\exists y : y \in \mathbb{Z} : Q))] \\
 \equiv & <wp.S.Q \wedge wp.S.R \equiv wp.S.(Q \wedge R)> \\
 & [P => (\forall y : y \in \mathbb{Z} : wp.S.(\exists y : y \in \mathbb{Z} : Q))] \\
 \equiv & <[wp.|[\text{var } y : \text{int } ; S |].Q \equiv (\forall y : y \in \mathbb{Z} : wp.S.Q)]> \\
 & [P => wp.|[\text{var } y : \text{int } ; S |].(\exists y : y \in \mathbb{Z} : Q)] \\
 \equiv & <\text{equivalencia TH / wp}> \\
 & \{P\} |[\text{var } y : \text{int } ; S |] \{(\exists y : y \in \mathbb{Z} : Q)\} \\
\end{align*}

% Pregunta 5
\section{Ejercicio 4.1.3}

Para empezar, observo que $i = A*B$ satisface $i$ $mod$ $A = 0 \wedge i$ $mod$ $B = 0$.
Como estamos buscando el mínimo, esto me da una cota superior para el resultado.
Pongo como primera invariante

$$
P_0: 1 \leq x \leq A*B
$$

Usando la técnica de tomar una conjunción como invariante, pongo la invariante

$$
P_1: x \text{ {\bf mod} } A = 0
$$

y el guardia

$$
B: x \text{ {\bf mod} } B \neq 0
$$

Cuando tengamos $P_0 \wedge P_1 \wedge \neg B$ tendremos un múltiplo de $A$ y $B$, mas
no necesariamente el mínimo. Añado entonces una condición más a la invariante:

$$
P_2: (\forall i : 1 \leq i < x \wedge i \text{ {\bf mod} } A = 0 : i \text{ {\bf mod} } B \neq 0)
$$

De esta manera, si se cumplen las invariantes y no el guardia llegamos a la postcondición.
Es claro que a inicialización $x := A$ satisface $P_0$ y $P_1$. También satisface
$P_2$, porque no hay número $r$ tal que $1 \leq r < x \wedge r \text{ {\bf mod} } A = 0$.
Con todo esto tenemos una aproximación a la solución (listado \ref{prog413_parcial}).

\begin{lstlisting}[float,caption={Solución parcial del ejercicio 4.1.3},label=prog413_parcial]
|[
    con A, B : int {$A > 0 \wedge B > 0$};
    var x : int;

    x := A;

    {inv $P_0 \wedge P_1 \wedge P_2$}
    do $x mod B \neq 0 \rightarrow$
        S
    od

    {$x = A \text{ {\bf lcm} } B$}
]|
\end{lstlisting}

En el rango de la cuantificación en $P_2$ sólo considero números divisibles por $A$.
El próximo número a considerar en un momento dado es $x + A$. Investigo para $S$ la
asignación $x := x + A$. Es claro que $P_1 \Rightarrow P_1(x := x + A)$. Por otra parte,
suponiendo la invariante, sea $p > 0$ tal que $x=p*A$. Tenemos $1 \leq A*p = x < A*B$ (porque
$x \text{ {\bf mod} } B \neq 0$. Por lo tanto, $p < B$ y entonces

\begin{align*}
 & P_0(x := x + A) \\
 \equiv & <\text{sustitución y definición de } P_0> \\
 & 1 \leq x + A \leq A*B \\
 \equiv & <\text{x = p*A}> \\
 & 1 \leq p*A + A \leq A*B \\
 \equiv & <\text{aritmética}> \\
 & 1 \leq A*(p+1) \leq A*B \\
 \equiv & <p < B, A*p \geq 1, \text{aritmética}> \\
 & true
\end{align*}

Además,

\begin{align*}
 & P_2(x := x + A) \\
 \equiv & <\text{sustitución y definición de } P_2> \\
 & (\forall i : 1 \leq i < x+A \wedge i \text{ {\bf mod} } A = 0 : i \text{ {\bf mod} } B \neq 0) \\
 \equiv & <\text{separación de rangos}> \\
 & (\forall i : 1 \leq i < x \wedge i \text{ {\bf mod} } A = 0 : i \text{ {\bf mod} } B \neq 0) \wedge \\
 & (\forall i : x \leq i < x+A \wedge i \text{ {\bf mod} } A = 0 : i \text{ {\bf mod} } B \neq 0) \\
 \equiv & <\text{definición } P_2> \\
 & P_2 \wedge (\forall i : x \leq i < x+A \wedge i \text{ {\bf mod} } A = 0 : i \text{ {\bf mod} } B \neq 0) \\
 \equiv & <\text{separación de rangos, suposición } P_2 \text{, neutro de } \wedge \text{, } P_1> \\
 & x \text{ {\bf mod} } B \neq 0 \wedge (\forall i : x+1 \leq i < x+A \wedge i \text{ {\bf mod} } A = 0 : i \text{ {\bf mod} } B \neq 0) \\
 \equiv & <\text{suposición guardia, neutro de } \wedge> \\
 & (\forall i : x+1 \leq i < x+A \wedge i \text{ {\bf mod} } A = 0 : i \text{ {\bf mod} } B \neq 0) \\
 \equiv & <\text{rango vacío por } P_1> \\
 & true
\end{align*}

Para ver la terminación del ciclo, pongo la variante $A*B-x$. Noto primero que
$P_0$ implica que la variante es no negativa. Si tenemos $X = A*B-x$, entonces

\begin{align*}
 & X(x := x + A) \\
 = & <\text{sustitución}> \\
 & A*B-x+A \\
 < & <A > 0, \text{aritmética, definición de X}> \\
 & X
\end{align*}

Queda entonces el programa final (listado \ref{prog413}).

\begin{lstlisting}[float,caption={Solución del ejercicio 4.1.3},label=prog413]
|[
    con A, B : int {$A > 0 \wedge B > 0$};
    var x : int;

    x := A;

    {inv $P_0 \wedge P_1 \wedge P_2$}
    {bound A*B-x}
    do $x mod B \neq 0 \rightarrow$
        x := x + A
    od

    {$x = A \text{ {\bf lcm} } B$}
]|
\end{lstlisting}

% Pregunta 6
\section{Ejercicio 4.2.4}

Voy a utilizar un lazo con la técnica de sustitución de constante por variable
para resolver el problema. Introduzco la variable $n$ y entonces tenemos
las invariantes

$$
P_0: r = (\sum i : 0 \leq i < n : f.i * X^i)
$$

y

$$
P_1: 0 \leq n \leq N
$$

Por construcción, $P_0 \wedge n = N$ implica la postcondición. Como la sumatoria
sobre un rango vacío es 0, se satisfacen $P_0$ y $P_1$ con $n, r := 0, 0$. El
guardia del lazo será $n \neq N$. Al incrementar $n$ en 1 y suponer
$P_0 \wedge P_1 \wedge n \neq N$ tenemos

\begin{align*}
 & (\sum i : 0 \leq i < n + 1: f.i * X^i) \\
 = & <\text{separación de rango con } 0 \leq n < n+ 1> \\
 & (\sum i : 0 \leq i < n : f.i * X^i) + f.n * X^n \\
 = & <P_0> \\
 & r + f.n * X^n
\end{align*}

En este punto noto que es necesaria una variable para el valor de $X^n$.
Introduzco entonces la variable $x$ con la invariante

$$
P_2: x = X^n
$$

Con esto queda que

\begin{align*}
 & (\sum i : 0 \leq i < n + 1: f.i * X^i) \\
 = & <\text{demostrado previamente}> \\
 & r + f.n * X^n \\
 = & <P_2> \\
 & r + f.n * x
\end{align*}

Por lo tanto, la asignación $n, r := n + 1, r + f.n * x$ basta para
satisfacer $P_0$. Queda por satisfacer $P_2$. Derivo

\begin{align*}
 & X^{(n+1)} \\
 = & <\text{definición de potencia}> \\
 & X^n * X \\
 = & <P_2> \\
 & x * X
\end{align*}

En conclusión, la asignación a ejecutar dentro del lazo es
$n, r, x := n + 1, r + f.n * x, x * X$. Agregar $x = 1$ a la sección
de inicialización permite que se cumpla $P_2$ al iniciar el lazo, ya
que $X^0 = 1$. Como función de cota usamos $N-n$. El programa final se
muestra en el listado \ref{prog424}.

\begin{lstlisting}[float,caption={Solución del ejercicio 4.2.4},label=prog424]
|[
    con N, X : int {$ N \geq 0$}; f : array[0..N) of int;
    var r : int;

    var n, x : int;

    n, r, x := 0, 0, 1;
    {inv $P_0 \wedge P_1 \wedge P_2$}
    {bound $N-n$}
    do $n \neq N \rightarrow$
        n, r, x := n + 1, r + f.n * x, x * X
    od

    {$r = (\sum i : 0 \leq i < N : f.i * X^i)$}
]|
\end{lstlisting}

% Pregunta 7
\section{Ejercicio 4.3.5}

Introduzco las variables $m, x$ e $y$. $m$ representará el crédito dado
por el segmento maximal, y $x$ e $y$ denotarán un segmento maximal. La
postcondición del programa a derivar es

$$
m = (\text{{\bf max} } p, q : 0 \leq p \leq q \leq N : credit.p.q) \wedge m = credit.x.y 
$$

Voy a utilizar un lazo con la técnica de sustitución de constante por variable. Para ello,
introduzco la variable $n$ en sustitución de $N$. Las invariantes del lazo son

$$
P_0: m = (\text{{\bf max} } p, q : 0 \leq p \leq q \leq n : credit.p.q)
$$

$$
P_1: m = credit.x.y
$$

$$
P_2: 0 \leq n \leq N
$$

La instrucción de inicialización $n, m, x, y := 0, 0, 0, 0$ satisface las invariantes
ya que $credit$.$0$.$0 = 0$. Al tratar de incrementar $n$ en $1$, tenemos

\begin{align*}
 & (\text{{\bf max} } p, q : 0 \leq p \leq q \leq n + 1 : credit.p.q) \\
 = & <\text{separación de rango con } n \geq 0, P_0> \\
 & m \text{ {\bf max} } (\text{{\bf max} } p : 0 \leq p \leq n + 1 : credit.p.(n+1)) \\
\end{align*}

Aquí noto la necesidad de una variable $s$ con invariante

$$
s = (\text{{\bf max} } p : 0 \leq p \leq n + 1 : credit.p.(n+1))
$$

Ese predicado no está definido para $n = N$, lo cual es admitido
por $P_2$. Entonces defino la invariante

$$
P_3: s = (\text{{\bf max} } p : 0 \leq p \leq n : credit.p.n)
$$

que es válida para $n = N$ e intento establecer $P_3(n := n+1)$ antes de la
instrucción de actualización de $m$. Para hacerlo, dados $P_3$ y $n \neq N$,
tenemos

\begin{align*}
 & (\text{{\bf max} } p : 0 \leq p \leq n + 1 : credit.p.(n+1)) \\
 = & <\text{separación de rango con } n \geq 0> \\
 & (\text{{\bf max} } p : 0 \leq p \leq n : credit.p.(n+1)) \text{ {\bf max} } credit.(n+1).(n+1) \\
 = & <\text{crédito con rango vacío es } 0> \\
 & (\text{{\bf max} } p : 0 \leq p \leq n : credit.p.(n+1)) \text{ {\bf max} } 0 \\
\end{align*}

Al llegar a este punto me doy cuenta de que hay tres posibilidades 
dependiendo del valor de $A.n$.

\begin{itemize}
 \item Caso 0 ($A.n = 0$):

 \begin{align*}
  & (\text{{\bf max} } p : 0 \leq p \leq n : credit.p.(n+1)) \text{ {\bf max} } 0 \\
  = & <\text{definición de } credit \text{, separación, aritmética, suposición } A.n = 0> \\
  & s \text{ {\bf max} } 0 \\
 \end{align*}

 \item Caso 1 ($A.n < 0$):

 \begin{align*}
  & (\text{{\bf max} } p : 0 \leq p \leq n : credit.p.(n+1)) \text{ {\bf max} } 0 \\
  = & <\text{definición de } credit \text{, separación, aritmética, suposición } A.n = 0> \\
  & s-1 \text{ {\bf max} } 0 \\
 \end{align*}

 \item Caso 2 ($A.n > 0$):

 \begin{align*}
  & (\text{{\bf max} } p : 0 \leq p \leq n : credit.p.(n+1)) \text{ {\bf max} } 0 \\
  = & <\text{definición de } credit \text{, separación, aritmética, suposición } A.n = 0> \\
  & s+1 \text{ {\bf max} } 0 \\
 \end{align*}

\end{itemize}

Para establecer $P_3(n := n + 1)$ puedo usar, entonces, la instrucción de selección. Con ese resultado,
intento de nuevo derivar la instrucción para mantener $P_0$.

\begin{align*}
 & (\text{{\bf max} } p, q : 0 \leq p \leq q \leq n + 1 : credit.p.q) \\
 = & <\text{separación de rango con } n \geq 0, P_0> \\
 & m \text{ {\bf max} } (\text{{\bf max} } p : 0 \leq p \leq n + 1 : credit.p.(n+1)) \\
 = & <\text{suposición de que se mantiene } P_3(n := n+1)> \\
 & m \text{ {\bf max} } s \\
\end{align*}

La inicialización $s := 0$ satisface $P_3$.
Hecho esto, escribo una aproximación parcial del programa (listado \ref{prog435_parcial}). Este programa
satisface $P_0$, $P_2$ y $P_3$ pero no necesariamente $P_1$. $P_1$ se puede ver afectado por la instrucción
que actualiza $m$ en caso que $m \neq m$ {\bf max} $s$.

\begin{lstlisting}[float,caption={Primera solución parcial del ejercicio 4.3.5},label=prog435_parcial]
|[
    con N : int { $N \geq 0$ };
    var A : array[0..N) of int;

    var n, m, x, y, s : int;

    n, m, x, y, s := 0, 0, 0, 0, 0;

    do $n \neq N \rightarrow$
        if $A.n = 0 \rightarrow$ $s$ := $s$ $max$ $0$
        |  $A.n < 0 \rightarrow$ $s$ := $(s-1)$ $max$ $0$
        |  $A.n > 0 \rightarrow$ $s$ := $(s+1)$ $max$ $0$
        fi;
        $m$ := $m$ $max$ $s$
    od
]|
\end{lstlisting}

Introduzco la nueva variable $e$ y la invariante

$$
P_4: s = credit.e.n
$$

Si $m \neq m$ {\bf max} $s$, entonces $s = m$ {\bf max} $s$. Luego de la asignación quedará $m = s$.
Es claro que, si realizamos $x, y := e, n+1$, entonces $P_1$ se satisfará si suponemos que $P_4(n := n+1)$ se
mantiene.

$P_4$ se cumple tras la inicialización $e := 0$. En el programa parcial que tengo se
puede ver afectada con la modificación de $s$. Observo que $s \geq 0$ se mantiene como
condición invariante y escribo una segunda aproximación del programa para ayudarme a lograr mantener
$P_4$ (listado \ref{prog435_parcial1}).

\begin{lstlisting}[float,caption={Segunda solución parcial del ejercicio 4.3.5},label=prog435_parcial1]
|[
    con N : int { $N \geq 0$ };
    var A : array[0..N) of int;

    var n, m, x, y, s : int;

    n, m, x, y, s, e := 0, 0, 0, 0, 0, 0;

    do $n \neq N \rightarrow$
        if $A.n = 0 \rightarrow$ skip
        |  $A.n < 0 \rightarrow$ $s$ :=
               if $(s-1) \neq (s-1)$ $max$ $0 \rightarrow$ s := 0
               |  $(s-1) = (s-1)$ $max$ $0 \rightarrow$ s := s-1
               fi
        |  $A.n > 0 \rightarrow$ $s$ := $(s+1)$
        fi;
        if $m = m$ $max$ $s \rightarrow$ skip
        |  $m \neq m$ $max$ $s \rightarrow$ m, x, y := s, e, n+1
        fi
    od
]|
\end{lstlisting}

Hay dos casos en los que cambia $s$. Por una parte tenemos $A.n > 0$:

\begin{align*}
 & (credit.e.n)(n := n+1) \\
 \equiv & <\text{sustitución, definición de credit}> \\
 & (\# i : e \leq i < n+1 : A.i > 0) - (\# i : e \leq i < n+1 : A.i < 0) \\
 \equiv & <\text{separación de rangos, suposición } A.n > 0> \\
 & (\# i : e \leq i < n : A.i > 0) + 1 - (\# i : e \leq i < n : A.i < 0) \\
 \equiv & <\text{aritmética, suposición } P_4> \\
 & s + 1 \\
\end{align*}

Como realizamos la asignación $s := s + 1$ en este caso, $P_4(n := n+1)$ se mantiene.
El segundo caso se presenta cuando $A.n < 0$ y se puede dividir en dos subcasos como se
ve en el listado \ref{prog435_parcial1}. Primero tenemos $(s - 1) \neq (s - 1)$ $max$ $0$.
Poner $e := n+1$ lo resuelve ya que el crédito de un rango vacío es $0$ y en este caso
asignamos $0$ a $s$. Cuando $(s - 1) = (s - 1)$ $max$ $0$ tenemos

\begin{align*}
 & (credit.e.n)(n := n+1) \\
 \equiv & <\text{sustitución, definición de credit}> \\
 & (\# i : e \leq i < n+1 : A.i > 0) - (\# i : e \leq i < n+1 : A.i < 0) \\
 \equiv & <\text{separación de rangos, suposición } A.n < 0> \\
 & (\# i : e \leq i < n : A.i > 0) - ((\# i : e \leq i < n : A.i < 0) + 1) \\
 \equiv & <\text{aritmética, suposición } P_4> \\
 & s - 1
\end{align*}

Como realizamos la asignación $s := s - 1$ en este caso, $P_4(n := n+1)$ se mantiene. Obtengo
así el programa final, el cual se puede ver en el listado \ref{prog435}.

\begin{lstlisting}[float,caption={Solución del ejercicio 4.3.5},label=prog435]
|[
    con N : int { $N \geq 0$ };
    var A : array[0..N) of int;

    var n, m, x, y, s : int;

    n, m, x, y, s, e := 0, 0, 0, 0, 0, 0;

    do $n \neq N \rightarrow$
        if $A.n = 0 \rightarrow$ skip
        |  $A.n < 0 \rightarrow$ $s$ :=
               if $(s-1) \neq (s-1)$ $max$ $0 \rightarrow$ s := 0, e := n+1
               |  $(s-1) = (s-1)$ $max$ $0 \rightarrow$ s := s-1
               fi
        |  $A.n > 0 \rightarrow$ $s$ := $(s+1)$
        fi;
        if $m = m$ $max$ $s \rightarrow$ skip
        |  $m \neq m$ $max$ $s \rightarrow$ m, x, y := s, e, n+1
        fi
    od
]|
\end{lstlisting}

% Pregunta 8
\section{Ejercicio 4.4.2}

El caso base para G.n es $n = N$. Por rango vacío de la sumatoria tenemos

$$
G.N = 0
$$

Para el caso recursivo voy a intentar relacionar $G.(n-1)$ con $G.(n)$ 
teniendo $0 < n \leq N$.

\begin{align*}
 & G.(n-1) \\
 = & <\text{definición de G.n}> \\
 & (\sum i : n-1 \leq i < N : f.i * X^{i-(n-1)}) \\
 = & <\text{separación de rango con } n \leq N> \\
 & f.(n-1) * X^{(n-1)-(n-1)} + (\sum i : n \leq i < N : f.i * X^{i-(n-1)}) \\
 = & <\text{aritmética y definición de potencia}> \\
 & f.(n-1) + (\sum i : n \leq i < N : f.i * X^{i-n}*X) \\
 = & <\text{distributividad de multiplicación sobre suma}> \\
 & f.(n-1) + X*(\sum i : n \leq i < N : f.i * X^{i-n}) \\
 = & <\text{definición de G.n}> \\
 & f.(n-1) + X*G.n \\
\end{align*}

Viendo que $r = G$.$0$ es equivalente a la postcondición solicitada, voy a derivar el
programa usando un ciclo. Introduzco la variable $n$ y pongo como invariante

$$
 r = G.n \wedge 0 \leq n \leq N
$$

y como variante $n$. La instrucción de inicialización es $n, r := N, 0$. Al intentar
reducir $n$ en 1 y suponiendo $n \neq 0$ tenemos

\begin{align*}
 & G.(n-1) \\
 = & <\text{visto previamente}> \\
 & f.(n-1) + X*G.n \\
 = & <\text{invariante}> \\
 & f.(n-1) + X*r \\
\end{align*}

Con esto puedo escribir el programa final (listado \ref{prog442}).

\begin{lstlisting}[float,caption={Solución del ejercicio 4.4.2},label=prog442]
|[
    con N, X : int { $N \geq 0$ }; f : array[0..N) of int;

    var r, n : int;

    n, r := N, 0;

    {inv $r = G.n$}
    {bound $n$}
    do $n \neq 0 \rightarrow$
        n, r := n-1, f.(n-1) + X*r
    od
]|
\end{lstlisting}

% Pregunta 9
\section{Ejercicio 10.2.0}

Introduzco la variable $m$ para derivar un programa con postcondición

$$
0 \leq m \leq N \wedge (\forall i : 0 \leq i < m : h.i \leq 0) \wedge (\forall i : m \leq i < N : h.i \geq 0)
$$

lo que satisfaría la postcondición de existencia de $p$ dada en el enunciado
del problema. Intento derivar el programa usando un lazo y la técnica de
sustitución de constante por variable. Las invariantes del lazo serían

$$
P_0: 0 \leq m \leq n \leq N
$$

$$
P_1: (\forall i : 0 \leq i < m : h.i \leq 0) \wedge (\forall i : m \leq i < n : h.i \geq 0)
$$

La instrucción de inicialización $m, n := 0, 0$ satisface las invariantes ya que $N \geq 0$ y
las cuantificaciones universales en $P_1$ quedan con rango vacío. El guardia $B$ del lazo será
$n \neq N$. Es claro que $P_0 \wedge P_1 \wedge \neg B$ implica la postcondición.
Al incrementar $n$ en $1$ podemos considerar el elemento $h.n$. Hay dos
posibilidades (no disjuntas): $h.n \leq 0$ y $h.n \geq 0$. Se puede ver una solución parcial en
el listado \ref{prog1020_parcial}.

\begin{lstlisting}[float,caption={Solución parcial del ejercicio 10.2.0},label=prog1020_parcial]
|[
    con N : int { $N \geq 0$ };
    var h : array[0..N) of int;

    var m, n : int;

    m, n := 0, 0;

    {inv $P_0 \wedge P_1$}
    {bound $N-n$}
    do $n \neq N \rightarrow$
        if $h.n \leq 0 \rightarrow$ $S_a$
        |  $h.n \geq 0 \rightarrow$ $S_b$
        fi
    od
]|
\end{lstlisting}

Ataco primero $S_b$ por ser el caso más fácil. Tenemos

\begin{align*}
 & (P_0 \wedge P_1)(n := n + 1) \\
 \equiv & <\text{n no ocurre en } P_0\text{, sustitución y definición de } P_1> \\
 & P_0 \wedge (\forall i : 0 \leq i < m : h.i \leq 0) \wedge (\forall i : m \leq i < n+1 : h.i \geq 0) \\
 \equiv & <\text{separación de rango con } n+1 > n \geq 0> \\
 & P_0 \wedge (\forall i : 0 \leq i < m : h.i \leq 0) \wedge (\forall i : m \leq i < n : h.i \geq 0) \wedge h.n \geq 0 \\
 \equiv & <\text{definición de } P_1> \\
 & P_0 \wedge P_1 \wedge h.n \geq 0
\end{align*}

Por lo tanto, dados que la invariante y el segundo guardia
son ciertos, basta con ejecutar $n := n + 1$ para mantener la invariante con
$n$ incrementada en $1$. Para conseguir $S_a$ considero la operación
$swap.m.n$, la cual no afecta la primera parte de $P_1$.
De la segunda parte de $P_1$ se infiere que
$m < n \Rightarrow h.m \geq 0$. Como suponemos que se mantiene $P_0$, eso
es equivalente a $m = n \vee h.m \geq 0$. En el primer caso, $m = n$ y
entonces

\begin{align*}
 & \{P_0 \wedge P_1 \wedge m = n \wedge h.n \leq 0\} \\
 & swap.m.n \\
 & \{P_0 \wedge (\forall i : 0 \leq i < m : h.i \leq 0) \wedge m = n \wedge h.m \leq 0\} \\
 & ;m, n := m+1, n+1 \\
 & \{P_0 \wedge (\forall i : 0 \leq i < m : h.i \leq 0) \wedge m = n\} \\
\end{align*}

Tenemos que

\begin{align*}
 & \{P_0 \wedge (\forall i : 0 \leq i < m : h.i \leq 0) \wedge m = n\} \\
 \Rightarrow & <\text{cuantificación universal con rango vacío, true es neutro de } \wedge> \\
 & \{P_0 \wedge (\forall i : 0 \leq i < m : h.i \leq 0) \wedge (\forall i : m \leq i < n : h.i \geq 0)\} \\
 \equiv & <\text{definición de } P_1> \\
 & \{P_0 \wedge P_1\} \\
\end{align*}

Por lo tanto, si $m = n$, es admisible que $S_a = swap.m.n ; m, n := m+1, n+1$. Queda ver el
segundo caso, en el que se mantiene que $h.m \geq 0$. Sean $P_{11}$ la primera parte de $P_1$
y $P_{12}$ la segunda. Entonces

\begin{align*}
 & \{P_0 \wedge P_1 \wedge h.m \geq 0 \wedge h.n \leq 0\} \\
 & swap.m.n \\
 & \{P_0 \wedge P_{11} \wedge h.m \leq 0 \wedge (\forall i : m + 1 \leq i < n : h.i \geq 0) \wedge h.n \geq 0\} \\
 & ;m, n := m+1, n+1 \\
 & \{P_0 \wedge P_{11} \wedge P_{12}\} \\
\end{align*}

La misma instrucción que cumple con lo necesario en el primer caso también sirve en el segundo.
Habiendo entonces resuelto $S_a$ se obtiene el programa final (listado \ref{prog1020}).

\begin{lstlisting}[float,caption={Solución del ejercicio 10.2.0},label=prog1020]
|[
    con N : int { $N \geq 0$ };
    var h : array[0..N) of int;

    var m, n : int;

    m, n := 0, 0;

    {inv $P_0 \wedge P_1$}
    {bound $N-n$}
    do $n \neq N \rightarrow$
        if $h.n \leq 0 \rightarrow$ swap.m.n ; m, n := m+1, n+1
        |  $h.n \geq 0 \rightarrow$ n := n+1
        fi
    od
]|
\end{lstlisting}

\end{document}

### 2008-11-21 construccion/src/jml/kaldewaij1020_jml.java

@@ -1,0 +1,27 @@
+public class kaldewaij1020_jml {
+
+    public static void main(String args[]) {
+        int h[] = new int[] {-1,2,-3,-4,5,6,7,-2,-8,10};
+        new kaldewaij1020_jml().algoritmo(10, h);
+    }
+
+    void algoritmo(final int N, int[] h) {
+        //@ assert N >= 0;
+
+        int m = 0, n = 0;
+
+        //@ maintaining 0 <= m && m <= n && n <= N && (\forall int i ; 0 <= i && i < m ; h[i] <= 0) && (\forall int i ; m <= i && i < n ; h[i] >= 0);
+        //@ decreasing N-n;
+        while (n != N) {
+            if (h[n] <= 0) {
+                {int r = h[m]; h[m] = h[n]; h[n] = r;}
+                m = m + 1;
+                n = n + 1;
+            } else if (h[n] >= 0) {
+                n = n+1;
+            }
+        }
+
+        //@ assert (\forall int i ; 0 <= i && i < m ; h[i] <= 0) && (\forall int i ; m <= i && i < n ; h[i] >= 0);
+    }
+}

### 2008-11-21 construccion/src/jml/kaldewaij413_jml.java

@@ -1,0 +1,25 @@
+public class kaldewaij413_jml {
+
+    public static void main(String args[]) {
+        int A = 9;
+        int B = 15;
+
+        new kaldewaij413_jml().algoritmo(A, B);
+    }
+
+    void algoritmo(final int A, final int B) {
+        //@ assert A > 0 && B > 0;
+
+        int x;
+
+        x = A;
+
+        //@ maintaining 1 <= x && x <= A*B && x % A == 0 && (\forall int i ; 1 <= i && i < x && i % A == 0 ; i % B != 0);
+        //@ decreasing A*B-x;
+        while (x % B != 0) {
+            x = x + A;
+        }
+
+        //@ assert x == (\min int i ; 1 <= i && i % A == 0 && i % B == 0 ; i);
+    }
+}

### 2008-11-21 construccion/src/jml/kaldewaij424_jml.java

@@ -1,0 +1,23 @@
+public class kaldewaij424_jml {
+
+    public static void main(String args[]) {
+        int f[] = new int[] {1,2,3,4,5,6,7,8,9,0};
+        new kaldewaij424_jml().algoritmo(10, 10, f);
+    }
+
+    void algoritmo(final int N, final int X, final int[] f) {
+        //@ assert N >= 0;
+        int r = 0, n = 0, x = 1;
+
+        //@ maintaining r == (\sum int i ; 0<=i && i<n ; f[i] * (\product int j ; 0<=j && j<i ; X));
+        //@ decreasing N-n;
+        while (n != N) {
+            r = r + f[n]*x;
+            x = x*X;
+            n = n + 1;
+        }
+
+        //@ assert r == (\sum int i ; 0<=i && i<N ; f[i] * (\product int j ; 0<=j && j<i ; X+1));
+        //@ assert (\forall int i ; 0 <=i && i<f.length ; f[i] == \old(f[i]));
+    }
+}

### 2008-11-21 construccion/src/jml/kaldewaij435_jml.java

@@ -1,0 +1,62 @@
+public class kaldewaij435_jml {
+
+    static int A[] = new int[] {-42,-3,8,-14,-6,-43,-123,-4,-439,3,2,5,0,-12,2,4,34,-50,-9,-23,-2,-18,-104,-22,-40,-40,-3,-34,0,-11,-11,-3,-43,-99,-66};
+
+    public static void main(String args[]) {
+        new kaldewaij435_jml().algoritmo(A.length, A);
+    }
+
+    /* defino este metodo para facilitar la escritura de condiciones */
+    //@ ensures \result == ((\num_of int i ; p<=i && i<q ; A[i] > 0) - (\num_of int i ; p<=i && i<q ; A[i] < 0));
+    /*@ pure @*/
+    int credit(int p, int q) {
+        int c=0;
+
+        for (int i = p; i < q; i++) {
+            if (A[i] > 0)
+                c++;
+
+            if (A[i] < 0)
+                c--;
+        }
+
+        return c;
+    }
+
+    void algoritmo(final int N, final int[] A) {
+        //@ assert N >= 0;
+
+        int x = 0, y = 0, m = 0, n = 0, s = 0, e = 0;
+
+        //@ maintaining m == (\max int p ; 0 <= p && p <= n ; (\max int q ; p <= q && q <= n ; credit(p, q))) &&
+        //@             s == (\max int p ; 0 <= p && p <= n ; credit(p, n)) &&
+        //@             0<=n && n<=N &&
+        //@             m == credit(x, y);
+        //@ decreasing N-n;
+        while (n != N) {
+            if (A[n] < 0) {
+                if (s-1 < 0) {
+                    s = 0;
+                    e = n + 1;
+                } else if (s-1 >= 0) {
+                    s = s -1;
+                }
+            } else if (A[n] == 0) {
+            } else if (A[n] > 0) {
+                s = s + 1;
+            }
+
+            if (m <= s) {
+                m = s;
+                x = e;
+                y = n + 1;
+            } else if (m > s) {
+            }
+
+            n = n + 1;
+        }
+
+        //@ assert m == (\max int p ; 0 <= p && p <= N ; (\max int q ; p <= q && q <= N ; credit(p, q))) &&
+        //@        m == credit(x,y);
+    }
+}

### 2008-11-21 construccion/src/jml/kaldewaij442_jml.java

@@ -1,0 +1,28 @@
+public class kaldewaij442_jml {
+
+    public static void main(String args[]) {
+        int N = 4;
+        int X = 2;
+        int f[] = new int[] {1,2,3,4};
+
+        new kaldewaij442_jml().algoritmo(N, X, f);
+    }
+
+    void algoritmo(final int N, final int X, final int[] f) {
+        //@ assert N >= 0;
+
+        int n, r;
+
+        n = N;
+        r = 0;
+
+        //@ maintaining r == (\sum int i ; n <= i && i < N ; f[i] * (\product int j ; 0 <= j && j < i-n ; X));
+        //@ decreasing n;
+        while (n != 0) {
+            r = f[n-1] + X*r;
+            n = n-1;
+        }
+
+        //@ assert r == (\sum int i ; 0 <= i && i < N ; f[i] * (\product int j ; 0 <= j && j < i ; X));
+    }
+}

### 2008-11-25 teoria_algoritmos/exposicion/presentacion.tex
algunas laminas de introduccion de la presentacion
makefile arreglado para que funcione en ubuntu mas laminas de la presentacion

@@
   \begin{itemize}
-  \item Marco teórico.
-  \item Diseño e implementación.
-  \item Resultados.
+  \item Definición del problema.
+  \item Opciones para resolver el problema.
+  \item Técnicas para mejorar backtracking.
   \end{itemize}
@@
 {
-  \frametitle{Introducción}
+  \frametitle{Ejemplos}
 
-  Este proyecto consiste del diseño, implementación y evaluación de un
-  compilador paralelizante.
+  Problema de satisfactibilidad: ¿es posible satisfacer las siguientes expresiones?
+
+  \begin{itemize}
+  \item $v_1 \vee v_2$
+  \item $(v_1 \vee \neg v_2) \wedge (v_2 \vee v_3)$
+  \item $(v_1 \wedge \neg v_2) \wedge (v_2 \wedge v_3)$
+  \end{itemize}
+}
+
+\frame
+{
+  \frametitle{Un ejemplo más complicado}
+
+  \begin{itemize}
+  \item $(\neg v_{1} \vee v_5 \vee v_3) \wedge (\neg v_{3} \vee \neg v_{4} \vee \neg v_{6}) \wedge (\neg v_{1} \vee \neg v_{8} \vee v_{7}) \wedge$
+        $(\neg v_{4} \vee \neg v_{1} \vee v_{5}) \wedge (v_4 \vee v_2 \vee v_8) \wedge (\neg v_3 \vee \neg v_9 \vee v_1) \wedge (v_7 \vee v_4 \vee v_8)$
+  \end{itemize}
+}
+
+\frame
+{
+  \frametitle{CNF}
+
+  \begin{itemize}
+  \item CNF: ``Conjunctive normal form'' o ``Forma normal conjuntiva''
+  \item Es una forma ``estándar'' de representar fórmulas.
+  \item La CNF de una fórmula consiste en una conjunción de disyunciones $(A \vee B \vee \ldots) \wedge (C \vee D \vee \ldots) \wedge \ldots$
+  \item Cualquier expresión proposicional se puede convertir a CNF preservando
+        equivalencia o satisfactibilidad.
+  \end{itemize}
+}
+
+\frame
+{
+  \frametitle{¿Cómo resolver el problema?}
+
+  \begin{itemize}
+  \item Heurísticas
+  \item Backtracking
+  \end{itemize}
+}
+
+\section{Backtracking}
+
+\frame
+{
+  \frametitle{Motivación}
+
+  Problema de satisfactibilidad: ¿es posible satisfacer las siguientes expresiones?
+
+  \begin{itemize}
+  \item $v_1 \vee v_2$
+  \item $(v_1 \vee \neg v_2) \wedge (v_2 \vee v_3)$
+  \item $(v_1 \wedge \neg v_2) \wedge (v_2 \wedge v_3)$
+  \end{itemize}
 }

### 2008-11-25 teoria_algoritmos/exposicion/presentacion.tex
algunas mejoras

@@
 {
-  \frametitle{¿Cómo resolver el problema?}
+  \frametitle{¿Cuál es el problema y cómo resolverlo?}
+
+  Dada una fórmula $F$ con $n$ variables $X_1, \ldots, X_n$, decir si es
+  posible asignar valores a las variables de manera que $F$ se satisfaga.
+
+  Enfoques:
 
@@
 {
-  \frametitle{Motivación}
+  \frametitle{Enfoque ``fuerza bruta''}
 
-  Problema de satisfactibilidad: ¿es posible satisfacer las siguientes expresiones?
+  Usar backtracking para intentar todas las posibles asignaciones de valores a variables. Por
+  ejemplo, para dos variables $X$ e $Y$ podemos probar:
 
-  \begin{itemize}
-  \item $v_1 \vee v_2$
-  \item $(v_1 \vee \neg v_2) \wedge (v_2 \vee v_3)$
-  \item $(v_1 \wedge \neg v_2) \wedge (v_2 \wedge v_3)$
-  \end{itemize}
+  $$
+  (X=0,Y=0), (X=0,Y=1), (X=1,Y=0), (X=1,Y=1)
+  $$
+
+  Problema con este enfoque: toma tiempo exponencial ($2^n$)
 }

### 2008-11-25 teoria_algoritmos/exposicion/presentacion.tex
todo bien hasta la parte de fuerza bruta inclusive

@@
 \begin{document}
-\title{Solución exacta de problemas difíciles: Backtraking para "Satisfiability"}
+\title{Solución exacta de problemas difíciles: Backtraking para ``Satisfiability"}
 \author{Daniel Izquierdo}
@@
 
-  Dada una fórmula $F$ con $n$ variables $X_1, \ldots, X_n$, decir si es
+  Dada una fórmula $F$ en CNF con $n$ variables $X_1, \ldots, X_n$, decir si es
   posible asignar valores a las variables de manera que $F$ se satisfaga.

### 2008-11-25 teoria_algoritmos/exposicion/presentacion.tex
mas laminas

@@
   \begin{itemize}
-  \item Heurísticas
+  \item Algoritmos probabilísticos
   \item Backtracking
@@
 
+\frame
+{
+  \frametitle{BCP}
+
+  \begin{itemize}
+   \item BCP: ``Boolean Constraint Propagation'' o ``Propagación de restricciones booleanas''
+   \item Al realizar una asignación puedo deducir sus efectos: cláusulas satisfechas, cláusulas unitarias y conflictos.
+   \item Ejemplo 1: $(X \vee Y) \wedge (\neg X \vee \neg W) \wedge (W \vee U)$
+   \item Ejemplo 2: $(X \vee Y) \wedge (X \vee Z) \wedge (\neg X \vee \neg W) \wedge (W)$
+  \end{itemize}
+}
+
+\frame
+{
+  \frametitle{GRASP y exploración del espacio de búsqueda}
+
+  GRASP
+
+  Exploramos un ``árbol de deciciones'' y utilizamos técnicas
+  para incrementar la velocidad de la exploración. Un nodo del árbol
+  de decisiones representa una asignación a una variable no asignada.
+  Proceso:
+
+  \begin{enumerate}
+   \item Tomo una decisión de asignación.
+   \item ``Extiendo'' el conjunto de asignaciones actual (BCP).
+   \item Si no pude satisfacer la fórmula, me reracto decisión de asignación tomada para intentar
+         otra asignación (backtracking).
+  \end{enumerate}
+}
+
+\frame
+{
+  \frametitle{Aprendizaje}
+}
+
+\frame
+{
+  \frametitle{Backtracking no cronológico}
+}
+
 \section{Preguntas}

### 2008-11-25 teoria_algoritmos/exposicion/presentacion.tex
mas laminas, casi lista

@@
 \usepackage[utf8]{inputenc}
+\usepackage{graphicx}
 
@@
         equivalencia o satisfactibilidad.
+  \item Cada elemento de la conjunción de una fórmula CNF se denomina cláusula.
   \end{itemize}
@@
 {
-  \frametitle{GRASP y exploración del espacio de búsqueda}
+  \frametitle{Exploración del espacio de búsqueda con GRASP}
 
-  GRASP
+  GRASP es un ``SAT solver''.
 
-  Exploramos un ``árbol de deciciones'' y utilizamos técnicas
+  Exploramos un ``árbol de decisiones'' y utilizamos técnicas
   para incrementar la velocidad de la exploración. Un nodo del árbol
@@
    \item ``Extiendo'' el conjunto de asignaciones actual (BCP).
-   \item Si no pude satisfacer la fórmula, me reracto decisión de asignación tomada para intentar
-         otra asignación (backtracking).
+   \item Si no pude satisfacer la fórmula, me retracto de la decisión de
+         asignación tomada para intentar otra asignación (backtracking).
   \end{enumerate}
@@
 {
+  \frametitle{Heurísticas para la decisión de asignación}
+
+  \begin{itemize}
+   \item Se pueden utilizar heurísticas para intentar tomar una buena decisión de asignación.
+   \item Ejemplo (GRASP): elegir la asignación que satisface el mayor número de cláusulas.
+  \end{itemize}
+}
+
+\frame
+{
+  \frametitle{Análisis de conflictos}
+
+  Podemos ir más allá de BCP y averiguar sobre las causas de los conflictos,
+  aprovechando esta información para optimizar la búsqueda.
+
+  \begin{enumerate}
+   \item Aprendizaje
+   \item Backtracking no cronológico
+  \end{enumerate}
+
+}
+
+\frame
+{
   \frametitle{Aprendizaje}
+
+   \begin{enumerate}
+    \item Podemos agregar cláusulas nuevas a partir de un conflicto para evitar conseguirlo de nuevo.
+    \item Se puede emplear una heurística para decidir si agregar cada cláusula o no, de manera de no
+          aumentar demasiado el tamaño de la base de datos.
+   \end{enumerate}
+}
+
+\frame
+{
+  \frametitle{Aprendizaje (cont.)}
+
+  \scalebox{0.38}{\includegraphics{ig.png}}
 }

### 2008-11-28 teoria_algoritmos/examen3/examen3.tex
.tex del examen 3 de teoria de algoritmos

--- 2008-11-06 teoria_algoritmos/examen2/examen2.tex
@@
 \begin{document}
+
 % Título
 
-\title{Tarea 2 de Teoría de Algoritmos}
+\title{Tarea 3 de Teoría de Algoritmos}
 \author{Daniel Izquierdo \\ \#08-86809}
-\date{6 de noviembre de 2008}
+\date{2 de diciembre de 2008}
 
@@
 
-% Pregunta 2
-\section{}
-
-En ambos casos, el algoritmo una matriz $T$ donde $T[d][c]$
-es el competidor contra el cual
-jugará el competidor $c$ en el día $d$.
-Supongo que
-$matriz(n,m)$ devuelve una matriz de tamaño $n x m$. También supongo que tengo las funciones
-$unir\_horizontal$ y $unir\_vertical$ tales que si $H_1$ y $H_2$ son matrices con la misma altura
-y $V_1$ y $V_2$ son matrices con el mismo ancho, entonces
-
-$$
-unir\_horizontal(H_1, H_2) = \left[
-\begin{array}{cc}
-H_1 & H_2 \\
-\end{array}
-\right]
-$$
-
-y
-
-$$
-unir\_vertical(V_1, V_2) = \left[
-\begin{array}{c}
-V_1 \\
-V_2 \\
-\end{array}
-\right]
-$$
-
-\begin{itemize}
-
-\item Caso $n$ potencia de dos:
-
-No considero los problemas de tamaño $1$ por no tener sentido un torneo con $1$ equipo y de $0$
-días de duración. Para resolver el problema creo el procedimiento auxiliar $tabla$ donde
-$tabla(n, inicio)$ es un torneo válido para los $n$ jugadores $inicio, inicio+1, ..., inicio+n-1$.
-Entonces la solución del problema es igual a $tabla(n, 1)$.
-
-El algoritmo divide un problema de tamaño $n$ en dos sub-problemas
-de tamaño $n/2$, lo cual se puede hacer por ser $n$ potencia de dos. El caso base $n = 2$
-se resuelve en un sólo día con un sólo partido donde los dos competidores juegan entre sí.
-En el caso recursivo, tengo dos sub-problemas $T_1$ y $T_2$, donde $T_1$ es un torneo de $n/2-1$ días
-de duración para los primeros $n/2$ jugadores y $T_2$ para los últimos $n/2$ jugadores. Tras resolver
-recursivamente ambos sub-problemas debo unir los resultados para obtener la solución. Observo que ningún
-jugador que aparece en $T_1$ aparece en $T_2$ y vice-versa. Por lo tanto puedo agregar $n/2-1$ juegos
-asignando juegos entre cada jugador $x$ del primer sub-problema y los contrincantes asignados a $x+n/2$
-en el segundo sub-problema. Eso se logra con la matriz
-
-$$
-\left[
-\begin{array}{cc}
-T_1 & T_2 \\
-T_2 & T_1 \\
-\end{array}
-\right]
-$$
-
-De esta manera me queda un torneo con $n/2-2$ juegos. Se puede ver que sólo faltan los juegos
-entre $x$ y $x+n/2$ con $x$ en $T_1$, ya que $x+n/2$ no aparece como contrincante de sí mismo
-en $T_2$. Entonces puedo agregar en el último día del torneo todos los juegos con esa forma para
-obtener un resultado válido de $n/2-1$ días. El algoritmo se puede observar en el listado
-\ref{alg:ejercicio3a}.
-
-\begin{lstlisting}[float,caption={Solución al ejercicio 3},label=alg:ejercicio3a]
-procedure torneo(n)
-  return tabla(n, 1)
-
-procedure tabla(n, inicio)
-  if n = 2 do
-    T = matriz(1,2)
-
-    T[1][inicio] $\asig inicio+1$
-    T[1][inicio+1] $\asig inicio$
-    return T
-
-  m1 $\asig tabla(n/2, inicio)$
-  m2 $\asig tabla(n/2, inicio+(n/2))$
-
-  vi $\asig unir_vertical(m1, m2)$
-  vd $\asig unir_vertical(m2, m1)$
-
-  piso $\asig matriz(1,n)$
-
-  for i $\asig 1$ to $n / 2$ do
-    piso[1][i] $\asig i+(n/2)$
-    piso[1][$i+(n/2)$] $\asig i$
-
-  total $\asig unir_horizontal(vi, vd)$
-
-  return $unir_vertical(total, piso)$
-\end{lstlisting}
-
-\item Caso $n > 1$:
-
-El caso base $n = 2$
-se resuelve en un sólo día con un sólo partido donde los dos competidores juegan entre sí.
-Cuando $n$ es impar agrego un jugador ficticio nuevo y resuelvo el problema
-para los $n+1$ jugadores, lo que me dará una solución de $(n+1)-1 = n$ jugadores
-(porque $n+1$ es par). Considero entonces que los juegos entre un jugador original
-y el jugador ficticio son días libres para dicho jugador original. Me queda
-entonces por resolver el caso en que $n$ es par, lo cual voy a suponer de ahora en
-adelante.
-
-Supongamos que debo organizar el torneo para los $n$ jugadores a partir del jugador $inicio$.
-Si $n/2$ es par, entonces el resultado de los sub-problemas $<inicio,inicio+1,\ldots,inicio+n/2-1>$
-y $<inicio+n/2,inicio+n/2+1,\ldots,inicio+n-1>$ no contiene jugadores con días libres ya que es
-posible organizarlos en $n/2-1$ días. En ese caso, y de manera análoga a la unión de sub-problemas
-hecha en el caso de $n$ potencia de dos, si $T_1$ es solución del primer sub-problema y $T_2$ del segundo,
-puedo organizar $n-2$ juegos de la siguiente manera:
-
-$$
-\left[
-\begin{array}{cc}
-T_1 & T_2 \\
-T_2 & T_1 \\
-\end{array}
-\right]
-$$
-
-Al igual que ocurrió con el caso de las potencias de 2, sólo quedan por agregar
-los juegos en los que un jugador $x$ del primer sub-problema
-juega con el jugador $x+n/2$. Esto añade un día de juegos y nos queda el torneo
-completo en $n-1$ días.
-
-Resta por ver qué pasa cuando $n/2$ es impar. Si $T_1$ es solución del primer sub-problema
-y $T_2$ del segundo, tenemos que $T_1$ y $T_2$ van a tener $n/2$ días cada uno, y en cada
-día aparece un jugador libre (porque $n$ es impar). Puedo entonces organizar $n/2$ días
-de torneo uniendo ambos resultados y poniendo en cada día a jugar los dos jugadores que
-en los sub-problemas tienen ese día libre. Con esto quedan $n/2-1$ días por organizar.
-
-Observemos que los juegos hechos entre jugadores que tenían días libres en los sub-problemas
-son tales que sus participantes son $x$ y $x+n/2$ con $x$ en los jugadores de $T_1$. Esto
-porque al resolver subproblemas asignamos días libres al mismo jugador relativo a $inicio$.
-Salvo por este caso no hay juegos entre jugadores de $T_1$ y de $T_2$.
-Entonces para los $n/2-1$ días que quedan realizo juegos entre los jugadores $x$ y $x+n/2+k$
-con $x$ en $T_1$ y $k > 0$. Esto termina por completar los $n/2-1$ días que faltan.
-
-En el listado \ref{alg:ejercicio3b} se puede observar
-el algoritmo desarrollado a partir de este razonamiento.
-
-\begin{lstlisting}[caption={Solución al ejercicio 3},label=alg:ejercicio3b]
-procedure torneo(n)
-  return tabla(n, 1)
-
-procedure tabla(n, inicio)
-  if n = 2 do
-    T = matriz(1,2)
-
-    T[1][inicio] $\asig inicio+1$
-    T[1][inicio+1] $\asig inicio$
-    return T
-
-  if n es impar do
-    return tabla(n+1, inicio)
-
-  m1 $\asig tabla(n/2, inicio)$
-  m2 $\asig tabla(n/2, inicio+(n/2))$
-
-  if n/2 es par do
-    vi $\asig unir\_vertical(m1, m2)$
-    vd $\asig unir\_vertical(m2, m1)$
-
-    piso $\asig matriz(1,n)$
-
-    for i $\asig 1$ to $n / 2$ do
-      piso[1][i] $\asig i+(n/2)$
-      piso[1][$i+(n/2)$] $\asig i$
-
-    total $\asig unir\_horizontal(vi, vd)$
-
-    return $unir\_vertical(total, piso)$
-
-  techo $\asig matriz(n/2, n)$
-
-  for $jug \asig 1$ to $n/2$
-    for $dia \asig 1$ to $n/2$
-        if m1[dia][jug] = $n/2+1$ do
-          techo[dia][jug] = jug+n/2
-          techo[dia][jug+n/2] = jug
-        else do
-          techo[dia][jug] = m1[dia][jug]
-          techo[dia][jug+n/2] = m2[dia][jug]
-
-  piso $\asig matriz(n/2-1, n)$
-
-  for $x \asig 1$ to $n/2-1$ do
-    for $y \asig 1$ to $n/2$ do
-      if y+(n/2+1) $\leq$ n do
-        piso[x][y] = y+(n/2+1)
-        piso[x][y+(n/2+1)] = y
-      else do
-        piso[x][y] = n/2+1
-        piso[x][n/2+1] = y
-
-  return $unir\_vertical(techo, piso)$
-
-\end{lstlisting}
-
-\end{itemize}
-
-% Pregunta 3
-\section{}
-
-\renewcommand{\labelenumi}{(\alph{enumi})}
-\begin{enumerate}
- \item El costo promedio en comparaciones a realizar es
-
-$$
-0.1 + (0.25+0.3)*2 + (0.2+0.05+0.05+0.05)*3 = 2.25
-$$
-
-El costo del árbol en la figura 8.8 del libro es menor y, por lo tanto, mejor.
-
- \item
-
-Para construir un arbol con un conjunto de elementos se elige el elemento con mayor probabilidad
-de ser accedido y se coloca como raíz. Su sub-árbol izquierdo es el resultado de aplicar el algoritmo
-sobre el subconjunto del conjunto original formado por los elementos menores a la raíz. El derecho es el
-resultado
-de aplicar el algoritmo sobre el subconjunto de elementos mayores a la raíz. En caso de que alguno
-de los conjuntos sea vacío, entonces el resultado para ese sub-árbol es una hoja.
-
- \item
-
-El siguiente árbol es más eficiente en promedio:
-
-\Tree [.12 6 [.34 [.20 18 27 ] 35 ] ]
-
-El costo promedio en comparaciones a realizar es
-
-$$
-0.25 + (0.2+0.3)*2 + (0.1+0.05)*3 + (0.05+0.05)*4 = 2.1
-$$
-
-Podemos ver que es menor que el costo promedio del árbol en la figura 8.8 del libro.
-Esto refuerza el hecho de que es usual que un problema no pueda ser resuelto con un
-algoritmo voraz (como el utilizado para construir dicho árbol).
-
-\end{enumerate}
-
-% Pregunta 4
-\section{}
-
-Supongo que --como aparece en el libro-- la secuencia de entrada está ordenada.
-Voy a usar la definición de $C_{i,j}$ dada en el libro con la nueva condición de que
-$C_{i,j} = 0$ si $j < i$ ya que un árbol sin elementos tiene como tiempo promedio de
-búsqueda $0$. Supongamos que tenemos una solución $X$
-óptima para la secuencia de elementos entre $i$ y $j$. Sea $k$ el índice del elemento
-raíz de esa solución. El sub-árbol izquierdo contiene los elementos menores que el elemento
-$k$, es decir, los elementos con índices entre $i$ y $k-1$ inclusive. El sub-árbol derecho contiene
-los elementos mayores que el elemento $k$, es decir, los elementos con índices entre $k+1$ y $j$
-inclusive. Sean $K_i$ y $K_d$ los costos de los sub-árboles izquierdo y
-derecho de la solución respectivamente. El costo de la solución $X$ es
-
-$$
-K_i + K_d + \sum_{k=i}^j p_k
-$$
-
-ya que debemos considerar el costo de ambos sub-árboles más el costo de acceder al
-elemento raíz más el costo añadido de incrementar la profundidad de cada nodo en
-cualquiera de los sub-árboles en uno (ya que estamos agregando la raíz a altura mayor).
-Si tenemos que $K_i \neq C_{i,k-1}$, entonces una solución mejor consiste en reemplazar
-el sub-árbol $K_i$ por uno tal que su costo sea igual a $C_{i,k-1}$, el cual es el costo
-mínimo para un sub-árbol con los elementos entre $i$ y $k-1$ inclusive. Esto contradice
-la suposición de que $X$ es una solución óptima.
-
-Por lo tanto tenemos obligatoriamente que el costo de $K_i$ es igual a $C_{i,k-1}$.
-Ocurre de manera análoga con el sub-árbol derecho y un árbol de costo $C_{k+1,j}$.
-Entonces el costo de la solución $X$ debe ser
-
-$$
-C_{i,k-1} + C_{k+1,j} + \sum_{k=i}^j p_k
-$$
-
-Ahora, supongamos al absurdo que existe $m \neq k$ tal que $i \leq m \leq j$ y
-$C_{i,m-1}+C_{m+1,j} < C_{i,k-1}+C_{k+1,j}$. Entonces el árbol que contiene los
-elementos entre $i$ y $j$ inclusive y tiene como raíz el elemento $m$ tiene menor
-costo, ya que
-
-$$
-C_{i,m-1} + C_{m+1,j} + \sum_{k=i}^j p_k < C_{i,k-1} + C_{k+1,j} + \sum_{k=i}^j p_k
-$$
-
-Esto contradice el hecho de que $X$ sea una solución óptima. Con esto, $k$ debe ser tal que
-no exista índice $m$ como el descrito. Así, he probado con el principio de optimalidad
-que
-
-$$
-C_{i,j} = (\text{{\bf min} } k : i \leq k \leq j : C_{i,k-1} + C_{k+1,j}) + \sum_{k=i}^j p_k
-$$
-
-Para obtener el algoritmo de programación dinámica solicitado,
-defino las tablas $S$ y $C$, ambas de tamaño $n$. Usaré la tabla $S$ para pre-calcular las sumatorias
-
-$$
-\sum_{k=i}^j p_k
-$$
-
-para todo $i < j$. Hago esto ya que necesito esos valores para resolver el problema. La tabla $C$ contendrá
-los valores $C_{i,j}$.
-
-Para obtener la tabla $S$ utilizo el siguiente algoritmo
-
-\begin{lstlisting}[caption={Cálculo de la tabla S},label=alg:tablaS]
-for i $\asig 1$ to $n$ do
-    suma $\asig p_i$
-
-    for j $\asig i+1$ to $n$ do
-        suma $\asig suma + p_j$
-        $S_{i,j} \asig suma$
-\end{lstlisting}
-
-En el algoritmo encontramos dos ciclos, uno dentro de otro, ambos realizando $O(n)$ operaciones. Entonces
-el costo en tiempo del algoritmo es $O(n^2)$. Hecho esto utilizaré programación dinámica para calcular
-$C_{1,n}$. Recordemos que los casos para $C$ son
-
-\begin{equation*}
-C_{i,j} = 
-\begin{cases}
-p_i & \text{si $i=j$,} \\
-0 &\text{si $i>j$,} \\
-C_{i,j} = (\text{{\bf min} } k : i \leq k \leq j : C_{i,k-1} + C_{k+1,j}) + \sum_{k=i}^j p_k &\text{si $i<j$.} \\
-\end{cases}
-\end{equation*}
-
-Para empezar relleno en la tabla los casos base ($i \geq j$) e inicializo los demás valores
-con $+\infty$ (porque estoy buscando el mínimo):
-
-\begin{lstlisting}[caption={Inicialización de casos base de la tabla C},label=alg:tablaC_base]
-for i $\asig 1$ to $n$ do
-    $S_{i,i} = p_i$
-
-for i $\asig 1$ to $n$ do
-    for j $\asig 1$ to $i-1$ do
-        $S_{i,j} \asig 0$
-
-for i $\asig 1$ to $n$ do
-    for j $\asig i+1$ to $n$ do
-        $S_{i,j} \asig +\infty$
-\end{lstlisting}
-
-Una vez resueltos los
-casos base intento llenar el resto de la tabla. Hay que observar que el valor de $C_{i,j}$
-depende únicamente de $S_{i,j}$ (ya calculado) y los $C_{x,y}$ tales que ($x=i \wedge y<j$) o
-($y=j \wedge x>i$). En otras palabras, depende de los elementos que están a la izquierda en su
-fila y de los que están abajo en su columna si en la tabla tenemos índices crecientes de izquierda
-a derecha y de arriba hacia abajo.
-
-Llamemos diagonal $p$ a la diagonal descendiente que comienza en la casilla $C_{i,i+p}$. Entonces
-puedo calcular la diagonal $p$ si y sólo si todas las diagonales $q < p$ están calculadas
-ya que para una casilla
-dada de esa diagonal $C_{i,i+p}$ es necesario y suficiente que estén calculadas todas las casillas
-entre $C_{i,i-1}$ y $C_{i,i+p-1}$ y las casillas entre $C_{i+1,i+p}$ y $C_{i+p+1,i+p}$.
-
-Entonces puedo calcular progresivamente todas las diagonales desde la mitad
-hacia la derecha hasta calcular $C_{1,n}$ en la última diagonal. Esto resulta en
-el siguiente algoritmo:
-
-\begin{lstlisting}[caption={Cálculo de la tabla C},label=alg:completoC]
-for j $\asig 1$ to $n-1$ do /* diagonal j */
-    for i $\asig 1$ to $n$ do /* fila i de la diagonal j */
-        for k $\asig i$ to $i+j$ do
-            actual $\asig$ $C_{i,k-1}+C_{k+1,i+j}+S_{i,i+j}$
-            $C_{i,i+j}$=min($C_{i,i+j}$,actual)
-\end{lstlisting}
-
-El algoritmo contiene tres ciclos anidados que pueden realizar $O(n)$ operaciones. Su
-complejidad es $O(n^3)$. Habiendo hallado el valor para $C_{1,n}$ debo construir un árbol
-de búsqueda óptimo con los valores $C_{i,j}$. Observemos que dados $i, j$
-debe existir $k$ tal que $C_{i,j} - \sum_{k=i}^j p_k = C_{i,k-1} + C_{k+1,j}$.
-Es claro que el costo de un árbol con raíz en $k$ tal que ambos sub-árboles son óptimos es
-igual a $C_{i,j}$. Entonces para construir un árbol óptimo simplemente debemos hallar $k$
-y construir recursivamente el árbol izquierdo (elementos entre $i$ y $k-1$) y el derecho
-(elementos entre $k+1$ y $j$). Suponiendo que $arbol(A,I,D)$ es un árbol cuyo valor es el elemento
-de la secuencia original con índice $A$ y cuyos
-sub-árboles son $I$ y $D$, el siguiente algoritmo recursivo halla el árbol óptimo para
-$i,j$ de la manera descrita:
-
-\begin{lstlisting}[caption={Cálculo del árbol},label=alg:arbol]
-procedure optimo(i, j)
-    if i > j do
-        return hoja
-
-    if i = j do
-        return arbol(i, hoja, hoja)
-
-    for k $\asig i$ to $j$ do
-        if $C_{i,j} - S_{i,j} = C_{i,k-1} + C_{k+1,j}$ do
-            return arbol(k, optimo(i,k-1), optimo(k+1,j))
-\end{lstlisting}
-
-A continuación presento el algoritmo completo para conseguir un árbol de búsqueda óptimo.
-$abo()$ será un árbol de búsqueda óptimo para los elementos de la secuencia con índice
-entre $1$ y $n$ inclusive. Este algoritmo necesita tiempo $O(n^3)$ para correr
-sobre una secuencia de $n$ elementos. Aplicar el algoritmo a la instancia del problema 8.33
-resultó en el siguiente árbol de costo promedio 2.1:
-
-\Tree [.12 6 [.34 [.20 18 27 ] 35 ] ]
-
-\begin{lstlisting}[float,caption={Solución del problema 8.34},label=alg:completo]
-procedure optimo(i, j)
-    if i > j do
-        return hoja
-
-    if i = j do
-        return arbol(i, hoja, hoja)
-
-    for k $\asig i$ to $j$ do
-        if $C_{i,j} - S_{i,j} = C_{i,k-1} + C_{k+1,j}$ do
-            return arbol(k, optimo(i,k-1), optimo(k+1,j))
-
-procedure abo()
-    /* calculo de S */
-    for i $\asig 1$ to $n$ do
-        suma $\asig p_i$
-
-        for j $\asig i+1$ to $n$ do
-            suma $\asig suma + p_j$
-            $S_{i,j} \asig suma$
-
-    /* calculo de C */
-    for i $\asig 1$ to $n$ do
-        $S_{i,i} = p_i$
-
-    for i $\asig 1$ to $n$ do
-        for j $\asig 1$ to $i-1$ do
-            $S_{i,j} \asig 0$
-
-    for i $\asig 1$ to $n$ do
-        for j $\asig i+1$ to $n$ do
-            $S_{i,j} \asig +\infty$
-
-    for j $\asig 1$ to $n-1$ do /* diagonal j */
-        for i $\asig 1$ to $n$ do /* fila i de la diagonal j */
-            for k $\asig i$ to $i+j$ do
-                actual $\asig$ $C_{i,k-1}+C_{k+1,i+j}+S_{i,i+j}$
-                $C_{i,i+j}$=min($C_{i,i+j}$,actual)
-
-    /* calculo del arbol */
-    return optimo(1,n)
-\end{lstlisting}
-
 \end{document}

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
gran parte de la pregunta 1

@@
 
+\renewcommand{\labelenumi}{\alph{enumi})}
+
 % Pregunta 1
@@
 
+\begin{enumerate}
+
+ \item
+
+Los vértices del grafo son subconjuntos de $A$, representados por una fila
+de unos y ceros. El vértice objetivo es $A$. La búsqueda comienza a partir del vértice
+que representa el conjunto vacío. Los sucesores de un vértice que representa el conjunto
+$V$ son todos los $V \bigcup U$ tales que $U \neq \emptyset \wedge U \in F \wedge V \bigcap U = \emptyset$.
+
+Este grafo no tiene ciclos. La explicación informal del por qué es que un nodo sólo tiene
+como sucesores conjuntos con cardinalidad estrictamente mayor a la suya.
+
+Si en un momento del backtracking consigo un nodo que ya visité anteriormente, entonces no hay solución
+por ese camino. Esto se debe al hecho de que seguí buscando luego de visitarlo, lo que implica que ni él ni
+ninguno de sus sucesores es el nodo final. Por lo tanto, al llegar a un nodo visitado previamente podemos
+ignorar su rama del árbol. Un criterio de poda puede ser, entonces, eliminar ramas previamente visitadas.
+Este criterio sólo es razonable para instancias pequeñas, ya que necesitamos espacio exponencial para marcar
+los nodos como visitados (puede haber hasta $2^n$ nodos distintos para un $A$ de cardinalidad $n$).
+
+ \item
+
+\begin{tabular}{cc}
+ Comienzo (0): & \{\} \\
+ Sucesor (1):  & \{c,e,f\} \\
+ Sucesor (2): & \{a,d,g,c,e,f\} \\
+ Backtracking (1): & \{c,e,f\} \\
+ Sucesor (2): & \{a,d,c,e,f\} \\
+ Sucesor (3): & \{a,d,c,e,f,b,g\} \\
+ Objetivo (3)
+\end{tabular}
+
+\end{enumerate}
+
+% Pregunta 2
+\section{}
+
+\begin{enumerate}
+ \item 
+ \item
+ \item
+\end{enumerate}
+
+% Pregunta 3
+\section{}
+
 \end{document}

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
corrida de la pregunta 1 de teoria

@@
 que representa el conjunto vacío. Los sucesores de un vértice que representa el conjunto
-$V$ son todos los $V \bigcup U$ tales que $U \neq \emptyset \wedge U \in F \wedge V \bigcap U = \emptyset$.
+$V$ son todos los $V \bigcup U$ tales que
+
+$$
+U \neq \emptyset \wedge U \in F \wedge V \bigcap U = \emptyset \text{ }(*)
+$$
 
@@
 
-\begin{tabular}{cc}
- Comienzo (0): & \{\} \\
- Sucesor (1):  & \{c,e,f\} \\
- Sucesor (2): & \{a,d,g,c,e,f\} \\
- Backtracking (1): & \{c,e,f\} \\
- Sucesor (2): & \{a,d,c,e,f\} \\
- Sucesor (3): & \{a,d,c,e,f,b,g\} \\
- Objetivo (3)
-\end{tabular}
+\renewcommand{\labelenumii}{\arabic{enumii}.}
+
+% F={{c,e,f}, {a,d,g}, {b,c,f}, {a,d}, {b,a,g}, {b,g}}
+\begin{enumerate}
+\item Comienzo la búsqueda en el conjunto vacío $\emptyset$. \\
+      Nodo actual: $\emptyset$
+
+\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{c,e,f\}
+      Nodo actual: $\{c, e, f\}$
+
+\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{a,d,g\}
+      Nodo actual: $\{c,e,f,a,d,g\}$
+
+\item No quedan sucesores del nodo actual; me retracto de mi última decisión.
+      Nodo actual: $\{c, e, f\}$
+
+\item Elijo para añadir el segundo elemento de $F$ elegible según $(*)$: \{a,d\}
+      Nodo actual: $\{c, e, f, a, d\}$
+
+\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{b,g\}
+      Nodo actual: $\{c, e, f, a, d, b, g\}$
+
+\item Llegué al objetivo. Fin. El algoritmo responde ``sí''.
+\end{enumerate}
 

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
varias cosas de la pregunta 2

@@
 \begin{enumerate}
+
  \item
+
+Tomemos $A$ como un arreglo (sin pérdida de generalidad). En el grafo implícito, un vértice es
+una asignación parcial $(A_1, A_2)$: cada elemento en $A[0 \ldots k)$ para algún $k \geq 0$ a uno de los dos multiconjuntos.
+Se tiene una asignación válida (aunque no necesariamente óptima) cuando se llega a un nodo en que $k = n$.
+Los sucesores de un vértice no final con $k$ variables asignadas
+son los resultantes de incluir $A[k]$ en $A_1$ o en $A_2$.
+
+Supongamos que, para una instancia del problema, $(S, T)$ es una solución óptima
+y $A[0] \in S$. Entonces $(T, S)$ también es una solución óptima, ya que
+
+$$
+\left|\sum_{e \in S} e - \sum_{e \in T} e\right| = \left|\sum_{e \in T} e - \sum_{e \in S} e\right|
+$$
+
+Por lo tanto, el multiconjunto en el cual se incluye el primer elemento del arreglo no cambia la solución.
+Entonces se puede elegir como nodo inicial del grafo $(A_1 = \{A[0]\}, A_2 = \emptyset)$ y no explorar
+la asignación de $A[0]$ a $A_2$.
+
  \item
+
  \item
+
 \end{enumerate}

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
parte a de pregunta 2 lista

@@
 Tomemos $A$ como un arreglo (sin pérdida de generalidad). En el grafo implícito, un vértice es
-una asignación parcial $(A_1, A_2)$: cada elemento en $A[0 \ldots k)$ para algún $k \geq 0$ a uno de los dos multiconjuntos.
-Se tiene una asignación válida (aunque no necesariamente óptima) cuando se llega a un nodo en que $k = n$.
+una asignación parcial $(A_1, A_2)$: cada elemento en $A[0 \ldots k)$ para algún $k \geq 0$ a uno de los dos multiconjuntos
+$A_1$ o $A_2$.
+Se tiene una asignación válida (aunque no necesariamente óptima) cuando se llega a un nodo final (lo que significa que $k = n$).
 Los sucesores de un vértice no final con $k$ variables asignadas
@@
 
+Para decidir cuál es el siguiente nodo a expandir en un momento dado se toman todos los
+sucesores no visitados del nodo actual y se elige alguno que tenga la mínima cota
+inferior según la función de acotamiento. En caso de no existir ninguno, se hace backtracking.
+
  \item

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
blargh

@@
 
-Para decidir cuál es el siguiente nodo a expandir en un momento dado se toman todos los
-sucesores no visitados del nodo actual y se elige alguno que tenga la mínima cota
-inferior según la función de acotamiento. En caso de no existir ninguno, se hace backtracking.
+Para decidir cuál es el siguiente nodo a expandir en un momento dado tomaré todos los
+sucesores no visitados del nodo actual y elegiré alguno que tenga la mínima cota
+superior según la función de acotamiento. En caso de no existir ninguno, haré backtracking.
+La función de acotamiento es descrita en la siguiente sección.
 
  \item
+
+La función de acotamiento $h$ aplicada a un nodo $x = (A_1, A_2)$ en que están asignados los $k$
+primeros elementos de $A$ cumple
+
+$$
+h(x) = \left| \left(\left|\sum_{e \in A_1} e - \sum_{e \in A_2} e\right|\right) - \sum_{i=k}^n A[i] \right|
+$$
+
+Es fácil ver que eso es una cota superior para cualquier asignación alcanzable desde $x$, ya que es el resultado
+de asignar todos los elementos que quedan por decidir al
 
@@
 
+Supongo que ordenamos los tipos de objeto de la manera sugerida en la sección 9.7.2 del libro.
+El grafo implícito sobre el cual se ramifica tiene como nodos asignaciones parciales de cantidades
+de objetos de cada tipo, de la forma $x_1 \ldots x_k, 0 \leq k \leq n$ como se describe en el libro. Los sucesores
+de un nodo son aquellos que resultan de extender su asignación parcial ya sea incrementando $x_k$ en uno o
+eligiendo si se puede $x_{k+1} = 1$, siempre y cuando se satisfaga la restricción de peso de la mochila.
+Usando el mismo argumento de la sección 9.6.1 del libro, no hay pérdida de generalidad
+si no consideramos incrementar los $x_j$ con $j < k$. Un nodo final es aquel que no tiene sucesores.
+
+La función de acotamiento es
+
+Criterio de escogencia
+
+A continuación el algoritmo aplicado paso a paso:
+
+\renewcommand{\labelenumi}{\arabic{enumi}.}
+
+\begin{enumerate}
+
+\item Comienzo la búsqueda sin
+
+\end{enumerate}
+
 \end{document}

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
erm

@@
 
-Para decidir cuál es el siguiente nodo a expandir en un momento dado tomaré todos los
-sucesores no visitados del nodo actual y elegiré alguno que tenga la mínima cota
-superior según la función de acotamiento. En caso de no existir ninguno, haré backtracking.
-La función de acotamiento es descrita en la siguiente sección.
+Al expandir un nodo, agrego todos sus sucesores a una cola de prioridad donde tiene mayor prioridad un nodo
+con menor cota superior según la función de acotamiento.
+Para decidir cuál es el siguiente nodo a expandir en un momento dado tomaré el nodo de mayor prioridad
+en la cola. La función de acotamiento es descrita en la siguiente sección.
 
@@
 
-Supongo que ordenamos los tipos de objeto de la manera sugerida en la sección 9.7.2 del libro.
+Supongo que ordenamos los tipos de objeto de la manera sugerida en la sección 9.7.2 del libro y que la manera
+de resolver empates es elegir el tipo de menor peso.
 El grafo implícito sobre el cual se ramifica tiene como nodos asignaciones parciales de cantidades
@@
 de un nodo son aquellos que resultan de extender su asignación parcial ya sea incrementando $x_k$ en uno o
-eligiendo si se puede $x_{k+1} = 1$, siempre y cuando se satisfaga la restricción de peso de la mochila.
+eligiendo si se puede $x_j = 1$ con $j > k$ y $x_i = 0$ con $k < i < j$, siempre y cuando se satisfaga la 
+restricción de peso de la mochila.
 Usando el mismo argumento de la sección 9.6.1 del libro, no hay pérdida de generalidad
@@
 
-La función de acotamiento es
+La función de acotamiento inferior es
+
+$$
+\sum_{i=1}^k x_i v_i + \left( W - \sum_{i=1}^k x_i w_i \right) v_{k+1}/w_{k+1}
+$$
 
@@
 
-A continuación el algoritmo aplicado paso a paso:
+A continuación el algoritmo aplicado paso a paso. Denotaré un nodo como $[x_1 = A_1, \ldots X_k = A_k] (V)(H)$ donde
+cada $x_i = A_i$ es la asignación de cantidades de objetos de cada tipo, $V$ es el valor obtenido y $H$ es el valor
+de la función de acotamiento inferior para ese nodo.
 
@@
 
-\item Comienzo la búsqueda sin
+\item Comienzo la búsqueda con $0$ elementos de cada tipo.
+\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1] (10)()$, $[x_2=1] (5)$, $[x_3=1] (3)$, $[x_4=1] (6)$.
+\item Escojo como próxima solución actual a $[x_1=1] (10)$.
+\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1,x_2=1] (15)()$, $[x_1=1,x_3=1] (13)$.
+\item Escojo como próxima solución actual a $[x_1=1,x_2=1] (15)$.
+\item No hay más sucesores, marco $[x_1=1,x_2=1] (15)$ como mejor solución global encontrada.
 

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
pregunta 3 lista

@@
 
-La función de acotamiento inferior es
+Estoy maximizando; la función de acotamiento es
 
 $$
-\sum_{i=1}^k x_i v_i + \left( W - \sum_{i=1}^k x_i w_i \right) v_{k+1}/w_{k+1}
+\sum_{i=1}^k x_i v_i + \left( W - \sum_{i=1}^k x_i w_i \right) v_{j}/w_{j}
 $$
 
-Criterio de escogencia
+donde $j$ es el mínimo índice de un tipo de objeto igual o a la derecha de $k$ que todavía
+puede ser insertado en la mochila. Para escoger el próximo nodo a expandir en un momento dado, el criterio es
+preferir el nodo cuyo valor de función de acotamiento es mayor.
 
-A continuación el algoritmo aplicado paso a paso. Denotaré un nodo como $[x_1 = A_1, \ldots X_k = A_k] (V)(H)$ donde
-cada $x_i = A_i$ es la asignación de cantidades de objetos de cada tipo, $V$ es el valor obtenido y $H$ es el valor
-de la función de acotamiento inferior para ese nodo.
+A continuación el algoritmo aplicado paso a paso. Denotaré un nodo como $[x_1=A_1, \ldots X_k=A_k] (V)(H)$ donde
+cada $x_i = A_i$ es la asignación de cantidades de objetos de cada tipo, $V$ es el valor obtenido por esa asignación y $H$ es el valor
+de la función de acotamiento para ese nodo.
 
@@
 \item Comienzo la búsqueda con $0$ elementos de cada tipo.
-\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1] (10)()$, $[x_2=1] (5)$, $[x_3=1] (3)$, $[x_4=1] (6)$.
-\item Escojo como próxima solución actual a $[x_1=1] (10)$.
-\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1,x_2=1] (15)()$, $[x_1=1,x_3=1] (13)$.
-\item Escojo como próxima solución actual a $[x_1=1,x_2=1] (15)$.
-\item No hay más sucesores, marco $[x_1=1,x_2=1] (15)$ como mejor solución global encontrada.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$[x_1=1] (10)(10+3*(5/3))$,\\
+$[x_2=1] (5)(5+5*(5/3))$,\\
+$[x_3=1] (3)(3+6*(3/2))$,\\
+$[x_4=1] (6)(6+4*(6/4))$.
+
+\item Escojo como próxima solución actual a $[x_1=1] (10)(10+3*(5/3))$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son:\\
+$[x_1=1,x_2=1] (15)(15+0)$,\\
+$[x_1=1,x_3=1] (13+0)$.
+
+\item Escojo como próxima solución actual a $[x_1=1,x_2=1] (15)(15+0)$.
+
+\item No hay más sucesores, marco $[x_1=1,x_2=1] (15)(15+0)$ como mejor solución global encontrada.
+
+\item Por tener cotas superiores menores, elimino de la búsqueda los nodos:\\
+$[x_2=1] (5)(5+5*(5/3))$,\\
+$[x_3=1] (3)(3+6*(3/2))$,\\
+$[x_4=1] (6)(6+4*(6/4))$,\\
+$[x_1=1,x_3=1] (13+0)$.
+
+\item No quedan nodos por expandir. Una solución óptima es la mejor solución global encontrada: $[x_1=1,x_2=1] (15)(15+0)$. Fin.
 

### 2008-12-02 teoria_algoritmos/examen3/examen3.tex
casi lista pregunta 2

@@
 Al expandir un nodo, agrego todos sus sucesores a una cola de prioridad donde tiene mayor prioridad un nodo
-con menor cota superior según la función de acotamiento.
+con menor cota según la función de acotamiento.
 Para decidir cuál es el siguiente nodo a expandir en un momento dado tomaré el nodo de mayor prioridad
@@
 
-Es fácil ver que eso es una cota superior para cualquier asignación alcanzable desde $x$, ya que es el resultado
-de asignar todos los elementos que quedan por decidir al
+ \item
 
- \item
+Denotaré una nodo como $(A_0,A_1) (H)$ donde los $A_i$ son los multiconjuntos parciales que particionan a
+$A$ y $H$ es  el valor de la función de acotamiento para ese nodo.
+
+\renewcommand{\labelenumii}{\arabic{enumii}.}
+
+\begin{enumerate}
+\item Ordeno $A$ de manera no creciente.
+
+\item Comienzo la búsqueda con la solución inicial $(\{A_0\},\emptyset) ()$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1]\},\emptyset) ()$,\\
+$(\{A[0]\},\{A[1]\}) ()$.
+
+\item Escojo como próxima solución actual a $(\{A[0],A[1]\},\emptyset) ()$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[2]\},\emptyset) ()$,\\
+$(\{A[0],A[1]\},\{A[2]\}) ()$.
+
+\item Escojo como próxima solución actual a $(\{A[0],A[1]\},\{A[2]\}) ()$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[3]\},\{A[2]\}) ()$,\\
+$(\{A[0],A[1]\},\{A[2],A[3]\}) ()$.
+
+\item Escojo como próxima solución actual a $(\{A[0],A[1]\},\{A[2],A[3]\}) ()$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[3],A[4]\},\{A[2]\}) ()$,\\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4]\}) ()$.
+
+\item Escojo como próxima solución actual a\\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4]\}) ()$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[3],A[4],A[5]\},\{A[2]\}) ()$,\\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4],A[5]\}) ()$.
+
+\item Escojo como próxima solución actual a \\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4],A[5]\}) ()$.
+
+\end{enumerate}
+
+Entonces la mejor solución tiene diferencia $1$.
 
@@
 
-\item No quedan nodos por expandir. Una solución óptima es la mejor solución global encontrada: $[x_1=1,x_2=1] (15)(15+0)$. Fin.
+\item No quedan nodos por expandir. La solución óptima es la mejor solución global encontrada: $[x_1=1,x_2=1] (15)(15+0)$. Fin.
 

### 2009-04-29 matematicas/tarea1.tex
directorio de matematicas para computacion, y .tex de primera tarea

--- 2008-12-02 teoria_algoritmos/examen3/examen3.tex
@@
 
-\title{Tarea 3 de Teoría de Algoritmos}
+\title{Tarea 1 de Matemáticas para Computación}
 \author{Daniel Izquierdo \\ \#08-86809}
-\date{2 de diciembre de 2008}
+\date{30 de abril de 2009}
 
 \maketitle
-
-\renewcommand{\labelenumi}{\alph{enumi})}
 
@@
 
-\begin{enumerate}
-
- \item
-
-Los vértices del grafo son subconjuntos de $A$, representados por una fila
-de unos y ceros. El vértice objetivo es $A$. La búsqueda comienza a partir del vértice
-que representa el conjunto vacío. Los sucesores de un vértice que representa el conjunto
-$V$ son todos los $V \bigcup U$ tales que
-
-$$
-U \neq \emptyset \wedge U \in F \wedge V \bigcap U = \emptyset \text{ }(*)
-$$
-
-Este grafo no tiene ciclos. La explicación informal del por qué es que un nodo sólo tiene
-como sucesores conjuntos con cardinalidad estrictamente mayor a la suya.
-
-Si en un momento del backtracking consigo un nodo que ya visité anteriormente, entonces no hay solución
-por ese camino. Esto se debe al hecho de que seguí buscando luego de visitarlo, lo que implica que ni él ni
-ninguno de sus sucesores es el nodo final. Por lo tanto, al llegar a un nodo visitado previamente podemos
-ignorar su rama del árbol. Un criterio de poda puede ser, entonces, eliminar ramas previamente visitadas.
-Este criterio sólo es razonable para instancias pequeñas, ya que necesitamos espacio exponencial para marcar
-los nodos como visitados (puede haber hasta $2^n$ nodos distintos para un $A$ de cardinalidad $n$).
-
- \item
-
-\renewcommand{\labelenumii}{\arabic{enumii}.}
-
-% F={{c,e,f}, {a,d,g}, {b,c,f}, {a,d}, {b,a,g}, {b,g}}
-\begin{enumerate}
-\item Comienzo la búsqueda en el conjunto vacío $\emptyset$. \\
-      Nodo actual: $\emptyset$
-
-\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{c,e,f\}
-      Nodo actual: $\{c, e, f\}$
-
-\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{a,d,g\}
-      Nodo actual: $\{c,e,f,a,d,g\}$
-
-\item No quedan sucesores del nodo actual; me retracto de mi última decisión.
-      Nodo actual: $\{c, e, f\}$
-
-\item Elijo para añadir el segundo elemento de $F$ elegible según $(*)$: \{a,d\}
-      Nodo actual: $\{c, e, f, a, d\}$
-
-\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{b,g\}
-      Nodo actual: $\{c, e, f, a, d, b, g\}$
-
-\item Llegué al objetivo. Fin. El algoritmo responde ``sí''.
-\end{enumerate}
-
-\end{enumerate}
-
-% Pregunta 2
-\section{}
-
-\begin{enumerate}
-
- \item
-
-Tomemos $A$ como un arreglo (sin pérdida de generalidad). En el grafo implícito, un vértice es
-una asignación parcial $(A_1, A_2)$: cada elemento en $A[0 \ldots k)$ para algún $k \geq 0$ a uno de los dos multiconjuntos
-$A_1$ o $A_2$.
-Se tiene una asignación válida (aunque no necesariamente óptima) cuando se llega a un nodo final (lo que significa que $k = n$).
-Los sucesores de un vértice no final con $k$ variables asignadas
-son los resultantes de incluir $A[k]$ en $A_1$ o en $A_2$.
-
-Supongamos que, para una instancia del problema, $(S, T)$ es una solución óptima
-y $A[0] \in S$. Entonces $(T, S)$ también es una solución óptima, ya que
-
-$$
-\left|\sum_{e \in S} e - \sum_{e \in T} e\right| = \left|\sum_{e \in T} e - \sum_{e \in S} e\right|
-$$
-
-Por lo tanto, el multiconjunto en el cual se incluye el primer elemento del arreglo no cambia la solución.
-Entonces se puede elegir como nodo inicial del grafo $(A_1 = \{A[0]\}, A_2 = \emptyset)$ y no explorar
-la asignación de $A[0]$ a $A_2$.
-
-Al expandir un nodo, agrego todos sus sucesores a una cola de prioridad donde tiene mayor prioridad un nodo
-con menor cota superior según la función de acotamiento.
-Para decidir cuál es el siguiente nodo a expandir en un momento dado tomaré el nodo de mayor prioridad
-en la cola. La función de acotamiento es descrita en la siguiente sección.
-
- \item
-
-La función de acotamiento $h$ aplicada a un nodo $x = (A_1, A_2)$ en que están asignados los $k$
-primeros elementos de $A$ cumple
-
-$$
-h(x) = \left| \left(\left|\sum_{e \in A_1} e - \sum_{e \in A_2} e\right|\right) - \sum_{i=k}^n A[i] \right|
-$$
-
-Es fácil ver que eso es una cota superior para cualquier asignación alcanzable desde $x$, ya que es el resultado
-de asignar todos los elementos que quedan por decidir al
-
- \item
-
-\end{enumerate}
-
-% Pregunta 3
-\section{}
-
-Supongo que ordenamos los tipos de objeto de la manera sugerida en la sección 9.7.2 del libro y que la manera
-de resolver empates es elegir el tipo de menor peso.
-El grafo implícito sobre el cual se ramifica tiene como nodos asignaciones parciales de cantidades
-de objetos de cada tipo, de la forma $x_1 \ldots x_k, 0 \leq k \leq n$ como se describe en el libro. Los sucesores
-de un nodo son aquellos que resultan de extender su asignación parcial ya sea incrementando $x_k$ en uno o
-eligiendo si se puede $x_j = 1$ con $j > k$ y $x_i = 0$ con $k < i < j$, siempre y cuando se satisfaga la 
-restricción de peso de la mochila.
-Usando el mismo argumento de la sección 9.6.1 del libro, no hay pérdida de generalidad
-si no consideramos incrementar los $x_j$ con $j < k$. Un nodo final es aquel que no tiene sucesores.
-
-La función de acotamiento inferior es
-
-$$
-\sum_{i=1}^k x_i v_i + \left( W - \sum_{i=1}^k x_i w_i \right) v_{k+1}/w_{k+1}
-$$
-
-Criterio de escogencia
-
-A continuación el algoritmo aplicado paso a paso. Denotaré un nodo como $[x_1 = A_1, \ldots X_k = A_k] (V)(H)$ donde
-cada $x_i = A_i$ es la asignación de cantidades de objetos de cada tipo, $V$ es el valor obtenido y $H$ es el valor
-de la función de acotamiento inferior para ese nodo.
-
-\renewcommand{\labelenumi}{\arabic{enumi}.}
-
-\begin{enumerate}
-
-\item Comienzo la búsqueda con $0$ elementos de cada tipo.
-\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1] (10)()$, $[x_2=1] (5)$, $[x_3=1] (3)$, $[x_4=1] (6)$.
-\item Escojo como próxima solución actual a $[x_1=1] (10)$.
-\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1,x_2=1] (15)()$, $[x_1=1,x_3=1] (13)$.
-\item Escojo como próxima solución actual a $[x_1=1,x_2=1] (15)$.
-\item No hay más sucesores, marco $[x_1=1,x_2=1] (15)$ como mejor solución global encontrada.
-
-\end{enumerate}
-
 \end{document}

### 2009-04-29 matematicas/tarea1.tex
pregunta 1 de tarea 1 de matematicas

@@
 
+Para empezar, cualquier elemento $a$ del CPO es cota superior del conjunto
+vacío, ya que se satisface trivialmente que
+$(\forall b : b \in \emptyset : b \leq a)$
+porque $b \in \emptyset \equiv \mathbf{false}$.
+Supongamos que existe $s$, el supremo del conjunto vacío. Por definición,
+para toda cota superior $c$ de $\emptyset$, $s \leq c$. Pero como todos los
+elementos del CPO son cota superior, entonces para cada elemento $x$ del CPO
+tenemos
+$s \leq x$. Entonces $s$ es el menor elemento del CPO.
+
+% Pregunta 2
+\section{}
+
+% Pregunta 3
+\section{}
+
+% Pregunta 4
+\section{}
+
+% Pregunta 5
+\section{}
+
+% Pregunta 6
+\section{}
+
 \end{document}

### 2009-04-29 matematicas/tarea1.tex
preguntas 2 y 4.1 de tarea 1 de matematicas

@@
 
+Sean $a, b$ dos letras distintas de $V$ tales que $a < b$. Existe al menos la
+cadena infinitamente decreciente
+
+$$
+ab > aab > aaab > aaaab > \ldots
+$$
+
+Por lo tanto, $(V^*,\preceq)$ no es Noetheriano.
+
 % Pregunta 3
@@
 \section{}
+
+% 4.1
+\subsection{}
+
+Llamaré $P$ al conjunto deseado. Supongo que $0$ es par.
+
+\begin{itemize}
+\item Base: $0 \in P$.
+
+\item Si $x,y$ son palabras tales que
+$xy \in P \wedge x \neq \epsilon$, entonces $x0y \in P$.
+
+\item Si $x,y$ son palabras tales que
+$xy \in P \wedge y \neq \epsilon$, entonces $x1y \in P$.
+\end{itemize}
+
+Un número es par si y sólo si su representación binaria termina en cero.
+Tenemos como caso base que $0$ representa un número par sin ceros inútiles.
+Supongamos que $x,y$ son palabras tales que $xy \in P \wedge x \neq \epsilon$.
+Entonces $x0y$ es par porque si $y = \epsilon$ entonces $x0y = x0$ y, si no,
+$y$ obligatoriamente termina en cero. $x0y$ no tiene ceros inútiles porque $x$
+no puede tenerlos. Por otra parte, si $x, y$ son palabras tales que
+$xy \in P \wedge y \neq \epsilon$, entonces $x1y$ es par porque $y$
+obligatoriamente lo es. Además, si $x = \epsilon$ entonces $x1y$ = $1y$ y
+entonces no tiene ceros inútiles. En caso de que $x \neq \epsilon$, $x$ no puede
+tener ceros inútiles y por lo tanto $x1y$ tampoco los tiene. Por todo esto, el
+esquema sólo construye representaciones de números pares sin ceros inútiles.
+
+El número más pequeño que estamos considerando es $0$, representado por el
+caso base. Sea $w$ la representación de un número par sin ceros inútiles de
+longitud $n > 1$ y supongamos que el esquema puede generar todas las
+representaciones con esa propiedad de longitud $m < n$. Entonces el esquema
+también puede generar $w$: si $w = x1y$ para algún $x, y$, $y$ no puede ser
+vacío y $xy$ también es par sin ceros inútiles, de menor longitud que $w$. Si
+$w = x0y$, $x$ no puede ser vacío y $xy$ también es par sin ceros inútiles, de
+menor longitud que $w$.
+Por todo esto, el esquema puede construir todas las representaciones de números
+pares sin ceros inútiles.
+
+% 4.2
+\subsection{}
 

### 2009-04-29 matematicas/tarea1.tex
pregunta 4.2 de tarea 1 de matematicas

@@
 
+Llamaré $P$ al conjunto deseado.
+
+\begin{itemize}
+\item Base: $\epsilon \in P$.
+\item Base: $0 \in P$.
+\item Base: $1 \in P$.
+\item Si $w \in P$, entonces $0w0 \in P$.
+\item Si $w \in P$, entonces $1w1 \in P$.
+\end{itemize}
+
+$\epsilon, 0, 1$ son palíndromes. Si $w$ es palíndrome, es claro que $0w0$ y
+$1w1$ también lo son. Entonces el esquema sólo genera palíndromes. Además, se ve
+que el esquema genera todos los palíndromes de longitud menor o igual a $1$.
+Supongamos que $w$ es palíndrome de longitud $n > 1$ y el esquema genera todas
+los palíndromes de longitud $m < n$. Tenemos $w = xpx$ donde $p$ es
+una palabra de longitud menor que $n$ y $x = 0 \vee x = 1$.
+Como el esquema puede generar $p$, es claro que el esquema también puede generar
+$w$.
+
 % Pregunta 5

### 2009-04-29 matematicas/tarea1.tex
pregunta 5 de tarea 1 de matematicas

@@
 
+Primero demostraré que $(\forall e \in E :: P_0(e))$ con
+$P_0(e) \equiv |e|_( = |e|_)$. Sea $e \in E$. El
+caso base es $e \in V$. Si es así, entonces $|e|_( = 0 = |e|_)$. Si no,
+supongamos que todas las expresiones de longitud menor que la de $e$ cumplen
+$P_0$. Hay dos posibilidades:
+
+\begin{itemize}
+\item Existe $d \in E$ tal que $e = d^*$. Supusimos que $|d|_( = |d|_)$.
+Entonces
+
+$$|e|_( = |d^*|_( = |d|_( = |d|_) = |d^*|_) = |e|_)$$
+
+\item Existen $f,g \in E$ tales que $e = (f+g)$. Supusimos que
+$|f|_( = |f|_)$ y $|g|_( = |g|_)$. Entonces
+
+$$
+|e|_( = |(f+g)|_( = 2 + |f|_( + |g|_( = 2 + |f|_) + |g|_) = |(f+g)|_) = |e|_)
+$$
+
+\end{itemize}
+
+Esto termina la primera parte de la demostración. Ahora falta ver que
+$(\forall e \in E :: P_1(e))$
+con
+$P_1(e) \equiv p \text{ es prefijo de } e \Rightarrow |p|_( \geq |p|_)$.
+Es claro que para prefijos vacíos se cumple $P_1$.
+El caso base es $e \in V$. Si es así, entonces el único posible prefijo $p$ de
+$e$ es $e$ y tenemos $|p|_( = |e|_( = 0 = |e|_) = |p|_)$. Para cualquier otro
+caso, supongamos que todas las expresiones de longitud menor que la de $e$
+cumplen $P_1$. Hay dos posibilidades:
+
+\begin{itemize}
+
+\item Existe $d \in E$ tal que $e = d^*$. Los únicos prefijos posibles de $e$
+son $d^*$ y $d$. Tenemos
+
+$$|d^*|_( = |d|_( = |d|_) = |d^*|_)$$
+
+y
+
+$$|d|_( = |d|_)$$
+
+\item Existen $f,g \in E$ tales que $e = (f+g)$. Hay dos posibilidades para un
+prefijo de $e$: incluye el último paréntesis, o no. En el primer caso tenemos
+
+$$
+|(f+g)|_( = |(f+g)|_)
+$$
+
+por $(f+g) \in E$ junto con la demostración anterior. En el segundo caso podemos
+tener
+
+\begin{eqnarray*}
+& & |(f+g|_( = 1 + |f|_( + |g|_( = 1 + |f|_) + |g|_) > |f|_) + |g|_) = |(f+g|_) \\
+& & |(f+|_( = 1 + |f|_( = 1 + |f|_) > |f|_) = |(f+|_) \\
+& & |(f|_( = 1 + |f|_( = 1 + |f|_) > |f|_) = |(f|_) \\
+& & |(|_( = 1 > 0 = |(f|_) \\
+\end{eqnarray*}
+
+Con esto queda completa la demostración. Resulta
+
+$$(\forall e \in E :: P_0(e) \wedge P_1(e))$$
+
+\end{itemize}
+
 % Pregunta 6

### 2009-04-29 matematicas/tarea1.tex
pregunta 6 de tarea 1 de matematicas

@@
 
+Primero observemos que el número de hojas de un árbol binario donde cada nodo
+interno tiene exactamente dos hijos debe tener $2^p$ hojas con $p \geq 0$.
+Entonces hacemos inducción sobre los $n$ potencia de dos. El caso base es un
+árbol con 1 hoja. Aquí tenemos $0 = 1 - 1$ nodos internos y se cumple la
+propiedad a demostrar.
+
+Ahora supongamos que la propiedad se cumple para los
+árboles estudiados con $n$ hojas. Un árbol con $2n$ hojas tiene un nodo raíz
+con dos hijos y cada hijo tiene $n$ hojas. Sabemos que el número de nodos de los
+hijos es $n-1$ por hipótesis inductiva. Por lo tanto, el número de
+nodos internos del árbol es $1$ por la raíz más dos veces $n-1$, es decir,
+
+$$
+1+2(n-1) = 1 + 2n - 2 = 2n - 1
+$$
+
+Con esto queda demostrada la proposición.
+
 \end{document}

### 2009-04-29 matematicas/tarea1.tex
pregunta 3 de tarea 1 de matematicas

@@
 \section{}
+
+Sea $f : (Z^+)^n \rightarrow \mathbb{N}$ tal que
+
+$$
+f((n_0,n_1,\ldots,n_{n-1})) = p_0^{n_0} p_1^{n_1} \ldots p_{n-1}^{n_{n-1}}
+$$
+
+para $p_0,\ldots,p_{n-1}$ primos distintos. Siempre podemos hacer esto porque
+hay infinitos números primos. Sean
+$q = (q_0,\ldots,q_{n-1}), r = (r_0,\ldots,r_{n-1}) \in (Z^+)^n$ tales que
+$q \neq r$. Entonces hay al menos un $i$ tal que $q_i \neq r_i$.
+Obligatoriamente $f(q) \neq f(r)$, porque la factorización prima de ambos
+números es distinta dado que difieren en la potencia de al menos uno de sus
+divisores primos. Entonces $f$ es inyectiva, por lo que $(Z^+)^n$ es contable.
 

### 2009-05-01 matematicas/tarea1.tex
pequenas correcciones a tarea 1 de matematicas

@@
 árboles estudiados con $n$ hojas. Un árbol con $2n$ hojas tiene un nodo raíz
-con dos hijos y cada hijo tiene $n$ hojas. Sabemos que el número de nodos de los
+con dos hijos y cada hijo tiene $n$ hojas. Sabemos que el número de nodos
+internos de los
 hijos es $n-1$ por hipótesis inductiva. Por lo tanto, el número de
-nodos internos del árbol es $1$ por la raíz más dos veces $n-1$, es decir,
+nodos internos del árbol es $1$ (debido a la raíz) más dos veces $n-1$,
+es decir,
 

### 2009-05-07 matematicas/tarea2.tex
.tex de tarea 2 de matematicas

--- 2009-04-29 matematicas/tarea1.tex
@@
 
-\title{Tarea 1 de Matemáticas para Computación}
+\title{Tarea 2 de Matemáticas para Computación}
 \author{Daniel Izquierdo \\ \#08-86809}
-\date{30 de abril de 2009}
+\date{12 de mayo de 2009}
 
@@
 
-Para empezar, cualquier elemento $a$ del CPO es cota superior del conjunto
-vacío, ya que se satisface trivialmente que
-$(\forall b : b \in \emptyset : b \leq a)$
-porque $b \in \emptyset \equiv \mathbf{false}$.
-Supongamos que existe $s$, el supremo del conjunto vacío. Por definición,
-para toda cota superior $c$ de $\emptyset$, $s \leq c$. Pero como todos los
-elementos del CPO son cota superior, entonces para cada elemento $x$ del CPO
-tenemos
-$s \leq x$. Entonces $s$ es el menor elemento del CPO.
-
-% Pregunta 2
-\section{}
-
-Sean $a, b$ dos letras distintas de $V$ tales que $a < b$. Existe al menos la
-cadena infinitamente decreciente
-
-$$
-ab > aab > aaab > aaaab > \ldots
-$$
-
-Por lo tanto, $(V^*,\preceq)$ no es Noetheriano.
-
-% Pregunta 3
-\section{}
-
-% Pregunta 4
-\section{}
-
-% 4.1
-\subsection{}
-
-Llamaré $P$ al conjunto deseado. Supongo que $0$ es par.
-
-\begin{itemize}
-\item Base: $0 \in P$.
-
-\item Si $x,y$ son palabras tales que
-$xy \in P \wedge x \neq \epsilon$, entonces $x0y \in P$.
-
-\item Si $x,y$ son palabras tales que
-$xy \in P \wedge y \neq \epsilon$, entonces $x1y \in P$.
-\end{itemize}
-
-Un número es par si y sólo si su representación binaria termina en cero.
-Tenemos como caso base que $0$ representa un número par sin ceros inútiles.
-Supongamos que $x,y$ son palabras tales que $xy \in P \wedge x \neq \epsilon$.
-Entonces $x0y$ es par porque si $y = \epsilon$ entonces $x0y = x0$ y, si no,
-$y$ obligatoriamente termina en cero. $x0y$ no tiene ceros inútiles porque $x$
-no puede tenerlos. Por otra parte, si $x, y$ son palabras tales que
-$xy \in P \wedge y \neq \epsilon$, entonces $x1y$ es par porque $y$
-obligatoriamente lo es. Además, si $x = \epsilon$ entonces $x1y$ = $1y$ y
-entonces no tiene ceros inútiles. En caso de que $x \neq \epsilon$, $x$ no puede
-tener ceros inútiles y por lo tanto $x1y$ tampoco los tiene. Por todo esto, el
-esquema sólo construye representaciones de números pares sin ceros inútiles.
-
-El número más pequeño que estamos considerando es $0$, representado por el
-caso base. Sea $w$ la representación de un número par sin ceros inútiles de
-longitud $n > 1$ y supongamos que el esquema puede generar todas las
-representaciones con esa propiedad de longitud $m < n$. Entonces el esquema
-también puede generar $w$: si $w = x1y$ para algún $x, y$, $y$ no puede ser
-vacío y $xy$ también es par sin ceros inútiles, de menor longitud que $w$. Si
-$w = x0y$, $x$ no puede ser vacío y $xy$ también es par sin ceros inútiles, de
-menor longitud que $w$.
-Por todo esto, el esquema puede construir todas las representaciones de números
-pares sin ceros inútiles.
-
-% 4.2
-\subsection{}
-
-Llamaré $P$ al conjunto deseado.
-
-\begin{itemize}
-\item Base: $\epsilon \in P$.
-\item Base: $0 \in P$.
-\item Base: $1 \in P$.
-\item Si $w \in P$, entonces $0w0 \in P$.
-\item Si $w \in P$, entonces $1w1 \in P$.
-\end{itemize}
-
-$\epsilon, 0, 1$ son palíndromes. Si $w$ es palíndrome, es claro que $0w0$ y
-$1w1$ también lo son. Entonces el esquema sólo genera palíndromes. Además, se ve
-que el esquema genera todos los palíndromes de longitud menor o igual a $1$.
-Supongamos que $w$ es palíndrome de longitud $n > 1$ y el esquema genera todas
-los palíndromes de longitud $m < n$. Tenemos $w = xpx$ donde $p$ es
-una palabra de longitud menor que $n$ y $x = 0 \vee x = 1$.
-Como el esquema puede generar $p$, es claro que el esquema también puede generar
-$w$.
-
-% Pregunta 5
-\section{}
-
-Primero demostraré que $(\forall e \in E :: P_0(e))$ con
-$P_0(e) \equiv |e|_( = |e|_)$. Sea $e \in E$. El
-caso base es $e \in V$. Si es así, entonces $|e|_( = 0 = |e|_)$. Si no,
-supongamos que todas las expresiones de longitud menor que la de $e$ cumplen
-$P_0$. Hay dos posibilidades:
-
-\begin{itemize}
-\item Existe $d \in E$ tal que $e = d^*$. Supusimos que $|d|_( = |d|_)$.
-Entonces
-
-$$|e|_( = |d^*|_( = |d|_( = |d|_) = |d^*|_) = |e|_)$$
-
-\item Existen $f,g \in E$ tales que $e = (f+g)$. Supusimos que
-$|f|_( = |f|_)$ y $|g|_( = |g|_)$. Entonces
-
-$$
-|e|_( = |(f+g)|_( = 2 + |f|_( + |g|_( = 2 + |f|_) + |g|_) = |(f+g)|_) = |e|_)
-$$
-
-\end{itemize}
-
-Esto termina la primera parte de la demostración. Ahora falta ver que
-$(\forall e \in E :: P_1(e))$
-con
-$P_1(e) \equiv p \text{ es prefijo de } e \Rightarrow |p|_( \geq |p|_)$.
-Es claro que para prefijos vacíos se cumple $P_1$.
-El caso base es $e \in V$. Si es así, entonces el único posible prefijo $p$ de
-$e$ es $e$ y tenemos $|p|_( = |e|_( = 0 = |e|_) = |p|_)$. Para cualquier otro
-caso, supongamos que todas las expresiones de longitud menor que la de $e$
-cumplen $P_1$. Hay dos posibilidades:
-
-\begin{itemize}
-
-\item Existe $d \in E$ tal que $e = d^*$. Los únicos prefijos posibles de $e$
-son $d^*$ y $d$. Tenemos
-
-$$|d^*|_( = |d|_( = |d|_) = |d^*|_)$$
-
-y
-
-$$|d|_( = |d|_)$$
-
-\item Existen $f,g \in E$ tales que $e = (f+g)$. Hay dos posibilidades para un
-prefijo de $e$: incluye el último paréntesis, o no. En el primer caso tenemos
-
-$$
-|(f+g)|_( = |(f+g)|_)
-$$
-
-por $(f+g) \in E$ junto con la demostración anterior. En el segundo caso podemos
-tener
-
-\begin{eqnarray*}
-& & |(f+g|_( = 1 + |f|_( + |g|_( = 1 + |f|_) + |g|_) > |f|_) + |g|_) = |(f+g|_) \\
-& & |(f+|_( = 1 + |f|_( = 1 + |f|_) > |f|_) = |(f+|_) \\
-& & |(f|_( = 1 + |f|_( = 1 + |f|_) > |f|_) = |(f|_) \\
-& & |(|_( = 1 > 0 = |(f|_) \\
-\end{eqnarray*}
-
-Con esto queda completa la demostración. Resulta
-
-$$(\forall e \in E :: P_0(e) \wedge P_1(e))$$
-
-\end{itemize}
-
-% Pregunta 6
-\section{}
-
-Primero observemos que el número de hojas de un árbol binario donde cada nodo
-interno tiene exactamente dos hijos debe tener $2^p$ hojas con $p \geq 0$.
-Entonces hacemos inducción sobre los $n$ potencia de dos. El caso base es un
-árbol con 1 hoja. Aquí tenemos $0 = 1 - 1$ nodos internos y se cumple la
-propiedad a demostrar.
-
-Ahora supongamos que la propiedad se cumple para los
-árboles estudiados con $n$ hojas. Un árbol con $2n$ hojas tiene un nodo raíz
-con dos hijos y cada hijo tiene $n$ hojas. Sabemos que el número de nodos de los
-hijos es $n-1$ por hipótesis inductiva. Por lo tanto, el número de
-nodos internos del árbol es $1$ por la raíz más dos veces $n-1$, es decir,
-
-$$
-1+2(n-1) = 1 + 2n - 2 = 2n - 1
-$$
-
-Con esto queda demostrada la proposición.
-
 \end{document}

### 2009-05-08 matematicas/tarea2.tex
pregunta 1 de tarea 2 de matematicas

@@
 
+Supongamos que $F$ no termina. Entonces existe una cadena
+infinita de llamadas
+
+$$
+P_0: F(x_0, y_0) \rightarrow F(x_1, y_1) \rightarrow F(x_2, y_2) \rightarrow \ldots
+$$
+
+donde defino $\rightarrow$ como ``llama a''.
+Aquí, no puede ser que haya $x_i$ tal que $x_i = 0$ o $y_i$ tal que $y_i = 0$
+porque
+en estos casos $F$ no hace llamadas recursivas.
+%Observemos que hay una
+%correspondencia uno a uno $c$ entre llamadas a $F$ y el producto cartesiano de
+%los naturales, de manera que $c(F(x,y)) = (x,y)$.
+Observemos que cuando
+$F(x,y) \rightarrow F(u,v)$, tenemos dos posibilidades:
+
+\begin{itemize}
+\item $(u,v) = (x-1,r)$ para algún $r$. Tenemos por definición del orden
+lexicográfico que $(x,y) \succeq (x-1,r)$ porque $x-1 < x$.
+\item $(u,v) = (x,y-1)$. Tenemos también por definición del orden lexicográfico
+que $(x,y) \succeq (x,y-1)$ porque $x = x \wedge y-1 < y$.
+\end{itemize}
+
+De esta manera, $F(x,y) \rightarrow F(u,v)$ implica que $(x,y) \succeq (u,v)$.
+Luego, la existencia de $P_0$ implica que existe una cadena infinita
+
+$$
+(x_0, y_0) \succeq (x_1, y_1) \succeq (x_2, y_2) \succeq \ldots
+$$
+
+lo cual es falso porque el orden es Noetheriano. Esta contradicción nos lleva a
+concluir que $F$ termina.
+
+% Pregunta 2
+\section{}
+
+% Pregunta 3
+\section{}
+
+% Pregunta 4
+\section{}
+
+% Pregunta 5
+\section{}
+
+% Pregunta 6
+\section{}
+
 \end{document}

### 2009-05-09 matematicas/tarea2.tex
avance de pregunta 2 de tarea 2 de matematicas

@@
 \newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}
+\newcommand{\piso}[1]{\lfloor #1 \rfloor}
 
@@
 
+Primero quiero ver qué debe cumplir $d$ para que $f(n) \leq dn$ siga de suponer
+que $n$ es suficientemente grande y $f(m) \leq dm$ para todo $m$ menor que $n$
+y también suficientemente grande. Si hacemos tal suposición y tenemos
+$\piso{pn}$ y $\piso{qn}$ suficientemente grandes, entonces necesitamos
+$f(n) \leq dn$ y tenemos
+
+\begin{align*}
+ & f(n) \\
+ = & \\
+ & f(\piso{pn}) + f(\piso{qn}) + bn \\
+ \leq & \\
+ & d\piso{pn} + d\piso{qn} + bn \\
+ \leq & \\
+ & dpn + dqn + bn \\
+ = & \\
+ & dn(p+q) + bn \\
+ = & \\
+ & n(d(p+q) + b) \\
+\end{align*}
+
+Para poder concluir que $f(n) \leq dn$ necesitamos que
+
+\begin{align*}
+ & n(d(p+q) + b) \leq dn \\
+ \equiv & \\
+ & d(p+q) + b \leq d \\
+ \equiv & \\
+ & b \leq d - d(p+q) \\
+ \equiv & \\
+ & b \leq d(1 - (p+q)) \\
+ \equiv & \\
+ & \frac{b}{1-(p+q)} \leq d \\
+\end{align*}
+
+El término de la izquierda en la última ecuación está definido y es positivo
+porque $p+q$ es positivo y menor que $1$ y $b$ es positivo.
+
 % Pregunta 3

### 2009-05-12 matematicas/tarea2.tex
pregunta 3 de tarea 2 de matematicas

@@
 \newcommand{\piso}[1]{\lfloor #1 \rfloor}
+\newcommand{\techo}[1]{\lceil #1 \rceil}
 
@@
 El término de la izquierda en la última ecuación está definido y es positivo
-porque $p+q$ es positivo y menor que $1$ y $b$ es positivo.
+porque $p+q$ es positivo y menor que $1$ y $b$ es positivo. Entonces tenemos
+que, bajo esa condición, $f(n) \leq dn$ es cierto si
+$f(\piso{pn}) \leq d\piso{pn}$ y $f(\piso{qn}) \leq d\piso{qn}$ también lo son.
+Para culminar la prueba buscaré números apropiados como base para aplicar
+inducción.
 
@@
 \section{}
+
+Estas funciones tienen la propiedad de que $\piso{f(x)} = \piso{f(\techo{x}}$.
+La prueba es similar a la dada en el libro para demostrar que
+$\techo{f(x)} = \techo{f(\techo{x})}$.
+Para todo $x$ entero se cumple trivialmente porque $x = \techo{x}$. Si $x$ no
+es entero, tenemos $x < \techo{x}$. Como $f$ es monótona decreciente, esto
+implica que $f(x) > f(\techo{x})$. Entonces
+$\piso{f(x)} \geq \piso{f(\techo{x})}$.
+Cuando es $\piso{f(x)} > \piso{f(\techo{x})}$, debe haber $y$ tal que
+$x \leq y < \techo{x}$ y $f(y) = \piso{f(x)}$ porque $f$ es continua.
+$y$ debe
+ser entero por la propiedad
+de $f$ de tomar valores enteros sólo con argumentos enteros.
+Pero no puede haber enteros $i$ tales que $x \leq i < \techo{x}$. 
+Con esto llegamos
+a una contradicción.
+Entonces debe ser $\piso{f(x)} = \piso{f(\techo{x})}$.
 
@@
 
-% Pregunta 6
-\section{}
-
 \end{document}

### 2009-05-12 matematicas/tarea2.tex
varias cosas de tarea 2 de matematicas

@@
 \newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}
-\newcommand{\piso}[1]{\lfloor #1 \rfloor}
-\newcommand{\techo}[1]{\lceil #1 \rceil}
+\newcommand{\piso}[1]{\left \lfloor #1 \right \rfloor}
+\newcommand{\techo}[1]{\left \lceil #1 \right \rceil}
 
@@
 $f(\piso{pn}) \leq d\piso{pn}$ y $f(\piso{qn}) \leq d\piso{qn}$ también lo son.
-Para culminar la prueba buscaré números apropiados como base para aplicar
+Para culminar la prueba necesito números apropiados como base para aplicar
 inducción.
@@
 
+Tenemos
+
+\begin{align*}
+ & \techo{\frac{n}{m}} = \piso{\frac{n+m-1}{m}} \\
+ \equiv & \\
+ & \techo{\frac{n}{m}} - \piso{\frac{n}{m}} = \piso{\frac{n+m-1}{m}} - \piso{\frac{n}{m}} \\
+ \equiv & \langle \piso{\frac{n}{m}} \text{ es entero} \rangle \\
+ & \techo{\frac{n}{m} - \piso{\frac{n}{m}}} = \piso{\frac{n+m-1}{m} - \piso{\frac{n}{m}}} \\
+ \equiv & \\
+ & \techo{\frac{n - m\piso{\frac{n}{m}}}{m}} = \piso{\frac{n+m-1 - m\piso{\frac{n}{m}}}{m}} \\
+ \equiv & \langle m \text{ positivo} \rangle \\
+ & \techo{\frac{n \mathtt{mod} m}{m}} = \piso{\frac{n \mathtt{mod} m +m-1}{m}} \\
+\end{align*}
+
+Esto último es equivalente a $n \mathtt{mod} m > 0$, porque debe ser
+$0 \leq n \mathtt{mod} m < m$.
+Entonces es equivalente a $m > 0$. Por lo tanto, $m > 0$ implica la igualdad que se quería
+demostrar.
+
 % Pregunta 5
@@
 
+Sea $N = \frac{3^77-1}{2}$. $N$ es impar porque 
+$N$ es compuesto porque es divisible entre $\frac{3^7-1}{2}$:
+
 \end{document}

### 2009-05-13 matematicas/tarea2.tex
terminada pregunta 5 de tarea 2 de matematicas

@@
 
-Sea $N = \frac{3^77-1}{2}$. $N$ es impar porque 
-$N$ es compuesto porque es divisible entre $\frac{3^7-1}{2}$:
+Sea $N = \frac{3^{77}-1}{2}$. Sabemos que $3^{77}$ tiene la forma
+$3^{2k+1}$ y que $3^{2k+1} \mathtt{mod} 4 = 3$. Entonces
+$3^{2k+1} - 1 \mathtt{mod} 4 = 2$, lo que nos dice que es $3^{77}-1$ es
+divisible entre $2$ pero no entre $4$. Al dividirlo entre $2$, el número resultante
+no puede ser vuelto a dividir entre $2$. Entonces es $\frac{3^{77}-1}{2}$ es impar.
+
+Un número de forma $3^n-1$ no puede ser primo si $n$ es compuesto, porque
+$3^{km}-1$ tiene como factor a $3^m-1$. Esto se puede ver de manera similar a
+la expuesta en el libro para mostrar que $2^{km}-1$ tiene a $2^m-1$ como factor:
+
+$$
+3^{km}-1 = (3^m-1)(3^{m(k-1)}+3^{m(k-2)}+\ldots+1)
+$$
+
+Entonces $N$ es compuesto porque es divisible al menos entre $\frac{3^7-1}{2}$
+y $\frac{3^{11}-1}{2}$ porque $77 = 7*11$.
 

### 2009-05-13 matematicas/tarea2.tex
parece terminada pregunta 2 de tarea 2 de matematicas

@@
 $f(\piso{pn}) \leq d\piso{pn}$ y $f(\piso{qn}) \leq d\piso{qn}$ también lo son.
+
 Para culminar la prueba necesito números apropiados como base para aplicar
-inducción.
+inducción. Pero observemos: sea $M$ el máximo valor tomado por $f(n)$ con
+$n \leq n_0$, que existe porque el dominio de la función es el conjunto de los
+naturales. Sea $T$ el máximo entre $M+1$ y $\frac{b}{1-(p+q)}$. Entonces se cumple
+
+$$
+f(n) \leq Tn
+$$
+
+para todo $n \leq n_0$ por la manera como elegimos $T$ (a partir de $M$). Además,
+como $T$ cumple la condición suficiente para $d$ que conseguimos anteriormente, entonces
+$f(n)$ sigue para todos los $n > n_0$ tomando $d = T$. Si se quiere, se puede
+tomar como base $n_0$.
 

### 2009-05-28 bdhetero/tarea2/tarea2.tex
esqueleto de tarea 2 de bdhetero

--- 2008-12-02 teoria_algoritmos/examen3/examen3.tex
@@
-\documentclass{article}
+\documentclass[12pt]{article}
 
-\usepackage{algorithm}
-\usepackage{algorithmic}
 \usepackage{amsmath}
@@
 \usepackage[utf8]{inputenc}
-\usepackage{listings}
 \usepackage[spanish]{babel}
-\usepackage{qtree}
-
-% Para codigo bonito
-\newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}
-
-\newcommand{\algoritmo}[2]
-{
-{
-  \singlespace
-  \begin{algorithm}[h]
-    \caption{#1}
-    \begin{algorithmic}[1]
-      #2
-    \end{algorithmic}
-  \end{algorithm}
-}
-}
-
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-\lstset{mathescape=true,frame=single}
 
 \newcommand{\asig}{\ensuremath{\leftarrow}}
-\renewcommand{\lstlistingname}{Listado} %listings en español
 
@@
 
-\title{Tarea 3 de Teoría de Algoritmos}
+\title{Bases de Datos Heterogéneas \\ Tarea 2}
 \author{Daniel Izquierdo \\ \#08-86809}
-\date{2 de diciembre de 2008}
 
@@
 
-\renewcommand{\labelenumi}{\alph{enumi})}
+\renewcommand{\labelenumi}{\alph{enumi}.}
+\begin{enumerate}
 
 % Pregunta 1
-\section{}
-
-\begin{enumerate}
-
- \item
-
-Los vértices del grafo son subconjuntos de $A$, representados por una fila
-de unos y ceros. El vértice objetivo es $A$. La búsqueda comienza a partir del vértice
-que representa el conjunto vacío. Los sucesores de un vértice que representa el conjunto
-$V$ son todos los $V \bigcup U$ tales que
-
-$$
-U \neq \emptyset \wedge U \in F \wedge V \bigcap U = \emptyset \text{ }(*)
-$$
-
-Este grafo no tiene ciclos. La explicación informal del por qué es que un nodo sólo tiene
-como sucesores conjuntos con cardinalidad estrictamente mayor a la suya.
-
-Si en un momento del backtracking consigo un nodo que ya visité anteriormente, entonces no hay solución
-por ese camino. Esto se debe al hecho de que seguí buscando luego de visitarlo, lo que implica que ni él ni
-ninguno de sus sucesores es el nodo final. Por lo tanto, al llegar a un nodo visitado previamente podemos
-ignorar su rama del árbol. Un criterio de poda puede ser, entonces, eliminar ramas previamente visitadas.
-Este criterio sólo es razonable para instancias pequeñas, ya que necesitamos espacio exponencial para marcar
-los nodos como visitados (puede haber hasta $2^n$ nodos distintos para un $A$ de cardinalidad $n$).
-
- \item
-
-\renewcommand{\labelenumii}{\arabic{enumii}.}
-
-% F={{c,e,f}, {a,d,g}, {b,c,f}, {a,d}, {b,a,g}, {b,g}}
-\begin{enumerate}
-\item Comienzo la búsqueda en el conjunto vacío $\emptyset$. \\
-      Nodo actual: $\emptyset$
-
-\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{c,e,f\}
-      Nodo actual: $\{c, e, f\}$
-
-\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{a,d,g\}
-      Nodo actual: $\{c,e,f,a,d,g\}$
-
-\item No quedan sucesores del nodo actual; me retracto de mi última decisión.
-      Nodo actual: $\{c, e, f\}$
-
-\item Elijo para añadir el segundo elemento de $F$ elegible según $(*)$: \{a,d\}
-      Nodo actual: $\{c, e, f, a, d\}$
-
-\item Elijo para añadir el primer elemento de $F$ elegible según $(*)$: \{b,g\}
-      Nodo actual: $\{c, e, f, a, d, b, g\}$
-
-\item Llegué al objetivo. Fin. El algoritmo responde ``sí''.
-\end{enumerate}
-
-\end{enumerate}
-
-% Pregunta 2
-\section{}
-
-\begin{enumerate}
-
- \item
-
-Tomemos $A$ como un arreglo (sin pérdida de generalidad). En el grafo implícito, un vértice es
-una asignación parcial $(A_1, A_2)$: cada elemento en $A[0 \ldots k)$ para algún $k \geq 0$ a uno de los dos multiconjuntos
-$A_1$ o $A_2$.
-Se tiene una asignación válida (aunque no necesariamente óptima) cuando se llega a un nodo final (lo que significa que $k = n$).
-Los sucesores de un vértice no final con $k$ variables asignadas
-son los resultantes de incluir $A[k]$ en $A_1$ o en $A_2$.
-
-Supongamos que, para una instancia del problema, $(S, T)$ es una solución óptima
-y $A[0] \in S$. Entonces $(T, S)$ también es una solución óptima, ya que
-
-$$
-\left|\sum_{e \in S} e - \sum_{e \in T} e\right| = \left|\sum_{e \in T} e - \sum_{e \in S} e\right|
-$$
-
-Por lo tanto, el multiconjunto en el cual se incluye el primer elemento del arreglo no cambia la solución.
-Entonces se puede elegir como nodo inicial del grafo $(A_1 = \{A[0]\}, A_2 = \emptyset)$ y no explorar
-la asignación de $A[0]$ a $A_2$.
-
-Al expandir un nodo, agrego todos sus sucesores a una cola de prioridad donde tiene mayor prioridad un nodo
-con menor cota superior según la función de acotamiento.
-Para decidir cuál es el siguiente nodo a expandir en un momento dado tomaré el nodo de mayor prioridad
-en la cola. La función de acotamiento es descrita en la siguiente sección.
-
- \item
-
-La función de acotamiento $h$ aplicada a un nodo $x = (A_1, A_2)$ en que están asignados los $k$
-primeros elementos de $A$ cumple
-
-$$
-h(x) = \left| \left(\left|\sum_{e \in A_1} e - \sum_{e \in A_2} e\right|\right) - \sum_{i=k}^n A[i] \right|
-$$
-
-Es fácil ver que eso es una cota superior para cualquier asignación alcanzable desde $x$, ya que es el resultado
-de asignar todos los elementos que quedan por decidir al
-
  \item
@@
 
-% Pregunta 3
-\section{}
-
-Supongo que ordenamos los tipos de objeto de la manera sugerida en la sección 9.7.2 del libro y que la manera
-de resolver empates es elegir el tipo de menor peso.
-El grafo implícito sobre el cual se ramifica tiene como nodos asignaciones parciales de cantidades
-de objetos de cada tipo, de la forma $x_1 \ldots x_k, 0 \leq k \leq n$ como se describe en el libro. Los sucesores
-de un nodo son aquellos que resultan de extender su asignación parcial ya sea incrementando $x_k$ en uno o
-eligiendo si se puede $x_j = 1$ con $j > k$ y $x_i = 0$ con $k < i < j$, siempre y cuando se satisfaga la 
-restricción de peso de la mochila.
-Usando el mismo argumento de la sección 9.6.1 del libro, no hay pérdida de generalidad
-si no consideramos incrementar los $x_j$ con $j < k$. Un nodo final es aquel que no tiene sucesores.
-
-La función de acotamiento inferior es
-
-$$
-\sum_{i=1}^k x_i v_i + \left( W - \sum_{i=1}^k x_i w_i \right) v_{k+1}/w_{k+1}
-$$
-
-Criterio de escogencia
-
-A continuación el algoritmo aplicado paso a paso. Denotaré un nodo como $[x_1 = A_1, \ldots X_k = A_k] (V)(H)$ donde
-cada $x_i = A_i$ es la asignación de cantidades de objetos de cada tipo, $V$ es el valor obtenido y $H$ es el valor
-de la función de acotamiento inferior para ese nodo.
-
-\renewcommand{\labelenumi}{\arabic{enumi}.}
-
-\begin{enumerate}
-
-\item Comienzo la búsqueda con $0$ elementos de cada tipo.
-\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1] (10)()$, $[x_2=1] (5)$, $[x_3=1] (3)$, $[x_4=1] (6)$.
-\item Escojo como próxima solución actual a $[x_1=1] (10)$.
-\item Expando mi solución actual. Las nuevas posibilidades son: $[x_1=1,x_2=1] (15)()$, $[x_1=1,x_3=1] (13)$.
-\item Escojo como próxima solución actual a $[x_1=1,x_2=1] (15)$.
-\item No hay más sucesores, marco $[x_1=1,x_2=1] (15)$ como mejor solución global encontrada.
-
-\end{enumerate}
-
 \end{document}

### 2009-05-28 bdhetero/proyecto1/informe.tex
introducción del informe del proyecto1 de bdhetero

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}

\newcommand{\asig}{\ensuremath{\leftarrow}}

\begin{document}

% Título

\title{Bases de Datos Heterogéneas \\ Proyecto 1 \\ Informe}
\author{Daniel Izquierdo \\ \#08-86809}
\date{28 de mayo de 2009}

\maketitle

\section{Introducción}

Una consulta sobre una base de datos generalmente puede ser evaluada o
implementada de varias maneras diferentes. Cada manera de evaluarla implica
costos de ejecución distintos. Puede haber grandes diferencias en los costos de
distintos planes de evaluación para la misma consulta, y puede haber un número
inmenso de planes distintos. Para cada consulta se desearía conocer su mejor
plan de evaluación en términos de recursos a consumir, pero esto usualmente no
es factible. Se intenta entonces encontrar un plan relativamente bueno, cuyo
costo de ejecución se aleje del peor costo posible.

Un optimizador es un programa que toma una consulta y produce el plan de
evaluación de menor costo o uno de costo relativamente bajo.
En este proyecto se diseñarán e implementarán optimizadores para ciertos tipos
de consultas. Se está considerando un subconjunto SQL' de las consultas
expresables en el lenguaje SQL. Los optimizadores usarán las meta-heurísticas de
programación dinámica y \emph{Simulated Annealing}.

\section{Diseño de la solución}

\section{Almacenamiento del catálogo}

\section{Estudio experimental}

\section{Conclusión}

\end{document}

### 2009-06-15 bdhetero/tarea2/tarea2.tex
pregunta a de tarea2 de bdhetero

@@
 
-% Pregunta 1
+% Pregunta a
+ \item
+
+\Tree
+[.$\Pi_{\text{PELICULA.NOMBRE}}$
+  [.{$\sigma$ \\
+   PRODUCE.PRESUPUESTO $ = 1000000$ \\
+   PRODUCE.AÑO $ = 2004$ {\bf OR} PRODUCE.AÑO $ = 2005$ \\
+   COMPAÑIA.NOMBRE $ = $ ``MGM'' \\
+   PELICULA.COD\_PELICULA $ = $ PRODUCE.COD\_PELICULA \\
+   PRODUCE.COD\_COMPAÑIA $ = $ COMPAÑIA.COD\_COMPAÑIA \\
+   }
+    [.$\times$
+      [.$\times$ PELICULA PRODUCE ]
+      COMPAÑIA
+    ]
+  ]
+]
+
+% Pregunta b
+ \item
+
+% Pregunta c
+ \item
+
+% Pregunta d
+ \item
+
+% Pregunta e
+ \item
+
+% Pregunta f
+ \item
+
+% Pregunta g
+ \item
+
+% Pregunta h
+ \item
+
+% Pregunta i
+ \item
+
+% Pregunta j
+ \item
+
+% Pregunta k
  \item

### 2009-06-15 teoria_algoritmos/examen3/examen3.tex
pregunta 2 mas desarrollada
modificaciones desconocidas muy viejas a tarea3 de teoria de algoritmos

--- 2008-12-02 teoria_algoritmos/examen3/examen3.tex
@@
 
-La función de acotamiento $h$ aplicada a un nodo $x = (A_1, A_2)$ en que están asignados los $k$
-primeros elementos de $A$ cumple
+Sean para un nodo cualquiera
 
 $$
-h(x) = \left| \left(\left|\sum_{e \in A_1} e - \sum_{e \in A_2} e\right|\right) - \sum_{i=k}^n A[i] \right|
+L = \left|\sum_{e \in A_1} e - \sum_{e \in A_2} e\right|
 $$
 
-Es fácil ver que eso es una cota superior para cualquier asignación alcanzable desde $x$, ya que es el resultado
-de asignar todos los elementos que quedan por decidir al
+y
+
+$$
+S = \sum_{i=k}^n A[i]
+$$
+
+Si $L = S$, hay una asignación que hace mínimo el valor absoluto de la diferencia de los multiconjunos, porque
+no puede ser menor que $0$ y se puede lograr diferencia $0$ asignando los elementos faltantes a la menor parte
+de la suma. Si $L < S$, entonces la mínima diferencia que se podría lograr es $S-L$, porque ya los elementos que
+generan $S$ están asignados y lo único que podemos hacer es tratar de acercarnos a $0$. Si $L > S$, entonces
+una cota inferior para la diferencia que podemos conseguir por el nodo es $1$ si $L-S$ es impar y $0$ si es par,
+lo que consistiría en insertar
+elementos en el menor componente hasta igualarlo con el mayor y luego repartir el resto equitativamente.
+
+Entonces la función de acotamiento sería:
+
+\begin{equation*}
+h(x) = 
+\begin{cases}
+0 & \text{si $S=L$ o $L>S$ y $L-S$ es par,} \\
+S-L &\text{si $S>L$,} \\
+1 &\text{si $L>S$ y $L-S$ es impar.} \\
+\end{cases}
+\end{equation*}
+
+Ordeno $A$ de manera no decreciente para obtener mejor rendimiento.
 
  \item
+
+Denotaré una nodo como $(A_0,A_1)$ donde los $A_i$ son los multiconjuntos parciales que particionan a
+$A$ y $H$ es  el valor de la función de acotamiento para ese nodo.
+
+\renewcommand{\labelenumii}{\arabic{enumii}.}
+
+\begin{enumerate}
+
+\item Ordeno $A$ de manera no creciente.
+
+\item Comienzo la búsqueda con la solución inicial $(\{A_0\},\emptyset)$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1]\},\emptyset)$,\\
+$(\{A[0]\},\{A[1]\})$.
+
+\item Escojo como próxima solución actual a $(\{A[0],A[1]\},\emptyset)$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[2]\},\emptyset)$,\\
+$(\{A[0],A[1]\},\{A[2]\})$.
+
+\item Escojo como próxima solución actual a $(\{A[0],A[1]\},\{A[2]\})$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[3]\},\{A[2]\})$,\\
+$(\{A[0],A[1]\},\{A[2],A[3]\})$.
+
+\item Escojo como próxima solución actual a $(\{A[0],A[1]\},\{A[2],A[3]\})$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[3],A[4]\},\{A[2]\})$,\\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4]\})$.
+
+\item Escojo como próxima solución actual a\\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4]\})$.
+
+\item Expando mi solución actual. Las nuevas posibilidades son: \\
+$(\{A[0],A[1],A[3],A[4],A[5]\},\{A[2]\})$,\\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4],A[5]\})$.
+
+\item Escojo como próxima solución actual a \\
+$(\{A[0],A[1]\},\{A[2],A[3],A[4],A[5]\})$.
+
+\item Puedo podar todos las demás ramas con esta solución. Fin.
+
+\end{enumerate}
+
+Entonces la mejor solución tiene diferencia $1$.
 

### 2009-06-16 bdhetero/tarea2/tarea2.tex
parte b de tarea 2 de bdhetero lista salvo pequenos arreglos

--- 2009-05-28 bdhetero/tarea2/tarea2.tex
@@
-\documentclass[12pt]{article}
+\documentclass[10pt]{article}
 
@@
 
-% Pregunta 1
+% Pregunta a
+ \item
+
+\Tree
+[.$\Pi_{\text{PELICULA.NOMBRE}}$
+  [.{$\sigma$ \\
+   PRODUCE.PRESUPUESTO $ = 1000000$ (C1) \\
+   PRODUCE.AÑO $ = 2004$ {\bf OR} PRODUCE.AÑO $ = 2005$ (C2) \\
+   COMPAÑIA.NOMBRE $ = $ ``MGM'' (C3) \\
+   PELICULA.COD\_PELICULA $ = $ PRODUCE.COD\_PELICULA (C4) \\
+   PRODUCE.COD\_COMPAÑIA $ = $ COMPAÑIA.COD\_COMPAÑIA (C5) \\
+   }
+    [.$\times$
+      [.$\times$ PELICULA PRODUCE ]
+      COMPAÑIA
+    ]
+  ]
+]
+
+% Pregunta b
+ \item
+
+\renewcommand{\labelenumii}{\arabic{enumii}.}
+\newcommand{\andsql}{\text{ {\bf and} }}
+
+\begin{enumerate}
+ \item Cascada de selecciones:
+       \begin{eqnarray*}
+       & \sigma_{C1 \andsql C2 \andsql C4 \andsql C3 \andsql C5}(\ldots) \\
+       \equiv & \\
+       & \sigma_{C1 \andsql C2}(\sigma_{C4}(\sigma_{C3}(\sigma_{C5}(\ldots))))
+       \end{eqnarray*}
+
+ \item Conversión del producto cartesiano en join:
+       \begin{eqnarray*}
+       & \sigma_{C5}((\text{PELICULA} \times \text{PRODUCE}) \times \text{COMPAÑIA}) \\
+       \equiv & \\
+       & (\text{PELICULA} \times \text{PRODUCE}) \Join_{C5} \text{COMPAÑIA} \\
+       \end{eqnarray*}
+
+\item Distributividad de la selección con respecto al join:
+       \begin{eqnarray*}
+       & \sigma_{(C1 \andsql C2 \andsql C4) \andsql C3}(\ldots \Join_{C5} \ldots)  \\
+       \equiv & \\
+       & \sigma_{C1 \andsql C2 \andsql C4}(\ldots) \Join_{C5} \sigma_{C3}(\text{COMPAÑIA})  \\
+       \end{eqnarray*}
+
+       Se puede verificar que se cumplen las condiciones sobre los atributos
+       para aplicar esta equivalencia.
+
+ \item Conversión del producto cartesiano en join:
+       \begin{eqnarray*}
+       & \sigma_{C4}(\text{PELICULA} \times \text{PRODUCE}) \\
+       \equiv & \\
+       & \text{PELICULA} \Join_{C4} \text{PRODUCE} \\
+       \end{eqnarray*}
+
+\item Distributividad de la selección con respecto al join:
+       \begin{eqnarray*}
+       & \sigma_{\mathtt{true} \andsql (C1 \andsql C2)}(\text{PELICULA} \Join_{C4} \text{PRODUCE})  \\
+       \equiv & \\
+       & PELICULA \Join_{C4} \sigma_{C1 \andsql C2}(\text{PRODUCE})  \\
+       \end{eqnarray*}
+
+       Se puede verificar que se cumplen las condiciones sobre los atributos
+       para aplicar esta equivalencia.
+
+ \item Cascada de selecciones:
+       \begin{eqnarray*}
+       & \sigma_{C1 \andsql C2}(\text{PRODUCE}) \\
+       \equiv & \\
+       & \sigma_{C1}(\sigma_{C2}(PRODUCE))
+       \end{eqnarray*}
+
+\end{enumerate}
+
+% Pregunta c
+ \item
+
+% Pregunta d
+ \item
+
+% Pregunta e
+ \item
+
+% Pregunta f
+ \item
+
+% Pregunta g
+ \item
+
+% Pregunta h
+ \item
+
+% Pregunta i
+ \item
+
+% Pregunta j
+ \item
+
+% Pregunta k
  \item

### 2009-06-18 matematicas/tarea5.tex
prepare .tex de tarea 5 de matematicas

--- 2009-05-12 matematicas/tarea2.tex
@@
 \usepackage{qtree}
+\usepackage{enumerate}
 
@@
 \newcommand{\techo}[1]{\left \lceil #1 \right \rceil}
+\newcommand{\stirlingU}[2]{\genfrac{[}{]}{0pt}{}{#1}{#2}}
+\newcommand{\stirlingD}[2]{\genfrac{\{}{\}}{0pt}{}{#1}{#2}}
 
@@
 
-\title{Tarea 2 de Matemáticas para Computación}
+\title{Tarea 5 de Matemáticas para Computación}
 \author{Daniel Izquierdo \\ \#08-86809}
-\date{12 de mayo de 2009}
+\date{25 de junio de 2009}
 
@@
 
-Supongamos que $F$ no termina. Entonces existe una cadena
-infinita de llamadas
-
-$$
-P_0: F(x_0, y_0) \rightarrow F(x_1, y_1) \rightarrow F(x_2, y_2) \rightarrow \ldots
-$$
-
-donde defino $\rightarrow$ como ``llama a''.
-Aquí, no puede ser que haya $x_i$ tal que $x_i = 0$ o $y_i$ tal que $y_i = 0$
-porque
-en estos casos $F$ no hace llamadas recursivas.
-%Observemos que hay una
-%correspondencia uno a uno $c$ entre llamadas a $F$ y el producto cartesiano de
-%los naturales, de manera que $c(F(x,y)) = (x,y)$.
-Observemos que cuando
-$F(x,y) \rightarrow F(u,v)$, tenemos dos posibilidades:
-
-\begin{itemize}
-\item $(u,v) = (x-1,r)$ para algún $r$. Tenemos por definición del orden
-lexicográfico que $(x,y) \succeq (x-1,r)$ porque $x-1 < x$.
-\item $(u,v) = (x,y-1)$. Tenemos también por definición del orden lexicográfico
-que $(x,y) \succeq (x,y-1)$ porque $x = x \wedge y-1 < y$.
-\end{itemize}
-
-De esta manera, $F(x,y) \rightarrow F(u,v)$ implica que $(x,y) \succeq (u,v)$.
-Luego, la existencia de $P_0$ implica que existe una cadena infinita
-
-$$
-(x_0, y_0) \succeq (x_1, y_1) \succeq (x_2, y_2) \succeq \ldots
-$$
-
-lo cual es falso porque el orden es Noetheriano. Esta contradicción nos lleva a
-concluir que $F$ termina.
-
-% Pregunta 2
-\section{}
-
-Primero quiero ver qué debe cumplir $d$ para que $f(n) \leq dn$ siga de suponer
-que $n$ es suficientemente grande y $f(m) \leq dm$ para todo $m$ menor que $n$
-y también suficientemente grande. Si hacemos tal suposición y tenemos
-$\piso{pn}$ y $\piso{qn}$ suficientemente grandes, entonces necesitamos
-$f(n) \leq dn$ y tenemos
-
-\begin{align*}
- & f(n) \\
- = & \\
- & f(\piso{pn}) + f(\piso{qn}) + bn \\
- \leq & \\
- & d\piso{pn} + d\piso{qn} + bn \\
- \leq & \\
- & dpn + dqn + bn \\
- = & \\
- & dn(p+q) + bn \\
- = & \\
- & n(d(p+q) + b) \\
-\end{align*}
-
-Para poder concluir que $f(n) \leq dn$ necesitamos que
-
-\begin{align*}
- & n(d(p+q) + b) \leq dn \\
- \equiv & \\
- & d(p+q) + b \leq d \\
- \equiv & \\
- & b \leq d - d(p+q) \\
- \equiv & \\
- & b \leq d(1 - (p+q)) \\
- \equiv & \\
- & \frac{b}{1-(p+q)} \leq d \\
-\end{align*}
-
-El término de la izquierda en la última ecuación está definido y es positivo
-porque $p+q$ es positivo y menor que $1$ y $b$ es positivo. Entonces tenemos
-que, bajo esa condición, $f(n) \leq dn$ es cierto si
-$f(\piso{pn}) \leq d\piso{pn}$ y $f(\piso{qn}) \leq d\piso{qn}$ también lo son.
-Para culminar la prueba necesito números apropiados como base para aplicar
-inducción.
-
-% Pregunta 3
-\section{}
-
-Estas funciones tienen la propiedad de que $\piso{f(x)} = \piso{f(\techo{x}}$.
-La prueba es similar a la dada en el libro para demostrar que
-$\techo{f(x)} = \techo{f(\techo{x})}$.
-Para todo $x$ entero se cumple trivialmente porque $x = \techo{x}$. Si $x$ no
-es entero, tenemos $x < \techo{x}$. Como $f$ es monótona decreciente, esto
-implica que $f(x) > f(\techo{x})$. Entonces
-$\piso{f(x)} \geq \piso{f(\techo{x})}$.
-Cuando es $\piso{f(x)} > \piso{f(\techo{x})}$, debe haber $y$ tal que
-$x \leq y < \techo{x}$ y $f(y) = \piso{f(x)}$ porque $f$ es continua.
-$y$ debe
-ser entero por la propiedad
-de $f$ de tomar valores enteros sólo con argumentos enteros.
-Pero no puede haber enteros $i$ tales que $x \leq i < \techo{x}$. 
-Con esto llegamos
-a una contradicción.
-Entonces debe ser $\piso{f(x)} = \piso{f(\techo{x})}$.
-
-% Pregunta 4
-\section{}
-
-Tenemos
-
-\begin{align*}
- & \techo{\frac{n}{m}} = \piso{\frac{n+m-1}{m}} \\
- \equiv & \\
- & \techo{\frac{n}{m}} - \piso{\frac{n}{m}} = \piso{\frac{n+m-1}{m}} - \piso{\frac{n}{m}} \\
- \equiv & \langle \piso{\frac{n}{m}} \text{ es entero} \rangle \\
- & \techo{\frac{n}{m} - \piso{\frac{n}{m}}} = \piso{\frac{n+m-1}{m} - \piso{\frac{n}{m}}} \\
- \equiv & \\
- & \techo{\frac{n - m\piso{\frac{n}{m}}}{m}} = \piso{\frac{n+m-1 - m\piso{\frac{n}{m}}}{m}} \\
- \equiv & \langle m \text{ positivo} \rangle \\
- & \techo{\frac{n \mathtt{mod} m}{m}} = \piso{\frac{n \mathtt{mod} m +m-1}{m}} \\
-\end{align*}
-
-Esto último es equivalente a $n \mathtt{mod} m > 0$, porque debe ser
-$0 \leq n \mathtt{mod} m < m$.
-Entonces es equivalente a $m > 0$. Por lo tanto, $m > 0$ implica la igualdad que se quería
-demostrar.
-
-% Pregunta 5
-\section{}
-
-Sea $N = \frac{3^77-1}{2}$. $N$ es impar porque 
-$N$ es compuesto porque es divisible entre $\frac{3^7-1}{2}$:
-
 \end{document}

### 2009-06-25 matematicas/tarea5.tex
agregue secciones para cada pregunta de tarea 5 de matematicas
pregunta 1 de tarea 5 de matematicas lista

@@
 
+$\stirlingD{n}{m}$ es el número de maneras de particionar un conjunto de $n$
+elementos en $m$ subconjuntos no vacíos. Ahora, sea $A$ un conjunto
+tal que $|A| = n$ y supongamos que tenemos un conjunto $B$ de contenedores
+llamados $b_i$ con $1 \leq i \leq m$. Observemos entonces que los $F_i$
+propuestos en el enunciado de la pregunta son las maneras de particionar $A$
+colocando cada elemento en alguno de los $b_k$ con $k \neq i$, lo que garantiza
+que $b_i$ queda vacío. Observemos también que si quisieramos particionar los $n$
+elementos en los $m$ subconjuntos olvidando la restricción de prohibir
+subconjuntos vacíos, hay $m^n$ opciones porque podemos elegir entre $m$
+subconjuntos para colocar cada elemento.
+
+Con todo esto tenemos que
+
+\[ \left| \bigcup_{i=1}^m F_i \right| \]
+
+es el número de maneras de particionar $A$ en los $b_i$ de manera de que exista
+al menos un $b_i$ vacío. Por principio de inclusión-exclusión y utilizando el
+hecho de que, para cualquier $q$, cualquier interseccón de $q$ distintos $F_i$
+tiene la misma
+cardinalidad, nos da
+
+\begin{align*}
+\left| \bigcup_{i=1}^m F_i \right| & = \sum_{k=1}^m \binom{m}{k} (-1)^{k-1} |F_1 \cap F_2 \cap \ldots \cap F_k|
+\end{align*}
+
+Ahora, sabemos que la intersección de $q$ $F_i$ distintos es el conjunto de las
+maneras de particionar $A$ en $m-q$ conjuntos, sin importar que queden conjuntos
+vacíos (porque estamos obligando a $q$ conjuntos a estar vacíos, y no nos
+importan los demás con tal de que cada elemento de $A$ quede asignado). Entonces
+la cardinalidad de una intersección tal es $(m-q)^n$. Por lo tanto queda
+
+\begin{align*}
+\left| \bigcup_{i=1}^m F_i \right| & = \sum_{k=1}^m \binom{m}{k} (-1)^{k-1} (m-k)^n
+\end{align*}
+
+Si lo que queremos es contar particiones en la que ningún $b_i$ esté
+vacío, debemos tomar todas las posibles maneras de particionar y quitarles el
+número que acabamos de calcular. Es decir, lo que buscamos es
+
+\begin{align*}
+m^n - \left| \bigcup_{i=1}^m F_i \right|
+  & = m^n - \sum_{k=1}^m \binom{m}{k} (-1)^{k-1} (m-k)^n \\
+  & = m^n + \sum_{k=1}^m \binom{m}{k} (-1)^{k} (m-k)^n \\
+  & = \sum_{k=0}^m \binom{m}{k} (-1)^{k} (m-k)^n \\
+  & = \sum_k \binom{m}{k} (-1)^{k} (m-k)^n \\
+  & = \sum_{m-k} \binom{m}{m-k} (-1)^{m-k} k^n \\
+  & = \sum_{m-k} \binom{m}{k} (-1)^{m-k} k^n \\
+  & = \sum_k \binom{m}{k} (-1)^{m-k} k^n \\
+\end{align*}
+
+Para terminar nos damos cuenta de que $\stirlingD{n}{m}$ es igual a este
+número salvo la numeración de los $b_i$. Hay $m!$ maneras de numerarlos
+(número de permutaciones distintas de $m$ elementos), factor que
+debemos eliminar del cálculo. Entonces nos queda
+
+\[
+\stirlingD{n}{m} = \frac{1}{m!} \sum_k \binom{m}{k} (-1)^{m-k} k^n \\
+\]
+
+% Pregunta 2
+\section{}
+
+% Pregunta 3
+\section{}
+
+% Pregunta 4
+\section{}
+
 \end{document}

### 2009-06-25 matematicas/tarea5.tex
parte a de pregunta 3 de tarea 5 de matematicas

@@
 
+\renewcommand{\labelenumi}{(\alph{enumi})}
+\begin{enumerate}
+
+\item Para un par de enteros $m,n$ con $1 \leq m \leq N$ y $1 \leq n \leq N$
+pueden haber tres posibilidades: $m=n$, $m<n$ o $m>n$. Si $m=n$ entonces
+$m \perp n \equiv m=n=1$, por lo que se cuenta un sólo caso para el resultado de
+$R(N)$.
+
+Analicemos qué ocurre si $m<n$. Para un $m \neq 1$, $\varphi(m)$ cuenta
+los coprimos $c < m$. O sea que $\Phi(N)$ cuenta todos los pares $1 \leq n < m
+\leq N$ con $n \perp m$ pero también añade el caso $n=m=1$, el cual debemos
+eliminar. Entonces cuando $m<n$ debemos agregar $\Phi(N)-1$ a $R(N)$.
+Al tener $m>n$ ocurre de manera similar, es decir, tenemos $\Phi(N)-1$ casos
+más. Por lo tanto tenemos
+
+\[
+R(N) = \Phi(N)-1 + \Phi(N)-1 + 1 = 2\Phi(N) - 1
+\]
+
+\item
+
+\end{enumerate}
+
 % Pregunta 4

### 2009-06-25 matematicas/tarea5.tex
partes b y c de pregunta 4 de tarea 5 de matematicas

@@
 
-\item
+\item Usando (4.62) del libro y el resultado anterior tenemos
+
+\begin{align*}
+R(N)
+  & = 2\Phi(N)-1 \\
+  & = 2 \left( \frac{1}{2} \sum_{d \geq 1} \mu(d) \piso{N/d} \piso{1+N/d} \right) -1 \\
+  & = \sum_{d \geq 1} \mu(d) \piso{N/d} \piso{1+N/d} - 1 \\
+  & = \ldots \\
+  & = \sum_{d \geq N} \piso{N/d}^2 \mu(d)  \\
+\end{align*}
 
@@
 
+\begin{enumerate}
+
+\item Función generatriz:
+
+\item Polinomio característico:
+
+La recurrencia indica la ecuación
+
+\begin{align*}
+& t_n - 3t_{n-1} + 2t_{n-2} = 3*2^{n-2} \\
+\equiv & \\
+& t_n - 3t_{n-1} + 2t_{n-2} = \frac{3}{4} 2^n \\
+\end{align*}
+
+Procedemos a multiplicar todo por $2$
+
+\begin{align*}
+2t_n - 6t_{n-1} + 4t_{n-2} = \frac{3}{4} 2^{n+1} \\
+\end{align*}
+
+y a sustituir $n$ por $n-1$
+
+\begin{align*}
+2t_{n-1} - 6t_{n-2} + 4t_{n-3} = \frac{3}{4} 2^{n} \\
+\end{align*}
+
+Restando de la ecuación original tenemos
+
+\begin{align*}
+& t_n - 3t_{n-1} +2t_{n-2} -2t_{n-1}+6t_{n-2}-4t_{n-3} = 0 \\
+\equiv & \\
+& t_n - 5t_{n-1} + 8t_{n-2}-4t_{n-3} = 0 \\
+\end{align*}
+
+De aquí obtenemos el polinomio característico:
+
+\[
+x^3-5x^2+8x-4 = (x-2)^2(x-1)
+\]
+
+lo que nos dice que todas las soluciones de la recurrencia general son de la
+forma
+
+\[
+t_n = c_1 + c_2 2^n + c_3 n 2^n
+\]
+
+La recurrencia del enunciado impone la condición $t_2 = 3t_1-2t_0+3$. Unido a
+los casos base tenemos el sistema de ecuaciones
+
+\begin{align*}
+c_1 + c_2 & = t_0 & n = 0 \\
+c_1+2c_2 + 2c_3 & = t_1 & n = 1 \\
+c_1+4c_2+8c_3 & = 3t_1-2t_0+3 & n = 2 \\
+\end{align*}
+
+Sustituyendo por los valores conocidos de los casos base queda
+
+\begin{align*}
+c_1 + c_2 & = 1 & \\
+c_1+2c_2 + 2c_3 & = 2 & \\
+c_1+4c_2+8c_3 & = 7 & \\
+\end{align*}
+
+Resolviendo obtenemos
+
+\begin{align*}
+c_1 & = 3 \\
+c_2 & = -2 \\
+c_3 & = \frac{3}{2} \\
+\end{align*}
+
+En conclusión, resulta que
+
+\[
+t_n = 3-2*2^n+\frac{3}{2}n2^n = 3-2^{n+1}+3n2^{n-1}
+\]
+
+\item $t_n \in \Theta(n2^n)$
+
+\end{enumerate}
+
 \end{document}

### 2009-06-25 matematicas/tarea5.tex
algunos avances de tarea 5 de matematicas
pregunta 2 de tarea 5 de matematicas lista (chucuta)

@@
 
+Supongamos que cada alfabeto tiene un orden y que trabajamos con permutaciones
+que preservan el orden de las letras de cada alfabeto. Es decir, permitimos
+permutaciones $l_1, l_2, \ldots, l_{5n}$ tales que si $l_i, l_j$ pertenecen
+al mismo alfabeto y $l_i < l_j$ según ese alfabeto, entonces $i < j$.
+Si definimos $F_i$ como el conjunto de tales permutaciones que además cumplan
+que hay alguna letra del alfabeto $i$ que no está acompañada de otra del mismo
+alfabeto, entonces
+
+\[ \left| \bigcup_{i=1}^n F_i \right| \]
+
+es el número de permutaciones que tienen alguna letra que no está acompañada. De
+manera similar al ejercicio anterior vemos que
+
+\begin{align*}
+\left| \bigcup_{i=1}^n F_i \right| & = \sum_{k=1}^n \binom{n}{k} (-1)^{k-1} |F_1 \cap F_2 \cap \ldots \cap F_k|
+\end{align*}
+
+Tenemos que la cardinalidad de la intersección de $q$ $F_i$ distintos es
+$(2n-q)!$. El número de permutaciones posibles total es $(2n)!$. Si eliminamos
+del número total aquellas que no nos interesan, queda
+
+\begin{align*}
+& (2n)! - \left| \bigcup_{i=1}^n F_i \right| \\
+= & \\
+& (2n)! - \sum_{k=1}^n \binom{n}{k} (-1)^{k-1} |F_1 \cap F_2 \cap \ldots \cap F_k| \\
+= & \\
+& (2n)! - \sum_{k=1}^n \binom{n}{k} (-1)^{k-1} (2n-k)! \\
+= & \\
+& (2n)! + \sum_{k=1}^n \binom{n}{k} (-1)^{k} (2n-k)! \\
+= & \\
+& \sum_{k=0}^n \binom{n}{k} (-1)^{k} (2n-k)! \\
+\end{align*}
+
+Este resultado es el número de permutaciones válidas dado un orden específico
+para cada alfabeto. Pero estamos buscando el número sin importar el orden. Hay
+$n$ alfabetos, cada uno de los cuales aporta $5!$ posibles órdenes
+(porque son las permutaciones de las cinco letras que tienen). Entonces debemos
+multiplicar el resultado por $(5!)^n = 120^n$ y queda
+
+\begin{align*}
+(120)^n (\sum_{k=0}^n \binom{n}{k} (-1)^{k} (2n-k)!) \\
+\end{align*}
+
 % Pregunta 3
@@
 \item Función generatriz:
+
+Como primer paso escribo la ecuación de la recurrencia de manera que sea válida
+para todo $n$ bajo la suposición de que los términos con índice negativo siempre
+se anulan. Entonces pongo
+
+\[
+t_n = 3t_{n-1} - 2t_{n-2} + 3*2^{n-2} [n \geq 0] - \frac{5}{2} [n=1] + \frac{1}{4}[n=0]
+\]
+
+El segundo paso es multiplicar por $z^n$, sumar sobre $n$ e intentar despejar $G(z)$.
+
+\begin{align*}
+G(z)
+  & = \sum_n t_n z^n \\
+  & = 3 \sum_n t_{n-1} z^n - 2 \sum_n t_{n-2} z^n + \frac{3}{4} \sum_n 2^n z^n [n \geq 0] - \\
+  &   \frac{5}{2} \sum_n z^n [n=1] + \frac{1}{4} \sum_n z^n [n=0] \\
+  & = 3zG(z) - 2z^2G(z) + \frac{3}{4}\frac{1}{1-2z} - \frac{5}{2}z + \frac{1}{4} \\
+\end{align*}
 

### 2009-07-04 matematicas/tarea5.tex
algunos ultimos arreglos de la tarea 5 de matematicas

@@
 
-\item Usando (4.62) del libro y el resultado anterior tenemos
+\item Usando (4.62), (4.61) del libro y el resultado anterior tenemos
 
@@
   & = \sum_{d \geq 1} \mu(d) \piso{N/d} \piso{1+N/d} - 1 \\
-  & = \ldots \\
   & = \sum_{d \geq N} \piso{N/d}^2 \mu(d)  \\
@@
 \end{align*}
+
+Como tercer paso despejamos $G(z)$, lo que resulta en la siguiente ecuación:
+
+\[
+G(z) = \frac{5z^2-3z+1}{-4z^3+8z^2-5z+1}
+\]
+
+El cuarto paso es expandir esto en series de potencia y conseguir
+ahí el coeficiente de $z^n$.
 

### 2009-07-11 bdhetero/tarea3/tarea3.tex
esqueleto correcto de tarea3 de bdhetero

--- 2009-06-25 matematicas/tarea5.tex
@@
-\documentclass{article}
+\documentclass[10pt]{article}
 
-\usepackage{algorithm}
-\usepackage{algorithmic}
 \usepackage{amsmath}
@@
 \usepackage[utf8]{inputenc}
-\usepackage{listings}
 \usepackage[spanish]{babel}
 \usepackage{qtree}
-\usepackage{enumerate}
-
-% Para codigo bonito
-\newcommand{\singlespace}{\renewcommand{\baselinestretch}{1.0}}
-\newcommand{\piso}[1]{\left \lfloor #1 \right \rfloor}
-\newcommand{\techo}[1]{\left \lceil #1 \right \rceil}
-\newcommand{\stirlingU}[2]{\genfrac{[}{]}{0pt}{}{#1}{#2}}
-\newcommand{\stirlingD}[2]{\genfrac{\{}{\}}{0pt}{}{#1}{#2}}
-
-\newcommand{\algoritmo}[2]
-{
-{
-  \singlespace
-  \begin{algorithm}[h]
-    \caption{#1}
-    \begin{algorithmic}[1]
-      #2
-    \end{algorithmic}
-  \end{algorithm}
-}
-}
-
-\newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-\lstset{mathescape=true,frame=single}
 
 \newcommand{\asig}{\ensuremath{\leftarrow}}
-\renewcommand{\lstlistingname}{Listado} %listings en español
 
@@
 
-\title{Tarea 5 de Matemáticas para Computación}
+\title{Bases de Datos Heterogéneas \\ Tarea 3}
 \author{Daniel Izquierdo \\ \#08-86809}
-\date{25 de junio de 2009}
 
@@
 
-$\stirlingD{n}{m}$ es el número de maneras de particionar un conjunto de $n$
-elementos en $m$ subconjuntos no vacíos. Ahora, sea $A$ un conjunto
-tal que $|A| = n$ y supongamos que tenemos un conjunto $B$ de contenedores
-llamados $b_i$ con $1 \leq i \leq m$. Observemos entonces que los $F_i$
-propuestos en el enunciado de la pregunta son las maneras de particionar $A$
-colocando cada elemento en alguno de los $b_k$ con $k \neq i$, lo que garantiza
-que $b_i$ queda vacío. Observemos también que si quisieramos particionar los $n$
-elementos en los $m$ subconjuntos olvidando la restricción de prohibir
-subconjuntos vacíos, hay $m^n$ opciones porque podemos elegir entre $m$
-subconjuntos para colocar cada elemento.
-
-Con todo esto tenemos que
-
-\[ \left| \bigcup_{i=1}^m F_i \right| \]
-
-es el número de maneras de particionar $A$ en los $b_i$ de manera de que exista
-al menos un $b_i$ vacío. Por principio de inclusión-exclusión y utilizando el
-hecho de que, para cualquier $q$, cualquier interseccón de $q$ distintos $F_i$
-tiene la misma
-cardinalidad, nos da
-
-\begin{align*}
-\left| \bigcup_{i=1}^m F_i \right| & = \sum_{k=1}^m \binom{m}{k} (-1)^{k-1} |F_1 \cap F_2 \cap \ldots \cap F_k|
-\end{align*}
-
-Ahora, sabemos que la intersección de $q$ $F_i$ distintos es el conjunto de las
-maneras de particionar $A$ en $m-q$ conjuntos, sin importar que queden conjuntos
-vacíos (porque estamos obligando a $q$ conjuntos a estar vacíos, y no nos
-importan los demás con tal de que cada elemento de $A$ quede asignado). Entonces
-la cardinalidad de una intersección tal es $(m-q)^n$. Por lo tanto queda
-
-\begin{align*}
-\left| \bigcup_{i=1}^m F_i \right| & = \sum_{k=1}^m \binom{m}{k} (-1)^{k-1} (m-k)^n
-\end{align*}
-
-Si lo que queremos es contar particiones en la que ningún $b_i$ esté
-vacío, debemos tomar todas las posibles maneras de particionar y quitarles el
-número que acabamos de calcular. Es decir, lo que buscamos es
-
-\begin{align*}
-m^n - \left| \bigcup_{i=1}^m F_i \right|
-  & = m^n - \sum_{k=1}^m \binom{m}{k} (-1)^{k-1} (m-k)^n \\
-  & = m^n + \sum_{k=1}^m \binom{m}{k} (-1)^{k} (m-k)^n \\
-  & = \sum_{k=0}^m \binom{m}{k} (-1)^{k} (m-k)^n \\
-  & = \sum_k \binom{m}{k} (-1)^{k} (m-k)^n \\
-  & = \sum_{m-k} \binom{m}{m-k} (-1)^{m-k} k^n \\
-  & = \sum_{m-k} \binom{m}{k} (-1)^{m-k} k^n \\
-  & = \sum_k \binom{m}{k} (-1)^{m-k} k^n \\
-\end{align*}
-
-Para terminar nos damos cuenta de que $\stirlingD{n}{m}$ es igual a este
-número salvo la numeración de los $b_i$. Hay $m!$ maneras de numerarlos
-(número de permutaciones distintas de $m$ elementos), factor que
-debemos eliminar del cálculo. Entonces nos queda
-
-\[
-\stirlingD{n}{m} = \frac{1}{m!} \sum_k \binom{m}{k} (-1)^{m-k} k^n \\
-\]
-
 % Pregunta 2
@@
 
-% Pregunta 3
-\section{}
-
-\renewcommand{\labelenumi}{(\alph{enumi})}
-\begin{enumerate}
-
-\item Para un par de enteros $m,n$ con $1 \leq m \leq N$ y $1 \leq n \leq N$
-pueden haber tres posibilidades: $m=n$, $m<n$ o $m>n$. Si $m=n$ entonces
-$m \perp n \equiv m=n=1$, por lo que se cuenta un sólo caso para el resultado de
-$R(N)$.
-
-Analicemos qué ocurre si $m<n$. Para un $m \neq 1$, $\varphi(m)$ cuenta
-los coprimos $c < m$. O sea que $\Phi(N)$ cuenta todos los pares $1 \leq n < m
-\leq N$ con $n \perp m$ pero también añade el caso $n=m=1$, el cual debemos
-eliminar. Entonces cuando $m<n$ debemos agregar $\Phi(N)-1$ a $R(N)$.
-Al tener $m>n$ ocurre de manera similar, es decir, tenemos $\Phi(N)-1$ casos
-más. Por lo tanto tenemos
-
-\[
-R(N) = \Phi(N)-1 + \Phi(N)-1 + 1 = 2\Phi(N) - 1
-\]
-
-\item Usando (4.62) del libro y el resultado anterior tenemos
-
-\begin{align*}
-R(N)
-  & = 2\Phi(N)-1 \\
-  & = 2 \left( \frac{1}{2} \sum_{d \geq 1} \mu(d) \piso{N/d} \piso{1+N/d} \right) -1 \\
-  & = \sum_{d \geq 1} \mu(d) \piso{N/d} \piso{1+N/d} - 1 \\
-  & = \ldots \\
-  & = \sum_{d \geq N} \piso{N/d}^2 \mu(d)  \\
-\end{align*}
-
-\end{enumerate}
-
-% Pregunta 4
-\section{}
-
-\begin{enumerate}
-
-\item Función generatriz:
-
-\item Polinomio característico:
-
-La recurrencia indica la ecuación
-
-\begin{align*}
-& t_n - 3t_{n-1} + 2t_{n-2} = 3*2^{n-2} \\
-\equiv & \\
-& t_n - 3t_{n-1} + 2t_{n-2} = \frac{3}{4} 2^n \\
-\end{align*}
-
-Procedemos a multiplicar todo por $2$
-
-\begin{align*}
-2t_n - 6t_{n-1} + 4t_{n-2} = \frac{3}{4} 2^{n+1} \\
-\end{align*}
-
-y a sustituir $n$ por $n-1$
-
-\begin{align*}
-2t_{n-1} - 6t_{n-2} + 4t_{n-3} = \frac{3}{4} 2^{n} \\
-\end{align*}
-
-Restando de la ecuación original tenemos
-
-\begin{align*}
-& t_n - 3t_{n-1} +2t_{n-2} -2t_{n-1}+6t_{n-2}-4t_{n-3} = 0 \\
-\equiv & \\
-& t_n - 5t_{n-1} + 8t_{n-2}-4t_{n-3} = 0 \\
-\end{align*}
-
-De aquí obtenemos el polinomio característico:
-
-\[
-x^3-5x^2+8x-4 = (x-2)^2(x-1)
-\]
-
-lo que nos dice que todas las soluciones de la recurrencia general son de la
-forma
-
-\[
-t_n = c_1 + c_2 2^n + c_3 n 2^n
-\]
-
-La recurrencia del enunciado impone la condición $t_2 = 3t_1-2t_0+3$. Unido a
-los casos base tenemos el sistema de ecuaciones
-
-\begin{align*}
-c_1 + c_2 & = t_0 & n = 0 \\
-c_1+2c_2 + 2c_3 & = t_1 & n = 1 \\
-c_1+4c_2+8c_3 & = 3t_1-2t_0+3 & n = 2 \\
-\end{align*}
-
-Sustituyendo por los valores conocidos de los casos base queda
-
-\begin{align*}
-c_1 + c_2 & = 1 & \\
-c_1+2c_2 + 2c_3 & = 2 & \\
-c_1+4c_2+8c_3 & = 7 & \\
-\end{align*}
-
-Resolviendo obtenemos
-
-\begin{align*}
-c_1 & = 3 \\
-c_2 & = -2 \\
-c_3 & = \frac{3}{2} \\
-\end{align*}
-
-En conclusión, resulta que
-
-\[
-t_n = 3-2*2^n+\frac{3}{2}n2^n = 3-2^{n+1}+3n2^{n-1}
-\]
-
-\item $t_n \in \Theta(n2^n)$
-
-\end{enumerate}
-
 \end{document}

### 2009-07-20 bdhetero/tarea3/tarea3.tex
pregunta 1 de tarea3 de bdhetero
arbol de pregunta2 de tarea3 de bdhetero
parte 2.1 de tarea3 de bdhetero
tarea 3 lista en teoria

@@
 
+\begin{itemize}
+ \item $v_1(x,y),v_6(x),v_7(y)$
+ \item $v_1(y,x),v_4(x,y),v_6(x),v_7(y)$
+ \item $v_3(y,x),v_3(x,y),v_6(x),v_7(y)$
+\end{itemize}
+
 % Pregunta 2
@@
 
+% Pregunta 2.1
+\subsection{}
+
+\Tree
+[.$\sigma_{\text{Premio.Premio}=\text{'Oscar'}, \text{Premio.Año}=2009}$
+ [.$\Join_{Pelicula.CodPelicula=Premio.CodPelicula}$
+   [.$\sigma_{\text{Evaluacion.Evaluacion}=\text{'Buena'}}$
+    [.$\Join_{\text{Pelicula.CodPelicula}=\text{Evaluacion.CodPelicula}}$
+     [.$\sigma_{\text{Pelicula.Director}=\text{'Almodovar'}}$
+      Pelicula
+     ]
+     Evaluacion
+    ]
+   ]
+   Premio
+ ]
+]
+
+\begin{itemize}
+ \item Costo selección sobre Pelicula: $10s+4*10s = 50s$.
+ Se obtienen en promedio $4\frac{1024}{100}$ películas.
+ \item Costo join Pelicula/Evaluacion: $50s + 4\frac{1024}{100}*(25s+1*20s) = 1895s$.
+ \item Costo selección sobre Evaluacion (en el mediador): $1895s + \lceil \frac{4(4\frac{1024}{100})}{1024} \rceil s = 1896s$.
+ \item Costo join Pelicula/Evaluacion/Premio: $1896s + 4\frac{1024}{100}*(27s+2*15s) = 4230s$.
+ \item Costo selección sobre Premio (en el mediador): $4230s + \lceil \frac{4(4\frac{1024}{100})}{1024} \rceil = 4231s$.
+\end{itemize}
+
+% Pregunta 2.2
+\subsection{}
+
+\begin{enumerate}[a.]
+ \item
+
+ \begin{enumerate}[1:]
+  % Iteracion 1
+  \item Posibles subplanes:
+    \begin{itemize}
+     \item Director $\rightarrow$  NombrePelicula, CodPelicula, Actores, Directores. Costo: $10+4*10s = 50s$.
+     \item NombrePelicula $\rightarrow$ Actores, Directores, CodPelicula. \\
+           No se tiene NombrePelicula.
+     \item CodPelicula $\rightarrow$  Evaluacion. No se tiene CodPelicula.
+     \item CodPelicula $\rightarrow$  Premio, Año. No se tiene CodPelicula.
+     \item Año $\rightarrow$ Premio, CodPelicula. Costo: $30s+3*45s = 165s$.
+    \end{itemize}
+
+     Se elige el subplan: Director $\rightarrow$  NombrePelicula, CodPelicula, Actores, Directores \\
+     Costo: $50s$
+
+  % Iteracion 2
+ \item Posibles subplanes:
+    \begin{itemize}
+     \item NombrePelicula $\rightarrow$ Actores, Directores, CodPelicula. \\
+     Ya tenemos Actores, Directores, CodPelicula.
+     \item CodPelicula $\rightarrow$  Evaluacion. Costo: $50s+(4\frac{1024}{100})(25+1*20)s+4 = 1893,2s+4$
+     \item CodPelicula $\rightarrow$  Premio, Año. Costo: $50s+(4\frac{1024}{100})(27+2*15)+4 = 2384,7s+4$
+     \item Año $\rightarrow$ Premio, CodPelicula. Costo (join): $50s+(30+3*45)s+4+\lceil \frac{4}{20} \rceil*(1600000*50/1024) = 215s+78129$.
+    \end{itemize}
+
+    Se elige el subplan: CodPelicula $\rightarrow$  Evaluacion \\
+    Costo: $1893,2s+4$
+
+  % Iteracion 3
+ \item Posibles subplanes:
+    \begin{itemize}
+     \item CodPelicula $\rightarrow$  Premio, Año. Costo: $1893,2s+4+(4\frac{1024}{100})(27+2*15)+4 = 4227,9s+8$
+     \item Año $\rightarrow$ Premio, CodPelicula. Costo (join): $1893,2s+4+(30+3*45)s+4+\lceil \frac{4}{20} \rceil*(1600000*50/1024) =
+           2058,2s + 78133$.
+    \end{itemize}
+
+    Se elige el subplan: CodPelicula $\rightarrow$  Premio, Año \\
+    Costo: $4227,9s+8$
+
+ \end{enumerate}
+
+ \item El costo del plan es $4227,9$ segundos y ocho veces el tiempo de lectura o escritura de un bloque.
+
+ \item El plan es peor que el propuesto anteriormente porque dura al menos más que el anterior en envío de páginas.
+
+\end{enumerate}
+
 \end{document}

### 2009-07-20 bdhetero/tarea2/tarea2.tex
parte b de tarea 2 de bdhetero lista
pregunta c de tarea2 de bdhetero
correccion de pregunta c de tarea2 de bdhetero

@@
 \renewcommand{\labelenumii}{\arabic{enumii}.}
-\newcommand{\andsql}{\text{ {\bf and} }}
+\newcommand{\andsql}{\text{ {\bf AND} }}
+
+Usaré conmutatividad y asociatividad de $\andsql$ y neutralidad de
+$\mathtt{true}$ sobre $\andsql$ libremente.
 
@@
        \begin{eqnarray*}
-       & \sigma_{C1 \andsql C2 \andsql C4 \andsql C3 \andsql C5}(\ldots) \\
+       & \sigma_{C1 \andsql C2 \andsql C3 \andsql C4 \andsql C5}(\ldots) \\
        \equiv & \\
-       & \sigma_{C1 \andsql C2}(\sigma_{C4}(\sigma_{C3}(\sigma_{C5}(\ldots))))
+       & \sigma_{C1 \andsql C2 \andsql C3 \andsql C4}(\sigma_{C5}(\ldots))
        \end{eqnarray*}
@@
        para aplicar esta equivalencia.
+
+ \item Cascada de selecciones:
+       \begin{eqnarray*}
+       & \sigma_{(C1 \andsql C2) \andsql C4}(\ldots) \\
+       \equiv & \\
+       & \sigma_{C1 \andsql C2}(\sigma_{C4}(\ldots))
+       \end{eqnarray*}
 
@@
        \equiv & \\
-       & \sigma_{C1}(\sigma_{C2}(PRODUCE))
+       & \sigma_{C1}(\sigma_{C2}(\text{PRODUCE}))
        \end{eqnarray*}
@@
 
+\begin{itemize}
+
+\item $\sigma_{C2}(\text{PRODUCE})$: 391 bloques, produce 2000 tuplas
+
+\item $\sigma_{C1}(\sigma_{C2}(\text{PRODUCE}))$: 79 bloques, produce hasta 2000 tuplas
+
+\item $PELICULA \Join_{C4} \sigma_{C1 \andsql C2}(\text{PRODUCE})$: \\
+      (hash join) 2931+237=3168 bloques, produce hasta 2000 tuplas
+
+\item $\sigma_{C4}(\text{COMPAÑIA})$: 25 bloques, produce hasta hasta 1 tupla
+
+\item $(...) \Join_{C5} (...)$: (hash join) 235+3=238 bloques
+
+\item Proyección: 238 bloques
+
+\end{itemize}
+
+El costo es $391+79+3168+25+238+238=4139$ accesos a bloques.
+
 % Pregunta d
  \item
+
+Hay que hacer proyección con eliminación de duplicados. Usando la técnica basada
+en \emph{hashing}, el costo es
+
+$391+79+3168+25+238+(2N)/1024$ donde $N$ es el número de tuplas resultado.
 

### 2009-07-20 bdhetero/tarea2/tarea2.tex
pregunta f de tarea2 de bdhetero

@@
 
-\item Proyección: 238 bloques
-
 \end{itemize}
 
-El costo es $391+79+3168+25+238+238=4139$ accesos a bloques.
+El costo es $391+79+3168+25+238=3901$ accesos a bloques.
 
@@
 
-Hay que hacer proyección con eliminación de duplicados. Usando la técnica basada
-en \emph{hashing}, el costo es
-
-$391+79+3168+25+238+(2N)/1024$ donde $N$ es el número de tuplas resultado.
-
 % Pregunta e
@@
 
+Para los operadores de selección no tenemos índices sobre los atributos de selección. Por esto
+se implementan como un recorrido completo de la relación. Para el join más profundo no tenemos índice
+sobre el atributo de join y tampoco cabe la relación izquierda en memoria principal, por lo que no podemos
+usar Block Nested Loop Join ni Index Nested Loop Join. Entre los operadores restantes, el más barato
+es Hash Join.
+
+Para el join menos profundo, no podemos usar Block Nested Loop Join por falta de memoria principal, y tampoco
+Index Nested Loop Join por ser una operación sobre resultados intermedios. Entonces el operador más barato
+es también Hash Join.
+
+El costo de esta solución es el visto en la parte c, es decir, 3901 accesos a bloques.
+Esta solución no es óptima porque --como veremos más adelante-- hay soluciones de menor costo.
+
 % Pregunta f
  \item
+
+\begin{itemize}
+ \item Tamaño 1: PELICULA: 977 bloques. Selección sobre PRODUCE: 470 bloques. Selección sobre COMPAÑIA: 25 bloques.
+       Se elige selección sobre COMPAÑIA. Costo hasta ahora: $25$ bloques, y tenemos una tupla de COMPAÑIA.
+ \item Tamaño 2: Sólo se puede hacer join con PRODUCE. Se hace join de la selección de COMPAÑIA con la selección de PRODUCE.
+       Costo hasta ahora: $25+3*1+3*470=25+1413$ bloques, y tenemos $4$ tuplas del join (por las estadísticas).
+ \item Tamaño 3: Sólo queda PELICULA. Podemos usar Block Nested Loop Join. Costo hasta ahora: $25+1413+1+1*2931=4370$ bloques.
+\end{itemize}
 

### 2009-07-21 bdhetero/tarea2/tarea2.tex
preguntas j,k de tarea2 de bdhetero
parte de prog. din. de tarea2 de bdhetero avanzada

@@
 
+\item Proyección: 238 bloques
+
 \end{itemize}
 
-El costo es $391+79+3168+25+238=3901$ accesos a bloques.
+El costo es $391+79+3168+25+238+238=4139$ accesos a bloques.
 
@@
  \item
+
+Hay que hacer proyección con eliminación de duplicados. Usando la técnica basada
+en \emph{hashing}, el costo es
+
+$391+79+3168+25+238+(2N)/1024$ donde $N$ es el número de tuplas resultado.
 
@@
 es también Hash Join.
-
-El costo de esta solución es el visto en la parte c, es decir, 3901 accesos a bloques.
-Esta solución no es óptima porque --como veremos más adelante-- hay soluciones de menor costo.
 
@@
 
+Criterio de equivalencia: incluye las mismas relaciones.
+
+\begin{itemize}
+ \item Iteración 1: PELICULA: 977 bloques. Selección sobre PRODUCE: 470 bloques. Selección sobre COMPAÑIA: 25 bloques.
+ \item Iteración 2: Hash join entre PELICULA y PRODUCE: 4341 bloques. Hash join entre PRODUCE y PELICULA: 4341 bloques.
+       Hash join entre PRODUCE y COMPAÑIA: 1413 bloques. Hash join entre COMPAÑIA y PRODUCE: 1413 bloques. Block Nested Loop Join
+       entre COMPAÑIA y PRODUCE: 471 bloques.
+ \item Iteración 3: Hash join entre PELICULA y (Block Nested Loop Join entre COMPAÑIA y PRODUCE): 4344 bloques. Hash join entre
+       COMPAÑIA y (Hash join entre PRODUCE y PELICULA): 13026 bloques.
+\end{itemize}
+
+Resultado: Hash join entre PELICULA y (Block Nested Loop Join entre COMPAÑIA y PRODUCE), costo 4344 bloques.
+
 % Pregunta h
  \item
+
+El algoritmo greedy es un caso particular de programación dinámica donde todos los
+planes son equivalentes. Como se vió en la corrida anterior, no se cumple el principio
+de optimalidad porque no llegamos a la solución óptima.
 
@@
 
+Sea $g_1 = 2^n$ y $g_k = 3g_{k-1}(g_{k-1}-1), k > 1$. Entonces una cota superior es
+
+$$
+\sum_{k=1}^n g_k
+$$
+
 % Pregunta k
  \item
+
+Sea $g_1 = 2^n$ y $g_k = 6g_{k-1}(g_{k-1}-1), k > 1$. Entonces una cota superior es
+
+$$
+\sum_{k=1}^n g_k
+$$
 

### 2009-09-21 administrativo/inscripcion_sd09.tex
directorio administrativo

\documentclass[]{letter}

\usepackage[spanish,activeacute]{babel}

\signature{Daniel Izquierdo}
\address{Daniel Izquierdo \\
Departamento de Computación y Tecnología de la Información \\
Edificio MYS. Segundo Piso, Ofic. 211-A \\
Ext. 3275}

\begin{document}

\pagestyle{empty}

\begin{letter}
\name{}
\date{Sartenejas, 21 de Septiembre de 2009}
\opening{Ing. Mar'ia G. Roiatti. \\ Coordinadora de C. \\ Presente}

Estimada Coordinadora,

Me dirijo a Ud., con el fin de solicitar mi inscripción extemporánea para el
trimestre Septiembre-Diciembre 2009. Es mi intención inscribir los códigos
CI-7312 y TG-8524.

Sin m'as a que hacer menci'on y agradeciendo toda la atenci'on que pueda prestar
a esta solicitud, se despide de Ud.,

\closing{Atentamente,}

\end{letter}

\end{document}

### 2009-09-21 administrativo/inscripcion_sd09.tex
carta de inscripcion sd09 lista

@@
 
-\usepackage[spanish,activeacute]{babel}
+\usepackage[spanish]{babel}
+\usepackage[utf8]{inputenc}
 
-\signature{Daniel Izquierdo}
+\signature{Daniel Izquierdo \\ Carnet \#08-86069 \\ C.I. 17.139.881}
 \address{Daniel Izquierdo \\
-Departamento de Computación y Tecnología de la Información \\
-Edificio MYS. Segundo Piso, Ofic. 211-A \\
+Departamento de Computación y T. I. \\
+Edificio MYS, Segundo Piso, Ofic. 211-A \\
 Ext. 3275}
@@
 \date{Sartenejas, 21 de Septiembre de 2009}
-\opening{Ing. Mar'ia G. Roiatti. \\ Coordinadora de C. \\ Presente}
+\opening{Prof. Soraya Abad \\ Coordinación de Postgrado en Computación \\ Universidad Simón Bolívar \\ Presente}
 
@@
 
-Me dirijo a Ud., con el fin de solicitar mi inscripción extemporánea para el
-trimestre Septiembre-Diciembre 2009. Es mi intención inscribir los códigos
-CI-7312 y TG-8524.
+Me dirijo a Ud. con el fin de solicitar mi inscripción extemporánea en la
+Maestría en Ciencias de la Computación para el trimestre Septiembre-Diciembre
+2009. El motivo de la extemporaneidad del trámite es que estaba indeciso sobre
+las materias a inscribir y quería consultarlo con mi posible tutor que se
+encontraba de viaje durante el período de inscripción normal.
 
-Sin m'as a que hacer menci'on y agradeciendo toda la atenci'on que pueda prestar
+Este trimestre deseo inscribir mi última materia e iniciar mi proyecto de grado.
+Los códigos correspondientes son CI-7312 y TG-8524. Esto es un total de ocho
+créditos.
+
+Sin más a que hacer mención y agradeciendo toda la atención que pueda prestar
 a esta solicitud, se despide de Ud.,

### 2009-09-22 mcdsat/dsgen/dsgen.py
generador de casos

@@ -1,0 +1,41 @@
+#!/usr/bin/env python
+
+import sys
+import random
+
+def sample_reps(pop, n):
+    assert(n > 0)
+    ret = []
+
+    while n > 0:
+        n = n - 1
+
+def query(subgoals, varlist):
+    print "q(%s) :- " % (",".join(varlist))
+
+def view(views, subgoals, varlist):
+    print "view"
+
+def main(views, subgoals, variables, predicates, distinguished):
+    varnum = int(variables)
+
+    varlist = ["X%d" % (i) for i in range(varnum)]
+    predlist = [("p%d" % (i), random.randint(1,varnum)) for i in range(int(predicates))]
+
+    print varlist
+    print predlist
+
+    query(subgoals, varlist)
+    view(views, subgoals, variables)
+
+def usage():
+    print "Usage: %s <views> <subgoals> <variables> <predicates> <distinguished>" % (sys.argv[0])
+
+if __name__ == '__main__':
+    #TODO necesario que views,subgoals,variables,predicates sean ints >= 1, y distinguished porcentaje
+    if len(sys.argv) != 6:
+        usage()
+        sys.exit(1)
+
+    random.seed()
+    main(sys.argv[1], sys.argv[2], sys.argv[3], sys.argv[4], sys.argv[5])

### 2009-09-22 mcdsat/dsgen/dsgen.py
generacion de queries lista

@@ -3,30 +3,52 @@
 import sys
 import random
 
-def sample_reps(pop, n):
-    assert(n > 0)
-    ret = []
+def sample_reps(population, k):
+    assert(k > 0)
 
-    while n > 0:
-        n = n - 1
+    result = [None] * k
 
-def query(subgoals, varlist):
-    print "q(%s) :- " % (",".join(varlist))
+    last = len(population)-1
 
-def view(views, subgoals, varlist):
-    print "view"
+    for i in xrange(k):
+        result[i] = population[random.randint(0,last)]
+
+    return result
+
+def gen_query(sgs_num, varlist, predlist, dist_pc):
+    sgs_predicates = sample_reps(predlist, sgs_num)
+    sgs = [(name,sample_reps(varlist,arity)) for (name,arity) in sgs_predicates]
+    sgs_strings = ["%s(%s)" % (name,",".join(list)) for (name,list) in sgs]
+
+    occurring_vars = []
+
+    for (_,l) in sgs:
+        for v in l:
+            if v not in occurring_vars:
+                #I know this is slow
+                occurring_vars.append(v)
+
+    dist_vars = random.sample(occurring_vars, int(len(occurring_vars)*(dist_pc/100.0)))
+
+    return "q(%s) :- %s" % (",".join(dist_vars), ",".join(sgs_strings))
+
+def gen_view(view_name):
+    pass
+
+def gen_views(views_num, subgoals, varlist):
+    return [gen_view(i) for i in xrange(views_num)]
 
 def main(views, subgoals, variables, predicates, distinguished):
-    varnum = int(variables)
+    sgs_num = int(subgoals)
+    vars_num = int(variables)
+    views_num = int(views)
+    dist_pc = float(distinguished)
 
-    varlist = ["X%d" % (i) for i in range(varnum)]
-    predlist = [("p%d" % (i), random.randint(1,varnum)) for i in range(int(predicates))]
+    varlist = ["X%d" % (i) for i in range(vars_num)]
+    predlist = [("p%d" % (i), random.randint(1,vars_num)) for i in range(int(predicates))]
 
-    print varlist
-    print predlist
-
-    query(subgoals, varlist)
-    view(views, subgoals, variables)
+    q = gen_query(sgs_num, varlist, predlist, dist_pc)
+    vs = gen_views(views_num, sgs_num, variables)
 
 def usage():
     print "Usage: %s <views> <subgoals> <variables> <predicates> <distinguished>" % (sys.argv[0])

### 2009-09-22 mcdsat/dsgen/dsgen.py
generacion de vistas

@@ -3,6 +3,9 @@
 import sys
 import random
 
+TEMPSUBGOALSNUMBERERASEME = 0
+TEMPDISTINGUISHEDPERCENTAGEERASEME = 0
+
 def sample_reps(population, k):
     assert(k > 0)
 
@@ -15,10 +18,13 @@ def sample_reps(population, k):
 
     return result
 
+def argstr(l):
+    return ",".join(l)
+    
 def gen_query(sgs_num, varlist, predlist, dist_pc):
     sgs_predicates = sample_reps(predlist, sgs_num)
     sgs = [(name,sample_reps(varlist,arity)) for (name,arity) in sgs_predicates]
-    sgs_strings = ["%s(%s)" % (name,",".join(list)) for (name,list) in sgs]
+    sgs_strings = ["%s(%s)" % (name, ",".join(l)) for (name,l) in sgs]
 
     occurring_vars = []
 
@@ -30,34 +36,61 @@ def gen_query(sgs_num, varlist, predlist, dist_pc):
 
     dist_vars = random.sample(occurring_vars, int(len(occurring_vars)*(dist_pc/100.0)))
 
-    return "q(%s) :- %s" % (",".join(dist_vars), ",".join(sgs_strings))
+    return "q(%s) :- %s" % (argstr(dist_vars), argstr(sgs_strings))
 
-def gen_view(view_name):
-    pass
+def gen_view(view_name, vars_list, preds_list):
+    global TEMPSUBGOALSNUMBERERASEME
+    global TEMPDISTINGUISHEDPERCENTAGEERASEME
+    sgs_predicates = sample_reps(preds_list, random.randint(1,TEMPSUBGOALSNUMBERERASEME*2))
+    sgs = [(name,sample_reps(vars_list,arity)) for (name,arity) in sgs_predicates]
+    sgs_strings = ["%s(%s)" % (name, argstr(l)) for (name,l) in sgs]
 
-def gen_views(views_num, subgoals, varlist):
-    return [gen_view(i) for i in xrange(views_num)]
+    occurring_vars = []
 
-def main(views, subgoals, variables, predicates, distinguished):
+    for (_,l) in sgs:
+        for v in l:
+            if v not in occurring_vars:
+                #I know this is slow
+                occurring_vars.append(v)
+
+    dist_vars = random.sample(occurring_vars, int(len(occurring_vars)*(TEMPDISTINGUISHEDPERCENTAGEERASEME/100.0)))
+
[…]

### 2009-10-10 ws2/reporte.tex
directorio de ws2, reporte de clase

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}

\newcommand{\tablaTripletas}[2]
{
\begin{tabular}{l|c}
 & Número de tripletas \\
 \hline
 Antes & #1 \\
 Después & #2 \\
\end{tabular}
}

\newcommand{\tablaPredicados}[1]
{
\begin{tabular}{l|c|c}
Predicado & Antes & Después \\
\hline
#1
\end{tabular}
}

\newcommand{\tablaInstancias}[1]
{
\begin{tabular}{l|c|c|c}
Clase & Antes & Después \\
\hline
#1
\end{tabular}
}

\newcommand{\resultadoDataSet}[5]
{
  \newpage

  \section{#1}

  \vspace{0.5cm}

  \tablaTripletas{#2}{#3}

  \vspace{1cm}

  \tablaPredicados{#4}

  \vspace{1cm}

  \tablaInstancias{#5}
}

\begin{document}

\resultadoDataSet{rdfPrimero}{89}{879}
{
    cursa & 14 & 70 \\
    dicta & 16 & 40 \\
    esProfesor & 0 & 27 \\
    esSupervisor & 9 & 45 \\
    esTutor & 0 & 36 \\
    rdf:type & 40 & 507 \\
    rdfs:subClassOf & 8 & 111 \\
    rdfs:subPropertyOf & 2 & 43 \\
}
{
    estudiante & 0 & 32 \\
    estudiantePostGrado & 4 & 20 \\
    estudiantePostGradoOcasional & 4 & 20 \\
    estudiantePreGrado & 4 & 20 \\
    materia & 0 & 32 \\
    materiaPostGrado & 4 & 20 \\
    materiaPreGrado & 4 & 20 \\
    persona & 0 & 40 \\
    profesor & 4 & 20 \\
    rdfs:Class & 11 & 66 \\
    rdfs:Property & 5 & 37 \\
    rdfs:Resource & 0 & 180 \\
}

\resultadoDataSet{rdfSegundo}{120}{1193}
{
    cursa & 14 & 70 \\
    dicta & 16 & 40 \\
    esProfesor & 0 & 54 \\
    esSupervisor & 18 & 90 \\
    esTutor & 0 & 72 \\
    rdf:type & 62 & 713 \\
    rdfs:subClassOf & 8 & 111 \\
    rdfs:subPropertyOf & 2 & 43 \\
}
{
    estudiante & 0 & 32 \\
    estudiantePostGrado & 4 & 20 \\
    estudiantePostGradoOcasional & 24 & 120 \\
    estudiantePreGrado & 4 & 20 \\
    materia & 0 & 40 \\
    materiaPostGrado & 5 & 25 \\
    materiaPreGrado & 5 & 25 \\
    persona & 0 & 40 \\
    profesor & 4 & 20 \\
    rdfs:Class & 11 & 66 \\
    rdfs:Property & 5 & 37 \\
    rdfs:Resource & 0 & 268 \\
}

\resultadoDataSet{rdfTercero}{162}{1489}
{
    cursa & 14 & 70 \\
    dicta & 16 & 40 \\
    esProfesor & 0 & 108 \\
    esSupervisor & 36 & 180 \\
    esTutor & 0 & 144 \\
    rdf:type & 86 & 793 \\
    rdfs:subClassOf & 8 & 111 \\
    rdfs:subPropertyOf & 2 & 43 \\
}
{
    estudiante & 0 & 32 \\
    estudiantePostGrado & 4 & 20 \\
    estudiantePostGradoOcasional & 48 & 160 \\
    estudiantePreGrado & 4 & 20 \\
    materia & 0 & 40 \\
    materiaPostGrado & 5 & 25 \\
    materiaPreGrado & 5 & 25 \\
    persona & 0 & 40 \\
    profesor & 4 & 20 \\
    rdfs:Class & 11 & 66 \\
    rdfs:Property & 5 & 37 \\
    rdfs:Resource & 0 & 308 \\
}

\end{document}

### 2009-10-10 ws2/reporte.tex
reporte terminado

@@
 \begin{document}
+
+\title{Aumento de tripletas RDF con axiomatización}
+\date{}
+\author{Daniel Izquierdo}
+
+\maketitle
+
+% DESCRIPCIÓN
+
+Con respecto al número de tripletas, con el pequeño conjunto de casos se observa
+aumento lineal de tripletas generadas por cantidad de tripletas preexistentes.
+El número de tripletas se multiplica aproximadamente por diez.
+
+Pasando al aumento de cláusulas por predicado usado podemos ver que los
+predicados que más contribuyen son los definidos por RDFS, particularmente
+{\tt rdfs:type}. Esto no sorprende porque la mayoría de los axiomas introducen
+cláusulas con este predicado. Los predicados definidos por el usuario registran
+aumento considerable pero no tan alto comparativamente. Algo importante es la
+introducción de muchas tripletas con predicados que no aparecían previamente.
+
+Viendo las tripletas por instancias también se nota que crecen más las tripletas
+relacionadas con clases de RDFS. El mayor incremento, por mucho, es el de {\tt
+rdfs:Resource}. Podemos observar que se introducen varias instancias de clases
+que previamente no estaban instanciadas.
+
+% RESULTADOS
 

### 2009-11-21 tesis/src/imcdsat/mcdsat/CQ/CQ.py
hack para que funcionen las reescrituras

@@ -4,6 +4,10 @@ from Seq import *
 CONSTANT_BASE = 1000000000
 
 def es_const(x):
+    #hack para las reescrituras que colocan la x antes
+    if type(x) == str:
+        x = x.lstrip('_XYZ')
+
     if type(x) != int:
         x = int(x)
 

### 2009-11-22 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
usando clausulas para constantes en teoria extendida

@@ -121,7 +121,8 @@ def clausulasCombinarMCD(transf, n, q, vistas):
                                 except KeyError:
                                     continue
 
-                                #clausulas2.append("%s %s 0\n" % (str(numVarT0), str(numVarT1)))
+                                #clausulas para constantes en teoria extendida
+                                clausulas2.append("%s %s 0\n" % (str(numVarT0), str(numVarT1)))
 
 
     return clausulas2

### 2009-11-23 ws2/tarea2.tex
.tex de tarea2 de ws2
codigo copiado de squin en tex de tarea2 de ws2

--- 2009-10-10 ws2/reporte.tex
@@
 \usepackage[spanish]{babel}
-
-\newcommand{\tablaTripletas}[2]
-{
-\begin{tabular}{l|c}
- & Número de tripletas \\
- \hline
- Antes & #1 \\
- Después & #2 \\
-\end{tabular}
-}
-
-\newcommand{\tablaPredicados}[1]
-{
-\begin{tabular}{l|c|c}
-Predicado & Antes & Después \\
-\hline
-#1
-\end{tabular}
-}
-
-\newcommand{\tablaInstancias}[1]
-{
-\begin{tabular}{l|c|c|c}
-Clase & Antes & Después \\
-\hline
-#1
-\end{tabular}
-}
-
-\newcommand{\resultadoDataSet}[5]
-{
-  \newpage
-
-  \section{#1}
-
-  \vspace{0.5cm}
-
-  \tablaTripletas{#2}{#3}
-
-  \vspace{1cm}
-
-  \tablaPredicados{#4}
-
-  \vspace{1cm}
-
-  \tablaInstancias{#5}
-}
+\usepackage{fancyvrb}
 
@@
 
-\resultadoDataSet{rdfPrimero}{89}{879}
+\title{Consultas }
+\date{}
+\author{Daniel Izquierdo \\ \#08-86809}
+
+\maketitle
+
+\section{Medicamentos alternativos}
+
+En el ejemplo de \emph{SQUIN} para obtener alternativas para medicamentos dados
+se utilizan tres consultas distintas.
+
+\fvset{fontsize=\footnotesize}
+
+\begin{SaveVerbatim}{VerbDrugLookupQuery}
+PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+PREFIX drugbank: <http://www4.wiwiss.fu-berlin.de/drugbank/resource/drugbank/>
+SELECT ?uri ?label WHERE {
+  ?uri a drugbank:drugs .
+  ?uri rdfs:label ?label .
+  FILTER( regex(STR(?label),'MEDICAMENTO','i') )
+}
+ORDER BY ?uri"
+\end{SaveVerbatim}
+
+\begin{SaveVerbatim}{VerbAltDrugQuery}
+PREFIX rdf:  <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
+PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+PREFIX owl:  <http://www.w3.org/2002/07/owl#>
+PREFIX drugbank: <http://www4.wiwiss.fu-berlin.de/drugbank/resource/drugbank/>
+PREFIX tcm:      <http://purl.org/net/tcm/tcm.lifescience.ntu.edu.tw/>
+SELECT DISTINCT ?disease ?diseaseLabel ?altMedicine ?altMedicineLabel WHERE {
+  <uri> drugbank:possibleDiseaseTarget ?disease.
+  ?disease owl:sameAs ?sameDisease.
+  ?sameDisease rdf:type tcm:Disease.
+  ?altMedicine tcm:treatment ?sameDisease.
+  ?altMedicine rdf:type tcm:Medicine.
+  ?disease rdfs:label ?diseaseLabel.
+  ?altMedicine rdfs:label ?altMedicineLabel.
+}
+ORDER BY ?sameDisease
+\end{SaveVerbatim}
+
+\begin{SaveVerbatim}{VerbSideEffectQuery}
+PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+PREFIX owl:  <http://www.w3.org/2002/07/owl#>
+PREFIX tcm:      <http://purl.org/net/tcm/tcm.lifescience.ntu.edu.tw/>
+PREFIX dailymed: <http://www4.wiwiss.fu-berlin.de/dailymed/resource/dailymed/>
+PREFIX sider:    <http://www4.wiwiss.fu-berlin.de/sider/resource/sider/>
+SELECT DISTINCT ?effect ?effectname WHERE {
+        <" + uri + "> tcm:ingredient ?ingredient .
+        ?dailymedIngredient owl:sameAs ?ingredient .
+        ?drug dailymed:activeIngredient ?dailymedIngredient .
+        ?drug owl:sameAs ?sider_drug .
+        ?sider_drug a sider:drugs .
+        ?sider_drug sider:sideEffect ?effect .
+        ?effect rdfs:label ?effectname .
+}
+ORDER BY ?effectname
+\end{SaveVerbatim}
+
+\begin{figure}[htbp]
+ \fbox
  {
-    cursa & 14 & 70 \\
-    dicta & 16 & 40 \\
-    esProfesor & 0 & 27 \\
-    esSupervisor & 9 & 45 \\
-    esTutor & 0 & 36 \\
-    rdf:type & 40 & 507 \\
-    rdfs:subClassOf & 8 & 111 \\
-    rdfs:subPropertyOf & 2 & 43 \\
+  \begin{minipage}{\textwidth}
+   \BUseVerbatim{VerbDrugLookupQuery}
+   \caption{Consulta de búsqueda de medicamentos}
+   \label{src:DrugLookupQuery}
+  \end{minipage}
  }
+\end{figure}
+
+\begin{figure}[htbp]
+ \fbox
  {
-    estudiante & 0 & 32 \\
-    estudiantePostGrado & 4 & 20 \\
-    estudiantePostGradoOcasional & 4 & 20 \\
-    estudiantePreGrado & 4 & 20 \\
-    materia & 0 & 32 \\
-    materiaPostGrado & 4 & 20 \\
-    materiaPreGrado & 4 & 20 \\
-    persona & 0 & 40 \\
-    profesor & 4 & 20 \\
-    rdfs:Class & 11 & 66 \\
-    rdfs:Property & 5 & 37 \\
-    rdfs:Resource & 0 & 180 \\
+  \begin{minipage}{\textwidth}
+   \BUseVerbatim{VerbAltDrugQuery}
+   \caption{Consulta de búsqueda de alternativas}
+   \label{src:AltDrugQuery}
+  \end{minipage}
  }
+\end{figure}
 
-\resultadoDataSet{rdfSegundo}{120}{1193}
+\begin{figure}[htbp]
+ \fbox
  {
-    cursa & 14 & 70 \\
-    dicta & 16 & 40 \\
-    esProfesor & 0 & 54 \\
-    esSupervisor & 18 & 90 \\
-    esTutor & 0 & 72 \\
-    rdf:type & 62 & 713 \\
-    rdfs:subClassOf & 8 & 111 \\
-    rdfs:subPropertyOf & 2 & 43 \\
+  \begin{minipage}{\textwidth}
+   \BUseVerbatim{VerbSideEffectQuery}
+   \caption{Consulta de búsqueda de efectos secundarios}
+   \label{src:SideEffectQuery}
+  \end{minipage}
  }
-{
-    estudiante & 0 & 32 \\
-    estudiantePostGrado & 4 & 20 \\
-    estudiantePostGradoOcasional & 24 & 120 \\
-    estudiantePreGrado & 4 & 20 \\
-    materia & 0 & 40 \\
-    materiaPostGrado & 5 & 25 \\
-    materiaPreGrado & 5 & 25 \\
-    persona & 0 & 40 \\
-    profesor & 4 & 20 \\
-    rdfs:Class & 11 & 66 \\
-    rdfs:Property & 5 & 37 \\
-    rdfs:Resource & 0 & 268 \\
-}
+\end{figure}
 
-\resultadoDataSet{rdfTercero}{162}{1489}
-{
-    cursa & 14 & 70 \\
-    dicta & 16 & 40 \\
-    esProfesor & 0 & 108 \\
-    esSupervisor & 36 & 180 \\
-    esTutor & 0 & 144 \\
-    rdf:type & 86 & 793 \\
-    rdfs:subClassOf & 8 & 111 \\
-    rdfs:subPropertyOf & 2 & 43 \\
-}
-{
-    estudiante & 0 & 32 \\
-    estudiantePostGrado & 4 & 20 \\
-    estudiantePostGradoOcasional & 48 & 160 \\
-    estudiantePreGrado & 4 & 20 \\
-    materia & 0 & 40 \\
-    materiaPostGrado & 5 & 25 \\
-    materiaPreGrado & 5 & 25 \\
-    persona & 0 & 40 \\
-    profesor & 4 & 20 \\
-    rdfs:Class & 11 & 66 \\
-    rdfs:Property & 5 & 37 \\
-    rdfs:Resource & 0 & 308 \\
-}
+\section{Medicamentos para enfermedades específicas}
 

### 2009-11-23 ws2/tarea2.tex
texto de primera parte de tarea2

@@
 \begin{document}
+
+\fvset{fontsize=\footnotesize}
 
@@
 En el ejemplo de \emph{SQUIN} para obtener alternativas para medicamentos dados
-se utilizan tres consultas distintas.
-
-\fvset{fontsize=\footnotesize}
+se utilizan tres consultas distintas. Para empezar se necesita buscar el
+\emph{uri} del nombre de medicamento dado por el usuario. No necesariamente el
+nombre corresponde a un medicamento y puede haber varios medicamentos cuyos
+nombres correspondan al patrón. Por esto se presenta al usuario una lista de
+medicamentos que corresponden. La lista se obtiene con la consulta que podemos
+observar en la figura \ref{src:DrugLookupQuery}. La palabra ``MEDICAMENTO'' se
+sustituye por el nombre o patrón dado.
 
@@
 ORDER BY ?uri"
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}{VerbAltDrugQuery}
-PREFIX rdf:  <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
-PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
-PREFIX owl:  <http://www.w3.org/2002/07/owl#>
-PREFIX drugbank: <http://www4.wiwiss.fu-berlin.de/drugbank/resource/drugbank/>
-PREFIX tcm:      <http://purl.org/net/tcm/tcm.lifescience.ntu.edu.tw/>
-SELECT DISTINCT ?disease ?diseaseLabel ?altMedicine ?altMedicineLabel WHERE {
-  <uri> drugbank:possibleDiseaseTarget ?disease.
-  ?disease owl:sameAs ?sameDisease.
-  ?sameDisease rdf:type tcm:Disease.
-  ?altMedicine tcm:treatment ?sameDisease.
-  ?altMedicine rdf:type tcm:Medicine.
-  ?disease rdfs:label ?diseaseLabel.
-  ?altMedicine rdfs:label ?altMedicineLabel.
-}
-ORDER BY ?sameDisease
-\end{SaveVerbatim}
-
-\begin{SaveVerbatim}{VerbSideEffectQuery}
-PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
-PREFIX owl:  <http://www.w3.org/2002/07/owl#>
-PREFIX tcm:      <http://purl.org/net/tcm/tcm.lifescience.ntu.edu.tw/>
-PREFIX dailymed: <http://www4.wiwiss.fu-berlin.de/dailymed/resource/dailymed/>
-PREFIX sider:    <http://www4.wiwiss.fu-berlin.de/sider/resource/sider/>
-SELECT DISTINCT ?effect ?effectname WHERE {
-        <" + uri + "> tcm:ingredient ?ingredient .
-        ?dailymedIngredient owl:sameAs ?ingredient .
-        ?drug dailymed:activeIngredient ?dailymedIngredient .
-        ?drug owl:sameAs ?sider_drug .
-        ?sider_drug a sider:drugs .
-        ?sider_drug sider:sideEffect ?effect .
-        ?effect rdfs:label ?effectname .
-}
-ORDER BY ?effectname
 \end{SaveVerbatim}
@@
 
+Teniendo un \emph{uri} específico del medicamento se puede comenzar a buscar sus
+alternativas. Esto se hace con la consulta de la figura \ref{src:AltDrugQuery}
+sustituyendo la palabra ``URI'' por el \emph{uri} correcto. Vemos que la
+consulta involucra varios documentos y básicamente consiste en listar las
+enfermedades tratadas por el medicamento dado, obtener medicamentos que tratan
+esa enfermedad y ordenar el resultado por enfermedad.
+
+\begin{SaveVerbatim}{VerbAltDrugQuery}
+PREFIX rdf:  <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
+PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+PREFIX owl:  <http://www.w3.org/2002/07/owl#>
+PREFIX drugbank: <http://www4.wiwiss.fu-berlin.de/drugbank/resource/drugbank/>
+PREFIX tcm:      <http://purl.org/net/tcm/tcm.lifescience.ntu.edu.tw/>
+SELECT DISTINCT ?disease ?diseaseLabel ?altMedicine ?altMedicineLabel WHERE {
+  <URI> drugbank:possibleDiseaseTarget ?disease.
+  ?disease owl:sameAs ?sameDisease.
+  ?sameDisease rdf:type tcm:Disease.
+  ?altMedicine tcm:treatment ?sameDisease.
+  ?altMedicine rdf:type tcm:Medicine.
+  ?disease rdfs:label ?diseaseLabel.
+  ?altMedicine rdfs:label ?altMedicineLabel.
+}
+ORDER BY ?sameDisease
+\end{SaveVerbatim}
+
 \begin{figure}[htbp]
@@
 \end{figure}
+
+Opcionalmente el usuario puede consultar efectos secundarios para cada una de
+las drogas alternativas obtenidas. En ese caso se realiza la consulta de la
+figura \ref{src:SideEffectQuery} haciendo la misma sustitución en ``URI''. Los
+documentos consultados nos proporcionan los ingredientes de la droga en
+cuestión y se procede a buscar efectos secundarios de medicamentos tales que
+alguno de sus ingredientes activos está en esa lista.
+
+\begin{SaveVerbatim}{VerbSideEffectQuery}
+PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+PREFIX owl:  <http://www.w3.org/2002/07/owl#>
+PREFIX tcm:      <http://purl.org/net/tcm/tcm.lifescience.ntu.edu.tw/>
+PREFIX dailymed: <http://www4.wiwiss.fu-berlin.de/dailymed/resource/dailymed/>
+PREFIX sider:    <http://www4.wiwiss.fu-berlin.de/sider/resource/sider/>
+SELECT DISTINCT ?effect ?effectname WHERE {
+        <URI> tcm:ingredient ?ingredient .
+        ?dailymedIngredient owl:sameAs ?ingredient .
+        ?drug dailymed:activeIngredient ?dailymedIngredient .
+        ?drug owl:sameAs ?sider_drug .
+        ?sider_drug a sider:drugs .
+        ?sider_drug sider:sideEffect ?effect .
+        ?effect rdfs:label ?effectname .
+}
+ORDER BY ?effectname
+\end{SaveVerbatim}
 

### 2009-11-25 ws2/src/SquinLib/src/ve/usb/squinlib/Main.java
squinlib completo

@@ -1,0 +1,34 @@
+package ve.usb.squinlib;
+
+import java.io.BufferedReader;
+import java.io.IOException;
+import java.io.InputStreamReader;
+import java.net.MalformedURLException;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+
+/**
+ * @author Daniel Izquierdo
+ */
+public class Main {
+
+    public static void main(String[] args) {
+        SparqlQuerySock sqs = new SparqlQuerySock("http://localhost:21786/SQUIN/query");
+
+        BufferedReader in = new BufferedReader(new InputStreamReader(System.in));
+        StringBuffer sb = new StringBuffer();
+        String line;
+
+        try {
+            while ((line = in.readLine()) != null) {
+                sb.append(line + '\n');
+            }
+
+            String result = sqs.call(sb.toString());
+
+            System.out.println(result);
+        } catch (Exception ex) {
+            System.out.println(ex);
+        }
+    }
+}

### 2009-11-25 ws2/src/SquinLib/src/ve/usb/squinlib/SparqlQuerySock.java

@@ -1,0 +1,78 @@
+package ve.usb.squinlib;
+
+import java.io.BufferedReader;
+import java.io.IOException;
+import java.io.InputStreamReader;
+import java.io.UnsupportedEncodingException;
+import java.net.MalformedURLException;
+import java.net.URL;
+import java.net.URLEncoder;
+import java.util.HashMap;
+import java.util.Map;
+import java.util.logging.Level;
+import java.util.logging.Logger;
+
+/**
+ * @author Daniel Izquierdo
+ */
+public class SparqlQuerySock {
+
+    private final String queryServer;
+    private final static Logger logger = Logger.getLogger(SparqlQuerySock.class.getCanonicalName());
+
+    public SparqlQuerySock(final String queryServer) {
+        logger.setLevel(Level.FINE);
+        this.queryServer = queryServer;
+    }
+
+    public String call(final String query) throws MalformedURLException, IOException {
+        logger.log(Level.INFO, "SparqlQuerySock call");
+
+        Map<String, String> request_params = new HashMap<String, String>();
+        request_params.put("query", query);
+
+        logger.log(Level.INFO, "query server: " + queryServer);
+        logger.log(Level.INFO, "query: " + query);
+
+        String requestUrl = createQueryURLString(this.queryServer, request_params);
+
+        logger.log(Level.INFO, "request url: " + requestUrl);
+
+        URL url = new URL(requestUrl);
+
+        BufferedReader in = new BufferedReader(new InputStreamReader(url.openStream()));
+        StringBuffer output = new StringBuffer();
+        String line;
+
+        while ((line = in.readLine()) != null) {
+            output.append(line);
+        }
+
+        return output.toString();
+    }
+
+    public static String createQueryURLString(String url, Map params) {
+        return url + '?' + createQueryString(params);
+    }
+
+    public static String createQueryString(Map params) {
+        StringBuffer queryString = new StringBuffer();
+        boolean first = true;
+
+        for (Object key : params.keySet()) {
+            if (first) {
+                first = false;
+            } else {
+                queryString.append('&');
+            }
+            try {
+                queryString.append(key.toString() + '=' + URLEncoder.encode(params.get(key).toString(), "UTF-8"));
+            } catch (UnsupportedEncodingException ex) {
+                /* realmente no debería pasar */
+                Logger.getLogger(SparqlQuerySock.class.getName()).log(Level.SEVERE, null, ex);
+            }
+        }
+
+        return queryString.toString();
+    }
+}

### 2009-11-26 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
soporte incompleto para generacion reducida de variables por transitividad

@@ -219,8 +219,8 @@ def generarTeoriaMCD(q, vistas):
     #print "clausulas d6  vi => -t_{x,A} si A no aparece en cuerpo de vi"
     #pprint.pprint(d6)
 
-    #clausulas = clausulas + d1 + d2 + d3 + d4 + d5 + d6
-    clausulas = clausulas + d1 + d2 + d3
+    clausulas = clausulas + d1 + d2 + d3 + d4 + d5 + d6
+    #clausulas = clausulas + d1 + d2 + d3
 
     return variables, clausulas
 
@@ -577,6 +577,48 @@ def clausula78a(varz, varg, varm, subObQ, subObV, vis, ltaux, c7, c8, vista):
 def clausulas_d4(q, vistas, variables_query, constantes_query):
     global varsT
     global varsV
+
+    #ejele
+
+    mappings = set()
+    newmappings = set()
+
+    for v in vistas:
+        for soq in q.cuerpo:
+            for sov in v.cuerpo:
+                if soq.predicado == sov.predicado:
+                    for (x, y) in zip(soq.orden, sov.orden):
+                        newmappings.add((x, y))
+
+    foundnewmappings = True
+
+    while foundnewmappings:
+        mappings.update(newmappings)
+        newmappings = set()
+
+        foundnewmappings = False
+
+        for (a, y0) in mappings:
+            if es_var(a) or es_const(y0):
+                continue
+
+            for (x1, y1) in mappings:
+                if es_const(x1) or (y0 != y1) or (x1 == y1):
+                    continue
+
+                for (x2, z) in mappings:
+                    if (x1 != x2) or es_const(z) or (x2 == z):
+                        continue
+
+                    if (a, z) not in mappings:
+                        foundnewmappings = True
+                        print "SI"
+                        print "%s por culpa de que %s y %s y %s" % (str((a,z)), str((a,y0)), str((x1,y1)), str((x2,z)))
+                        newmappings.add((a, z))
+
+    import sys
+    sys.exit(1)
+    return 0
 
     d4set = set([])
 

### 2009-11-29 mcdsat/paper/paper.tex
paper inicial de mcdsat rechazado

\documentclass{sig-alternate}

\usepackage{enumerate}

\makeatletter
\let\@copyrightspace\relax
\makeatother

\def\sharedaffiliation{%
\end{tabular}
\begin{tabular}{c}}

\newcommand{\denselist}{\topsep 0pt \itemsep -4pt}
\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\vvec}[1]{\mathbf{#1}}
\newcommand{\join}{\bowtie}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\body}{{body}}
\newcommand{\head}{{head}}
\newcommand{\qrule}{:\!\!-}
\newcommand{\arc}{\text{arc}}

\begin{document}
\allowdisplaybreaks

\title{Computing Minimum-Cost Rewritings Using Circuits}
\subtitle{[Extended Abstract]}
\numberofauthors{3}
\author{
\alignauthor Daniel Izquierdo \email{idaniel@ldc.usb.ve}
\alignauthor Mar\'{\i}a Esther Vidal \email{mvidal@ldc.usb.ve}
\alignauthor Blai Bonet \email{bonet@ldc.usb.ve}
\sharedaffiliation
\affaddr{Departamento de Computaci\'on} \\
\affaddr{Universidad Sim\'on Bol\'{\i}var} \\
\affaddr{Caracas, Venezuela}
}
\maketitle

\newtheorem{theorem}{Theorem}

\begin{abstract}
We show how the logical translation used in the McdSat algorithm
can be used as well to compute the minimum-cost rewritings with
respect to a simple additive cost model. Our algorithm exploits
a known relation between d-DNNF theories and arithmetic circuits.
\end{abstract}

\section{Introduction}

Recently, we considered the problem of computing all the rewritings
of a query in terms of materialized views \cite{halevy:survey}.
This problem is important in the context of data integration
\cite{duschka:plan,kwok:plan,lambrecht:plan},
and query optimization and data maintenance
\cite{levy:bucket,zaharioudakis:table,mitra:svbt}.
Our approach consisted in translating the problem of rewriting
into the problem of enumerating the models of a propositional
theory \cite{arvelo:aaai06} which is compiled into deterministic
and decomposable negation normal form (d-DNNF) in order to
enumerate the models efficiently \cite{darwiche:dnnf}.

The theory is a CNF formula whose models are in correspondence
with the rewritings of the query. This theory is constructed
following the MiniCon Descriptions (MCDs) and their possible compositions
\cite{pottinger:minicon}.
The logical approach demonstrated to scale better than the MiniCon
algorithm over a large number of benchmarks often showing performance
improvements of several orders of magnitude.

However, the McdSat algorithm, the name of the logical approach,
was not desgined for rewriting problems involving explicit constants,
nor to compute the best rewritings with respect to a given cost model.
In this work, we address these shortcommings by provinding an
extended logical formulation capable of handling the constants,
and by showing how the d-DNNF can be exploited in order to compute
the best rewritings for a simple yet interesting cost models.

The paper is organized as follows.
We define the problem, the translation, and the cost models (Sect.~2).
We show how the theory can be extended to support explicit constants (Sect.~3)
and how the d-DNNF is exploited to find best rewritings (Sect.~4).
We end with some experiments (Sect.~5) and a discussion (Sect.~6).

\section{Problem, Theories and Costs}

We consider databases of the form $D=\tup{P,T}$ where
$P$ be a set of predicates and $T=\{T_p\}_{p\in P}$ is a collection
of tables that represents the predicates in extensional form.
A conjunctive query $Q$ over $P$ is of the form 
\begin{alignat*}{1}
Q(\vvec{x})\ \qrule\ \  p_1(\vvec{x}_1),\, p_1(\vvec{x}_2),\, \ldots,\, p_m(\vvec{x}_m)\,,
\end{alignat*}
where $p_i\in P$, $\vvec{x}$ is a vector of variables, and each
$\vvec{x}_i$ is a vector of variables and constants.
The result of $Q$ over $D$, denoted as $Q(D)$, is the table with
$|\vvec{x}|$ columns that result of the projection of the relational
join $\join\!\!\{T_{p_i}\}_{i=1}^m$ over $\vvec{x}$.
The atoms in the body of $Q$ are called the (sub)goals of $Q$.

A view $V$ over $D$ is a query over $P$. In the context of data
integration, the database $D$ is an idealized description of 
the information contained across multiple data sources described
as views.
Given a database $D$, a query $Q$ and a collection of 
views $E=\tup{\{V_i\}_i,\{E_i\}_i}$, we are required to find
all the tuples in $Q(D)$ obtainable from the views in $E$.
That is, we need to find all the \emph{rewritings} of the form
\begin{alignat*}{1}
R(\vvec{x})\ \qrule\ \ V_{i_1}(\vvec{x}_1),\, V_{i_2}(\vvec{x}_2),\, \ldots,\, V_{i_n}(\vvec{x}_n)
\end{alignat*}
such that $R(E) \subseteq Q(D)$.
A query rewriting problem (QRP) is a tuple $\tup{P,Q,\{V_i\}}$ where $P$
is a set of predicates, $Q$ is a query over $P$ and $\{V_i\}$ is a collection
of views. We assume \emph{safe} problems in the sense that all 
variables mentioned in the head of the query (resp.\ in the head of each view)
appear in the body of the query (resp.\ in the body of each view).
Further, we only deal with QRPs with no arithmetic predicates inside the
query or views.
A rewriting $R$ is \emph{valid} if for all databases $D=\tup{P,T}$
and extensions $\{E_i\}$, $R(E) \subseteq Q(D)$.
A collection $\R$ of valid rewritings is a solution if
for all databases $D=\tup{P,T}$ and extensions $\{E_i\}$, there
is no other $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(D)$.
We are interested in finding a rewriting $\R$.

\subsection{Logical Theories}

In \cite{arvelo:aaai06}, we showed that a rewriting
for a query $Q$ with $m$ goals can be obtained by enumerating the
models of a logical theory
$\Delta=\Delta_{rew}\cup\Delta_{mcd}^1\cup\cdots\Delta_{mcd}^m$
where $\Delta_{rew}$ specified how to combine $m$ independent
copies of MCD theories $\Delta_{mcd}$ that cover all goals in $Q$.
Each $\Delta^i_{mcd}$ is a copy of the theory $\Delta_{mcd}$
in which each literal $p$ is tagged as $p^i$.
The theory $\Delta_{mcd}$ consists of different groups of 
clauses that guarantess that its models are in correspondence
with the MCDs, while the theory $\Delta_{rew}$ contains additional
clauses to guarantee a sound and complete composition of the MCDs.
The reader is referred to \cite{arvelo:aaai06} for a comprehensive
description of the propositional theory.

Let us describe the difficulties that arise when constants
are presents in a QRP with the following example:
\begin{alignat*}{1}
Q(x,z)\ &\qrule\ \ p_1(x,y),\, p_2(y,z)\,, \\
V_1(x_1)\ &\qrule\ \ p_1(x_1,A)\,, \\
V_2(x_2)\ &\qrule\ \ p_2(B,x_2)\,, \\
V_3(x_3,y_3)\ &\qrule\ \ p_2(x_3,y_3)
%V_4(x_4,y_4)\ &\qrule\ \ p_1(x_1,A),\, p_2(B,y_4)\,,
\end{alignat*}
where $Q$ is the query and $\{V_1,V_2,V_3\}$ are the 
views. In this case, the only rewriting is 
$R(x,z)\qrule V_1(x),V_3(A,z)$ because the candidate 
$R(x,z)\qrule V_1(x),V_2(z)$ is not valid as it maps
$y$ into constants $A$ and $B$ that denote different
objects.

The main problem when handling constants is to
be sure that different constants are not mapped into
each other either directly or indirectly (via transitivity).

\subsection{Cost Models}

We assume a simple additive cost model in which each view
$V_i$ is associated with a cost $c(V_i)$, and the cost of
a rewriting is the sum of the costs for the views in it.
An optimal or best rewriting is one with minimum cost,
and the optimal cost of a QRP is the cost of a best
rewriting. A QRP has always a well-defined optimal
cost (if there are no rewritings, its cost is $\infty$),
but it may have multiple best rewritings.
The rewriting problem with costs consists in finding
all the rewritings of minimum cost.

%The cost model is simple yet expressive; e.g., if all
%costs are unit, then best rewritings are those with
%minimum number of views, yet in cases where views
%represent different sources, it is natural
%to associate them with different costs.

\section{Extended Theories}

The theory $\Delta_{mcd}$ makes use of propositions $t_{x,y}$
to denote that the variable/constant $x$ in the query is mapped
into the variable/constant $y$ in the view, and propositions $v_i$
to indicate that the MCD uses view $V_i$.
We obtain a sound and complete theory for QRPs with constants if
$\Delta_{mcd}$ is extended with the clauses:
\begin{enumerate}[C1.]\denselist
\item (Inconsistent-1): $t_{x,A} \Rightarrow \neg t_{x,B}$,
\item (Inconsistent-2): $t_{A,x} \Rightarrow \neg t_{B,x}$,
\item (Inconsistent-3): $\neg t_{A,B}$,
\item (Transitivity-1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$,
\item (Transitivity-2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
\end{enumerate}
Clauses C1--C3 prune candidate rewritings in which one constant
is directly mapped into a different one, and the last two implement
a restricted propagation of transitivity.
Similarly, $\Delta_{rew}$ must be extended with the clauses:
\begin{enumerate}[C1.]\denselist
\item[C6.] (Inconsistent-4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
\end{enumerate}
XXXX showed that when the query has $m$ goals and there
are at most $n$ dfferent constant symbols, it is enough
to consider rewritings of length at most XXXXXX \cite{ref}.

Given a QRP problem $\tup{Q,\{V_i\}}$ possibly with constant 
symbols, we construct logical theories whose models are in
correspondence with the rewritings. These theories are then
compiled into d-DNNF from which all rewritings are extracted
efficiently.
Likewise, best rewritings can be computed by performing model
enumeration twice: the first pass computes the optimal cost,
and the second filters out suboptimal rewritings.
However, there is a better way.

\section{Minimum-Cost Models}

Darwiche \& Marquis \cite{darwiche:weighted} show how to compute the
rank $r^*(\Delta)$ of a propositional theory $\Delta$
efficiently when $r$ is a literal ranking function and
$\Delta$ is in d-DNNF. A literal ranking function ranks
the models of the theory in terms of the rank of the
literals $l$ that are true in the model, and ranks the
theory as the best rank of its models:
\[ r(\omega) = \sum_{\omega\vDash l} r(l),\quad r^*(\Delta) = \min_{\omega\vDash\Delta} r(\omega)\,. \]
Furthermore, not only $r^*(\Delta)$ can be computed
efficiently from the d-DNNF but also the models that
have such rank.
The procedure for computing the rank converts the d-DNNF
into a circuit in which the literals are replaced by their
ranks, the `or' nodes by `min', and the `and' nodes by `sum'.
The evaluation of the circuit computes the rank of $\Delta$.

We thus obtain a simple method for computing the best
rewritings when the literal ranking function $r_c$ induced
by the cost model $c$ is used; $r_c$ is defined by
$r_c(l)=c(V_i)$ if $l=v^t_i$ for some $t$, and $r_c(l)=0$
otherwise.

\section{Preliminary Experiments}

[brief explanation of results; no tables, no graphs; leave them for the poster]

\section{Discussion}

We have shown how the propositional theory used in McdSat can
be extended to support constants, and how it can be used to 
compute the best rewritings with respect to an additive
cost model. The cost model is simple yet expressive.
The computation of the best models relies on an arithmetic
circuit that is obtained from the d-DNNF of the propositional
theory.

\small
\bibliography{../../control}
\bibliographystyle{plain}

\end{document}

### 2009-11-29 tesis/src/qrpgen/generators/flights.py
refactoring para mejorar qrpgen

@@ -1,6 +1,14 @@
 from qrp_structures import *
 
 def samecompany(argv):
+    companies = int(argv[0])
+    steps = int(argv[1])
+
+def specificcompany(argv):
+    companies = int(argv[0])
+    steps = int(argv[1])
+
+def samecompany(argv):
     if len(argv) != 2:
         #TODO
         sys.exit(1)
@@ -13,7 +21,7 @@ def samecompany(argv):
     # generate query
     company_var = Argument('var', steps)
     query_pred = "q"
-    query_args = [Argument('var', i) for i in range(1, steps+1)]
+    query_args = [Argument('var', i) for i in range(1, steps)]
     query_table = Table(query_pred, query_args)
     cityA, cityB = (Argument('const', 1), Argument('const', 2))
     query_sos = [Table(flights_pred, [cityA, Argument('var', 1), company_var])]

### 2009-11-29 tesis/src/qrpgen/qrpgen.py

@@ -3,6 +3,11 @@
 import sys
 import generators.flights
 
+def writelist(l, filename):
+    f = open(filename, 'w')
+    f.writelines([str(e) + '\n' for e in l])
+    f.close()
+
 def main():
     if len(sys.argv) < 4:
         usage_and_quit()
@@ -23,13 +28,8 @@ def main():
     
     query, views = generator(sys.argv[4:])
 
-    viewsfile = open(viewsfilename, 'w')
-    viewsfile.writelines([str(v) + '\n' for v in views])
-    viewsfile.close()
-
-    queryfile = open(queryfilename, 'w')
-    queryfile.write(str(query) + '\n')
-    queryfile.close()
+    writelist(views, viewsfilename)
+    writelist([query], queryfilename)
 
 def print_usage():
     print "Usage: %s <views_file> <query_file> <command> [command_arguments...]" % (sys.argv[0])

### 2009-11-29 mcdsat/mcdsat-1.0/models/nnf/nnf.cc
algunos cambios de prueba en mcdsat viejo para calcular mincost

@@ -5,6 +5,9 @@
 #include <unistd.h>
 #include <math.h>
 
+#define INF 0x7f
+#define MIN(a,b) ((a)<(b)?(a):(b))
+
 namespace nnf {
 
 size_t
@@ -318,6 +321,37 @@ Manager::compute_value_sorted_smooth( const Node *n, float *output, const float 
   return( result );
 }
 #endif
+
+int
+Manager::mincost_recursively( const Node *n, Model &m) const
+{
+  bool stat = true;
+  if( n->type_ == And ) {
+    int mm=INF;
+
+    for( NodePtr *p = n->children_; (*p != 0); ++p ) {
+      mm=MIN(mm,mincost_recursively(*p,m));
+    }
+
+    return mm;
+  }
+  else if( n->type_ == Or ) {
+    int sum=0;
+
+    for( NodePtr *p = n->children_; (*p != 0); ++p ) {
+      sum+=mincost_recursively(*p,m);
+    }
+
+    return sum;
+  }
+  else if( n->type_ == Variable ) {
+      //if n vista, devolver costo, else 0
+      return 0;
+  }
+  else if( n->type_ == Value ) {
+    return 0;
+  }
+}
 
 std::pair<bool,const Node*>
 Manager::enumerate_models_recursively( const Node *n, Model &m, const Node *last_or ) const

### 2009-11-29 mcdsat/mcdsat-1.0/models/nnf/nnf.h

@@ -444,6 +444,8 @@ namespace nnf {
 #ifdef MP
     MP_num* mp_count_models( MP_num *output = 0, const int *litmap = 0 ) const;
 #endif
+
+    int mincost_recursively( const Node *n, Model &m) const;
     std::pair<bool,const Node*> enumerate_models_recursively( const Node *n, Model &m, const Node *last_or ) const;
     void enumerate_models( std::ostream &os, bool all = false ) const;
     void enumerate_models( ModelList &models ) const;

### 2009-12-01 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
cambiado parcialmente Traductor3 para reducir t generadas

@@ -583,13 +583,47 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
     mappings = set()
     newmappings = set()
 
+    viewmappings = {}
+
     for v in vistas:
+        currentmappings = set()
+        newmappings = set()
+
         for soq in q.cuerpo:
             for sov in v.cuerpo:
                 if soq.predicado == sov.predicado:
                     for (x, y) in zip(soq.orden, sov.orden):
                         newmappings.add((x, y))
 
+        updated = True
+
+        while updated:
+            currentmappings.update(newmappings)
+            newmappings = set()
+            updated = False
+
+            for (a, y0) in currentmappings:
+                if es_var(a) or es_const(y0):
+                    continue
+
+                for (x1, y1) in currentmappings:
+                    if es_const(x1) or (y0 != y1) or (x1 == y1):
+                        continue
+
+                    for (x2, z) in currentmappings:
+                        if (x1 != x2) or es_const(z) or (x2 == z):
+                            continue
+
+                        if (a, z) not in currentmappings:
+                            updated = True
+                            print "SI"
+                            print "%s por culpa de que %s y %s y %s (en vista %s)" % (str((a,z)), str((a,y0)), str((x1,y1)), str((x2,z)), str(v))
+                            newmappings.add((a, z))
+
+
+    import sys
+    sys.exit(1)
+    #BBB
     foundnewmappings = True
 
     while foundnewmappings:
@@ -613,8 +647,9 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
                     if (a, z) not in mappings:
                         foundnewmappings = True
                         print "SI"
-                        print "%s por culpa de que %s y %s y %s" % (str((a,z)), str((a,y0)), str((x1,y1)), str((x2,z)))
+                        print "%s por culpa de que %s y %s y %s (en vista %s)" % (str((a,z)), str((a,y0)), str((x1,y1)), str((x2,z)), str(v))
                         newmappings.add((a, z))
+    #!BBB
 
     import sys
     sys.exit(1)

### 2009-12-02 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
cosas de traductor3 que no funcionan

@@ -578,7 +578,24 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
     global varsT
     global varsV
 
-    #ejele
+    lt_d4 = []
+
+    #return compatible mappings in a, b, c
+    #TODO more efficient implementation
+    def compatibles(a, b, c):
+        ret = set()
+
+        for (aq, av) in a:
+            for (bq, bv) in b:
+                for (cq, cv) in c:
+                    if (aq == bq and av == bv) or (aq != bq and av != bv):
+                        if (aq == cq and av == cv) or (aq != cq and av != cv):
+                            if (bq == cq and bv == cv) or (bq != cq and bv != cv):
+                                ret.add((aq, av))
+                                ret.add((bq, bv))
+                                ret.add((cq, cv))
+
+        return ret
 
     mappings = set()
     newmappings = set()
@@ -589,11 +606,21 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
         currentmappings = set()
         newmappings = set()
 
+        providers = {}
+        soqn = 0
+
         for soq in q.cuerpo:
+            sovn = 0
+
             for sov in v.cuerpo:
                 if soq.predicado == sov.predicado:
                     for (x, y) in zip(soq.orden, sov.orden):
                         newmappings.add((x, y))
+                        providers.setdefault((x, y), set()).add((soqn, sovn))
+
+                sovn += 1
+
+            soqn += 1
 
         updated = True
 
@@ -611,49 +638,29 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
                         continue
 
                     for (x2, z) in currentmappings:
-                        if (x1 != x2) or es_const(z) or (x2 == z):
+                        if (x1 != x2) or (x2 == z):
                             continue
 
+                        print "hey %s" % (str((a,z,)))
+                        print currentmappings
                         if (a, z) not in currentmappings:
-                            updated = True
-                            print "SI"
-                            print "%s por culpa de que %s y %s y %s (en vista %s)" % (str((a,z)), str((a,y0)), str((x1,y1)), str((x2,z)), str(v))
-                            newmappings.add((a, z))
+                            cs = compatibles(providers[(a,y0)], providers[(x1,y1)], providers[(x2,z)])
 
+                            print "probe para %s %s %s" % (str((a,y0)), str((x1,y1)), str((x2,z)))
[…]

### 2009-12-02 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
listas clausulas d4 con restricciones de generacion

@@ -641,12 +641,9 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
                         if (x1 != x2) or (x2 == z):
                             continue
 
-                        print "hey %s" % (str((a,z,)))
-                        print currentmappings
                         if (a, z) not in currentmappings:
                             cs = compatibles(providers[(a,y0)], providers[(x1,y1)], providers[(x2,z)])
 
-                            print "probe para %s %s %s" % (str((a,y0)), str((x1,y1)), str((x2,z)))
                             if len(cs) > 0:
                                 updated = True
                                 newmappings.add((a, z))
@@ -656,7 +653,6 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
                                 varT = varsT.get((a,z))
 
                                 if varT is None:
-                                    print "generando para %s, %s" % (str(a), str(z))
                                     varT = VariableSat(True, 't', [int(a), int(z)])
                                     varsT[(int(a), int(z))]=varT
                                     lt_d4.append(varT)
@@ -675,10 +671,8 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
                             if y == z:
                                 continue
 
-                            print "pat %s %s %s %s" % (a,x,y,z)
                             if es_const(y):
                                 continue
-                            print "pot %s %s %s %s" % (a,x,y,z)
 
                             d4set.add((varsV[numVista],a,y,x,z))
 
@@ -687,14 +681,12 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
     d4 = []
 
     for (vv,a,y,x,z) in d4set:
-        vtay = varsT.get((a,y))
-        vtxy = varsT.get((x,y))
-        vtxz = varsT.get((x,z))
-        vtaz = varsT.get((a,z))
+        vtay = varsT.get((int(a),int(y)))
+        vtxy = varsT.get((int(x),int(y)))
+        vtxz = varsT.get((int(x),int(z)))
+        vtaz = varsT.get((int(a),int(z)))
 
-        print vtaz
         if vtaz and vtay and vtxy and vtxz:
-            print "transit"
             d4.append([vv.negarVar(), vtay.negarVar(), vtxy.negarVar(), vtxz.negarVar(), vtaz])
 
     return d4, lt_d4
@@ -730,7 +722,7 @@ def clausulas_d5(q, vistas, variables_query, constantes_query):
         varT = varsT.get((i,j))
 
[…]

### 2009-12-02 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
reduccion de clausulas aplicada a las tipo d5 tambien

@@ -421,12 +421,11 @@ def clausulas_d1d2d4d5d6(q, vistas):
 
     d1 = clausulas_d1(variables_query, constantes_query, variables_vistas, constantes_vistas)
     d2 = clausulas_d2(variables_query, constantes_query, variables_vistas, constantes_vistas)
-    d4, lt_d4 = clausulas_d4(q, vistas, variables_query, constantes_query)
-    d5, lt_d5 = clausulas_d5(q, vistas, variables_query, constantes_query)
+    d4, d5, lt_d4d5 = clausulas_d4d5(q, vistas, variables_query, constantes_query)
 
     d6 = clausulas_d6(vistas, variables_query, constantes_vistas)
 
-    return d1, d2, d4, list(set(lt_d4 + lt_d5)), d5, d6
+    return d1, d2, d4, list(set(lt_d4d5)), d5, d6
 
 def clausulas_d1(variables_query, constantes_query, variables_vistas, constantes_vistas):
     global varsT
@@ -574,11 +573,11 @@ def clausula78a(varz, varg, varm, subObQ, subObV, vis, ltaux, c7, c8, vista):
     c8.append([varz.negarVar(), varm])
     return lt
 
-def clausulas_d4(q, vistas, variables_query, constantes_query):
+def clausulas_d4d5(q, vistas, variables_query, constantes_query):
     global varsT
     global varsV
 
-    lt_d4 = []
+    lt_d4d5 = []
 
     #return compatible mappings in a, b, c
     #TODO more efficient implementation
@@ -655,10 +654,11 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
                                 if varT is None:
                                     varT = VariableSat(True, 't', [int(a), int(z)])
                                     varsT[(int(a), int(z))]=varT
-                                    lt_d4.append(varT)
+                                    lt_d4d5.append(varT)
 
 
     d4set = set([])
+    d5set = set([])
 
     numVista = 0
 
@@ -678,25 +678,6 @@ def clausulas_d4(q, vistas, variables_query, constantes_query):
 
         numVista += 1
 
-    d4 = []
-
-    for (vv,a,y,x,z) in d4set:
-        vtay = varsT.get((int(a),int(y)))
-        vtxy = varsT.get((int(x),int(y)))
[…]

### 2009-12-02 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
soporte para reduccion de clausulas d5 (eso no estaba comiteado ya?)

@@ -628,6 +628,7 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query):
             newmappings = set()
             updated = False
 
+            # PARA D4
             for (a, y0) in currentmappings:
                 if es_var(a) or es_const(y0):
                     continue
@@ -654,6 +655,35 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query):
                                 if varT is None:
                                     varT = VariableSat(True, 't', [int(a), int(z)])
                                     varsT[(int(a), int(z))]=varT
+                                    lt_d4d5.append(varT)
+
+            # PARA D5
+            for (y0, a) in currentmappings:
+                if es_var(a) or es_const(y0):
+                    continue
+
+                for (y1, x1) in currentmappings:
+                    if es_const(x1) or (y0 != y1) or (x1 == y1):
+                        continue
+
+                    for (z, x2) in currentmappings:
+                        if (x1 != x2) or (x2 == z):
+                            continue
+
+                        if (z, a) not in currentmappings:
+                            cs = compatibles(providers[(y0,a)], providers[(y1,x1)], providers[(z,x2)])
+
+                            if len(cs) > 0:
+                                updated = True
+                                newmappings.add((z, a))
+                                providers[(z,a)] = cs
+
+                                #create the mapping var
+                                varT = varsT.get((z,a))
+
+                                if varT is None:
+                                    varT = VariableSat(True, 't', [int(z), int(a)])
+                                    varsT[(int(z), int(a))]=varT
                                     lt_d4d5.append(varT)
 
 
@@ -692,6 +722,7 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query):
                             if es_var(a) or es_const(x):
                                 continue
 
+                            #print "add %s %s %s %s" % (str(a), str(y), str(x), str(z))
                             d5set.add((varsV[numVista],a,y,x,z))
 
         numVista += 1
@@ -710,11 +741,13 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query):
     d5 = []
 
     for (vv,a,y,x,z) in d5set:
-        vtya = varsT.get((y,a))
-        vtyx = varsT.get((y,x))
-        vtzx = varsT.get((z,x))
-        vtza = varsT.get((z,a))
[…]

### 2009-12-02 tesis/src/imcdsat/mcdsat/CQ/CQ.py
nombres de variables diferentes para cada vista

@@ -2,6 +2,7 @@ from SubObjetivo import *
 from Seq import *
 
 CONSTANT_BASE = 1000000000
+VAR_BASE = 400000000
 
 def es_const(x):
     #hack para las reescrituras que colocan la x antes

### 2009-12-02 tesis/src/imcdsat/mcdsat/Traductor/GenerarReescrituras.py

@@ -97,6 +97,12 @@ def crearReescritura(mcdc, query):
     res = CQ(cab, cue, [])
     return res    
 
+def nombreReal(var):
+    if es_const(var):
+        return str(int(var)-CONSTANT_BASE)
+
+    #VARRENAME
+    return 'X'+str(int(var)%10000)
 
 def crearMCD(mod, vistas):
     phictodo = {}
@@ -110,7 +116,10 @@ def crearMCD(mod, vistas):
                 vista = vistas[int(numVista)]
         elif var[0] == 't':
             i = var.find(',')
-            phictodo.setdefault('X'+var[2:i],[]).append('X'+var[i+2:len(var)-1])
+            vq = nombreReal(var[2:i])
+            vv = nombreReal(var[i+2:len(var)-1])
+
+            phictodo.setdefault(vq,[]).append(vv)
         elif var[0] == 'g':
             gc.add(var[2:len(var)-1])
     return MCD(phictodo, gc, vista)

### 2009-12-05 tesis/experimentos/0/runner-grafico0.sh
directorio de experimentos

@@ -1,0 +1,18 @@
+#!/bin/bash
+
+OUTFILE=grafico0.out
+
+echo '########' | tee -a ${OUTFILE}
+
+for vistas in `seq 5 5 80`; do
+    for pasos in `seq 2 5`; do
+        INDIR=./casos-grafico0
+        VIEWSFILE=${INDIR}/views-${vistas}-${pasos}.txt
+        QUERYFILE=${INDIR}/query-${vistas}-${pasos}.txt
+
+        echo ${vistas} ${pasos} | tee -a ${OUTFILE}
+
+        /usr/bin/time -a -p -o ${OUTFILE} ./imcdsat/mcdsat/compileonly RW ${VIEWSFILE} ${QUERYFILE}
+
+    done
+done

### 2009-12-06 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
quite algunos flags extra que quedaban de Traductor3

@@ -103,26 +103,17 @@ def clausulasCombinarMCD(transf, n, q, vistas):
 
     for v in vistas:
         for so in v.cuerpo:
-            print "varsVs.update de %s en so %s" % (set(so.orden),so)
             varsVs.update(set(so.orden))
 
-    print "so?"
     for x in varsQ:
-        print "0"
         for a in varsVs:
-            print "1"
             for b in varsVs:
-                print "2 a=%s b=%s" % (a,b)
                 if es_const(a) and es_const(b) and (a!=b):
-                    print "3"
                     for numCopiaT0 in xrange(n):
-                        print "4"
                         for numCopiaT1 in xrange(n):
                             if numCopiaT0 != numCopiaT1:
                                 varT0 = VariableSat(False, 't', [int(x), int(a)])
                                 varT1 = VariableSat(False, 't', [int(x), int(b)])
-
-                                print "varT0: %s, varT1: %s" % (varT0, varT1)
 
                                 try:
                                     numVarT0 = transf.var2NumSimple(varT0, numCopiaT0)
@@ -679,8 +670,6 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query, ltaux, lz):
 
                             for (soqn, sovn) in providers[(a,z)]:
                                 varZ = varsZ.get((soqn,sovn,numVista))
-                                #print "agregue para %s : %s la varz %s" % (varT,varV,varZ)
-                                #print "antes estaba asi: %s" % (ltaux.get(varT,varV))
                                 ltaux.setdefault((varT,varV),set([])).add(varZ)    
 
             # PARA D5
@@ -720,8 +709,6 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query, ltaux, lz):
 
                             for (soqn, sovn) in providers[(z,a)]:
                                 varZ = varsZ.get((soqn,sovn,numVista))
-                                #print "agregue para %s : %s la varz %s" % (varT,varV,varZ)
-                                #print "antes estaba asi: %s" % (ltaux.get(varT,varV))
                                 ltaux.setdefault((varT,varV),set([])).add(varZ)    
 
     d4set = set([])
@@ -759,7 +746,6 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query, ltaux, lz):
                             if es_var(a) or es_const(x):
                                 continue
 
-                            #print "add %s %s %s %s" % (str(a), str(y), str(x), str(z))
[…]

### 2009-12-06 tesis/src/models/bestmodel.cc
fuente de models para ser editada

@@ -1,0 +1,173 @@
+#include <nnf.h>
+#include <satlib.h>
+
+#include <stdio.h>
+#include <string.h>
+#include <stdlib.h>
+#include <sys/resource.h>
+#include <iostream>
+#include <fstream>
+#include <functional>
+#include <set>
+
+int verbosity_level = 0;
+std::vector<int> fvars;
+
+void
+banner( std::ostream &os )
+{
+  //os << "Compute best model from NNF." << std::endl;
+}
+
+void
+usage( std::ostream &os )
+{
+  os << "usage: bestmodel [-f <file>] [-c <file>] <nnf>" << std::endl;
+}
+
+inline float
+read_time_in_seconds( void )
+{
+  struct rusage r_usage;
+  getrusage( RUSAGE_SELF, &r_usage );
+  return( (float)r_usage.ru_utime.tv_sec +
+	  (float)r_usage.ru_stime.tv_sec +
+	  (float)r_usage.ru_utime.tv_usec / (float)1000000 +
+	  (float)r_usage.ru_stime.tv_usec / (float)1000000 );
+}
+
+void
+read_simple_file( std::istream &is, std::vector<int> &contents )
+{
+  size_t n;
+  is >> n;
+  contents.clear();
+  contents.reserve(n);
+  for( size_t i = 0; i < n; ++i ) {
+    int entry;
+    is >> entry;
+    contents.push_back(entry);
+  }
+}
+
+void
+read_pairs_file( std::istream &is, std::vector<std::pair<int,int> > &contents )
+{
+  size_t n;
+  is >> n;
+  contents.clear();
+  contents.reserve(n);
+  for( size_t i = 0; i < n; ++i ) {
+    int p, q;
+    is >> p >> q;
+    contents.push_back(std::make_pair(p,q));
+  }
+}
+
+int
+main( int argc, const char **argv )
+{
+  float i_seconds, l_seconds, seconds;
+  const char *forget_file = 0, *cost_file = 0;
+  banner(std::cout);
+
+  // parse arguments
+  ++argv;
+  --argc;
+  if( argc == 0 ) {
+   print_usage:
+    usage(std::cout);
+    exit(-1);
+  }
+  for( ; argc && ((*argv)[0] == '-'); --argc, ++argv ) {
+    switch( (*argv)[1] ) {
+      case 'f':
+        forget_file = *++argv;
+        --argc;
+        break;
+      case 'c':
+        cost_file = *++argv;
+        --argc;
+        break;
+      case '?':
[…]

### 2009-12-06 tesis/src/models/models.cc

@@ -1,0 +1,162 @@
+/*
+ *  Copyright (C) 2006 Universidad Simon Bolivar
+ * 
+ *  Permission is hereby granted to distribute this software for
+ *  non-commercial research purposes, provided that this copyright
+ *  notice is included with any such distribution.
+ *  
+ *  THIS SOFTWARE IS PROVIDED "AS IS" WITHOUT WARRANTY OF ANY KIND,
+ *  EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE
+ *  IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+ *  PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
+ *  SOFTWARE IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU
+ *  ASSUME THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+ *  
+ *  Blai Bonet, bonet@ldc.usb.ve
+ *
+ */
+
+#include <nnf.h>
+#include <satlib.h>
+#include <utils.h>
+
+#include <iostream>
+#include <fstream>
+
+void
+banner( std::ostream &os )
+{
+  os << "Model enumerator for d-DNNF theories." << std::endl
+     << "Universidad Simon Bolivar, 2006." << std::endl;
+}
+
+void
+usage( std::ostream &os )
+{
+  os << "usage: models [--mp] [--literal-counts] [--num <n>] [--verbose] [--write-models] [--write-all-literals] <nnf>" << std::endl;
+}
+
+int
+main( int argc, char **argv )
+{
+  float i_seconds, l_seconds, seconds;
+  bool use_mp = false;
+  bool literal_counts = false;
+  bool all = false;
+  bool output = false;
+  bool verbose = false;
+  size_t count = 0;
+  banner(std::cout);
+
+  ++argv;
+  --argc;
+  if( argc == 0 ) {
+   print_usage:
+    usage(std::cout);
+    exit(-1);
+  }
+  for( ; argc && ((*argv)[0] == '-'); --argc, ++argv ) {
+    if( !strcmp(*argv,"--mp") ) {
+      use_mp = true;
+    }
+    else if( !strcmp(*argv,"--literal-counts") ) {
+      literal_counts = true;
+    }
[…]

### 2009-12-06 tesis/src/models/nnf/mp.h

@@ -1,0 +1,72 @@
+/*
+ *  Copyright (C) 2006 Universidad Simon Bolivar
+ * 
+ *  Permission is hereby granted to distribute this software for
+ *  non-commercial research purposes, provided that this copyright
+ *  notice is included with any such distribution.
+ *  
+ *  THIS SOFTWARE IS PROVIDED "AS IS" WITHOUT WARRANTY OF ANY KIND,
+ *  EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE
+ *  IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+ *  PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
+ *  SOFTWARE IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU
+ *  ASSUME THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+ *  
+ *  Blai Bonet, bonet@ldc.usb.ve
+ *
+ */
+
+#ifndef MP_INCLUDE
+#define MP_INCLUDE
+
+#ifdef MP
+
+#include <gmp.h>
+
+namespace mp {
+  class Int {
+    mpz_t z_;
+  public:
+    Int( int n = 0 ) { mpz_init_set_si(z_,n); }
+    Int( const Int &a ) { mpz_init_set(z_,a.z_); }
+    ~Int() { mpz_clear(z_); }
+    void operator=( const Int &a ) { mpz_set(z_,a.z_); }
+    void operator=( int n ) { mpz_set_si(z_,n); }
+    bool operator==( const Int &a ) { return( mpz_cmp(z_,a.z_) == 0 ); }
+    bool operator==( int n ) { return( mpz_cmp_si(z_,n) == 0 ); }
+    bool operator<( const Int &a ) { return( mpz_cmp(z_,a.z_) < 0 ); }
+    bool operator<( int n ) { return( mpz_cmp_si(z_,n) < 0 ); }
+    void operator+=( const Int &a ) { mpz_add(z_,z_,a.z_); }
+    void operator*=( const Int &a ) { mpz_mul(z_,z_,a.z_); }
+    void operator^=( unsigned n ) { mpz_pow_ui(z_,z_,n); }
[…]

### 2009-12-06 tesis/src/models/nnf/nnf.cc

@@ -1,0 +1,427 @@
+/*
+ *  Copyright (C) 2006 Universidad Simon Bolivar
+ * 
+ *  Permission is hereby granted to distribute this software for
+ *  non-commercial research purposes, provided that this copyright
+ *  notice is included with any such distribution.
+ *  
+ *  THIS SOFTWARE IS PROVIDED "AS IS" WITHOUT WARRANTY OF ANY KIND,
+ *  EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE
+ *  IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+ *  PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
+ *  SOFTWARE IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU
+ *  ASSUME THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+ *  
+ *  Blai Bonet, bonet@ldc.usb.ve
+ *
+ */
+
+#include "nnf.h"
+#include "satlib.h"
+#include <fstream>
+#include <stdio.h>
+#include <unistd.h>
+#include <math.h>
+
+namespace nnf {
+
+size_t
+Manager::count_edges() const
+{
+  for( size_t i = 0, isz = node_pools_.size(); i < isz; ++i ) {
+    NodePool *pool = node_pools_[i];
+    for( size_t j = 0, jsz = pool->boundary_; j < jsz; ++j ) {
+      const Node *n = &pool->table_[j];
+      assert( n->cache_.first_ == 0 );
+      if( (n->type_ == And) || (n->type_ == Or) ) {
+        size_t count = 0;
+        for( const NodePtr *p = n->children_; *p != 0; ++p ) {
+          count += 1 + (size_t)(*p)->cache_.first_;
+          (*p)->cache_.set();
+        }
+        n->cache_.set( (const void*)count );
+      }
+    }
+  }
+  size_t count = (size_t)root_->cache_.first_;
+  clean_node_cache();
+  return(count);
+}
+
+void
+Manager::sort()
+{
+  if( sorted_ ) return;
+  char tname[20] = "/tmp/nnf.XXXXXX";
+  char *fname = mktemp(tname);
+
+  if( verbose_ ) std::cout << "nnf: sorting: dumping" << std::flush;
[…]

### 2009-12-06 tesis/src/models/nnf/nnf.h

@@ -1,0 +1,623 @@
+/*
+ *  Copyright (C) 2006 Universidad Simon Bolivar
+ * 
+ *  Permission is hereby granted to distribute this software for
+ *  non-commercial research purposes, provided that this copyright
+ *  notice is included with any such distribution.
+ *  
+ *  THIS SOFTWARE IS PROVIDED "AS IS" WITHOUT WARRANTY OF ANY KIND,
+ *  EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE
+ *  IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+ *  PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
+ *  SOFTWARE IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU
+ *  ASSUME THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+ *  
+ *  Blai Bonet, bonet@ldc.usb.ve
+ *
+ */
+
+#ifndef NNF_INCLUDE
+#define NNF_INCLUDE
+
+#include <iostream>
+#include <iomanip>
+#include <ext/hash_map>
+#include <list>
+#include <set>
+#include <vector>
+#include <stdio.h>
+#include <stdlib.h>
+#include <assert.h>
+
+#include "mp.h"
+#include "utils.h"
+
+namespace nnf {
+
+  template<typename T> struct Pool {
+    T *table_;
+    size_t size_;
+    size_t used_;
+    size_t boundary_;
+    std::list<std::pair<size_t,T*> > free_list_;
+    size_t size_free_list_;
+    Pool( size_t size ) : size_(size), used_(0), boundary_(0), size_free_list_(0) { table_ = new T[size]; }
+    ~Pool() { delete[] table_; }
+    bool verify() const { return( (boundary_ <= size_) && (used_ + size_free_list_ == boundary_) ); }
+    bool from_here( const T *e ) const { return( (table_ <= e) && (e < &table_[size_]) ); }
+    T* allocate( size_t n = 1 ) { if( n+boundary_ <= size_ ) { T *e = &table_[boundary_]; boundary_ += n; used_ += n; return(e); } return(0); }
+    void clear() { used_ = 0; boundary_ = 0; free_list_.clear(); size_free_list_ = 0; }
[…]

### 2009-12-06 tesis/src/models/nnf/satlib.cc

@@ -1,0 +1,148 @@
+/*
+ *  Copyright (C) 2006 Universidad Simon Bolivar
+ * 
+ *  Permission is hereby granted to distribute this software for
+ *  non-commercial research purposes, provided that this copyright
+ *  notice is included with any such distribution.
+ *  
+ *  THIS SOFTWARE IS PROVIDED "AS IS" WITHOUT WARRANTY OF ANY KIND,
+ *  EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE
+ *  IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+ *  PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
+ *  SOFTWARE IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU
+ *  ASSUME THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+ *  
+ *  Blai Bonet, bonet@ldc.usb.ve
+ *
+ */
+
+#include "nnf.h"
+#include "satlib.h"
+
+#include <iostream>
+#include <ios>
+#include <set>
+#include <vector>
+#include <stdio.h>
+#include <ctype.h>
+#include <fstream>
+#include <string>
+
+namespace satlib {
+
+bool
+read_line( std::istream &is, std::string &line )
+{
+  char c;
+  line.clear();
+  while( is.get(c) ) {
+    if( c == '\n' ) return(true);
+    line += c;
+  }
+  return(false);
+}
+
+void
+read_nnf_file( std::istream &is, nnf::Manager &manager )
+{
+  size_t n = 0;
+  int num_vars, num_nodes, num_edges;
+  std::string line;
+  const nnf::Node **nodes = 0;
+
+  // first pass: determine parents' sizes
+  size_t *psize = 0;
+  while( read_line(is,line) ) {
+    if( line[0] == 'c' )
+      continue;
+    else if( line[0] == 'n' ) {
+      sscanf( line.c_str(), "nnf %d %d %d", &num_nodes, &num_edges, &num_vars );
+      psize = new size_t[num_nodes];
+      for( int i = 0; i < num_nodes; ++i ) psize[i] = 0;
+    }
+    else if( line[0] == '%' )
+      break;
+    else {
+      char *str = strdup(line.c_str());
[…]

### 2009-12-06 tesis/src/models/models.cc
fixes para que compile en 64 bits

@@ -20,6 +20,8 @@
 #include <satlib.h>
 #include <utils.h>
 
+#include <string.h>
+
 #include <iostream>
 #include <fstream>
 

### 2009-12-06 tesis/src/models/nnf/nnf.cc

@@ -92,9 +92,9 @@ Manager::count_models( float *output, const int *lit_map ) const
           }
         }
         else if( n->type_ == Value )
-          count = ((int)n->children_?1:0);
+          count = ((long)n->children_?1:0);
         else if( n->type_ == Variable ) {
-          if( lit_map && lit_map[(int)n->children_^1] )
+          if( lit_map && lit_map[(long)n->children_^1] )
             count = 0;
           else
             count = 1;
@@ -131,10 +131,10 @@ Manager::count_models( float *output, const int *lit_map ) const
         }
         n->cache_.second_ = *(const float**)&pd;
         if( n->type_ == Variable ) {
-          if( ((int)n->children_%2 == 0) && (!lit_map || !lit_map[(int)n->children_^1]) )
-            output[(int)n->children_>>1] = pd;
-          else if( output[(int)n->children_>>1] == -1 )
-            output[(int)n->children_>>1] = 0;
+          if( ((long)n->children_%2 == 0) && (!lit_map || !lit_map[(long)n->children_^1]) )
+            output[(long)n->children_>>1] = pd;
+          else if( output[(long)n->children_>>1] == -1 )
+            output[(long)n->children_>>1] = 0;
         }
       }
     }
@@ -167,9 +167,9 @@ Manager::mp_count_models( mp::Int **output, const int *lit_map ) const
           }
         }
         else if( n->type_ == Value )
-          *count = ((int)n->children_?1:0);
+          *count = ((long)n->children_?1:0);
         else if( n->type_ == Variable ) {
-          if( lit_map && lit_map[(int)n->children_^1] )
+          if( lit_map && lit_map[(long)n->children_^1] )
             *count = 0;
           else
             *count = 1;
@@ -206,10 +206,10 @@ Manager::mp_count_models( mp::Int **output, const int *lit_map ) const
         }
         n->cache_.second_ = (const void*)pd;
         if( n->type_ == Variable ) {
-          if( ((int)n->children_%2 == 0) && (!lit_map || !lit_map[(int)n->children_^1]) )
-            output[(int)n->children_>>1] = new mp::Int(*pd);
-          else if( output[(int)n->children_>>1] == 0 )
-            output[(int)n->children_>>1] = new mp::Int;
[…]

### 2009-12-06 tesis/src/models/nnf/nnf.h

@@ -28,6 +28,7 @@
 #include <stdio.h>
 #include <stdlib.h>
 #include <assert.h>
+#include <limits.h>
 
 #include "mp.h"
 #include "utils.h"
@@ -104,9 +105,9 @@ namespace nnf {
     {
       os << "node: id=" << this << ", type=" << type_ << std::flush;
       if( type_ == Value )
-        os << ", value=" << (int)children_;
+        os << ", value=" << (long)children_;
       else if( type_ == Variable )
-        os << ", literal=" << (int)children_;
+        os << ", literal=" << (long)children_;
       else if( (type_ == And) || (type_ == Or) ) {
         os << ", children={";
         if( children_ ) for( const NodePtr *p = children_; *p != 0; ++p ) os << *p << ",";
@@ -135,8 +136,8 @@ namespace nnf {
     size_t hits() const { return(hits_); }
     float hit_rate() const { return((float)hits_/(float)lookups_); }
     const Node* lookup( int lit ) const { ++lookups_; const_iterator hi = find(lit); if( hi == end() ) return(0); ++hits_; return((*hi).second); }
-    void insert( const Node *n ) { assert( n->type_ == Variable ); basetype::insert( std::make_pair((int)n->children_,n) ); }
-    void remove( const Node *n ) { iterator hi = find((int)n->children_); erase(hi); }
+    void insert( const Node *n ) { assert( n->type_ == Variable ); basetype::insert( std::make_pair((long)n->children_,n) ); }
+    void remove( const Node *n ) { iterator hi = find((long)n->children_); erase(hi); }
     void clear() { basetype::clear(); lookups_ = 0; hits_ = 0; }
   };
 
@@ -459,8 +460,8 @@ namespace nnf {
 
     const Node* true_node() { if( !true_node_ ) true_node_ = get_node(Value,true); return(true_node_); }
     const Node* false_node() { if( !false_node_ ) false_node_ = get_node(Value,false); return(false_node_); }
-    bool is_constant_false( const Node* n ) const { return( (n->type_ == Value) && ((int)n->children_ == 0) ); }
-    bool is_constant_true( const Node* n ) const { return( (n->type_ == Value) && ((int)n->children_ == 1) ); }
+    bool is_constant_false( const Node* n ) const { return( (n->type_ == Value) && ((long)n->children_ == 0) ); }
+    bool is_constant_true( const Node* n ) const { return( (n->type_ == Value) && ((long)n->children_ == 1) ); }
 
     size_t count_nodes() const
     {

### 2009-12-06 tesis/src/models/nnf/satlib.cc

@@ -27,6 +27,7 @@
 #include <ctype.h>
 #include <fstream>
 #include <string>
+#include <string.h>
 
 namespace satlib {
 

### 2009-12-07 ws2/tarea2.tex
cosas de debugging de squinlib

@@
 
+Si queremos encontrar los medicamentos que pueden atacar una enfermedad
+específica podemos utilizar la consulta que se ve en la figura
+\ref{src:DrugForDiseaseQuery}. Allí simplemente se obtienen todos las drogas que
+atacan la enfermedad según la propiedad {\bf drugbank:possibleDiseaseTarget} y
+sus etiquetas.
+
+\begin{SaveVerbatim}{VerbDrugForDiseaseQuery}
+PREFIX rdf:  <http://www.w3.org/1999/02/22-rdf-syntax-ns#>
+PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
+PREFIX owl:  <http://www.w3.org/2002/07/owl#>
+PREFIX drugbank: <http://www4.wiwiss.fu-berlin.de/drugbank/resource/drugbank/>
+PREFIX tcm:      <http://purl.org/net/tcm/tcm.lifescience.ntu.edu.tw/>
+SELECT DISTINCT ?drug ?drugLabel WHERE {
+  ?drug drugbank:possibleDiseaseTarget <URI>.
+  ?drug rdfs:label ?drugLabel.
+}
+ORDER BY ?drugLabel
+\end{SaveVerbatim}
+
+\begin{figure}[htbp]
+ \fbox
+ {
+  \begin{minipage}{\textwidth}
+   \BUseVerbatim{VerbDrugForDiseaseQuery}
+   \caption{Consulta de búsqueda de medicamentos para enfermedades}
+   \label{src:DrugForDiseaseQuery}
+  \end{minipage}
+ }
+\end{figure}
+
 \end{document}

### 2009-12-07 ws2/src/SquinLib/src/ve/usb/squinlib/Main.java

@@ -13,7 +13,8 @@ import java.util.logging.Logger;
 public class Main {
 
     public static void main(String[] args) {
-        SparqlQuerySock sqs = new SparqlQuerySock("http://localhost:21786/SQUIN/query");
+        //SparqlQuerySock sqs = new SparqlQuerySock("http://localhost:21786/SQUIN/query");
+        SparqlQuerySock sqs = new SparqlQuerySock("http://ijaz.ldc.usb.ve:8080/SQUIN/query");
 
         BufferedReader in = new BufferedReader(new InputStreamReader(System.in));
         StringBuffer sb = new StringBuffer();

### 2009-12-07 ws2/src/SquinLib/src/ve/usb/squinlib/SparqlQuerySock.java

@@ -26,17 +26,17 @@ public class SparqlQuerySock {
     }
 
     public String call(final String query) throws MalformedURLException, IOException {
-        logger.log(Level.INFO, "SparqlQuerySock call");
+        logger.log(Level.INFO, "SparqlQuerySock.call(...)");
 
         Map<String, String> request_params = new HashMap<String, String>();
         request_params.put("query", query);
 
-        logger.log(Level.INFO, "query server: " + queryServer);
-        logger.log(Level.INFO, "query: " + query);
+        logger.log(Level.FINE, "query server: " + queryServer);
+        logger.log(Level.FINE, "query: " + query);
 
         String requestUrl = createQueryURLString(this.queryServer, request_params);
 
-        logger.log(Level.INFO, "request url: " + requestUrl);
+        logger.log(Level.FINE, "request url: " + requestUrl);
 
         URL url = new URL(requestUrl);
 

### 2009-12-10 tesis/src/imcdsat/mcdsat/Traductor/GenerarReescrituras.py
soporte para el nuevo formato de models listo

@@ -146,10 +146,26 @@ def generarReescRW(numeros, stdin, lenQuery, query, vistas):
     infile = stdin
     modelos = []
     n = len(numeros)
+
+    in_models = False
+
     for x in infile:
+        if not in_models:
+            if x.startswith('--- models begin ---'):
+                in_models = True
+
+            continue
+
+        if in_models:
+            if x.startswith('---- models end ----'):
+                break
+
         l = x.strip().split()
-        if l[0] != 'main:':
-            print crearReescritura(obtModeloRW(numeros, n, l, lenQuery, vistas), query)
+
+        model = l[1:len(l)-1]
+
+        print crearReescritura(obtModeloRW(numeros, n, model, lenQuery, vistas), query)
+
     return modelos
 
 def generarMejorReescRW(numeros, stdin, lenQuery, query, vistas, archCostos):
@@ -158,9 +174,14 @@ def generarMejorReescRW(numeros, stdin, lenQuery, query, vistas, archCostos):
     n = len(numeros)
     for x in infile:
         l = x.strip().split()
-        print l
-        continue
-        print crearReescritura(obtModeloRW(numeros, n, l, lenQuery, vistas), query)
+
+        if l[0] == 'main:':
+            continue
+
+        cost = int(l[0])
+        model = l[2:len(l)-1]
+
+        print (cost, crearReescritura(obtModeloRW(numeros, n, model, lenQuery, vistas), query))
         break
 
 def obtModeloRW(numeros, n, lista, lenQuery, vistas):
@@ -186,16 +207,21 @@ def generarReescMCD(numeros, stdin, vistas):
     infile = stdin
     modelos = []
     count = 0
-    x = infile.readline()
-    x = infile.readline()
+    for i in range(1,6):
+        x = infile.readline()
     while 1:
         if not(x):
             break
         x = infile.readline()
+
+        if x.startswith('---- models end ----'):
+            break
+
         l = x.strip().split()
-        n = len(l)
+        model = l[1:len(l)-1]
+        n = len(model)
         if n > 1: 
-            mod = obtModelo(numeros, l, n)
+            mod = obtModelo(numeros, model, n)
             mcd = crearMCD(mod, vistas)
             if mcd != None:
                 #print mod

### 2009-12-10 tesis/src/qrpgen/generators/flights.py
generacion de multiples vistas por aerolinea

@@ -1,22 +1,23 @@
 from qrp_structures import *
 
-def samecompany(argv):
-    companies = int(argv[0])
+def sameairline(argv):
+    airlines = int(argv[0])
     steps = int(argv[1])
+    return nsteps(airlines, steps, 'sameairline')
 
-def specificcompany(argv):
-    companies = int(argv[0])
+def specificairline(argv):
+    airlines = int(argv[0])
     steps = int(argv[1])
+    return nsteps(airlines, steps, 'specificairline')
 
-def samecompany(argv):
-    if len(argv) != 2:
-        #TODO
-        sys.exit(1)
-
-    flights_pred = "r0"
-
-    companies = int(argv[0])
+def sameairline_manyviews(argv):
+    airlines = int(argv[0])
     steps = int(argv[1])
+    views_per_airline = int(argv[2])
+    return nsteps(airlines, steps, views_per_airline)
+
+def nsteps(airlines, steps, views_per_airline=1):
+    flights_pred = "r1"
 
     # generate query
     company_var = Argument('var', steps)
@@ -36,17 +37,21 @@ def samecompany(argv):
     # generate views
     view_pred_prefix = 'v'
 
-    def gen_view(i):
+    def gen_view(num, a):
         varS = Argument('var', 1)
         varT = Argument('var', 2)
-        constCompany = Argument('const', i)
+        constAirline = Argument('const', a)
 
-        view_table = Table(view_pred_prefix + str(i), [varS, varT])
-        view_sos = [Table(flights_pred, [varS, varT, constCompany])]
+        view_table = Table(view_pred_prefix + str(num), [varS, varT])
+        view_sos = [Table(flights_pred, [varS, varT, constAirline])]
         view = Query(view_table, view_sos)
 
         return view
 
-    views = [gen_view(i) for i in range(100, 100+companies)]
+    views_base = 100
+    views = []
+
+    for n in range(views_per_airline):
+        views += [gen_view(n*airlines+i, i) for i in range(views_base, views_base+airlines)]
 
     return query, views

### 2009-12-10 tesis/src/qrpgen/qrpgen.py

@@ -13,7 +13,8 @@ def main():
         usage_and_quit()
 
     commands = {}
-    commands['samecompany'] = generators.flights.samecompany
+    commands['sameairline'] = generators.flights.sameairline
+    commands['sameairline_manyviews'] = generators.flights.sameairline_manyviews
 
     viewsfilename = sys.argv[1]
     queryfilename = sys.argv[2]

### 2009-12-11 tesis/src/qrpgen/generators/flights.py
formato de ayuda mas bonito para qrpgen

@@ -14,9 +14,9 @@ def sameairline_manyviews(argv):
     airlines = int(argv[0])
     steps = int(argv[1])
     views_per_airline = int(argv[2])
-    return nsteps(airlines, steps, views_per_airline)
+    return nsteps(airlines, steps, 'sameairline', views_per_airline)
 
-def nsteps(airlines, steps, views_per_airline=1):
+def nsteps(airlines, steps, type, views_per_airline=1):
     flights_pred = "r1"
 
     # generate query

### 2009-12-11 tesis/src/qrpgen/generators/structured.py

@@ -1,0 +1,5 @@
+from qrp_structures import *
+
+def chainquery(argv):
+    nviews = int(sys.argv[0])
+    

### 2009-12-11 tesis/src/qrpgen/qrpgen.py

@@ -2,6 +2,12 @@
 
 import sys
 import generators.flights
+import generators.structured
+
+commands = {}
+commands['sameairline'] = generators.flights.sameairline
+commands['sameairline_manyviews'] = generators.flights.sameairline_manyviews
+commands['chainquery'] = generators.structured.chainquery
 
 def writelist(l, filename):
     f = open(filename, 'w')
@@ -11,10 +17,6 @@ def writelist(l, filename):
 def main():
     if len(sys.argv) < 4:
         usage_and_quit()
-
-    commands = {}
-    commands['sameairline'] = generators.flights.sameairline
-    commands['sameairline_manyviews'] = generators.flights.sameairline_manyviews
 
     viewsfilename = sys.argv[1]
     queryfilename = sys.argv[2]
@@ -34,6 +36,11 @@ def main():
 
 def print_usage():
     print "Usage: %s <views_file> <query_file> <command> [command_arguments...]" % (sys.argv[0])
+    print
+    print "Available commands:"
+
+    for c in commands.keys():
+        print "  %s" % (c)
 
 if __name__ == "__main__":
     main()

### 2009-12-11 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
pase de archivo de costos para bestmodel

@@ -24,7 +24,7 @@ varsT = {}
 varsZ = {}
 contribVT = {}
 
-def traducir(exp, archV, archC, archVars, archTiempo, archSalida):
+def traducir(exp, archV, archC, archVars, archTiempo, archSalida, archCostosSat):
     #tiempo = timeit.Timer('traducir1()', "from __main__ import traducir1; import psyco; psyco.full()").timeit(1)/1
     tiempoi = resource.getrusage(resource.RUSAGE_SELF)[0]
     traducir1(exp, archV, archC, archVars, archSalida)

### 2009-12-11 tesis/experimentos/3/runner.sh
experimento 3

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+echo '########' | tee -a ${OUTFILE}
+
+INDIR=casos
+OUTFILE=datos-grafico3.out
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/mcdsat
+
+for vistas in `seq 5 5 40`; do
+    for pasos in `seq 2 5`; do
+        for copias in 2; do
+            VIEWSFILE=${INDIR}/views-${vistas}-${pasos}-${copias}.txt
+            QUERYFILE=${INDIR}/query-${vistas}-${pasos}-${copias}.txt
+
+            echo ${vistas} ${pasos} ${copias} | tee -a ${OUTFILE}
+
+            for i in `seq 1`; do #n times to get average
+                ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+            done
+        done
+    done
+done

### 2009-12-11 tesis/src/imcdsat/mcdsat/Main.py
fix para que no haya problemas en casos sin costos

@@ -29,11 +29,15 @@ if __name__ == "__main__":
 
         if len(sys.argv) > 8:
             archCostosSat = sys.argv[8]
+        else:
+            archCostosSat = None
 
         traducir(exp, archVistas, archCons, archVars, archTiempo, archSalida, archCostosSat)
     elif op == 'G':
         if len(sys.argv) > 7:
             archCostos = sys.argv[7]
+        else:
+            archCostos = None
 
         generarReescrituras(exp, archVistas, archCons, archVars, archTiempo, archCostos, sys.stdin)
 

### 2009-12-11 tesis/src/qrpgen/generators/structured.py
generador aleatorio 0

@@ -1,5 +1,44 @@
+import random
+
 from qrp_structures import *
 
 def chainquery(argv):
-    nviews = int(sys.argv[0])
-    
+    #TODO
+    return [], []
+
+def randomviews(argv):
+    nviews = int(argv[0])
+    nview_sos = int(argv[1])
+
+    nquery_sos = int(argv[2])
+
+    n_preds = int(argv[3])
+    n_vars = int(argv[4])
+
+    n_varsperpred = int(argv[5])
+
+    perc_constants = float(argv[6])
+
+    def rand_constvar(i):
+        r = random.random()
+
+        if r < perc_constants:
+            return Argument('const', i)
+
+        return Argument('var', i)
+
+    predicates = ["r%d" % i for i in xrange(1, n_preds+1)]
+
+    query_table = Table("q", [Argument('var', i+1) for i in xrange(n_vars)])
+    query_sos = [Table(pred, [rand_constvar(i+1) for i in random.sample(range(n_vars)*n_vars, n_varsperpred)]) for pred in random.sample(predicates * n_preds, nquery_sos)]
+    query = Query(query_table, query_sos)
+
+    views = []
+
+    for n in xrange(1, nviews+1):
+        view_table = Table("v%s" % n, [Argument('var', i+1) for i in xrange(n_vars)])
+        view_sos = [Table(pred, [rand_constvar(i+1) for i in random.sample(range(n_vars)*n_vars, n_varsperpred)]) for pred in random.sample(predicates * n_preds, nquery_sos)]
+        view = Query(view_table, view_sos)
+        views.append(view)
+
+    return query, views

### 2009-12-11 tesis/src/qrpgen/qrpgen.py

@@ -5,9 +5,10 @@ import generators.flights
 import generators.structured
 
 commands = {}
-commands['sameairline'] = generators.flights.sameairline
-commands['sameairline_manyviews'] = generators.flights.sameairline_manyviews
-commands['chainquery'] = generators.structured.chainquery
+commands['flights.sameairline'] = generators.flights.sameairline
+commands['flights.sameairline_manyviews'] = generators.flights.sameairline_manyviews
+commands['structured.chainquery'] = generators.structured.chainquery
+commands['randomviews'] = generators.structured.randomviews
 
 def writelist(l, filename):
     f = open(filename, 'w')

### 2009-12-11 tesis/propuesta/propuesta_iae.tex
propuesta_iae.tex

  \documentclass{llncs}
\usepackage{amsmath}
\usepackage{txfonts}
\usepackage[boxed]{algorithm2e}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage{enumerate}
\DeclareGraphicsExtensions{.jpg, .png, .mps, .bmp, .pdf}
\usepackage{epsfig}
\usepackage{subfigure}
\newtheorem{Definition}{Definition} 
\newtheorem{Example}{Example} 
\newtheorem{Theorem}{Theorem} 
\newtheorem{Proof}{Proof} 
\newcommand{\denselist}{\topsep 0pt \itemsep -4pt}
\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\vvec}[1]{\mathbf{#1}}
\newcommand{\join}{\bowtie}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\body}{{body}}
\newcommand{\head}{{head}}
\newcommand{\qrule}{:\!\!-}
\newcommand{\arc}{\text{arc}}
\newcommand{\Theory}[1]{T(#1)}
\newcommand{\mcdsat}{\textsc{McdSat}}
\newcommand{\minicon}{{MiniCon}}
\newcommand{\Omit}[1]{}
\newcommand{\citeX}[1]{\citeauthor{#1}~\citeyear{#1}}

\begin{document}
\allowdisplaybreaks
\title{Selecting the Maximum-Utility  Web Services  Using Circuits}
\author{Daniel Izquierdo \inst{1}  and Blai Bonet \inst{1}  and Mar\'{\i}a-Esther Vidal \inst{1}}

\institute{Universidad Sim\'on Bol\'{\i}var\\
                Caracas, Venezuela\\
                \{idaniel,bonet,mvidal\}ldc.usb.ve} \maketitle
\begin{abstract}
In Service Oriented Architectures (SOA), concrete services are described in terms of functionality   and QoS parameters, while software applications are specified as workflows of abstract services and non-functional criteria that establish the  application desirable functionality and  expected quality, respectively.  During workflow execution time,  abstract services  are rewritten 
with the concrete services  that meet the workflow  functional and non-functional restrictions. Since, service selection is done on-the-flight and the space of concrete services may be very large, efficiency and scalability are required.  In this paper we propose the {\it McdSat}$^c$ framework to efficiently solve the problem of service selection in presence of a large number of concrete services.
First, we use the Local As View approach to describe the functionality  of the concrete services in terms of views of abstract services, the quality as an overall utility function that combines the different QoS values that describe the service, and  workflows as  conjunctive queries on abstract services. Based on this representation, we 
model the problem of service selection as a known problem in the area of integration systems referred as query rewriting using views.  We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and maximizes the overall utility function; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.  We report on the performance of our proposed approach, and we observe that  {\it McdSat}$^c$  efficiently scales up to large instances of the service selection problem.

\end{abstract}                
\section{Introduction}
Under the umbrella of the Semantic Web and supported by Service Oriented architectures, the number of Web data sources and services has exploded during the last few years.  For example,  currently, the molecular biology databases
collection includes 1,170 databases~\cite{Galperin09}, that is
95 more than the previous year~\cite{Galperin2008} and 110 more than two years ago~\cite{Galperin2007};  tools
and services as well as the number of instances published by these resources, follow a similar progression~\cite{Benson07}. Thanks
to this wealth, users rely more on various
digital tasks such as data retrieval from public data sources and
data analysis with Web tools or services organized in
complex workflows.

In order to design an application, users usually specify the functionality in terms of a workflow of abstract services, while the desirable quality  is defined by restricting the values of the QoS parameters that describe the concrete available services. To evaluate the proposed workflow,  abstract services are instantiated on concrete services in a way restrictions on the QoS parameters are satisfied.  However, the performance of this tasks is impacted by the number of available sources,  techniques enable to efficiently scale up to a large number of services are required.

In this paper we aim to provide a solution to this problem and provide techniques that efficiently traverse the space of available services and select combinations that meet the functional and non-functional requirements that describe a designed application.  We name our proposed solution {\it MCDSat}$^c$.

In {\it MCDSat}$^c$, concrete service functionalities are decribed as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.  In addition, rules that define workflows and  concrete services in the {\it MCDSat}$^c$ catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied. To represent the QoS measures, each concrete service description is annotated with a real number that represents an overall value of the QoS measures of this service.  Finally, {\it MCDSat}$^c$ models the web service selection problem as the problem of query rewriting using views \cite{halevy:survey}. This problem is important in the context of data integration
\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance \cite{AfratiLU07,levy:bucket}, and several approaches have been defined to efficiently scale up to a large number of views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}. In this paper we propose the {\it MCDSat}$^c$ framework which extends the approach proposed in ~\cite{arvelo:aaai06},  with the capability of dealing with constants and identifying the composition of services that maximizes a given utility function.

{\it MCDSat}$^c$  translates an instance of the service selection problem into the problem of enumerating the models of a propositional theory that encodes information about the workflow, the concrete services and the conditions that need to be satisfied in order to generate a composition that maximizes the utility function. This theory is represented as a CNF formula which is compiled into a deterministic
and decomposable negation normal form d-DNNF representation, from which the models that maximizes the overall utility function are obtained in linear time using modern SAT techniques~\cite{darwiche:dnnf}.   
We empirically studied the quality of the {\it MCDSat}$^c$ on variety of benchmarks and we observed  that the proposed technique provides an efficient solution to the service selection problem which is able to scale up to large workflows as well to a large number of views.

 This paper is comprised of six additional sections. The next section motivates our approach by using an example, and  we summarize existing related work in section 3. Section 4 describes the {\it MCDSat}$^c$ system architecture, and  the experimental study is reported in section 5. 
 In section 6 we present the formalization of the web service selection problem as a Propositional Logic theory, and finally, section 7 outlines our conclusions and future work.

\section{Motivating Example}
Consider the abstract and concrete services in Table ~\ref{table:services}  which represent applications in the flight domain.

\begin{table}
\begin{tabular}{| l| l |}
\hline
Abstract Service & Description \\
\hline
$fight(x1,x2)$ & relates cities x1 and x2 if there is a flight between them\\
\hline
$uscity(x1)$ &  determines if a city x1 is a US city\\
\hline \hline
Concrete  Service & Description \\
\hline
$national(x1,y1)$&  returns  two US cities that are connected by  direct flights \\
\hline
$oneway(x1,y1)$ & returns two cites that are connected by one way direct flights \\
\hline
$onestop(x1,y1)$ &  returns  two cites that are connected by one-stop flights \\
\hline
$ flightToCCS(x1,"CCS")$ & returns a city that is connected to  Caracas by a direct flight\\
 \hline
$ onestopToCCS(x1,y1)$ & returns two cites x1 and y1, such that there is  one-stop flight\\ 
 & from city x1 to Caracas stopping at city y1.\\
 \hline
 $fromWAS("WAS",y1)$ &  returns a city y1 to which flights from Washington arrive. \\
 \hline
\end{tabular}
\label{table:services}
\caption{Abstract and Concrete Services}
\end{table}

Using the  Local As View (LAV) approach~\cite{AmbiteISWC09}, concrete services in Table 1 are semantically described as views on the abstract services {\it flight} and {\it uscity} as follows:

\begin{alignat*}{1}
national(x1,y1)\ &\qrule\ flight(x1,y1),\,  uscity(x1),\,  uscity(y1)\,. \\
oneway(x2,y2)\ &\qrule\ flight(x2,y2)\,. \\
onestop(x3,z3)\ &\qrule\ flight(x3,y3),\, flight(y3,z3)\,. \\
flightToCCS(x2,"CCS")\ &\qrule\ flight(x2,"CCS")\,. \\
onestopToCCS(x3,y3)\ &\qrule\ flight(x3,y3),\, flight(y3,"CCS")\,. \\
fromWAS("WAS",y3)\ &\qrule\ flight("WAS",x3)\,. \\
\end{alignat*}

Now suppose a user is interested in implementing a workflow able to retrieve the one-stop round trip flights from US cities to any city in the world, such that flights can stop at any city. 
Consider the following conjunctive query that represents the workflow that defines this request in terms of abstract services. This workflow was defined in way the input parameters of each abstract service are bound by  constants in the workflow, or by attributes projected out by  previous abstract services.

\begin{alignat*}{1}
w(x,y1,y2,y3)\ &\qrule\ flight(x,y1),\, flight(y1,y2),\,  flight(y2,y3),\,  flight(y3,x),\,  \\
                   \ & uscity(x)\,. \\
\end{alignat*}

Implementations of this workflow correspond to compositions of the concrete services where each concrete service can implement a subset of abstract services of the workflow, but each abstract service can be implemented by exactly one concrete service. The following composition of concrete services corresponds to one of the XX implementations of the workflow:

\begin{alignat*}{1}
wi1(x,y1,"CCS",y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\,  oneway("CCS",y3),\,  \\
\ & national(y3,x)\,.  \\
\end{alignat*}

However, note that the following compositions are not valid:

\begin{alignat*}{1}
wi2(x,y1,"CCS"y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\, \\
\ & fromWAS("WAS",y3),\,  national(y3,x)\,. \\
wi3(x,y1,y2,y3)\ &\qrule\ national(x1,y1),\, onestop(x1,y2),\,  oneway(y2,y3),\,  \\
\ & national(y3,x)\,. \\
\end{alignat*}

The composition {\it wi2} is not valid because  it maps the workflow variable
{\it y2} into constants {\it "CCS"} and {\it "WAS"} which denote different cities.
On the other hand, the composition {\it wi3} does not implement the workflow because the concrete service  {\it onestop} does not receive as input or produce as output, the city where the flight stops (represented by the variable {\it y1}); thus, it is not possible to ensure that the city where this flight stops and the end city  retrieved by the service {\it national}, are the same.

To avoid producing not valid workflow implementations, it is required to handling constants in a way that different constants are not mapped into
each other either directly or indirectly (via transitivity). In addition,  all the attributes in the workflow output  or needed to join other services in the workflow, have to be produced by the selected concrete service.

In addition, suppose each concrete service is annotated with a utility function that aggregates the values of different QoS that characterize the behavior of the service. Then, since the space of valid service compositions of an abstract workflow can be very large, it is required a technique able to identify the ones that maximize the utility function  without having to enumerate the whole set of valid compositions. In this paper, we propose a Propositional logic-based  approach that takes advantage of the  power of  modern SAT solvers, to efficiently enumerate the service compositions that correspond to valid implementations of a given abstract workflow, as well as the compositions that maximize a given utility function.

\section{Related Work}
In this section we summarize existing approaches that provide solutions to the problems of service selection, knowledge representation and query rewriting.

\begin{description}
\item[Service Selection Solutions] \mbox{}\\
The problem of selecting the services that implement an abstract workflow and best meet the QoS-based criteria  is known as the QoS-aware service selection or composition problem, which has been shown to be NP-hard~\cite{Hiroshi2008}. This problem is a combinatorial optimization problem and several heuristics have been proposed to find a relatively good solution in a reasonably short period of time.  A  distance metric-based heuristic  to drive a backward search algorithm  is proposed~\cite{rahmani08}; this metric induces an order of the services in a way that sink nodes are unlikely to be visited. In ~\cite{berardi05,berardi08,berardi06}, services and workflows are described in terms of deterministic finite state machines that are encoded as a Description Logic theory whose models correspond to solutions of the problem;  although reasoning methods for Description Logics formalisms could be exploited, scalability or performance of the proposed solution has not been reported.  ~\cite{myoung08} propose a constraint-based approach that encodes the non-functional permissible values as a set of constraints whose violation needs to be minimized; to traverse the space of possibly optimal solutions, a hybrid algorithm that combines the tabu search and simulating annealing meta-heuristics is implemented; experimental results show that the proposed solution is able to scale up to a large number of services and abstract processes.  In ~\cite{cardellini07} the QoS-aware service composition problem is encoded as a Linear Programming problem providing a scalable solution to the problem.  In ~\cite{Hiroshi2008}  this problem  is defined as a multi-objective optimization problem where the different QoS parameters are considered equally important and there is not an aggregated function to combine all of them; a genetic-based algorithm is proposed to identify a set of non-dominated service compositions that best meet all the QoS parameters.  ~\cite{alrifaiR09} propose a two-fold solution that uses a hybrid integer programming algorithm to find the decomposition of global QoS into local constraint, and then,  selects the services that best meet the local constraints.   
Recently,  two new planning-based approaches have been proposed~\cite{kuterG09,sohrabiM09}.  ~\cite{kuterG09} extend the SHOP2 planning algorithm to select the trustworthy composition of services that implement a given OWL-S process model, while ~\cite{sohrabiM09}  propose a HTN planning-based solution where user preference metrics and domain regulations are used to guide the planner into the space of relevant compositions. Finally, ~\cite{lecue09} proposes a genetic-based algorithm to identify the composition of services that best meet the quality criteria for a set of QoS parameters.

Although these solutions are able to efficiently solve the optimization problem and scale up to a large number of abstract processes, none of them are tailored to semantically describe services in terms of the abstract process, or use these descriptions to identify the services that  implement a given workflow and best meet user non-functional criteria.

\item[Knowledge Compilation Languages] \mbox{}\\
Knowledge compilation is the area in AI concerned with the
problem of mapping logical theories into suitable fragments
that make certain desired operations tractable
\cite{cadoli:compilation}. Different compilation languages have been defined, for instance,  
Ordered Binary Decision Diagrams (OBDDs)~\cite{bryant:obdd},  Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation Normal Form (DNNF)  \cite{darwiche:dnnf}.
In this work we make use of the properties of the deterministic DNNFs  (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable and efficient solution to the service selection problem.

A Negation Normal Form (NNF) theory is  constructed from literals using only conjunctions
and disjunctions \cite{barwise:handbook}, and it can be represented as a directed acyclic graph
in which the leaves are labeled with literals and the
internal nodes are labeled with $\land$ and $\lor$;
see Fig.~\ref{fig:dnnf} for an example. An NNF is said to be decomposable (DNNF) \cite{darwiche:dnnf}
if for each conjunction $\bigwedge_i\phi_i$, the set of
variables in each conjunct are pairwise disjoint;
i.e,. $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i<j$.
A DNNF supports a number of operations in
polynomial time in the size of its DAG. For example,
we can test whether a DNNF is satisfiable by a single
bottom-up pass over its DAG in linear time. A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each disjunction $\bigvee_i\phi_i$, the disjuncts
are pairwise logically contradictory; i.e.,
$\phi_i\Rightarrow\neg\phi_j$ for $i<j$.
The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable
and deterministic. A d-DNNF supports the model counting and enumeration in polynomial time
in the size of its DAG.  {\it MCDSat}$^c$  exploits the properties of a d-DNNF theory, to efficiently provide a solution to the problem of enumerating the service compositions that correspond to models of the  theory, i.e., the service compositions that implement the abstract  workflow and best meet the utility function. DECIR ALGO SOBRE EL ALGORITMO QUE PRODUCE LOS MEJORES.

\begin{figure}
\centering
\includegraphics[width=8cm]{odd}
\caption{A decomposable and deterministic NNF.}
\label{fig:dnnf}
\end{figure}

\item[Query Rewriting Solutions] \mbox{}\\
A number of algorithms have been developed to find the rewritings of a given query; the most prominent being the bucket algorithm~\cite{levy:bucket},  the inverse rules algorithm ~\cite{duschka:answer,Qian96}, the minicon algorithm ~\cite{pottinger:minicon}, and the {\it MCDSat} ~\cite{arvelo:aaai06}. Generally, query rewriting algorithms work in two phases: first, they identify the views that rewrite at least one  subgoal of the query; second they combine the selected views to produce a rewriting. The main difference between existing approaches, is the criteria used to choose the relevant views and reduce the space of non-useful rewritings.

The bucket algorithm reduces the number of possibilities just considering each subgoal in the query in isolation, and selecting the views that are able to  produce at least the attributes projected by the query. Since, attributes involved in query joins are not verified, a large number of rewritings comprised of  Cartesian products may  be generated. 	
The Inverse Rules algorithm constructs a set of rules that invert the view definition and establish how to compute tuples for the database relations from tuples of the views. Similarly, to the bucket algorith it can produce a large number non-useful of rewritings.

The MiniCon algorithm  overcomes the limitations of the previous algorithms by identifying only views that rewrite a set of the query goals and that  can be combined with the rest of the query subgoals. The key idea is to identify the mappings between  the variables in each subgoal of the query to the variables  in one or more subgoals of the views, in a way that, join variables in the query are mapped to join variables in the body of a view or to the distinguished variables of the view. Mappings between variables and subgoals are represented in MiniCon Descriptions (MCD's)\cite{pottinger:minicon}.

Finally,  the
{\it MCDSat} algorithm is able to identify the query rewritings of a query by translating the problem of rewriting
into the problem of enumerating the models of a propositional
theory  $\Theory{\Q}$ whose models are in correspondence
with the rewritings of the query. 
The {\it MCDSat} algorithm exploits the properties of d-DNNFs to efficiently 
compute the MCDs associated with theory $\Theory{\Q}$.
The  {\it MCDSat} algorithm has demonstrated to scale better than the MiniCon
algorithm over a large number of benchmarks often showing performance
improvements of several orders of magnitude. However, the McdSat algorithm was not desgined for rewriting problems involving explicit constants,
nor to compute the best rewritings with respect to a given utility function or cost model, and this paper we propose a new encoding  that overcomes these limitations.

\end{description}

\section{The McdSat$^c$ Architecture}

In order to do so, the theory $\Theory{\Q}$ is transformed into a d-DNNF
theory $\Delta(\Q)$ with a CNF to d-DNNF compiler.
The compiler's algorithm is similar to the DPLL algorithm
for SAT but enhanced with advanced techniques such as clause
learning and caching.
\begin{figure}
\centering
\includegraphics[width=10cm]{mdcsatC.pdf}
\caption{The McdSat$^c$ Architecture}
\label{fig:dnnf}
\end{figure}

\section{Experimental Results}
We have conducted an empirical analysis  on the benefits of the techniques implemented in the   McdSat$^c$ system.

\begin{description}
\item[Dataset and Query Benchmark:]  we conducted our experiments over a benchmark that includes XX queries and a set of  YY views. Queries and views are starts and chains, and they have between ZZ and fTT constants; queries have QQ sub-subgoals.
\item[Evaluation Metrics:] we report on runtime performance which corresponds to the  {\it user time} produced by the {\it time} command of the Unix operation system. 
\end{description}

\begin{figure}
\centering
\includegraphics[width=8cm]{plot1.pdf}
\caption{Compilation Time Benchmark I}
\label{fig:plot1}
\end{figure}

\section{Formalization of the Service Selection Problem}

We consider databases of the form $D=\tup{P,T}$ where
$P$ is a set of predicates and $T=\{T_p\}_{p\in P}$ is a collection
of tables that represents the predicates in extensional form.
A conjunctive query $Q$ over $P$ is of the form 
\begin{alignat*}{1}
Q(\vvec{x})\ \qrule\ \  p_1(\vvec{x}_1),\, p_1(\vvec{x}_2),\, \ldots,\, p_m(\vvec{x}_m)\,,
\end{alignat*}
where $p_i\in P$, $\vvec{x}$ is a vector of variables, and each
$\vvec{x}_i$ is a vector of variables and constants.
The result of $Q$ over $D$, denoted as $Q(D)$, is the table with
$|\vvec{x}|$ columns that result of the projection of the relational
join $\join\!\!\{T_{p_i}\}_{i=1}^m$ over $\vvec{x}$.
The atoms in the body of $Q$ are called the (sub)goals of $Q$.

A view $V$ over $D$ is a query over $P$. In the context of the service selection problem, 
the database $D$ is an idealized description of 
the output produced abstract workflow  implemented by multiple concrete services described
as views.
Given a database $D$, a query $Q$ and a collection of 
views $E=\tup{\{V_i\}_i,\{E_i\}_i}$, we are required to find
all the tuples in $Q(D)$ obtainable from the views in $E$.
That is, we need to find all the \emph{compositions} of the form
\begin{alignat*}{1}
R(\vvec{x})\ \qrule\ \ V_{i_1}(\vvec{x}_1),\, V_{i_2}(\vvec{x}_2),\, \ldots,\, V_{i_n}(\vvec{x}_n)
\end{alignat*}
such that $R(E) \subseteq Q(D)$.
A service selection problem (SSP) is a tuple $\tup{P,Q,\{V_i\}}$ where $P$
is a set of predicates that represent abstract services, $Q$ is a query over $P$ and $\{V_i\}$ is a collection of views that define the concrete services in terms of  abstract services. We assume \emph{safe} problems in the sense that all 
variables mentioned in the head of the query (resp.\ in the head of each view)
appear in the body of the query (resp.,\ in the body of each view); also the input and output restrictions of the abstract services used in the query are satisfied.
Further, we only deal with SSPs with no arithmetic predicates inside the
query or views.
A composition $C$ is \emph{valid} if for all databases $D=\tup{P,T}$
and extensions $\{E_i\}$, $R(E) \subseteq Q(D)$.
A collection $\R$ of valid compositions is a solution if
for all databases $D=\tup{P,T}$ and extensions $\{E_i\}$, there
is no other $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(D)$.
We are interested in finding a composition $\R$.

\subsection{Logical Theories}

In \cite{arvelo:aaai06}, we showed that a rewriting
for a query $Q$ with $m$ goals can be obtained by enumerating the
models of a logical theory
$\Delta=\Delta_{rew}\cup\Delta_{mcd}^1\cup\cdots\Delta_{mcd}^m$
where $\Delta_{rew}$ specified how to combine $m$ independent
copies of MCD theories $\Delta_{mcd}$ that cover all goals in $Q$.
Each $\Delta^i_{mcd}$ is a copy of the theory $\Delta_{mcd}$
in which each literal $p$ is tagged as $p^i$.
The theory $\Delta_{mcd}$ consists of different groups of 
clauses that guarantees that its models are in correspondence
with the MCDs, while the theory $\Delta_{rew}$ contains additional
clauses to guarantee a sound and complete composition of the MCDs.
The reader is referred to \cite{arvelo:aaai06} for a comprehensive
description of the propositional theory.

Let us describe the difficulties that arise when constants
are presents in a SSP with the following example:
\begin{alignat*}{1}
Q(x,z)\ &\qrule\ \ p_1(x,y),\, p_2(y,z)\,, \\
V_1(x_1)\ &\qrule\ \ p_1(x_1,A)\,, \\
V_2(x_2)\ &\qrule\ \ p_2(B,x_2)\,, \\
V_3(x_3,y_3)\ &\qrule\ \ p_2(x_3,y_3)
%V_4(x_4,y_4)\ &\qrule\ \ p_1(x_1,A),\, p_2(B,y_4)\,,
\end{alignat*}
where $Q$ is the query and $\{V_1,V_2,V_3\}$ are the 
views. In this case, the only rewriting is 
$R(x,z)\qrule V_1(x),V_3(A,z)$ because the candidate 
$R(x,z)\qrule V_1(x),V_2(z)$ is not valid as it maps
$y$ into constants $A$ and $B$ that denote different
objects.

The main problem when handling constants is to
be sure that different constants are not mapped into
each other either directly or indirectly (via transitivity).

\subsection{Utility Functions}

We assume a simple additive utility function in which each view
$V_i$ is associated with a utility measure $c(V_i)$, and the overall utility value of
a composition is the sum of the utility values for the views in it.
An optimal or best service composition is one with maximum utility value,
and the optimal utility value of a SSP is the utility value of a best
service composition. A SSP has always a well-defined optimal
utility value (if there are no compositions, its utility is $0$),
but it may have multiple best service compositions.
The service composition problem with utility consists in finding
all the compositions of maximum utility value.

%The cost model is simple yet expressive; e.g., if all
%costs are unit, then best rewritings are those with
%minimum number of views, yet in cases where views
%represent different sources, it is natural
%to associate them with different costs.

\subsection{Extended Theories}

The theory $\Delta_{mcd}$ makes use of propositions $t_{x,y}$
to denote that the variable/constant $x$ in the query is mapped
into the variable/constant $y$ in the view, and propositions $v_i$
to indicate that the MCD uses view $V_i$.
We obtain a sound and complete theory for SSPs with constants if
$\Delta_{mcd}$ is extended with the clauses:

\begin{enumerate}[C1.]\denselist
\item (Inconsistent-1): $t_{x,A} \Rightarrow \neg t_{x,B}$,
\item (Inconsistent-2): $t_{A,x} \Rightarrow \neg t_{B,x}$,
\item (Inconsistent-3): $\neg t_{A,B}$,
\item (Transitivity-1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$,
\item (Transitivity-2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
\end{enumerate}
Clauses C1--C3 prune candidate service compositions in which one constant
is directly mapped into a different one, and the last two implement
a restricted propagation of transitivity.
Similarly, $\Delta_{rew}$ must be extended with the clauses:
\begin{enumerate}[C1.]\denselist
\item[C6.] (Inconsistent-4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
\end{enumerate}
~\cite{RajaramanSU95} showed that for queries without negation or arithmetic comparisons,  but with constants, and $m$ goals and $q$ variables, it is enough
to consider service compositions of length at most  $m$ plus  $q$ subgoals ~\cite{Ullman00}.

Given a SSP problem $\tup{Q,\{V_i\}}$ possibly with constant 
symbols, we construct logical theories whose models are in
correspondence with the service compositions. These theories are then
compiled into d-DNNF from which all rewritings are extracted
efficiently.
Likewise, best service compositions can be computed by performing model
enumeration twice: the first pass computes the optimal utility function,
and the second filters out suboptimal service compositions.
However, there is a better way, which we will describe in the next section.

\subsection{Maximum-Utility Function}

Darwiche \& Marquis \cite{darwiche:weighted} show how to compute the
rank $r^*(\Delta)$ of a propositional theory $\Delta$
efficiently when $r$ is a literal ranking function and
$\Delta$ is in d-DNNF. A literal ranking function ranks
the models of the theory in terms of the rank of the
literals $l$ that are true in the model, and ranks the
theory as the best rank of its models:
\[ r(\omega) = \sum_{\omega\vDash l} r(l),\quad r^*(\Delta) = \max_{\omega\vDash\Delta} r(\omega)\,. \]
Furthermore, not only $r^*(\Delta)$ can be computed
efficiently from the d-DNNF but also the models that
have such rank.
The procedure for computing the rank converts the d-DNNF
into a circuit in which the literals are replaced by their
ranks, the `or' nodes by `max', and the `and' nodes by `sum'.
The evaluation of the circuit computes the rank of $\Delta$.

We thus obtain a simple method for computing the best
rewritings when the literal ranking function $r_c$ induced
by the cost model $c$ is used; $r_c$ is defined by
$r_c(l)=c(V_i)$ if $l=v^t_i$ for some $t$, and $r_c(l)=0$
otherwise.

\section{Conclusions and Future Work}

We have shown how the propositional theory used in McdSat can
be extended to support constants, and how it can be used to 
compute the best rewritings with respect to an additive
cost model. The cost model is simple yet expressive.
The computation of the best models relies on an arithmetic
circuit that is obtained from the d-DNNF of the propositional
theory.

\bibliographystyle{abbrv}
\bibliography{ref}

\end{document}

### 2009-12-13 publicaciones/eswc10/paperESWC10.tex
pequeñas correcciones a paper de mvidal

--- 2009-12-11 tesis/propuesta/propuesta_iae.tex
@@
 model the problem of service selection as a known problem in the area of integration systems referred as query rewriting using views.
-We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and maximizes the overall utility function; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.
+We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and have the same cost as the corresponding composition; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.
 We report on the performance of our proposed approach, and we observe that  {\it McdSat}$^c$  efficiently scales up to large instances of the service selection problem.
@@
 To evaluate the proposed workflow,  abstract services are instantiated on concrete services in a way restrictions on the QoS parameters are satisfied.
-However, the performance of this tasks is impacted by the number of available sources,  techniques enable to efficiently scale up to a large number of services are required.
+However, since the number of available sources has an impact on the performance of this task, techniques that enable us to efficiently scale up to a large number of services are required.
 
@@
 
-In {\it MCDSat}$^c$, concrete service functionalities are decribed as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.
+In {\it MCDSat}$^c$, concrete service functionalities are described as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.
 In addition, rules that define workflows and  concrete services in the {\it MCDSat}$^c$ catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied.
@@
 {\it y2} into constants {\it "CCS"} and {\it "WAS"} which denote different cities.
-On the other hand, the composition {\it wi3} does not implement the workflow because the concrete service  {\it onestop} does not receive as input or produce as output, the city where the flight stops (represented by the variable {\it y1}); thus, it is not possible to ensure that the city where this flight stops and the end city  retrieved by the service {\it national}, are the same.
+On the other hand, the composition {\it wi3} does not implement the workflow because the concrete service  {\it onestop} does not receive as input or produce as output the city where the flight stops (represented by the variable {\it y1}); thus, it is not possible to ensure that the city where this flight stops and the end city  retrieved by the service {\it national}, are the same.
 
-To avoid producing not valid workflow implementations, it is required to handling constants in a way that different constants are not mapped into
+To avoid producing invalid workflow implementations, it is required to handle constants in a way that different constants are not mapped into
 each other either directly or indirectly (via transitivity).
@@
 In addition, suppose each concrete service is annotated with a utility function that aggregates the values of different QoS that characterize the behavior of the service.
-Then, since the space of valid service compositions of an abstract workflow can be very large, it is required a technique able to identify the ones that maximize the utility function  without having to enumerate the whole set of valid compositions.
+Then, since the space of valid service compositions of an abstract workflow can be very large, we require a technique able to identify the ones that maximize the utility function  without having to enumerate the whole set of valid compositions.
 In this paper, we propose a Propositional logic-based  approach that takes advantage of the  power of  modern SAT solvers, to efficiently enumerate the service compositions that correspond to valid implementations of a given abstract workflow, as well as the compositions that maximize a given utility function.
@@
 In ~\cite{berardi05,berardi08,berardi06}, services and workflows are described in terms of deterministic finite state machines that are encoded as a Description Logic theory whose models correspond to solutions of the problem;  although reasoning methods for Description Logics formalisms could be exploited, scalability or performance of the proposed solution has not been reported.
-~\cite{myoung08} propose a constraint-based approach that encodes the non-functional permissible values as a set of constraints whose violation needs to be minimized; to traverse the space of possibly optimal solutions, a hybrid algorithm that combines the tabu search and simulating annealing meta-heuristics is implemented; experimental results show that the proposed solution is able to scale up to a large number of services and abstract processes.
+~\cite{myoung08} propose a constraint-based approach that encodes the non-functional permissible values as a set of constraints whose violation needs to be minimized; to traverse the space of possibly optimal solutions, a hybrid algorithm that combines the tabu search and simulated annealing meta-heuristics is implemented; experimental results show that the proposed solution is able to scale up to a large number of services and abstract processes.
 In ~\cite{cardellini07} the QoS-aware service composition problem is encoded as a Linear Programming problem providing a scalable solution to the problem.
@@
 \item[Dataset and Query Benchmark:]  we conducted our experiments over a benchmark that includes XX queries and a set of  YY views.
-Queries and views are starts and chains, and they have between ZZ and fTT constants; queries have QQ sub-subgoals.
+Queries and views are starts and chains, and they have between ZZ and TT constants; queries have QQ sub-subgoals.
 \item[Evaluation Metrics:] we report on runtime performance which corresponds to the  {\it user time} produced by the {\it time} command of the Unix operation system. 

### 2009-12-13 tesis/propuesta/propuesta_iae.tex
avance de traduccion a español de propuesta hasta motivacion

@@
 \usepackage[pdftex]{graphicx}
+\usepackage[utf8x]{inputenc}
 \usepackage{enumerate}
@@
 \allowdisplaybreaks
-\title{Selecting the Maximum-Utility  Web Services  Using Circuits}
-\author{Daniel Izquierdo \inst{1}  and Blai Bonet \inst{1}  and Mar\'{\i}a-Esther Vidal \inst{1}}
+\title{Selección de Servicios Web De Máxima Utilidad Usando Circuitos}
+\author{Daniel Izquierdo}
 
@@
                 Caracas, Venezuela\\
-                \{idaniel,bonet,mvidal\}ldc.usb.ve} \maketitle
+                idaniel@ldc.usb.ve} \maketitle
 \begin{abstract}
-In Service Oriented Architectures (SOA), concrete services are described in terms of functionality   and QoS parameters, while software applications are specified as workflows of abstract services and non-functional criteria that establish the  application desirable functionality and  expected quality, respectively.
-During workflow execution time,  abstract services  are rewritten
-with the concrete services  that meet the workflow  functional and non-functional restrictions.
-Since, service selection is done on-the-flight and the space of concrete services may be very large, efficiency and scalability are required.
-In this paper we propose the {\it McdSat}$^c$ framework to efficiently solve the problem of service selection in presence of a large number of concrete services.
-First, we use the Local As View approach to describe the functionality  of the concrete services in terms of views of abstract services, the quality as an overall utility function that combines the different QoS values that describe the service, and  workflows as  conjunctive queries on abstract services.
-Based on this representation, we
-model the problem of service selection as a known problem in the area of integration systems referred as query rewriting using views.
-We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and maximizes the overall utility function; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.
-We report on the performance of our proposed approach, and we observe that  {\it McdSat}$^c$  efficiently scales up to large instances of the service selection problem.
+En las arquitecturas orientadas a servicios (SOA), los servicios concretos se
+describen en términos de parámetros de funcionalidad y calidad de servicio,
+mientras que las aplicaciones de \emph{software} se especifican como flujos de
+trabajo de servicios abstractos y criterios no-funcionales que establecen la
+funcionalidad deseada y la calidad esperada respectivamente. En tiempo de
+ejecución de un flujo de trabajo, los servicios abstractos se reescriben con los
+servicios concretos que cumplen con las restricciones funcionales y no
+funcionales del flujo de trabajo. Dado que la selección de los servicios es
+hecha al momento y el espacio de servicios concretos puede ser muy grande, se
+requiere eficiencia y escalabilidad. En este trabajo se propone la
+infraestructura {\it McdSat}$^c$ para resolver eficientemente el problema de
+selección de servicios en presencia de un número grande de servicios concretos.
+En primer lugar se utiliza la aproximación \emph{Local As View} para describir
+la funcionalidad de los servicios concretos en términos de vistas de servicios
+abstractos, la calidad como una función de utilidad general que combina los
+distintos valores de calidad que describen cada servicio, y los flujos de
+trabajo como consultas conjuntivas sobre servicios abstractos. Basándose en esta
+representación, se modela el problema de selección de servicios como un problema
+conocido en el área de integración de sistemas conocido como reescritura de
+consultas usando vistas. Codificamos el problema de selección de servicios como
+una teoría proposicional cuyos modelos corresponden a la composición de los
+servicios que implementan el flujo abstracto; se explotan relaciones conocidas
+entre teorías d-DNNF y circuitos aritméticos para proveer una solución eficiente
+y escalable. Se reporta sobre el rendimiento de la aproximación propuesta y se
+observa que  {\it McdSat}$^c$  escala a instancias grandes del problema de
+selección de servicios.
 
 \end{abstract}                
-\section{Introduction}
-Under the umbrella of the Semantic Web and supported by Service Oriented architectures, the number of Web data sources and services has exploded during the last few years.
-For example,  currently, the molecular biology databases
-collection includes 1,170 databases~\cite{Galperin09}, that is
-95 more than the previous year~\cite{Galperin2008} and 110 more than two years ago~\cite{Galperin2007};  tools
-and services as well as the number of instances published by these resources, follow a similar progression~\cite{Benson07}. Thanks
-to this wealth, users rely more on various
-digital tasks such as data retrieval from public data sources and
-data analysis with Web tools or services organized in
-complex workflows.
+\section{Introducción}
+Bajo la sombra de la Web Semántica y con el apoyo de arquitecturas orientadas a
+servicios, el número de fuentes de datos y servicios Web ha explotado en los
+últimos años. Por ejemplo, actualmente la colección de bases de datos de
+biología molecular incluye 1.170 bases de datos~\cite{Galperin09}, lo cual es 95
+más que el año pasado~\cite{Galperin2008} y 110 más que hace dos
+años~\cite{Galperin2007}. Tanto las herramientas y servicios como el número de
+instancias publicadas por estos recursos siguen una progresión
+similar~\cite{Benson07}. Gracias a este tesoro de datos, los usuarios se apoyan
+más en varias tareas digitales como obtención de datos de fuentes públicas y
+análisis de datos con herramientas o servicios Web organizados en flujos de
+trabajo complejos.
 
-In order to design an application, users usually specify the functionality in terms of a workflow of abstract services, while the desirable quality  is defined by restricting the values of the QoS parameters that describe the concrete available services.
-To evaluate the proposed workflow,  abstract services are instantiated on concrete services in a way restrictions on the QoS parameters are satisfied.
-However, the performance of this tasks is impacted by the number of available sources,  techniques enable to efficiently scale up to a large number of services are required.
+Para diseñar una aplicación, los usuarios frecuentemente especifican la
+funcionalidad en términos de un flujo de trabajo de servicios abstractos,
+mientras que la calidad deseada se define restringiendo los valores de los
+parámetros QoS (calidad de servicio) que describen los servicios concretos
+disponibles. Para evaluar el flujo propuesto, se instancian los servicios
+abstractos con concretos de manera que se satisfagan las restricciones sobre
+parámetros QoS. Sin embargo, el rendimiento de estas tareas es impactado por el
+número de fuentes disponibles. Se requieren técnicas que permitan escalar
+eficientemente a un número grande de servicios.
 
-In this paper we aim to provide a solution to this problem and provide techniques that efficiently traverse the space of available services and select combinations that meet the functional and non-functional requirements that describe a designed application.
-We name our proposed solution {\it MCDSat}$^c$.
+En este trabajo se intenta proponer una solución a este problema y proveer
+técnicas que recorran eficientemente el espacio de servicios disponibles y
+seleccionen combinaciones que cumplan con los requerimientos funcionales y
+no-funcionales que describen una aplicación diseñada. Se llama 
+{\it MCDSat}$^c$ a la solución propuesta.
 
-In {\it MCDSat}$^c$, concrete service functionalities are decribed as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.
-In addition, rules that define workflows and  concrete services in the {\it MCDSat}$^c$ catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied.
-To represent the QoS measures, each concrete service description is annotated with a real number that represents an overall value of the QoS measures of this service.
-Finally, {\it MCDSat}$^c$ models the web service selection problem as the problem of query rewriting using views \cite{halevy:survey}.
-This problem is important in the context of data integration
-\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance \cite{AfratiLU07,levy:bucket}, and several approaches have been defined to efficiently scale up to a large number of views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-In this paper we propose the {\it MCDSat}$^c$ framework which extends the approach proposed in ~\cite{arvelo:aaai06},  with the capability of dealing with constants and identifying the composition of services that maximizes a given utility function.
+En {\it MCDSat}$^c$, las funcionalidades de servicios concretos se describen
+como vistas sobre servicios abstractos y los flujos de trabajo se representan
+como consultas conjuntivas sobre servicios abstractos; esta representación de
+los servicios es similar a la generada de manera semiautomática para el
+sistema {\cal DEIMOS}~\cite{AmbiteISWC09}. Adicionalmente, las reglas
+que definen flujos de trabajo y servicios concretos en el catálogo de
+{\it MCDSat}$^c$ son creadas de manera que las restricciones de entrada y salida
+de los servicios abstractos combinados sean satisfechas. Para representar las
+medidas de QoS, cada descripción de servicio concreto es anotada con un número
+real que representa un valor general para las medidas de QoS del servicio en
+particular. Finalmente, {\it MCDSat}$^c$ modela el problema de selección de
+servicios web como el problema de reescritura de consultas usando vistas
+\cite{halevy:survey}. Este problema es importante en el contexto de integración
+de datos \cite{Chen05,JaudoinPRST05} y de optimización de consultas y
+mantenimiento de datos \cite{AfratiLU07,levy:bucket} y se han definido varias
+aproximaciones para escalar eficientemente a un número grande de
+vistas~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
+En este trabajo se propone la infraestructura {\it MCDSat}$^c$ que extiende la
+aproximación propuesta en  ~\cite{arvelo:aaai06}, con la capacidad de trabajar
+con constantes y de identificar la composición de servicios que maximice una
+función de utilidad dada.
 
-{\it MCDSat}$^c$  translates an instance of the service selection problem into the problem of enumerating the models of a propositional theory that encodes information about the workflow, the concrete services and the conditions that need to be satisfied in order to generate a composition that maximizes the utility function.
-This theory is represented as a CNF formula which is compiled into a deterministic
-and decomposable negation normal form d-DNNF representation, from which the models that maximizes the overall utility function are obtained in linear time using modern SAT techniques~\cite{darwiche:dnnf}.
-We empirically studied the quality of the {\it MCDSat}$^c$ on variety of benchmarks and we observed  that the proposed technique provides an efficient solution to the service selection problem which is able to scale up to large workflows as well to a large number of views.
+{\it MCDSat}$^c$  traduce una instancia del problema de selección de servicios
+en el problema de enumerar los modelos de una teoría proposicional que codifica
+información sobre el flujo de trabajo, los servicios concretos y las condiciones
+que deben ser satisfechas para generar una composición que maximice la función
+de utilidad. Esta teoría se representa como una fórmula CNF que es compilada a
+su representación d-DNNF, a partir de la cual los modelos que maximicen la
+función global de utilidad se obtienen en tiempo lineal usando técnicas modernas
+de SAT~\cite{darwiche:dnnf}. Se estudia empíricamente la calidad de {\it
+MCDSat}$^c$ sobre una variedad de \emph{benchmarks} y se observa que la técnica
+propuesta provee una solución eficiente al problema de selección de servicios
+que es capaz de escalar a flujos de trabajo grandes así como a números de vistas
+grandes.
 
@@
 
-\section{Motivating Example}
-Consider the abstract and concrete services in Table ~\ref{table:services}  which represent applications in the flight domain.
+\section{Motivación}
+
+Consideremos los servicios abstractos y concretos en la tabla
+~\ref{table:services} que representan aplicaciones en el dominio de vuelos.
 
@@
 \hline
-Abstract Service & Description \\
+Servicio Abstracto & Descripción \\
 \hline
-$fight(x1,x2)$ & relates cities x1 and x2 if there is a flight between them\\
+$fight(x1,x2)$ & relaciona las ciudades x1 y x2 si hay un vuelo entre ellas\\
 \hline
-$uscity(x1)$ &  determines if a city x1 is a US city\\
+$uscity(x1)$ &  determina si una ciudad x1 es una ciudad de Estados Unidos\\
 \hline \hline
-Concrete  Service & Description \\
+Servicio Concreto & Descripción \\
 \hline
-$national(x1,y1)$&  returns  two US cities that are connected by  direct flights \\
+$national(x1,y1)$&  retorna dos ciudades de Estados Unidos que están conectadas
+por vuelos directos \\
 \hline
-$oneway(x1,y1)$ & returns two cites that are connected by one way direct flights \\
+$oneway(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos directos
+de ida\\
 \hline
-$onestop(x1,y1)$ &  returns  two cites that are connected by one-stop flights \\
+$onestop(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos con una
+escala\\
 \hline
-$ flightToCCS(x1,"CCS")$ & returns a city that is connected to  Caracas by a direct flight\\
+$ flightToCCS(x1,"CCS")$ & retorna una ciudad que está conectada a Caracas por
+un vuelo directo\\
  \hline
-$ onestopToCCS(x1,y1)$ & returns two cites x1 and y1, such that there is  one-stop flight\\ 
- & from city x1 to Caracas stopping at city y1.\\
+$ onestopToCCS(x1,y1)$ & retorna dos ciudades x1 y y1 tales que hay un vuelo con
+una escala desde la ciudad x1 hasta Caracas con parada en la ciudad y1\\
  \hline
- $fromWAS("WAS",y1)$ &  returns a city y1 to which flights from Washington arrive. \\
+ $fromWAS("WAS",y1)$ &  retorna una ciudad y1 a la cual llegan vuelos desde
+Washington.\\
  \hline
@@
 \label{table:services}
-\caption{Abstract and Concrete Services}
+\caption{Servicios Abstractos y Concretos}
 \end{table}
@@
 \centering
-\includegraphics[width=8cm]{odd}
+%\includegraphics[width=8cm]{odd}
 \caption{A decomposable and deterministic NNF.}
@@
 \centering
-\includegraphics[width=10cm]{mdcsatC.pdf}
+%COMENTADO\includegraphics[width=10cm]{mdcsatC.pdf}
 \caption{The McdSat$^c$ Architecture}
@@
 \centering
-\includegraphics[width=8cm]{plot1.pdf}
+%COMENTADO\includegraphics[width=8cm]{plot1.pdf}
 \caption{Compilation Time Benchmark I}

### 2009-12-13 tesis/propuesta/propuesta_iae.tex
mas cambios a la propuesta camino a cambiarla al formato correcto

@@
 \end{abstract}
+
 \section{Introducción}
@@
 
-Using the  Local As View (LAV) approach~\cite{AmbiteISWC09}, concrete services in Table 1 are semantically described as views on the abstract services {\it flight} and {\it uscity} as follows:
+Usando la aproximación \emph{Local As View} (LAV) ~\cite{AmbiteISWC09}, los
+servicios concretos en la tabla \ref{table:services} son descritos
+semánticamente como vistas sobre los servicios abstractos {\it flight} y
+{\it uscity} como sigue:
 
@@
 
-Now suppose a user is interested in implementing a workflow able to retrieve the one-stop round trip flights from US cities to any city in the world, such that flights can stop at any city. 
-Consider the following conjunctive query that represents the workflow that defines this request in terms of abstract services.
-This workflow was defined in way the input parameters of each abstract service are bound by  constants in the workflow, or by attributes projected out by  previous abstract services.
+Ahora supongamos que un usuario está interesado en implementar un flujo de
+trabajo capaz de obtener los vuelos ida y vuelta con una escala de ciudades de
+Estados Unidos a cualquier ciudad en el mundo donde las escalas pueden ser
+hechas en cualquier ciudad. Consideremos la siguiente consulta conjuntiva que
+representa el flujo de trabajo que define esta solicitud en términos de
+servicios abstractos. Este flujo fue definido de manera que los parámetros de
+entrada de cada servicio abstracto están ligados a constantes o a
+atributos proyectados por servicios abstractos previos.
 
@@
 
-Implementations of this workflow correspond to compositions of the concrete services where each concrete service can implement a subset of abstract services of the workflow, but each abstract service can be implemented by exactly one concrete service.
-The following composition of concrete services corresponds to one of the XX implementations of the workflow:
+Las implementaciones de este flujo de trabajo corresponden a composiciones de
+servicios concretos donde cada servicio concreto puede implementar un
+subconjunto de los servicios abstractos del flujo de trabajo, pero cada servicio
+abstracto puede ser implementado por exactamente un servicio concreto. La
+siguiente composición de servicios concretos corresponde a una de las XX
+implementaciones del flujo de trabajo.
 
@@
 
-However, note that the following compositions are not valid:
+Sin embargo, nótese que las siguientes composiciones no son válidas:
 
@@
 
-The composition {\it wi2} is not valid because  it maps the workflow variable
-{\it y2} into constants {\it "CCS"} and {\it "WAS"} which denote different cities.
-On the other hand, the composition {\it wi3} does not implement the workflow because the concrete service  {\it onestop} does not receive as input or produce as output, the city where the flight stops (represented by the variable {\it y1}); thus, it is not possible to ensure that the city where this flight stops and the end city  retrieved by the service {\it national}, are the same.
+La composición {\it wi2} no es válida porque asocia la variable del flujo de
+trabajo {\it y2} a las constantes {\it "CCS"} y {\it "WAS"} que denotan ciudades
+distintas. Por otra parte, la composición {\it wi3} no implementa el flujo de
+trabajo porque el servicio concreto {\it onestop} no recibe como entrada o
+produce como salida la ciudad donde el vuelo se detiene (representada por la
+variable {\it y1}); por esto, no es posible garantizar que la ciudad donde el
+vuelo se detiene y la ciudad final obtenida por el servicio {\it national} sean
+las mismas.
 
-To avoid producing not valid workflow implementations, it is required to handling constants in a way that different constants are not mapped into
-each other either directly or indirectly (via transitivity).
-In addition,  all the attributes in the workflow output  or needed to join other services in the workflow, have to be produced by the selected concrete service.
+Para evitar producir implementaciones del flujo de trabajo inválidas, se
+requiere manejar constantes de manera que constantes distintas no sean asociadas
+entre sí directa o indirectamente (por transitividad). Además, todos los
+atributos en la salida del flujo o necesarios para unir otros servicios en el
+flujo de trabajo deben ser producidos por el servicio concreto seleccionado.
 
-In addition, suppose each concrete service is annotated with a utility function that aggregates the values of different QoS that characterize the behavior of the service.
-Then, since the space of valid service compositions of an abstract workflow can be very large, it is required a technique able to identify the ones that maximize the utility function  without having to enumerate the whole set of valid compositions.
-In this paper, we propose a Propositional logic-based  approach that takes advantage of the  power of  modern SAT solvers, to efficiently enumerate the service compositions that correspond to valid implementations of a given abstract workflow, as well as the compositions that maximize a given utility function.
+Adicionalmente, supongamos que cada servicio concreto está anotado con una
+función de utilidad que agrega los valores de parámetros diferentes de QoS que
+caracterizan el comportamiento del servicio. Entonces, como el espacio de
+composiciones de servicios válidas de un flujo abstracto puede ser muy grande,
+se requiere una técnica capaz de identificar aquellos que maximicen la función
+de utilidad sin tener que enumerar el conjunto completo de composiciones
+válidas. En este trabajo se propone una aproximación proposicional basada en
+lógica que toma provecho del poder de los \emph{SAT solvers} modernos para
+enumerar eficientemente las composiciones de servicios que corresponden a
+implementaciones válidas de un flujo de trabajo dado, así como las composiciones
+que maximizan una función de utilidad dada.
 
@@
 
-We have shown how the propositional theory used in McdSat can
-be extended to support constants, and how it can be used to 
 compute the best rewritings with respect to an additive
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% PARTE DEL FORMATO
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% TODO PORTADA
+
+% TODO indice
+
+\section{Planteamiento del problema}
+\section{Justificación e importancia}
+\section{Objetivos}
+\section{Metodología propuesta}
+\section{Cronograma de actividades}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Referencias
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\addcontentsline{toc}{section}{Referencias} 
+\bibliographystyle{abbrv}
+\bibliography{ref}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% FIN PARTE DEL FORMATO
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2009-12-13 tesis/propuesta/portada.tex
anadi archivo portada.tex

\begin{titlepage}

\begin{flushleft}
\textsf{
Universidad Simón Bolívar \\
Coordinación de Ciencias de la Computación \\
Maestría en Ciencias de la Computación \\
}
\end{flushleft}

\begin{center}

{\large \bf \textsf{Título tentativo del trabajo (max 100. chars)}}

\end{center}

\vspace{5.0cm}

% Author(s)
\textsf{
Por \\
pepe \\
}

\vspace{\fill}

\vspace{\fill}

\vspace{1.0cm}
\textsf{Sartenejas, diciembre de 2009}

\end{titlepage}

### 2009-12-13 tesis/propuesta/planteamiento.tex
separacion de partes de la propuesta en .texs

--- 2009-12-13 tesis/propuesta/propuesta_iae.tex
@@
-  \documentclass{llncs}
-\usepackage{amsmath}
-\usepackage{txfonts}
-\usepackage[boxed]{algorithm2e}
-\usepackage{amssymb}
-\usepackage[pdftex]{graphicx}
-\usepackage[utf8x]{inputenc}
-\usepackage{enumerate}
-\DeclareGraphicsExtensions{.jpg, .png, .mps, .bmp, .pdf}
-\usepackage{epsfig}
-\usepackage{subfigure}
-\newtheorem{Definition}{Definition} 
-\newtheorem{Example}{Example} 
-\newtheorem{Theorem}{Theorem} 
-\newtheorem{Proof}{Proof} 
-\newcommand{\denselist}{\topsep 0pt \itemsep -4pt}
-\newcommand{\tup}[1]{\langle #1 \rangle}
-\newcommand{\vvec}[1]{\mathbf{#1}}
-\newcommand{\join}{\bowtie}
-\newcommand{\R}{\mathcal{R}}
-\newcommand{\Q}{\mathcal{Q}}
-\newcommand{\body}{{body}}
-\newcommand{\head}{{head}}
-\newcommand{\qrule}{:\!\!-}
-\newcommand{\arc}{\text{arc}}
-\newcommand{\Theory}[1]{T(#1)}
-\newcommand{\mcdsat}{\textsc{McdSat}}
-\newcommand{\minicon}{{MiniCon}}
-\newcommand{\Omit}[1]{}
-\newcommand{\citeX}[1]{\citeauthor{#1}~\citeyear{#1}}
+\section{Planteamiento del problema}
 
-\begin{document}
-\allowdisplaybreaks
-\title{Selección de Servicios Web De Máxima Utilidad Usando Circuitos}
-\author{Daniel Izquierdo}
-
-\institute{Universidad Sim\'on Bol\'{\i}var\\
-                Caracas, Venezuela\\
-                idaniel@ldc.usb.ve} \maketitle
-\begin{abstract}
-En las arquitecturas orientadas a servicios (SOA), los servicios concretos se
-describen en términos de parámetros de funcionalidad y calidad de servicio,
-mientras que las aplicaciones de \emph{software} se especifican como flujos de
-trabajo de servicios abstractos y criterios no-funcionales que establecen la
-funcionalidad deseada y la calidad esperada respectivamente. En tiempo de
-ejecución de un flujo de trabajo, los servicios abstractos se reescriben con los
-servicios concretos que cumplen con las restricciones funcionales y no
-funcionales del flujo de trabajo. Dado que la selección de los servicios es
-hecha al momento y el espacio de servicios concretos puede ser muy grande, se
-requiere eficiencia y escalabilidad. En este trabajo se propone la
-infraestructura {\it McdSat}$^c$ para resolver eficientemente el problema de
-selección de servicios en presencia de un número grande de servicios concretos.
-En primer lugar se utiliza la aproximación \emph{Local As View} para describir
-la funcionalidad de los servicios concretos en términos de vistas de servicios
-abstractos, la calidad como una función de utilidad general que combina los
-distintos valores de calidad que describen cada servicio, y los flujos de
-trabajo como consultas conjuntivas sobre servicios abstractos. Basándose en esta
-representación, se modela el problema de selección de servicios como un problema
-conocido en el área de integración de sistemas conocido como reescritura de
-consultas usando vistas. Codificamos el problema de selección de servicios como
-una teoría proposicional cuyos modelos corresponden a la composición de los
-servicios que implementan el flujo abstracto; se explotan relaciones conocidas
-entre teorías d-DNNF y circuitos aritméticos para proveer una solución eficiente
-y escalable. Se reporta sobre el rendimiento de la aproximación propuesta y se
-observa que  {\it McdSat}$^c$  escala a instancias grandes del problema de
-selección de servicios.
-
-\end{abstract}
-
-\section{Introducción}
 Bajo la sombra de la Web Semántica y con el apoyo de arquitecturas orientadas a
@@
 disponibles. Para evaluar el flujo propuesto, se instancian los servicios
-abstractos con concretos de manera que se satisfagan las restricciones sobre
+abstractos con concretos de manera que se satisfagan las restricciones sobre los
 parámetros QoS. Sin embargo, el rendimiento de estas tareas es impactado por el
@@
 no-funcionales que describen una aplicación diseñada. Se llama 
-{\it MCDSat}$^c$ a la solución propuesta.
+\mcdsatc a la solución propuesta.
 
-En {\it MCDSat}$^c$, las funcionalidades de servicios concretos se describen
+En \mcdsatc, las funcionalidades de servicios concretos se describen
 como vistas sobre servicios abstractos y los flujos de trabajo se representan
@@
 que definen flujos de trabajo y servicios concretos en el catálogo de
-{\it MCDSat}$^c$ son creadas de manera que las restricciones de entrada y salida
+\mcdsatc son creadas de manera que las restricciones de entrada y salida
 de los servicios abstractos combinados sean satisfechas. Para representar las
@@
 real que representa un valor general para las medidas de QoS del servicio en
-particular. Finalmente, {\it MCDSat}$^c$ modela el problema de selección de
+particular. Finalmente, \mcdsatc modela el problema de selección de
 servicios web como el problema de reescritura de consultas usando vistas
@@
 vistas~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-En este trabajo se propone la infraestructura {\it MCDSat}$^c$ que extiende la
+En este trabajo se propone la infraestructura \mcdsatc que extiende la
 aproximación propuesta en  ~\cite{arvelo:aaai06}, con la capacidad de trabajar
@@
 
-{\it MCDSat}$^c$  traduce una instancia del problema de selección de servicios
+\mcdsatc  traduce una instancia del problema de selección de servicios
 en el problema de enumerar los modelos de una teoría proposicional que codifica
@@
 función global de utilidad se obtienen en tiempo lineal usando técnicas modernas
-de SAT~\cite{darwiche:dnnf}. Se estudia empíricamente la calidad de {\it
-MCDSat}$^c$ sobre una variedad de \emph{benchmarks} y se observa que la técnica
+de SAT~\cite{darwiche:dnnf}. Se estudia empíricamente la calidad de \mcdsatc
+sobre una variedad de \emph{benchmarks} y se observa que la técnica
 propuesta provee una solución eficiente al problema de selección de servicios
@@
 
-This paper is comprised of six additional sections.
-The next section motivates our approach by using an example, and  we summarize existing related work in section 3.
-Section 4 describes the {\it MCDSat}$^c$ system architecture, and  the experimental study is reported in section 5.
+This paper is comprised of six additional sections. The next section motivates
+our approach by using an example, and  we summarize existing related work in
+section 3. Section 4 describes the \mcdsatc system architecture, and  the experimental study is reported in section 5. 
 In section 6 we present the formalization of the web service selection problem as a Propositional Logic theory, and finally, section 7 outlines our conclusions and future work.
@@
 La composición {\it wi2} no es válida porque asocia la variable del flujo de
-trabajo {\it y2} a las constantes {\it "CCS"} y {\it "WAS"} que denotan ciudades
+trabajo {\it y2} a las constantes {\it ``CCS''} y {\it ``WAS''} que denotan
+ciudades
 distintas. Por otra parte, la composición {\it wi3} no implementa el flujo de
@@
 que maximizan una función de utilidad dada.
-
-\section{Related Work}
-In this section we summarize existing approaches that provide solutions to the problems of service selection, knowledge representation and query rewriting.
-
-\begin{description}
-\item[Service Selection Solutions] \mbox{}\\
-The problem of selecting the services that implement an abstract workflow and best meet the QoS-based criteria  is known as the QoS-aware service selection or composition problem, which has been shown to be NP-hard~\cite{Hiroshi2008}.
-This problem is a combinatorial optimization problem and several heuristics have been proposed to find a relatively good solution in a reasonably short period of time.
-A  distance metric-based heuristic  to drive a backward search algorithm  is proposed~\cite{rahmani08}; this metric induces an order of the services in a way that sink nodes are unlikely to be visited.
-In ~\cite{berardi05,berardi08,berardi06}, services and workflows are described in terms of deterministic finite state machines that are encoded as a Description Logic theory whose models correspond to solutions of the problem;  although reasoning methods for Description Logics formalisms could be exploited, scalability or performance of the proposed solution has not been reported.
-~\cite{myoung08} propose a constraint-based approach that encodes the non-functional permissible values as a set of constraints whose violation needs to be minimized; to traverse the space of possibly optimal solutions, a hybrid algorithm that combines the tabu search and simulating annealing meta-heuristics is implemented; experimental results show that the proposed solution is able to scale up to a large number of services and abstract processes.
-In ~\cite{cardellini07} the QoS-aware service composition problem is encoded as a Linear Programming problem providing a scalable solution to the problem.
-In ~\cite{Hiroshi2008}  this problem  is defined as a multi-objective optimization problem where the different QoS parameters are considered equally important and there is not an aggregated function to combine all of them; a genetic-based algorithm is proposed to identify a set of non-dominated service compositions that best meet all the QoS parameters.
-~\cite{alrifaiR09} propose a two-fold solution that uses a hybrid integer programming algorithm to find the decomposition of global QoS into local constraint, and then,  selects the services that best meet the local constraints.
-Recently,  two new planning-based approaches have been proposed~\cite{kuterG09,sohrabiM09}.
-~\cite{kuterG09} extend the SHOP2 planning algorithm to select the trustworthy composition of services that implement a given OWL-S process model, while ~\cite{sohrabiM09}  propose a HTN planning-based solution where user preference metrics and domain regulations are used to guide the planner into the space of relevant compositions.
-Finally, ~\cite{lecue09} proposes a genetic-based algorithm to identify the composition of services that best meet the quality criteria for a set of QoS parameters.
-
-Although these solutions are able to efficiently solve the optimization problem and scale up to a large number of abstract processes, none of them are tailored to semantically describe services in terms of the abstract process, or use these descriptions to identify the services that  implement a given workflow and best meet user non-functional criteria.
-
-\item[Knowledge Compilation Languages] \mbox{}\\
-Knowledge compilation is the area in AI concerned with the
-problem of mapping logical theories into suitable fragments
-that make certain desired operations tractable
-\cite{cadoli:compilation}. Different compilation languages have been defined, for instance,  
-Ordered Binary Decision Diagrams (OBDDs)~\cite{bryant:obdd},  Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation Normal Form (DNNF)  \cite{darwiche:dnnf}.
-In this work we make use of the properties of the deterministic DNNFs  (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable and efficient solution to the service selection problem.
-
-A Negation Normal Form (NNF) theory is  constructed from literals using only conjunctions
-and disjunctions \cite{barwise:handbook}, and it can be represented as a directed acyclic graph
-in which the leaves are labeled with literals and the
-internal nodes are labeled with $\land$ and $\lor$;
-see Fig.~\ref{fig:dnnf} for an example. An NNF is said to be decomposable (DNNF) \cite{darwiche:dnnf}
-if for each conjunction $\bigwedge_i\phi_i$, the set of
-variables in each conjunct are pairwise disjoint;
-i.e,. $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i<j$.
-A DNNF supports a number of operations in
-polynomial time in the size of its DAG. For example,
-we can test whether a DNNF is satisfiable by a single
-bottom-up pass over its DAG in linear time. A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each disjunction $\bigvee_i\phi_i$, the disjuncts
-are pairwise logically contradictory; i.e.,
-$\phi_i\Rightarrow\neg\phi_j$ for $i<j$.
-The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable
-and deterministic. A d-DNNF supports the model counting and enumeration in polynomial time
-in the size of its DAG.
-{\it MCDSat}$^c$  exploits the properties of a d-DNNF theory, to efficiently provide a solution to the problem of enumerating the service compositions that correspond to models of the  theory, i.e., the service compositions that implement the abstract  workflow and best meet the utility function.
-DECIR ALGO SOBRE EL ALGORITMO QUE PRODUCE LOS MEJORES.
-
-\begin{figure}
-\centering
-%\includegraphics[width=8cm]{odd}
-\caption{A decomposable and deterministic NNF.}
-\label{fig:dnnf}
-\end{figure}
-
-\item[Query Rewriting Solutions] \mbox{}\\
-A number of algorithms have been developed to find the rewritings of a given query; the most prominent being the bucket algorithm~\cite{levy:bucket},  the inverse rules algorithm ~\cite{duschka:answer,Qian96}, the minicon algorithm ~\cite{pottinger:minicon}, and the {\it MCDSat} ~\cite{arvelo:aaai06}.
-Generally, query rewriting algorithms work in two phases: first, they identify the views that rewrite at least one  subgoal of the query; second they combine the selected views to produce a rewriting.
-The main difference between existing approaches, is the criteria used to choose the relevant views and reduce the space of non-useful rewritings.
-
-The bucket algorithm reduces the number of possibilities just considering each subgoal in the query in isolation, and selecting the views that are able to  produce at least the attributes projected by the query.
-Since, attributes involved in query joins are not verified, a large number of rewritings comprised of  Cartesian products may  be generated.
-The Inverse Rules algorithm constructs a set of rules that invert the view definition and establish how to compute tuples for the database relations from tuples of the views.
-Similarly, to the bucket algorith it can produce a large number non-useful of rewritings.
-
-The MiniCon algorithm  overcomes the limitations of the previous algorithms by identifying only views that rewrite a set of the query goals and that  can be combined with the rest of the query subgoals.
-The key idea is to identify the mappings between  the variables in each subgoal of the query to the variables  in one or more subgoals of the views, in a way that, join variables in the query are mapped to join variables in the body of a view or to the distinguished variables of the view.
-Mappings between variables and subgoals are represented in MiniCon Descriptions (MCD's)\cite{pottinger:minicon}.
-
-Finally,  the
-{\it MCDSat} algorithm is able to identify the query rewritings of a query by translating the problem of rewriting
-into the problem of enumerating the models of a propositional
-theory  $\Theory{\Q}$ whose models are in correspondence
-with the rewritings of the query. 
-The {\it MCDSat} algorithm exploits the properties of d-DNNFs to efficiently 
-compute the MCDs associated with theory $\Theory{\Q}$.
-The  {\it MCDSat} algorithm has demonstrated to scale better than the MiniCon
-algorithm over a large number of benchmarks often showing performance
-improvements of several orders of magnitude. However, the McdSat algorithm was not desgined for rewriting problems involving explicit constants,
-nor to compute the best rewritings with respect to a given utility function or cost model, and this paper we propose a new encoding  that overcomes these limitations.
-
-\end{description}
-
-\section{The McdSat$^c$ Architecture}
-
-In order to do so, the theory $\Theory{\Q}$ is transformed into a d-DNNF
-theory $\Delta(\Q)$ with a CNF to d-DNNF compiler.
-The compiler's algorithm is similar to the DPLL algorithm
-for SAT but enhanced with advanced techniques such as clause
-learning and caching.
-\begin{figure}
-\centering
-%COMENTADO\includegraphics[width=10cm]{mdcsatC.pdf}
-\caption{The McdSat$^c$ Architecture}
-\label{fig:dnnf}
-\end{figure}
-
-\section{Experimental Results}
-We have conducted an empirical analysis  on the benefits of the techniques implemented in the   McdSat$^c$ system.
-
-\begin{description}
-\item[Dataset and Query Benchmark:]  we conducted our experiments over a benchmark that includes XX queries and a set of  YY views.
-Queries and views are starts and chains, and they have between ZZ and fTT constants; queries have QQ sub-subgoals.
-\item[Evaluation Metrics:] we report on runtime performance which corresponds to the  {\it user time} produced by the {\it time} command of the Unix operation system. 
-\end{description}
-
-\begin{figure}
-\centering
-%COMENTADO\includegraphics[width=8cm]{plot1.pdf}
-\caption{Compilation Time Benchmark I}
-\label{fig:plot1}
-\end{figure}
-
-\section{Formalization of the Service Selection Problem}
-
-We consider databases of the form $D=\tup{P,T}$ where
-$P$ is a set of predicates and $T=\{T_p\}_{p\in P}$ is a collection
-of tables that represents the predicates in extensional form.
-A conjunctive query $Q$ over $P$ is of the form 
-\begin{alignat*}{1}
-Q(\vvec{x})\ \qrule\ \  p_1(\vvec{x}_1),\, p_1(\vvec{x}_2),\, \ldots,\, p_m(\vvec{x}_m)\,,
-\end{alignat*}
-where $p_i\in P$, $\vvec{x}$ is a vector of variables, and each
-$\vvec{x}_i$ is a vector of variables and constants.
-The result of $Q$ over $D$, denoted as $Q(D)$, is the table with
-$|\vvec{x}|$ columns that result of the projection of the relational
-join $\join\!\!\{T_{p_i}\}_{i=1}^m$ over $\vvec{x}$.
-The atoms in the body of $Q$ are called the (sub)goals of $Q$.
-
-A view $V$ over $D$ is a query over $P$. In the context of the service selection problem, 
-the database $D$ is an idealized description of 
-the output produced abstract workflow  implemented by multiple concrete services described
-as views.
-Given a database $D$, a query $Q$ and a collection of 
-views $E=\tup{\{V_i\}_i,\{E_i\}_i}$, we are required to find
-all the tuples in $Q(D)$ obtainable from the views in $E$.
-That is, we need to find all the \emph{compositions} of the form
-\begin{alignat*}{1}
-R(\vvec{x})\ \qrule\ \ V_{i_1}(\vvec{x}_1),\, V_{i_2}(\vvec{x}_2),\, \ldots,\, V_{i_n}(\vvec{x}_n)
-\end{alignat*}
-such that $R(E) \subseteq Q(D)$.
-A service selection problem (SSP) is a tuple $\tup{P,Q,\{V_i\}}$ where $P$
-is a set of predicates that represent abstract services, $Q$ is a query over $P$ and $\{V_i\}$ is a collection of views that define the concrete services in terms of  abstract services.
-We assume \emph{safe} problems in the sense that all
-variables mentioned in the head of the query (resp.\ in the head of each view)
-appear in the body of the query (resp.,\ in the body of each view); also the input and output restrictions of the abstract services used in the query are satisfied.
-Further, we only deal with SSPs with no arithmetic predicates inside the
-query or views.
-A composition $C$ is \emph{valid} if for all databases $D=\tup{P,T}$
-and extensions $\{E_i\}$, $R(E) \subseteq Q(D)$.
-A collection $\R$ of valid compositions is a solution if
-for all databases $D=\tup{P,T}$ and extensions $\{E_i\}$, there
-is no other $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(D)$.
-We are interested in finding a composition $\R$.
-
-\subsection{Logical Theories}
-
-In \cite{arvelo:aaai06}, we showed that a rewriting
-for a query $Q$ with $m$ goals can be obtained by enumerating the
-models of a logical theory
-$\Delta=\Delta_{rew}\cup\Delta_{mcd}^1\cup\cdots\Delta_{mcd}^m$
-where $\Delta_{rew}$ specified how to combine $m$ independent
-copies of MCD theories $\Delta_{mcd}$ that cover all goals in $Q$.
-Each $\Delta^i_{mcd}$ is a copy of the theory $\Delta_{mcd}$
-in which each literal $p$ is tagged as $p^i$.
-The theory $\Delta_{mcd}$ consists of different groups of 
-clauses that guarantees that its models are in correspondence
-with the MCDs, while the theory $\Delta_{rew}$ contains additional
-clauses to guarantee a sound and complete composition of the MCDs.
-The reader is referred to \cite{arvelo:aaai06} for a comprehensive
-description of the propositional theory.
-
-Let us describe the difficulties that arise when constants
-are presents in a SSP with the following example:
-\begin{alignat*}{1}
-Q(x,z)\ &\qrule\ \ p_1(x,y),\, p_2(y,z)\,, \\
-V_1(x_1)\ &\qrule\ \ p_1(x_1,A)\,, \\
-V_2(x_2)\ &\qrule\ \ p_2(B,x_2)\,, \\
-V_3(x_3,y_3)\ &\qrule\ \ p_2(x_3,y_3)
-%V_4(x_4,y_4)\ &\qrule\ \ p_1(x_1,A),\, p_2(B,y_4)\,,
-\end{alignat*}
-where $Q$ is the query and $\{V_1,V_2,V_3\}$ are the 
-views. In this case, the only rewriting is 
-$R(x,z)\qrule V_1(x),V_3(A,z)$ because the candidate 
-$R(x,z)\qrule V_1(x),V_2(z)$ is not valid as it maps
-$y$ into constants $A$ and $B$ that denote different
-objects.
-
-The main problem when handling constants is to
-be sure that different constants are not mapped into
-each other either directly or indirectly (via transitivity).
-
-\subsection{Utility Functions}
-
-We assume a simple additive utility function in which each view
-$V_i$ is associated with a utility measure $c(V_i)$, and the overall utility value of
-a composition is the sum of the utility values for the views in it.
-An optimal or best service composition is one with maximum utility value,
-and the optimal utility value of a SSP is the utility value of a best
-service composition. A SSP has always a well-defined optimal
-utility value (if there are no compositions, its utility is $0$),
-but it may have multiple best service compositions.
-The service composition problem with utility consists in finding
-all the compositions of maximum utility value.
-
-%The cost model is simple yet expressive; e.g., if all
-%costs are unit, then best rewritings are those with
-%minimum number of views, yet in cases where views
-%represent different sources, it is natural
-%to associate them with different costs.
-
-\subsection{Extended Theories}
-
-The theory $\Delta_{mcd}$ makes use of propositions $t_{x,y}$
-to denote that the variable/constant $x$ in the query is mapped
-into the variable/constant $y$ in the view, and propositions $v_i$
-to indicate that the MCD uses view $V_i$.
-We obtain a sound and complete theory for SSPs with constants if
-$\Delta_{mcd}$ is extended with the clauses:
-
-\begin{enumerate}[C1.]\denselist
-\item (Inconsistent-1): $t_{x,A} \Rightarrow \neg t_{x,B}$,
-\item (Inconsistent-2): $t_{A,x} \Rightarrow \neg t_{B,x}$,
-\item (Inconsistent-3): $\neg t_{A,B}$,
-\item (Transitivity-1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$,
-\item (Transitivity-2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
-\end{enumerate}
-Clauses C1--C3 prune candidate service compositions in which one constant
-is directly mapped into a different one, and the last two implement
-a restricted propagation of transitivity.
-Similarly, $\Delta_{rew}$ must be extended with the clauses:
-\begin{enumerate}[C1.]\denselist
-\item[C6.] (Inconsistent-4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
-\end{enumerate}
-~\cite{RajaramanSU95} showed that for queries without negation or arithmetic comparisons,  but with constants, and $m$ goals and $q$ variables, it is enough
-to consider service compositions of length at most  $m$ plus  $q$ subgoals ~\cite{Ullman00}.
-
-Given a SSP problem $\tup{Q,\{V_i\}}$ possibly with constant 
-symbols, we construct logical theories whose models are in
-correspondence with the service compositions. These theories are then
-compiled into d-DNNF from which all rewritings are extracted
-efficiently.
-Likewise, best service compositions can be computed by performing model
-enumeration twice: the first pass computes the optimal utility function,
-and the second filters out suboptimal service compositions.
-However, there is a better way, which we will describe in the next section.
-
-\subsection{Maximum-Utility Function}
-
-Darwiche \& Marquis \cite{darwiche:weighted} show how to compute the
-rank $r^*(\Delta)$ of a propositional theory $\Delta$
-efficiently when $r$ is a literal ranking function and
-$\Delta$ is in d-DNNF. A literal ranking function ranks
-the models of the theory in terms of the rank of the
-literals $l$ that are true in the model, and ranks the
-theory as the best rank of its models:
-\[ r(\omega) = \sum_{\omega\vDash l} r(l),\quad r^*(\Delta) = \max_{\omega\vDash\Delta} r(\omega)\,. \]
-Furthermore, not only $r^*(\Delta)$ can be computed
-efficiently from the d-DNNF but also the models that
-have such rank.
-The procedure for computing the rank converts the d-DNNF
-into a circuit in which the literals are replaced by their
-ranks, the `or' nodes by `max', and the `and' nodes by `sum'.
-The evaluation of the circuit computes the rank of $\Delta$.
-
-We thus obtain a simple method for computing the best
-rewritings when the literal ranking function $r_c$ induced
-by the cost model $c$ is used; $r_c$ is defined by
-$r_c(l)=c(V_i)$ if $l=v^t_i$ for some $t$, and $r_c(l)=0$
-otherwise.
-
-\section{Conclusions and Future Work}
-
-compute the best rewritings with respect to an additive
-cost model. The cost model is simple yet expressive.
-The computation of the best models relies on an arithmetic
-circuit that is obtained from the d-DNNF of the propositional
-theory.
-
-\bibliographystyle{abbrv}
-\bibliography{ref}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% PARTE DEL FORMATO
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% TODO PORTADA
-
-% TODO indice
-
-\section{Planteamiento del problema}
-\section{Justificación e importancia}
-\section{Objetivos}
-\section{Metodología propuesta}
-\section{Cronograma de actividades}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Referencias
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\addcontentsline{toc}{section}{Referencias} 
-\bibliographystyle{abbrv}
-\bibliography{ref}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% FIN PARTE DEL FORMATO
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\end{document}

### 2009-12-13 tesis/propuesta/propuesta_iae.tex
separacion de partes de la propuesta en .texs

--- 2009-12-11 tesis/propuesta/propuesta_iae.tex
@@
-  \documentclass{llncs}
+\documentclass{article}
+
+\usepackage[spanish]{babel}
 \usepackage{amsmath}
@@
 
+\newcommand{\comment}[1]{}
+
+\newcommand{\mcdsatc}{{\it MCDSat}$^c$}
+
 \begin{document}
+
+\comment{
 \allowdisplaybreaks
-\title{Selecting the Maximum-Utility  Web Services  Using Circuits}
-\author{Daniel Izquierdo \inst{1}  and Blai Bonet \inst{1}  and Mar\'{\i}a-Esther Vidal \inst{1}}
+\title{Selección de Servicios Web De Máxima Utilidad Usando Circuitos}
+\author{Daniel Izquierdo}
 
-\institute{Universidad Sim\'on Bol\'{\i}var\\
+{Universidad Sim\'on Bol\'{\i}var\\
                 Caracas, Venezuela\\
-                \{idaniel,bonet,mvidal\}ldc.usb.ve} \maketitle
-\begin{abstract}
-In Service Oriented Architectures (SOA), concrete services are described in terms of functionality   and QoS parameters, while software applications are specified as workflows of abstract services and non-functional criteria that establish the  application desirable functionality and  expected quality, respectively.
-During workflow execution time,  abstract services  are rewritten
-with the concrete services  that meet the workflow  functional and non-functional restrictions.
-Since, service selection is done on-the-flight and the space of concrete services may be very large, efficiency and scalability are required.
-In this paper we propose the {\it McdSat}$^c$ framework to efficiently solve the problem of service selection in presence of a large number of concrete services.
-First, we use the Local As View approach to describe the functionality  of the concrete services in terms of views of abstract services, the quality as an overall utility function that combines the different QoS values that describe the service, and  workflows as  conjunctive queries on abstract services.
-Based on this representation, we
-model the problem of service selection as a known problem in the area of integration systems referred as query rewriting using views.
-We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and maximizes the overall utility function; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.
-We report on the performance of our proposed approach, and we observe that  {\it McdSat}$^c$  efficiently scales up to large instances of the service selection problem.
-
-\end{abstract}                
-\section{Introduction}
-Under the umbrella of the Semantic Web and supported by Service Oriented architectures, the number of Web data sources and services has exploded during the last few years.
-For example,  currently, the molecular biology databases
-collection includes 1,170 databases~\cite{Galperin09}, that is
-95 more than the previous year~\cite{Galperin2008} and 110 more than two years ago~\cite{Galperin2007};  tools
-and services as well as the number of instances published by these resources, follow a similar progression~\cite{Benson07}. Thanks
-to this wealth, users rely more on various
-digital tasks such as data retrieval from public data sources and
-data analysis with Web tools or services organized in
-complex workflows.
-
-In order to design an application, users usually specify the functionality in terms of a workflow of abstract services, while the desirable quality  is defined by restricting the values of the QoS parameters that describe the concrete available services.
-To evaluate the proposed workflow,  abstract services are instantiated on concrete services in a way restrictions on the QoS parameters are satisfied.
-However, the performance of this tasks is impacted by the number of available sources,  techniques enable to efficiently scale up to a large number of services are required.
-
-In this paper we aim to provide a solution to this problem and provide techniques that efficiently traverse the space of available services and select combinations that meet the functional and non-functional requirements that describe a designed application.
-We name our proposed solution {\it MCDSat}$^c$.
-
-In {\it MCDSat}$^c$, concrete service functionalities are decribed as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.
-In addition, rules that define workflows and  concrete services in the {\it MCDSat}$^c$ catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied.
-To represent the QoS measures, each concrete service description is annotated with a real number that represents an overall value of the QoS measures of this service.
-Finally, {\it MCDSat}$^c$ models the web service selection problem as the problem of query rewriting using views \cite{halevy:survey}.
-This problem is important in the context of data integration
-\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance \cite{AfratiLU07,levy:bucket}, and several approaches have been defined to efficiently scale up to a large number of views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-In this paper we propose the {\it MCDSat}$^c$ framework which extends the approach proposed in ~\cite{arvelo:aaai06},  with the capability of dealing with constants and identifying the composition of services that maximizes a given utility function.
-
-{\it MCDSat}$^c$  translates an instance of the service selection problem into the problem of enumerating the models of a propositional theory that encodes information about the workflow, the concrete services and the conditions that need to be satisfied in order to generate a composition that maximizes the utility function.
-This theory is represented as a CNF formula which is compiled into a deterministic
-and decomposable negation normal form d-DNNF representation, from which the models that maximizes the overall utility function are obtained in linear time using modern SAT techniques~\cite{darwiche:dnnf}.
-We empirically studied the quality of the {\it MCDSat}$^c$ on variety of benchmarks and we observed  that the proposed technique provides an efficient solution to the service selection problem which is able to scale up to large workflows as well to a large number of views.
-
- This paper is comprised of six additional sections.
-The next section motivates our approach by using an example, and  we summarize existing related work in section 3.
-Section 4 describes the {\it MCDSat}$^c$ system architecture, and  the experimental study is reported in section 5.
- In section 6 we present the formalization of the web service selection problem as a Propositional Logic theory, and finally, section 7 outlines our conclusions and future work.
-
-\section{Motivating Example}
-Consider the abstract and concrete services in Table ~\ref{table:services}  which represent applications in the flight domain.
-
-\begin{table}
-\begin{tabular}{| l| l |}
-\hline
-Abstract Service & Description \\
-\hline
-$fight(x1,x2)$ & relates cities x1 and x2 if there is a flight between them\\
-\hline
-$uscity(x1)$ &  determines if a city x1 is a US city\\
-\hline \hline
-Concrete  Service & Description \\
-\hline
-$national(x1,y1)$&  returns  two US cities that are connected by  direct flights \\
-\hline
-$oneway(x1,y1)$ & returns two cites that are connected by one way direct flights \\
-\hline
-$onestop(x1,y1)$ &  returns  two cites that are connected by one-stop flights \\
-\hline
-$ flightToCCS(x1,"CCS")$ & returns a city that is connected to  Caracas by a direct flight\\
- \hline
-$ onestopToCCS(x1,y1)$ & returns two cites x1 and y1, such that there is  one-stop flight\\ 
- & from city x1 to Caracas stopping at city y1.\\
- \hline
- $fromWAS("WAS",y1)$ &  returns a city y1 to which flights from Washington arrive. \\
- \hline
-\end{tabular}
-\label{table:services}
-\caption{Abstract and Concrete Services}
-\end{table}
-
-Using the  Local As View (LAV) approach~\cite{AmbiteISWC09}, concrete services in Table 1 are semantically described as views on the abstract services {\it flight} and {\it uscity} as follows:
-
-\begin{alignat*}{1}
-national(x1,y1)\ &\qrule\ flight(x1,y1),\,  uscity(x1),\,  uscity(y1)\,. \\
-oneway(x2,y2)\ &\qrule\ flight(x2,y2)\,. \\
-onestop(x3,z3)\ &\qrule\ flight(x3,y3),\, flight(y3,z3)\,. \\
-flightToCCS(x2,"CCS")\ &\qrule\ flight(x2,"CCS")\,. \\
-onestopToCCS(x3,y3)\ &\qrule\ flight(x3,y3),\, flight(y3,"CCS")\,. \\
-fromWAS("WAS",y3)\ &\qrule\ flight("WAS",x3)\,. \\
-\end{alignat*}
-
-Now suppose a user is interested in implementing a workflow able to retrieve the one-stop round trip flights from US cities to any city in the world, such that flights can stop at any city. 
-Consider the following conjunctive query that represents the workflow that defines this request in terms of abstract services.
-This workflow was defined in way the input parameters of each abstract service are bound by  constants in the workflow, or by attributes projected out by  previous abstract services.
-
-\begin{alignat*}{1}
-w(x,y1,y2,y3)\ &\qrule\ flight(x,y1),\, flight(y1,y2),\,  flight(y2,y3),\,  flight(y3,x),\,  \\
-                   \ & uscity(x)\,. \\
-\end{alignat*}
-
-Implementations of this workflow correspond to compositions of the concrete services where each concrete service can implement a subset of abstract services of the workflow, but each abstract service can be implemented by exactly one concrete service.
-The following composition of concrete services corresponds to one of the XX implementations of the workflow:
-
-\begin{alignat*}{1}
-wi1(x,y1,"CCS",y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\,  oneway("CCS",y3),\,  \\
-\ & national(y3,x)\,.  \\
-\end{alignat*}
-
-However, note that the following compositions are not valid:
-
-\begin{alignat*}{1}
-wi2(x,y1,"CCS"y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\, \\
-\ & fromWAS("WAS",y3),\,  national(y3,x)\,. \\
-wi3(x,y1,y2,y3)\ &\qrule\ national(x1,y1),\, onestop(x1,y2),\,  oneway(y2,y3),\,  \\
-\ & national(y3,x)\,. \\
-\end{alignat*}
-
-The composition {\it wi2} is not valid because  it maps the workflow variable
-{\it y2} into constants {\it "CCS"} and {\it "WAS"} which denote different cities.
-On the other hand, the composition {\it wi3} does not implement the workflow because the concrete service  {\it onestop} does not receive as input or produce as output, the city where the flight stops (represented by the variable {\it y1}); thus, it is not possible to ensure that the city where this flight stops and the end city  retrieved by the service {\it national}, are the same.
-
-To avoid producing not valid workflow implementations, it is required to handling constants in a way that different constants are not mapped into
-each other either directly or indirectly (via transitivity).
-In addition,  all the attributes in the workflow output  or needed to join other services in the workflow, have to be produced by the selected concrete service.
-
-In addition, suppose each concrete service is annotated with a utility function that aggregates the values of different QoS that characterize the behavior of the service.
-Then, since the space of valid service compositions of an abstract workflow can be very large, it is required a technique able to identify the ones that maximize the utility function  without having to enumerate the whole set of valid compositions.
-In this paper, we propose a Propositional logic-based  approach that takes advantage of the  power of  modern SAT solvers, to efficiently enumerate the service compositions that correspond to valid implementations of a given abstract workflow, as well as the compositions that maximize a given utility function.
+                idaniel@ldc.usb.ve} \maketitle
 
@@
 theory.
+}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\input{portada}
+
+\tableofcontents
+\newpage
+
+\input{planteamiento}
+\input{justificacion}
+\input{objetivos}
+\input{metodologia}
+\input{cronograma}
+
+\addcontentsline{toc}{section}{Referencias} 
 \bibliographystyle{abbrv}

### 2009-12-14 tesis/propuesta/portada.tex
portada con abstrac lista

@@
 \begin{flushleft}
-\textsf{
+\includegraphics[scale=0.3]{logo_usb.png} \\
 Universidad Simón Bolívar \\
@@
 Maestría en Ciencias de la Computación \\
-}
 \end{flushleft}
 
+\vspace{0.8cm}
+
 \begin{center}
-
-{\large \bf \textsf{Título tentativo del trabajo (max 100. chars)}}
-
+{\large \bf \textsf{Selección de Servicios Web De Máxima Utilidad Usando Circuitos}}
 \end{center}
 
-\vspace{5.0cm}
+\vspace{0.2cm}
 
-% Author(s)
-\textsf{
-Por \\
-pepe \\
-}
+\begin{abstract}
+En las arquitecturas orientadas a servicios (SOA), los servicios concretos se
+describen en términos de parámetros de funcionalidad y calidad de servicio,
+mientras que las aplicaciones de \emph{software} se especifican como flujos de
+trabajo de servicios abstractos y criterios no-funcionales que establecen la
+funcionalidad deseada y la calidad esperada respectivamente. En tiempo de
+ejecución de un flujo de trabajo, los servicios abstractos se reescriben con los
+servicios concretos que cumplen con las restricciones funcionales y no
+funcionales del flujo de trabajo. Dado que la selección de los servicios es
+hecha al momento y el espacio de servicios concretos puede ser muy grande, se
+requiere eficiencia y escalabilidad. En este trabajo se propone la
+infraestructura \mcdsatc para resolver eficientemente el problema de
+selección de servicios en presencia de un número grande de servicios concretos.
+En primer lugar se utiliza la aproximación \emph{Local As View} para describir
+la funcionalidad de los servicios concretos en términos de vistas de servicios
+abstractos, la calidad como una función de utilidad general que combina los
+distintos valores de calidad que describen cada servicio, y los flujos de
+trabajo como consultas conjuntivas sobre servicios abstractos. Basándose en esta
+representación, se modela el problema de selección de servicios como un problema
+conocido en el área de integración de sistemas conocido como reescritura de
+consultas usando vistas. Se codifica el problema de selección de servicios como
+una teoría proposicional cuyos modelos corresponden a la composición de los
+servicios que implementan el flujo abstracto; se explotan relaciones conocidas
+entre teorías d-DNNF y circuitos aritméticos para proveer una solución eficiente
+y escalable. Se reporta sobre el rendimiento de la aproximación propuesta y se
+observa que \mcdsatc escala a instancias grandes del problema de
+selección de servicios.
+\end{abstract}
 
@@
 
-\vspace{\fill}
-
-\vspace{1.0cm}
-\textsf{Sartenejas, diciembre de 2009}
+\begin{minipage}{0.3\textwidth}
+\begin{flushleft}
+\emph{Estudiante:} \\
+Daniel \textsc{Izquierdo} \\
+Carnet 08-86809 \\
+\end{flushleft}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+\begin{center}
+Fecha estimada de culminación: \\
+Marzo de 2010
+\end{center}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+\begin{flushright}
+\emph{Asesor:} \\
+Prof. María Esther \textsc{Vidal}
+\end{flushright}
+\end{minipage}
 

### 2009-12-14 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py
fix para cuando aparecen constantes en la transitividad

@@ -666,7 +666,7 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query, ltaux, lz):
                         varT = varsT.get((int(a),int(z)))
 
                         if varT:
-                            providers[(a,z)].update(cs)
+                            providers.setdefault((a,z), set()).update(cs)
 
                             for (soqn, sovn) in providers[(a,z)]:
                                 varZ = varsZ.get((soqn,sovn,numVista))
@@ -705,7 +705,7 @@ def clausulas_d4d5(q, vistas, variables_query, constantes_query, ltaux, lz):
                         varT = varsT.get((int(z),int(a)))
 
                         if varT:
-                            providers[(z,a)].update(cs)
+                            providers.setdefault((z,a), set()).update(cs)
 
                             for (soqn, sovn) in providers[(z,a)]:
                                 varZ = varsZ.get((soqn,sovn,numVista))

### 2009-12-14 tesis/src/imcdsat/mcdsat/Parser/CQparser.py
presentacion de modelos de RW y MCD bonitos

@@ -75,6 +75,22 @@ def getVariableName(x):
 
     return x
 
+def getRealConstantName(n):
+    global renameArguments
+
+    if renameArguments:
+        return str(int(n)-1000000000)
+
+    return n
+
+def getRealVariableName(x):
+    global renameArguments
+
+    if renameArguments:
+        return 'X'+str(int(x) % 10000)
+
+    return x
+
 def crearSO(pred, lista):
     ord = []
     arg = {}

### 2009-12-14 tesis/src/imcdsat/mcdsat/Traductor/GenerarReescrituras.py

@@ -224,7 +224,6 @@ def generarReescMCD(numeros, stdin, vistas):
             mod = obtModelo(numeros, model, n)
             mcd = crearMCD(mod, vistas)
             if mcd != None:
-                #print mod
                 print mcd
 
 

### 2009-12-14 tesis/src/imcdsat/mcdsat/Traductor/MCD.py

@@ -1,3 +1,6 @@
+from CQ.CQ import es_const
+from Parser.CQparser import getRealConstantName, getRealVariableName
+
 class MCD:
     ## hc: head homomorfismo (dict)
     ## phic: mapping vars(Q) en vars(V) (dict)
@@ -9,11 +12,33 @@ class MCD:
         self.vistaH = vista.map_variables(self.hc)
         self.gc = gc
         self.clave = "V:"  + self.vista.cabeza.predicado + " tau:" + str(self.phictodo) + " phic:" + str(self.phic) + " hc:" + str(self.hc) + " gc:" + str(list(self.gc))
-        self.string = "V:"  + self.vista.cabeza.predicado + " tau:" + str(self.phictodo) + " gc:" + str(list(self.gc))
         self.ec, self.phi_1 = self.calcularEc()
+
+        self.string = None
 
     # OJO CON ESTO PQ EL STRING DE UNA TABLA DE HASH NO ES NECESARIAMENTE EL MISMO SIEMPRE
     def __str__(self):
+        if not self.string:
+            display_phictodo = {}
+
+            for a in self.phictodo:
+                if es_const(a):
+                    na = getRealConstantName(a)
+                else:
+                    na = getRealVariableName(a)
+
+                display_phictodo[na] = []
+
+                for b in self.phictodo[a]:
+                    if es_const(b):
+                        nb = getRealConstantName(b)
+                    else:
+                        nb = getRealVariableName(b)
+
+                    display_phictodo[na].append(nb)
+
+            self.string = "V:"  + self.vista.cabeza.predicado + " tau:" + str(display_phictodo) + " gc:" + str(list(self.gc))
+
         return self.string
 
     __repr__ = __str__
@@ -23,7 +48,6 @@ class MCD:
 
     def __eq__(self, other):
         return self.clave == other.clave
-
 
 #     def calcularPhiH(self):
 #         h = {}

### 2009-12-14 ws2/final/examen.tex
.tex de examen final de ws2

\documentclass[12pt]{article}

\usepackage[spanish]{babel}
\usepackage[utf8x]{inputenc}

\begin{document}

% Título

\author{Daniel Izquierdo \\ \#08-86809}
\title{Examen Final \\ Web Semántica 2}
\maketitle

% Contenido

Con el rápido crecimiento que actualmente experimenta el ``Cloud of Linked
Data'' surgen muchas nuevas aplicaciones que involucran la recuperación de datos
de varias fuentes distintas, cada una con su propia estructura. Usuarios y
aplicaciones necesitan cada vez más tener la capacidad de evaluar muchas consultas
contra fuentes de datos heterogéneas. Las fuentes de datos se pueden
caracterizar por tener estructuras distintas y contener resultados parciales de
la consulta que se realiza. Además, escogiendo un criterio de calidad para las
fuentes, resulta que todas pueden tener valores distintos para los parámetros de
calidad. Para evaluar una consulta se deben conocer sus posibles reescrituras en
términos de las fuentes disponibles.

Por otra parte también existe la necesidad de componer servicios web
descubiertos para implementar realizar flujos de trabajo. Un usuario puede
definir un flujo de trabajo como una composición de servicios abstractos que son
implementados por servicios concretos. Al hacerlo, requiere conocer las maneras
de implementar los servicios abstractos con los concretos existentes. También
existen criterios de calidad que se deben tomar en cuenta a la hora de elegir
una implementación dada.

Precisamente por la explosión de datos y servicios que se vive en los momentos,
en ambos casos el espacio de soluciones a la pregunta de cuáles son las
reescrituras buscadas puede llegar a ser muy grande. Se requiere entonces de un
sistema que pueda resolver el problema de manera eficiente y escalable. El
proyecto que estoy desarrollando es un intento de crearlo.

La capacidad de escalar a números realistas de vistas (fuentes de datos o
servicios concretos) y sub-objetivos (pasos del flujo de trabajo) al momento de
reescribir las consultas dadas ya fue lograda por \ref{yarvelo}. Mi contribución
ha sido añadir la capacidad a esta solución de tomar en cuenta constantes en las
consultas y vistas, así como de manejar el concepto de rendimiento para
seleccionar las mejores reescrituras según los parámetros de calidad en un
problema dado.

%En base a la tendencia actual que existe en la Web Semantica de publicar grandes datasets en el "Cloud of Linked Data", identifique las bondades y limitaciones del proyecto que esta desarrollando que permitan o limiten la manipulacion de este tipo de datasets.

%Su respuesta debe estar documentada con referencias bibliograficas actuales, resultados experimentales y analisis del comportamiento observado durante el proceso de experimentacion. El documento debe tener al menos dos paginas o 2000 palabras.

\end{document}

### 2009-12-14 tesis/experimentos/4/runner.sh
experimento 4

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+echo '########' | tee -a ${OUTFILE}
+
+INDIR=casos
+OUTFILE=experimento4.out
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/mcdsat
+
+for nviews in `seq 45 5 80`; do
+    for nview_sos in `seq 2 5`; do
+        VIEWSFILE=${INDIR}/views-$nviews-$nview_sos-4-10-10-3-0.5.txt
+        QUERYFILE=${INDIR}/query-$nviews-$nview_sos-4-10-10-3-0.5.txt
+
+        echo ${nviews} ${nview_sos} 4 10 10 3 0.5 | tee -a ${OUTFILE}
+
+        for i in `seq 1`; do #n times to get average
+            ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+        done
+    done
+done

### 2009-12-14 ws2/final/examen.tex
examen terminado salvo por referencias

@@
-\documentclass[12pt]{article}
+\documentclass{article}
 
@@
 \usepackage[utf8x]{inputenc}
+\usepackage{graphicx}
 
@@
 \title{Examen Final \\ Web Semántica 2}
+\date{14 de diciembre de 2009}
 \maketitle
@@
 
-%En base a la tendencia actual que existe en la Web Semantica de publicar grandes datasets en el "Cloud of Linked Data", identifique las bondades y limitaciones del proyecto que esta desarrollando que permitan o limiten la manipulacion de este tipo de datasets.
+A continuación muestro resultados experimentales que demuestran que la solución
+desarrollada puede manejar tomando en cuenta constantes y parámetros de calidad.
+Para evaluar el proyecto se ejecutaron pruebas con problemas reescritura de
+flujos de trabajo en el área de vuelos y aerolíneas con las siguientes
+características:
 
-%Su respuesta debe estar documentada con referencias bibliograficas actuales, resultados experimentales y analisis del comportamiento observado durante el proceso de experimentacion.
-El documento debe tener al menos dos paginas o 2000 palabras.
+\begin{itemize}
+
+\item Servicio abstracto {\it vuelo(x,y,a)} que relaciona las ciudades $x$ e $y$
+si hay un vuelo de la aerolínea $a$ entre ellas.
+
+\item Consultas que retornan escalas para vuelos entre una ciudad $a$ y otra
+ciudad $b$ dadas, con la condición de que todos los vuelos usan la misma
+aerolínea. Tienen la forma:
+
+$$
+q(x_0,x_1,\ldots,x_{n-1}) :- vuelo(a,x_0,z),\, vuelo(x_0,x_1,z), \ldots, vuelo(x_{n-2},x_{n-1},z), vuelo(x_{n-1},b,z)
+$$
+
+\item Servicios concretos $v_A$ que retornan vuelos con una aerolínea dada. Por
+ejemplo:
+
+$$
+v_{\mathrm{CO}} :- vuelo(x,y,\mathrm{CO})
+$$
+
+\end{itemize}
+
+Los parámetros variables en los experimentos fueron el número de vistas (de 5 a
+80 en incrementos de 5) y el
+número de objetivos de las consultas (de 2 a 5). Podemos ver los resultados en la
+figura \ref{grafico:plot1}.
+
+\begin{figure}[h!]
+  \caption{Resultados del primer experimento}
+  \centering
+   \includegraphics[width=0.5\textwidth]{plot1}
+  \label{grafico:plot1}
+\end{figure}
+
+A partir de estos resultados tenemos observaciones importantes. En primer lugar
+se nota la progresión polinomial de los tiempos de ejecución, por lo que podemos
+hablar de alta eficiencia, al menos para este tipo de problemas. Además vemos
+tiempos muy razonables para instancias con tamaños de la vida real. El caso más
+grande tiene ochenta aerolíneas y vuelos con cinco escalas --números bastante
+ajustados a lo que se vería en una agencia de viajes, por ejemplo-- fue resuelto
+en menos de cuatro minutos. Cabe acotar que, por la arquitectura de la solución
+desarrollada, se puede ``compilar'' una instancia del problema de manera de
+satisfacer múltiples consultas sobre la misma instancia pero con parámetros de
+calidad distinto de forma muy veloz; para el caso más grande toma menos de un
+segundo resolver una consulta una vez obtenida la versión compilada.
+
+A pesar de los buenos resultados, hay que aclarar que el sistema desarrollado
+está limitado en cuanto a escalabilidad a miles de vistas y decenas de objetivos
+de consulta. Al realizar pruebas con problemas de tamaño mucho mayor se
+encontraron tiempos de ejecución inmanejablemente grandes. Parte del trabajo
+futuro sobre el tema se puede realizar en la dirección de hacer accesible
+instancias de problemas con estos tamaños.
 

### 2009-12-14 ws2/final/examen.tex
archivo bib de examen final de ws2, con ref de yarvelo

@@
 Con el rápido crecimiento que actualmente experimenta el ``Cloud of Linked
-Data'' surgen muchas nuevas aplicaciones que involucran la recuperación de datos
+Data'' (\cite{Galperin09}, \cite{Benson07})
+surgen muchas nuevas aplicaciones que involucran la recuperación de datos
 de varias fuentes distintas, cada una con su propia estructura. Usuarios y
@@
 servicios concretos) y sub-objetivos (pasos del flujo de trabajo) al momento de
-reescribir las consultas dadas ya fue lograda por \ref{yarvelo}. Mi contribución
+reescribir las consultas dadas ya fue lograda por \cite{arvelo:aaai06}. Mi contribución
 ha sido añadir la capacidad a esta solución de tomar en cuenta constantes en las
@@
 
+\bibliographystyle{abbrv}
+\bibliography{ref}
+
 \end{document}

### 2009-12-15 publicaciones/eswc10/paperESWC10.tex
añadi la ultima version del paper enviada por mvidal

@@
 Since, service selection is done on-the-flight and the space of concrete services may be very large, efficiency and scalability are required.
-In this paper we propose the {\it McdSat}$^c$ framework to efficiently solve the problem of service selection in presence of a large number of concrete services.
+In this paper we propose the {\it SsdSat} framework to efficiently solve the problem of service selection in presence of a large number of concrete services.
 First, we use the Local As View approach to describe the functionality  of the concrete services in terms of views of abstract services, the quality as an overall utility function that combines the different QoS values that describe the service, and  workflows as  conjunctive queries on abstract services.
@@
 We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and have the same cost as the corresponding composition; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.
-We report on the performance of our proposed approach, and we observe that  {\it McdSat}$^c$  efficiently scales up to large instances of the service selection problem.
+We report on the performance of our proposed approach, and we observe that  {\it SsdSat}  efficiently scales up to large instances of the service selection problem.
 
@@
 In this paper we aim to provide a solution to this problem and provide techniques that efficiently traverse the space of available services and select combinations that meet the functional and non-functional requirements that describe a designed application.
-We name our proposed solution {\it MCDSat}$^c$.
+We name our proposed solution {\it SsdSat}.
 
-In {\it MCDSat}$^c$, concrete service functionalities are described as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.
-In addition, rules that define workflows and  concrete services in the {\it MCDSat}$^c$ catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied.
+In {\it SsdSat}, concrete service functionalities are described as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one semi-automatically generated for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.
+In addition, rules that define workflows and  concrete services in the {\it SsdSat} catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied.
 To represent the QoS measures, each concrete service description is annotated with a real number that represents an overall value of the QoS measures of this service.
-Finally, {\it MCDSat}$^c$ models the web service selection problem as the problem of query rewriting using views \cite{halevy:survey}.
+Finally, {\it SsdSat} models the web service selection problem as the problem of query rewriting using views \cite{halevy:survey}.
 This problem is important in the context of data integration
 \cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance \cite{AfratiLU07,levy:bucket}, and several approaches have been defined to efficiently scale up to a large number of views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-In this paper we propose the {\it MCDSat}$^c$ framework which extends the approach proposed in ~\cite{arvelo:aaai06},  with the capability of dealing with constants and identifying the composition of services that maximizes a given utility function.
+In this paper we propose the {\it SsdSat} framework which extends the approach proposed in ~\cite{arvelo:aaai06},  with the capability of dealing with constants and identifying the composition of services that maximizes a given utility function.
 
-{\it MCDSat}$^c$  translates an instance of the service selection problem into the problem of enumerating the models of a propositional theory that encodes information about the workflow, the concrete services and the conditions that need to be satisfied in order to generate a composition that maximizes the utility function.
+{\it SsdSat}  translates an instance of the service selection problem into the problem of enumerating the models of a propositional theory that encodes information about the workflow, the concrete services and the conditions that need to be satisfied in order to generate a composition that maximizes the utility function.
 This theory is represented as a CNF formula which is compiled into a deterministic
 and decomposable negation normal form d-DNNF representation, from which the models that maximizes the overall utility function are obtained in linear time using modern SAT techniques~\cite{darwiche:dnnf}.
-We empirically studied the quality of the {\it MCDSat}$^c$ on variety of benchmarks and we observed  that the proposed technique provides an efficient solution to the service selection problem which is able to scale up to large workflows as well to a large number of views.
+We empirically studied the quality of the {\it SsdSat} on variety of benchmarks and we observed  that the proposed technique provides an efficient solution to the service selection problem which is able to scale up to large workflows as well to a large number of views.
 
@@
 The next section motivates our approach by using an example, and  we summarize existing related work in section 3.
-Section 4 describes the {\it MCDSat}$^c$ system architecture, and  the experimental study is reported in section 5.
+Section 4 describes the {\it SsdSat} system architecture, and  the experimental study is reported in section 5.
  In section 6 we present the formalization of the web service selection problem as a Propositional Logic theory, and finally, section 7 outlines our conclusions and future work.
@@
 \section{Motivating Example}
-Consider the abstract and concrete services in Table ~\ref{table:services}  which represent applications in the flight domain.
+Consider the abstract and concrete services in Table ~\ref{table:services}  that represent applications in the flight domain.
 
@@
 in the size of its DAG.
-{\it MCDSat}$^c$  exploits the properties of a d-DNNF theory, to efficiently provide a solution to the problem of enumerating the service compositions that correspond to models of the  theory, i.e., the service compositions that implement the abstract  workflow and best meet the utility function.
+{\it SsdSat}  exploits the properties of a d-DNNF theory, to efficiently provide a solution to the problem of enumerating the service compositions that correspond to models of the  theory, i.e., the service compositions that implement the abstract  workflow and best meet the utility function.
 DECIR ALGO SOBRE EL ALGORITMO QUE PRODUCE LOS MEJORES.
@@
 algorithm over a large number of benchmarks often showing performance
-improvements of several orders of magnitude. However, the McdSat algorithm was not desgined for rewriting problems involving explicit constants,
+improvements of several orders of magnitude. However, the {\it MCDSat} algorithm was not desgined for rewriting problems involving explicit constants,
 nor to compute the best rewritings with respect to a given utility function or cost model, and this paper we propose a new encoding  that overcomes these limitations.
@@
 
-\section{The McdSat$^c$ Architecture}
+\section{The {\it SsdSat} Architecture}
+Figure ~\ref{fig:architecture} presents the architecture of{\it SsdSat}; it is comprised of a {\it Catalog} of service descriptions, the {\it SSP Translator},  the {\it c2d}, the {\it Minimal Model Enumerator}, and the {\it Model Translator}.
 
-In order to do so, the theory $\Theory{\Q}$ is transformed into a d-DNNF
-theory $\Delta(\Q)$ with a CNF to d-DNNF compiler.
-The compiler's algorithm is similar to the DPLL algorithm
-for SAT but enhanced with advanced techniques such as clause
-learning and caching.
 \begin{figure}
 \centering
-\includegraphics[width=10cm]{mdcsatC.pdf}
-\caption{The McdSat$^c$ Architecture}
-\label{fig:dnnf}
+\includegraphics[height=90mm,width=3.5in]{mdcsatC.pdf}
+\caption{The {\it SsdSat} Architecture}
+\label{fig:architecture}
 \end{figure}
 
+First, an instance of the Service Selection Problem is defined as an abstract workflow represented by a conjunctive query on abstract services which is given as input, and a set of concrete services defined by views of abstract services.
+
+The {\it SsdSat}  Catalog is populated with descriptions of abstract and concrete services; each service is described in terms of input and output attributes of the service, and a real value representing a utility function that aggregates the QoS parameters of the service.
+Additionally, concrete services are defined as views of abstract services following the paradigm Local As View proposed in ~\cite{levy:bucket}; these descriptions can be generated by the {\cal DEIMOS} system proposed by~\cite{AmbiteISWC09}.
+
+The {\it SSP Translator} encodes an instance of the Service Selection Problem as a Conjunctive Normal Form formula that corresponds to the propositional logic theory that represents the instance of the problem.
+{\it c2d}~\cite{c2d} is an off-the-shell component that compiles  the CNF formula that represents the input instance of the service selection problem into  d-DNNF.
+{\it c2d} uses conflict direct backtracking with clause learning, and information about the state of clauses to  index into a cache intermediate results.
+Thus, although the complexity of  compiling CNF into d-DNNF exponentially increases  with the connectivity  of the  CNF, {\it c2d} is able to compile large instances of SSP is few minutes.
+The {\it Minimal Model Enumerator} extracts the models from  the d-DNNF in linear time; also models counting and enumeration of the model that maximizes the utility function can be performed in linear time.
+Finally,
+the {\it Model Translator} decodes the models into the service compositions that solve the input problem.
+
 \section{Experimental Results}
-We have conducted an empirical analysis  on the benefits of the techniques implemented in the   McdSat$^c$ system.
+We have conducted an empirical analysis  on the benefits of the techniques implemented in the  {\it SsdSat} system.
 
@@
 
-\section{Formalization of the Service Selection Problem}
+\section{Formalization of the Service Selection Problem (SSP)}
+We consider service catalogs of the form $C=\tup{S,T}$ where
+$S$ is a set of predicates that represent abstract services and $T=\{T_s\}_{s\in S}$ is a collection
+of tables that represents the result of the evaluation of the abstract services. In the context of the Service Selection Problem, 
+a service catalog $C$ is an idealized description of 
+the output produced by abstract workflows  implemented by multiple concrete services described
+as views.
 
-We consider databases of the form $D=\tup{P,T}$ where
-$P$ is a set of predicates and $T=\{T_p\}_{p\in P}$ is a collection
-of tables that represents the predicates in extensional form.
-A conjunctive query $Q$ over $P$ is of the form 
+A workflow  $Q$ over $S$ is represented by a conjunctive query of the form 
 \begin{alignat*}{1}
-Q(\vvec{x})\ \qrule\ \  p_1(\vvec{x}_1),\, p_1(\vvec{x}_2),\, \ldots,\, p_m(\vvec{x}_m)\,,
+Q(\vvec{x})\ \qrule\ \  s_1(\vvec{x}_1),\, s_1(\vvec{x}_2),\, \ldots,\, s_m(\vvec{x}_m)\,,
 \end{alignat*}
-where $p_i\in P$, $\vvec{x}$ is a vector of variables, and each
+where $s_i\in S$, $\vvec{x}$ is a vector of variables, and each
 $\vvec{x}_i$ is a vector of variables and constants.
-The result of $Q$ over $D$, denoted as $Q(D)$, is the table with
+The result of $Q$ over $C$, denoted as $Q(C)$, is the table with
 $|\vvec{x}|$ columns that result of the projection of the relational
@@
 
-A view $V$ over $D$ is a query over $P$. In the context of the service selection problem, 
-the database $D$ is an idealized description of 
-the output produced abstract workflow  implemented by multiple concrete services described
-as views.
-Given a database $D$, a query $Q$ and a collection of 
+A view $V$ over $C$ is a query over $S$. Given a database $C$, a query $Q$ and a collection of 
 views $E=\tup{\{V_i\}_i,\{E_i\}_i}$, we are required to find
-all the tuples in $Q(D)$ obtainable from the views in $E$.
-That is, we need to find all the \emph{compositions} of the form
+all the tuples in $Q(C)$ obtainable from the views in $E$.
+That is, we need to find all the \emph{service compositions} of the form
 \begin{alignat*}{1}
@@
 \end{alignat*}
-such that $R(E) \subseteq Q(D)$.
-A service selection problem (SSP) is a tuple $\tup{P,Q,\{V_i\}}$ where $P$
-is a set of predicates that represent abstract services, $Q$ is a query over $P$ and $\{V_i\}$ is a collection of views that define the concrete services in terms of  abstract services.
+such that $R(E) \subseteq Q(C)$.
+A service selection problem (SSP) is a tuple $\tup{S,Q,\{V_i\}}$ where $S$
+is a set of predicates that represent abstract services, $Q$ is a query over $S$ and $\{V_i\}$ is a collection of views that define the concrete services in terms of  abstract services.
 We assume \emph{safe} problems in the sense that all
@@
 query or views.
-A composition $C$ is \emph{valid} if for all databases $D=\tup{P,T}$
-and extensions $\{E_i\}$, $R(E) \subseteq Q(D)$.
+A service composition $SC$ is \emph{valid} if for all catalogs $C=\tup{S,T}$
+and extensions $\{E_i\}$, $R(E) \subseteq Q(C)$.
 A collection $\R$ of valid compositions is a solution if
-for all databases $D=\tup{P,T}$ and extensions $\{E_i\}$, there
-is no other $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(D)$.
-We are interested in finding a composition $\R$.
+for all service catalogs $C=\tup{S,T}$ and extensions $\{E_i\}$, there
+is no other $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(C)$.
+We are interested in finding a service composition $\R$.
 
 \subsection{Logical Theories}
+We use an approach similar to one described in ~\cite{arvelo:aaai06} to encode the Service Selection Problem, the reader is referred to \cite{arvelo:aaai06} for a comprehensive
+description of the propositional theory. We identify the service compositions by enumerating the
+models of a logical theory $\Delta=\Delta_{com}\cup\Delta_{ssd}^1\cup\cdots\Delta_{ssd}^m$
+where $\Delta_{com}$ specified how to combine $m$ independent
+copies of SSD theories $\Delta_{SSd}$ that cover all goals in $Q$.
+Each $\Delta^i_{ssd}$ is a copy of the theory $\Delta_{ssd}$
+in which each literal $s$ is tagged as $s^i$.
+The theory $\Delta_{ssd}$ consists of different groups of 
+clauses that guarantees that its models are in correspondence
+with the SSDs, while the theory $\Delta_{com}$ contains additional
+clauses to guarantee a sound and complete composition of the SSDs.
+A Selection Service Description (SSD) is  $\tup{V,\tau,SS}$ for $\Q$, where $V$ is the concrete service used to cover the subset of abstract services $SS$, and  $\tau\subseteq Vars(Q)\times Vars(V)$ such that all variables in
+$\tau_x=\{ y\in Vars(V) : (x,y)\in\tau\}$ are distinguished in $V$.
+A SSD is a simplification of the MCDs proposed by ~\cite{pottinger:minicon} to characterize the views that can be used to subsets of sub-goals of a query.
 
-In \cite{arvelo:aaai06}, we showed that a rewriting
-for a query $Q$ with $m$ goals can be obtained by enumerating the
-models of a logical theory
-$\Delta=\Delta_{rew}\cup\Delta_{mcd}^1\cup\cdots\Delta_{mcd}^m$
-where $\Delta_{rew}$ specified how to combine $m$ independent
-copies of MCD theories $\Delta_{mcd}$ that cover all goals in $Q$.
-Each $\Delta^i_{mcd}$ is a copy of the theory $\Delta_{mcd}$
-in which each literal $p$ is tagged as $p^i$.
-The theory $\Delta_{mcd}$ consists of different groups of 
-clauses that guarantees that its models are in correspondence
-with the MCDs, while the theory $\Delta_{rew}$ contains additional
-clauses to guarantee a sound and complete composition of the MCDs.
-The reader is referred to \cite{arvelo:aaai06} for a comprehensive
-description of the propositional theory.
+The SSP theory for $\Q$,  denoted by $\Theory{\Q}$, consists
+of the following propositional variables:
+\begin{enumerate}\denselist
+\item $\{v_0,\ldots,v_n\}$ to indicate the SSD uses view $V_i$;
+      $v_0$ is used to indicate the \emph{null} SSD,
+\item $\{g_1,\ldots,g_m\}$ to indicate the goals in $SS$,
+\item $\{z_{j,k,i}\}$ to indicate the $j$th goal in $Q$ is
+      covered by the $k$th goal in $V_i$, and
+\item $\{t_{x,y}\}$ to indicate that $(x,y)\in\tau$.
+\item $t_{x,y}$
+to denote that the variable/constant $x$ in the query is mapped
+into the variable/constant $y$ in the view, and propositions $v_i$
+to indicate that the SSD uses view $V_i$.
+\end{enumerate}
+The range of indices for the $z$ variables depend on $\Q$;
+e.g. if $V_i$ has $K$ goals of the same type as goal
+$j$ in $Q$, then $1\leq k\leq K$ for $z_{j,k,i}$.
+For $t$ variables, $x$ ranges in the number of variables
+in $Q$ and $y$ ranges in the maximum number of variables
+in the views.
 
-Let us describe the difficulties that arise when constants
-are presents in a SSP with the following example:
-\begin{alignat*}{1}
-Q(x,z)\ &\qrule\ \ p_1(x,y),\, p_2(y,z)\,, \\
-V_1(x_1)\ &\qrule\ \ p_1(x_1,A)\,, \\
-V_2(x_2)\ &\qrule\ \ p_2(B,x_2)\,, \\
-V_3(x_3,y_3)\ &\qrule\ \ p_2(x_3,y_3)
-%V_4(x_4,y_4)\ &\qrule\ \ p_1(x_1,A),\, p_2(B,y_4)\,,
-\end{alignat*}
-where $Q$ is the query and $\{V_1,V_2,V_3\}$ are the 
-views. In this case, the only rewriting is 
-$R(x,z)\qrule V_1(x),V_3(A,z)$ because the candidate 
-$R(x,z)\qrule V_1(x),V_2(z)$ is not valid as it maps
-$y$ into constants $A$ and $B$ that denote different
-objects.
+The following  clauses encode the SSP problem in terms of the SSDs.
+~\cite{RajaramanSU95} showed that for queries without negation or arithmetic comparisons,  but with constants, and $m$ goals and $q$ variables, it is enough
+to consider service compositions of length at most  $m$ plus  $q$ subgoals ~\cite{Ullman00}.
 
-The main problem when handling constants is to
-be sure that different constants are not mapped into
-each other either directly or indirectly (via transitivity).
+\begin{enumerate}[C1.]\denselist
+\item (At least of view is used): $\bigvee_{i=0}^n v_i$
+\item (At most one view is used): $v_i \Rightarrow \neg v_j$ for $0\leq i,j \leq n$,
+\item (Null view equals null): $v_0 \Rightarrow \neg g_j$ for $1\leq j\leq m$,
+\item (Views are useful): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ for $1\leq i\leq n$,
+\item (Subgoals are covered at most once): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ for
+      appropriate $i,j,k,l$ with $k\neq l$,
+\item (Scope of views): $v_i \Rightarrow \neg g_j$ for such goals that can't be covered by $V_i$,
+\item (Consistency): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ for appropriate $i,j,k$.
+\item[C8.] (Dead variables): $v_i \Rightarrow \neg t_{x,y}$ for all $x,y$ with $y\notin V_i$,
+\item[C9.] (1-1 for $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ for all
+           existential variables $y,y'\in V_i$,
+\item[C10.] (Distinguished): $v_i \Rightarrow \neg t_{x,y}$ for all distinguished $x\in Q$ and
+            existential $y\in V_i$,
+\item[C11.] (Existential): $v_i \land t_{x,y} \Rightarrow g_j$ for all existential $y\in V_i$ and
+            goals $j$ that contain $x\in Q$,
+\item[C12.] (Matching): $v_i \land z_{j,k,i} \Rightarrow t_{x,y}$ for all $x$ and $y$ that
+            should match if goal $j$ in $Q$ is covered by goal $k$ in $V_i$.
+\item[C13.] (If all variables in the view are distinguished, then the SSD covers only one goal):
+            $v_i \land g_j \Rightarrow \neg g_k$.
+\end{enumerate}
 
-\subsection{Utility Functions}
+These first thirteen are the same clauses used by the {\it MCDSat} to encode the Query Rewriting Problem in propositional logic ~\cite{arvelo:aaai06}.
+The following clauses are required to allow constants in the workflows and the views.
+As we showed in Section 2, the main problem when handling constants is to
+be sure that different constants are not mapped into each other either directly or indirectly (via transitivity). These clauses ensure that two constants are mapped iff they are the same.
 
+\begin{enumerate}[S1.]\denselist
+\item (Inconsistent-1): $t_{x,A} \Rightarrow \neg t_{x,B}$,
+\item (Inconsistent-2): $t_{A,x} \Rightarrow \neg t_{B,x}$,
+\item (Inconsistent-3): $\neg t_{A,B}$,
+\item (Transitivity-1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$,
+\item (Transitivity-2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
+\end{enumerate}
+Clauses S1--S3 prune candidate service compositions in which one constant
+is directly mapped into a different one, and the last two implement
+a restricted propagation of transitivity.
+Similarly, $\Delta_{rew}$ must be extended with the clauses:
+\begin{enumerate}[S1.]\denselist
+\item[S6.] (Inconsistent-4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
+\end{enumerate}
+
+\subsection{Maximum-Utility Function}
 We assume a simple additive utility function in which each view
@@
 all the compositions of maximum utility value.
-
-%The cost model is simple yet expressive; e.g., if all
-%costs are unit, then best rewritings are those with
-%minimum number of views, yet in cases where views
-%represent different sources, it is natural
-%to associate them with different costs.
-
-\subsection{Extended Theories}
-
-The theory $\Delta_{mcd}$ makes use of propositions $t_{x,y}$
-to denote that the variable/constant $x$ in the query is mapped
-into the variable/constant $y$ in the view, and propositions $v_i$
-to indicate that the MCD uses view $V_i$.
-We obtain a sound and complete theory for SSPs with constants if
-$\Delta_{mcd}$ is extended with the clauses:
-
-\begin{enumerate}[C1.]\denselist
-\item (Inconsistent-1): $t_{x,A} \Rightarrow \neg t_{x,B}$,
-\item (Inconsistent-2): $t_{A,x} \Rightarrow \neg t_{B,x}$,
-\item (Inconsistent-3): $\neg t_{A,B}$,
-\item (Transitivity-1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$,
-\item (Transitivity-2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
-\end{enumerate}
-Clauses C1--C3 prune candidate service compositions in which one constant
-is directly mapped into a different one, and the last two implement
-a restricted propagation of transitivity.
-Similarly, $\Delta_{rew}$ must be extended with the clauses:
-\begin{enumerate}[C1.]\denselist
-\item[C6.] (Inconsistent-4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
-\end{enumerate}
-~\cite{RajaramanSU95} showed that for queries without negation or arithmetic comparisons,  but with constants, and $m$ goals and $q$ variables, it is enough
-to consider service compositions of length at most  $m$ plus  $q$ subgoals ~\cite{Ullman00}.
-
-Given a SSP problem $\tup{Q,\{V_i\}}$ possibly with constant 
-symbols, we construct logical theories whose models are in
-correspondence with the service compositions. These theories are then
-compiled into d-DNNF from which all rewritings are extracted
-efficiently.
-Likewise, best service compositions can be computed by performing model
-enumeration twice: the first pass computes the optimal utility function,
-and the second filters out suboptimal service compositions.
-However, there is a better way, which we will describe in the next section.
-
-\subsection{Maximum-Utility Function}
-
 Darwiche \& Marquis \cite{darwiche:weighted} show how to compute the
@@
 
-We have shown how the propositional theory used in McdSat can
+We have shown how the propositional theory used in {\it MCDSat} can
 be extended to support constants, and how it can be used to 

### 2009-12-16 publicaciones/eswc10/paperESWC10.tex
parte de experimental results rellena

@@
 \newcommand{\mcdsat}{\textsc{McdSat}}
+\newcommand{\ssdsat}{{\it SsdSat} }
 \newcommand{\minicon}{{MiniCon}}
@@
 \section{Experimental Results}
-We have conducted an empirical analysis  on the benefits of the techniques implemented in the  {\it SsdSat} system.
+We have conducted an empirical analysis on the benefits of the techniques
+implemented in the  {\it SsdSat} system. The characteristics of the experiment
+follow:
 
 \begin{description}
-\item[Dataset and Query Benchmark:]  we conducted our experiments over a benchmark that includes XX queries and a set of  YY views.
-Queries and views are starts and chains, and they have between ZZ and TT constants; queries have QQ sub-subgoals.
-\item[Evaluation Metrics:] we report on runtime performance which corresponds to the  {\it user time} produced by the {\it time} command of the Unix operation system. 
+\item[Dataset and Query Benchmark:] we evaluated the system on a set of three
+different benchmarks as described below. The objective was to assess the
+performance of \ssdsat on varying conditions.
+\item[Tests:] A benefit of the \ssdsat approach is that one can compile the
+logical theory of a given problem instance and then calculate all rewritings or
+the best ones any number of times in linear time. The cost model for finding
+best rewritings can even be changed with no need to recompile the logical
+theory, which is the costly phase of the process. Because of this, for each run
+benchmark we make a clear distinction on whether the reported times comprise
+theory compilation only or compilation plus enumeration rewriting.
+\item[Evaluation Metrics:] we report on runtime performance which corresponds to the {\it user time} produced by the {\it time} command of the Unix operating system. 
+\item[Environment:] the experiments were run on a machine with an Intel Core 2 Duo 2GHz processor and 4GB of RAM.
 \end{description}
+
+The first benchmark consisted of a problem instance modeled on air
+travel queries. Concrete services were of the form
+
+\begin{alignat*}{1}
+v_i(\vvec{x}, \vvec{y})\ \qrule\ \  flight(\vvec{x},\vvec{y},i)
+\end{alignat*}
+
+Here, we suppose a service returns all flights between two cities with a specific
+airline. View $i$ describes flights for the airline $i$, since the abstract service
+$flight(x,y,a)$ is satisfied when a direct flight exists departing from city $x$
+and arriving at city $y$ with airline $a$. Workflows had the form
+
+\begin{alignat*}{1}
+q(\vvec{x}_1, \ldots, \vvec{x}_n)\ \qrule\ flight(CCS, \vvec{x}_1, \vvec{a}),
+flight(\vvec{x}_1, \vvec{x}_2, \vvec{a}),
+\ldots,
+flight(\vvec{x}_{n-1}, \vvec{x}_n, \vvec{a}),
+flight(\vvec{x}_n, WAS, \vvec{a})
+\end{alignat*}
+
+Our benchmark for this type of problem included workflows with two to five
+sub-goals and sets of five to eighty concrete services. We measured both compilation time
+and time to list all rewritings. As it turned out, listing all rewritings took
+less than half a second in the worst case. The results for compilation time can
+be seen in figure \ref{fig:plot1}.
 
@@
 \label{fig:plot1}
+\end{figure}
+
+In an attempt to induce exponential growth in the number of rewritings so as to
+get longer rewriting generation times, we modified this benchmark setting the
+number of concrete services per airline to two. We ran tests with this modification and the
+same numbers of workflow sub-goals and concrete services (meaning we were now handling five to
+forty airlines). For this case we recorded full rewriting enumeration times,
+resulting in the graph in figure \ref{fig:plot2}.
+
+\begin{figure}
+\centering
+\includegraphics[width=8cm]{plot1.pdf}
+\caption{Rewriting Enumeration Time Benchmark I}
+\label{fig:plot2}
+\end{figure}
+
+Although both tests previously described showed good results, they suffered from
+a considerable issue which was lack of multiple sub-goals in concrete services.
+We therefore began running tests with unstructured, randomly generated instances
+with three variables per abstract service, ten distinct variables and ten
+distinct constants, four sub-goals in workflows and concrete services and a
+$50\%$ probability of constants appearing in each abstract service argument. We
+can observe results in figure \ref{fig:plot3}.
+
+\begin{figure}
+\centering
+\includegraphics[width=8cm]{plot1.pdf}
+\caption{Compilation Time Benchmark I}
+\label{fig:plot3}
 \end{figure}

### 2009-12-22 publicaciones/eswc10/paperESWC10.tex
paper submiteado

--- 2009-12-13 publicaciones/eswc10/paperESWC10.tex
@@
 \documentclass{llncs}
+
 \usepackage{amsmath}
-\usepackage{txfonts}
-\usepackage[boxed]{algorithm2e}
 \usepackage{amssymb}
-\usepackage[pdftex]{graphicx}
+\usepackage{graphicx}
 \usepackage{enumerate}
-\DeclareGraphicsExtensions{.jpg, .png, .mps, .bmp, .pdf}
-\usepackage{epsfig}
-\usepackage{subfigure}
-\newtheorem{Definition}{Definition} 
-\newtheorem{Example}{Example} 
-\newtheorem{Theorem}{Theorem} 
-\newtheorem{Proof}{Proof} 
-\newcommand{\denselist}{\topsep 0pt \itemsep -4pt}
+\usepackage{url}
+
 \newcommand{\tup}[1]{\langle #1 \rangle}
@@
 \newcommand{\Q}{\mathcal{Q}}
-\newcommand{\body}{{body}}
-\newcommand{\head}{{head}}
 \newcommand{\qrule}{:\!\!-}
-\newcommand{\arc}{\text{arc}}
-\newcommand{\Theory}[1]{T(#1)}
 \newcommand{\mcdsat}{\textsc{McdSat}}
@@
 \newcommand{\Omit}[1]{}
-\newcommand{\citeX}[1]{\citeauthor{#1}~\citeyear{#1}}
+
+% flights
+\newcommand{\flight}{\text{\it flight}}
+\newcommand{\UScity}{\text{\it uscity}}
+\newcommand{\national}{\text{\it national}}
+\newcommand{\oneway}{\text{\it one-way}}
+\newcommand{\onestop}{\text{\it one-stop}}
+\newcommand{\flightPA}{\text{\it flight-to-pa}}
+\newcommand{\onestopPA}{\text{\it onestop-to-pa}}
+\newcommand{\fromNY}{\text{\it from-ny}}
+\newcommand{\PA}{\text{PA}}
+\newcommand{\NY}{\text{NY}}
+\newcommand{\AL}{\text{AL}}
 
 \begin{document}
+
 \allowdisplaybreaks
-\title{Selecting the Maximum-Utility  Web Services  Using Circuits}
-\author{Daniel Izquierdo \inst{1}  and Blai Bonet \inst{1}  and Mar\'{\i}a-Esther Vidal \inst{1}}
+\title{Optimal Instantiation of Abstract Workflows using Logic and Circuits}
+\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
+\institute{Departamento de Computaci\'on \\
+           Universidad Sim\'on Bol\'{\i}var \\
+           Caracas 89000, Venezuela \\
+           \url{{idaniel,bonet,mvidal}@ldc.usb.ve}}
+\maketitle
 
-\institute{Universidad Sim\'on Bol\'{\i}var\\
-                Caracas, Venezuela\\
-                \{idaniel,bonet,mvidal\}ldc.usb.ve} \maketitle
 \begin{abstract}
-In Service Oriented Architectures (SOA), concrete services are described in terms of functionality   and QoS parameters, while software applications are specified as workflows of abstract services and non-functional criteria that establish the  application desirable functionality and  expected quality, respectively.
-During workflow execution time,  abstract services  are rewritten
-with the concrete services  that meet the workflow  functional and non-functional restrictions.
-Since, service selection is done on-the-flight and the space of concrete services may be very large, efficiency and scalability are required.
-In this paper we propose the {\it McdSat}$^c$ framework to efficiently solve the problem of service selection in presence of a large number of concrete services.
-First, we use the Local As View approach to describe the functionality  of the concrete services in terms of views of abstract services, the quality as an overall utility function that combines the different QoS values that describe the service, and  workflows as  conjunctive queries on abstract services.
-Based on this representation, we
-model the problem of service selection as a known problem in the area of integration systems referred as query rewriting using views.
-We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and have the same cost as the corresponding composition; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.
-We report on the performance of our proposed approach, and we observe that  {\it McdSat}$^c$  efficiently scales up to large instances of the service selection problem.
+The typical Service Oriented Architecture consists of two main layers.
+The concrete layer with concrete services whose functionalities are
+described in terms of pre- and post-conditions and non-functional
+properties in terms of QoS parameters, and the abstract layer with
+software applications whose functionalities are described in terms
+of abstract workflows and non-functional properties in terms of
+restrictions on QoS.
+During the execution of a workflow, the abstract services are instantiated
+into concrete services that meet the functional and non-functional
+requirements. This instantiation, that must be done on-the-flight,
+consists in a search on a combinatorial space of possibilities.
+In this paper, we propose a framework to efficiently solve the
+instantiation problem that is firmly grounded on logic.
+The framework adopts the Local-As-View approach in which
+the functionality of concrete services is described using views
+of abstract services, the quality of an instantiation as an
+overall utility function that combines the different QoS parameters,
+and the abstract workflows as conjunctive queries on abstract services.
+Using this representation, the problem of workflow instantiation 
+is cast as a problem of query rewriting, from the area of integration
+systems.
+Then, building on related work, we devise an encoding of the
+workflow instantiation problem as a logical theory whose models
+are in correspondence with the instantiations of the workflow,
+and the best ranked models are in correspondence with the
+optimal instantiations of the workflow.
+Thus, by exploiting known properties of logical theories in 
+d-DNNF format, we provide an efficient and scalable solution
+to the workflow instantiation problem.
+The approach does not only scale up to large instances, as the
+experimental results show, but is also sound and complete since,
+being founded on logic, is amenable to formal analysis.
+\end{abstract}
 
-\end{abstract}                
 \section{Introduction}
-Under the umbrella of the Semantic Web and supported by Service Oriented architectures, the number of Web data sources and services has exploded during the last few years.
-For example,  currently, the molecular biology databases
-collection includes 1,170 databases~\cite{Galperin09}, that is
-95 more than the previous year~\cite{Galperin2008} and 110 more than two years ago~\cite{Galperin2007};  tools
-and services as well as the number of instances published by these resources, follow a similar progression~\cite{Benson07}. Thanks
-to this wealth, users rely more on various
-digital tasks such as data retrieval from public data sources and
-data analysis with Web tools or services organized in
-complex workflows.
 
-In order to design an application, users usually specify the functionality in terms of a workflow of abstract services, while the desirable quality  is defined by restricting the values of the QoS parameters that describe the concrete available services.
-To evaluate the proposed workflow,  abstract services are instantiated on concrete services in a way restrictions on the QoS parameters are satisfied.
-However, since the number of available sources has an impact on the performance of this task, techniques that enable us to efficiently scale up to a large number of services are required.
+Under the umbrella of the Semantic Web, and supported by Service Oriented
+Architectures (SOA), the number of Web data sources and services has exploded
+during the last few years. For example, the molecular biology database
+collection currently contains 1,170 databases~\cite{Galperin09}, a number
+that is 95 more than the previous year~\cite{Galperin2008} and 110 more
+than two years ago~\cite{Galperin2007}; tools and services, as well as the
+number of instances published by these resources, follow a similar
+progression~\cite{Benson07}. Thanks to this wealth of resources, users'
+tendency is to rely more on automatic methods for handling them such as
+data retrieval from public sources and analysis using Web
+tools or services composed in complex workflows.
 
-In this paper we aim to provide a solution to this problem and provide techniques that efficiently traverse the space of available services and select combinations that meet the functional and non-functional requirements that describe a designed application.
-We name our proposed solution {\it MCDSat}$^c$.
+The typical SOA consists of two layers. The concrete layers that is
+made of concrete services whose functionalities are described in
+terms of pre- and post-conditions and non-functional properties in
+terms of QoS parameters, and the abstract layer that is made of 
+software applications whose functionalities are described in terms
+of abstract workflows and non-functional properties in terms of
+restrictions on QoS.
+The execution of an abstract workflow involves the instantiation
+of the abstract services into concrete services that meet the
+functional and non-functional requirements. This instantiation
+process can be seen as a search for a target instantiation on
+the combinatorial space of all valid instantiations.
+Thus, one is interested in efficient techniques for performing
+this search that are able to scale up as the number of concrete
+services or the complexity of the workflow increases.
+We call the problem of instantiating a given abstract workflow
+with concrete services, from a given pool of concrete services,
+such that certain QoS demands are meet, the Workflow Instantiation
+Problem (WIP).
 
-In {\it MCDSat}$^c$, concrete service functionalities are described as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.
-In addition, rules that define workflows and  concrete services in the {\it MCDSat}$^c$ catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied.
-To represent the QoS measures, each concrete service description is annotated with a real number that represents an overall value of the QoS measures of this service.
-Finally, {\it MCDSat}$^c$ models the web service selection problem as the problem of query rewriting using views \cite{halevy:survey}.
+In this paper, we consider a restricted version of WIP that adopts
+the Local-As-View (LAV) approach \cite{levy:bucket}.
+In LAV, all elements in a problem are specified with a common language
+that is grounded on \emph{abstract services} such that concrete
+services are described as views of abstract services, the quality
+of an instantiation as an overall utility function that combine the
+different QoS parameters, and the abstract workflow as a conjunctive
+query on the abstract services; this representation is similar to the
+one that is semi-automatically generated for the DEIMOS system
+\cite{AmbiteISWC09}.
+In this version of WIP, the rules that define workflows (conjunctive queries)
+and concrete services (views) are created in a way that all the 
+functional restrictions on pre- and post-conditions of services
+and their combinations are satisfied, and the QoS measures are
+represented by annotating each concrete service description with
+a real number that represent the overall QoS utility of the service.
+
+Under these assumptions, the WIP can be cast as the well-known Query
+Rewriting Problem (QRP) for LAV which is central to integration systems
+\cite{halevy:survey}. The QRP consists of a conjunctive query that must be
+answered in terms of views in which the query and the views are described
+using LAV with abstract relations.
 This problem is important in the context of data integration
-\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance \cite{AfratiLU07,levy:bucket}, and several approaches have been defined to efficiently scale up to a large number of views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-In this paper we propose the {\it MCDSat}$^c$ framework which extends the approach proposed in ~\cite{arvelo:aaai06},  with the capability of dealing with constants and identifying the composition of services that maximizes a given utility function.
+\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance
+\cite{AfratiLU07,levy:bucket}, and several approaches have been defined
+that scale to a large number of views
+\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
 
-{\it MCDSat}$^c$  translates an instance of the service selection problem into the problem of enumerating the models of a propositional theory that encodes information about the workflow, the concrete services and the conditions that need to be satisfied in order to generate a composition that maximizes the utility function.
-This theory is represented as a CNF formula which is compiled into a deterministic
-and decomposable negation normal form d-DNNF representation, from which the models that maximizes the overall utility function are obtained in linear time using modern SAT techniques~\cite{darwiche:dnnf}.
-We empirically studied the quality of the {\it MCDSat}$^c$ on variety of benchmarks and we observed  that the proposed technique provides an efficient solution to the service selection problem which is able to scale up to large workflows as well to a large number of views.
+The recent approach of Arvelo et al.\ \cite{arvelo:aaai06} is based 
+on the efficient enumeration of models of a propositional logic theory. 
+Given a QRP, a logical theory is constructed such that each model of
+the theory encodes a valid rewriting, and thus all valid rewritings
+are obtained by enumerating all models of the theory. This enumeration
+can be efficiently performed if the logical theory is in certain 
+normal form called deterministic and decomposable negation normal
+form (d-DNNF) \cite{darwiche:d-dnnfs}. Thus, the approach consists in
+transforming (called compiling in the field of knowledge compilation)
+the logical theory into d-DNNF format for enumerating its models efficiently.
 
- This paper is comprised of six additional sections.
-The next section motivates our approach by using an example, and  we summarize existing related work in section 3.
-Section 4 describes the {\it MCDSat}$^c$ system architecture, and  the experimental study is reported in section 5.
- In section 6 we present the formalization of the web service selection problem as a Propositional Logic theory, and finally, section 7 outlines our conclusions and future work.
+Yet d-DNNF theories not only support the efficient enumeration of
+models but other operations too. Among them, the enumeration of
+the best ranked models can also be performed efficiently on d-DNNF.
+Given a literal ranking function $r(\ell)$ that assign ranks
+to each literal $\ell$, one defines the rank $r(\omega)$ of a
+model $\omega$ as the sum of the ranks of the literals made true
+by $\omega$ (i.e., $r(\omega)\doteq\sum_{\omega\vDash\ell}r(\ell)$),
+and say that $\omega$ is a best (ranked) model if there is no model
+$\omega'$ such that $r(\omega')<r(\omega)$.
+
+Given a theory in d-DNNF, one computes the rank of the best models,
+and a best model, in time linear in the size of the d-DNNF.
+This computation transforms the DAG of the d-DNNF into a arithmetic
+circuit by replacing the AND nodes with `+' and the OR nodes with `min'.
+The literal ranking function assigns
+values to the leaves of the circuit that are propagated to the
+root in linear time. The value of the root is the rank of the
+best model \cite{darwiche:weighted}.
+
+In this paper, we exploit the properties of d-DNNFs by constructing
+a logical theory whose models encode the valid workflow instantiations
+and best models encode the optimal (best) instantiations.
+Thus, the combinatorial search is reduced to the computation
+of a best model of a logic theory which can be efficiently
+performed once the theory is transformed into d-DNNF format.
+
+The paper contains six more sections. The next section illustrates and
+motivates the WIP with a simple yet typical example. Section~3 summarizes
+the existing related work in three areas of selection of services,
+query rewriting, and knowledge compilation.
+Then, Sections~4 and 5 describe the architecture of the system and 
+report our empirical results over different benchmark problems
+respectively.
+For the interested reader, Section~6 presents a formal description
+of the proposed approach together with an analysis.
+The paper concludes with a final discussion in Section~7.
 
 \section{Motivating Example}
-Consider the abstract and concrete services in Table ~\ref{table:services}  which represent applications in the flight domain.
 
-\begin{table}
-\begin{tabular}{| l| l |}
-\hline
-Abstract Service & Description \\
-\hline
-$fight(x1,x2)$ & relates cities x1 and x2 if there is a flight between them\\
-\hline
-$uscity(x1)$ &  determines if a city x1 is a US city\\
-\hline \hline
-Concrete  Service & Description \\
-\hline
-$national(x1,y1)$&  returns  two US cities that are connected by  direct flights \\
-\hline
-$oneway(x1,y1)$ & returns two cites that are connected by one way direct flights \\
-\hline
-$onestop(x1,y1)$ &  returns  two cites that are connected by one-stop flights \\
-\hline
-$ flightToCCS(x1,"CCS")$ & returns a city that is connected to  Caracas by a direct flight\\
- \hline
-$ onestopToCCS(x1,y1)$ & returns two cites x1 and y1, such that there is  one-stop flight\\ 
- & from city x1 to Caracas stopping at city y1.\\
- \hline
- $fromWAS("WAS",y1)$ &  returns a city y1 to which flights from Washington arrive. \\
- \hline
-\end{tabular}
-\label{table:services}
-\caption{Abstract and Concrete Services}
-\end{table}
+Consider a simple flight-information system which contains information
+about flights between cities and information about which cities are in
+the US. Such a system can be described using LAV with the two abstract
+services $\flight(x,y)$ and $\UScity(x)$. The former relates two cities
+$x$ and $y$ if there is a direct flight between them, and the latter tells
+whether $x$ is a US city or not.
 
-Using the  Local As View (LAV) approach~\cite{AmbiteISWC09}, concrete services in Table 1 are semantically described as views on the abstract services {\it flight} and {\it uscity} as follows:
+For the concrete services, assume that the available data sources on
+the Internet contain the following information:
+\begin{enumerate}[--]
+\item $\national(x,y)$ relates two US cities that are connected by a direct flight,
+\item $\oneway(x,y)$ relates two cities that are connected by a one-way flight,
+\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
+\item $\flightPA(x)$ tells if there is a direct flight from $x$ to Paris,
+\item $\onestopPA(x,y)$ relates $x$ and $y$ if there is a flight from $x$ to Paris
+      with a stop at $y$, and
+\item $\fromNY(x)$ tells if there is a flight from New York into $x$.
+\end{enumerate}
+Furthermore, the concrete services are described using the abstract services:
+\begin{alignat*}{1}
+\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
+\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
+\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
+\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
+\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
+\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
+\end{alignat*}
+Now suppose that a user is interested in constructing a workflow able to retrieve
+the one-stop round-trip flights from US cities to any city in the world, such that
+flights can stop at any city. 
+\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
+The following conjunctive query represents the workflow that defines this request
+in terms of abstract services. The workflow is defined in way that all issues
+about the binding of input/output parameters between services had been resolved
+such that any instantiation of the abstract services in terms of concrete services
+is a valid implementation of the workflow. 
+Implementations correspond to compositions of concrete services in which a concrete
+service may implement one or more abstract services from the workflow, but each
+abstract service can be implemented by exactly one concrete service.
+For example, the following composition corresponds to one such implementation.
+\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
+However, the following two compositions are not valid.
+\begin{alignat*}{1}
+I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
+I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
+\end{alignat*}
+The first composition is not valid because it maps the workflow variable $y$
+into constants \PA\ and \NY\ that denote different cities.
+On the other hand, $I''$ does not implement the workflow because the concrete
+service $\onestop(x,y)$ does not receive as input, or produce as output, the
+middle city where the flight stops and thus is not possible to ensure that 
+this city is bound to the city $w$ that is returned by the workflow.
 
-\begin{alignat*}{1}
-national(x1,y1)\ &\qrule\ flight(x1,y1),\,  uscity(x1),\,  uscity(y1)\,. \\
-oneway(x2,y2)\ &\qrule\ flight(x2,y2)\,. \\
-onestop(x3,z3)\ &\qrule\ flight(x3,y3),\, flight(y3,z3)\,. \\
-flightToCCS(x2,"CCS")\ &\qrule\ flight(x2,"CCS")\,. \\
-onestopToCCS(x3,y3)\ &\qrule\ flight(x3,y3),\, flight(y3,"CCS")\,. \\
-fromWAS("WAS",y3)\ &\qrule\ flight("WAS",x3)\,. \\
-\end{alignat*}
+These examples show that a proper workflow instantiations must handle
+constants such that not two different constants are mapped to each
+other either directly or indirectly via transitivity, and that all
+attributes that appear in a join or in the output need to be produced
+by the selected concrete services.
 
-Now suppose a user is interested in implementing a workflow able to retrieve the one-stop round trip flights from US cities to any city in the world, such that flights can stop at any city. 
-Consider the following conjunctive query that represents the workflow that defines this request in terms of abstract services.
-This workflow was defined in way the input parameters of each abstract service are bound by  constants in the workflow, or by attributes projected out by  previous abstract services.
-
-\begin{alignat*}{1}
-w(x,y1,y2,y3)\ &\qrule\ flight(x,y1),\, flight(y1,y2),\,  flight(y2,y3),\,  flight(y3,x),\,  \\
-                   \ & uscity(x)\,. \\
-\end{alignat*}
-
-Implementations of this workflow correspond to compositions of the concrete services where each concrete service can implement a subset of abstract services of the workflow, but each abstract service can be implemented by exactly one concrete service.
-The following composition of concrete services corresponds to one of the XX implementations of the workflow:
-
-\begin{alignat*}{1}
-wi1(x,y1,"CCS",y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\,  oneway("CCS",y3),\,  \\
-\ & national(y3,x)\,.  \\
-\end{alignat*}
-
-However, note that the following compositions are not valid:
-
-\begin{alignat*}{1}
-wi2(x,y1,"CCS"y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\, \\
-\ & fromWAS("WAS",y3),\,  national(y3,x)\,. \\
-wi3(x,y1,y2,y3)\ &\qrule\ national(x1,y1),\, onestop(x1,y2),\,  oneway(y2,y3),\,  \\
-\ & national(y3,x)\,. \\
-\end{alignat*}
-
-The composition {\it wi2} is not valid because  it maps the workflow variable
-{\it y2} into constants {\it "CCS"} and {\it "WAS"} which denote different cities.
-On the other hand, the composition {\it wi3} does not implement the workflow because the concrete service  {\it onestop} does not receive as input or produce as output the city where the flight stops (represented by the variable {\it y1}); thus, it is not possible to ensure that the city where this flight stops and the end city  retrieved by the service {\it national}, are the same.
-
-To avoid producing invalid workflow implementations, it is required to handle constants in a way that different constants are not mapped into
-each other either directly or indirectly (via transitivity).
-In addition,  all the attributes in the workflow output  or needed to join other services in the workflow, have to be produced by the selected concrete service.
-
-In addition, suppose each concrete service is annotated with a utility function that aggregates the values of different QoS that characterize the behavior of the service.
-Then, since the space of valid service compositions of an abstract workflow can be very large, we require a technique able to identify the ones that maximize the utility function  without having to enumerate the whole set of valid compositions.
-In this paper, we propose a Propositional logic-based  approach that takes advantage of the  power of  modern SAT solvers, to efficiently enumerate the service compositions that correspond to valid implementations of a given abstract workflow, as well as the compositions that maximize a given utility function.
+The QoS parameters are modeled by annotating the concrete services with
+utilities that characterize their behavior which are then aggregated
+during instantiation. Thus, as said before, the best instantiation is
+the one that minimizes (or maximizes) the aggregation of utilities.
 
 \section{Related Work}
-In this section we summarize existing approaches that provide solutions to the problems of service selection, knowledge representation and query rewriting.
 
-\begin{description}
-\item[Service Selection Solutions] \mbox{}\\
-The problem of selecting the services that implement an abstract workflow and best meet the QoS-based criteria  is known as the QoS-aware service selection or composition problem, which has been shown to be NP-hard~\cite{Hiroshi2008}.
-This problem is a combinatorial optimization problem and several heuristics have been proposed to find a relatively good solution in a reasonably short period of time.
-A  distance metric-based heuristic  to drive a backward search algorithm  is proposed~\cite{rahmani08}; this metric induces an order of the services in a way that sink nodes are unlikely to be visited.
-In ~\cite{berardi05,berardi08,berardi06}, services and workflows are described in terms of deterministic finite state machines that are encoded as a Description Logic theory whose models correspond to solutions of the problem;  although reasoning methods for Description Logics formalisms could be exploited, scalability or performance of the proposed solution has not been reported.
-~\cite{myoung08} propose a constraint-based approach that encodes the non-functional permissible values as a set of constraints whose violation needs to be minimized; to traverse the space of possibly optimal solutions, a hybrid algorithm that combines the tabu search and simulated annealing meta-heuristics is implemented; experimental results show that the proposed solution is able to scale up to a large number of services and abstract processes.
-In ~\cite{cardellini07} the QoS-aware service composition problem is encoded as a Linear Programming problem providing a scalable solution to the problem.
-In ~\cite{Hiroshi2008}  this problem  is defined as a multi-objective optimization problem where the different QoS parameters are considered equally important and there is not an aggregated function to combine all of them; a genetic-based algorithm is proposed to identify a set of non-dominated service compositions that best meet all the QoS parameters.
-~\cite{alrifaiR09} propose a two-fold solution that uses a hybrid integer programming algorithm to find the decomposition of global QoS into local constraint, and then,  selects the services that best meet the local constraints.
-Recently,  two new planning-based approaches have been proposed~\cite{kuterG09,sohrabiM09}.
-~\cite{kuterG09} extend the SHOP2 planning algorithm to select the trustworthy composition of services that implement a given OWL-S process model, while ~\cite{sohrabiM09}  propose a HTN planning-based solution where user preference metrics and domain regulations are used to guide the planner into the space of relevant compositions.
-Finally, ~\cite{lecue09} proposes a genetic-based algorithm to identify the composition of services that best meet the quality criteria for a set of QoS parameters.
+In this section we summarize the existing approaches that provide
+solutions to the problems of service selection, query rewriting and
+discuss related work in the area of Artificial Intelligence called
+knowledge compilation.
 
-Although these solutions are able to efficiently solve the optimization problem and scale up to a large number of abstract processes, none of them are tailored to semantically describe services in terms of the abstract process, or use these descriptions to identify the services that  implement a given workflow and best meet user non-functional criteria.
+\subsection{Service Selection}
 
-\item[Knowledge Compilation Languages] \mbox{}\\
-Knowledge compilation is the area in AI concerned with the
-problem of mapping logical theories into suitable fragments
-that make certain desired operations tractable
-\cite{cadoli:compilation}. Different compilation languages have been defined, for instance,  
-Ordered Binary Decision Diagrams (OBDDs)~\cite{bryant:obdd},  Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation Normal Form (DNNF)  \cite{darwiche:dnnf}.
-In this work we make use of the properties of the deterministic DNNFs  (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable and efficient solution to the service selection problem.
+The problem of selecting the services that implement an abstract workflow
+and best fit the QoS-based criteria is known as the QoS-aware service
+selection or composition problem, which has been shown to be NP-hard \cite{Hiroshi2008}.
+This problem is a combinatorial optimization problem and several heuristics 
+have been proposed to find a relatively good solution in a reasonably short
+period of time.
 
-A Negation Normal Form (NNF) theory is  constructed from literals using only conjunctions
-and disjunctions \cite{barwise:handbook}, and it can be represented as a directed acyclic graph
-in which the leaves are labeled with literals and the
-internal nodes are labeled with $\land$ and $\lor$;
-see Fig.~\ref{fig:dnnf} for an example. An NNF is said to be decomposable (DNNF) \cite{darwiche:dnnf}
-if for each conjunction $\bigwedge_i\phi_i$, the set of
-variables in each conjunct are pairwise disjoint;
-i.e,. $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i<j$.
-A DNNF supports a number of operations in
-polynomial time in the size of its DAG. For example,
-we can test whether a DNNF is satisfiable by a single
-bottom-up pass over its DAG in linear time. A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each disjunction $\bigvee_i\phi_i$, the disjuncts
-are pairwise logically contradictory; i.e.,
-$\phi_i\Rightarrow\neg\phi_j$ for $i<j$.
-The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable
-and deterministic. A d-DNNF supports the model counting and enumeration in polynomial time
-in the size of its DAG.
-{\it MCDSat}$^c$  exploits the properties of a d-DNNF theory, to efficiently provide a solution to the problem of enumerating the service compositions that correspond to models of the  theory, i.e., the service compositions that implement the abstract  workflow and best meet the utility function.
-DECIR ALGO SOBRE EL ALGORITMO QUE PRODUCE LOS MEJORES.
+Rahmani et al.\ \cite{rahmani08} present a distance metric-based heuristic
+that focus a backward search algorithm; this metric induces an order of the
+services in a way that sink nodes are unlikely to be visited.
+In a series of papers, Berardi and others \cite{berardi05,berardi08,berardi06}
+describe services and workflows in terms of deterministic finite-state machines
+that are encoded using Description Logic theories whose models correspond to
+solutions of the problem. Although reasoning methods for Description Logics
+formalisms could be exploited, scalability or performance of the proposed
+solution has not been reported.
 
-\begin{figure}
+Ko et al.\ \cite{myoung08} propose a constraint-based approach that encodes the
+non-functional permissible values as a set of constraints whose violation needs
+to be minimized; to traverse the space of possibly optimal solutions, a hybrid
+algorithm that combines tabu search and simulated annealing meta-heuristics is
+implemented. Experimental results show that the proposed solution is able to
+scale up to a large number of services and abstract processes.
+Cardellini et al.\ \cite{cardellini07} encode one part of the QoS-aware service
+composition problem as a Linear Programming problem \cite{cardellini07}.
+On the other hand, Wada el at.\ \cite{Hiroshi2008} treat the problem as a 
+multi-objective optimization problem where the different QoS parameters are
+considered equally important instead of aggregating them into a single function.
+Then, a genetic-based algorithm is proposed to identify a set of non-dominated
+service compositions that best fit all the QoS requirements.
+
+Alrifai and Risse \cite{alrifaiR09} propose a two-fold solution that uses a
+hybrid integer programming algorithm to find the decomposition of global QoS
+into local constraints, and then, selects the services that best meet the local
+constraints.
+
+Recently, two planning-based approaches have been proposed.
+Kuter and Golbeck \cite{kuterG09} extend the SHOP2 planning algorithm to select
+the trustworthy composition of services that implement a given OWL-S process model,
+while Sohrabi and McIlraith \cite{sohrabiM09} propose a HTN planning-based solution
+where user preference metrics and domain regulations are used to guide the planner
+into the space of relevant compositions.
+Finally, L\'ecu\'e \cite{lecue09} proposes a genetic-based algorithm to identify
+the composition of services that best meet the quality criteria for a set of QoS
+parameters.
+
+Although these solutions are able to solve the optimization problem and scale up
+to a number of abstract processes, none of them are tailored to semantically describe
+the services in terms of abstract process, nor to use these descriptions to 
+identify the services that implement a given workflow or best meet the user's
+non-functional criteria.
+
+\subsection{Query Rewriting}
+
+A number of algorithms have been developed to find the rewritings of a given query;
+the most prominent being the bucket algorithm \cite{levy:bucket}, the inverse rules
+algorithm \cite{pods:DuschkaG97,Qian96}, the \minicon\ algorithm \cite{pottinger:minicon},
+and the \mcdsat\ algorithm \cite{arvelo:aaai06}.
+Generally, query rewriting algorithms work in two phases. During the first phase,
+the algorithms identify the views that rewrite at least one sub-goal of the query, and
+during the second, these partial rewritings are combined into complete rewritings.
+The main difference between the algorithms is the criteria used to choose the relevant
+views to reduce the space of non-useful rewritings.
+
+The bucket algorithm reduces the number of possibilities by just considering each
+sub-goal in the query in isolation, and selecting the views that are able to produce
+at least the attributes projected by the query. Since the attributes involved in the
+joins in the query are not verified, a large number of rewritings comprised of Cartesian
+products may be generated.
+
+The Inverse Rules algorithm constructs a set of rules that invert the view definitions
+and establish how to compute tuples for the database relations from the tuples of the
+views. Similarly to the bucket algorithm, it can produce a large number of non-useful
+rewritings.
+
+The \minicon\ algorithm overcomes the limitations of the previous algorithms by identifying
+only views that rewrite a set of the sub-goals of the query, and that can be combined with
+the rest of the sub-goals. The key idea is to identify the mappings between the variables in
+each sub-goal to the variables in one or more sub-goals in the views, in a way
+that, join variables in the query are mapped to join variables in the body of a view or to
+the distinguished variables of the view. Mappings between variables and sub-goals are
+represented in the so-called MiniCon Descriptions (MCDs) \cite{pottinger:minicon}.
+
+Finally, the \mcdsat\ algorithm is able to identify the query rewritings of a query by
+translating the problem of rewriting into the problem of enumerating the models of a
+propositional theory whose models are in correspondence with the rewritings of the query.
+The algorithm exploits the properties of d-DNNFs to efficiently enumerate the models of
+the theory.
+The \mcdsat\ algorithm has demonstrated to scale better than the \minicon\ algorithm over
+a large number of benchmarks often showing performance improvements of several orders of
+magnitude. However, the \mcdsat\ algorithm was not designed for rewriting problems
+involving explicit constants, nor to compute the best rewritings with respect to a given
+utility function or cost model. In this paper, we propose an extended encoding that 
+overcomes these limitations and apply the encoding to the Workflow Instantiation Problem.
+
+\subsection{Knowledge Compilation}
+
+Knowledge compilation is the area in AI concerned with the problem of mapping
+logical theories into suitable fragments that make certain desired operations
+tractable \cite{cadoli:compilation}. Different compilation languages have been
+defined, for instance, Ordered Binary Decision Diagrams (OBDDs) \cite{bryant:obdd},
+Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation
+Normal Form (DNNF) \cite{darwiche:map}. In this work, we make use of the properties
+of the deterministic DNNFs (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable
+and efficient solution to the service selection problem.
+
+\begin{figure}[t]
 \centering
-\includegraphics[width=8cm]{odd}
+\includegraphics[width=.7\textwidth]{odd}
 \caption{A decomposable and deterministic NNF.}
@@
 
-\item[Query Rewriting Solutions] \mbox{}\\
-A number of algorithms have been developed to find the rewritings of a given query; the most prominent being the bucket algorithm~\cite{levy:bucket},  the inverse rules algorithm ~\cite{duschka:answer,Qian96}, the minicon algorithm ~\cite{pottinger:minicon}, and the {\it MCDSat} ~\cite{arvelo:aaai06}.
-Generally, query rewriting algorithms work in two phases: first, they identify the views that rewrite at least one  subgoal of the query; second they combine the selected views to produce a rewriting.
-The main difference between existing approaches, is the criteria used to choose the relevant views and reduce the space of non-useful rewritings.
+A Negation Normal Form (NNF) theory is constructed from literals using only conjunctions
+and disjunctions \cite{barwise:handbook}, and it can be represented as a directed
+acyclic graph in which the leaves are labeled with literals and the internal nodes
+are labeled with $\land$ and $\lor$; see Fig.~\ref{fig:dnnf} for an example.
+This fragment is universal meaning that for every logical formula there is an
+equivalent one in NNF format.
+An NNF is said to be decomposable (DNNF) \cite{darwiche:d-dnnfs} if for each conjunction
+$\bigwedge_i\phi_i$, the set of variables in each conjunct are pairwise disjoint; i.e,.
+$Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i\neq j$.
+A DNNF supports a number of operations in polytime in the size of its DAG.
+For example, we can test whether a DNNF is satisfiable by a single bottom-up pass
+over its DAG in linear time.
+A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each
+disjunction $\bigvee_i\phi_i$, the disjuncts are pairwise logically contradictory;
+i.e., $\phi_i\land\phi_j\equiv\textbf{false}$ for $i\neq j$.
+The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable and deterministic.
+A d-DNNF supports model counting in polytime in the size of its DAG, and model
+enumeration in polytime in the size of the output.
+Furthermore, given a literal ranking function $r$, one can compute the rank
+of the best model in polytime for DNNFs \cite{darwiche:weighted}.
 
-The bucket algorithm reduces the number of possibilities just considering each subgoal in the query in isolation, and selecting the views that are able to  produce at least the attributes projected by the query.
-Since, attributes involved in query joins are not verified, a large number of rewritings comprised of  Cartesian products may  be generated.
-The Inverse Rules algorithm constructs a set of rules that invert the view definition and establish how to compute tuples for the database relations from tuples of the views.
-Similarly, to the bucket algorith it can produce a large number non-useful of rewritings.
+The fragments DNNF and d-DNNF are universal yet translating a CNF theory 
+to DNNF format has an exponential cost in the worst case. This translation
+is referred to as compilation in the field. There is a publicly available
+compiler, called c2d,\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}
+that perform this compilation process and that make
+use of modern SAT techniques such as conflict-directed backtracking,
+clause learning and caching \cite{darwiche:compiler}.
+This compiler incurs in the worst case in exponential space in a parameter
+called the width of the decomposition tree that is related to the ``connectivity''
+of the CNF theory. However, in our experiments, the CNF theories that are
+compiled are of low width.
 
-The MiniCon algorithm  overcomes the limitations of the previous algorithms by identifying only views that rewrite a set of the query goals and that  can be combined with the rest of the query subgoals.
-The key idea is to identify the mappings between  the variables in each subgoal of the query to the variables  in one or more subgoals of the views, in a way that, join variables in the query are mapped to join variables in the body of a view or to the distinguished variables of the view.
-Mappings between variables and subgoals are represented in MiniCon Descriptions (MCD's)\cite{pottinger:minicon}.
+%DNNFs had also been used elsewhere to tackle other combinatorial optimization
+%problems such as problems in planning \cite{palacios:icaps05} and diagnosis
+%\cite{REF}.
 
-Finally,  the
-{\it MCDSat} algorithm is able to identify the query rewritings of a query by translating the problem of rewriting
-into the problem of enumerating the models of a propositional
-theory  $\Theory{\Q}$ whose models are in correspondence
-with the rewritings of the query. 
-The {\it MCDSat} algorithm exploits the properties of d-DNNFs to efficiently 
-compute the MCDs associated with theory $\Theory{\Q}$.
-The  {\it MCDSat} algorithm has demonstrated to scale better than the MiniCon
-algorithm over a large number of benchmarks often showing performance
-improvements of several orders of magnitude. However, the McdSat algorithm was not desgined for rewriting problems involving explicit constants,
-nor to compute the best rewritings with respect to a given utility function or cost model, and this paper we propose a new encoding  that overcomes these limitations.
-
-\end{description}
-
-\section{The McdSat$^c$ Architecture}
-
-In order to do so, the theory $\Theory{\Q}$ is transformed into a d-DNNF
-theory $\Delta(\Q)$ with a CNF to d-DNNF compiler.
-The compiler's algorithm is similar to the DPLL algorithm
-for SAT but enhanced with advanced techniques such as clause
-learning and caching.
-\begin{figure}
+\begin{figure}[t]
 \centering
-\includegraphics[width=10cm]{mdcsatC.pdf}
-\caption{The McdSat$^c$ Architecture}
-\label{fig:dnnf}
+\includegraphics[width=.9\textwidth]{architecture}
+\caption{System Architecture}
+\label{fig:architecture}
 \end{figure}
 
+\section{System Architecture}
+
+We use an architecture that is comprised of a Catalog of service descriptions,
+the Encoder, the compiler c2d, the best model Finder, and the Decoder.
+Figure~\ref{fig:architecture} depicts the overall architecture of the system.
+In this framework, an instance of Workflow Instantiation Problem is defined as an
+abstract workflow represented by a conjunctive query on abstract services which
+is given as input together with a set set of concrete services defined by views
+of abstract services.
+
+The Catalog is populated with descriptions of abstract and concrete services;
+each service is described in terms of input and output attributes, and annotated
+with a real value that represents the QoS utility of the service.
+The description of the concrete services, that are defined as views of abstract
+services, can be generated semi- or automatically using tools such as the DEIMOS
+system \cite{AmbiteISWC09}.
+
+An input instance of WIP is encoded as a CNF theory whose models corresponds to
+the instantiations of the workflow by the Encoder. The compiler c2d, an off-the-shelf
+component, compiles the CNF formula into d-DNNF.
+The Encoder translates the WIP instances into CNF theories, that are then converted
+into d-DNNF using c2d. The Finder computes a best model given the QoS parameters
+in linear time in the size of the resulting d-DNNF. It is important to remark,
+that the compilation process needs to be performed only once as it does not
+depend on the value of the QoS parameters. Thus, even if the compilation happens
+to be costly in terms of time, this cost can be amortized since the resulting
+d-DNNF can be used to find best instantiations with respect to multiple values
+of the QoS parameters.
+Finally, the Decoder translates the best model returned by the Finder into
+a workflow instantiation that solves the WIP.
+
+Given a CNF that encodes a WIP, its d-DNNF is a compact representation of
+all the workflow instantiations. That is, one can generate in a backtrack-free
+manner all the instantiations of the workflow. If the user is interested in
+a best instantiation given the QoS parameters, then it can be computed in
+linear time in the size of the d-DNNF. If the user is interested in the 
+all the best instantiations, these can be computed in linear time in the
+number of them. Finally, if the user is interested in all instantiations,
+these can be also computed in linear time in the number of them.
+In the latter two cases, if such number is exponential (in the size of the
+input), the enumeration of the instantiations is also exponential but
+this complexity is intrinsic to the problem and thus cannot be avoided.
+
+In order to make our results more accessible to a general audience, we decided
+to present the experimental results in the next section, and leave the formal
+and theoretical results for the following section. In this way, the interested
+reader may skip the formal details of the approach on a first reading.
+
 \section{Experimental Results}
-We have conducted an empirical analysis  on the benefits of the techniques implemented in the   McdSat$^c$ system.
 
-\begin{description}
-\item[Dataset and Query Benchmark:]  we conducted our experiments over a benchmark that includes XX queries and a set of  YY views.
-Queries and views are starts and chains, and they have between ZZ and TT constants; queries have QQ sub-subgoals.
-\item[Evaluation Metrics:] we report on runtime performance which corresponds to the  {\it user time} produced by the {\it time} command of the Unix operation system. 
-\end{description}
+We conducted an empirical analysis on three benchmarks.
+All the experiments were run on a desktop machine with an Intel
+Core 2 Duo 2GHz CPU and 4Gb of memory, and the time was measured
+with the Unix time command.
 
-\begin{figure}
+The objective of the experiment is to assess the performance
+of the proposal on varying conditions. The main benefit of our approach
+is that one can compile the logical theory for a problem instance
+and then calculate all the instantiations, or the best ones, any 
+number of times. The cost model for finding best instantiations
+can be changed with no need to recompile the logical theory.
+Therefore, the time complexity of our approach is basically the
+time to encode the WIP as a CNF plus the time to compile the
+CNF into a d-DNNF and the time to decode the models. 
+The times to encode and decode are negligible compared to
+the time to compile the CNF. Because of this, we focus on the
+time to compile the benchmark problems.
+
+The first benchmark consists of problems for air-travel queries.
+Concrete services are of the form $V_i(x,y)\qrule\ \flight(x,y,\AL_i)$
+where $\AL_i$ is a constant that denotes the name of an airline, 
+and $\flight(x,y,\AL_i)$ relates the cities $x$ and $y$ such that
+there is a flight between $x$ and $y$ served by $\AL_i$. This concrete
+service is assumed to return all flights between two cities with an
+specific airline. The workflow has the form 
+\[ W(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,z),\,\flight(x_1,x_2,z),\,\ldots,\,\flight(x_n,\NY,z)\,. \]
+
+\begin{figure}[t]
 \centering
-\includegraphics[width=8cm]{plot1.pdf}
-\caption{Compilation Time Benchmark I}
+\includegraphics[width=1\textwidth]{plots/plot1}
+\caption{Compilation times for experiments I and II for different
+number of goals and different number of views. The plots are in
+logarithmic scale, and the time is in seconds.}
 \label{fig:plot1}
@@
 
-\section{Formalization of the Service Selection Problem}
+The benchmark includes instances for workflows with 2 to 5 sub-goals and
+sets of 10 to 100 concrete services. The results for the compilation are
+shown in panel (a) of Fig.~\ref{fig:plot1}. This is a plot in logarithmic
+scale that suggests a sub-exponential behavior. In any case, the results
+show good performance since realistic instances of the problem (sets of 100
+airlines with 5-stop flights) can be compiled in 328 seconds.
+The size in disk of the d-DNNF for 100 airlines and 5-stop flights is 3.4Mb.
+On this d-DNNF, the best model can be computed in 0.29 seconds,
+and the enumeration of all models in 0.47 seconds.
 
-We consider databases of the form $D=\tup{P,T}$ where
-$P$ is a set of predicates and $T=\{T_p\}_{p\in P}$ is a collection
-of tables that represents the predicates in extensional form.
-A conjunctive query $Q$ over $P$ is of the form 
-\begin{alignat*}{1}
-Q(\vvec{x})\ \qrule\ \  p_1(\vvec{x}_1),\, p_1(\vvec{x}_2),\, \ldots,\, p_m(\vvec{x}_m)\,,
-\end{alignat*}
-where $p_i\in P$, $\vvec{x}$ is a vector of variables, and each
-$\vvec{x}_i$ is a vector of variables and constants.
-The result of $Q$ over $D$, denoted as $Q(D)$, is the table with
-$|\vvec{x}|$ columns that result of the projection of the relational
-join $\join\!\!\{T_{p_i}\}_{i=1}^m$ over $\vvec{x}$.
-The atoms in the body of $Q$ are called the (sub)goals of $Q$.
+In an attempt to induce an exponential growth in the compilation time,
+in the second benchmark we add a second concrete service for each airline.
+This modification increases the number of valid instantiations from linear
+to exponential since each leg of a flight can now be instantiated by two
+concrete services and thus a flight with $n$ legs may have up to $2^n$
+instantiations.
+We ran the compiler for instances comprising the same number of workflow
+sub-goals and total number of concrete services. The results plotted in
+logarithmic scale are shown in panel (b) of Fig.~\ref{fig:plot1}.
 
-A view $V$ over $D$ is a query over $P$. In the context of the service selection problem, 
-the database $D$ is an idealized description of 
-the output produced abstract workflow  implemented by multiple concrete services described
-as views.
-Given a database $D$, a query $Q$ and a collection of 
-views $E=\tup{\{V_i\}_i,\{E_i\}_i}$, we are required to find
-all the tuples in $Q(D)$ obtainable from the views in $E$.
-That is, we need to find all the \emph{compositions} of the form
-\begin{alignat*}{1}
-R(\vvec{x})\ \qrule\ \ V_{i_1}(\vvec{x}_1),\, V_{i_2}(\vvec{x}_2),\, \ldots,\, V_{i_n}(\vvec{x}_n)
-\end{alignat*}
-such that $R(E) \subseteq Q(D)$.
-A service selection problem (SSP) is a tuple $\tup{P,Q,\{V_i\}}$ where $P$
-is a set of predicates that represent abstract services, $Q$ is a query over $P$ and $\{V_i\}$ is a collection of views that define the concrete services in terms of  abstract services.
-We assume \emph{safe} problems in the sense that all
-variables mentioned in the head of the query (resp.\ in the head of each view)
-appear in the body of the query (resp.,\ in the body of each view); also the input and output restrictions of the abstract services used in the query are satisfied.
-Further, we only deal with SSPs with no arithmetic predicates inside the
-query or views.
-A composition $C$ is \emph{valid} if for all databases $D=\tup{P,T}$
-and extensions $\{E_i\}$, $R(E) \subseteq Q(D)$.
-A collection $\R$ of valid compositions is a solution if
-for all databases $D=\tup{P,T}$ and extensions $\{E_i\}$, there
-is no other $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(D)$.
-We are interested in finding a composition $\R$.
+These tests show good performance for this type of problems, but they do not
+involve concrete services with multiple sub-goals. We therefore designed a
+third experiment that consists unstructured, randomly generated instances.
+Each instance contains three variables per abstract service, ten distinct
+variables and ten distinct constants, six sub-goals in the workflows,
+2 to 5 sub-goals in the concrete services, and a varying number of services.
+The chance that an argument of an abstract service is bound to a constant is 50\%.
+The results are depicted in Fig.~\ref{fig:plot3}.
+The compilation time for these instances does not grow monotonically
+since they are randomly generated. The same happens for the size of
+the theories and the number of models. For example, the d-DNNF for 
+a problem with 45 views each with 5 sub-goals is of size 5.1Mb and has
+$1.26\times 10^8$ models. The time to find the best model for
+this d-DNNF is 0.46 seconds while the time to enumerate all models
+is about 17 hours.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=.8\textwidth]{plots/plot3}
+\caption{Compilation times for experiments III for different number of goals
+and different number of views. The plots are in logarithmic scale, and the time
+is in seconds.}
+\label{fig:plot3}
+\end{figure}
+
+These are preliminary experiments, yet the results show that the
+proposed approach efficiently scales for problems with several goals
+and views. We believe that these results are encouraging and motivate
+us to continue this research. We plan to conduct additional experiments
+with other types and sizes of workflows and, when possible, compare with
+other approaches.
+
+\section{Formalization}
+
+We consider service catalogs of the form $C=\tup{S,T}$ where $S$ is a set
+of predicates that represent abstract services and $T=\{T_s\}_{s\in S}$ is
+a collection of tables that represents the result of evaluating the abstract
+services. In the context of the WIP, a service catalog $C$ is an idealized
+description of the output produced by abstract workflows implemented by
+concrete services described as views.
+A workflow $W$ over $S$ is a conjunctive query of the form 
+\[ W(\vvec{x})\ \qrule\ s_1(\vvec{x}_1),\,s_1(\vvec{x}_2),\,\ldots,\,s_m(\vvec{x}_m)\,, \]
+where $s_i\in S$, $\vvec{x}$ is a vector of variables, and each $\vvec{x}_i$
+is a vector of variables and constants. The result of $W$ over $C$, denoted
+as $W(C)$, is the table with $|\vvec{x}|$ columns that result of the projection
+of the relational join $\join\!\!\{T_{s_i}\}_{i=1}^n$ over $\vvec{x}$.
+The atoms in the body of $W$ are called the (sub)goals of $W$, and the variables
+in the head of $W$ are called distinguished.
+
+A concrete service is described as a view $V$ over $C$ that, following LAV, is
+a query over $S$. Given a catalog $C$, a workflow $W$ and a collection of $n$ views
+$E=\tup{\{V_i\}_{i=1}^n,\{E_i\}_{i=1}^n}$, we are asked to find all the tuples in $W(C)$
+that can be obtained from the views in $E$. That is, we need to find all the
+\emph{instantiations}
+\[ R(\vvec{x})\ \qrule\ V_{i_1}(\vvec{x}_1),\,V_{i_2}(\vvec{x}_2),\,\ldots,\,V_{i_{m_i}}(\vvec{x}_{m_i}) \]
+such that $R(E) \subseteq Q(C)$.
+A Workflow Instantiation Problem is a tuple $\tup{S,W,\{V_i\}_i}$ where $S$
+is a set of predicates that represent abstract services, $W$ is a workflow over
+$S$, and $\{V_i\}_i$ is a collection of views that define the concrete services.
+We assume \emph{safe} problems in the sense that all variables mentioned in the
+head of the workflow (resp.\ in the head of each view) appear in the body of the
+workflow (resp.\ in the body of each view).
+Further, we only deal with WIPs with no arithmetic predicates inside the workflow
+or views. An instantiation $R$ is \emph{valid} if for all catalogs $C=\tup{S,T}$
+and extensions $\{E_i\}_i$, $R(E)\subseteq Q(C)$. A collection $\R$ of valid
+instantiations is a solution if for all service catalogs $C=\tup{S,T}$ and
+extensions $\{E_i\}_i$, there is no $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(C)$.
 
@@
 
-In \cite{arvelo:aaai06}, we showed that a rewriting
-for a query $Q$ with $m$ goals can be obtained by enumerating the
-models of a logical theory
-$\Delta=\Delta_{rew}\cup\Delta_{mcd}^1\cup\cdots\Delta_{mcd}^m$
-where $\Delta_{rew}$ specified how to combine $m$ independent
-copies of MCD theories $\Delta_{mcd}$ that cover all goals in $Q$.
-Each $\Delta^i_{mcd}$ is a copy of the theory $\Delta_{mcd}$
-in which each literal $p$ is tagged as $p^i$.
-The theory $\Delta_{mcd}$ consists of different groups of 
-clauses that guarantees that its models are in correspondence
-with the MCDs, while the theory $\Delta_{rew}$ contains additional
-clauses to guarantee a sound and complete composition of the MCDs.
-The reader is referred to \cite{arvelo:aaai06} for a comprehensive
-description of the propositional theory.
+We use an approach similar to the one described in \cite{arvelo:aaai06} to
+encode the WIP. We have limited space to make a comprehensive description of
+the logical theory so the reader is referred there for details and formal results.
 
-Let us describe the difficulties that arise when constants
-are presents in a SSP with the following example:
-\begin{alignat*}{1}
-Q(x,z)\ &\qrule\ \ p_1(x,y),\, p_2(y,z)\,, \\
-V_1(x_1)\ &\qrule\ \ p_1(x_1,A)\,, \\
-V_2(x_2)\ &\qrule\ \ p_2(B,x_2)\,, \\
-V_3(x_3,y_3)\ &\qrule\ \ p_2(x_3,y_3)
-%V_4(x_4,y_4)\ &\qrule\ \ p_1(x_1,A),\, p_2(B,y_4)\,,
-\end{alignat*}
-where $Q$ is the query and $\{V_1,V_2,V_3\}$ are the 
-views. In this case, the only rewriting is 
-$R(x,z)\qrule V_1(x),V_3(A,z)$ because the candidate 
-$R(x,z)\qrule V_1(x),V_2(z)$ is not valid as it maps
-$y$ into constants $A$ and $B$ that denote different
-objects.
+We identify workflow instantiations by enumerating the models of a logical
+theory $\Delta=\Delta_{com}\cup\Delta_{id}^1\cup\cdots\Delta_{id}^N$
+where $\Delta_{com}$ specifies how to combine $N$ independent copies theories
+$\Delta_{id}$ that cover the goals in $W$.
+Each $\Delta^i_{id}$ is a tagged copy $\Delta_{id}$ in which each literal
+$\ell$ is tagged as $\ell^i$.
+The Instantiation Description (ID) theory $\Delta_{id}$ consists of different
+groups of clauses that guarantees that its models are in correspondence with
+partial instantiations, while the theory $\Delta_{com}$ contains additional
+clauses to guarantee a sound and complete combination of partial instantiations
+into a complete instantiation.
+The ID theory $\Delta_{id}$ consists of the following variables:
+\begin{enumerate}[--]
+\item $\{v_0,\ldots,v_n\}$ to indicate which $V_i$ is used, or $v_0$ to indicate the null ID.
+\item $\{g_1,\ldots,g_m\}$ to indicate the goals covered by the view.
+\item $\{z_{j,k,i}\}$ to indicate that the $j$th goal in $W$ is covered by the $k$th goal in $V_i$.
+\item $\{t_{x,y}\}$ to indicate that the variable/constant $x$ in $W$ is mapped into the
+      variable/constant $y$ in the view.
+\end{enumerate}
+The ranges of the indices for the $z$ and $t$ variables depend on the problem.
 
-The main problem when handling constants is to
-be sure that different constants are not mapped into
-each other either directly or indirectly (via transitivity).
+The following clauses encode the WIP problem in terms of the WIDs.
+Rajaraman et al.\ \cite{RajaramanSU95} showed that for queries without negation or
+arithmetic comparisons, but with constants, and $m$ goals and $k$ variables in the
+head of the workflow, it is enough to consider instantiations of length at most $N=m+k$.
+\begin{enumerate}[C10.]
+\item[C1.] (At least of view is used): $\bigvee_{i=0}^n v_i$.
+\item[C2.] (At most one view is used): $\neg v_i\lor\neg v_j$ for $i\neq j$.
+\item[C3.] (Null view equals null): $v_0 \Rightarrow \neg g_j$ for $1\leq j\leq m$.
+\item[C4.] (Views are useful): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ for $1\leq i\leq n$.
+\item[C5.] (Subgoals covered at most once): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ for appropriate $i,j,k,l$.
+\item[C6.] (Scope of views): $v_i \Rightarrow \neg g_j$ for goals that cannot be covered by $V_i$.
+\item[C7.] (Consistency): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ for appropriate $i,j,k$.
+\item[C8.] (Dead variables): $v_i \Rightarrow \neg t_{x,y}$ for all $x,y$ with $y\notin V_i$.
+\item[C9.] (1-1 for $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ for all
+           existential variables $y,y'\in V_i$.
+\item[C10.] (Distinguished): $v_i \Rightarrow \neg t_{x,y}$ for distinguished $x$ and existential
+            $y\in V_i$.
+\item[C11.] (Existential): $v_i\land t_{x,y}\Rightarrow g_j$ for exist.\ $y\in V_i$ and
+            goals $g_j$ that contain $x$.
+\item[C12.] (Match): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ for all $x,y$ that must match
+            if $g_j$ is covered by goal $k$ in $V_i$.
+\item[C13.] (If all vars in $V_i$ are distinguished, it covers only one goal):
+            $v_i \land g_j \Rightarrow \neg g_k$ for appropriate views $v_i$.
+\end{enumerate}
+These clauses are the same clauses used by \mcdsat\ to encode QRPs \cite{arvelo:aaai06}.
+In order to properly manage constant symbols, the clauses must be enhanced with:
+\begin{enumerate}[C10.]
+\item[C14.] (Direct inconsistency 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
+\item[C15.] (Direct inconsistency 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
+\item[C16.] (Direct inconsistency 3): $\neg t_{A,B}$.
+\item[C17.] (Transitivity 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
+\item[C18.] (Transitivity 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
+\end{enumerate}
+Recall that, as shown in Section~2, the main issue when handling
+constants is to be sure that not two different constants are mapped
+into each other either directly or indirectly.
+Clauses C14--C16 prune direct inconsistent mappings, while the last
+two clauses implement a restricted propagation of mappings that prune
+indirect inconsistencies.
+Likewise, the theory $\Delta_{com}$ that specifies how a complete
+instantiation contains the following clauses:
+\begin{enumerate}[C10.]
+\item[C19.] (Cover all goals): $\bigvee_{j=1}^m g^i_j$ for $1\leq i\leq N$.
+\item[C20.] (Disjunctive cover): $g^i_k \Rightarrow \neg g^j_k$ for $i\neq j$.
+\item[C21.] (Prune symmetries): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
+            for $1\leq j\leq m$ and $1\leq i\leq N$.
+\item[C22.] (Direct inconsistency 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
+\end{enumerate}
+These clauses provide sound and complete characterization of WIPs in the
+sense that their models are in correspondence with the
+instantiations of the workflow.
 
-\subsection{Utility Functions}
+\subsection{QoS Parameters}
 
-We assume a simple additive utility function in which each view
-$V_i$ is associated with a utility measure $c(V_i)$, and the overall utility value of
-a composition is the sum of the utility values for the views in it.
-An optimal or best service composition is one with maximum utility value,
-and the optimal utility value of a SSP is the utility value of a best
-service composition. A SSP has always a well-defined optimal
-utility value (if there are no compositions, its utility is $0$),
-but it may have multiple best service compositions.
-The service composition problem with utility consists in finding
-all the compositions of maximum utility value.
-
-%The cost model is simple yet expressive; e.g., if all
-%costs are unit, then best rewritings are those with
-%minimum number of views, yet in cases where views
-%represent different sources, it is natural
-%to associate them with different costs.
-
-\subsection{Extended Theories}
-
-The theory $\Delta_{mcd}$ makes use of propositions $t_{x,y}$
-to denote that the variable/constant $x$ in the query is mapped
-into the variable/constant $y$ in the view, and propositions $v_i$
-to indicate that the MCD uses view $V_i$.
-We obtain a sound and complete theory for SSPs with constants if
-$\Delta_{mcd}$ is extended with the clauses:
-
-\begin{enumerate}[C1.]\denselist
-\item (Inconsistent-1): $t_{x,A} \Rightarrow \neg t_{x,B}$,
-\item (Inconsistent-2): $t_{A,x} \Rightarrow \neg t_{B,x}$,
-\item (Inconsistent-3): $\neg t_{A,B}$,
-\item (Transitivity-1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$,
-\item (Transitivity-2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
-\end{enumerate}
-Clauses C1--C3 prune candidate service compositions in which one constant
-is directly mapped into a different one, and the last two implement
-a restricted propagation of transitivity.
-Similarly, $\Delta_{rew}$ must be extended with the clauses:
-\begin{enumerate}[C1.]\denselist
-\item[C6.] (Inconsistent-4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
-\end{enumerate}
-~\cite{RajaramanSU95} showed that for queries without negation or arithmetic comparisons,  but with constants, and $m$ goals and $q$ variables, it is enough
-to consider service compositions of length at most  $m$ plus  $q$ subgoals ~\cite{Ullman00}.
-
-Given a SSP problem $\tup{Q,\{V_i\}}$ possibly with constant 
-symbols, we construct logical theories whose models are in
-correspondence with the service compositions. These theories are then
-compiled into d-DNNF from which all rewritings are extracted
-efficiently.
-Likewise, best service compositions can be computed by performing model
-enumeration twice: the first pass computes the optimal utility function,
-and the second filters out suboptimal service compositions.
-However, there is a better way, which we will describe in the next section.
-
-\subsection{Maximum-Utility Function}
-
-Darwiche \& Marquis \cite{darwiche:weighted} show how to compute the
-rank $r^*(\Delta)$ of a propositional theory $\Delta$
-efficiently when $r$ is a literal ranking function and
-$\Delta$ is in d-DNNF. A literal ranking function ranks
-the models of the theory in terms of the rank of the
-literals $l$ that are true in the model, and ranks the
-theory as the best rank of its models:
-\[ r(\omega) = \sum_{\omega\vDash l} r(l),\quad r^*(\Delta) = \max_{\omega\vDash\Delta} r(\omega)\,. \]
-Furthermore, not only $r^*(\Delta)$ can be computed
-efficiently from the d-DNNF but also the models that
-have such rank.
-The procedure for computing the rank converts the d-DNNF
-into a circuit in which the literals are replaced by their
-ranks, the `or' nodes by `max', and the `and' nodes by `sum'.
-The evaluation of the circuit computes the rank of $\Delta$.
-
-We thus obtain a simple method for computing the best
-rewritings when the literal ranking function $r_c$ induced
-by the cost model $c$ is used; $r_c$ is defined by
-$r_c(l)=c(V_i)$ if $l=v^t_i$ for some $t$, and $r_c(l)=0$
-otherwise.
+For the QoS parameters, we assume a simple additive aggregation model in
+which each view $V_i$ is associated with a cost $c(V_i)$ (negative if utility),
+and a complete instantiation with the sum of the cost of its views.
+An optimal or best instantiation is one with minimum cost, and the
+optimal value of the WIP is the cost of an optimal instantiation.
+A WIP always has a well-defined optimal value (if there are no instantiations,
+its cost is $\infty$), but it may have multiple best instantiations.
+The WIP with costs consists in finding all optimal instantiations or
+one instantiation, this depends on the particular application.
+In our formulation, this can be done from the d-DNNF that encodes $\Delta$
+using the literal ranking function $r$ that assigns $r(\ell)=c(V_i)$ if
+$\ell=v_i$, and $r(\ell)=0$ if $\ell\notin\{v_1,\ldots,v_n\}$.
 
@@
 
-We have shown how the propositional theory used in McdSat can
-be extended to support constants, and how it can be used to 
-compute the best rewritings with respect to an additive
-cost model. The cost model is simple yet expressive.
-The computation of the best models relies on an arithmetic
-circuit that is obtained from the d-DNNF of the propositional
-theory.
+We have shown how the propositional theory used in \mcdsat\ for
+computing rewritings of queries can be extended to support constants
+symbols and thus adapted to the problem of instantiating abstract
+workflows. In this work, we adopted a LAV formulation of the
+workflow instantiation problem typically found in SOAs.
+Our formulation assumes that all concrete services are described
+using views over abstract services and that the workflow is 
+expressed as a conjunctive queries over the abstract services.
+This workflow has resolved all the issues related to the
+binding of input and output parameters of services. What
+remains is the problem of finding the best instantiation
+of the workflow in terms of the QoS parameters.
+This formulation of the problem is supported by systems 
+like DEIMOS that are able to generate in semi- or
+automatic manner LAV descriptions as the one required.
+
+The experimental results show that the approach can be applied to
+real-sized workflow problems. The whole approach is only possible
+when the compilation of the CNF theory into d-DNNF format succeeds.
+In the future, we plan to address this limitation by adapting a
+branch-and-bound optimization algorithm that searches the combinatorial
+space of instantiations but prunes suboptimal branches with admissible
+heuristics.
+
+We also are interested in using other target compilation languages
+like Ordered Binary Decision Diagrams.
 

### 2010-01-10 tesis/propuesta/traduccion/traduccion.tex
abstract

\documentclass{article}

\usepackage[utf8x]{inputenc}
\usepackage[spanish]{babel}

\begin{document}

\title{Instanciación Óptima de Flujos de Trabajo Abstractos Usando Lógica y Circuitos}
\maketitle

\section{Resumen}

La Arquitectura Orientada a Servicios típica consiste de dos capas principales.
La capa concreta con servicios concretos cuyas funcionalidades son descritas en
términos de pre- y post-condiciones y propiedades no funcionales en términos de
parámetros de calidad de servicio (Quality Of Service, QoS), y la capa abstracta
con aplicaciones de software cuyas funcionalidades son descritas en términos de
flujos de trabajo abstractos y propiedades no funcionales en términos de
restricciones sobre QoS. Durante la ejecución de un flujo de trabajo, los
servicios abstractos son instanciados en servicios concretos que cumplen con los
requerimientos funcionales y no funcionales. Esta instanciación, que debe ser
hecha en el momento, consiste de una búsqueda en un espacio de posibilidades
combinatorio. En este trabajo se propone una infraestructura para resolver
eficientemente el problema de instanciación que está firmemente basada en
lógica. La infraestructura adopta la aproximación Local-As-View en la cual la
funcionalidad de servicios concretos se describe usando vistas de servicios
abstractos, la calidad de una instanciación como una función de utilidad global
que combina los diferentes parámetros QoS, y los flujos de trabajo abstractos
como consultas conjuntivas sobre servicios abstractos. Usando esta
representación, el problema de instanciación de flujos de trabajo se convierte
en un problema de reescritura de consultas, del área de sistemas de integración.
Entonces, construyendo sobre trabajo relacionado, se idea una codificación del
problema de instanciación de flujos de trabajo como una teoría lógica cuyos
modelos están en correspondencia con las instanciaciones del flujo de trabajo, y
los modelos mejor puntuados en correspondencia con instanciaciones óptimas del
flujo de trabajo. Así, explotando propiedades conocidas de teorías lógicas en
formato d-DNNF, se provee una solución eficiente y escalable al problema de
instanciación de flujos de trabajo. Esta aproximación no sólo escala a
instancias grandes como muestran resultados experimentales ya realizados, sino
que también es correcta y completa dado que, estándo basada en la lógica, es
amigable al análisis formal.

\end{document}

### 2010-01-11 tesis/propuesta/traduccion/traduccion.tex
intro avanzada

@@
 
+\section{Introducción}
+
+Bajo la sombra de la Web Semántica, y soportado por Arquitecturas Orientadas a
+Servicios (SOA), el número de fuentes de datos y servicios Web ha explotado en
+los últimos años. Por ejemplo, la colección de bases de datos de biología
+molecular actualmente contiene 1170 bases de datos XREF22W, un número que es
+mayor que el del último año XREF21W por 95, y que el de hace dos años por 110
+XREF20W; las herramientas y servicios, así como el número de instancias
+publicados por estos recursos, siguen una progresión similar XREF6W. Gracias a
+este tesoro de datos, la tendencia de los usuarios es hacia depender cada vez
+más en métodos automáticos para manejarlos tales como recuperación de datos de
+fuentes públicas y análisis utilizando herramientas o servicios Web compuestos
+en flujos de trabajo complejos.
+
+La SOA típica consiste de dos capas. La capa concreta que está hecha de
+servicios concretos cuyas funcionalidades son descritas en términos de pre- y
+post-condiciones y propiedades no funcionales en términos de parámetros QoS, y
+la capa abstracta compuesta de aplicaciones de software cuyas funcionalidades
+son descritas en términos de flujos de trabajo abstractos y propiedades no
+funcionales en términos de restricciones de QoS. La ejecución de un flujo de
+trabajo abstracto involucra la instanciación de los servicios abstractos en
+servicios concretos que cumplen con los requerimientos funcionales y no
+funcionales. Este proceso de instanciación puede ser visto como una búsqueda de
+una instanciación objetivo en el espacio combinatorio de todas las
+instanciaciones válidas. Así, uno está interesado en técnicas eficientes para
+realizar esta búsqueda que sean capaces de escalar a medida que el número de
+servicios concretos o la complejidad del flujo de trabajo aumenta. Se llama al
+problema de instanciar un flujo de trabajo dado con servicios concretos, de un
+conjunto dado de servicios concretos, de manera que ciertas demandas de QoS sean
+cumplidas, el Workflow Instantiation Problem (WIP, "Problema de Instanciacion de
+Flujos de Trabajo").
+
+En este trabajo se considera una versión restringida de WIP que adopta la
+aproximación Local-As-View (LAV) XREF28W. En LAV, todos los elementos de un
+problema son especificados con un lenguaje común basado en servicios abstractos
+tal que los servicios concretos se describen como vistas de servicios
+abstractos, la calidad de una instanciación como una función de utilidad global
+que combina los diferentes parámetros QoS, y el flujo de trabajo abstracto como
+una consulta conjuntvia sobre los servicios abstractos; esta representación es
+similar a la que es generada de manera semiautomática para el sistema DEIMOS
+XREF3W. En esta versión de WIP, las reglas que definen flujos de trabajo
+(consultas conjuntivas) y servicios concretos (vistas) son creados de manera que
+todas las restricciones funcionales sobre pre- y post-condiciones de los
+servicios y sus combinaciones sean satisfechas, y las medidas de QoS son
+representadas anotando cada descripción de servicio concreto con un número real
+que representa la utilidad QoS global del servicio.
+
+Bajo estas suposiciones, WIP puede ser convertido en el bien conocido Problema
+de Reescritura de Consultas (QRP) para LAV que es central a los sistemas de
+integración XREF23W. QRP consiste de una consulta conjuntiva que debe ser
+respondida en términos de vistas donde la consulta y las vistas son descritas
+usando LAV con relaciones abstractas. Este problema es importante en el contexto
+de integración de datos XREF13X24W, y optimización de consultas y mantenimiento
+de datos XREF1,28W, y varias aproximaciones que escalan a un número grande de
+vistas han sido definidas XREF4,18,19,28,29W.
+
+La aproximación recente de Arvelo et al. XREF4W está basada en la enumeración
+eficiente de modelos de una teoría lógica proposicional. Dado unQRP, una teoría
+lógica es construída de manera que cada modelo de la teoría codifica una
+reescritura válida, y así todas las reescrituras son obtenidas enumerando los
+modelos de la teoría. Esta enumeración puede ser realizada eficientemente si la
+teoría lógica está en cierta forma normal llamada la forma normal de negación
+determinística y descomponible (d-DNNF por sus siglas en inglés) XREF14W. Así,
+la aproximación consiste en transformar (llamado compilar en el campo de
+compilación del conocimiento) la teoría lógica en el formato d-DNNF para
+enumerar sus modelos eficientemente.
+
+Pero las teorías d-DNNF no sólo soportan la enumeración eficiente de sus
+modelos, sino también otras operaciones. Entre ellas se encuentra la enumeración
+de los modelos con mejor puntuación. Dada una función de puntuación de literales
+$r(l)$ que asigne puntuaciones a cada literal $l$, se define la puntuación
+$r(w)$ de un modelo $w$ como la suma de las puntuaciones de los literales
+activados por $w$ (es decir, $r(w) = suma...$, y se dice que $w$ es un modelo
+mejor puntuado si no hay modelo $wpriemrie$ tal que $r(w)...$.
+
+Dada una teoría en d-DNNF, se computa la puntuación de los mejores modelos, y
+los mejores modelos, en tiempo lineal en el tamaño del d-DNNF. Este cómputo
+transforma el GAD del d-DNNF en un circuito aritmético reemplazando los nodos
+AND con '+' y los nodos OR con 'min'. La función de puntuación de literales
+asigna valores a las hojas del circuito que son propagados a la raíz en tiempo
+lineal. El valor de la raíz es la puntuación del mejor modelo XREF17W.
+
+En este trabajo, se explotan las propiedades de los d-DNNF construyendo una
+teoría lógica cuyos modelos codifican las instanciaciones del flujo de trabajo
+
 \end{document}

### 2010-01-11 tesis/propuesta/traduccion/traduccion.tex
mas avanzado

@@
 teoría lógica cuyos modelos codifican las instanciaciones del flujo de trabajo
+y cuyos mejores modelos codifican las instanciaciones óptimas (mejores). Así, la
+búsqueda combinatoria se reduce a la computación de un mejor modelo de una
+teoría lógica que puede ser realizada eficientemente una vez que la teoría sea
+transformada en formato d-DNNF.
+
+Esta propuesta contiene seis secciones más. La próxima sección ilustra y motiva
+WIP con un ejemplo simple pero típico. La sección 3 resume el trabajo existente
+relacionado en tres áreas de selección de servicios, reescritura de consultas y
+compilación del conocimiento. Entonces, las secciones 4 y 5 describen la
+arquitectura del sistema y reportan los resultados empíricos obtenidos hasta
+ahora sobre distintos problemas de prueba respectivamente. La sección 6 presenta
+una descripción formal de la aproximación propuesta junto con un análisis. La
+propuesta concluye con una discusión final en la sección 7.
+
+\section{Ejemplo motivante}
+
+Considere un sistema simple de información de vuelos que contiene información
+sobre vuelos entre ciudades e información sobre qué ciudades están en los
+Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
+abstractos $flight$ y $uscity$. El primero relaciona dos ciudades $x$ e $y$ si
+hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
+E.E.U.U.  o no. Para los servicios concretos, supongamos que las fuentes de
+datos disponibles en Internet contienen la siguiente información:
+
+\begin{itemize}
+TODO
+-national
+\end{itemize}
+
+Adicionalmente, los servicios concretos se describen con los siguientes
+servicios abstractos:
+
+\begin{itemize}
+TODO
+-national
+\end{itemize}
+
+Ahora supongamos que un usuario está interesado en construir un flujo de trabajo
+capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
+en el mundo, tales que los vuelos puedan detenerse en cualquier ciudad.
+
+$$
+W(x,w,y,z)
+$$
+
+La siguiente consulta conjuntiva representa el flujo de trabajo que define esta
+solicitud en términos de servicios abstractos. El flujo de trabajo es definido
+de manera que todo lo relacionado sobre el ligamiento de parámetros de
+entrada/salida es resuelto. Así, cualquier instanciación de los servicios
+abstractos en términos de los servicios concretos es una implementación válida
+del flujo de trabajo. Las implementaciones corresponden a composiciones de
+ervicios concretos en las cuales un servicio concreto puede implementar uno o
+más servicios abstractos del flujo, pero cada servicio abstracto puede ser
+implementado por exactamente un servicio concreto. Por ejemplo, la siguiente
+composición corresponde a una de estas implementaciones.
+
+$$
+I(x
+$$
+
+Sin embargo, las siguientes dos composiciones no son válidas.
+
+$$
+I'
+I''
+$$
+
+La primera composición no es válida porque asocia la variable del flujo $y$ a
+las constantes PA y NY que denotan ciudades diferentes. En la otra mano, $I''$
+no lo implementa porque el servicio concreto $ones-stop$ no recibe como entrada
+o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces
+no es posible asegurarse de que esa ciudad esté asociada a la ciudad $w$ que es
+retornada por el flujo de trabajo.
+
+Estos ejemplos muestran que las instanciaciones correctas de un flujo deben
+manejar constantes de manera que ods constantes distintas no sean asociadas
+entre sí de manera directa o indirectamente vía transitividad, y que todos los
+atributos que aprezcan en un TODOemphjoin o en la salida sean producidos por los
+servicios concretos seleccionados.
+
+Los parámetros QoS son modelados anotando los servicios concretos con utilidades
+que caracterizan su comportamiento que luego son agregados durante la
+instanciación. Así, como se mencionó previamente, la mejor instanciación es la
+que minimiza (o maximiza) la agregación de utilidades.
+
+\emph{Trabajo Relacionado}
+
+En esta sección se resumen las aproximaciones existentes que proveen soluciones
+a los problemas de selección de servicios y reescritura de consultas y se
+discute el trabajo relacionado en el área de inteligencia artificial llamado
+compilación del conocimiento.
+
+\subsection{Selección de Servicios}
+
+El problema de seleccionar los servicios que implementan un flujo de trabajo
+abstracto y cumplen mejor los criterios basados en QoS es conocido como el
+problema de selección o composición de servicios consciente de QoS, que ha sido
+demostrado ser NP-difícil XREF34W. Este es un problema de optimización
+combinatoria y varias heurísticas han sido propuestas para conseguir una
+solución relativamente buena en un período razonablemente corto de tiempo.
+
+Rahmani et al. XREF31W presenta una heurística basada en métrica de distancia
+que guía un algoritmo de búsqueda hacia atrás; esta métrica induce un orden de
+los servicios en una manera en que los nodos vertedero tienen baja probabilidad
+de ser visitados. En una serie de artículos, Berardi y otros XREF7-9W describen
+servicios y flujos de trabajo en términos de máquinas de estados finitos
+determinísticas que son codificadas usando teorías de Lógica de Descripción
+cuyos modelos corresponden a soluciones del problema. Aunque se podrían explotar
+métodos para formalismos de Lógicas de Descripción, no han sido reportados la
+escalabilidad o el rendimiento de la solución propuesta.
+
+Ko et al. XREF25W propone una aproximación basada en restriccciones que codifica
+los valores permisibles no funcionales como un conjunto de restricciones cuya
+violación debe ser minimizada; para recorrer el espacio de soluciones
+posiblemente óptima, se implementa un algoritmo híbrido que combina las
+metaheurísticas TODOemphtabusearch y TODOemphsimulatedannealing. Los resultados
+experimentales muestran que la solución propuesta es capaz de escalar a un
+número grande de servicios y procesos abstractos. Cardellini et al. XREF12W
+codifican una parte del problema de composición de servicios consciente de QoS
+como un problema de programación lineal XREF12W. Por otra parte, Wada et al.
+XREF34W trata el problema como uno de optimización con múltiples objetivos donde
+los diferentes parámetros de QoS son considerados igualmente importanes en vez
+de ser agregados en una función. Luego se propone un algoritmo basado en
+algoritmos genéticos que identifica un conjunto de composiciones de servicios
+no dominadas que mejor encajan en los requerimientos QoS.
+
+Alrifai y Risse XREF2W proponen una solución de doble cara que usa un
+algoritmo de programación entera híbrido para obtener la descomposición de QoS
+globales en restricciones locales y luego selecciona los servicios que mejor
+cumplen las restricciones locales.
+
+Recientemente, dos aproximaciones basadas en planificación han sido propuestas.
+Kuter y Golbek XREF26W extienden el algoritmo de planificación SHOP2 para
+seleccionar la composición confiable de servicios que implementan un modelo de
+procesos OWL-S dado, mientras que Sohrabi y McIlraith XREF33W proponen una
+solución basada en planificación HTN donde las métricas de preferencia del
+usario y regulaciones de dominio son utilizadas para guíar el planificador hacia
+el espacio de composiciones relevantes. Finalmente, Lécué XREF27W propone un
+algoritmo basado en genética para identificar la composición de servicios que
+mejor cumple con los criterios de calidad para un conjunto de parámetros QoS.
+
+Aunque estas soluciones son capaces de resolver el problema de optimización y
+escalar a un número de procesos abstractos, ninguno de ellos está ajustado para
+describir semánticamente los servicios en términos de procesos abstractos, ni
+para usar estas descripciones para identificar los servicios que implementan un
+flujo de trabajo dado o que mejor cumplen los criterios no funcionales del
+usuario.
 

### 2010-01-11 tesis/propuesta/traduccion/traduccion.tex
pag6

@@
 
+\subsection{Reescritura de Consultas}
+
+Un número de algoritmos han sido desarrollados para encontrar las reescrituras
+de una consulta dada; los más prominentes son el algoritmo bucket XREF28W, el
+algoritmo de reglas inversas XREF18,30W, el algoritmo MiniCon XREF29, y el
+algoritmo McdSat XREF4W. Generalmente, la reescritura de consultas trabaja en
+dos fases. Durante la primera, el algoritmo identifica las vistas que reescriben
+al menos un subobjetivo de la consulta, y durante el segundo estas reescrituras
+parciales son combinadas en reescrituras completas. La diferencia principal
+entre los algoritmos es el conjunto de criterios usado para elegir las vistas
+relevantes para reducir el espacio de reescrituras no útiles.
+
+El algoritmo bucket reduce el número de posibilidades considerando sólo cada
+subobjetivo de la consulta aislado, y seleccionando las vistas que son capaces
+de producir al menos los atributos proyectados por la consulta. Como los
+atributos involucrados en los joins de la consulta no son verificados, un número
+grande de reescrituras comprendido por productos cartesianos puede ser generado.
+
+El algoritmo de reglas inversas construye un conjunto de reglas que invierten
+las definiciones de vistas y establecen cómo computar tuplas para las relaciones
+de la base de datos desde las tuplas de las vistas. Al igual que el algoritmo
+bucket, puede producir un gran número de reescrituras no útiles.
+
+El algoritmo MiniCon supera las limitaciones de los algoritmos previos
+identificando solamente vistas que reescriben un conjunto de los subobjetivos de
+la consulta, y que pueden ser combinados con el resto de los subobjetivos. La
+idea principal es identificar las asociaciones entre variables de cada
+subobjetivo a las variables en uno o más subojbetivos en las vistas, de manera
+que las variables de join de la consulta son asociadas a las variables de join
+del cuerpo de una vista o a variables distinguidas de la vista. Asociaciones
+entre variables y subojbetivos son representadas en las llamadas TODOemphMiniCon
+Descriptions (MCDs) XREF29W.
+
+Finalmente, el algoritmo McdSat es capaz de identificar las reescrituras de una
+consulta traduciendo el problema de reescritura al problema de enumeración de
+modelos de una teoría proposicional cuyos modelos están en correspondencia con
+las reescrituras de la consulta. El algoritmo explota las propiedades de los
+d-DNNF para enumerar eficientemente los modelos de la teoría. El algoritmo
+McdSat ha demostrado escalar mejor que el algoritmo MiniCon sobre un número
+grande de experimentos donde frecuentemente muestra mejoras de rendimiento de
+varios órdenes de magnitud. Sin embargo, el algoritmo McdSat no fue diseñado
+para problemas de reescritura que involucren constantes explícitas, ni para
+computar las mejores reescrituras respecto a una función de utilidad o modelo de
+costo dado. En este trabajo se propone una codificación extendida que supera
+estas limitaciones y se aplica la codificación al problema de instanciación de
+flujos de trabajo.
+
 \end{document}

### 2010-01-11 tesis/propuesta/traduccion/traduccion.tex
pag7

@@
 
+\subsection{Compilación del conocimiento}
+TODO Fig1
+
+Compilación del conocimiento es el área en inteligencia artificial que se ocupa
+del problema de traducir teorías lógicas en fragmentos apropiados que hagan
+tratables ciertas operaciones deseadas XREF11W. Diferentes lenguajes de
+compilación han sido definidos. Por ejemplo están los TODOemphtodoesto Ordered
+Binary Decision Diagrams (OBDDs) XREF10W, Negation Normal Form (NNF) XREF5W, y
+Decomposable Negation Normal Form (DNNF) XREF16W. En este trabajo se usan las
+propiedades de los DNNFs determinísticos (d-DNNF) XREF14W para proveer una
+solución eficiente y escalable al problema de selección de serviccios.
+
+Una teoría en Negation Normal Form es construida desde literales usando sólo
+conjunciones y disjunciones XREF5W, y puede ser representada como un grafo
+acíclico dirigido en que las hojas son etiquetadas con literales y los nodos
+internos con $TODOand$ y $TODOor$; ver ejemplo en la figura  1TODOLABELREF. Este
+fragmento es universal, lo que significa que para cada fórmula lógica hay una
+equivalente en formato NNF. Se dice que un NNF es descomponible (DNNF) XREF14W
+si para cada conjunción $TODOformula$, sus variables son disjuntos por pares; es
+decir, $TODO formula vars y vars$ para $i TODOnoigual j$. Un DNNF soporta un
+número de operaciones en tiempo polinomial en el tamaño de su GAD. Por ejemplo,
+podemos probar si un DNNF es satisfactible haciendo una simple pasada de abajo
+hacia arriba sobre su GAD en tiempo linear. Se dice que un DNNF es
+determinístico (d-DNNF) XREF14W si para cada disjunción $TODOformula$, los
+disjuntos sob lógicamente contradictorios por pares, es decir, $TODO formula
+false idistintoj$. El NNF en la figura 1TODOLABELREF, por ejemplo, es
+decomposableTODOemph y determinístico. Un d-DNNF soporta conteo de modelos en
+tiempo polinomial en el tamaño de su GAD, y enumeración de modelos en tiempo
+polinomial en el tamaño de la salida. Además, dada una función de puntuación de
+literales $r$, se puede computar la puntuación del mejor modelo en tiempo
+polinomial para DNNFs XREF17W.
+
+Los fragmentos DNNF y d-DNNF son universales pero traducir una teoría CNF a
+formrato DNNF tiene costo exponencial en el peor caso. Esta traducción se
+llama compilación en este campo. Hay un compilador públicamente disponible,
+llamado c2d TODOfootnote que realiza este proceso de compilación y que hace uso
+de técnicas de SAT modernas como backtracking dirigido por conflictos,
+aprendizaje de cláusulas y caching XREF15W. Este compilador entra en el peor
+caso en espacio exponencial en un parámetro llamado el ancho del árbol de
+descomposición que está relacionado a la TODO"conectividad" de la teoría CNF.
+Sin embargo, en nuestros experimentos, las teorías CNF que son compiladas son de
+ancho pequeño.
+
 \end{document}

### 2010-01-11 tesis/propuesta/traduccion/traduccion.tex
pag 11

@@
 
+\section{Arquitectura del sistema}
+
+TODOfig2 system arquitecthuere
+
+Se utiliza una arquitectura comprendida por un catálogo de descripciones de
+servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
+el Decodificador. La figura TODOREF2 muestra la arquitectura global del sistema.
+En estta infraestructura, una instancia del problema de instanciación de lfujos
+de trabajo se define como un flujo abstracto representado por una consulta
+conjuntiva sobre servicios abstractos que es dada como entrada junto con un
+conjunto de servicios concretos definidos por vistas de servicios abstractos.
+
+El catálogo se pobla con descripciones de servicios abstractos y concretos; cada
+servicio es descrito en términos de atributos de entrada y salida y anotado con
+un valor real que representa la utilidad QoS del servicio. La descripción de los
+servicios concretos, que son definidos como vistas de los servicios abstractos,
+puede ser generada de manera semiautomática o automática usando herramientas
+tales como el sistema DEIMOS XREF3W.
+
+Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
+corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
+compilador c2d, un componente TODOemphofftheshelf, compila la fórmula CNF a
+d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
+convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
+parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
+importante remarcar que el proceso de compilación necesita ser realizado sólo
+una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
+compilación resulta ser costosa en términos de tiempo, este costo puede ser
+amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
+instanciaciones con respecto a múltiples valores de los parámetros QoS.
+Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscardor
+a una instanciación de flujo de trabajo que resuelve el WIP.
+
+Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
+todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
+una manera libre de TODOemphbacktracking todas las instanciaciones del flujo de
+trabajo. Si el usuario está interesado en una mejor instanciación dados
+parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamñao
+del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
+estas pueden ser computadas en tiempo lineal en su número. FInalmente, si el
+usuario está interesado en todas las instanciaciones, estas también pueden ser
+computadas en tiempo lineal en su número. En los últimos dos casos, si ese
+número es exponencial (en el tamaño de la entrada), la enumeración de las
+instanciaciones también lo es pero esta complejidad es intrínsica al problema y
+por lo tanto no puede ser evitada.
+
+\section{Resultados Experimentales}
+
+Se realizó un análisis empírico sobre tres experimentos. Todos lso experimentos
+fueron realizados en una máquina de escritorio con un CPU Intel Core 2 Duo de
+2GHz y 4GB de memoria, y el tiempo fue medido con el comando de Unix
+timeTODOemph.
+
+El objetivo del experimento es determinar el rendimiento de la propuesta con
+condiciones variantes. El principal beneficio de la aproximación propuesta es
+que se puede compilar la teoría lógica para una instancia del problema y luego
+calcular todas las instanciaciones, o las mejores, cualquier número de veces. El
+modelo de costos para conseguir mejores instanciaciones puede ser cambiado sin
+necesidad de recompilar la teoría lógica. Por lo tanto, la complejidad en tiempo
+de nuestra aproximación es básicamente el tiempo para codificar el WIP como un
+CNF más el tiempo para compilar el CNF en un d-DNNF y el tiempo para decodificar
+los modelos. Los tiempos para codificar y decodificar son despreciables
+comparados con el tiempo para compilar el CNF. Por esto, se enfoca en el tiempo
+necesario para compilar los problemas de los experimentos.
+
+El primer experimento consiste de problemas para consultas para viajes aéreos.
+Los servicios concretos son de la forma $VTODO formula$ donde $AL_iTODO$ es una
+constante que denota el nombre de una aerolínea, y $flightTODO$ relaciona las
+ciudades $x$ e $y$ tales que hay un vuelo entre $x$ e $y$ servido por $ALITODO$.
+Se supone que este servicio concreto retorna todos los vuelos entre dos ciudades
+con una aerolínea específica. El flujo de trabajo tiene la forma
+
+$$
+Wx1...TODO
+$$
+
+El experimento incluye instancias para flujos de trabajo con 2 a 5 subobjetivos
+y conjuntos de 10 a 100 servicios concretos. Los resultados para la compilación
+son mostrados en el panel (a) de la figura TODO3. Este es un gráfico en escala
+logarítmica que sugiere comportamiento sub-exponencial. En cualquier caso, los
+resultados muestran buen rendimiento dado que instancias realísticas del
+problema (conjuntos de 100 aerolíneas con vuelos de cinco paradas) pueden ser
+compiladas en 328 segundos. El tamaño en disco del d-DNNF para 100 aerolíneas y
+vuelos con cinco paradas es 3,4MB. En este d-DNNF, el mejor modelo puede ser
+computado en 0.29 segundos, y la enumeración de todos los modelos en 0.47
+segundos.
+
+TODOfig3 experimentos
+
+En un intento de inducir crecimiento exponencial en el tamaño de compilación, en
+el segundo experimento se agrega un segundo servicio concreto para cada
+aerolínea. Esta modificación incrementa el número de instanciaciones válidas de
+lineal a exponencial dado que cada tramo de un vuelo puede ser ahora instanciado
+por dos servicios concretos y entonces un vuelo con $n$ tramos puede tener hasta
+$2^n$ instanciaciones. Se corrió el compilador para instancias que comprendían
+el mismo número de subobjetivos del flujo de trabajo y el número total de
+servicios concretos. Los resultados graficados en escala logarítmica se muestran
+en el panel (b) de la figura TODO3.
+
+Estas pruebas muestran buen rendimiento para este tipo de problemas, pero no
+involucran servicios concretos con múltiples subobjetivos. Por lo tanto se
+diseñó un tercer experimento que consiste de instancias no estructuradas y
+generadas aleatoriamente. Cada instancia contiene tres variables por servicio
+abstracto, diez variables distintas y diez constantes distintas, seis
+subobjetivos en los flujos de trabajo, 2 a 5 subobjetivos en los servicios
+concretos, y número de servicios variante. La probabilidad de que un argumento
+de un servicio abstracto esté ligado a una constante es 50\%. Los resultados se
+muestran en la figura TODO4. El tiempo de compilación de estas instancias no
+crece monótonamente dado que son generadas aleatoriamente. Lo mismo ocurre para
+el tamaño de las teorías y los números de modelos. Por ejemplo, el d-DNNF para
+un problema con 45 vistas cada una con 5 subobjetivos fue de tamaño 5,1Mb y tuvo
+$1.26 TODO 20101028$ modelos. El tiempo para buscar el mejor modelo para este
+d-DNNF es 0.46 segundos mientras que el tiempo para enumerar todos los modelos
+es alredededor de 17 horas.
+
+Estos son experimentos preliminares, pero los resultados muestran que la
+aproximación propuesta escala eficientemente para problemas con varios objetivos
+y vistas. Se cree que estos resultados son alentadores y motivan a continuar
+esta línea de investigación.
+
+\section{Formalización}
+
+Consideramos cat[alogos de servicio de la forma $C = ST$ donde $S$ es un
+conjunto de predicados que representa servicios abstractos y $T=TTODO$ es una
+colección de tablas que representa el resultado de evaluar los servicios
+abstractos. En el contexto del WIP, un catálogo de servicio $C$ es una
+descripción idealizada de la salida producida por flujos abstractos
+implementados por servicios concretos descritos como vistas. Un flujo de trabajo
+$W$ sobre $S$ es una consulta conjuntiva de la forma
+
+$$
+WX
+$$
+
+donde $si,x$ es un vector de variables y cada $xi$ es un vector de variables y
+constantes. El resultado de $W$ sobre $C$, denotado como $W(C)$, es la tabla con
+$|x|$ columnas que resulta de la proyección del TODOemphjoin relacional
+$JOINTODO$ sobre $x$. Los átomos en el cuerpo de $W$ son llamados los
+(sub)objetivos de $W$, y las variables en la cabeza de $W$ son llamados
+distinguidos.
+
+Un servicio concreto es descrito como una vista $V$ sobre $C$ que, siguiendo
+LAV, es una consulta sobre $S$. Dado un catálogo $C$, un flujo de trabajo $W$ y
+una colección de $n$ vistas $E=$, se nos pide conseguir todas las tuplas en
+$W(C)$ que puedan ser obtenidas de vistas en $E$. Es decir, se necesitan
+conseguir las instanciaciones
+
+$$
+R(x)
+$$
+
+tales que $R(E)...$. Un problema de instanciación de flujo de trabajo es una
+tupla $tUTPLA$ donde $S$ es un conjunto de predicados que representa servicios
+abstractos, $W$ es un flujo de trabajo sobre $S$, y $VI$ una colección de vistas
+que define servicios concretos. Suponemos trabajo con problemas \emph{seguros}
+en el sentido en que todas las variables mencionadas en la cabeza del flujo de
+trabajo (respectivamente en la cabeza de cada vista) aparecen en el cuerpo del
+flujo de trabajo (respectivamente en el cuerpo de cada vista). Además, sólo
+lidiamos con WIPs sin predicados aritméticos dentro del flujo de trabajo o
+vistas. Una instanciación $R$ es válida si para todos los catálogos $C=$ y
+extensiones $EIII$, $R(E#)$. Una colección $R$ de instanciaciones válidas es una
+solución si para todos los catálogos de servicio $C=$ y extensiones $EII$, no
+existe $R'$ tal que $R(E= C$.
+
 \end{document}

### 2010-01-12 tesis/propuesta/secciones/planteamiento.tex
reorganizacion de propuesta

--- 2009-12-13 tesis/propuesta/planteamiento.tex
@@
 parámetros QoS (calidad de servicio) que describen los servicios concretos
-disponibles. Para evaluar el flujo propuesto, se instancian los servicios
-abstractos con concretos de manera que se satisfagan las restricciones sobre los
-parámetros QoS. Sin embargo, el rendimiento de estas tareas es impactado por el
+disponibles. Para evaluar el flujo propuesto, se deben instanciar los servicios
+abstractos con servicios concretos de manera que se satisfagan las restricciones
+sobre los
+parámetros QoS. Sin embargo, el rendimiento de esta tarea es impactado por el
 número de fuentes disponibles. Se requieren técnicas que permitan escalar
@@
 seleccionen combinaciones que cumplan con los requerimientos funcionales y
-no-funcionales que describen una aplicación diseñada. Se llama 
-\mcdsatc a la solución propuesta.
+no-funcionales que describen una aplicación diseñada. A la solución propuesta se
+le da el nombre de \mcdsatc.
 
@@
 En este trabajo se propone la infraestructura \mcdsatc que extiende la
-aproximación propuesta en  ~\cite{arvelo:aaai06}, con la capacidad de trabajar
+aproximación propuesta en  ~\cite{arvelo:aaai06} con la capacidad de trabajar
 con constantes y de identificar la composición de servicios que maximice una
@@
 
-This paper is comprised of six additional sections. The next section motivates
-our approach by using an example, and  we summarize existing related work in
-section 3. Section 4 describes the \mcdsatc system architecture, and  the experimental study is reported in section 5. 
-In section 6 we present the formalization of the web service selection problem as a Propositional Logic theory, and finally, section 7 outlines our conclusions and future work.
+\subsection{Ejemplo}
 
-\section{Motivación}
-
-Consideremos los servicios abstractos y concretos en la tabla
-~\ref{table:services} que representan aplicaciones en el dominio de vuelos.
+Consideremos los servicios abstractos y concretos en el
+cuadro~\ref{table:services} que representan aplicaciones en el dominio de
+vuelos.
 
 \begin{table}
-\begin{tabular}{| l| l |}
+\begin{tabular}{|l|p{8.00cm}|}
 \hline
@@
 \hline
-$fight(x1,x2)$ & relaciona las ciudades x1 y x2 si hay un vuelo entre ellas\\
+$vuelo(x1,x2)$ & relaciona las ciudades $x1$ y $x2$ si hay un vuelo entre ellas\\
 \hline
-$uscity(x1)$ &  determina si una ciudad x1 es una ciudad de Estados Unidos\\
+$ciudadEU(x1)$ &  determina si una ciudad $x1$ es una ciudad de Estados Unidos\\
 \hline \hline
@@
 \hline
-$national(x1,y1)$&  retorna dos ciudades de Estados Unidos que están conectadas
+$nacional(x1,y1)$&  retorna dos ciudades de Estados Unidos que están conectadas
 por vuelos directos \\
 \hline
-$oneway(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos directos
+$ida(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos directos
 de ida\\
 \hline
-$onestop(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos con una
+$unaescala(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos con una
 escala\\
 \hline
-$ flightToCCS(x1,"CCS")$ & retorna una ciudad que está conectada a Caracas por
+$vueloHaciaCCS(x1, ``CCS")$ & retorna una ciudad que está conectada a Caracas por
 un vuelo directo\\
  \hline
-$ onestopToCCS(x1,y1)$ & retorna dos ciudades x1 y y1 tales que hay un vuelo con
-una escala desde la ciudad x1 hasta Caracas con parada en la ciudad y1\\
+$unaescalaHaciaCCS(x1,y1)$ & retorna dos ciudades $x1$ y $y1$ tales que hay un vuelo con
+una escala desde la ciudad $x1$ hasta Caracas con parada en la ciudad $y1$\\
 \hline
- $fromWAS("WAS",y1)$ &  retorna una ciudad y1 a la cual llegan vuelos desde
+$desdeWAS(``WAS",y1)$ &  retorna una ciudad $y1$ a la cual llegan vuelos desde
 Washington.\\
@@
 Usando la aproximación \emph{Local As View} (LAV) ~\cite{AmbiteISWC09}, los
-servicios concretos en la tabla \ref{table:services} son descritos
-semánticamente como vistas sobre los servicios abstractos {\it flight} y
-{\it uscity} como sigue:
+servicios concretos en el cuadro \ref{table:services} son descritos
+semánticamente como vistas sobre los servicios abstractos {\it vuelo} y
+{\it ciudadEU} como sigue:
 
 \begin{alignat*}{1}
-national(x1,y1)\ &\qrule\ flight(x1,y1),\,  uscity(x1),\,  uscity(y1)\,. \\
-oneway(x2,y2)\ &\qrule\ flight(x2,y2)\,. \\
-onestop(x3,z3)\ &\qrule\ flight(x3,y3),\, flight(y3,z3)\,. \\
-flightToCCS(x2,"CCS")\ &\qrule\ flight(x2,"CCS")\,. \\
-onestopToCCS(x3,y3)\ &\qrule\ flight(x3,y3),\, flight(y3,"CCS")\,. \\
-fromWAS("WAS",y3)\ &\qrule\ flight("WAS",x3)\,. \\
+nacional(x1,y1)\ &\qrule\ vuelo(x1,y1),\,  ciudadEU(x1),\,  ciudadEU(y1)\,. \\
+ida(x2,y2)\ &\qrule\ vuelo(x2,y2)\,. \\
+unaescala(x3,z3)\ &\qrule\ vuelo(x3,y3),\, vuelo(y3,z3)\,. \\
+vueloHaciaCCS(x2,``CCS")\ &\qrule\ vuelo(x2,``CCS")\,. \\
+unaescalaHaciaCCS(x3,y3)\ &\qrule\ vuelo(x3,y3),\, vuelo(y3,``CCS")\,. \\
+desdeWAS(``WAS",y3)\ &\qrule\ vuelo(``WAS",x3)\,. \\
 \end{alignat*}
@@
 \begin{alignat*}{1}
-w(x,y1,y2,y3)\ &\qrule\ flight(x,y1),\, flight(y1,y2),\,  flight(y2,y3),\,  flight(y3,x),\,  \\
-                   \ & uscity(x)\,. \\
+w(x,y1,y2,y3)\ &\qrule\ vuelo(x,y1),\, vuelo(y1,y2),\,  vuelo(y2,y3),\,  vuelo(y3,x),\,  \\
+                   \ & ciudadEU(x)\,. \\
 \end{alignat*}
@@
 \begin{alignat*}{1}
-wi1(x,y1,"CCS",y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\,  oneway("CCS",y3),\,  \\
-\ & national(y3,x)\,.  \\
+wi1(x,y1,``CCS",y3)\ &\qrule\ nacional(x,y1),\, vueloHaciaCCS(y1,``CCS"),\, ida(``CCS",y3),\,  \\
+\ & nacional(y3,x)\,.  \\
 \end{alignat*}
@@
 \begin{alignat*}{1}
-wi2(x,y1,"CCS"y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\, \\
-\ & fromWAS("WAS",y3),\,  national(y3,x)\,. \\
-wi3(x,y1,y2,y3)\ &\qrule\ national(x1,y1),\, onestop(x1,y2),\,  oneway(y2,y3),\,  \\
-\ & national(y3,x)\,. \\
+wi2(x,y1,``CCS"y3)\ &\qrule\ nacional(x,y1),\, vueloHaciaCCS(y1,``CCS"),\, \\
+\ & desdeWAS(``WAS",y3),\,  nacional(y3,x)\,. \\
+wi3(x,y1,y2,y3)\ &\qrule\ nacional(x1,y1),\, unaescala(x1,y2),\,  ida(y2,y3),\,  \\
+\ & nacional(y3,x)\,. \\
 \end{alignat*}
@@
 distintas. Por otra parte, la composición {\it wi3} no implementa el flujo de
-trabajo porque el servicio concreto {\it onestop} no recibe como entrada o
+trabajo porque el servicio concreto {\it unaescala} no recibe como entrada o
 produce como salida la ciudad donde el vuelo se detiene (representada por la
 variable {\it y1}); por esto, no es posible garantizar que la ciudad donde el
-vuelo se detiene y la ciudad final obtenida por el servicio {\it national} sean
+vuelo se detiene y la ciudad final obtenida por el servicio {\it nacional} sean
 las mismas.

### 2010-01-12 tesis/propuesta/secciones/planteamiento.tex
planteamiento listo

@@
 
-Bajo la sombra de la Web Semántica y con el apoyo de arquitecturas orientadas a
-servicios, el número de fuentes de datos y servicios Web ha explotado en los
-últimos años. Por ejemplo, actualmente la colección de bases de datos de
-biología molecular incluye 1.170 bases de datos~\cite{Galperin09}, lo cual es 95
-más que el año pasado~\cite{Galperin2008} y 110 más que hace dos
-años~\cite{Galperin2007}. Tanto las herramientas y servicios como el número de
-instancias publicadas por estos recursos siguen una progresión
-similar~\cite{Benson07}. Gracias a este tesoro de datos, los usuarios se apoyan
-más en varias tareas digitales como obtención de datos de fuentes públicas y
-análisis de datos con herramientas o servicios Web organizados en flujos de
-trabajo complejos.
+Bajo la sombra de la Web Semántica, y con el soporte de las Arquitecturas Orientadas a
+Servicios (SOA), el número de fuentes de datos y servicios Web ha explotado en
+los últimos años. Por ejemplo, la colección de bases de datos de biología
+molecular actualmente contiene 1.170 bases de datos~\cite{Galperin09}, un número
+que es
+mayor que el del último año por 95~\cite{Galperin2008}, y que el de hace dos
+años por 110~\cite{Galperin2007};
+las herramientas y servicios, así como el número de instancias
+publicados por estos recursos, siguen una progresión similar~\cite{Benson07}.
+Gracias a
+este tesoro de datos, la tendencia de los usuarios es hacia depender cada vez
+más de métodos automáticos para manejarlos tales como recuperación de datos de
+fuentes públicas y análisis utilizando herramientas o servicios Web compuestos
+en flujos de trabajo complejos.
 
-Para diseñar una aplicación, los usuarios frecuentemente especifican la
-funcionalidad en términos de un flujo de trabajo de servicios abstractos,
-mientras que la calidad deseada se define restringiendo los valores de los
-parámetros QoS (calidad de servicio) que describen los servicios concretos
-disponibles. Para evaluar el flujo propuesto, se deben instanciar los servicios
-abstractos con servicios concretos de manera que se satisfagan las restricciones
-sobre los
-parámetros QoS. Sin embargo, el rendimiento de esta tarea es impactado por el
-número de fuentes disponibles. Se requieren técnicas que permitan escalar
-eficientemente a un número grande de servicios.
+La SOA típica consiste de dos capas. La capa concreta que está hecha de
+servicios concretos cuyas funcionalidades son descritas en términos de pre- y
+post-condiciones y propiedades no funcionales en términos de parámetros de
+calidad de servicio (QoS), y
+la capa abstracta compuesta de aplicaciones de software cuyas funcionalidades
+son descritas en términos de flujos de trabajo abstractos y propiedades no
+funcionales en términos de restricciones de QoS. La ejecución de un flujo de
+trabajo abstracto involucra la instanciación de los servicios abstractos en
+servicios concretos que cumplan con los requerimientos funcionales y no
+funcionales. Este proceso de instanciación puede ser visto como una búsqueda de
+una instanciación objetivo en el espacio combinatorio de todas las
+instanciaciones válidas. Entonces se está interesado en técnicas eficientes para
+realizar esta búsqueda que sean capaces de escalar a medida que el número de
+servicios concretos o la complejidad del flujo de trabajo aumenta. Se llama al
+problema de instanciar un flujo de trabajo dado con servicios concretos de un
+conjunto de servicios concretos dado, de manera que ciertas demandas de QoS sean
+cumplidas, el Problema de Instanciación de Flujos de Trabajo (WIP).
 
-En este trabajo se intenta proponer una solución a este problema y proveer
-técnicas que recorran eficientemente el espacio de servicios disponibles y
-seleccionen combinaciones que cumplan con los requerimientos funcionales y
-no-funcionales que describen una aplicación diseñada. A la solución propuesta se
-le da el nombre de \mcdsatc.
+En este trabajo se considera una versión restringida de WIP que adopta la
+aproximación Local-As-View (LAV) \cite{levy:bucket}. En LAV, todos los elementos de un
+problema son especificados con un lenguaje común basado en servicios abstractos
+tal que los servicios concretos se describen como vistas de servicios
+abstractos, la calidad de una instanciación como una función de utilidad global
+que combina los diferentes parámetros QoS, y el flujo de trabajo abstracto como
+una consulta conjuntiva sobre los servicios abstractos; esta representación es
+similar a la que es generada de manera semiautomática para el sistema DEIMOS
+\cite{AmbiteISWC09}. En esta versión de WIP, las reglas que definen flujos de trabajo
+(consultas conjuntivas) y servicios concretos (vistas) son creados de manera que
+todas las restricciones funcionales sobre pre- y post-condiciones de los
+servicios y sus combinaciones sean satisfechas, y las medidas de QoS son
+representadas anotando cada descripción de servicio concreto con un número real
+que representa la utilidad QoS global del servicio.
 
-En \mcdsatc, las funcionalidades de servicios concretos se describen
-como vistas sobre servicios abstractos y los flujos de trabajo se representan
-como consultas conjuntivas sobre servicios abstractos; esta representación de
-los servicios es similar a la generada de manera semiautomática para el
-sistema {\cal DEIMOS}~\cite{AmbiteISWC09}. Adicionalmente, las reglas
-que definen flujos de trabajo y servicios concretos en el catálogo de
-\mcdsatc son creadas de manera que las restricciones de entrada y salida
-de los servicios abstractos combinados sean satisfechas. Para representar las
-medidas de QoS, cada descripción de servicio concreto es anotada con un número
-real que representa un valor general para las medidas de QoS del servicio en
-particular. Finalmente, \mcdsatc modela el problema de selección de
-servicios web como el problema de reescritura de consultas usando vistas
-\cite{halevy:survey}. Este problema es importante en el contexto de integración
-de datos \cite{Chen05,JaudoinPRST05} y de optimización de consultas y
-mantenimiento de datos \cite{AfratiLU07,levy:bucket} y se han definido varias
-aproximaciones para escalar eficientemente a un número grande de
-vistas~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-En este trabajo se propone la infraestructura \mcdsatc que extiende la
-aproximación propuesta en  ~\cite{arvelo:aaai06} con la capacidad de trabajar
-con constantes y de identificar la composición de servicios que maximice una
-función de utilidad dada.
+Bajo estas suposiciones, WIP puede ser convertido en el bien conocido Problema
+de Reescritura de Consultas (QRP) para LAV que es central a los sistemas de
+integración \cite{halevy:survey}. QRP consiste de una consulta conjuntiva que debe ser
+respondida en términos de vistas donde la consulta y las vistas son descritas
+usando LAV con relaciones abstractas. Este problema es importante en el contexto
+de integración de datos \cite{Chen05,JaudoinPRST05}, y optimización de consultas y mantenimiento
+de datos \cite{AfratiLU07,levy:bucket}, y varias aproximaciones que escalan a un número grande de
+vistas han sido definidas \cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
 
-\mcdsatc  traduce una instancia del problema de selección de servicios
-en el problema de enumerar los modelos de una teoría proposicional que codifica
-información sobre el flujo de trabajo, los servicios concretos y las condiciones
-que deben ser satisfechas para generar una composición que maximice la función
-de utilidad. Esta teoría se representa como una fórmula CNF que es compilada a
-su representación d-DNNF, a partir de la cual los modelos que maximicen la
-función global de utilidad se obtienen en tiempo lineal usando técnicas modernas
-de SAT~\cite{darwiche:dnnf}. Se estudia empíricamente la calidad de \mcdsatc
-sobre una variedad de \emph{benchmarks} y se observa que la técnica
-propuesta provee una solución eficiente al problema de selección de servicios
-que es capaz de escalar a flujos de trabajo grandes así como a números de vistas
-grandes.
+La aproximación recente de Arvelo et al. \cite{arvelo:aaai06} está basada en la enumeración
+eficiente de modelos de una teoría lógica proposicional. Dado un QRP, una teoría
+lógica es construída de manera que cada modelo de la teoría codifica una
+reescritura válida, y así todas las reescrituras son obtenidas enumerando los
+modelos de la teoría. Esta enumeración puede ser realizada eficientemente si la
+teoría lógica está en cierta forma normal llamada la forma normal de negación
+determinística y descomponible (d-DNNF) \cite{darwiche:d-dnnfs}. Así,
+la aproximación consiste en transformar (llamado compilar en el campo de
+compilación del conocimiento) la teoría lógica en el formato d-DNNF para
+enumerar sus modelos eficientemente.
 
-\subsection{Ejemplo}
+Pero las teorías d-DNNF no sólo soportan la enumeración eficiente de sus
+modelos, sino también otras operaciones. Entre ellas se encuentra la enumeración
+de los modelos con mejor puntuación. Dada una función de puntuación de literales
+$r(\ell)$ que asigne puntuaciones a cada literal $\ell$, se define la puntuación
+$r(\omega)$ de un modelo $\omega$ como la suma de las puntuaciones de los literales
+activados por $\omega$ (es decir,
+$r(\omega)\doteq\sum_{\omega\vDash\ell}r(\ell)$), y se dice que $\omega$ es un modelo
+mejor puntuado si no hay modelo $\omega'$ tal que $r(\omega')<r(\omega)$.
 
-Consideremos los servicios abstractos y concretos en el
-cuadro~\ref{table:services} que representan aplicaciones en el dominio de
-vuelos.
+Dada una teoría en d-DNNF, se computa la puntuación de los mejores modelos, y
+los mejores modelos, en tiempo lineal en el tamaño del d-DNNF. Este cómputo
+transforma el GAD del d-DNNF en un circuito aritmético reemplazando los nodos
+AND con `+' y los nodos OR con `min'. La función de puntuación de literales
+asigna valores a las hojas del circuito que son propagados a la raíz en tiempo
+lineal. El valor de la raíz es la puntuación del mejor modelo \cite{darwiche:weighted}.
 
-\begin{table}
-\begin{tabular}{|l|p{8.00cm}|}
-\hline
-Servicio Abstracto & Descripción \\
-\hline
-$vuelo(x1,x2)$ & relaciona las ciudades $x1$ y $x2$ si hay un vuelo entre ellas\\
-\hline
-$ciudadEU(x1)$ &  determina si una ciudad $x1$ es una ciudad de Estados Unidos\\
-\hline \hline
-Servicio Concreto & Descripción \\
-\hline
-$nacional(x1,y1)$&  retorna dos ciudades de Estados Unidos que están conectadas
-por vuelos directos \\
-\hline
-$ida(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos directos
-de ida\\
-\hline
-$unaescala(x1,y1)$ & retorna dos ciudades que están conectadas por vuelos con una
-escala\\
-\hline
-$vueloHaciaCCS(x1, ``CCS")$ & retorna una ciudad que está conectada a Caracas por
-un vuelo directo\\
- \hline
-$unaescalaHaciaCCS(x1,y1)$ & retorna dos ciudades $x1$ y $y1$ tales que hay un vuelo con
-una escala desde la ciudad $x1$ hasta Caracas con parada en la ciudad $y1$\\
-\hline
-$desdeWAS(``WAS",y1)$ &  retorna una ciudad $y1$ a la cual llegan vuelos desde
-Washington.\\
- \hline
-\end{tabular}
-\label{table:services}
-\caption{Servicios Abstractos y Concretos}
-\end{table}
-
-Usando la aproximación \emph{Local As View} (LAV) ~\cite{AmbiteISWC09}, los
-servicios concretos en el cuadro \ref{table:services} son descritos
-semánticamente como vistas sobre los servicios abstractos {\it vuelo} y
-{\it ciudadEU} como sigue:
-
-\begin{alignat*}{1}
-nacional(x1,y1)\ &\qrule\ vuelo(x1,y1),\,  ciudadEU(x1),\,  ciudadEU(y1)\,. \\
-ida(x2,y2)\ &\qrule\ vuelo(x2,y2)\,. \\
-unaescala(x3,z3)\ &\qrule\ vuelo(x3,y3),\, vuelo(y3,z3)\,. \\
-vueloHaciaCCS(x2,``CCS")\ &\qrule\ vuelo(x2,``CCS")\,. \\
-unaescalaHaciaCCS(x3,y3)\ &\qrule\ vuelo(x3,y3),\, vuelo(y3,``CCS")\,. \\
-desdeWAS(``WAS",y3)\ &\qrule\ vuelo(``WAS",x3)\,. \\
-\end{alignat*}
-
-Ahora supongamos que un usuario está interesado en implementar un flujo de
-trabajo capaz de obtener los vuelos ida y vuelta con una escala de ciudades de
-Estados Unidos a cualquier ciudad en el mundo donde las escalas pueden ser
-hechas en cualquier ciudad. Consideremos la siguiente consulta conjuntiva que
-representa el flujo de trabajo que define esta solicitud en términos de
-servicios abstractos. Este flujo fue definido de manera que los parámetros de
-entrada de cada servicio abstracto están ligados a constantes o a
-atributos proyectados por servicios abstractos previos.
-
-\begin{alignat*}{1}
-w(x,y1,y2,y3)\ &\qrule\ vuelo(x,y1),\, vuelo(y1,y2),\,  vuelo(y2,y3),\,  vuelo(y3,x),\,  \\
-                   \ & ciudadEU(x)\,. \\
-\end{alignat*}
-
-Las implementaciones de este flujo de trabajo corresponden a composiciones de
-servicios concretos donde cada servicio concreto puede implementar un
-subconjunto de los servicios abstractos del flujo de trabajo, pero cada servicio
-abstracto puede ser implementado por exactamente un servicio concreto. La
-siguiente composición de servicios concretos corresponde a una de las XX
-implementaciones del flujo de trabajo.
-
-\begin{alignat*}{1}
-wi1(x,y1,``CCS",y3)\ &\qrule\ nacional(x,y1),\, vueloHaciaCCS(y1,``CCS"),\, ida(``CCS",y3),\,  \\
-\ & nacional(y3,x)\,.  \\
-\end{alignat*}
-
-Sin embargo, nótese que las siguientes composiciones no son válidas:
-
-\begin{alignat*}{1}
-wi2(x,y1,``CCS"y3)\ &\qrule\ nacional(x,y1),\, vueloHaciaCCS(y1,``CCS"),\, \\
-\ & desdeWAS(``WAS",y3),\,  nacional(y3,x)\,. \\
-wi3(x,y1,y2,y3)\ &\qrule\ nacional(x1,y1),\, unaescala(x1,y2),\,  ida(y2,y3),\,  \\
-\ & nacional(y3,x)\,. \\
-\end{alignat*}
-
-La composición {\it wi2} no es válida porque asocia la variable del flujo de
-trabajo {\it y2} a las constantes {\it ``CCS''} y {\it ``WAS''} que denotan
-ciudades
-distintas. Por otra parte, la composición {\it wi3} no implementa el flujo de
-trabajo porque el servicio concreto {\it unaescala} no recibe como entrada o
-produce como salida la ciudad donde el vuelo se detiene (representada por la
-variable {\it y1}); por esto, no es posible garantizar que la ciudad donde el
-vuelo se detiene y la ciudad final obtenida por el servicio {\it nacional} sean
-las mismas.
-
-Para evitar producir implementaciones del flujo de trabajo inválidas, se
-requiere manejar constantes de manera que constantes distintas no sean asociadas
-entre sí directa o indirectamente (por transitividad). Además, todos los
-atributos en la salida del flujo o necesarios para unir otros servicios en el
-flujo de trabajo deben ser producidos por el servicio concreto seleccionado.
-
-Adicionalmente, supongamos que cada servicio concreto está anotado con una
-función de utilidad que agrega los valores de parámetros diferentes de QoS que
-caracterizan el comportamiento del servicio. Entonces, como el espacio de
-composiciones de servicios válidas de un flujo abstracto puede ser muy grande,
-se requiere una técnica capaz de identificar aquellos que maximicen la función
-de utilidad sin tener que enumerar el conjunto completo de composiciones
-válidas. En este trabajo se propone una aproximación proposicional basada en
-lógica que toma provecho del poder de los \emph{SAT solvers} modernos para
-enumerar eficientemente las composiciones de servicios que corresponden a
-implementaciones válidas de un flujo de trabajo dado, así como las composiciones
-que maximizan una función de utilidad dada.
+En este trabajo se propone explotar las propiedades de los d-DNNF construyendo una
+teoría lógica cuyos modelos codifican las instanciaciones del flujo de trabajo
+y cuyos mejores modelos codifican las instanciaciones óptimas (mejores). Así, la
+búsqueda combinatoria se reduce a la computación de un mejor modelo de una
+teoría lógica que puede ser realizada eficientemente una vez que la teoría sea
+transformada al formato d-DNNF.

### 2010-01-12 tesis/propuesta/secciones/portada.tex
planteamiento listo

--- 2010-01-11 tesis/propuesta/traduccion/traduccion.tex
@@
-\documentclass{article}
+\begin{titlepage}
 
-\usepackage[utf8x]{inputenc}
-\usepackage[spanish]{babel}
+\begin{flushleft}
+\includegraphics[scale=0.3]{logo_usb.png} \\
+Universidad Simón Bolívar \\
+Coordinación de Ciencias de la Computación \\
+Maestría en Ciencias de la Computación \\
+\end{flushleft}
 
-\begin{document}
+\vspace{0.8cm}
 
-\title{Instanciación Óptima de Flujos de Trabajo Abstractos Usando Lógica y Circuitos}
-\maketitle
+\begin{center}
+{\large \bf \textsf{Instanciación Óptima de Flujos de Trabajo Abstractos Usando Lógica y Circuitos}}
+\end{center}
 
-\section{Resumen}
+\vspace{0.2cm}
 
+\begin{abstract}
 La Arquitectura Orientada a Servicios típica consiste de dos capas principales.
@@
 
-\section{Introducción}
+\end{abstract}
 
-Bajo la sombra de la Web Semántica, y soportado por Arquitecturas Orientadas a
-Servicios (SOA), el número de fuentes de datos y servicios Web ha explotado en
-los últimos años. Por ejemplo, la colección de bases de datos de biología
-molecular actualmente contiene 1170 bases de datos XREF22W, un número que es
-mayor que el del último año XREF21W por 95, y que el de hace dos años por 110
-XREF20W; las herramientas y servicios, así como el número de instancias
-publicados por estos recursos, siguen una progresión similar XREF6W. Gracias a
-este tesoro de datos, la tendencia de los usuarios es hacia depender cada vez
-más en métodos automáticos para manejarlos tales como recuperación de datos de
-fuentes públicas y análisis utilizando herramientas o servicios Web compuestos
-en flujos de trabajo complejos.
+\vspace{\fill}
 
-La SOA típica consiste de dos capas. La capa concreta que está hecha de
-servicios concretos cuyas funcionalidades son descritas en términos de pre- y
-post-condiciones y propiedades no funcionales en términos de parámetros QoS, y
-la capa abstracta compuesta de aplicaciones de software cuyas funcionalidades
-son descritas en términos de flujos de trabajo abstractos y propiedades no
-funcionales en términos de restricciones de QoS. La ejecución de un flujo de
-trabajo abstracto involucra la instanciación de los servicios abstractos en
-servicios concretos que cumplen con los requerimientos funcionales y no
-funcionales. Este proceso de instanciación puede ser visto como una búsqueda de
-una instanciación objetivo en el espacio combinatorio de todas las
-instanciaciones válidas. Así, uno está interesado en técnicas eficientes para
-realizar esta búsqueda que sean capaces de escalar a medida que el número de
-servicios concretos o la complejidad del flujo de trabajo aumenta. Se llama al
-problema de instanciar un flujo de trabajo dado con servicios concretos, de un
-conjunto dado de servicios concretos, de manera que ciertas demandas de QoS sean
-cumplidas, el Workflow Instantiation Problem (WIP, "Problema de Instanciacion de
-Flujos de Trabajo").
+\begin{minipage}{0.3\textwidth}
+\begin{flushleft}
+\emph{Estudiante:} \\
+Daniel \textsc{Izquierdo} \\
+Carnet 08-86809 \\
+\end{flushleft}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+\begin{center}
+Fecha estimada de culminación: \\
+Marzo de 2010
+\end{center}
+\end{minipage}
+\begin{minipage}{0.3\textwidth}
+\begin{flushright}
+\emph{Asesor:} \\
+Prof. María Esther \textsc{Vidal}
+\end{flushright}
+\end{minipage}
 
-En este trabajo se considera una versión restringida de WIP que adopta la
-aproximación Local-As-View (LAV) XREF28W. En LAV, todos los elementos de un
-problema son especificados con un lenguaje común basado en servicios abstractos
-tal que los servicios concretos se describen como vistas de servicios
-abstractos, la calidad de una instanciación como una función de utilidad global
-que combina los diferentes parámetros QoS, y el flujo de trabajo abstracto como
-una consulta conjuntvia sobre los servicios abstractos; esta representación es
-similar a la que es generada de manera semiautomática para el sistema DEIMOS
-XREF3W. En esta versión de WIP, las reglas que definen flujos de trabajo
-(consultas conjuntivas) y servicios concretos (vistas) son creados de manera que
-todas las restricciones funcionales sobre pre- y post-condiciones de los
-servicios y sus combinaciones sean satisfechas, y las medidas de QoS son
-representadas anotando cada descripción de servicio concreto con un número real
-que representa la utilidad QoS global del servicio.
-
-Bajo estas suposiciones, WIP puede ser convertido en el bien conocido Problema
-de Reescritura de Consultas (QRP) para LAV que es central a los sistemas de
-integración XREF23W. QRP consiste de una consulta conjuntiva que debe ser
-respondida en términos de vistas donde la consulta y las vistas son descritas
-usando LAV con relaciones abstractas. Este problema es importante en el contexto
-de integración de datos XREF13X24W, y optimización de consultas y mantenimiento
-de datos XREF1,28W, y varias aproximaciones que escalan a un número grande de
-vistas han sido definidas XREF4,18,19,28,29W.
-
-La aproximación recente de Arvelo et al. XREF4W está basada en la enumeración
-eficiente de modelos de una teoría lógica proposicional. Dado unQRP, una teoría
-lógica es construída de manera que cada modelo de la teoría codifica una
-reescritura válida, y así todas las reescrituras son obtenidas enumerando los
-modelos de la teoría. Esta enumeración puede ser realizada eficientemente si la
-teoría lógica está en cierta forma normal llamada la forma normal de negación
-determinística y descomponible (d-DNNF por sus siglas en inglés) XREF14W. Así,
-la aproximación consiste en transformar (llamado compilar en el campo de
-compilación del conocimiento) la teoría lógica en el formato d-DNNF para
-enumerar sus modelos eficientemente.
-
-Pero las teorías d-DNNF no sólo soportan la enumeración eficiente de sus
-modelos, sino también otras operaciones. Entre ellas se encuentra la enumeración
-de los modelos con mejor puntuación. Dada una función de puntuación de literales
-$r(l)$ que asigne puntuaciones a cada literal $l$, se define la puntuación
-$r(w)$ de un modelo $w$ como la suma de las puntuaciones de los literales
-activados por $w$ (es decir, $r(w) = suma...$, y se dice que $w$ es un modelo
-mejor puntuado si no hay modelo $wpriemrie$ tal que $r(w)...$.
-
-Dada una teoría en d-DNNF, se computa la puntuación de los mejores modelos, y
-los mejores modelos, en tiempo lineal en el tamaño del d-DNNF. Este cómputo
-transforma el GAD del d-DNNF en un circuito aritmético reemplazando los nodos
-AND con '+' y los nodos OR con 'min'. La función de puntuación de literales
-asigna valores a las hojas del circuito que son propagados a la raíz en tiempo
-lineal. El valor de la raíz es la puntuación del mejor modelo XREF17W.
-
-En este trabajo, se explotan las propiedades de los d-DNNF construyendo una
-teoría lógica cuyos modelos codifican las instanciaciones del flujo de trabajo
-y cuyos mejores modelos codifican las instanciaciones óptimas (mejores). Así, la
-búsqueda combinatoria se reduce a la computación de un mejor modelo de una
-teoría lógica que puede ser realizada eficientemente una vez que la teoría sea
-transformada en formato d-DNNF.
-
-Esta propuesta contiene seis secciones más. La próxima sección ilustra y motiva
-WIP con un ejemplo simple pero típico. La sección 3 resume el trabajo existente
-relacionado en tres áreas de selección de servicios, reescritura de consultas y
-compilación del conocimiento. Entonces, las secciones 4 y 5 describen la
-arquitectura del sistema y reportan los resultados empíricos obtenidos hasta
-ahora sobre distintos problemas de prueba respectivamente. La sección 6 presenta
-una descripción formal de la aproximación propuesta junto con un análisis. La
-propuesta concluye con una discusión final en la sección 7.
-
-\section{Ejemplo motivante}
-
-Considere un sistema simple de información de vuelos que contiene información
-sobre vuelos entre ciudades e información sobre qué ciudades están en los
-Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
-abstractos $flight$ y $uscity$. El primero relaciona dos ciudades $x$ e $y$ si
-hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
-E.E.U.U.  o no. Para los servicios concretos, supongamos que las fuentes de
-datos disponibles en Internet contienen la siguiente información:
-
-\begin{itemize}
-TODO
--national
-\end{itemize}
-
-Adicionalmente, los servicios concretos se describen con los siguientes
-servicios abstractos:
-
-\begin{itemize}
-TODO
--national
-\end{itemize}
-
-Ahora supongamos que un usuario está interesado en construir un flujo de trabajo
-capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
-en el mundo, tales que los vuelos puedan detenerse en cualquier ciudad.
-
-$$
-W(x,w,y,z)
-$$
-
-La siguiente consulta conjuntiva representa el flujo de trabajo que define esta
-solicitud en términos de servicios abstractos. El flujo de trabajo es definido
-de manera que todo lo relacionado sobre el ligamiento de parámetros de
-entrada/salida es resuelto. Así, cualquier instanciación de los servicios
-abstractos en términos de los servicios concretos es una implementación válida
-del flujo de trabajo. Las implementaciones corresponden a composiciones de
-ervicios concretos en las cuales un servicio concreto puede implementar uno o
-más servicios abstractos del flujo, pero cada servicio abstracto puede ser
-implementado por exactamente un servicio concreto. Por ejemplo, la siguiente
-composición corresponde a una de estas implementaciones.
-
-$$
-I(x
-$$
-
-Sin embargo, las siguientes dos composiciones no son válidas.
-
-$$
-I'
-I''
-$$
-
-La primera composición no es válida porque asocia la variable del flujo $y$ a
-las constantes PA y NY que denotan ciudades diferentes. En la otra mano, $I''$
-no lo implementa porque el servicio concreto $ones-stop$ no recibe como entrada
-o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces
-no es posible asegurarse de que esa ciudad esté asociada a la ciudad $w$ que es
-retornada por el flujo de trabajo.
-
-Estos ejemplos muestran que las instanciaciones correctas de un flujo deben
-manejar constantes de manera que ods constantes distintas no sean asociadas
-entre sí de manera directa o indirectamente vía transitividad, y que todos los
-atributos que aprezcan en un TODOemphjoin o en la salida sean producidos por los
-servicios concretos seleccionados.
-
-Los parámetros QoS son modelados anotando los servicios concretos con utilidades
-que caracterizan su comportamiento que luego son agregados durante la
-instanciación. Así, como se mencionó previamente, la mejor instanciación es la
-que minimiza (o maximiza) la agregación de utilidades.
-
-\emph{Trabajo Relacionado}
-
-En esta sección se resumen las aproximaciones existentes que proveen soluciones
-a los problemas de selección de servicios y reescritura de consultas y se
-discute el trabajo relacionado en el área de inteligencia artificial llamado
-compilación del conocimiento.
-
-\subsection{Selección de Servicios}
-
-El problema de seleccionar los servicios que implementan un flujo de trabajo
-abstracto y cumplen mejor los criterios basados en QoS es conocido como el
-problema de selección o composición de servicios consciente de QoS, que ha sido
-demostrado ser NP-difícil XREF34W. Este es un problema de optimización
-combinatoria y varias heurísticas han sido propuestas para conseguir una
-solución relativamente buena en un período razonablemente corto de tiempo.
-
-Rahmani et al. XREF31W presenta una heurística basada en métrica de distancia
-que guía un algoritmo de búsqueda hacia atrás; esta métrica induce un orden de
-los servicios en una manera en que los nodos vertedero tienen baja probabilidad
-de ser visitados. En una serie de artículos, Berardi y otros XREF7-9W describen
-servicios y flujos de trabajo en términos de máquinas de estados finitos
-determinísticas que son codificadas usando teorías de Lógica de Descripción
-cuyos modelos corresponden a soluciones del problema. Aunque se podrían explotar
-métodos para formalismos de Lógicas de Descripción, no han sido reportados la
-escalabilidad o el rendimiento de la solución propuesta.
-
-Ko et al. XREF25W propone una aproximación basada en restriccciones que codifica
-los valores permisibles no funcionales como un conjunto de restricciones cuya
-violación debe ser minimizada; para recorrer el espacio de soluciones
-posiblemente óptima, se implementa un algoritmo híbrido que combina las
-metaheurísticas TODOemphtabusearch y TODOemphsimulatedannealing. Los resultados
-experimentales muestran que la solución propuesta es capaz de escalar a un
-número grande de servicios y procesos abstractos. Cardellini et al. XREF12W
-codifican una parte del problema de composición de servicios consciente de QoS
-como un problema de programación lineal XREF12W. Por otra parte, Wada et al.
-XREF34W trata el problema como uno de optimización con múltiples objetivos donde
-los diferentes parámetros de QoS son considerados igualmente importanes en vez
-de ser agregados en una función. Luego se propone un algoritmo basado en
-algoritmos genéticos que identifica un conjunto de composiciones de servicios
-no dominadas que mejor encajan en los requerimientos QoS.
-
-Alrifai y Risse XREF2W proponen una solución de doble cara que usa un
-algoritmo de programación entera híbrido para obtener la descomposición de QoS
-globales en restricciones locales y luego selecciona los servicios que mejor
-cumplen las restricciones locales.
-
-Recientemente, dos aproximaciones basadas en planificación han sido propuestas.
-Kuter y Golbek XREF26W extienden el algoritmo de planificación SHOP2 para
-seleccionar la composición confiable de servicios que implementan un modelo de
-procesos OWL-S dado, mientras que Sohrabi y McIlraith XREF33W proponen una
-solución basada en planificación HTN donde las métricas de preferencia del
-usario y regulaciones de dominio son utilizadas para guíar el planificador hacia
-el espacio de composiciones relevantes. Finalmente, Lécué XREF27W propone un
-algoritmo basado en genética para identificar la composición de servicios que
-mejor cumple con los criterios de calidad para un conjunto de parámetros QoS.
-
-Aunque estas soluciones son capaces de resolver el problema de optimización y
-escalar a un número de procesos abstractos, ninguno de ellos está ajustado para
-describir semánticamente los servicios en términos de procesos abstractos, ni
-para usar estas descripciones para identificar los servicios que implementan un
-flujo de trabajo dado o que mejor cumplen los criterios no funcionales del
-usuario.
-
-\subsection{Reescritura de Consultas}
-
-Un número de algoritmos han sido desarrollados para encontrar las reescrituras
-de una consulta dada; los más prominentes son el algoritmo bucket XREF28W, el
-algoritmo de reglas inversas XREF18,30W, el algoritmo MiniCon XREF29, y el
-algoritmo McdSat XREF4W. Generalmente, la reescritura de consultas trabaja en
-dos fases. Durante la primera, el algoritmo identifica las vistas que reescriben
-al menos un subobjetivo de la consulta, y durante el segundo estas reescrituras
-parciales son combinadas en reescrituras completas. La diferencia principal
-entre los algoritmos es el conjunto de criterios usado para elegir las vistas
-relevantes para reducir el espacio de reescrituras no útiles.
-
-El algoritmo bucket reduce el número de posibilidades considerando sólo cada
-subobjetivo de la consulta aislado, y seleccionando las vistas que son capaces
-de producir al menos los atributos proyectados por la consulta. Como los
-atributos involucrados en los joins de la consulta no son verificados, un número
-grande de reescrituras comprendido por productos cartesianos puede ser generado.
-
-El algoritmo de reglas inversas construye un conjunto de reglas que invierten
-las definiciones de vistas y establecen cómo computar tuplas para las relaciones
-de la base de datos desde las tuplas de las vistas. Al igual que el algoritmo
-bucket, puede producir un gran número de reescrituras no útiles.
-
-El algoritmo MiniCon supera las limitaciones de los algoritmos previos
-identificando solamente vistas que reescriben un conjunto de los subobjetivos de
-la consulta, y que pueden ser combinados con el resto de los subobjetivos. La
-idea principal es identificar las asociaciones entre variables de cada
-subobjetivo a las variables en uno o más subojbetivos en las vistas, de manera
-que las variables de join de la consulta son asociadas a las variables de join
-del cuerpo de una vista o a variables distinguidas de la vista. Asociaciones
-entre variables y subojbetivos son representadas en las llamadas TODOemphMiniCon
-Descriptions (MCDs) XREF29W.
-
-Finalmente, el algoritmo McdSat es capaz de identificar las reescrituras de una
-consulta traduciendo el problema de reescritura al problema de enumeración de
-modelos de una teoría proposicional cuyos modelos están en correspondencia con
-las reescrituras de la consulta. El algoritmo explota las propiedades de los
-d-DNNF para enumerar eficientemente los modelos de la teoría. El algoritmo
-McdSat ha demostrado escalar mejor que el algoritmo MiniCon sobre un número
-grande de experimentos donde frecuentemente muestra mejoras de rendimiento de
-varios órdenes de magnitud. Sin embargo, el algoritmo McdSat no fue diseñado
-para problemas de reescritura que involucren constantes explícitas, ni para
-computar las mejores reescrituras respecto a una función de utilidad o modelo de
-costo dado. En este trabajo se propone una codificación extendida que supera
-estas limitaciones y se aplica la codificación al problema de instanciación de
-flujos de trabajo.
-
-\end{document}
+\end{titlepage}

### 2010-01-12 tesis/propuesta/secciones/justificacion.tex
parte de ejemplo motivante agregada a la propuesta

\section{Justificación e importancia}

\subsection{Ejemplo motivante}

% flights
\newcommand{\flight}{\text{\it flight}}
\newcommand{\UScity}{\text{\it uscity}}
\newcommand{\national}{\text{\it national}}
\newcommand{\oneway}{\text{\it one-way}}
\newcommand{\onestop}{\text{\it one-stop}}
\newcommand{\flightPA}{\text{\it flight-to-pa}}
\newcommand{\onestopPA}{\text{\it onestop-to-pa}}
\newcommand{\fromNY}{\text{\it from-ny}}
\newcommand{\PA}{\text{PA}}
\newcommand{\NY}{\text{NY}}
\newcommand{\AL}{\text{AL}}
% end flights

Considere un sistema simple de información de vuelos que contiene información
sobre vuelos entre ciudades e información sobre qué ciudades están en los
Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
abstractos $\flight(x,y)$ y $\UScity(x)$. El primero relaciona dos ciudades $x$ e $y$ si
hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
E.E.U.U.  o no. Para los servicios concretos, supongamos que las fuentes de
datos disponibles en Internet contienen la siguiente información:

\begin{enumerate}[--]
\item $\national(x,y)$ relaciona dos ciudades de E.E.U.U. conectadas por un vuelo directo,
\item $\oneway(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,
\item $\onestop(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,
\item $\flightPA(x)$ dice si hay un vuelo directo desde $x$ a París,
\item $\onestopPA(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a París con una escala en $y$, y
\item $\fromNY(x)$ dice si hay un vuelo de Nueva York a $x$.
\end{enumerate}

Adicionalmente, los servicios concretos se describen con los siguientes
servicios abstractos:

\begin{alignat*}{1}
\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
\end{alignat*}

Ahora supongamos que un usuario está interesado en construir un flujo de trabajo
capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
en el mundo, tales que los vuelos puedan detenerse en cualquier ciudad.

\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]

La siguiente consulta conjuntiva representa el flujo de trabajo que define esta
solicitud en términos de servicios abstractos. El flujo de trabajo es definido
de manera que todo lo relacionado sobre el ligamiento de parámetros de
entrada/salida es resuelto. Así, cualquier instanciación de los servicios
abstractos en términos de los servicios concretos es una implementación válida
del flujo de trabajo. Las implementaciones corresponden a composiciones de
servicios concretos en las cuales un servicio concreto puede implementar uno o
más servicios abstractos del flujo, pero cada servicio abstracto puede ser
implementado por exactamente un servicio concreto. Por ejemplo, la siguiente
composición corresponde a una de estas implementaciones.

\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]

Sin embargo, las siguientes dos composiciones no son válidas.

\begin{alignat*}{1}
I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
\end{alignat*}

La primera composición no es válida porque asocia la variable del flujo $y$ a
las constantes \PA\ y \NY\ que denotan ciudades diferentes. En la otra mano, $I''$
no lo implementa porque el servicio concreto $\onestop(x,y)$ no recibe como entrada
o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces
no es posible asegurarse de que esa ciudad esté asociada a la ciudad $w$ que es
retornada por el flujo de trabajo.

Estos ejemplos muestran que las instanciaciones correctas de un flujo deben
manejar constantes de manera que dos constantes distintas no sean asociadas
entre sí de manera directa o indirectamente vía transitividad, y que todos los
atributos que aparezcan en un \emph{join} o en la salida sean producidos por los
servicios concretos seleccionados.

Los parámetros QoS son modelados anotando los servicios concretos con utilidades
que caracterizan su comportamiento y que luego son agregados durante la
instanciación. Así, como se mencionó previamente, la mejor instanciación es la
que minimiza (o maximiza) la agregación de utilidades.

### 2010-01-12 tesis/propuesta/traduccion/traduccion.tex
traduccion sin las partes ya copiadas

@@
 
-\title{Instanciación Óptima de Flujos de Trabajo Abstractos Usando Lógica y Circuitos}
-\maketitle
+\tableofcontents
 
-\section{Resumen}
-
-La Arquitectura Orientada a Servicios típica consiste de dos capas principales.
-La capa concreta con servicios concretos cuyas funcionalidades son descritas en
-términos de pre- y post-condiciones y propiedades no funcionales en términos de
-parámetros de calidad de servicio (Quality Of Service, QoS), y la capa abstracta
-con aplicaciones de software cuyas funcionalidades son descritas en términos de
-flujos de trabajo abstractos y propiedades no funcionales en términos de
-restricciones sobre QoS. Durante la ejecución de un flujo de trabajo, los
-servicios abstractos son instanciados en servicios concretos que cumplen con los
-requerimientos funcionales y no funcionales. Esta instanciación, que debe ser
-hecha en el momento, consiste de una búsqueda en un espacio de posibilidades
-combinatorio. En este trabajo se propone una infraestructura para resolver
-eficientemente el problema de instanciación que está firmemente basada en
-lógica. La infraestructura adopta la aproximación Local-As-View en la cual la
-funcionalidad de servicios concretos se describe usando vistas de servicios
-abstractos, la calidad de una instanciación como una función de utilidad global
-que combina los diferentes parámetros QoS, y los flujos de trabajo abstractos
-como consultas conjuntivas sobre servicios abstractos. Usando esta
-representación, el problema de instanciación de flujos de trabajo se convierte
-en un problema de reescritura de consultas, del área de sistemas de integración.
-Entonces, construyendo sobre trabajo relacionado, se idea una codificación del
-problema de instanciación de flujos de trabajo como una teoría lógica cuyos
-modelos están en correspondencia con las instanciaciones del flujo de trabajo, y
-los modelos mejor puntuados en correspondencia con instanciaciones óptimas del
-flujo de trabajo. Así, explotando propiedades conocidas de teorías lógicas en
-formato d-DNNF, se provee una solución eficiente y escalable al problema de
-instanciación de flujos de trabajo. Esta aproximación no sólo escala a
-instancias grandes como muestran resultados experimentales ya realizados, sino
-que también es correcta y completa dado que, estándo basada en la lógica, es
-amigable al análisis formal.
-
-\section{Introducción}
-
-Bajo la sombra de la Web Semántica, y soportado por Arquitecturas Orientadas a
-Servicios (SOA), el número de fuentes de datos y servicios Web ha explotado en
-los últimos años. Por ejemplo, la colección de bases de datos de biología
-molecular actualmente contiene 1170 bases de datos XREF22W, un número que es
-mayor que el del último año XREF21W por 95, y que el de hace dos años por 110
-XREF20W; las herramientas y servicios, así como el número de instancias
-publicados por estos recursos, siguen una progresión similar XREF6W. Gracias a
-este tesoro de datos, la tendencia de los usuarios es hacia depender cada vez
-más en métodos automáticos para manejarlos tales como recuperación de datos de
-fuentes públicas y análisis utilizando herramientas o servicios Web compuestos
-en flujos de trabajo complejos.
-
-La SOA típica consiste de dos capas. La capa concreta que está hecha de
-servicios concretos cuyas funcionalidades son descritas en términos de pre- y
-post-condiciones y propiedades no funcionales en términos de parámetros QoS, y
-la capa abstracta compuesta de aplicaciones de software cuyas funcionalidades
-son descritas en términos de flujos de trabajo abstractos y propiedades no
-funcionales en términos de restricciones de QoS. La ejecución de un flujo de
-trabajo abstracto involucra la instanciación de los servicios abstractos en
-servicios concretos que cumplen con los requerimientos funcionales y no
-funcionales. Este proceso de instanciación puede ser visto como una búsqueda de
-una instanciación objetivo en el espacio combinatorio de todas las
-instanciaciones válidas. Así, uno está interesado en técnicas eficientes para
-realizar esta búsqueda que sean capaces de escalar a medida que el número de
-servicios concretos o la complejidad del flujo de trabajo aumenta. Se llama al
-problema de instanciar un flujo de trabajo dado con servicios concretos, de un
-conjunto dado de servicios concretos, de manera que ciertas demandas de QoS sean
-cumplidas, el Workflow Instantiation Problem (WIP, "Problema de Instanciacion de
-Flujos de Trabajo").
-
-En este trabajo se considera una versión restringida de WIP que adopta la
-aproximación Local-As-View (LAV) XREF28W. En LAV, todos los elementos de un
-problema son especificados con un lenguaje común basado en servicios abstractos
-tal que los servicios concretos se describen como vistas de servicios
-abstractos, la calidad de una instanciación como una función de utilidad global
-que combina los diferentes parámetros QoS, y el flujo de trabajo abstracto como
-una consulta conjuntvia sobre los servicios abstractos; esta representación es
-similar a la que es generada de manera semiautomática para el sistema DEIMOS
-XREF3W. En esta versión de WIP, las reglas que definen flujos de trabajo
-(consultas conjuntivas) y servicios concretos (vistas) son creados de manera que
-todas las restricciones funcionales sobre pre- y post-condiciones de los
-servicios y sus combinaciones sean satisfechas, y las medidas de QoS son
-representadas anotando cada descripción de servicio concreto con un número real
-que representa la utilidad QoS global del servicio.
-
-Bajo estas suposiciones, WIP puede ser convertido en el bien conocido Problema
-de Reescritura de Consultas (QRP) para LAV que es central a los sistemas de
-integración XREF23W. QRP consiste de una consulta conjuntiva que debe ser
-respondida en términos de vistas donde la consulta y las vistas son descritas
-usando LAV con relaciones abstractas. Este problema es importante en el contexto
-de integración de datos XREF13X24W, y optimización de consultas y mantenimiento
-de datos XREF1,28W, y varias aproximaciones que escalan a un número grande de
-vistas han sido definidas XREF4,18,19,28,29W.
-
-La aproximación recente de Arvelo et al. XREF4W está basada en la enumeración
-eficiente de modelos de una teoría lógica proposicional. Dado unQRP, una teoría
-lógica es construída de manera que cada modelo de la teoría codifica una
-reescritura válida, y así todas las reescrituras son obtenidas enumerando los
-modelos de la teoría. Esta enumeración puede ser realizada eficientemente si la
-teoría lógica está en cierta forma normal llamada la forma normal de negación
-determinística y descomponible (d-DNNF por sus siglas en inglés) XREF14W. Así,
-la aproximación consiste en transformar (llamado compilar en el campo de
-compilación del conocimiento) la teoría lógica en el formato d-DNNF para
-enumerar sus modelos eficientemente.
-
-Pero las teorías d-DNNF no sólo soportan la enumeración eficiente de sus
-modelos, sino también otras operaciones. Entre ellas se encuentra la enumeración
-de los modelos con mejor puntuación. Dada una función de puntuación de literales
-$r(l)$ que asigne puntuaciones a cada literal $l$, se define la puntuación
-$r(w)$ de un modelo $w$ como la suma de las puntuaciones de los literales
-activados por $w$ (es decir, $r(w) = suma...$, y se dice que $w$ es un modelo
-mejor puntuado si no hay modelo $wpriemrie$ tal que $r(w)...$.
-
-Dada una teoría en d-DNNF, se computa la puntuación de los mejores modelos, y
-los mejores modelos, en tiempo lineal en el tamaño del d-DNNF. Este cómputo
-transforma el GAD del d-DNNF en un circuito aritmético reemplazando los nodos
-AND con '+' y los nodos OR con 'min'. La función de puntuación de literales
-asigna valores a las hojas del circuito que son propagados a la raíz en tiempo
-lineal. El valor de la raíz es la puntuación del mejor modelo XREF17W.
-
-En este trabajo, se explotan las propiedades de los d-DNNF construyendo una
-teoría lógica cuyos modelos codifican las instanciaciones del flujo de trabajo
-y cuyos mejores modelos codifican las instanciaciones óptimas (mejores). Así, la
-búsqueda combinatoria se reduce a la computación de un mejor modelo de una
-teoría lógica que puede ser realizada eficientemente una vez que la teoría sea
-transformada en formato d-DNNF.
-
-Esta propuesta contiene seis secciones más. La próxima sección ilustra y motiva
-WIP con un ejemplo simple pero típico. La sección 3 resume el trabajo existente
-relacionado en tres áreas de selección de servicios, reescritura de consultas y
-compilación del conocimiento. Entonces, las secciones 4 y 5 describen la
-arquitectura del sistema y reportan los resultados empíricos obtenidos hasta
-ahora sobre distintos problemas de prueba respectivamente. La sección 6 presenta
-una descripción formal de la aproximación propuesta junto con un análisis. La
-propuesta concluye con una discusión final en la sección 7.
-
-\section{Ejemplo motivante}
-
-Considere un sistema simple de información de vuelos que contiene información
-sobre vuelos entre ciudades e información sobre qué ciudades están en los
-Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
-abstractos $flight$ y $uscity$. El primero relaciona dos ciudades $x$ e $y$ si
-hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
-E.E.U.U.  o no. Para los servicios concretos, supongamos que las fuentes de
-datos disponibles en Internet contienen la siguiente información:
-
-\begin{itemize}
-TODO
--national
-\end{itemize}
-
-Adicionalmente, los servicios concretos se describen con los siguientes
-servicios abstractos:
-
-\begin{itemize}
-TODO
--national
-\end{itemize}
-
-Ahora supongamos que un usuario está interesado en construir un flujo de trabajo
-capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
-en el mundo, tales que los vuelos puedan detenerse en cualquier ciudad.
-
-$$
-W(x,w,y,z)
-$$
-
-La siguiente consulta conjuntiva representa el flujo de trabajo que define esta
-solicitud en términos de servicios abstractos. El flujo de trabajo es definido
-de manera que todo lo relacionado sobre el ligamiento de parámetros de
-entrada/salida es resuelto. Así, cualquier instanciación de los servicios
-abstractos en términos de los servicios concretos es una implementación válida
-del flujo de trabajo. Las implementaciones corresponden a composiciones de
-ervicios concretos en las cuales un servicio concreto puede implementar uno o
-más servicios abstractos del flujo, pero cada servicio abstracto puede ser
-implementado por exactamente un servicio concreto. Por ejemplo, la siguiente
-composición corresponde a una de estas implementaciones.
-
-$$
-I(x
-$$
-
-Sin embargo, las siguientes dos composiciones no son válidas.
-
-$$
-I'
-I''
-$$
-
-La primera composición no es válida porque asocia la variable del flujo $y$ a
-las constantes PA y NY que denotan ciudades diferentes. En la otra mano, $I''$
-no lo implementa porque el servicio concreto $ones-stop$ no recibe como entrada
-o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces
-no es posible asegurarse de que esa ciudad esté asociada a la ciudad $w$ que es
-retornada por el flujo de trabajo.
-
-Estos ejemplos muestran que las instanciaciones correctas de un flujo deben
-manejar constantes de manera que ods constantes distintas no sean asociadas
-entre sí de manera directa o indirectamente vía transitividad, y que todos los
-atributos que aprezcan en un TODOemphjoin o en la salida sean producidos por los
-servicios concretos seleccionados.
-
-Los parámetros QoS son modelados anotando los servicios concretos con utilidades
-que caracterizan su comportamiento que luego son agregados durante la
-instanciación. Así, como se mencionó previamente, la mejor instanciación es la
-que minimiza (o maximiza) la agregación de utilidades.
-
-\emph{Trabajo Relacionado}
+\section{Trabajo Relacionado}
 
@@
 vistas. Una instanciación $R$ es válida si para todos los catálogos $C=$ y
-extensiones $EIII$, $R(E#)$. Una colección $R$ de instanciaciones válidas es una
+extensiones $EIII$, $R(E\#)$. Una colección $R$ de instanciaciones válidas es una
 solución si para todos los catálogos de servicio $C=$ y extensiones $EII$, no
@@
 
+\subsection{Teorías Lógicas}
+
+Se usa una aproximación similar a la descrita en XREF4W para codificar el WIP.
+Se identifican instanciaciones del flujo de trabajo enumerando los modelos de
+una teoría lógica.
+
 \end{document}

### 2010-01-12 tesis/propuesta/secciones/justificacion.tex
parte de resultados experimentales integrada

@@
 que minimiza (o maximiza) la agregación de utilidades.
+
+\subsection{Resultados Experimentales}
+
+Se realizó un análisis empírico sobre tres experimentos. Todos los experimentos
+fueron realizados en una máquina de escritorio con un CPU Intel Core 2 Duo de
+2GHz y 4GB de memoria, y el tiempo fue medido con el comando de Unix
+\emph{time}.
+
+El objetivo de los experimentos es determinar el rendimiento de la propuesta con
+condiciones variantes. El principal beneficio de la aproximación propuesta es
+que se puede compilar la teoría lógica para una instancia del problema y luego
+calcular todas las instanciaciones, o las mejores, cualquier número de veces. El
+modelo de costos para conseguir mejores instanciaciones puede ser cambiado sin
+necesidad de recompilar la teoría lógica. Por lo tanto, la complejidad en tiempo
+de la aproximación es básicamente el tiempo para codificar el WIP como un
+CNF más el tiempo para compilar el CNF en un d-DNNF y el tiempo para decodificar
+los modelos. Los tiempos para codificar y decodificar son despreciables
+comparados con el tiempo para compilar el CNF. Por esto, el enfoque es en el tiempo
+necesario para compilar los problemas de los experimentos.
+
+El primer experimento consiste de problemas para consultas para viajes aéreos.
+Los servicios concretos son de la forma $V_i(x,y)\qrule\ \flight(x,y,\AL_i)$
+donde $\AL_i$ es una
+constante que denota el nombre de una aerolínea, y $\flight(x,y,\AL_i)$ relaciona las
+ciudades $x$ e $y$ tales que hay un vuelo entre $x$ e $y$ servido por $\AL_i$.
+Se supone que este servicio concreto retorna todos los vuelos entre dos ciudades
+con una aerolínea específica. El flujo de trabajo tiene la forma
+
+\[ W(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,z),\,\flight(x_1,x_2,z),\,\ldots,\,\flight(x_n,\NY,z)\,. \]
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1\textwidth]{graphics/plot1}
+\caption{Tiempos de compilación para los experimentos I y II para diferentes
+números de objetivos y de vistas. Los gráficos están en escala logarítmica, y el
+tiempo es en segundos.}
+\label{fig:plot1}
+\end{figure}
+
+El experimento incluye instancias para flujos de trabajo con 2 a 5 subobjetivos
+y conjuntos de 10 a 100 servicios concretos. Los resultados para la compilación
+son mostrados en el panel (a) de la figura ~\ref{fig:plot1}. Este es un gráfico en escala
+logarítmica que sugiere comportamiento sub-exponencial. En cualquier caso, los
+resultados muestran buen rendimiento dado que instancias realísticas del
+problema (conjuntos de 100 aerolíneas con vuelos de cinco paradas) pueden ser
+compiladas en 328 segundos. El tamaño en disco del d-DNNF para 100 aerolíneas y
+vuelos con cinco paradas es 3,4MB. En este d-DNNF, el mejor modelo puede ser
+computado en 0.29 segundos, y la enumeración de todos los modelos en 0.47
+segundos.
+
+En un intento de inducir crecimiento exponencial en el tamaño de compilación, en
+el segundo experimento se agrega un segundo servicio concreto para cada
+aerolínea. Esta modificación incrementa el número de instanciaciones válidas de
+lineal a exponencial dado que cada tramo de un vuelo puede ser ahora instanciado
+por dos servicios concretos y entonces un vuelo con $n$ tramos puede tener hasta
+$2^n$ instanciaciones. Se corrió el compilador para instancias que comprendían
+el mismo número de objetivos del flujo de trabajo y el número total de
+servicios concretos. Los resultados graficados en escala logarítmica se muestran
+en el panel (b) de la figura ~\ref{fig:plot1}.
+
+Estas pruebas muestran buen rendimiento para este tipo de problemas, pero no
+involucran servicios concretos con múltiples objetivos. Por lo tanto se
+diseñó un tercer experimento que consiste de instancias no estructuradas y
+generadas aleatoriamente. Cada instancia contiene tres variables por servicio
+abstracto, diez variables distintas y diez constantes distintas, seis
+objetivos en los flujos de trabajo, 2 a 5 objetivos en los servicios
+concretos, y número de servicios variante. La probabilidad de que un argumento
+de un servicio abstracto esté ligado a una constante es 50\%. Los resultados se
+muestran en la figura ~\ref{fig:plot3}. El tiempo de compilación de estas instancias no
+crece monótonamente dado que son generadas aleatoriamente. Lo mismo ocurre para
+el tamaño de las teorías y los números de modelos. Por ejemplo, el d-DNNF para
+un problema con 45 vistas cada una con 5 objetivos fue de tamaño 5,1Mb y tuvo
+$1.26\times 10^8$ modelos. El tiempo para buscar el mejor modelo para este
+d-DNNF es 0.46 segundos mientras que el tiempo para enumerar todos los modelos
+es alredededor de 17 horas.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=.8\textwidth]{graphics/plot3}
+\caption{Tiempos de compilación para el experimento III para diferentes números
+de objetivos y de vistas. Los gráficos están en escala logarítmica, y el
+tiempo es en segundos.}
+\label{fig:plot3}
+\end{figure}
+
+Estos son experimentos preliminares, pero los resultados muestran que la
+aproximación propuesta escala eficientemente para problemas con varios objetivos
+y vistas. Se cree que estos resultados son alentadores y motivan a continuar
+esta línea de investigación.

### 2010-01-12 tesis/propuesta/traduccion/traduccion.tex
parte de resultados experimentales integrada

@@
 
-\section{Resultados Experimentales}
-
-Se realizó un análisis empírico sobre tres experimentos. Todos lso experimentos
-fueron realizados en una máquina de escritorio con un CPU Intel Core 2 Duo de
-2GHz y 4GB de memoria, y el tiempo fue medido con el comando de Unix
-timeTODOemph.
-
-El objetivo del experimento es determinar el rendimiento de la propuesta con
-condiciones variantes. El principal beneficio de la aproximación propuesta es
-que se puede compilar la teoría lógica para una instancia del problema y luego
-calcular todas las instanciaciones, o las mejores, cualquier número de veces. El
-modelo de costos para conseguir mejores instanciaciones puede ser cambiado sin
-necesidad de recompilar la teoría lógica. Por lo tanto, la complejidad en tiempo
-de nuestra aproximación es básicamente el tiempo para codificar el WIP como un
-CNF más el tiempo para compilar el CNF en un d-DNNF y el tiempo para decodificar
-los modelos. Los tiempos para codificar y decodificar son despreciables
-comparados con el tiempo para compilar el CNF. Por esto, se enfoca en el tiempo
-necesario para compilar los problemas de los experimentos.
-
-El primer experimento consiste de problemas para consultas para viajes aéreos.
-Los servicios concretos son de la forma $VTODO formula$ donde $AL_iTODO$ es una
-constante que denota el nombre de una aerolínea, y $flightTODO$ relaciona las
-ciudades $x$ e $y$ tales que hay un vuelo entre $x$ e $y$ servido por $ALITODO$.
-Se supone que este servicio concreto retorna todos los vuelos entre dos ciudades
-con una aerolínea específica. El flujo de trabajo tiene la forma
-
-$$
-Wx1...TODO
-$$
-
-El experimento incluye instancias para flujos de trabajo con 2 a 5 subobjetivos
-y conjuntos de 10 a 100 servicios concretos. Los resultados para la compilación
-son mostrados en el panel (a) de la figura TODO3. Este es un gráfico en escala
-logarítmica que sugiere comportamiento sub-exponencial. En cualquier caso, los
-resultados muestran buen rendimiento dado que instancias realísticas del
-problema (conjuntos de 100 aerolíneas con vuelos de cinco paradas) pueden ser
-compiladas en 328 segundos. El tamaño en disco del d-DNNF para 100 aerolíneas y
-vuelos con cinco paradas es 3,4MB. En este d-DNNF, el mejor modelo puede ser
-computado en 0.29 segundos, y la enumeración de todos los modelos en 0.47
-segundos.
-
-TODOfig3 experimentos
-
-En un intento de inducir crecimiento exponencial en el tamaño de compilación, en
-el segundo experimento se agrega un segundo servicio concreto para cada
-aerolínea. Esta modificación incrementa el número de instanciaciones válidas de
-lineal a exponencial dado que cada tramo de un vuelo puede ser ahora instanciado
-por dos servicios concretos y entonces un vuelo con $n$ tramos puede tener hasta
-$2^n$ instanciaciones. Se corrió el compilador para instancias que comprendían
-el mismo número de subobjetivos del flujo de trabajo y el número total de
-servicios concretos. Los resultados graficados en escala logarítmica se muestran
-en el panel (b) de la figura TODO3.
-
-Estas pruebas muestran buen rendimiento para este tipo de problemas, pero no
-involucran servicios concretos con múltiples subobjetivos. Por lo tanto se
-diseñó un tercer experimento que consiste de instancias no estructuradas y
-generadas aleatoriamente. Cada instancia contiene tres variables por servicio
-abstracto, diez variables distintas y diez constantes distintas, seis
-subobjetivos en los flujos de trabajo, 2 a 5 subobjetivos en los servicios
-concretos, y número de servicios variante. La probabilidad de que un argumento
-de un servicio abstracto esté ligado a una constante es 50\%. Los resultados se
-muestran en la figura TODO4. El tiempo de compilación de estas instancias no
-crece monótonamente dado que son generadas aleatoriamente. Lo mismo ocurre para
-el tamaño de las teorías y los números de modelos. Por ejemplo, el d-DNNF para
-un problema con 45 vistas cada una con 5 subobjetivos fue de tamaño 5,1Mb y tuvo
-$1.26 TODO 20101028$ modelos. El tiempo para buscar el mejor modelo para este
-d-DNNF es 0.46 segundos mientras que el tiempo para enumerar todos los modelos
-es alredededor de 17 horas.
-
-Estos son experimentos preliminares, pero los resultados muestran que la
-aproximación propuesta escala eficientemente para problemas con varios objetivos
-y vistas. Se cree que estos resultados son alentadores y motivan a continuar
-esta línea de investigación.
-
 \section{Formalización}

### 2010-01-12 tesis/propuesta/secciones/planteamiento.tex
parte de ejemplo motivante agregada a la propuesta
parte de trabajo relacionado integrada a la propuesta

@@
 transformada al formato d-DNNF.
+
+\subsection{Trabajo Relacionado}
+
+En esta sección se resumen las aproximaciones existentes que proveen soluciones
+a los problemas de selección de servicios y reescritura de consultas y se
+discute el trabajo relacionado en el área de inteligencia artificial llamado
+compilación del conocimiento.
+
+\subsubsection{Selección de Servicios}
+
+El problema de seleccionar los servicios que implementan un flujo de trabajo
+abstracto y cumplen mejor los criterios basados en QoS es conocido como el
+problema de selección o composición de servicios consciente de QoS, que ha sido
+demostrado ser NP-difícil \cite{Hiroshi2008}. Este es un problema de optimización
+combinatoria y varias heurísticas han sido propuestas para conseguir una
+solución relativamente buena en un período razonablemente corto de tiempo.
+
+Rahmani et al. \cite{rahmani08} presentan una heurística basada en métrica de distancia
+que guía un algoritmo de búsqueda hacia atrás; esta métrica induce un orden de
+los servicios en una manera en que los nodos vertedero tienen baja probabilidad
+de ser visitados. En una serie de artículos, Berardi y otros \cite{berardi05,berardi08,berardi06}
+describen
+servicios y flujos de trabajo en términos de máquinas de estados finitos
+determinísticas que son codificadas usando teorías de Lógica de Descripción
+cuyos modelos corresponden a soluciones del problema. Aunque se podrían explotar
+métodos para formalismos de Lógicas de Descripción, no han sido reportados la
+escalabilidad o el rendimiento de la solución propuesta.
+
+Ko et al. \cite{myoung08} proponen una aproximación basada en restriccciones que codifica
+los valores permisibles no funcionales como un conjunto de restricciones cuya
+violación debe ser minimizada; para recorrer el espacio de soluciones
+posiblemente óptima, se implementa un algoritmo híbrido que combina las
+metaheurísticas \emph{tabu search} y \emph{simulated annealing}. Los resultados
+experimentales muestran que la solución propuesta es capaz de escalar a un
+número grande de servicios y procesos abstractos. Cardellini et al. \cite{cardellini07}
+codifican una parte del problema de composición de servicios consciente de QoS
+como un problema de programación lineal \cite{cardellini07}. Por otra parte, Wada et al.
+\cite{Hiroshi2008} trata el problema como uno de optimización con múltiples objetivos donde
+los diferentes parámetros de QoS son considerados igualmente importantes en vez
+de ser agregados en una función. Luego se propone un algoritmo basado en
+algoritmos genéticos que identifica un conjunto de composiciones de servicios
+no dominadas que mejor encajan en los requerimientos QoS.
+
+Alrifai y Risse \cite{alrifaiR09} proponen una solución de doble cara que usa un
+algoritmo de programación entera híbrido para obtener la descomposición de QoS
+globales en restricciones locales y luego selecciona los servicios que mejor
+cumplen las restricciones locales.
+
+Recientemente, dos aproximaciones basadas en planificación han sido propuestas.
+Kuter y Golbeck \cite{kuterG09} extienden el algoritmo de planificación SHOP2 para
+seleccionar la composición confiable de servicios que implementan un modelo de
+procesos OWL-S dado, mientras que Sohrabi y McIlraith \cite{sohrabiM09} proponen una
+solución basada en planificación HTN donde las métricas de preferencia del
+usario y regulaciones de dominio son utilizadas para guíar el planificador hacia
+el espacio de composiciones relevantes. Finalmente, Lécué \cite{lecue09} propone un
+algoritmo basado en genética para identificar la composición de servicios que
+mejor cumple con los criterios de calidad para un conjunto de parámetros QoS.
+
+Aunque estas soluciones son capaces de resolver el problema de optimización y
+escalar a un número de procesos abstractos, ninguno de ellos está ajustado para
+describir semánticamente los servicios en términos de procesos abstractos, ni
+para usar estas descripciones para identificar los servicios que implementan un
+flujo de trabajo dado o que mejor cumplen los criterios no funcionales del
+usuario.
+
+\subsubsection{Reescritura de Consultas}
+
+Un número de algoritmos han sido desarrollados para encontrar las reescrituras
+de una consulta dada; los más prominentes son el algoritmo bucket \cite{levy:bucket}, el
+algoritmo de reglas inversas \cite{pods:DuschkaG97,Qian96}, el algoritmo
+\minicon \cite{pottinger:minicon}, y el
+algoritmo \mcdsat \cite{arvelo:aaai06}. Generalmente, la reescritura de consultas trabaja en
+dos fases. Durante la primera, el algoritmo identifica las vistas que reescriben
+al menos un subobjetivo de la consulta, y durante el segundo estas reescrituras
+parciales son combinadas en reescrituras completas. La diferencia principal
+entre los algoritmos es el conjunto de criterios usado para elegir las vistas
+relevantes para reducir el espacio de reescrituras no útiles.
+
+El algoritmo bucket reduce el número de posibilidades considerando sólo cada
+subobjetivo de la consulta aislado, y seleccionando las vistas que son capaces
+de producir al menos los atributos proyectados por la consulta. Como los
+atributos involucrados en los \emph{joins} de la consulta no son verificados, un número
+grande de reescrituras comprendido por productos cartesianos puede ser generado.
+
+El algoritmo de reglas inversas construye un conjunto de reglas que invierten
+las definiciones de vistas y establecen cómo computar tuplas para las relaciones
+de la base de datos a partir de las tuplas de las vistas. Al igual que el algoritmo
+bucket, puede producir un gran número de reescrituras no útiles.
+
+El algoritmo \minicon\ supera las limitaciones de los algoritmos previos
+identificando solamente vistas que reescriben un conjunto de los subobjetivos de
+la consulta, y que pueden ser combinados con el resto de los subobjetivos. La
+idea principal es identificar las asociaciones entre variables de cada
+subobjetivo a las variables en uno o más subojbetivos en las vistas, de manera
+que las variables de \emph{join} de la consulta son asociadas a las variables de
+\emph{join}
+del cuerpo de una vista o a variables distinguidas de la vista. Asociaciones
+entre variables y subojbetivos son representadas en las llamadas \emph{MiniCon
+Descriptions} (MCDs) \cite{pottinger:minicon}.
+
+Finalmente, el algoritmo \mcdsat\ es capaz de identificar las reescrituras de una
+consulta traduciendo el problema de reescritura al problema de enumeración de
+modelos de una teoría proposicional cuyos modelos están en correspondencia con
+las reescrituras de la consulta. El algoritmo explota las propiedades de los
+d-DNNF para enumerar eficientemente los modelos de la teoría.
+El algoritmo
+\mcdsat\ ha demostrado escalar mejor que el algoritmo \minicon\ sobre un número
+grande de experimentos donde frecuentemente muestra mejoras de rendimiento de
+varios órdenes de magnitud. Sin embargo, el algoritmo \mcdsat\ no fue diseñado
+para problemas de reescritura que involucren constantes explícitas, ni para
+computar las mejores reescrituras respecto a una función de utilidad o modelo de
+costo dado. En este trabajo se propone una codificación extendida que supera
+estas limitaciones y se aplica la codificación al problema de instanciación de
+flujos de trabajo.
+
+\subsubsection{Compilación del Conocimiento}
+
+Compilación del conocimiento es el área en inteligencia artificial que se ocupa
+del problema de traducir teorías lógicas en fragmentos apropiados que hagan
+tratables ciertas operaciones deseadas \cite{cadoli:compilation}. Diferentes lenguajes de
+compilación han sido definidos. Por ejemplo están los \emph{Ordered
+Binary Decision Diagrams} (OBDDs) \cite{bryant:obdd},
+\emph{Negation Normal Form} (NNF) \cite{barwise:handbook}, y
+\emph{Decomposable Negation Normal Form} (DNNF) \cite{darwiche:map}. En este trabajo se usan las
+propiedades de los DNNF determinísticos (d-DNNF) \cite{darwiche:d-dnnfs} para proveer una
+solución eficiente y escalable al problema de selección de serviccios.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=.7\textwidth]{graphics/odd}
+\caption{Un NNF descomponible y determinístico.}
+\label{fig:dnnf}
+\end{figure}
+
+Una teoría en Negation Normal Form es construida desde literales usando sólo
+conjunciones y disjunciones \cite{barwise:handbook}, y puede ser representada como un grafo
+acíclico dirigido en que las hojas son etiquetadas con literales y los nodos
+internos con $\land$ y $\lor$; ver ejemplo en la figura~\ref{fig:dnnf}. Este
+fragmento es universal, lo que significa que para cada fórmula lógica hay una
+equivalente en formato NNF. Se dice que un NNF es descomponible (DNNF)
+\cite{darwiche:d-dnnfs}
+si para cada conjunción $\bigwedge_i\phi_i$, sus variables son disjuntas por pares; es
+decir, $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ para $i\neq j$. Un DNNF soporta un
+número de operaciones en tiempo polinomial en el tamaño de su GAD. Por ejemplo,
+podemos probar si un DNNF es satisfactible haciendo una simple pasada de abajo
+hacia arriba sobre su GAD en tiempo linear. Se dice que un DNNF es
+determinístico (d-DNNF) \cite{darwiche:d-dnnfs} si para cada disjunción $\bigvee_i\phi_i$, los
+disjuntos son lógicamente contradictorios por pares, es decir,
+$\phi_i\land\phi_j\equiv\textbf{false}$ para $i\neq j$.
+El NNF en la figura~\ref{fig:dnnf}, por ejemplo, es
+descomponible y determinístico. Un d-DNNF soporta conteo de modelos en
+tiempo polinomial en el tamaño de su GAD, y enumeración de modelos en tiempo
+polinomial en el tamaño de la salida. Además, dada una función de puntuación de
+literales $r$, se puede computar la puntuación del mejor modelo en tiempo
+polinomial para DNNFs \cite{darwiche:weighted}.
+
+Los fragmentos DNNF y d-DNNF son universales pero traducir una teoría CNF a
+formato DNNF tiene costo exponencial en el peor caso. Esta traducción se
+llama compilación en este campo. Hay un compilador públicamente disponible,
+llamado c2d\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}, que realiza este proceso de compilación y que hace uso
+de técnicas de SAT modernas como backtracking dirigido por conflictos,
+aprendizaje de cláusulas y caching \cite{darwiche:compiler}. Este compilador entra en el peor
+caso en espacio exponencial en un parámetro llamado el ancho del árbol de
+descomposición que está relacionado a la ``conectividad'' de la teoría CNF.
+Sin embargo, en nuestros experimentos, las teorías CNF que son compiladas son de
+ancho pequeño.

### 2010-01-12 tesis/propuesta/secciones/portada.tex
parte de trabajo relacionado integrada a la propuesta

@@
 \begin{flushleft}
-\includegraphics[scale=0.3]{logo_usb.png} \\
+\includegraphics[scale=0.3]{graphics/logo_usb.png} \\
 Universidad Simón Bolívar \\

### 2010-01-12 tesis/propuesta/traduccion/traduccion.tex
parte de trabajo relacionado integrada a la propuesta

@@
 \tableofcontents
-
-\section{Trabajo Relacionado}
-
-En esta sección se resumen las aproximaciones existentes que proveen soluciones
-a los problemas de selección de servicios y reescritura de consultas y se
-discute el trabajo relacionado en el área de inteligencia artificial llamado
-compilación del conocimiento.
-
-\subsection{Selección de Servicios}
-
-El problema de seleccionar los servicios que implementan un flujo de trabajo
-abstracto y cumplen mejor los criterios basados en QoS es conocido como el
-problema de selección o composición de servicios consciente de QoS, que ha sido
-demostrado ser NP-difícil XREF34W. Este es un problema de optimización
-combinatoria y varias heurísticas han sido propuestas para conseguir una
-solución relativamente buena en un período razonablemente corto de tiempo.
-
-Rahmani et al. XREF31W presenta una heurística basada en métrica de distancia
-que guía un algoritmo de búsqueda hacia atrás; esta métrica induce un orden de
-los servicios en una manera en que los nodos vertedero tienen baja probabilidad
-de ser visitados. En una serie de artículos, Berardi y otros XREF7-9W describen
-servicios y flujos de trabajo en términos de máquinas de estados finitos
-determinísticas que son codificadas usando teorías de Lógica de Descripción
-cuyos modelos corresponden a soluciones del problema. Aunque se podrían explotar
-métodos para formalismos de Lógicas de Descripción, no han sido reportados la
-escalabilidad o el rendimiento de la solución propuesta.
-
-Ko et al. XREF25W propone una aproximación basada en restriccciones que codifica
-los valores permisibles no funcionales como un conjunto de restricciones cuya
-violación debe ser minimizada; para recorrer el espacio de soluciones
-posiblemente óptima, se implementa un algoritmo híbrido que combina las
-metaheurísticas TODOemphtabusearch y TODOemphsimulatedannealing. Los resultados
-experimentales muestran que la solución propuesta es capaz de escalar a un
-número grande de servicios y procesos abstractos. Cardellini et al. XREF12W
-codifican una parte del problema de composición de servicios consciente de QoS
-como un problema de programación lineal XREF12W. Por otra parte, Wada et al.
-XREF34W trata el problema como uno de optimización con múltiples objetivos donde
-los diferentes parámetros de QoS son considerados igualmente importanes en vez
-de ser agregados en una función. Luego se propone un algoritmo basado en
-algoritmos genéticos que identifica un conjunto de composiciones de servicios
-no dominadas que mejor encajan en los requerimientos QoS.
-
-Alrifai y Risse XREF2W proponen una solución de doble cara que usa un
-algoritmo de programación entera híbrido para obtener la descomposición de QoS
-globales en restricciones locales y luego selecciona los servicios que mejor
-cumplen las restricciones locales.
-
-Recientemente, dos aproximaciones basadas en planificación han sido propuestas.
-Kuter y Golbek XREF26W extienden el algoritmo de planificación SHOP2 para
-seleccionar la composición confiable de servicios que implementan un modelo de
-procesos OWL-S dado, mientras que Sohrabi y McIlraith XREF33W proponen una
-solución basada en planificación HTN donde las métricas de preferencia del
-usario y regulaciones de dominio son utilizadas para guíar el planificador hacia
-el espacio de composiciones relevantes. Finalmente, Lécué XREF27W propone un
-algoritmo basado en genética para identificar la composición de servicios que
-mejor cumple con los criterios de calidad para un conjunto de parámetros QoS.
-
-Aunque estas soluciones son capaces de resolver el problema de optimización y
-escalar a un número de procesos abstractos, ninguno de ellos está ajustado para
-describir semánticamente los servicios en términos de procesos abstractos, ni
-para usar estas descripciones para identificar los servicios que implementan un
-flujo de trabajo dado o que mejor cumplen los criterios no funcionales del
-usuario.
-
-\subsection{Reescritura de Consultas}
-
-Un número de algoritmos han sido desarrollados para encontrar las reescrituras
-de una consulta dada; los más prominentes son el algoritmo bucket XREF28W, el
-algoritmo de reglas inversas XREF18,30W, el algoritmo MiniCon XREF29, y el
-algoritmo McdSat XREF4W. Generalmente, la reescritura de consultas trabaja en
-dos fases. Durante la primera, el algoritmo identifica las vistas que reescriben
-al menos un subobjetivo de la consulta, y durante el segundo estas reescrituras
-parciales son combinadas en reescrituras completas. La diferencia principal
-entre los algoritmos es el conjunto de criterios usado para elegir las vistas
-relevantes para reducir el espacio de reescrituras no útiles.
-
-El algoritmo bucket reduce el número de posibilidades considerando sólo cada
-subobjetivo de la consulta aislado, y seleccionando las vistas que son capaces
-de producir al menos los atributos proyectados por la consulta. Como los
-atributos involucrados en los joins de la consulta no son verificados, un número
-grande de reescrituras comprendido por productos cartesianos puede ser generado.
-
-El algoritmo de reglas inversas construye un conjunto de reglas que invierten
-las definiciones de vistas y establecen cómo computar tuplas para las relaciones
-de la base de datos desde las tuplas de las vistas. Al igual que el algoritmo
-bucket, puede producir un gran número de reescrituras no útiles.
-
-El algoritmo MiniCon supera las limitaciones de los algoritmos previos
-identificando solamente vistas que reescriben un conjunto de los subobjetivos de
-la consulta, y que pueden ser combinados con el resto de los subobjetivos. La
-idea principal es identificar las asociaciones entre variables de cada
-subobjetivo a las variables en uno o más subojbetivos en las vistas, de manera
-que las variables de join de la consulta son asociadas a las variables de join
-del cuerpo de una vista o a variables distinguidas de la vista. Asociaciones
-entre variables y subojbetivos son representadas en las llamadas TODOemphMiniCon
-Descriptions (MCDs) XREF29W.
-
-Finalmente, el algoritmo McdSat es capaz de identificar las reescrituras de una
-consulta traduciendo el problema de reescritura al problema de enumeración de
-modelos de una teoría proposicional cuyos modelos están en correspondencia con
-las reescrituras de la consulta. El algoritmo explota las propiedades de los
-d-DNNF para enumerar eficientemente los modelos de la teoría. El algoritmo
-McdSat ha demostrado escalar mejor que el algoritmo MiniCon sobre un número
-grande de experimentos donde frecuentemente muestra mejoras de rendimiento de
-varios órdenes de magnitud. Sin embargo, el algoritmo McdSat no fue diseñado
-para problemas de reescritura que involucren constantes explícitas, ni para
-computar las mejores reescrituras respecto a una función de utilidad o modelo de
-costo dado. En este trabajo se propone una codificación extendida que supera
-estas limitaciones y se aplica la codificación al problema de instanciación de
-flujos de trabajo.
-
-\subsection{Compilación del conocimiento}
-TODO Fig1
-
-Compilación del conocimiento es el área en inteligencia artificial que se ocupa
-del problema de traducir teorías lógicas en fragmentos apropiados que hagan
-tratables ciertas operaciones deseadas XREF11W. Diferentes lenguajes de
-compilación han sido definidos. Por ejemplo están los TODOemphtodoesto Ordered
-Binary Decision Diagrams (OBDDs) XREF10W, Negation Normal Form (NNF) XREF5W, y
-Decomposable Negation Normal Form (DNNF) XREF16W. En este trabajo se usan las
-propiedades de los DNNFs determinísticos (d-DNNF) XREF14W para proveer una
-solución eficiente y escalable al problema de selección de serviccios.
-
-Una teoría en Negation Normal Form es construida desde literales usando sólo
-conjunciones y disjunciones XREF5W, y puede ser representada como un grafo
-acíclico dirigido en que las hojas son etiquetadas con literales y los nodos
-internos con $TODOand$ y $TODOor$; ver ejemplo en la figura  1TODOLABELREF. Este
-fragmento es universal, lo que significa que para cada fórmula lógica hay una
-equivalente en formato NNF. Se dice que un NNF es descomponible (DNNF) XREF14W
-si para cada conjunción $TODOformula$, sus variables son disjuntos por pares; es
-decir, $TODO formula vars y vars$ para $i TODOnoigual j$. Un DNNF soporta un
-número de operaciones en tiempo polinomial en el tamaño de su GAD. Por ejemplo,
-podemos probar si un DNNF es satisfactible haciendo una simple pasada de abajo
-hacia arriba sobre su GAD en tiempo linear. Se dice que un DNNF es
-determinístico (d-DNNF) XREF14W si para cada disjunción $TODOformula$, los
-disjuntos sob lógicamente contradictorios por pares, es decir, $TODO formula
-false idistintoj$. El NNF en la figura 1TODOLABELREF, por ejemplo, es
-decomposableTODOemph y determinístico. Un d-DNNF soporta conteo de modelos en
-tiempo polinomial en el tamaño de su GAD, y enumeración de modelos en tiempo
-polinomial en el tamaño de la salida. Además, dada una función de puntuación de
-literales $r$, se puede computar la puntuación del mejor modelo en tiempo
-polinomial para DNNFs XREF17W.
-
-Los fragmentos DNNF y d-DNNF son universales pero traducir una teoría CNF a
-formrato DNNF tiene costo exponencial en el peor caso. Esta traducción se
-llama compilación en este campo. Hay un compilador públicamente disponible,
-llamado c2d TODOfootnote que realiza este proceso de compilación y que hace uso
-de técnicas de SAT modernas como backtracking dirigido por conflictos,
-aprendizaje de cláusulas y caching XREF15W. Este compilador entra en el peor
-caso en espacio exponencial en un parámetro llamado el ancho del árbol de
-descomposición que está relacionado a la TODO"conectividad" de la teoría CNF.
-Sin embargo, en nuestros experimentos, las teorías CNF que son compiladas son de
-ancho pequeño.
 

### 2010-01-12 tesis/propuesta/secciones/planteamiento.tex
parte de formulacion teorica

@@
-\section{Planteamiento del problema}
+\section{Planteamiento del Problema}
 
@@
 ancho pequeño.
+
+\subsection{Formalización}
+
+Consideramos cat[alogos de servicio de la forma $C=\tup{S,T}$ donde $S$ es un
+conjunto de predicados que representa servicios abstractos y $T=\{T_s\}_{s\in S}$ es una
+colección de tablas que representa el resultado de evaluar los servicios
+abstractos. En el contexto del WIP, un catálogo de servicio $C$ es una
+descripción idealizada de la salida producida por flujos abstractos
+implementados por servicios concretos descritos como vistas. Un flujo de trabajo
+$W$ sobre $S$ es una consulta conjuntiva de la forma
+
+\[ W(\vvec{x})\ \qrule\ s_1(\vvec{x}_1),\,s_1(\vvec{x}_2),\,\ldots,\,s_m(\vvec{x}_m)\,, \]
+
+donde $s_i\in S$, $\vvec{x}$ es un vector de variables y cada $\vvec{x}_i$ es un vector de variables y
+constantes. El resultado de $W$ sobre $C$, denotado como $W(C)$, es la tabla con
+ $|\vvec{x}|$ columnas que resulta de la proyección del \emph{join} relacional
+$\join\!\!\{T_{s_i}\}_{i=1}^n$ sobre $\vvec{x}$. Los átomos en el cuerpo de $W$ son llamados los
+(sub)objetivos de $W$, y las variables en la cabeza de $W$ son llamadas
+distinguidas.
+
+Un servicio concreto es descrito como una vista $V$ sobre $C$ que, siguiendo
+LAV, es una consulta sobre $S$. Dado un catálogo $C$, un flujo de trabajo $W$ y
+una colección de $n$ vistas $E=\tup{\{V_i\}_{i=1}^n,\{E_i\}_{i=1}^n}$, se pide
+conseguir todas las tuplas en
+$W(C)$ que puedan ser obtenidas de vistas en $E$. Es decir, se necesitan
+conseguir las \emph{instanciaciones}
+
+\[ R(\vvec{x})\ \qrule\ V_{i_1}(\vvec{x}_1),\,V_{i_2}(\vvec{x}_2),\,\ldots,\,V_{i_{m_i}}(\vvec{x}_{m_i}) \]
+
+tales que $R(E) \subseteq Q(C)$.
+Un problema de instanciación de flujo de trabajo es una
+tupla $\tup{S,W,\{V_i\}_i}$ donde $S$ es un conjunto de predicados
+que representa servicios
+abstractos, $W$ es un flujo de trabajo sobre $S$, y $\{V_i\}_i$ una colección de vistas
+que define servicios concretos. Suponemos trabajo con problemas \emph{seguros}
+en el sentido en que todas las variables mencionadas en la cabeza del flujo de
+trabajo (respectivamente en la cabeza de cada vista) aparecen en el cuerpo del
+flujo de trabajo (respectivamente en el cuerpo de cada vista). Además, sólo
+lidiamos con WIPs sin predicados aritméticos dentro del flujo de trabajo o
+vistas. Una instanciación $R$ es \emph{válida} si para todos los catálogos $C=\tup{S,T}$ y
+extensiones $\{E_i\}_i$, $R(E)\subseteq Q(C)$. Una colección $\R$ de instanciaciones válidas es una
+solución si para todos los catálogos de servicio $C=\tup{S,T}$ y extensiones $\{E_i\}_i$, no
+existe $\R'$ tal que $\R(E)\subset\R'(E)\subseteq Q(C)$.
+
+\subsubsection{Teorías Lógicas}
+
+Se usa una aproximación similar a la descrita en
+\cite{arvelo:aaai06} para codificar el WIP. 
+Se identifican instanciaciones del flujo de trabajo enumerando los modelos de
+una teoría lógica
+$\Delta=\Delta_{com}\cup\Delta_{id}^1\cup\cdots\Delta_{id}^N$
+donde $\Delta_{com}$ especifica cómo combinar $N$ copias independientes de la teoría
+$\Delta_{id}$ que cubren los objetivos en $W$.
+Cada $\Delta^i_{id}$ es una copia etiquetada $\Delta_{id}$ en la cual cada
+literal 
+$\ell$ está etiquetado como $\ell^i$.
+La teoría \emph{Instantiation Description} (ID) $\Delta_{id}$ consiste de
+diferentes grupos de cláusulas que garantizan que sus modelos están en
+correspondencia con instanciaciones parciales, mientras que la teoría
+$\Delta_{com}$ contiene cláusulas adicionales para garantizar una combinación
+correcta y completa de instanciaciones parciales para una instanciación
+completa.
+La teoría ID $\Delta_{id}$ consiste de las siguientes variables:
+
+\begin{enumerate}[--]
+\item $\{v_0,\ldots,v_n\}$ para indicar qué $V_i$ se usa, o $v_0$ para indicar el ID nulo.
+\item $\{g_1,\ldots,g_m\}$ para indicar objetivos cubiertos por la vista.
+\item $\{z_{j,k,i}\}$ para indicar que el objetivo $j$ en $W$ está cubierto por el objetivo $k$ en $V_i$.
+\item $\{t_{x,y}\}$ para indicar que la variable/constante $x$ en $W$ está asociada a la variable/constante $y$ en la vista.
+\end{enumerate}
+
+Los rangos de los índices para las variables $z$ y $t$ dependen del problema.
+
+Las siguientes cláusulas codifican el problema WIP en términos de los WIDs.
+Rajaraman et al.\ \cite{RajaramanSU95} mostraron que para consultas sin negación
+o comparaciones aritméticas, pero con constantes, y $m$ objetivos y $k$
+variables en la cabeza del flujo de trabajo, es suficiente considerar
+instanciaciones de longitud a lo más $N=m+k$.
+
+\begin{enumerate}[C10.]
+\item[C1.] (Al menos una vista es usada): $\bigvee_{i=0}^n v_i$.
+\item[C2.] (A lo más una vista es usada): $\neg v_i\lor\neg v_j$ para $i\neq j$.
+\item[C3.] (Vista nula implica nulidad): $v_0 \Rightarrow \neg g_j$ para $1\leq j\leq m$.
+\item[C4.] (Las vistas son útiles): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ para $1\leq i\leq n$.
+\item[C5.] (Objetivos cubiertos máximo una vez): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ para $i,j,k,l$ apropiados.
+\item[C6.] (Alcance de las vistas): $v_i \Rightarrow \neg g_j$ para objetivos que no pueden ser cubiertos por $V_i$.
+\item[C7.] (Consistencia): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ para $i,j,k$ apropiados.
+\item[C8.] (Variables muertas): $v_i \Rightarrow \neg t_{x,y}$ para todo $x,y$ con $y\notin V_i$.
+\item[C9.] (1-1 para $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ para todas las variables existenciales $y,y'\in V_i$.
+\item[C10.] (Distinguidos): $v_i \Rightarrow \neg t_{x,y}$ para $x$ distinguido y $y\in V_i$ existencial .
+\item[C11.] (Existencial): $v_i\land t_{x,y}\Rightarrow g_j$ para exist.\ $y\in V_i$ y objetivos $g_j$ que contengan $x$.
+\item[C12.] (Coincidencia): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ para todo
+$x,y$ que deban coincidir si $g_j$ es cubierto por el objetivo $k$ en $V_i$.
+\item[C13.] (Si todas las variables en $V_i$ son distinguidas, sólo cubre un objetivo): $v_i \land g_j \Rightarrow \neg g_k$ para vistas $v_i$ apropiadas.
+\end{enumerate}
+
+Estas son las mismas cláusulas usadas por \mcdsat\ para codificar QRPs \cite{arvelo:aaai06}.
+Para poder manejar símbolos constantes apropiadamente, las cláusulas deben ser
+mejoradas con:
+
+\begin{enumerate}[C10.]
+\item[C14.] (Inconsistencia directa 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
+\item[C15.] (Inconsistencia directa 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
+\item[C16.] (Inconsistencia directa 3): $\neg t_{A,B}$.
+\item[C17.] (Transitividad 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
+\item[C18.] (Transitividad 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
+\end{enumerate}
+
+El problema principal al manejar constantes es estar seguro de que no haya dos
+constantes distintas asociadas entre sí directa o indirectamente. Las cláusulas
+C14--C16 quitan asociaciones directamente inconsistentes, mientras que las
+últimas dos cláusulas implementan propagación de asociaciones restringida que
+quita inconsistencias indirectas.
+De la misma manera, la teoría $\Delta_{com}$ que especifica instanciaciones
+completas contiene las siguientes cláusulas:
+
+\begin{enumerate}[C10.]
+\item[C19.] (Cubrir todos los objetivos): $\bigvee_{j=1}^m g^i_j$ para $1\leq i\leq N$.
+\item[C20.] (Cobertura disjuntiva): $g^i_k \Rightarrow \neg g^j_k$ para $i\neq j$.
+\item[C21.] (Quitar simetrías): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
+            para $1\leq j\leq m$ y $1\leq i\leq N$.
+\item[C22.] (Inconsistencia directa 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
+\end{enumerate}
+
+Estas cláusulas proveen una caracterización correcta y completa de WIPs en el
+sentido en que sus modelos están en correspondencia con las instanciaciones del
+flujo de trabajo.
+
+\subsubsection{Parámetros QoS}
+
+Para los parámetros QoS, se supone un modelo de agregación aditiva simple en el
+que cada vista
+$V_i$ está asociada con un costo $c(V_i)$ (negativo si tratamos de utilidad),
+y una instanciación completa con la suma de los costos de sus vistas.
+Una instanciación óptima o mejor es una con costo mínimo, y el valor óptimo del
+WIP es el costo de una instanciación óptima.
+Un WIP siempre tiene un valor óptimo bien definido (si no hay instanciaciones,
+su costo es $\infty$), pero puede tener múltiples mejores instanciaciones.
+El WIP con costos consiste en encontrar todas las instanciaciones óptimas o una
+instanciación, esto depende de la aplicación particular.
+En la formulación propuesta, esto puede hacerse a partir del d-DNNF que codifica
+$\Delta$
+usando la función de puntuación de literales $r$ que asigna
+$r(\ell)=c(V_i)$ si
+$\ell=v_i$, y $r(\ell)=0$ if $\ell\notin\{v_1,\ldots,v_n\}$.

### 2010-01-12 tesis/propuesta/traduccion/traduccion.tex
parte de formulacion teorica

@@
 
-\section{Formalización}
-
-Consideramos cat[alogos de servicio de la forma $C = ST$ donde $S$ es un
-conjunto de predicados que representa servicios abstractos y $T=TTODO$ es una
-colección de tablas que representa el resultado de evaluar los servicios
-abstractos. En el contexto del WIP, un catálogo de servicio $C$ es una
-descripción idealizada de la salida producida por flujos abstractos
-implementados por servicios concretos descritos como vistas. Un flujo de trabajo
-$W$ sobre $S$ es una consulta conjuntiva de la forma
-
-$$
-WX
-$$
-
-donde $si,x$ es un vector de variables y cada $xi$ es un vector de variables y
-constantes. El resultado de $W$ sobre $C$, denotado como $W(C)$, es la tabla con
-$|x|$ columnas que resulta de la proyección del TODOemphjoin relacional
-$JOINTODO$ sobre $x$. Los átomos en el cuerpo de $W$ son llamados los
-(sub)objetivos de $W$, y las variables en la cabeza de $W$ son llamados
-distinguidos.
-
-Un servicio concreto es descrito como una vista $V$ sobre $C$ que, siguiendo
-LAV, es una consulta sobre $S$. Dado un catálogo $C$, un flujo de trabajo $W$ y
-una colección de $n$ vistas $E=$, se nos pide conseguir todas las tuplas en
-$W(C)$ que puedan ser obtenidas de vistas en $E$. Es decir, se necesitan
-conseguir las instanciaciones
-
-$$
-R(x)
-$$
-
-tales que $R(E)...$. Un problema de instanciación de flujo de trabajo es una
-tupla $tUTPLA$ donde $S$ es un conjunto de predicados que representa servicios
-abstractos, $W$ es un flujo de trabajo sobre $S$, y $VI$ una colección de vistas
-que define servicios concretos. Suponemos trabajo con problemas \emph{seguros}
-en el sentido en que todas las variables mencionadas en la cabeza del flujo de
-trabajo (respectivamente en la cabeza de cada vista) aparecen en el cuerpo del
-flujo de trabajo (respectivamente en el cuerpo de cada vista). Además, sólo
-lidiamos con WIPs sin predicados aritméticos dentro del flujo de trabajo o
-vistas. Una instanciación $R$ es válida si para todos los catálogos $C=$ y
-extensiones $EIII$, $R(E\#)$. Una colección $R$ de instanciaciones válidas es una
-solución si para todos los catálogos de servicio $C=$ y extensiones $EII$, no
-existe $R'$ tal que $R(E= C$.
-
-\subsection{Teorías Lógicas}
-
-Se usa una aproximación similar a la descrita en XREF4W para codificar el WIP.
-Se identifican instanciaciones del flujo de trabajo enumerando los modelos de
-una teoría lógica.
-
 \end{document}

### 2010-01-12 tesis/propuesta/secciones/metodologia.tex
parte de arquitectura del sistema integrada

--- 2010-01-12 tesis/propuesta/traduccion/traduccion.tex
@@
-\documentclass{article}
+\section{Metodología Propuesta}
 
-\usepackage[utf8x]{inputenc}
-\usepackage[spanish]{babel}
+\subsection{Arquitectura del Sistema}
 
-\begin{document}
-
-\tableofcontents
-
-\section{Arquitectura del sistema}
-
-TODOfig2 system arquitecthuere
+\begin{figure}[t]
+\centering
+\includegraphics[width=.9\textwidth]{graphics/architecture}
+\caption{Arquitectura del Sistema}
+\label{fig:architecture}
+\end{figure}
 
@@
 servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
-el Decodificador. La figura TODOREF2 muestra la arquitectura global del sistema.
-En estta infraestructura, una instancia del problema de instanciación de lfujos
+el Decodificador. La figura~\ref{fig:architecture} muestra la arquitectura global del sistema.
+En esta infraestructura, una instancia del problema de instanciación de flujos
 de trabajo se define como un flujo abstracto representado por una consulta
@@
 puede ser generada de manera semiautomática o automática usando herramientas
-tales como el sistema DEIMOS XREF3W.
+tales como el sistema DEIMOS \cite{AmbiteISWC09}.
 
@@
 corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
-compilador c2d, un componente TODOemphofftheshelf, compila la fórmula CNF a
+compilador c2d, un componente \emph{off-the-shelf}, compila la fórmula CNF a
 d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
@@
 instanciaciones con respecto a múltiples valores de los parámetros QoS.
-Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscardor
+Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscador
 a una instanciación de flujo de trabajo que resuelve el WIP.
@@
 todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
-una manera libre de TODOemphbacktracking todas las instanciaciones del flujo de
+una manera libre de \emph{backtracking} todas las instanciaciones del flujo de
 trabajo. Si el usuario está interesado en una mejor instanciación dados
-parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamñao
+parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamaño
 del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
-estas pueden ser computadas en tiempo lineal en su número. FInalmente, si el
+estas pueden ser computadas en tiempo lineal en su número. Finalmente, si el
 usuario está interesado en todas las instanciaciones, estas también pueden ser
@@
 por lo tanto no puede ser evitada.
-
-\end{document}

### 2010-01-12 tesis/propuesta/traduccion/traduccion.tex
parte de arquitectura del sistema integrada

@@
 
-\section{Arquitectura del sistema}
-
-TODOfig2 system arquitecthuere
-
-Se utiliza una arquitectura comprendida por un catálogo de descripciones de
-servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
-el Decodificador. La figura TODOREF2 muestra la arquitectura global del sistema.
-En estta infraestructura, una instancia del problema de instanciación de lfujos
-de trabajo se define como un flujo abstracto representado por una consulta
-conjuntiva sobre servicios abstractos que es dada como entrada junto con un
-conjunto de servicios concretos definidos por vistas de servicios abstractos.
-
-El catálogo se pobla con descripciones de servicios abstractos y concretos; cada
-servicio es descrito en términos de atributos de entrada y salida y anotado con
-un valor real que representa la utilidad QoS del servicio. La descripción de los
-servicios concretos, que son definidos como vistas de los servicios abstractos,
-puede ser generada de manera semiautomática o automática usando herramientas
-tales como el sistema DEIMOS XREF3W.
-
-Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
-corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
-compilador c2d, un componente TODOemphofftheshelf, compila la fórmula CNF a
-d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
-convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
-parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
-importante remarcar que el proceso de compilación necesita ser realizado sólo
-una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
-compilación resulta ser costosa en términos de tiempo, este costo puede ser
-amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
-instanciaciones con respecto a múltiples valores de los parámetros QoS.
-Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscardor
-a una instanciación de flujo de trabajo que resuelve el WIP.
-
-Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
-todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
-una manera libre de TODOemphbacktracking todas las instanciaciones del flujo de
-trabajo. Si el usuario está interesado en una mejor instanciación dados
-parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamñao
-del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
-estas pueden ser computadas en tiempo lineal en su número. FInalmente, si el
-usuario está interesado en todas las instanciaciones, estas también pueden ser
-computadas en tiempo lineal en su número. En los últimos dos casos, si ese
-número es exponencial (en el tamaño de la entrada), la enumeración de las
-instanciaciones también lo es pero esta complejidad es intrínsica al problema y
-por lo tanto no puede ser evitada.
-
 \end{document}

### 2010-01-12 tesis/propuesta/secciones/resultados.tex
seccion de resultados iniciales

--- 2010-01-12 tesis/propuesta/secciones/justificacion.tex
@@
-\section{Justificación e importancia}
+\section{Resultados Iniciales}
 
-\subsection{Ejemplo motivante}
+\subsection{Arquitectura del Sistema}
 
-% flights
-\newcommand{\flight}{\text{\it flight}}
-\newcommand{\UScity}{\text{\it uscity}}
-\newcommand{\national}{\text{\it national}}
-\newcommand{\oneway}{\text{\it one-way}}
-\newcommand{\onestop}{\text{\it one-stop}}
-\newcommand{\flightPA}{\text{\it flight-to-pa}}
-\newcommand{\onestopPA}{\text{\it onestop-to-pa}}
-\newcommand{\fromNY}{\text{\it from-ny}}
-\newcommand{\PA}{\text{PA}}
-\newcommand{\NY}{\text{NY}}
-\newcommand{\AL}{\text{AL}}
-% end flights
+\begin{figure}[t]
+\centering
+\includegraphics[width=.9\textwidth]{graphics/architecture}
+\caption{Arquitectura del Sistema}
+\label{fig:architecture}
+\end{figure}
 
-Considere un sistema simple de información de vuelos que contiene información
-sobre vuelos entre ciudades e información sobre qué ciudades están en los
-Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
-abstractos $\flight(x,y)$ y $\UScity(x)$. El primero relaciona dos ciudades $x$ e $y$ si
-hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
-E.E.U.U.  o no. Para los servicios concretos, supongamos que las fuentes de
-datos disponibles en Internet contienen la siguiente información:
+Se utiliza una arquitectura comprendida por un catálogo de descripciones de
+servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
+el Decodificador. La figura~\ref{fig:architecture} muestra la arquitectura global del sistema.
+En esta infraestructura, una instancia del problema de instanciación de flujos
+de trabajo se define como un flujo abstracto representado por una consulta
+conjuntiva sobre servicios abstractos que es dada como entrada junto con un
+conjunto de servicios concretos definidos por vistas de servicios abstractos.
 
-\begin{enumerate}[--]
-\item $\national(x,y)$ relaciona dos ciudades de E.E.U.U. conectadas por un vuelo directo,
-\item $\oneway(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,
-\item $\onestop(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,
-\item $\flightPA(x)$ dice si hay un vuelo directo desde $x$ a París,
-\item $\onestopPA(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a París con una escala en $y$, y
-\item $\fromNY(x)$ dice si hay un vuelo de Nueva York a $x$.
-\end{enumerate}
+El catálogo se pobla con descripciones de servicios abstractos y concretos; cada
+servicio es descrito en términos de atributos de entrada y salida y anotado con
+un valor real que representa la utilidad QoS del servicio. La descripción de los
+servicios concretos, que son definidos como vistas de los servicios abstractos,
+puede ser generada de manera semiautomática o automática usando herramientas
+tales como el sistema DEIMOS \cite{AmbiteISWC09}.
 
-Adicionalmente, los servicios concretos se describen con los siguientes
-servicios abstractos:
+Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
+corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
+compilador c2d, un componente \emph{off-the-shelf}, compila la fórmula CNF a
+d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
+convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
+parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
+importante remarcar que el proceso de compilación necesita ser realizado sólo
+una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
+compilación resulta ser costosa en términos de tiempo, este costo puede ser
+amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
+instanciaciones con respecto a múltiples valores de los parámetros QoS.
+Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscador
+a una instanciación de flujo de trabajo que resuelve el WIP.
 
-\begin{alignat*}{1}
-\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
-\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
-\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
-\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
-\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
-\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
-\end{alignat*}
-
-Ahora supongamos que un usuario está interesado en construir un flujo de trabajo
-capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
-en el mundo, tales que los vuelos puedan detenerse en cualquier ciudad.
-
-\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
-
-La siguiente consulta conjuntiva representa el flujo de trabajo que define esta
-solicitud en términos de servicios abstractos. El flujo de trabajo es definido
-de manera que todo lo relacionado sobre el ligamiento de parámetros de
-entrada/salida es resuelto. Así, cualquier instanciación de los servicios
-abstractos en términos de los servicios concretos es una implementación válida
-del flujo de trabajo. Las implementaciones corresponden a composiciones de
-servicios concretos en las cuales un servicio concreto puede implementar uno o
-más servicios abstractos del flujo, pero cada servicio abstracto puede ser
-implementado por exactamente un servicio concreto. Por ejemplo, la siguiente
-composición corresponde a una de estas implementaciones.
-
-\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
-
-Sin embargo, las siguientes dos composiciones no son válidas.
-
-\begin{alignat*}{1}
-I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
-I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
-\end{alignat*}
-
-La primera composición no es válida porque asocia la variable del flujo $y$ a
-las constantes \PA\ y \NY\ que denotan ciudades diferentes. En la otra mano, $I''$
-no lo implementa porque el servicio concreto $\onestop(x,y)$ no recibe como entrada
-o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces
-no es posible asegurarse de que esa ciudad esté asociada a la ciudad $w$ que es
-retornada por el flujo de trabajo.
-
-Estos ejemplos muestran que las instanciaciones correctas de un flujo deben
-manejar constantes de manera que dos constantes distintas no sean asociadas
-entre sí de manera directa o indirectamente vía transitividad, y que todos los
-atributos que aparezcan en un \emph{join} o en la salida sean producidos por los
-servicios concretos seleccionados.
-
-Los parámetros QoS son modelados anotando los servicios concretos con utilidades
-que caracterizan su comportamiento y que luego son agregados durante la
-instanciación. Así, como se mencionó previamente, la mejor instanciación es la
-que minimiza (o maximiza) la agregación de utilidades.
+Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
+todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
+una manera libre de \emph{backtracking} todas las instanciaciones del flujo de
+trabajo. Si el usuario está interesado en una mejor instanciación dados
+parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamaño
+del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
+estas pueden ser computadas en tiempo lineal en su número. Finalmente, si el
+usuario está interesado en todas las instanciaciones, estas también pueden ser
+computadas en tiempo lineal en su número. En los últimos dos casos, si ese
+número es exponencial (en el tamaño de la entrada), la enumeración de las
+instanciaciones también lo es pero esta complejidad es intrínsica al problema y
+por lo tanto no puede ser evitada.
 

### 2010-01-12 tesis/propuesta/secciones/metodologia.tex
metodologia comenzada

@@
 
-\subsection{Arquitectura del Sistema}
+La fase inicial de este trabajo consiste en realizar una investigación sobre el
+problema a ser tratado y el estado del arte en cuanto a las últimas
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=.9\textwidth]{graphics/architecture}
-\caption{Arquitectura del Sistema}
-\label{fig:architecture}
-\end{figure}
+Una vez conocidas las propuestas actuales se procede a identificar sus
+fortalezas y particularmente sus limitaciones. Se analizarán las causas de
+dichas limitaciones y se propondrán maneras de superarlas. Las propuestas deben
+estar acompañadas de un basamento formal que permita demostrar su correctitud y
+deben ser implementables en un lapso razonable de manera de evaluar su
+rendimiento y su capacidad de resolver el problema.
 
-Se utiliza una arquitectura comprendida por un catálogo de descripciones de
-servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
-el Decodificador. La figura~\ref{fig:architecture} muestra la arquitectura global del sistema.
-En esta infraestructura, una instancia del problema de instanciación de flujos
-de trabajo se define como un flujo abstracto representado por una consulta
-conjuntiva sobre servicios abstractos que es dada como entrada junto con un
-conjunto de servicios concretos definidos por vistas de servicios abstractos.
+Luego comenzará la implementación de la nueva solución
 
-El catálogo se pobla con descripciones de servicios abstractos y concretos; cada
-servicio es descrito en términos de atributos de entrada y salida y anotado con
-un valor real que representa la utilidad QoS del servicio. La descripción de los
-servicios concretos, que son definidos como vistas de los servicios abstractos,
-puede ser generada de manera semiautomática o automática usando herramientas
-tales como el sistema DEIMOS \cite{AmbiteISWC09}.
-
-Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
-corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
-compilador c2d, un componente \emph{off-the-shelf}, compila la fórmula CNF a
-d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
-convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
-parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
-importante remarcar que el proceso de compilación necesita ser realizado sólo
-una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
-compilación resulta ser costosa en términos de tiempo, este costo puede ser
-amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
-instanciaciones con respecto a múltiples valores de los parámetros QoS.
-Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscador
-a una instanciación de flujo de trabajo que resuelve el WIP.
-
-Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
-todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
-una manera libre de \emph{backtracking} todas las instanciaciones del flujo de
-trabajo. Si el usuario está interesado en una mejor instanciación dados
-parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamaño
-del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
-estas pueden ser computadas en tiempo lineal en su número. Finalmente, si el
-usuario está interesado en todas las instanciaciones, estas también pueden ser
-computadas en tiempo lineal en su número. En los últimos dos casos, si ese
-número es exponencial (en el tamaño de la entrada), la enumeración de las
-instanciaciones también lo es pero esta complejidad es intrínsica al problema y
-por lo tanto no puede ser evitada.
+La finalización del proceso de implementación da paso a la fase de evaluación de
+la propuesta.

### 2010-01-12 tesis/propuesta/secciones/metodologia.tex
cosas de la propuesta

@@
 La fase inicial de este trabajo consiste en realizar una investigación sobre el
-problema a ser tratado y el estado del arte en cuanto a las últimas
+problema a ser tratado. Se debe conocer el estado del arte en cuanto a resultados obtenidos
+relacionados al problema y las líneas de investigación que se están siguiendo.
 
 Una vez conocidas las propuestas actuales se procede a identificar sus
-fortalezas y particularmente sus limitaciones. Se analizarán las causas de
+fortalezas y especialmente sus limitaciones. Se analizarán las causas de
 dichas limitaciones y se propondrán maneras de superarlas. Las propuestas deben
@@
 
-Luego comenzará la implementación de la nueva solución
+Luego comenzará la implementación de la nueva solución. Para ello se podrá
+utilizar como punto de partida sistemas existentes y modificarlos para mejorar
+sus características siguiendo las ideas propuestas.
 
-La finalización del proceso de implementación da paso a la fase de evaluación de
-la propuesta.
+La finalización del proceso de implementación dará paso a la fase de evaluación de
+la propuesta. Se propone realizar series de experimentos diversos sobre el
+producto final que ilustren sus capacidades y rendimiento. En casos en
+que soluciones ya existentes tengan las mismas capacidades, se debe comparar el
+rendimiento de ambas sobre una variedad de casos de prueba que persigan revelar
+sus peores comportamientos. Esto permitirá determinar empíricamente las virtudes
+y desventajas de la solución propuesta.

### 2010-01-13 tesis/propuesta/secciones/justificacion.tex
parte de formulacion teorica
cosas de la propuesta
anadi tabla de lenzerini a justificacion

--- 2010-01-12 tesis/propuesta/secciones/justificacion.tex
@@
-\section{Justificación e importancia}
+\section{Justificación e Importancia}
 
-\subsection{Ejemplo motivante}
+\subsection{Ejemplo Motivante}
 
@@
 que minimiza (o maximiza) la agregación de utilidades.
+
+\subsection{Complejidad del Problema}
+
+Estamos tratando con un problema difícil \cite{lenzerini:dataintegration}. 
+El cuadro~\ref{cqcomplexity} ilustra la complejidad de distintos tipos de
+problemas de resolución de consultas basadas en vistas. El caso particular que
+se está tratando en este trabajo es CQ con búsqueda exacta, el cual está en
+coNP.
+
+\begin{table}
+\begin{center}
+  \begin{tabular}{|c||c|c|c|c|c|}
+\hline
+Sound  &  CQ  & CQ$\neq$  & PQ  &Datalog &FOL \\
+\hline
+  CQ  & PTIME &coNP  &PTIME &PTIME  &undec. \\
+ CQ$\neq$  & PTIME &coNP  &PTIME &PTIME  &undec. \\
+  PQ  &  coNP &coNP  & coNP & coNP  &undec. \\
+Datalog &coNP &undec.& coNP &undec. &undec. \\
+ FOL  & undec.&undec.&undec.&undec. &undec. \\
+\hline
+ Exact  & CQ &  CQ$\neq$     &  PQ  &Datalog &FOL \\
+\hline
+  CQ    &coNP &coNP  & coNP & coNP  &undec. \\
+ CQ$\neq$    &coNP &coNP  & coNP & coNP  &undec. \\
+  PQ    &coNP &coNP  & coNP & coNP  &undec. \\
+Datalog&undec.&undec.&undec.&undec. &undec. \\
+ FOL   &undec.&undec.&undec.&undec. &undec. \\
+\hline
+  \end{tabular}
+  \caption{Complejidad de responder consultas basadas en vistas (obtenido de
+\cite{lenzerini:dataintegration})}
+  \label{cqcomplexity}
+\end{center}
+\end{table}

### 2010-01-18 tesis/propuesta/secciones/cronograma.tex
cambio orden de asesores, y aproximacion por solucion

\section{Cronograma de Actividades}

\begin{tabular}{|p{8cm}|c|c|}
\hline
Actividad & Duración & Fecha estimada \\
\hline
Investigación sobre el estado del arte & 2 semanas & \\
Identificación de limitaciones de propuestas actuales & 2 semanas \\
Implementación de generación de cláusulas nuevas para constantes & 6 semanas & \\
Experimentos con generación de cláusulas nuevas & & \\
Implementación de búsqueda de mejor reescritura & & \\
Experimentos con búsqueda de mejor reescritura & & \\
Implementación de sistema de eliminación de cláusulas para uso de
\emph{Branch and Bound} & & \\
\hline
\end{tabular}

### 2010-01-18 tesis/propuesta/secciones/planteamiento.tex
cambio orden de asesores, y aproximacion por solucion

@@
 En este trabajo se considera una versión restringida de WIP que adopta la
-aproximación Local-As-View (LAV) \cite{levy:bucket}. En LAV, todos los elementos de un
+solución Local-As-View (LAV) \cite{levy:bucket}. En LAV, todos los elementos de un
 problema son especificados con un lenguaje común basado en servicios abstractos
@@
 de integración de datos \cite{Chen05,JaudoinPRST05}, y optimización de consultas y mantenimiento
-de datos \cite{AfratiLU07,levy:bucket}, y varias aproximaciones que escalan a un número grande de
+de datos \cite{AfratiLU07,levy:bucket}, y varias soluciones que escalan a un número grande de
 vistas han sido definidas \cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
 
-La aproximación recente de Arvelo et al. \cite{arvelo:aaai06} está basada en la enumeración
+La solución recente de Arvelo et al. \cite{arvelo:aaai06} está basada en la enumeración
 eficiente de modelos de una teoría lógica proposicional. Dado un QRP, una teoría
@@
 determinística y descomponible (d-DNNF) \cite{darwiche:d-dnnfs}. Así,
-la aproximación consiste en transformar (llamado compilar en el campo de
+la solución consiste en transformar (llamado compilar en el campo de
 compilación del conocimiento) la teoría lógica en el formato d-DNNF para
@@
 
-En esta sección se resumen las aproximaciones existentes que proveen soluciones
+En esta sección se resumen las soluciones existentes que proveen soluciones
 a los problemas de selección de servicios y reescritura de consultas y se
@@
 
-Ko et al. \cite{myoung08} proponen una aproximación basada en restriccciones que codifica
+Ko et al. \cite{myoung08} proponen una solución basada en restriccciones que codifica
 los valores permisibles no funcionales como un conjunto de restricciones cuya
@@
 
-Recientemente, dos aproximaciones basadas en planificación han sido propuestas.
+Recientemente, dos soluciones basadas en planificación han sido propuestas.
 Kuter y Golbeck \cite{kuterG09} extienden el algoritmo de planificación SHOP2 para
@@
 
-Se usa una aproximación similar a la descrita en
+Se usa una solución similar a la descrita en
 \cite{arvelo:aaai06} para codificar el WIP. 

### 2010-01-18 tesis/propuesta/secciones/portada.tex
cosas de la propuesta
cambio orden de asesores, y aproximacion por solucion

@@
 \begin{abstract}
-La Arquitectura Orientada a Servicios típica consiste de dos capas principales.
+Una Arquitectura Orientada a Servicios típica consiste de dos capas principales.
 La capa concreta con servicios concretos cuyas funcionalidades son descritas en
@@
 eficientemente el problema de instanciación que está firmemente basada en
-lógica. La infraestructura adopta la aproximación Local-As-View en la cual la
+lógica. La infraestructura adopta la solución Local-As-View en la cual la
 funcionalidad de servicios concretos se describe usando vistas de servicios
@@
 formato d-DNNF, se provee una solución eficiente y escalable al problema de
-instanciación de flujos de trabajo. Esta aproximación no sólo escala a
+instanciación de flujos de trabajo. Esta solución no sólo escala a
 instancias grandes como muestran resultados experimentales ya realizados, sino
@@
 \begin{flushright}
-\emph{Asesor:} \\
-Prof. María Esther \textsc{Vidal}
+\emph{Asesores:} \\
+Prof. Blai \textsc{Bonet} \\
+Prof. María Esther \textsc{Vidal} \\
 \end{flushright}

### 2010-01-18 tesis/propuesta/secciones/resultados.tex
cambio orden de asesores, y aproximacion por solucion

@@
 El objetivo de los experimentos es determinar el rendimiento de la propuesta con
-condiciones variantes. El principal beneficio de la aproximación propuesta es
+condiciones variantes. El principal beneficio de la solución propuesta es
 que se puede compilar la teoría lógica para una instancia del problema y luego
@@
 necesidad de recompilar la teoría lógica. Por lo tanto, la complejidad en tiempo
-de la aproximación es básicamente el tiempo para codificar el WIP como un
+de la solución es básicamente el tiempo para codificar el WIP como un
 CNF más el tiempo para compilar el CNF en un d-DNNF y el tiempo para decodificar
@@
 Estos son experimentos preliminares, pero los resultados muestran que la
-aproximación propuesta escala eficientemente para problemas con varios objetivos
+solución propuesta escala eficientemente para problemas con varios objetivos
 y vistas. Se cree que estos resultados son alentadores y motivan a continuar

### 2010-01-23 tesis/src/eswc_demo/main.py
eswc_demo

@@ -1,0 +1,15 @@
+import sys
+from PyQt4 import QtCore, QtGui
+from ui.main import Ui_Main
+
+class MyForm(QtGui.QMainWindow):
+    def __init__(self, parent=None):
+        QtGui.QWidget.__init__(self, parent)
+        self.ui = Ui_Main()
+        self.ui.setupUi(self)
+
+if __name__ == "__main__":
+    app = QtGui.QApplication(sys.argv)
+    myapp = MyForm()
+    myapp.show()
+    sys.exit(app.exec_())

### 2010-01-23 tesis/src/eswc_demo/main.py
makefile, y prueba.py hecha a mano

@@ -1,15 +1,20 @@
 import sys
-from PyQt4 import QtCore, QtGui
-from ui.main import Ui_Main
 
-class MyForm(QtGui.QMainWindow):
+from PyQt4.QtGui import *
+
+from ui.prueba import Ui_Main
+
+class MyForm(QMainWindow):
     def __init__(self, parent=None):
-        QtGui.QWidget.__init__(self, parent)
+        QWidget.__init__(self, parent)
         self.ui = Ui_Main()
         self.ui.setupUi(self)
 
+        #queryDisplay = self.ui.CentralWid.findChild(QLineEdit, "queryDisplay")
+        #queryDisplay.setText("pepe")
+
 if __name__ == "__main__":
-    app = QtGui.QApplication(sys.argv)
+    app = QApplication(sys.argv)
     myapp = MyForm()
     myapp.show()
     sys.exit(app.exec_())

### 2010-01-23 tesis/src/eswc_demo/ui/prueba.py

@@ -1,0 +1,73 @@
+# -*- coding: utf-8 -*-
+
+from PyQt4 import QtCore
+
+from PyQt4.QtGui import *
+
+class Ui_Main(object):
+    def setupUi(self, Main):
+        Main.setObjectName("Main")
+        Main.resize(602, 402)
+
+        #CentralWid
+        self.CentralWid = QWidget(Main)
+        self.CentralWid.setObjectName("CentralWid")
+
+        # top-level horizontal splitter
+        self.splitter_2 = QSplitter(self.CentralWid)
+        self.splitter_2.setOrientation(QtCore.Qt.Horizontal)
+        self.splitter_2.setObjectName("splitter_2")
+
+        self.listView = QListView(self.splitter_2)
+#        sizePolicy = QSizePolicy(QSizePolicy.Expanding, QSizePolicy.Expanding)
+#        sizePolicy.setHorizontalStretch(0)
+#        sizePolicy.setVerticalStretch(0)
+#        sizePolicy.setHeightForWidth(self.listView.sizePolicy().hasHeightForWidth())
+#        self.listView.setSizePolicy(sizePolicy)
+        self.listView.setObjectName("listView")
+
+        # right side vertical splitter
+        self.splitter = QSplitter(self.splitter_2)
+        self.splitter.setOrientation(QtCore.Qt.Vertical)
+        self.splitter.setObjectName("splitter")
+
+        # main :: main right side :: instance info
+        self.instanceInfo_gridLayoutWidget = QWidget(self.splitter)
+        self.instanceInfo_gridLayout = QGridLayout(self.instanceInfo_gridLayoutWidget)
+
+        # main :: main right side :: instance info :: query display
+        self.queryDisplay = QLineEdit(self.instanceInfo_gridLayoutWidget)
+        self.queryDisplay.setReadOnly(True)
+        self.queryDisplay.setObjectName("queryDisplay")
+        self.instanceInfo_gridLayout.addWidget(self.queryDisplay)
+
+#        self.horizontalLayoutWidget = QWidget(self.splitter)
+#        self.horizontalLayoutWidget.setObjectName("horizontalLayoutWidget")
+#        self.horizontalLayout = QHBoxLayout(self.horizontalLayoutWidget)
+#        self.horizontalLayout.setObjectName("horizontalLayout")
+#        self.label = QLabel(self.horizontalLayoutWidget)
[…]

### 2010-01-24 tesis/src/eswc_demo/main.py
avances x

@@ -10,8 +10,17 @@ class MyForm(QMainWindow):
         self.ui = Ui_Main()
         self.ui.setupUi(self)
 
-        #queryDisplay = self.ui.CentralWid.findChild(QLineEdit, "queryDisplay")
-        #queryDisplay.setText("pepe")
+        queryDisplay = self.ui.CentralWidget.findChild(QLineEdit, "queryDisplayBox")
+        queryDisplay.setText("q(X1,X2) :- r(X1,X3),r(X3,X4),r(X4,X5),r(X5,X2)")
+
+        listView = self.ui.CentralWidget.findChild(QListWidget, "listView")
+
+        if listView == None:
+            print "sno"
+
+        listView.insertItem(0, QListWidgetItem("60 views", listView))
+        listView.insertItem(0, QListWidgetItem("80 views", listView))
+        listView.insertItem(0, QListWidgetItem("100 views", listView))
 
 if __name__ == "__main__":
     app = QApplication(sys.argv)

### 2010-01-24 tesis/src/eswc_demo/ui/prueba.py

@@ -9,16 +9,15 @@ class Ui_Main(object):
         Main.setObjectName("Main")
         Main.resize(602, 402)
 
-        #CentralWid
-        self.CentralWid = QWidget(Main)
-        self.CentralWid.setObjectName("CentralWid")
+        # CentralWidget
+        self.CentralWidget = QWidget(Main)
 
-        # top-level horizontal splitter
-        self.splitter_2 = QSplitter(self.CentralWid)
-        self.splitter_2.setOrientation(QtCore.Qt.Horizontal)
-        self.splitter_2.setObjectName("splitter_2")
+        # top
+        self.topSplitter = QSplitter(self.CentralWidget)
+        self.topSplitter.setOrientation(QtCore.Qt.Horizontal)
 
-        self.listView = QListView(self.splitter_2)
+        # top :: left side
+        self.listView = QListWidget(self.topSplitter)
 #        sizePolicy = QSizePolicy(QSizePolicy.Expanding, QSizePolicy.Expanding)
 #        sizePolicy.setHorizontalStretch(0)
 #        sizePolicy.setVerticalStretch(0)
@@ -26,22 +25,86 @@ class Ui_Main(object):
 #        self.listView.setSizePolicy(sizePolicy)
         self.listView.setObjectName("listView")
 
-        # right side vertical splitter
-        self.splitter = QSplitter(self.splitter_2)
-        self.splitter.setOrientation(QtCore.Qt.Vertical)
-        self.splitter.setObjectName("splitter")
+        # top :: right side
+        self.rightSplitter = QSplitter(self.topSplitter)
+        self.rightSplitter.setOrientation(QtCore.Qt.Vertical)
 
-        # main :: main right side :: instance info
-        self.instanceInfo_gridLayoutWidget = QWidget(self.splitter)
-        self.instanceInfo_gridLayout = QGridLayout(self.instanceInfo_gridLayoutWidget)
+        # top :: right side :: instance info
+        self.instanceInfo_LayoutWidget = QWidget(self.rightSplitter)
+        self.instanceInfo_Layout = QGridLayout(self.instanceInfo_LayoutWidget)
 
-        # main :: main right side :: instance info :: query display
-        self.queryDisplay = QLineEdit(self.instanceInfo_gridLayoutWidget)
-        self.queryDisplay.setReadOnly(True)
[…]

### 2010-01-24 tesis/src/eswc_demo/main.py
listviews con datos

@@ -22,6 +22,16 @@ class MyForm(QMainWindow):
         listView.insertItem(0, QListWidgetItem("80 views", listView))
         listView.insertItem(0, QListWidgetItem("100 views", listView))
 
+        #parte de archivo TODO
+        viewsDisplayBox = self.ui.CentralWidget.findChild(QListWidget, "viewsDisplayBox")
+
+        f = open("views-60-5.txt")
+
+        for line in f:
+            viewsDisplayBox.insertItem(0, QListWidgetItem(line))
+
+        f.close()
+
 if __name__ == "__main__":
     app = QApplication(sys.argv)
     myapp = MyForm()

### 2010-01-24 tesis/src/eswc_demo/ui/prueba.py

@@ -46,7 +46,7 @@ class Ui_Main(object):
         self.viewsDisplayLabel = QLabel(self.instanceInfo_LayoutWidget)
         self.instanceInfo_Layout.addWidget(self.viewsDisplayLabel, 1, 0)
 
-        self.viewsDisplayBox = QListView(self.instanceInfo_LayoutWidget)
+        self.viewsDisplayBox = QListWidget(self.instanceInfo_LayoutWidget)
         self.viewsDisplayBox.setObjectName("viewsDisplayBox")
         self.instanceInfo_Layout.addWidget(self.viewsDisplayBox, 1, 1)
 
@@ -84,7 +84,7 @@ class Ui_Main(object):
         self.dnnf_ButtonsLayout = QHBoxLayout()
         self.dnnf_ButtonsWidget.setLayout(self.dnnf_ButtonsLayout)
 
-        # top :: right side :: solutions :: dnnf :: dnnf buttons layout :: dnnf genandload button
+        # top :: right side :: solutions :: dnnf :: dnnf buttons layout :: dnnf buttons
         self.dnnfGenAndLoadButton = QPushButton("Generate && Load")
         self.dnnf_ButtonsLayout.addWidget(self.dnnfGenAndLoadButton)
 
@@ -99,10 +99,22 @@ class Ui_Main(object):
         self.rewritings_Layout = QVBoxLayout()
         self.rewritings_Widget.setLayout(self.rewritings_Layout)
 
-        # top :: right side :: solutions :: rewritings :: irx label
-        self.irxLabel = QLabel(self.dnnf_Widget)
-        self.rewritings_Layout.addWidget(self.irxLabel)
+        # top :: right side :: solutions :: rewritings :: rewritings buttons layout
+        self.rewritings_ButtonsWidget = QWidget()
+        self.rewritings_Layout.addWidget(self.rewritings_ButtonsWidget)
 
+        self.rewritings_ButtonsLayout = QHBoxLayout()
+        self.rewritings_ButtonsWidget.setLayout(self.rewritings_ButtonsLayout)
+
+        # top :: right side :: solutions :: rewritings :: rewritings buttons layout :: rewritings buttons
+        self.rewritingsGetAllButton = QPushButton("Get All")
+        self.rewritings_ButtonsLayout.addWidget(self.rewritingsGetAllButton)
+
+        self.rewritingsGetBestButton = QPushButton("Get Best")
[…]

### 2010-01-24 tesis/src/eswc_demo/main.py
"concrete services" alineado arriba, servicios en orden correcto

@@ -26,9 +26,13 @@ class MyForm(QMainWindow):
         viewsDisplayBox = self.ui.CentralWidget.findChild(QListWidget, "viewsDisplayBox")
 
         f = open("views-60-5.txt")
+        lines = [l for l in f]
+        f.close()
 
-        for line in f:
-            viewsDisplayBox.insertItem(0, QListWidgetItem(line))
+        lines.reverse()
+
+        for s in lines:
+            viewsDisplayBox.insertItem(0, QListWidgetItem(s.strip()))
 
         f.close()
 

### 2010-01-24 tesis/src/eswc_demo/ui/prueba.py

@@ -10,7 +10,11 @@ class Ui_Main(object):
         Main.resize(602, 402)
 
         # CentralWidget
-        self.CentralWidget = QWidget(Main)
+        self.CentralWidget = QFrame(Main)
+
+        # menubar
+        menubar = Main.menuBar()
+        file = menubar.addMenu('&File')
 
         # top
         self.topSplitter = QSplitter(self.CentralWidget)
@@ -44,7 +48,7 @@ class Ui_Main(object):
 
         # top :: right side :: instance info :: views display label and box
         self.viewsDisplayLabel = QLabel(self.instanceInfo_LayoutWidget)
-        self.instanceInfo_Layout.addWidget(self.viewsDisplayLabel, 1, 0)
+        self.instanceInfo_Layout.addWidget(self.viewsDisplayLabel, 1, 0, QtCore.Qt.AlignTop)
 
         self.viewsDisplayBox = QListWidget(self.instanceInfo_LayoutWidget)
         self.viewsDisplayBox.setObjectName("viewsDisplayBox")
@@ -90,6 +94,9 @@ class Ui_Main(object):
 
         self.dnnfLoadFromFileButton = QPushButton("Load From File")
         self.dnnf_ButtonsLayout.addWidget(self.dnnfLoadFromFileButton)
+
+        self.dnnfUnloadButton = QPushButton("Unload")
+        self.dnnf_ButtonsLayout.addWidget(self.dnnfUnloadButton)
 
         # top :: right side :: solutions :: rewritings
         self.rewritings_Widget = QGroupBox()

### 2010-01-25 tesis/propuesta/secciones/portada.tex
portada lista

@@
 \begin{abstract}
-Una Arquitectura Orientada a Servicios típica consiste de dos capas principales.
+Una arquitectura orientada a servicios típica se separa en dos capas --grupos de
+componentes-- principales.
 La capa concreta con servicios concretos cuyas funcionalidades son descritas en
@@
 flujos de trabajo abstractos y propiedades no funcionales en términos de
-restricciones sobre QoS. Durante la ejecución de un flujo de trabajo, los
+restricciones sobre QoS. Durante la ejecución de un flujo de trabajo o \emph{workflow}, los
 servicios abstractos son instanciados en servicios concretos que cumplen con los
@@
 combinatorio. En este trabajo se propone una infraestructura para resolver
-eficientemente el problema de instanciación que está firmemente basada en
-lógica. La infraestructura adopta la solución Local-As-View en la cual la
+eficientemente el problema de instanciación que está basada en
+la lógica proposicional. La infraestructura adopta el enfoque de definición denominado Local-As-View en la cual la
 funcionalidad de servicios concretos se describe usando vistas de servicios

### 2010-01-25 tesis/propuesta/secciones/justificacion.tex
gran parte de las correcciones de ma esther

@@
 % flights
-\newcommand{\flight}{\text{\it flight}}
-\newcommand{\UScity}{\text{\it uscity}}
-\newcommand{\national}{\text{\it national}}
-\newcommand{\oneway}{\text{\it one-way}}
-\newcommand{\onestop}{\text{\it one-stop}}
-\newcommand{\flightPA}{\text{\it flight-to-pa}}
-\newcommand{\onestopPA}{\text{\it onestop-to-pa}}
-\newcommand{\fromNY}{\text{\it from-ny}}
-\newcommand{\PA}{\text{PA}}
-\newcommand{\NY}{\text{NY}}
+\newcommand{\vuelo}{\text{\it vuelo}}
+\newcommand{\ciudadVE}{\text{\it ciudadVE}}
+\newcommand{\nacional}{\text{\it nacional}}
+\newcommand{\ida}{\text{\it ida}}
+\newcommand{\unaescala}{\text{\it una-escala}}
+\newcommand{\vueloCCS}{\text{\it vuelo-hacia-ccs}}
+\newcommand{\unaparadaCCS}{\text{\it unaescala-hacia-ccs}}
+\newcommand{\desdeVAL}{\text{\it desde-val}}
+\newcommand{\PA}{\text{CCS}}
+\newcommand{\NY}{\text{VAL}}
 \newcommand{\AL}{\text{AL}}
@@
 Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
-abstractos $\flight(x,y)$ y $\UScity(x)$. El primero relaciona dos ciudades $x$ e $y$ si
+abstractos $\vuelo(x,y)$ y $\ciudadVE(x)$. El primero relaciona dos ciudades $x$ e $y$ si
 hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
-E.E.U.U.  o no. Para los servicios concretos, supongamos que las fuentes de
+Venezuela  o no. Para los servicios concretos, supongamos que las fuentes de
 datos disponibles en Internet contienen la siguiente información:
@@
 \begin{enumerate}[--]
-\item $\national(x,y)$ relaciona dos ciudades de E.E.U.U. conectadas por un vuelo directo,
-\item $\oneway(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,
-\item $\onestop(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,
-\item $\flightPA(x)$ dice si hay un vuelo directo desde $x$ a París,
-\item $\onestopPA(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a París con una escala en $y$, y
-\item $\fromNY(x)$ dice si hay un vuelo de Nueva York a $x$.
+\item $\nacional(x,y)$ relaciona dos ciudades de Venezuela conectadas por un vuelo directo,
+\item $\ida(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,
+\item $\unaescala(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,
+\item $\vueloCCS(x)$ dice si hay un vuelo directo desde $x$ a Caracas,
+\item $\unaparadaCCS(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a Caracas con una escala en $y$, y
+\item $\desdeVAL(x)$ dice si hay un vuelo de Valencia a $x$.
 \end{enumerate}
@@
 \begin{alignat*}{1}
-\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
-\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
-\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
-\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
-\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
-\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
+\nacional(x,y)\   &\qrule\ \vuelo(x,y),\,\ciudadVE(x),\,\ciudadVE(y)\,. \\
+\ida(x,y)\     &\qrule\ \vuelo(x,y)\,. \\
+\unaescala(x,z)\    &\qrule\ \vuelo(x,y),\,\vuelo(y,z)\,. \\
+\vueloCCS(x)\     &\qrule\ \vuelo(x,\PA)\,. \\
+\unaparadaCCS(x,y)\  &\qrule\ \vuelo(x,y),\,\vuelo(y,\PA)\,. \\
+\desdeVAL(x)\       &\qrule\ \vuelo(\NY,x)\,.
 \end{alignat*}
 
-Ahora supongamos que un usuario está interesado en construir un flujo de trabajo
+Ahora supongamos que un usuario está interesado en construir una aplicación
 capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
@@
 
-\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
+\[ W(x,w,y,z) \qrule \ciudadVE(x),\,\vuelo(x,w),\,\vuelo(w,y),\,\vuelo(y,z),\,\vuelo(z,x)\,. \]
 
@@
 
-\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
+\[ I(x,w,y,z)\ \qrule\ \nacional(x,w),\,\vueloCCS(w),\,\ida(\PA,z),\,\nacional(z,x)\,. \]
 
@@
 \begin{alignat*}{1}
-I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
-I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
+I'(x,w,y,z)\  \qrule\ &\nacional(x,y),\,\vueloCCS(y),\,\desdeVAL(z),\,\nacional(z,x)\,. \\
+I''(x,w,y,z)\ \qrule\ &\unaescala(x,y),\,\ida(y,z),\,\nacional(z,x)\,.
 \end{alignat*}
@@
 las constantes \PA\ y \NY\ que denotan ciudades diferentes. En la otra mano, $I''$
-no lo implementa porque el servicio concreto $\onestop(x,y)$ no recibe como entrada
+no lo implementa porque el servicio concreto $\unaescala(x,y)$ no recibe como entrada
 o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces

### 2010-01-25 tesis/propuesta/secciones/planteamiento.tex
gran parte de las correcciones de ma esther

@@
 
-Bajo la sombra de la Web Semántica, y con el soporte de las Arquitecturas Orientadas a
+En el contexto de la Web Semántica, y con el soporte de las Arquitecturas Orientadas a
 Servicios (SOA), el número de fuentes de datos y servicios Web ha explotado en
@@
 Gracias a
-este tesoro de datos, la tendencia de los usuarios es hacia depender cada vez
+esta variedad de datos, la tendencia de los usuarios es hacia depender cada vez
 más de métodos automáticos para manejarlos tales como recuperación de datos de
@@
 servicios concretos cuyas funcionalidades son descritas en términos de pre- y
-post-condiciones y propiedades no funcionales en términos de parámetros de
+post-condiciones y propiedades no funcionales en conjuntos de parámetros de
 calidad de servicio (QoS), y
@@
 
-En este trabajo se considera una versión restringida de WIP que adopta la
-solución Local-As-View (LAV) \cite{levy:bucket}. En LAV, todos los elementos de un
+En este trabajo se considera una versión restringida de WIP que adopta el
+enfoque Local-As-View (LAV) \cite{levy:bucket}. En LAV, todos los elementos de un
 problema son especificados con un lenguaje común basado en servicios abstractos
@@
 
-La solución recente de Arvelo et al. \cite{arvelo:aaai06} está basada en la enumeración
-eficiente de modelos de una teoría lógica proposicional. Dado un QRP, una teoría
+La solución propuesta por Arvelo et al. \cite{arvelo:aaai06} está basada en la enumeración
+eficiente de modelos de una teoría lógica proposicional. Dada una consulta conjuntiva, una teoría
 lógica es construída de manera que cada modelo de la teoría codifica una
@@
 modelos, sino también otras operaciones. Entre ellas se encuentra la enumeración
-de los modelos con mejor puntuación. Dada una función de puntuación de literales
-$r(\ell)$ que asigne puntuaciones a cada literal $\ell$, se define la puntuación
-$r(\omega)$ de un modelo $\omega$ como la suma de las puntuaciones de los literales
+de los modelos con mejor calidad. Dada una función de calidad de literales
+$r(\ell)$ que asigne calidades a cada literal $\ell$, se define la calidad
+$r(\omega)$ de un modelo $\omega$ como la suma de las calidades de los literales
 activados por $\omega$ (es decir,
 $r(\omega)\doteq\sum_{\omega\vDash\ell}r(\ell)$), y se dice que $\omega$ es un modelo
-mejor puntuado si no hay modelo $\omega'$ tal que $r(\omega')<r(\omega)$.
+de mejor calidad si no hay modelo $\omega'$ tal que $r(\omega')<r(\omega)$.
 
-Dada una teoría en d-DNNF, se computa la puntuación de los mejores modelos, y
+Dada una teoría en d-DNNF, se computa la calidad de los mejores modelos, y
 los mejores modelos, en tiempo lineal en el tamaño del d-DNNF. Este cómputo
 transforma el GAD del d-DNNF en un circuito aritmético reemplazando los nodos
-AND con `+' y los nodos OR con `min'. La función de puntuación de literales
-asigna valores a las hojas del circuito que son propagados a la raíz en tiempo
-lineal. El valor de la raíz es la puntuación del mejor modelo \cite{darwiche:weighted}.
+AND con `+' y los nodos OR con `min'. La función de calidad de literales
+asigna valores a las hojas del circuito aritmético (grafo acíclico dirigido usado
+para calcular polinomios) que son propagados a la raíz en tiempo
+lineal. El valor de la raíz es la calidad del mejor modelo \cite{darwiche:weighted}.
 
@@
 búsqueda combinatoria se reduce a la computación de un mejor modelo de una
-teoría lógica que puede ser realizada eficientemente una vez que la teoría sea
+teoría lógica que puede ser realizada eficientemente una vez que la teoría es
 transformada al formato d-DNNF.
@@
 
-En esta sección se resumen las soluciones existentes que proveen soluciones
+En esta sección se resumen los enfoques existentes que proveen soluciones
 a los problemas de selección de servicios y reescritura de consultas y se
@@
 problema de selección o composición de servicios consciente de QoS, que ha sido
-demostrado ser NP-difícil \cite{Hiroshi2008}. Este es un problema de optimización
+demostrado ser \emph{NP-hard} \cite{Hiroshi2008}. Este es un problema de optimización
 combinatoria y varias heurísticas han sido propuestas para conseguir una
-solución relativamente buena en un período razonablemente corto de tiempo.
+solución relativamente buena en un período razonable de tiempo.
 
-Rahmani et al. \cite{rahmani08} presentan una heurística basada en métrica de distancia
-que guía un algoritmo de búsqueda hacia atrás; esta métrica induce un orden de
-los servicios en una manera en que los nodos vertedero tienen baja probabilidad
+Rahmani et al. \cite{rahmani08} presentan una heurística basada en una métrica de distancia
+que guía un algoritmo de búsqueda; esta métrica induce un orden de
+los servicios en una manera en que los nodos vertedero (sin salida) tienen baja probabilidad
 de ser visitados. En una serie de artículos, Berardi y otros \cite{berardi05,berardi08,berardi06}
@@
 servicios y flujos de trabajo en términos de máquinas de estados finitos
-determinísticas que son codificadas usando teorías de Lógica de Descripción
+determinísticas que son codificadas usando teorías de Lógicas Descriptivas
 cuyos modelos corresponden a soluciones del problema. Aunque se podrían explotar
-métodos para formalismos de Lógicas de Descripción, no han sido reportados la
+métodos para formalismos de Lógicas Descriptivas, no han sido reportados la
 escalabilidad o el rendimiento de la solución propuesta.
 
-Ko et al. \cite{myoung08} proponen una solución basada en restriccciones que codifica
+Ko et al. \cite{myoung08} proponen un enfoque basado en restriccciones que codifica
 los valores permisibles no funcionales como un conjunto de restricciones cuya
@@
 número grande de servicios y procesos abstractos. Cardellini et al. \cite{cardellini07}
-codifican una parte del problema de composición de servicios consciente de QoS
+codifican una parte del problema de composición de servicios caracterizados por QoS
 como un problema de programación lineal \cite{cardellini07}. Por otra parte, Wada et al.
@@
 algoritmos genéticos que identifica un conjunto de composiciones de servicios
-no dominadas que mejor encajan en los requerimientos QoS.
+no dominadas que mejor satisfacen los requerimientos QoS.
 
@@
 Un número de algoritmos han sido desarrollados para encontrar las reescrituras
-de una consulta dada; los más prominentes son el algoritmo bucket \cite{levy:bucket}, el
+de una consulta dada; los más prominentes son el algoritmo \emph{bucket} \cite{levy:bucket}, el
 algoritmo de reglas inversas \cite{pods:DuschkaG97,Qian96}, el algoritmo
@@
 
-El algoritmo bucket reduce el número de posibilidades considerando sólo cada
+El algoritmo denominado \emph{bucket} reduce el número de posibilidades considerando sólo cada
 subobjetivo de la consulta aislado, y seleccionando las vistas que son capaces
@@
 propiedades de los DNNF determinísticos (d-DNNF) \cite{darwiche:d-dnnfs} para proveer una
-solución eficiente y escalable al problema de selección de serviccios.
+solución eficiente y escalable al problema de selección de servicios.
 
@@
 acíclico dirigido en que las hojas son etiquetadas con literales y los nodos
-internos con $\land$ y $\lor$; ver ejemplo en la figura~\ref{fig:dnnf}. Este
-fragmento es universal, lo que significa que para cada fórmula lógica hay una
+internos con $\land$ y $\lor$; ver ejemplo en la figura~\ref{fig:dnnf}. Esta
+forma normal es universal, lo que significa que para cada fórmula lógica hay una
 equivalente en formato NNF. Se dice que un NNF es descomponible (DNNF)
@@
 número de operaciones en tiempo polinomial en el tamaño de su GAD. Por ejemplo,
-podemos probar si un DNNF es satisfactible haciendo una simple pasada de abajo
-hacia arriba sobre su GAD en tiempo linear. Se dice que un DNNF es
+podemos probar si un DNNF es satisfactible haciendo un sólo recorrido 
+sobre su GAD en tiempo linear. Se dice que un DNNF es
 determinístico (d-DNNF) \cite{darwiche:d-dnnfs} si para cada disjunción $\bigvee_i\phi_i$, los
@@
 tiempo polinomial en el tamaño de su GAD, y enumeración de modelos en tiempo
-polinomial en el tamaño de la salida. Además, dada una función de puntuación de
-literales $r$, se puede computar la puntuación del mejor modelo en tiempo
+polinomial en el tamaño de la salida. Además, dada una función de calidad de
+literales $r$, se puede computar la calidad del mejor modelo en tiempo
 polinomial para DNNFs \cite{darwiche:weighted}.
 
-Los fragmentos DNNF y d-DNNF son universales pero traducir una teoría CNF a
+Toda teoría puede representarse en las formas DNNF y d-DNNF pero traducir una teoría CNF a
 formato DNNF tiene costo exponencial en el peor caso. Esta traducción se
@@
 de técnicas de SAT modernas como backtracking dirigido por conflictos,
-aprendizaje de cláusulas y caching \cite{darwiche:compiler}. Este compilador entra en el peor
-caso en espacio exponencial en un parámetro llamado el ancho del árbol de
+aprendizaje de cláusulas y caching \cite{darwiche:compiler}. Este compilador incurre en el peor
+caso (peor gasto) en espacio exponencial en un parámetro llamado el ancho del árbol de
 descomposición que está relacionado a la ``conectividad'' de la teoría CNF.
@@
 
-Consideramos cat[alogos de servicio de la forma $C=\tup{S,T}$ donde $S$ es un
+Consideramos catálogos de servicio de la forma $C=\tup{S,T}$ donde $S$ es un
 conjunto de predicados que representa servicios abstractos y $T=\{T_s\}_{s\in S}$ es una
@@
 abstractos. En el contexto del WIP, un catálogo de servicio $C$ es una
-descripción idealizada de la salida producida por flujos abstractos
+descripción virtual de la salida producida por flujos abstractos
 implementados por servicios concretos descritos como vistas. Un flujo de trabajo
@@
 $\Delta$
-usando la función de puntuación de literales $r$ que asigna
+usando la función de calidad de literales $r$ que asigna
 $r(\ell)=c(V_i)$ si

### 2010-01-25 tesis/propuesta/secciones/portada.tex
gran parte de las correcciones de ma esther

@@
 instancias grandes como muestran resultados experimentales ya realizados, sino
-que también es correcta y completa dado que, estándo basada en la lógica, es
+que también es correcta y completa dado que, estando basada en la lógica, es
 amigable al análisis formal.

### 2010-01-25 tesis/propuesta/secciones/resultados.tex
gran parte de las correcciones de ma esther

@@
 El primer experimento consiste de problemas para consultas para viajes aéreos.
-Los servicios concretos son de la forma $V_i(x,y)\qrule\ \flight(x,y,\AL_i)$
+Los servicios concretos son de la forma $V_i(x,y)\qrule\ \vuelo(x,y,\AL_i)$
 donde $\AL_i$ es una
-constante que denota el nombre de una aerolínea, y $\flight(x,y,\AL_i)$ relaciona las
+constante que denota el nombre de una aerolínea, y $\vuelo(x,y,\AL_i)$ relaciona las
 ciudades $x$ e $y$ tales que hay un vuelo entre $x$ e $y$ servido por $\AL_i$.
@@
 
-\[ W(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,z),\,\flight(x_1,x_2,z),\,\ldots,\,\flight(x_n,\NY,z)\,. \]
+\[ W(x_1,\ldots,x_n)\ \qrule\ \vuelo(\PA,x_1,z),\,\vuelo(x_1,x_2,z),\,\ldots,\,\vuelo(x_n,\NY,z)\,. \]
 

### 2010-01-25 tesis/propuesta/secciones/metodologia.tex
todas las correcciones faciles de ma esther listas

@@
 Luego comenzará la implementación de la nueva solución. Para ello se podrá
-utilizar como punto de partida sistemas existentes y modificarlos para mejorar
+utilizar como punto de partida sistemas existentes como el propuesto por Arvelo
+et al. \cite{arvelo:aaai06} y modificarlos para mejorar
 sus características siguiendo las ideas propuestas.

### 2010-01-25 tesis/propuesta/secciones/resultados.tex
todas las correcciones faciles de ma esther listas

@@
 
-Se utiliza una arquitectura comprendida por un catálogo de descripciones de
+Se propone una arquitectura comprendida por un catálogo de descripciones de
 servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
@@
 
-El catálogo se pobla con descripciones de servicios abstractos y concretos; cada
+El catálogo se carga con descripciones de servicios abstractos y concretos; cada
 servicio es descrito en términos de atributos de entrada y salida y anotado con
@@
 
-\subsection{Resultados Experimentales}
+\subsection{Resultados Experimentales Iniciales}
 

### 2010-01-25 tesis/propuesta/secciones/justificacion.tex
reordenamiento de la propuesta segun ma esther. solo faltan correcciones de pags 2,3,9

--- 2010-01-13 tesis/propuesta/secciones/justificacion.tex
@@
 \section{Justificación e Importancia}
-
-\subsection{Ejemplo Motivante}
-
-% flights
-\newcommand{\flight}{\text{\it flight}}
-\newcommand{\UScity}{\text{\it uscity}}
-\newcommand{\national}{\text{\it national}}
-\newcommand{\oneway}{\text{\it one-way}}
-\newcommand{\onestop}{\text{\it one-stop}}
-\newcommand{\flightPA}{\text{\it flight-to-pa}}
-\newcommand{\onestopPA}{\text{\it onestop-to-pa}}
-\newcommand{\fromNY}{\text{\it from-ny}}
-\newcommand{\PA}{\text{PA}}
-\newcommand{\NY}{\text{NY}}
-\newcommand{\AL}{\text{AL}}
-% end flights
-
-Considere un sistema simple de información de vuelos que contiene información
-sobre vuelos entre ciudades e información sobre qué ciudades están en los
-Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
-abstractos $\flight(x,y)$ y $\UScity(x)$. El primero relaciona dos ciudades $x$ e $y$ si
-hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
-E.E.U.U.  o no. Para los servicios concretos, supongamos que las fuentes de
-datos disponibles en Internet contienen la siguiente información:
-
-\begin{enumerate}[--]
-\item $\national(x,y)$ relaciona dos ciudades de E.E.U.U. conectadas por un vuelo directo,
-\item $\oneway(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,
-\item $\onestop(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,
-\item $\flightPA(x)$ dice si hay un vuelo directo desde $x$ a París,
-\item $\onestopPA(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a París con una escala en $y$, y
-\item $\fromNY(x)$ dice si hay un vuelo de Nueva York a $x$.
-\end{enumerate}
-
-Adicionalmente, los servicios concretos se describen con los siguientes
-servicios abstractos:
-
-\begin{alignat*}{1}
-\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
-\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
-\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
-\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
-\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
-\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
-\end{alignat*}
-
-Ahora supongamos que un usuario está interesado en construir un flujo de trabajo
-capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
-en el mundo, tales que los vuelos puedan detenerse en cualquier ciudad.
-
-\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
-
-La siguiente consulta conjuntiva representa el flujo de trabajo que define esta
-solicitud en términos de servicios abstractos. El flujo de trabajo es definido
-de manera que todo lo relacionado sobre el ligamiento de parámetros de
-entrada/salida es resuelto. Así, cualquier instanciación de los servicios
-abstractos en términos de los servicios concretos es una implementación válida
-del flujo de trabajo. Las implementaciones corresponden a composiciones de
-servicios concretos en las cuales un servicio concreto puede implementar uno o
-más servicios abstractos del flujo, pero cada servicio abstracto puede ser
-implementado por exactamente un servicio concreto. Por ejemplo, la siguiente
-composición corresponde a una de estas implementaciones.
-
-\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
-
-Sin embargo, las siguientes dos composiciones no son válidas.
-
-\begin{alignat*}{1}
-I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
-I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
-\end{alignat*}
-
-La primera composición no es válida porque asocia la variable del flujo $y$ a
-las constantes \PA\ y \NY\ que denotan ciudades diferentes. En la otra mano, $I''$
-no lo implementa porque el servicio concreto $\onestop(x,y)$ no recibe como entrada
-o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces
-no es posible asegurarse de que esa ciudad esté asociada a la ciudad $w$ que es
-retornada por el flujo de trabajo.
-
-Estos ejemplos muestran que las instanciaciones correctas de un flujo deben
-manejar constantes de manera que dos constantes distintas no sean asociadas
-entre sí de manera directa o indirectamente vía transitividad, y que todos los
-atributos que aparezcan en un \emph{join} o en la salida sean producidos por los
-servicios concretos seleccionados.
-
-Los parámetros QoS son modelados anotando los servicios concretos con utilidades
-que caracterizan su comportamiento y que luego son agregados durante la
-instanciación. Así, como se mencionó previamente, la mejor instanciación es la
-que minimiza (o maximiza) la agregación de utilidades.
 

### 2010-01-25 tesis/propuesta/secciones/planteamiento.tex
reordenamiento de la propuesta segun ma esther. solo faltan correcciones de pags 2,3,9

@@
 
+\subsection{Ejemplo Motivante}
+
+% flights
+\newcommand{\vuelo}{\text{\it vuelo}}
+\newcommand{\ciudadVE}{\text{\it ciudadVE}}
+\newcommand{\nacional}{\text{\it nacional}}
+\newcommand{\ida}{\text{\it ida}}
+\newcommand{\unaescala}{\text{\it una-escala}}
+\newcommand{\vueloCCS}{\text{\it vuelo-hacia-ccs}}
+\newcommand{\unaparadaCCS}{\text{\it unaescala-hacia-ccs}}
+\newcommand{\desdeVAL}{\text{\it desde-val}}
+\newcommand{\PA}{\text{CCS}}
+\newcommand{\NY}{\text{VAL}}
+\newcommand{\AL}{\text{AL}}
+% end flights
+
+Considere un sistema simple de información de vuelos que contiene información
+sobre vuelos entre ciudades e información sobre qué ciudades están en los
+Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
+abstractos $\vuelo(x,y)$ y $\ciudadVE(x)$. El primero relaciona dos ciudades $x$ e $y$ si
+hay un vuelo directo entre ellas, y el segundo dice si $x$ es una ciudad de
+Venezuela  o no. Para los servicios concretos, supongamos que las fuentes de
+datos disponibles en Internet contienen la siguiente información:
+
+\begin{enumerate}[--]
+\item $\nacional(x,y)$ relaciona dos ciudades de Venezuela conectadas por un vuelo directo,
+\item $\ida(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,
+\item $\unaescala(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,
+\item $\vueloCCS(x)$ dice si hay un vuelo directo desde $x$ a Caracas,
+\item $\unaparadaCCS(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a Caracas con una escala en $y$, y
+\item $\desdeVAL(x)$ dice si hay un vuelo de Valencia a $x$.
+\end{enumerate}
+
+Adicionalmente, los servicios concretos se describen con los siguientes
+servicios abstractos:
+
+\begin{alignat*}{1}
+\nacional(x,y)\   &\qrule\ \vuelo(x,y),\,\ciudadVE(x),\,\ciudadVE(y)\,. \\
+\ida(x,y)\     &\qrule\ \vuelo(x,y)\,. \\
+\unaescala(x,z)\    &\qrule\ \vuelo(x,y),\,\vuelo(y,z)\,. \\
+\vueloCCS(x)\     &\qrule\ \vuelo(x,\PA)\,. \\
+\unaparadaCCS(x,y)\  &\qrule\ \vuelo(x,y),\,\vuelo(y,\PA)\,. \\
+\desdeVAL(x)\       &\qrule\ \vuelo(\NY,x)\,.
+\end{alignat*}
+
+Ahora supongamos que un usuario está interesado en construir una aplicación
+capaz de recuperar los vuelos ida y vuelta con una parada hacia cualquier ciudad
+en el mundo, tales que los vuelos puedan detenerse en cualquier ciudad.
+
+\[ W(x,w,y,z) \qrule \ciudadVE(x),\,\vuelo(x,w),\,\vuelo(w,y),\,\vuelo(y,z),\,\vuelo(z,x)\,. \]
+
+La siguiente consulta conjuntiva representa el flujo de trabajo que define esta
+solicitud en términos de servicios abstractos. El flujo de trabajo es definido
+de manera que todo lo relacionado sobre el ligamiento de parámetros de
+entrada/salida es resuelto. Así, cualquier instanciación de los servicios
+abstractos en términos de los servicios concretos es una implementación válida
+del flujo de trabajo. Las implementaciones corresponden a composiciones de
+servicios concretos en las cuales un servicio concreto puede implementar uno o
+más servicios abstractos del flujo, pero cada servicio abstracto puede ser
+implementado por exactamente un servicio concreto. Por ejemplo, la siguiente
+composición corresponde a una de estas implementaciones.
+
+\[ I(x,w,y,z)\ \qrule\ \nacional(x,w),\,\vueloCCS(w),\,\ida(\PA,z),\,\nacional(z,x)\,. \]
+
+Sin embargo, las siguientes dos composiciones no son válidas.
+
+\begin{alignat*}{1}
+I'(x,w,y,z)\  \qrule\ &\nacional(x,y),\,\vueloCCS(y),\,\desdeVAL(z),\,\nacional(z,x)\,. \\
+I''(x,w,y,z)\ \qrule\ &\unaescala(x,y),\,\ida(y,z),\,\nacional(z,x)\,.
+\end{alignat*}
+
+La primera composición no es válida porque asocia la variable del flujo $y$ a
+las constantes \PA\ y \NY\ que denotan ciudades diferentes. En la otra mano, $I''$
+no lo implementa porque el servicio concreto $\unaescala(x,y)$ no recibe como entrada
+o produce como salida la ciudad intermedia donde el vuelo se detiene y entonces
+no es posible asegurarse de que esa ciudad esté asociada a la ciudad $w$ que es
+retornada por el flujo de trabajo.
+
+Estos ejemplos muestran que las instanciaciones correctas de un flujo deben
+manejar constantes de manera que dos constantes distintas no sean asociadas
+entre sí de manera directa o indirectamente vía transitividad, y que todos los
+atributos que aparezcan en un \emph{join} o en la salida sean producidos por los
+servicios concretos seleccionados.
+
+Los parámetros QoS son modelados anotando los servicios concretos con utilidades
+que caracterizan su comportamiento y que luego son agregados durante la
+instanciación. Así, como se mencionó previamente, la mejor instanciación es la
+que minimiza (o maximiza) la agregación de utilidades.
+
 \subsection{Trabajo Relacionado}
@@
 
-Alrifai y Risse \cite{alrifaiR09} proponen una solución de doble cara que usa un
-algoritmo de programación entera híbrido para obtener la descomposición de QoS
+Alrifai y Risse \cite{alrifaiR09} proponen una solución de dos fases que usa un
+algoritmo de programación entera para obtener la descomposición de QoS
 globales en restricciones locales y luego selecciona los servicios que mejor
@@
 existe $\R'$ tal que $\R(E)\subset\R'(E)\subseteq Q(C)$.
-
-\subsubsection{Teorías Lógicas}
-
-Se usa una solución similar a la descrita en
-\cite{arvelo:aaai06} para codificar el WIP. 
-Se identifican instanciaciones del flujo de trabajo enumerando los modelos de
-una teoría lógica
-$\Delta=\Delta_{com}\cup\Delta_{id}^1\cup\cdots\Delta_{id}^N$
-donde $\Delta_{com}$ especifica cómo combinar $N$ copias independientes de la teoría
-$\Delta_{id}$ que cubren los objetivos en $W$.
-Cada $\Delta^i_{id}$ es una copia etiquetada $\Delta_{id}$ en la cual cada
-literal 
-$\ell$ está etiquetado como $\ell^i$.
-La teoría \emph{Instantiation Description} (ID) $\Delta_{id}$ consiste de
-diferentes grupos de cláusulas que garantizan que sus modelos están en
-correspondencia con instanciaciones parciales, mientras que la teoría
-$\Delta_{com}$ contiene cláusulas adicionales para garantizar una combinación
-correcta y completa de instanciaciones parciales para una instanciación
-completa.
-La teoría ID $\Delta_{id}$ consiste de las siguientes variables:
-
-\begin{enumerate}[--]
-\item $\{v_0,\ldots,v_n\}$ para indicar qué $V_i$ se usa, o $v_0$ para indicar el ID nulo.
-\item $\{g_1,\ldots,g_m\}$ para indicar objetivos cubiertos por la vista.
-\item $\{z_{j,k,i}\}$ para indicar que el objetivo $j$ en $W$ está cubierto por el objetivo $k$ en $V_i$.
-\item $\{t_{x,y}\}$ para indicar que la variable/constante $x$ en $W$ está asociada a la variable/constante $y$ en la vista.
-\end{enumerate}
-
-Los rangos de los índices para las variables $z$ y $t$ dependen del problema.
-
-Las siguientes cláusulas codifican el problema WIP en términos de los WIDs.
-Rajaraman et al.\ \cite{RajaramanSU95} mostraron que para consultas sin negación
-o comparaciones aritméticas, pero con constantes, y $m$ objetivos y $k$
-variables en la cabeza del flujo de trabajo, es suficiente considerar
-instanciaciones de longitud a lo más $N=m+k$.
-
-\begin{enumerate}[C10.]
-\item[C1.] (Al menos una vista es usada): $\bigvee_{i=0}^n v_i$.
-\item[C2.] (A lo más una vista es usada): $\neg v_i\lor\neg v_j$ para $i\neq j$.
-\item[C3.] (Vista nula implica nulidad): $v_0 \Rightarrow \neg g_j$ para $1\leq j\leq m$.
-\item[C4.] (Las vistas son útiles): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ para $1\leq i\leq n$.
-\item[C5.] (Objetivos cubiertos máximo una vez): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ para $i,j,k,l$ apropiados.
-\item[C6.] (Alcance de las vistas): $v_i \Rightarrow \neg g_j$ para objetivos que no pueden ser cubiertos por $V_i$.
-\item[C7.] (Consistencia): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ para $i,j,k$ apropiados.
-\item[C8.] (Variables muertas): $v_i \Rightarrow \neg t_{x,y}$ para todo $x,y$ con $y\notin V_i$.
-\item[C9.] (1-1 para $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ para todas las variables existenciales $y,y'\in V_i$.
-\item[C10.] (Distinguidos): $v_i \Rightarrow \neg t_{x,y}$ para $x$ distinguido y $y\in V_i$ existencial .
-\item[C11.] (Existencial): $v_i\land t_{x,y}\Rightarrow g_j$ para exist.\ $y\in V_i$ y objetivos $g_j$ que contengan $x$.
-\item[C12.] (Coincidencia): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ para todo
-$x,y$ que deban coincidir si $g_j$ es cubierto por el objetivo $k$ en $V_i$.
-\item[C13.] (Si todas las variables en $V_i$ son distinguidas, sólo cubre un objetivo): $v_i \land g_j \Rightarrow \neg g_k$ para vistas $v_i$ apropiadas.
-\end{enumerate}
-
-Estas son las mismas cláusulas usadas por \mcdsat\ para codificar QRPs \cite{arvelo:aaai06}.
-Para poder manejar símbolos constantes apropiadamente, las cláusulas deben ser
-mejoradas con:
-
-\begin{enumerate}[C10.]
-\item[C14.] (Inconsistencia directa 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
-\item[C15.] (Inconsistencia directa 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
-\item[C16.] (Inconsistencia directa 3): $\neg t_{A,B}$.
-\item[C17.] (Transitividad 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
-\item[C18.] (Transitividad 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
-\end{enumerate}
-
-El problema principal al manejar constantes es estar seguro de que no haya dos
-constantes distintas asociadas entre sí directa o indirectamente. Las cláusulas
-C14--C16 quitan asociaciones directamente inconsistentes, mientras que las
-últimas dos cláusulas implementan propagación de asociaciones restringida que
-quita inconsistencias indirectas.
-De la misma manera, la teoría $\Delta_{com}$ que especifica instanciaciones
-completas contiene las siguientes cláusulas:
-
-\begin{enumerate}[C10.]
-\item[C19.] (Cubrir todos los objetivos): $\bigvee_{j=1}^m g^i_j$ para $1\leq i\leq N$.
-\item[C20.] (Cobertura disjuntiva): $g^i_k \Rightarrow \neg g^j_k$ para $i\neq j$.
-\item[C21.] (Quitar simetrías): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
-            para $1\leq j\leq m$ y $1\leq i\leq N$.
-\item[C22.] (Inconsistencia directa 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
-\end{enumerate}
-
-Estas cláusulas proveen una caracterización correcta y completa de WIPs en el
-sentido en que sus modelos están en correspondencia con las instanciaciones del
-flujo de trabajo.
-
-\subsubsection{Parámetros QoS}
-
-Para los parámetros QoS, se supone un modelo de agregación aditiva simple en el
-que cada vista
-$V_i$ está asociada con un costo $c(V_i)$ (negativo si tratamos de utilidad),
-y una instanciación completa con la suma de los costos de sus vistas.
-Una instanciación óptima o mejor es una con costo mínimo, y el valor óptimo del
-WIP es el costo de una instanciación óptima.
-Un WIP siempre tiene un valor óptimo bien definido (si no hay instanciaciones,
-su costo es $\infty$), pero puede tener múltiples mejores instanciaciones.
-El WIP con costos consiste en encontrar todas las instanciaciones óptimas o una
-instanciación, esto depende de la aplicación particular.
-En la formulación propuesta, esto puede hacerse a partir del d-DNNF que codifica
-$\Delta$
-usando la función de calidad de literales $r$ que asigna
-$r(\ell)=c(V_i)$ si
-$\ell=v_i$, y $r(\ell)=0$ if $\ell\notin\{v_1,\ldots,v_n\}$.

### 2010-01-25 tesis/propuesta/secciones/resultados.tex
reordenamiento de la propuesta segun ma esther. solo faltan correcciones de pags 2,3,9

--- 2010-01-25 tesis/propuesta/secciones/resultados.tex
@@
 \section{Resultados Iniciales}
-
-\subsection{Arquitectura del Sistema}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=.9\textwidth]{graphics/architecture}
-\caption{Arquitectura del Sistema}
-\label{fig:architecture}
-\end{figure}
-
-Se utiliza una arquitectura comprendida por un catálogo de descripciones de
-servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
-el Decodificador. La figura~\ref{fig:architecture} muestra la arquitectura global del sistema.
-En esta infraestructura, una instancia del problema de instanciación de flujos
-de trabajo se define como un flujo abstracto representado por una consulta
-conjuntiva sobre servicios abstractos que es dada como entrada junto con un
-conjunto de servicios concretos definidos por vistas de servicios abstractos.
-
-El catálogo se pobla con descripciones de servicios abstractos y concretos; cada
-servicio es descrito en términos de atributos de entrada y salida y anotado con
-un valor real que representa la utilidad QoS del servicio. La descripción de los
-servicios concretos, que son definidos como vistas de los servicios abstractos,
-puede ser generada de manera semiautomática o automática usando herramientas
-tales como el sistema DEIMOS \cite{AmbiteISWC09}.
-
-Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
-corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
-compilador c2d, un componente \emph{off-the-shelf}, compila la fórmula CNF a
-d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
-convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
-parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
-importante remarcar que el proceso de compilación necesita ser realizado sólo
-una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
-compilación resulta ser costosa en términos de tiempo, este costo puede ser
-amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
-instanciaciones con respecto a múltiples valores de los parámetros QoS.
-Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscador
-a una instanciación de flujo de trabajo que resuelve el WIP.
-
-Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
-todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
-una manera libre de \emph{backtracking} todas las instanciaciones del flujo de
-trabajo. Si el usuario está interesado en una mejor instanciación dados
-parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamaño
-del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
-estas pueden ser computadas en tiempo lineal en su número. Finalmente, si el
-usuario está interesado en todas las instanciaciones, estas también pueden ser
-computadas en tiempo lineal en su número. En los últimos dos casos, si ese
-número es exponencial (en el tamaño de la entrada), la enumeración de las
-instanciaciones también lo es pero esta complejidad es intrínsica al problema y
-por lo tanto no puede ser evitada.
-
-\subsection{Resultados Experimentales}
 

### 2010-01-25 tesis/propuesta/secciones/solucion.tex
reordenamiento de la propuesta segun ma esther. solo faltan correcciones de pags 2,3,9

\section{Solución Propuesta}

\subsection{Arquitectura del Sistema}

\begin{figure}[t]
\centering
\includegraphics[width=.9\textwidth]{graphics/architecture}
\caption{Arquitectura del Sistema}
\label{fig:architecture}
\end{figure}

Se propone una arquitectura comprendida por un catálogo de descripciones de
servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
el Decodificador. La figura~\ref{fig:architecture} muestra la arquitectura global del sistema.
En esta infraestructura, una instancia del problema de instanciación de flujos
de trabajo se define como un flujo abstracto representado por una consulta
conjuntiva sobre servicios abstractos que es dada como entrada junto con un
conjunto de servicios concretos definidos por vistas de servicios abstractos.

El catálogo se carga con descripciones de servicios abstractos y concretos; cada
servicio es descrito en términos de atributos de entrada y salida y anotado con
un valor real que representa la utilidad QoS del servicio. La descripción de los
servicios concretos, que son definidos como vistas de los servicios abstractos,
puede ser generada de manera semiautomática o automática usando herramientas
tales como el sistema DEIMOS \cite{AmbiteISWC09}.

Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
compilador c2d, un componente \emph{off-the-shelf}, compila la fórmula CNF a
d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
importante remarcar que el proceso de compilación necesita ser realizado sólo
una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
compilación resulta ser costosa en términos de tiempo, este costo puede ser
amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
instanciaciones con respecto a múltiples valores de los parámetros QoS.
Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscador
a una instanciación de flujo de trabajo que resuelve el WIP.

Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
una manera libre de \emph{backtracking} todas las instanciaciones del flujo de
trabajo. Si el usuario está interesado en una mejor instanciación dados
parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamaño
del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
estas pueden ser computadas en tiempo lineal en su número. Finalmente, si el
usuario está interesado en todas las instanciaciones, estas también pueden ser
computadas en tiempo lineal en su número. En los últimos dos casos, si ese
número es exponencial (en el tamaño de la entrada), la enumeración de las
instanciaciones también lo es pero esta complejidad es intrínsica al problema y
por lo tanto no puede ser evitada.

\subsection{Teorías Lógicas}

Se usa una solución similar a la descrita en
\cite{arvelo:aaai06} para codificar el WIP. 
Se identifican instanciaciones del flujo de trabajo enumerando los modelos de
una teoría lógica
$\Delta=\Delta_{com}\cup\Delta_{id}^1\cup\cdots\Delta_{id}^N$
donde $\Delta_{com}$ especifica cómo combinar $N$ copias independientes de la teoría
$\Delta_{id}$ que cubren los objetivos en $W$.
Cada $\Delta^i_{id}$ es una copia etiquetada $\Delta_{id}$ en la cual cada
literal 
$\ell$ está etiquetado como $\ell^i$.
La teoría \emph{Instantiation Description} (ID) $\Delta_{id}$ consiste de
diferentes grupos de cláusulas que garantizan que sus modelos están en
correspondencia con instanciaciones parciales, mientras que la teoría
$\Delta_{com}$ contiene cláusulas adicionales para garantizar una combinación
correcta y completa de instanciaciones parciales para una instanciación
completa.
La teoría ID $\Delta_{id}$ consiste de las siguientes variables:

\begin{enumerate}[--]
\item $\{v_0,\ldots,v_n\}$ para indicar qué $V_i$ se usa, o $v_0$ para indicar el ID nulo.
\item $\{g_1,\ldots,g_m\}$ para indicar objetivos cubiertos por la vista.
\item $\{z_{j,k,i}\}$ para indicar que el objetivo $j$ en $W$ está cubierto por el objetivo $k$ en $V_i$.
\item $\{t_{x,y}\}$ para indicar que la variable/constante $x$ en $W$ está asociada a la variable/constante $y$ en la vista.
\end{enumerate}

Los rangos de los índices para las variables $z$ y $t$ dependen del problema.

Las siguientes cláusulas codifican el problema WIP en términos de los WIDs.
Rajaraman et al.\ \cite{RajaramanSU95} mostraron que para consultas sin negación
o comparaciones aritméticas, pero con constantes, y $m$ objetivos y $k$
variables en la cabeza del flujo de trabajo, es suficiente considerar
instanciaciones de longitud a lo más $N=m+k$.

\begin{enumerate}[C10.]
\item[C1.] (Al menos una vista es usada): $\bigvee_{i=0}^n v_i$.
\item[C2.] (A lo más una vista es usada): $\neg v_i\lor\neg v_j$ para $i\neq j$.
\item[C3.] (Vista nula implica nulidad): $v_0 \Rightarrow \neg g_j$ para $1\leq j\leq m$.
\item[C4.] (Las vistas son útiles): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ para $1\leq i\leq n$.
\item[C5.] (Objetivos cubiertos máximo una vez): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ para $i,j,k,l$ apropiados.
\item[C6.] (Alcance de las vistas): $v_i \Rightarrow \neg g_j$ para objetivos que no pueden ser cubiertos por $V_i$.
\item[C7.] (Consistencia): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ para $i,j,k$ apropiados.
\item[C8.] (Variables muertas): $v_i \Rightarrow \neg t_{x,y}$ para todo $x,y$ con $y\notin V_i$.
\item[C9.] (1-1 para $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ para todas las variables existenciales $y,y'\in V_i$.
\item[C10.] (Distinguidos): $v_i \Rightarrow \neg t_{x,y}$ para $x$ distinguido y $y\in V_i$ existencial .
\item[C11.] (Existencial): $v_i\land t_{x,y}\Rightarrow g_j$ para exist.\ $y\in V_i$ y objetivos $g_j$ que contengan $x$.
\item[C12.] (Coincidencia): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ para todo
$x,y$ que deban coincidir si $g_j$ es cubierto por el objetivo $k$ en $V_i$.
\item[C13.] (Si todas las variables en $V_i$ son distinguidas, sólo cubre un objetivo): $v_i \land g_j \Rightarrow \neg g_k$ para vistas $v_i$ apropiadas.
\end{enumerate}

Estas son las mismas cláusulas usadas por \mcdsat\ para codificar QRPs \cite{arvelo:aaai06}.
Para poder manejar símbolos constantes apropiadamente, las cláusulas deben ser
mejoradas con:

\begin{enumerate}[C10.]
\item[C14.] (Inconsistencia directa 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
\item[C15.] (Inconsistencia directa 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
\item[C16.] (Inconsistencia directa 3): $\neg t_{A,B}$.
\item[C17.] (Transitividad 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
\item[C18.] (Transitividad 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
\end{enumerate}

El problema principal al manejar constantes es estar seguro de que no haya dos
constantes distintas asociadas entre sí directa o indirectamente. Las cláusulas
C14--C16 quitan asociaciones directamente inconsistentes, mientras que las
últimas dos cláusulas implementan propagación de asociaciones restringida que
quita inconsistencias indirectas.
De la misma manera, la teoría $\Delta_{com}$ que especifica instanciaciones
completas contiene las siguientes cláusulas:

\begin{enumerate}[C10.]
\item[C19.] (Cubrir todos los objetivos): $\bigvee_{j=1}^m g^i_j$ para $1\leq i\leq N$.
\item[C20.] (Cobertura disjuntiva): $g^i_k \Rightarrow \neg g^j_k$ para $i\neq j$.
\item[C21.] (Quitar simetrías): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
            para $1\leq j\leq m$ y $1\leq i\leq N$.
\item[C22.] (Inconsistencia directa 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
\end{enumerate}

Estas cláusulas proveen una caracterización correcta y completa de WIPs en el
sentido en que sus modelos están en correspondencia con las instanciaciones del
flujo de trabajo.

\subsection{Parámetros QoS}

Para los parámetros QoS, se supone un modelo de agregación aditiva simple en el
que cada vista
$V_i$ está asociada con un costo $c(V_i)$ (negativo si tratamos de utilidad),
y una instanciación completa con la suma de los costos de sus vistas.
Una instanciación óptima o mejor es una con costo mínimo, y el valor óptimo del
WIP es el costo de una instanciación óptima.
Un WIP siempre tiene un valor óptimo bien definido (si no hay instanciaciones,
su costo es $\infty$), pero puede tener múltiples mejores instanciaciones.
El WIP con costos consiste en encontrar todas las instanciaciones óptimas o una
instanciación, esto depende de la aplicación particular.
En la formulación propuesta, esto puede hacerse a partir del d-DNNF que codifica
$\Delta$
usando la función de calidad de literales $r$ que asigna
$r(\ell)=c(V_i)$ si
$\ell=v_i$, y $r(\ell)=0$ if $\ell\notin\{v_1,\ldots,v_n\}$.

### 2010-01-25 tesis/propuesta/secciones/planteamiento.tex
todas las correcciones menos la de complejidad del problema listas

@@
 realizar esta búsqueda que sean capaces de escalar a medida que el número de
-servicios concretos o la complejidad del flujo de trabajo aumenta. Se llama al
+servicios concretos o la complejidad (número
+de objetivos, tamaño) de la estructura del flujo de trabajo  aumenta. Se llama al
 problema de instanciar un flujo de trabajo dado con servicios concretos de un
@@
 La solución propuesta por Arvelo et al. \cite{arvelo:aaai06} está basada en la enumeración
-eficiente de modelos de una teoría lógica proposicional. Dada una consulta conjuntiva, una teoría
-lógica es construída de manera que cada modelo de la teoría codifica una
+eficiente de modelos de una teoría lógica proposicional. Dada una consulta conjuntiva
+y un conjunto de fuentes definidas como vistas, se construye una teoría
+lógica de manera que cada modelo de la teoría codifica una
 reescritura válida, y así todas las reescrituras son obtenidas enumerando los

### 2010-01-25 publicaciones/demo_eswc10/paperdemo.tex
demo_eswc10

--- 2009-12-22 publicaciones/eswc10/paperESWC10.tex
@@
  \documentclass{llncs}
-
  \usepackage{amsmath}
@@
 \begin{document}
-
 \allowdisplaybreaks
-\title{Optimal Instantiation of Abstract Workflows using Logic and Circuits}
+\title{SatWins: A Sat-based tool to Optimal Instantiations of Abstract Workflows}
 \author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
@@
            Caracas 89000, Venezuela \\
-           \url{{idaniel,bonet,mvidal}@ldc.usb.ve}}
+           \url{{idaniel,mvidal,bonet}@ldc.usb.ve}}
 \maketitle
@@
 \begin{abstract}
-The typical Service Oriented Architecture consists of two main layers.
-The concrete layer with concrete services whose functionalities are
-described in terms of pre- and post-conditions and non-functional
-properties in terms of QoS parameters, and the abstract layer with
-software applications whose functionalities are described in terms
-of abstract workflows and non-functional properties in terms of
-restrictions on QoS.
-During the execution of a workflow, the abstract services are instantiated
-into concrete services that meet the functional and non-functional
-requirements. This instantiation, that must be done on-the-flight,
-consists in a search on a combinatorial space of possibilities.
-In this paper, we propose a framework to efficiently solve the
+We present SatWins a system to efficiently solve the abstract workflow
 instantiation problem that is firmly grounded on logic.
-The framework adopts the Local-As-View approach in which
+SatWins adopts the Local-As-View approach in which
 the functionality of concrete services is described using views
@@
 and the abstract workflows as conjunctive queries on abstract services.
-Using this representation, the problem of workflow instantiation 
-is cast as a problem of query rewriting, from the area of integration
-systems.
-Then, building on related work, we devise an encoding of the
-workflow instantiation problem as a logical theory whose models
+We discuss the encoding of the workflow instantiation problem as a logical theory whose models
 are in correspondence with the instantiations of the workflow,
 and the best ranked models are in correspondence with the
-optimal instantiations of the workflow.
-Thus, by exploiting known properties of logical theories in 
-d-DNNF format, we provide an efficient and scalable solution
+optimal instantiations of the workflow. We demonstrate how by exploiting known properties of logical theories in 
+d-DNNF format, SatWins provides an efficient and scalable solution
 to the workflow instantiation problem. 
-The approach does not only scale up to large instances, as the
-experimental results show, but is also sound and complete since,
-being founded on logic, is amenable to formal analysis.
 \end{abstract}
@@
 \section{Introduction}
-
-Under the umbrella of the Semantic Web, and supported by Service Oriented
-Architectures (SOA), the number of Web data sources and services has exploded
-during the last few years. For example, the molecular biology database
-collection currently contains 1,170 databases~\cite{Galperin09}, a number
-that is 95 more than the previous year~\cite{Galperin2008} and 110 more
-than two years ago~\cite{Galperin2007}; tools and services, as well as the
-number of instances published by these resources, follow a similar
-progression~\cite{Benson07}. Thanks to this wealth of resources, users'
-tendency is to rely more on automatic methods for handling them such as
-data retrieval from public sources and analysis using Web
-tools or services composed in complex workflows.
-
-The typical SOA consists of two layers. The concrete layers that is
-made of concrete services whose functionalities are described in
-terms of pre- and post-conditions and non-functional properties in
-terms of QoS parameters, and the abstract layer that is made of 
-software applications whose functionalities are described in terms
-of abstract workflows and non-functional properties in terms of
-restrictions on QoS.
+Emerging technologies support the basis for the wealth of available Web resources, as well as current users' tendency to rely more on automatic methods for handling existing Web resources or services composed in complex workflows.
 The execution of an abstract workflow involves the instantiation
-of the abstract services into concrete services that meet the
+of the abstract services into concrete services that meet the workflow
 functional and non-functional requirements. This instantiation
@@
 this search that are able to scale up as the number of concrete
-services or the complexity of the workflow increases.
-We call the problem of instantiating a given abstract workflow
-with concrete services, from a given pool of concrete services,
+services or the complexity of the workflow increases. We call the problem of instantiating a given abstract workflow with concrete services, from a given pool of concrete services,
 such that certain QoS demands are meet, the Workflow Instantiation
@@
 
-In this paper, we consider a restricted version of WIP that adopts
-the Local-As-View (LAV) approach \cite{levy:bucket}.
-In LAV, all elements in a problem are specified with a common language
-that is grounded on \emph{abstract services} such that concrete
-services are described as views of abstract services, the quality
-of an instantiation as an overall utility function that combine the
-different QoS parameters, and the abstract workflow as a conjunctive
-query on the abstract services; this representation is similar to the
+In this paper, we describe SatWins a system that considers
+ a restricted version of WIP that adopts
+the Local-As-View (LAV) approach \cite{levy:bucket} to specify the semantics of the concrete services as views of abstract services; this representation is similar to the
 one that is semi-automatically generated for the DEIMOS system
-\cite{AmbiteISWC09}.
+\cite{AmbiteISWC09}. Additionally,  abstract workflows correspond  to conjunctive
+queries on the abstract services, and  the quality
+of an instantiation is an overall utility function that combine the
+different QoS parameters of the selected concrete services.
 In this version of WIP, the rules that define workflows (conjunctive queries)
@@
 
-Under these assumptions, the WIP can be cast as the well-known Query
-Rewriting Problem (QRP) for LAV which is central to integration systems
-\cite{halevy:survey}. The QRP consists of a conjunctive query that must be
-answered in terms of views in which the query and the views are described
-using LAV with abstract relations.
-This problem is important in the context of data integration
-\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance
-\cite{AfratiLU07,levy:bucket}, and several approaches have been defined
-that scale to a large number of views
-\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-
-The recent approach of Arvelo et al.\ \cite{arvelo:aaai06} is based 
-on the efficient enumeration of models of a propositional logic theory. 
-Given a QRP, a logical theory is constructed such that each model of
-the theory encodes a valid rewriting, and thus all valid rewritings
+In addition to provide a novel approach to describe the WIP problem, SATWins is able to  efficiently identify an abstract workflow instantiations. 
+Given a WIP, a logical theory is constructed such that each model of
+the theory encodes a valid instantiation, and thus all valid instantiations
 are obtained by enumerating all models of the theory. This enumeration
-can be efficiently performed if the logical theory is in certain 
+can be efficiently performed if the logical theory is in a certain 
 normal form called deterministic and decomposable negation normal
@@
 transforming (called compiling in the field of knowledge compilation)
-the logical theory into d-DNNF format for enumerating its models efficiently.
+the logical theory into d-DNNF format for enumerating its models efficiently; also, the  enumeration 
+the best ranked models can also be performed efficiently on d-DNNF and the complexity is independent of the utility function used to rank the optimal instantiation.
 
-Yet d-DNNF theories not only support the efficient enumeration of
-models but other operations too. Among them, the enumeration of
-the best ranked models can also be performed efficiently on d-DNNF.
-Given a literal ranking function $r(\ell)$ that assign ranks
-to each literal $\ell$, one defines the rank $r(\omega)$ of a
-model $\omega$ as the sum of the ranks of the literals made true
-by $\omega$ (i.e., $r(\omega)\doteq\sum_{\omega\vDash\ell}r(\ell)$),
-and say that $\omega$ is a best (ranked) model if there is no model
-$\omega'$ such that $r(\omega')<r(\omega)$.
+In this demonstration, we  attempt to show the following key aspects of the approach implemented by SATWins. We will use a real-world example based on a simple flight-information system.
+\begin{itemize}
+\item We show the expressiveness of representing concrete services as views on abstract services.
+\item We demonstrate the performance of the proposed techniques by computing the optimal instantiation of a given WIP in terms of different utility functions.  
+\item We show the scalability of our solution by demonstrating how SATWins is able to enumerate e realistic instances of the problem in less of one second.
+\end{itemize}
 
-Given a theory in d-DNNF, one computes the rank of the best models,
-and a best model, in time linear in the size of the d-DNNF.
-This computation transforms the DAG of the d-DNNF into a arithmetic
-circuit by replacing the AND nodes with `+' and the OR nodes with `min'.
-The literal ranking function assigns
-values to the leaves of the circuit that are propagated to the
-root in linear time. The value of the root is the rank of the
-best model \cite{darwiche:weighted}.
-
-In this paper, we exploit the properties of d-DNNFs by constructing
-a logical theory whose models encode the valid workflow instantiations
-and best models encode the optimal (best) instantiations.
-Thus, the combinatorial search is reduced to the computation
-of a best model of a logic theory which can be efficiently
-performed once the theory is transformed into d-DNNF format.
-
-The paper contains six more sections. The next section illustrates and
-motivates the WIP with a simple yet typical example. Section~3 summarizes
-the existing related work in three areas of selection of services,
-query rewriting, and knowledge compilation.
-Then, Sections~4 and 5 describe the architecture of the system and 
-report our empirical results over different benchmark problems
-respectively.
-For the interested reader, Section~6 presents a formal description
-of the proposed approach together with an analysis.
-The paper concludes with a final discussion in Section~7.
-
-\section{Motivating Example}
-
-Consider a simple flight-information system which contains information
-about flights between cities and information about which cities are in
-the US. Such a system can be described using LAV with the two abstract
-services $\flight(x,y)$ and $\UScity(x)$. The former relates two cities
-$x$ and $y$ if there is a direct flight between them, and the latter tells
-whether $x$ is a US city or not.
-
-For the concrete services, assume that the available data sources on
-the Internet contain the following information:
-\begin{enumerate}[--]
-\item $\national(x,y)$ relates two US cities that are connected by a direct flight,
-\item $\oneway(x,y)$ relates two cities that are connected by a one-way flight,
-\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
-\item $\flightPA(x)$ tells if there is a direct flight from $x$ to Paris,
-\item $\onestopPA(x,y)$ relates $x$ and $y$ if there is a flight from $x$ to Paris
-      with a stop at $y$, and
-\item $\fromNY(x)$ tells if there is a flight from New York into $x$.
-\end{enumerate}
-Furthermore, the concrete services are described using the abstract services:
-\begin{alignat*}{1}
-\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
-\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
-\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
-\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
-\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
-\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
-\end{alignat*}
-Now suppose that a user is interested in constructing a workflow able to retrieve
-the one-stop round-trip flights from US cities to any city in the world, such that
-flights can stop at any city. 
-\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
-The following conjunctive query represents the workflow that defines this request
-in terms of abstract services. The workflow is defined in way that all issues
-about the binding of input/output parameters between services had been resolved
-such that any instantiation of the abstract services in terms of concrete services
-is a valid implementation of the workflow. 
-Implementations correspond to compositions of concrete services in which a concrete
-service may implement one or more abstract services from the workflow, but each
-abstract service can be implemented by exactly one concrete service.
-For example, the following composition corresponds to one such implementation.
-\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
-However, the following two compositions are not valid.
-\begin{alignat*}{1}
-I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
-I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
-\end{alignat*}
-The first composition is not valid because it maps the workflow variable $y$
-into constants \PA\ and \NY\ that denote different cities.
-On the other hand, $I''$ does not implement the workflow because the concrete
-service $\onestop(x,y)$ does not receive as input, or produce as output, the
-middle city where the flight stops and thus is not possible to ensure that 
-this city is bound to the city $w$ that is returned by the workflow.
-
-These examples show that a proper workflow instantiations must handle
-constants such that not two different constants are mapped to each
-other either directly or indirectly via transitivity, and that all
-attributes that appear in a join or in the output need to be produced
-by the selected concrete services.
-
-The QoS parameters are modeled by annotating the concrete services with
-utilities that characterize their behavior which are then aggregated
-during instantiation. Thus, as said before, the best instantiation is
-the one that minimizes (or maximizes) the aggregation of utilities.
-
-\section{Related Work}
-
-In this section we summarize the existing approaches that provide
-solutions to the problems of service selection, query rewriting and
-discuss related work in the area of Artificial Intelligence called
-knowledge compilation.
-
-\subsection{Service Selection}
-
-The problem of selecting the services that implement an abstract workflow
-and best fit the QoS-based criteria is known as the QoS-aware service
-selection or composition problem, which has been shown to be NP-hard \cite{Hiroshi2008}.
-This problem is a combinatorial optimization problem and several heuristics 
-have been proposed to find a relatively good solution in a reasonably short
-period of time.
-
-Rahmani et al.\ \cite{rahmani08} present a distance metric-based heuristic
-that focus a backward search algorithm; this metric induces an order of the
-services in a way that sink nodes are unlikely to be visited.
-In a series of papers, Berardi and others \cite{berardi05,berardi08,berardi06}
-describe services and workflows in terms of deterministic finite-state machines
-that are encoded using Description Logic theories whose models correspond to
-solutions of the problem. Although reasoning methods for Description Logics
-formalisms could be exploited, scalability or performance of the proposed
-solution has not been reported.
-
-Ko et al.\ \cite{myoung08} propose a constraint-based approach that encodes the
-non-functional permissible values as a set of constraints whose violation needs
-to be minimized; to traverse the space of possibly optimal solutions, a hybrid
-algorithm that combines tabu search and simulated annealing meta-heuristics is
-implemented. Experimental results show that the proposed solution is able to
-scale up to a large number of services and abstract processes.
-Cardellini et al.\ \cite{cardellini07} encode one part of the QoS-aware service
-composition problem as a Linear Programming problem \cite{cardellini07}.
-On the other hand, Wada el at.\ \cite{Hiroshi2008} treat the problem as a 
-multi-objective optimization problem where the different QoS parameters are
-considered equally important instead of aggregating them into a single function.
-Then, a genetic-based algorithm is proposed to identify a set of non-dominated
-service compositions that best fit all the QoS requirements.
-
-Alrifai and Risse \cite{alrifaiR09} propose a two-fold solution that uses a
-hybrid integer programming algorithm to find the decomposition of global QoS
-into local constraints, and then, selects the services that best meet the local
-constraints.
-
-Recently, two planning-based approaches have been proposed.
-Kuter and Golbeck \cite{kuterG09} extend the SHOP2 planning algorithm to select
-the trustworthy composition of services that implement a given OWL-S process model,
-while Sohrabi and McIlraith \cite{sohrabiM09} propose a HTN planning-based solution
-where user preference metrics and domain regulations are used to guide the planner
-into the space of relevant compositions.
-Finally, L\'ecu\'e \cite{lecue09} proposes a genetic-based algorithm to identify
-the composition of services that best meet the quality criteria for a set of QoS
-parameters.
-
-Although these solutions are able to solve the optimization problem and scale up
-to a number of abstract processes, none of them are tailored to semantically describe
-the services in terms of abstract process, nor to use these descriptions to 
-identify the services that implement a given workflow or best meet the user's
-non-functional criteria.
-
-\subsection{Query Rewriting}
-
-A number of algorithms have been developed to find the rewritings of a given query;
-the most prominent being the bucket algorithm \cite{levy:bucket}, the inverse rules
-algorithm \cite{pods:DuschkaG97,Qian96}, the \minicon\ algorithm \cite{pottinger:minicon},
-and the \mcdsat\ algorithm \cite{arvelo:aaai06}.
-Generally, query rewriting algorithms work in two phases. During the first phase,
-the algorithms identify the views that rewrite at least one sub-goal of the query, and
-during the second, these partial rewritings are combined into complete rewritings.
-The main difference between the algorithms is the criteria used to choose the relevant
-views to reduce the space of non-useful rewritings.
-
-The bucket algorithm reduces the number of possibilities by just considering each
-sub-goal in the query in isolation, and selecting the views that are able to produce
-at least the attributes projected by the query. Since the attributes involved in the
-joins in the query are not verified, a large number of rewritings comprised of Cartesian
-products may be generated.
-
-The Inverse Rules algorithm constructs a set of rules that invert the view definitions
-and establish how to compute tuples for the database relations from the tuples of the
-views. Similarly to the bucket algorithm, it can produce a large number of non-useful
-rewritings.
-
-The \minicon\ algorithm overcomes the limitations of the previous algorithms by identifying
-only views that rewrite a set of the sub-goals of the query, and that can be combined with
-the rest of the sub-goals. The key idea is to identify the mappings between the variables in
-each sub-goal to the variables in one or more sub-goals in the views, in a way
-that, join variables in the query are mapped to join variables in the body of a view or to
-the distinguished variables of the view. Mappings between variables and sub-goals are
-represented in the so-called MiniCon Descriptions (MCDs) \cite{pottinger:minicon}.
-
-Finally, the \mcdsat\ algorithm is able to identify the query rewritings of a query by
-translating the problem of rewriting into the problem of enumerating the models of a
-propositional theory whose models are in correspondence with the rewritings of the query.
-The algorithm exploits the properties of d-DNNFs to efficiently enumerate the models of
-the theory.
-The \mcdsat\ algorithm has demonstrated to scale better than the \minicon\ algorithm over
-a large number of benchmarks often showing performance improvements of several orders of
-magnitude. However, the \mcdsat\ algorithm was not designed for rewriting problems
-involving explicit constants, nor to compute the best rewritings with respect to a given
-utility function or cost model. In this paper, we propose an extended encoding that 
-overcomes these limitations and apply the encoding to the Workflow Instantiation Problem.
-
-\subsection{Knowledge Compilation}
-
-Knowledge compilation is the area in AI concerned with the problem of mapping
-logical theories into suitable fragments that make certain desired operations
-tractable \cite{cadoli:compilation}. Different compilation languages have been
-defined, for instance, Ordered Binary Decision Diagrams (OBDDs) \cite{bryant:obdd},
-Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation
-Normal Form (DNNF) \cite{darwiche:map}. In this work, we make use of the properties
-of the deterministic DNNFs (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable
-and efficient solution to the service selection problem.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=.7\textwidth]{odd}
-\caption{A decomposable and deterministic NNF.}
-\label{fig:dnnf}
-\end{figure}
-
-A Negation Normal Form (NNF) theory is constructed from literals using only conjunctions
-and disjunctions \cite{barwise:handbook}, and it can be represented as a directed
-acyclic graph in which the leaves are labeled with literals and the internal nodes
-are labeled with $\land$ and $\lor$; see Fig.~\ref{fig:dnnf} for an example.
-This fragment is universal meaning that for every logical formula there is an
-equivalent one in NNF format.
-An NNF is said to be decomposable (DNNF) \cite{darwiche:d-dnnfs} if for each conjunction
-$\bigwedge_i\phi_i$, the set of variables in each conjunct are pairwise disjoint; i.e,.
-$Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i\neq j$.
-A DNNF supports a number of operations in polytime in the size of its DAG.
-For example, we can test whether a DNNF is satisfiable by a single bottom-up pass
-over its DAG in linear time.
-A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each
-disjunction $\bigvee_i\phi_i$, the disjuncts are pairwise logically contradictory;
-i.e., $\phi_i\land\phi_j\equiv\textbf{false}$ for $i\neq j$.
-The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable and deterministic.
-A d-DNNF supports model counting in polytime in the size of its DAG, and model
-enumeration in polytime in the size of the output.
-Furthermore, given a literal ranking function $r$, one can compute the rank
-of the best model in polytime for DNNFs \cite{darwiche:weighted}.
-
-The fragments DNNF and d-DNNF are universal yet translating a CNF theory 
-to DNNF format has an exponential cost in the worst case. This translation
-is referred to as compilation in the field. There is a publicly available
-compiler, called c2d,\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}
-that perform this compilation process and that make
-use of modern SAT techniques such as conflict-directed backtracking,
-clause learning and caching \cite{darwiche:compiler}.
-This compiler incurs in the worst case in exponential space in a parameter
-called the width of the decomposition tree that is related to the ``connectivity''
-of the CNF theory. However, in our experiments, the CNF theories that are
-compiled are of low width.
-
-%DNNFs had also been used elsewhere to tackle other combinatorial optimization
-%problems such as problems in planning \cite{palacios:icaps05} and diagnosis
-%\cite{REF}.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=.9\textwidth]{architecture}
-\caption{System Architecture}
-\label{fig:architecture}
-\end{figure}
-
-\section{System Architecture}
-
-We use an architecture that is comprised of a Catalog of service descriptions,
+\section{SatWins Architecture}
+ The SatWins architecture  is comprised of a Catalog of service descriptions,
 the Encoder, the compiler c2d, the best model Finder, and the Decoder.
@@
 of abstract services.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=.9\textwidth]{architecture}
+\caption{The SatWins Architecture}
+\label{fig:architecture}
+\end{figure}
 
@@
 
-An input instance of WIP is encoded as a CNF theory whose models corresponds to
-the instantiations of the workflow by the Encoder. The compiler c2d, an off-the-shelf
+An input instance of WIP is encoded as a CNF theory whose models correspond to
+the instantiations of the workflow by the Encoder. 
+The compiler c2d, an off-the-shelf
 component, compiles the CNF formula into d-DNNF.
@@
 into d-DNNF using c2d. The Finder computes a best model given the QoS parameters
-in linear time in the size of the resulting d-DNNF. It is important to remark,
+in linear time in the size of the resulting d-DNNF. It is important to remark
 that the compilation process needs to be performed only once as it does not
@@
 input), the enumeration of the instantiations is also exponential but
-this complexity is intrinsic to the problem and thus cannot be avoided.
-
-In order to make our results more accessible to a general audience, we decided
-to present the experimental results in the next section, and leave the formal
-and theoretical results for the following section. In this way, the interested
-reader may skip the formal details of the approach on a first reading.
-
-\section{Experimental Results}
-
-We conducted an empirical analysis on three benchmarks.
-All the experiments were run on a desktop machine with an Intel
-Core 2 Duo 2GHz CPU and 4Gb of memory, and the time was measured
-with the Unix time command.
-
-The objective of the experiment is to assess the performance
-of the proposal on varying conditions. The main benefit of our approach
-is that one can compile the logical theory for a problem instance
-and then calculate all the instantiations, or the best ones, any 
-number of times. The cost model for finding best instantiations
-can be changed with no need to recompile the logical theory.
-Therefore, the time complexity of our approach is basically the
-time to encode the WIP as a CNF plus the time to compile the
-CNF into a d-DNNF and the time to decode the models. 
-The times to encode and decode are negligible compared to
-the time to compile the CNF. Because of this, we focus on the
-time to compile the benchmark problems.
-
-The first benchmark consists of problems for air-travel queries.
-Concrete services are of the form $V_i(x,y)\qrule\ \flight(x,y,\AL_i)$
-where $\AL_i$ is a constant that denotes the name of an airline, 
-and $\flight(x,y,\AL_i)$ relates the cities $x$ and $y$ such that
-there is a flight between $x$ and $y$ served by $\AL_i$. This concrete
-service is assumed to return all flights between two cities with an
-specific airline. The workflow has the form 
-\[ W(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,z),\,\flight(x_1,x_2,z),\,\ldots,\,\flight(x_n,\NY,z)\,. \]
+this complexity is intrinsic to the problem and thus cannot be avoided. Figure \~ref{fig:snapshot} shows a snapshot of an input instance of WIP and the functionality of SatWins.
 
@@
 \centering
-\includegraphics[width=1\textwidth]{plots/plot1}
-\caption{Compilation times for experiments I and II for different
-number of goals and different number of views. The plots are in
-logarithmic scale, and the time is in seconds.}
-\label{fig:plot1}
+\includegraphics[height=30mm,width=.7\textwidth]{demo}
+\caption{A snapshot of the SatWins interface}
+\label{fig:snapshot}
 \end{figure}
 
-The benchmark includes instances for workflows with 2 to 5 sub-goals and
-sets of 10 to 100 concrete services. The results for the compilation are
-shown in panel (a) of Fig.~\ref{fig:plot1}. This is a plot in logarithmic
-scale that suggests a sub-exponential behavior. In any case, the results
-show good performance since realistic instances of the problem (sets of 100
-airlines with 5-stop flights) can be compiled in 328 seconds.
-The size in disk of the d-DNNF for 100 airlines and 5-stop flights is 3.4Mb.
-On this d-DNNF, the best model can be computed in 0.29 seconds,
-and the enumeration of all models in 0.47 seconds.
+\section{Demonstration Use Cases}
+In this demonstration, we will go through the different steps of the proposed approach. We illustrate our approach on  a simple flight-information system which contains information
+about flights between cities and information about which cities are in
+the US. Such a system can be described using LAV with the two abstract
+services $\flight(x,y)$ and $\UScity(x)$. The former relates two cities
+$x$ and $y$ if there is a direct flight between them, and the latter tells
+whether $x$ is a US city or not.
+We also consider different sets of concrete services; functionality, complexity and size will be changed during the demonstration to show expressiveness, performance and scalability of our approach.
+An example of some of the concrete services to be used is as follows:
+\begin{enumerate}[--]
+\item $\national(x,y)$ relates two US cities that are connected by a direct flight,
+\item $\oneway(x,y)$ relates two cities that are connected by a one-way flight,
+\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
+\item $\flightPA(x)$ tells if there is a direct flight from $x$ to Paris,
+\item $\onestopPA(x,y)$ relates $x$ and $y$ if there is a flight from $x$ to Paris
+      with a stop at $y$, and
+\item $\fromNY(x)$ tells if there is a flight from New York into $x$.
+\end{enumerate}
+The descriptions of these concrete services using the abstract services:
+\begin{alignat*}{1}
+\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
+\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
+\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
+\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
+\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
+\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
+\end{alignat*}
 
-In an attempt to induce an exponential growth in the compilation time,
-in the second benchmark we add a second concrete service for each airline.
-This modification increases the number of valid instantiations from linear
-to exponential since each leg of a flight can now be instantiated by two
-concrete services and thus a flight with $n$ legs may have up to $2^n$
-instantiations.
-We ran the compiler for instances comprising the same number of workflow
-sub-goals and total number of concrete services. The results plotted in
-logarithmic scale are shown in panel (b) of Fig.~\ref{fig:plot1}.
+In addition, the following abstract workflow to retrieve
+the one-stop round-trip flights from US cities to any city in the world, such that
+flights can stop at any city, is represented as a conjunctive query on the abstract services as follows: 
+\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
 
-These tests show good performance for this type of problems, but they do not
-involve concrete services with multiple sub-goals. We therefore designed a
-third experiment that consists unstructured, randomly generated instances.
-Each instance contains three variables per abstract service, ten distinct
-variables and ten distinct constants, six sub-goals in the workflows,
-2 to 5 sub-goals in the concrete services, and a varying number of services.
-The chance that an argument of an abstract service is bound to a constant is 50\%.
-The results are depicted in Fig.~\ref{fig:plot3}.
+Any instantiation of the abstract services in terms of
+concrete services is a valid implementation of the workflow or concrete service compositions.
+For example, the following composition corresponds to one such implementation.
+\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
+However, the following composition is not valid.
+\begin{alignat*}{1}
+I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
+\end{alignat*}
+
+Using this domain, we will demonstrate the following scenarios:
+\begin{itemize}
+\item To show the scalability of the proposed approach, we use the behavior of SATWins in three different benchmarks.
+In all the cases, we show time for compilation, size of the  d-DNNF, and time to enumerate all the models as well as the optimal one.
+We start with a benchmark of instances for workflows with 2 to 5 sub-goals and sets of 10 to 100 concrete services.
+Although, the observed time suggests a sub-exponential behavior,  in any case, the results
+show good performance. For instance, sets of 100
+airlines with 5-stop flights can be compiled in 328 seconds; the size in disk of the d-DNNF is 3.4Mb;
+the best model can be computed in 0.29 seconds, and the enumeration of all models in 0.47 seconds.
+Then, in an attempt to induce an exponential growth in the compilation time, we add a second concrete service for each airline and we will show that SATWins behavior remains similar.
+Finally, we  involve concrete services with multiple sub-goals which are randomly generated.
 The compilation time for these instances does not grow monotonically
@@
 is about 17 hours.
+\item To show the performance of the approach and reusability of existing d-DNNF, we will allow users to choose different utility metrics to compute the best solution.
+We will show that the existing d-DNNF can be used to compute the best model with different utility functions, while the execution time remains the same for the different functions.
+\end{itemize}
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=.8\textwidth]{plots/plot3}
-\caption{Compilation times for experiments III for different number of goals
-and different number of views. The plots are in logarithmic scale, and the time
-is in seconds.}
-\label{fig:plot3}
-\end{figure}
-
-These are preliminary experiments, yet the results show that the
-proposed approach efficiently scales for problems with several goals
-and views. We believe that these results are encouraging and motivate
-us to continue this research. We plan to conduct additional experiments
-with other types and sizes of workflows and, when possible, compare with
-other approaches.
-
-\section{Formalization}
-
-We consider service catalogs of the form $C=\tup{S,T}$ where $S$ is a set
-of predicates that represent abstract services and $T=\{T_s\}_{s\in S}$ is
-a collection of tables that represents the result of evaluating the abstract
-services. In the context of the WIP, a service catalog $C$ is an idealized
-description of the output produced by abstract workflows implemented by
-concrete services described as views.
-A workflow $W$ over $S$ is a conjunctive query of the form 
-\[ W(\vvec{x})\ \qrule\ s_1(\vvec{x}_1),\,s_1(\vvec{x}_2),\,\ldots,\,s_m(\vvec{x}_m)\,, \]
-where $s_i\in S$, $\vvec{x}$ is a vector of variables, and each $\vvec{x}_i$
-is a vector of variables and constants. The result of $W$ over $C$, denoted
-as $W(C)$, is the table with $|\vvec{x}|$ columns that result of the projection
-of the relational join $\join\!\!\{T_{s_i}\}_{i=1}^n$ over $\vvec{x}$.
-The atoms in the body of $W$ are called the (sub)goals of $W$, and the variables
-in the head of $W$ are called distinguished.
-
-A concrete service is described as a view $V$ over $C$ that, following LAV, is
-a query over $S$. Given a catalog $C$, a workflow $W$ and a collection of $n$ views
-$E=\tup{\{V_i\}_{i=1}^n,\{E_i\}_{i=1}^n}$, we are asked to find all the tuples in $W(C)$
-that can be obtained from the views in $E$. That is, we need to find all the
-\emph{instantiations}
-\[ R(\vvec{x})\ \qrule\ V_{i_1}(\vvec{x}_1),\,V_{i_2}(\vvec{x}_2),\,\ldots,\,V_{i_{m_i}}(\vvec{x}_{m_i}) \]
-such that $R(E) \subseteq Q(C)$.
-A Workflow Instantiation Problem is a tuple $\tup{S,W,\{V_i\}_i}$ where $S$
-is a set of predicates that represent abstract services, $W$ is a workflow over
-$S$, and $\{V_i\}_i$ is a collection of views that define the concrete services.
-We assume \emph{safe} problems in the sense that all variables mentioned in the
-head of the workflow (resp.\ in the head of each view) appear in the body of the
-workflow (resp.\ in the body of each view).
-Further, we only deal with WIPs with no arithmetic predicates inside the workflow
-or views. An instantiation $R$ is \emph{valid} if for all catalogs $C=\tup{S,T}$
-and extensions $\{E_i\}_i$, $R(E)\subseteq Q(C)$. A collection $\R$ of valid
-instantiations is a solution if for all service catalogs $C=\tup{S,T}$ and
-extensions $\{E_i\}_i$, there is no $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(C)$.
-
-\subsection{Logical Theories}
-
-We use an approach similar to the one described in \cite{arvelo:aaai06} to
-encode the WIP. We have limited space to make a comprehensive description of
-the logical theory so the reader is referred there for details and formal results.
-
-We identify workflow instantiations by enumerating the models of a logical
-theory $\Delta=\Delta_{com}\cup\Delta_{id}^1\cup\cdots\Delta_{id}^N$
-where $\Delta_{com}$ specifies how to combine $N$ independent copies theories
-$\Delta_{id}$ that cover the goals in $W$.
-Each $\Delta^i_{id}$ is a tagged copy $\Delta_{id}$ in which each literal
-$\ell$ is tagged as $\ell^i$.
-The Instantiation Description (ID) theory $\Delta_{id}$ consists of different
-groups of clauses that guarantees that its models are in correspondence with
-partial instantiations, while the theory $\Delta_{com}$ contains additional
-clauses to guarantee a sound and complete combination of partial instantiations
-into a complete instantiation.
-The ID theory $\Delta_{id}$ consists of the following variables:
-\begin{enumerate}[--]
-\item $\{v_0,\ldots,v_n\}$ to indicate which $V_i$ is used, or $v_0$ to indicate the null ID.
-\item $\{g_1,\ldots,g_m\}$ to indicate the goals covered by the view.
-\item $\{z_{j,k,i}\}$ to indicate that the $j$th goal in $W$ is covered by the $k$th goal in $V_i$.
-\item $\{t_{x,y}\}$ to indicate that the variable/constant $x$ in $W$ is mapped into the
-      variable/constant $y$ in the view.
-\end{enumerate}
-The ranges of the indices for the $z$ and $t$ variables depend on the problem.
-
-The following clauses encode the WIP problem in terms of the WIDs.
-Rajaraman et al.\ \cite{RajaramanSU95} showed that for queries without negation or
-arithmetic comparisons, but with constants, and $m$ goals and $k$ variables in the
-head of the workflow, it is enough to consider instantiations of length at most $N=m+k$.
-\begin{enumerate}[C10.]
-\item[C1.] (At least of view is used): $\bigvee_{i=0}^n v_i$.
-\item[C2.] (At most one view is used): $\neg v_i\lor\neg v_j$ for $i\neq j$.
-\item[C3.] (Null view equals null): $v_0 \Rightarrow \neg g_j$ for $1\leq j\leq m$.
-\item[C4.] (Views are useful): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ for $1\leq i\leq n$.
-\item[C5.] (Subgoals covered at most once): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ for appropriate $i,j,k,l$.
-\item[C6.] (Scope of views): $v_i \Rightarrow \neg g_j$ for goals that cannot be covered by $V_i$.
-\item[C7.] (Consistency): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ for appropriate $i,j,k$.
-\item[C8.] (Dead variables): $v_i \Rightarrow \neg t_{x,y}$ for all $x,y$ with $y\notin V_i$.
-\item[C9.] (1-1 for $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ for all
-           existential variables $y,y'\in V_i$.
-\item[C10.] (Distinguished): $v_i \Rightarrow \neg t_{x,y}$ for distinguished $x$ and existential
-            $y\in V_i$.
-\item[C11.] (Existential): $v_i\land t_{x,y}\Rightarrow g_j$ for exist.\ $y\in V_i$ and
-            goals $g_j$ that contain $x$.
-\item[C12.] (Match): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ for all $x,y$ that must match
-            if $g_j$ is covered by goal $k$ in $V_i$.
-\item[C13.] (If all vars in $V_i$ are distinguished, it covers only one goal):
-            $v_i \land g_j \Rightarrow \neg g_k$ for appropriate views $v_i$.
-\end{enumerate}
-These clauses are the same clauses used by \mcdsat\ to encode QRPs \cite{arvelo:aaai06}.
-In order to properly manage constant symbols, the clauses must be enhanced with:
-\begin{enumerate}[C10.]
-\item[C14.] (Direct inconsistency 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
-\item[C15.] (Direct inconsistency 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
-\item[C16.] (Direct inconsistency 3): $\neg t_{A,B}$.
-\item[C17.] (Transitivity 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
-\item[C18.] (Transitivity 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
-\end{enumerate}
-Recall that, as shown in Section~2, the main issue when handling
-constants is to be sure that not two different constants are mapped
-into each other either directly or indirectly.
-Clauses C14--C16 prune direct inconsistent mappings, while the last
-two clauses implement a restricted propagation of mappings that prune
-indirect inconsistencies.
-Likewise, the theory $\Delta_{com}$ that specifies how a complete
-instantiation contains the following clauses:
-\begin{enumerate}[C10.]
-\item[C19.] (Cover all goals): $\bigvee_{j=1}^m g^i_j$ for $1\leq i\leq N$.
-\item[C20.] (Disjunctive cover): $g^i_k \Rightarrow \neg g^j_k$ for $i\neq j$.
-\item[C21.] (Prune symmetries): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
-            for $1\leq j\leq m$ and $1\leq i\leq N$.
-\item[C22.] (Direct inconsistency 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
-\end{enumerate}
-These clauses provide sound and complete characterization of WIPs in the
-sense that their models are in correspondence with the
-instantiations of the workflow.
-
-\subsection{QoS Parameters}
-
-For the QoS parameters, we assume a simple additive aggregation model in
-which each view $V_i$ is associated with a cost $c(V_i)$ (negative if utility),
-and a complete instantiation with the sum of the cost of its views.
-An optimal or best instantiation is one with minimum cost, and the
-optimal value of the WIP is the cost of an optimal instantiation.
-A WIP always has a well-defined optimal value (if there are no instantiations,
-its cost is $\infty$), but it may have multiple best instantiations.
-The WIP with costs consists in finding all optimal instantiations or
-one instantiation, this depends on the particular application.
-In our formulation, this can be done from the d-DNNF that encodes $\Delta$
-using the literal ranking function $r$ that assigns $r(\ell)=c(V_i)$ if
-$\ell=v_i$, and $r(\ell)=0$ if $\ell\notin\{v_1,\ldots,v_n\}$.
-
-\section{Conclusions and Future Work}
-
-We have shown how the propositional theory used in \mcdsat\ for
-computing rewritings of queries can be extended to support constants
-symbols and thus adapted to the problem of instantiating abstract
-workflows. In this work, we adopted a LAV formulation of the
-workflow instantiation problem typically found in SOAs.
-Our formulation assumes that all concrete services are described
-using views over abstract services and that the workflow is 
-expressed as a conjunctive queries over the abstract services.
-This workflow has resolved all the issues related to the
-binding of input and output parameters of services. What
-remains is the problem of finding the best instantiation
-of the workflow in terms of the QoS parameters.
-This formulation of the problem is supported by systems 
-like DEIMOS that are able to generate in semi- or
-automatic manner LAV descriptions as the one required.
-
-The experimental results show that the approach can be applied to
-real-sized workflow problems. The whole approach is only possible
+\section{Conclusions}
+In this demonstration, we will present a Local-As-View (LAV)-based approach \cite{levy:bucket} to WIP.
+We will show how by using simple conjunctive rules the functionality of concrete services and abstract workflows can be defined.
+In addition, we will show how a propositional-based approach can be used to represent instances of WIP and different scenarios where the performance and scalability of the technique will be observed by the users.
+We will demonstrate that the approach can be applied to
+real-sized workflow problems, while the whole approach is only possible
 when the compilation of the CNF theory into d-DNNF format succeeds.
-In the future, we plan to address this limitation by adapting a
-branch-and-bound optimization algorithm that searches the combinatorial
-space of instantiations but prunes suboptimal branches with admissible
-heuristics.
-
-We also are interested in using other target compilation languages
-like Ordered Binary Decision Diagrams.
-
 \bibliographystyle{abbrv}
 \bibliography{ref}
-
  \end{document}

### 2010-01-25 publicaciones/eswc10/paperESWC10.tex
paper con correcciones

@@
 
-The typical SOA consists of two layers. The concrete layers that is
+The typical SOA consists of two layers. The concrete layer that is
 made of concrete services whose functionalities are described in
@@
 are obtained by enumerating all models of the theory. This enumeration
-can be efficiently performed if the logical theory is in certain 
+can be efficiently performed if the logical theory is in a certain 
 normal form called deterministic and decomposable negation normal
@@
 composition problem as a Linear Programming problem \cite{cardellini07}.
-On the other hand, Wada el at.\ \cite{Hiroshi2008} treat the problem as a 
+On the other hand, Wada et al.\ \cite{Hiroshi2008} treat the problem as a 
 multi-objective optimization problem where the different QoS parameters are
@@
 each sub-goal to the variables in one or more sub-goals in the views, in a way
-that, join variables in the query are mapped to join variables in the body of a view or to
+that join variables in the query are mapped to join variables in the body of a view or to
 the distinguished variables of the view. Mappings between variables and sub-goals are
@@
 An NNF is said to be decomposable (DNNF) \cite{darwiche:d-dnnfs} if for each conjunction
-$\bigwedge_i\phi_i$, the set of variables in each conjunct are pairwise disjoint; i.e,.
+$\bigwedge_i\phi_i$, its variables are pairwise disjoint; i.e,.
 $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i\neq j$.
@@
 compiler, called c2d,\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}
-that perform this compilation process and that make
+that performs this compilation process and that makes
 use of modern SAT techniques such as conflict-directed backtracking,
@@
 
-An input instance of WIP is encoded as a CNF theory whose models corresponds to
+An input instance of WIP is encoded as a CNF theory whose models correspond to
 the instantiations of the workflow by the Encoder. The compiler c2d, an off-the-shelf
@@
 \begin{enumerate}[C10.]
-\item[C1.] (At least of view is used): $\bigvee_{i=0}^n v_i$.
+\item[C1.] (At least one view is used): $\bigvee_{i=0}^n v_i$.
 \item[C2.] (At most one view is used): $\neg v_i\lor\neg v_j$ for $i\neq j$.
@@
 indirect inconsistencies.
-Likewise, the theory $\Delta_{com}$ that specifies how a complete
-instantiation contains the following clauses:
+Likewise, the theory $\Delta_{com}$ that specifies complete
+instantiations contains the following clauses:
 \begin{enumerate}[C10.]
@@
 \end{enumerate}
-These clauses provide sound and complete characterization of WIPs in the
+These clauses provide a sound and complete characterization of WIPs in the
 sense that their models are in correspondence with the

### 2010-01-25 publicaciones/demo_eswc10/paperdemo.tex
algunas correcciones que le hice

@@
 queries on the abstract services, and  the quality
-of an instantiation is an overall utility function that combine the
+of an instantiation is an overall utility function that combines the
 different QoS parameters of the selected concrete services.
@@
 represented by annotating each concrete service description with
-a real number that represent the overall QoS utility of the service.
+a real number that represents the overall QoS utility of the service.
 
-In addition to provide a novel approach to describe the WIP problem, SATWins is able to  efficiently identify an abstract workflow instantiations. 
+In addition to providing a novel approach to describe the WIP problem, SATWins is able to  efficiently identify an abstract workflow's instantiations. 
 Given a WIP, a logical theory is constructed such that each model of
@@
 transforming (called compiling in the field of knowledge compilation)
-the logical theory into d-DNNF format for enumerating its models efficiently; also, the  enumeration 
-the best ranked models can also be performed efficiently on d-DNNF and the complexity is independent of the utility function used to rank the optimal instantiation.
+the logical theory into d-DNNF format for enumerating its models efficiently; also, the  enumeration  of
+the best ranked models can also be performed efficiently on a theory in d-DNNF and
+the complexity is independent of the utility function used to rank the optimal instantiation.
 
@@
 \item We demonstrate the performance of the proposed techniques by computing the optimal instantiation of a given WIP in terms of different utility functions.  
-\item We show the scalability of our solution by demonstrating how SATWins is able to enumerate e realistic instances of the problem in less of one second.
+\item We show the scalability of our solution by demonstrating how SATWins is able to enumerate realistic instances of the problem in less than one second.
 \end{itemize}
@@
 Figure~\ref{fig:architecture} depicts the overall architecture of the system.
-In this framework, an instance of Workflow Instantiation Problem is defined as an
+In this framework, an instance of the Workflow Instantiation Problem is defined as an
 abstract workflow represented by a conjunctive query on abstract services which
-is given as input together with a set set of concrete services defined by views
+is given as input together with a set of concrete services defined by views
 of abstract services.
@@
 input), the enumeration of the instantiations is also exponential but
-this complexity is intrinsic to the problem and thus cannot be avoided. Figure \~ref{fig:snapshot} shows a snapshot of an input instance of WIP and the functionality of SatWins.
+this complexity is intrinsic to the problem and thus cannot be avoided.
+Figure \~ref{fig:snapshot} shows a snapshot of an input instance of WIP and the functionality of SatWins.
 
@@
 \section{Demonstration Use Cases}
-In this demonstration, we will go through the different steps of the proposed approach. We illustrate our approach on  a simple flight-information system which contains information
+In this demonstration, we will go through the different steps of the proposed approach.
+We illustrate our approach on  a simple flight-information system which contains information
 about flights between cities and information about which cities are in
@@
 $x$ and $y$ if there is a direct flight between them, and the latter tells
-whether $x$ is a US city or not.
-We also consider different sets of concrete services; functionality, complexity and size will be changed during the demonstration to show expressiveness, performance and scalability of our approach.
-An example of some of the concrete services to be used is as follows:
+whether $x$ is a US city or not. We also consider different sets of concrete services; functionality, complexity and size will be changed during the demonstration to show
+the expressiveness, performance and scalability of our approach. An example of some of the concrete services to be used is as follows:
 \begin{enumerate}[--]
@@
 \begin{itemize}
-\item To show the scalability of the proposed approach, we use the behavior of SATWins in three different benchmarks.
-In all the cases, we show time for compilation, size of the  d-DNNF, and time to enumerate all the models as well as the optimal one.
-We start with a benchmark of instances for workflows with 2 to 5 sub-goals and sets of 10 to 100 concrete services.
-Although, the observed time suggests a sub-exponential behavior,  in any case, the results
+\item To show the scalability of the proposed approach, we use the behavior of
+SATWins in three different benchmarks. In all the cases, we show time for
+compilation, size of the  d-DNNF, and time to enumerate all the models as well
+as the optimal one. We start with a benchmark of instances for workflows with 2
+to 5 sub-goals and sets of 10 to 100 concrete services. Although the observed
+time suggests a sub-exponential behavior, in any case, the results
 show good performance. For instance, sets of 100

### 2010-01-26 publicaciones/demo_eswc10/paperdemo.tex
paper demo con mas correcciones

@@
 \allowdisplaybreaks
-\title{SatWins: A Sat-based tool to Optimal Instantiations of Abstract Workflows}
+\title{SatWins: A Sat-based Tool to Optimal Instantiations of Abstract Workflows}
 \author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
@@
 \begin{abstract}
-We present SatWins a system to efficiently solve the abstract workflow
+We present SatWins, a system to efficiently solve the abstract workflow
 instantiation problem that is firmly grounded on logic.
@@
 of abstract services, the quality of an instantiation as an
-overall utility function that combines the different QoS parameters,
+overall utility function that combines different QoS parameters,
 and the abstract workflows as conjunctive queries on abstract services.
 We discuss the encoding of the workflow instantiation problem as a logical theory whose models
-are in correspondence with the instantiations of the workflow,
+correspond to the instantiations of the workflow,
 and the best ranked models are in correspondence with the
-optimal instantiations of the workflow. We demonstrate how by exploiting known properties of logical theories in 
+optimal instantiations. We demonstrate how by exploiting known properties of logical theories in 
 d-DNNF format, SatWins provides an efficient and scalable solution
@@
 services or the complexity of the workflow increases. We call the problem of instantiating a given abstract workflow with concrete services, from a given pool of concrete services,
-such that certain QoS demands are meet, the Workflow Instantiation
+such that certain QoS demands are met, the Workflow Instantiation
 Problem (WIP).
 
-In this paper, we describe SatWins a system that considers
+In this paper we describe SatWins, a system that considers
  a restricted version of WIP that adopts
@@
 different QoS parameters of the selected concrete services.
-In this version of WIP, the rules that define workflows (conjunctive queries)
-and concrete services (views) are created in a way that all the 
+Rules that define workflows (conjunctive queries)
+and concrete services (views) are created in a way such that all the 
 functional restrictions on pre- and post-conditions of services
-and their combinations are satisfied, and the QoS measures are
-represented by annotating each concrete service description with
-a real number that represents the overall QoS utility of the service.
+and their combinations are satisfied; the QoS measures correspond to real numbers that represent the overall QoS utility of the services.
 
-In addition to providing a novel approach to describe the WIP problem, SATWins is able to  efficiently identify an abstract workflow's instantiations. 
+In addition to providing a novel approach to describe the WIP problem, SATWins is able to  efficiently identify abstract workflow's instantiations. 
 Given a WIP, a logical theory is constructed such that each model of
@@
 transforming (called compiling in the field of knowledge compilation)
-the logical theory into d-DNNF format for enumerating its models efficiently; also, the  enumeration  of
-the best ranked models can also be performed efficiently on a theory in d-DNNF and
-the complexity is independent of the utility function used to rank the optimal instantiation.
+the logical theory into d-DNNF format for enumerating its models efficiently. The  enumeration  of
+the best models can also be performed efficiently on a theory in d-DNNF, and the same d-DNNF can be reused for different utility functions.
 
-In this demonstration, we  attempt to show the following key aspects of the approach implemented by SATWins. We will use a real-world example based on a simple flight-information system.
+In the demonstration, we will use a simple flight-information system to show the following key issues of the proposed approach.
 \begin{itemize}
-\item We show the expressiveness of representing concrete services as views on abstract services.
-\item We demonstrate the performance of the proposed techniques by computing the optimal instantiation of a given WIP in terms of different utility functions.  
-\item We show the scalability of our solution by demonstrating how SATWins is able to enumerate realistic instances of the problem in less than one second.
+\item We demonstrate expressiveness of representing concrete services as views on abstract services.
+\item We discuss scalability  by demonstrating how SATWins is able to enumerate realistic instances of the problem in less than one second.
+\item We demonstrate the performance  by enumerating in the same d-DNNF, the optimal models based on different utility functions.
+
 \end{itemize}
@@
 Figure~\ref{fig:architecture} depicts the overall architecture of the system.
-In this framework, an instance of the Workflow Instantiation Problem is defined as an
+In SATWins, an instance of the Workflow Instantiation Problem is defined as an
 abstract workflow represented by a conjunctive query on abstract services which
@@
 \centering
-\includegraphics[width=.9\textwidth]{architecture}
+\includegraphics[height=20mm,width=.6\textwidth]{architecture}
 \caption{The SatWins Architecture}
@@
 this complexity is intrinsic to the problem and thus cannot be avoided.
-Figure \~ref{fig:snapshot} shows a snapshot of an input instance of WIP and the functionality of SatWins.
+Figure ~\ref{fig:snapshot} shows a snapshot of an input instance of WIP and the functionality of SatWins.
 
@@
 \centering
-\includegraphics[height=30mm,width=.7\textwidth]{demo}
+\includegraphics[height=85mm,width=.8\textwidth]{demo}
 \caption{A snapshot of the SatWins interface}
@@
 $x$ and $y$ if there is a direct flight between them, and the latter tells
-whether $x$ is a US city or not. We also consider different sets of concrete services; functionality, complexity and size will be changed during the demonstration to show
-the expressiveness, performance and scalability of our approach. An example of some of the concrete services to be used is as follows:
+whether $x$ is a US city or not. Some concrete services are as follows:
 \begin{enumerate}[--]
@@
 
-Using this domain, we will demonstrate the following scenarios:
+Using this flight-based system, we will demonstrate the following scenarios:
 \begin{itemize}
-\item To show the scalability of the proposed approach, we use the behavior of
-SATWins in three different benchmarks. In all the cases, we show time for
-compilation, size of the  d-DNNF, and time to enumerate all the models as well
-as the optimal one. We start with a benchmark of instances for workflows with 2
+\item To show the scalability of the proposed approach, we run SATWins in three different benchmarks. In each scenario, we show time for
+compilation, size of the  d-DNNF, and time to enumerate all the models and the optimal one. We start with a benchmark of instances for workflows with 2
 to 5 sub-goals and sets of 10 to 100 concrete services. Although the observed
@@
 \item To show the performance of the approach and reusability of existing d-DNNF, we will allow users to choose different utility metrics to compute the best solution.
-We will show that the existing d-DNNF can be used to compute the best model with different utility functions, while the execution time remains the same for the different functions.
+We will show that the existing d-DNNF can be used to compute the best model with different utility functions, while the execution time remains similar for the different functions.
 \end{itemize}
@@
 In this demonstration, we will present a Local-As-View (LAV)-based approach \cite{levy:bucket} to WIP.
-We will show how by using simple conjunctive rules the functionality of concrete services and abstract workflows can be defined.
-In addition, we will show how a propositional-based approach can be used to represent instances of WIP and different scenarios where the performance and scalability of the technique will be observed by the users.
+We will show how by using simple conjunctive rules, the functionality of concrete services and abstract workflows can be defined.
+In addition, we will show how a propositional logic-based approach can be used to represent instances of WIP, and different scenarios where the performance and scalability of the technique will be observed by  users.
 We will demonstrate that the approach can be applied to

### 2010-01-27 tesis/propuesta/secciones/cronograma.tex
casi todas las correcciones del segundo lote de ma esther

@@
 \hline
-Investigación sobre el estado del arte & 2 semanas & \\
-Identificación de limitaciones de propuestas actuales & 2 semanas \\
-Implementación de generación de cláusulas nuevas para constantes & 6 semanas & \\
-Experimentos con generación de cláusulas nuevas & & \\
-Implementación de búsqueda de mejor reescritura & & \\
-Experimentos con búsqueda de mejor reescritura & & \\
-Implementación de sistema de eliminación de cláusulas para uso de
-\emph{Branch and Bound} & & \\
+\hline
+Investigación sobre el estado del arte & 2 semanas & 2 de octubre de 2009\\
+\hline
+Identificación de limitaciones de propuestas actuales & 2 semanas & 16 de octubre de 2009\\
+\hline
+Propuesta de formalización del problema incluyendo manejo de constantes y costos & 2 semanas & 30 de octubre de 2009 \\
+\hline
+Pruebas formales sobre nueva formalización & 2 semanas &  13 de noviembre de 2009 \\
+\hline
+Implementación de generación de cláusulas nuevas para constantes & 2 semanas & 27 de noviembre de 2009 \\
+\hline
+Experimentos con generación de cláusulas nuevas & 1 semana & 04 de diciembre de 2009 \\
+\hline
+Implementación de búsqueda de mejor reescritura & 2 semanas & 11 de diciembre de 2009 \\
+\hline
+Experimentos con búsqueda de mejor reescritura & 1 semana & 18 de diciembre de 2009 \\
+\hline
+Implementación de sistema de eliminación de cláusulas para uso de \emph{Branch and Bound} & 5 semanas & 12 de febrero de 2010 \\
+\hline
+Experimentos sobre método \emph{Branch and Bound} & 3 semanas & 8 de marzo de 2010 \\
 \hline

### 2010-01-27 tesis/propuesta/secciones/justificacion.tex
casi todas las correcciones del segundo lote de ma esther

@@
 
-\subsection{Complejidad del Problema}
-
-Estamos tratando con un problema difícil \cite{lenzerini:dataintegration}. 
-El cuadro~\ref{cqcomplexity} ilustra la complejidad de distintos tipos de
+Estamos tratando con un problema difícil \cite{lenzerini:dataintegration} en el
+sentido de que es computacionalmente complejo. Más formalmente, el problema
+pertenece a coNP.
+coNP es la clase de los problemas tales que se puede verificar eficientemente (en
+tiempo polinomial) demostraciones de que no tienen ninguna solucion.
+El cuadro~\ref{cqcomplexity}
+ilustra la complejidad de distintos tipos de
 problemas de resolución de consultas basadas en vistas. El caso particular que
-se está tratando en este trabajo es CQ con búsqueda exacta, el cual está en
-coNP.
+se está tratando en este trabajo es CQ con búsqueda exacta.
 
@@
 \end{table}
+
+Debido

### 2010-01-27 tesis/propuesta/secciones/metodologia.tex
casi todas las correcciones del segundo lote de ma esther

@@
 La finalización del proceso de implementación dará paso a la fase de evaluación de
-la propuesta. Se propone realizar series de experimentos diversos sobre el
-producto final que ilustren sus capacidades y rendimiento. En casos en
-que soluciones ya existentes tengan las mismas capacidades, se debe comparar el
-rendimiento de ambas sobre una variedad de casos de prueba que persigan revelar
-sus peores comportamientos. Esto permitirá determinar empíricamente las virtudes
-y desventajas de la solución propuesta.
+la propuesta. Para determinar las virtudes y desventajas de la solución se utilizarán
+dos estrategias.
+
+En primer lugar es necesario verificar las propiedades de completitud y
+correctitud del conjunto de soluciones dado para cada instancia del problema.
+Estas propiedades se probarán analíticamente, demostrando que las instanciaciones
+de \emph{workflows} enumeradas haciendo uso del enfoque propuesto son todas válidas
+y que no queda ninguna válida sin enumerar.
+
+Hecho esto se conducirá un estudio empírico que ilustre la eficiencia
+y la eficacia del producto final. En casos en
+que soluciones ya existentes en la literatura tengan las mismas capacidades, se
+debe comparar el rendimiento de ambas sobre una variedad de casos de prueba que
+persigan revelar sus peores comportamientos.

### 2010-01-27 tesis/propuesta/secciones/planteamiento.tex
casi todas las correcciones del segundo lote de ma esther

@@
 problema de instanciar un flujo de trabajo dado con servicios concretos de un
-conjunto de servicios concretos dado, de manera que ciertas demandas de QoS sean
+conjunto de servicios concretos dado, de manera que ciertas restricciones sobre QoS sean
 cumplidas, el Problema de Instanciación de Flujos de Trabajo (WIP).
@@
 que minimiza (o maximiza) la agregación de utilidades.
-
-\subsection{Trabajo Relacionado}
-
-En esta sección se resumen los enfoques existentes que proveen soluciones
-a los problemas de selección de servicios y reescritura de consultas y se
-discute el trabajo relacionado en el área de inteligencia artificial llamado
-compilación del conocimiento.
-
-\subsubsection{Selección de Servicios}
-
-El problema de seleccionar los servicios que implementan un flujo de trabajo
-abstracto y cumplen mejor los criterios basados en QoS es conocido como el
-problema de selección o composición de servicios consciente de QoS, que ha sido
-demostrado ser \emph{NP-hard} \cite{Hiroshi2008}. Este es un problema de optimización
-combinatoria y varias heurísticas han sido propuestas para conseguir una
-solución relativamente buena en un período razonable de tiempo.
-
-Rahmani et al. \cite{rahmani08} presentan una heurística basada en una métrica de distancia
-que guía un algoritmo de búsqueda; esta métrica induce un orden de
-los servicios en una manera en que los nodos vertedero (sin salida) tienen baja probabilidad
-de ser visitados. En una serie de artículos, Berardi y otros \cite{berardi05,berardi08,berardi06}
-describen
-servicios y flujos de trabajo en términos de máquinas de estados finitos
-determinísticas que son codificadas usando teorías de Lógicas Descriptivas
-cuyos modelos corresponden a soluciones del problema. Aunque se podrían explotar
-métodos para formalismos de Lógicas Descriptivas, no han sido reportados la
-escalabilidad o el rendimiento de la solución propuesta.
-
-Ko et al. \cite{myoung08} proponen un enfoque basado en restriccciones que codifica
-los valores permisibles no funcionales como un conjunto de restricciones cuya
-violación debe ser minimizada; para recorrer el espacio de soluciones
-posiblemente óptima, se implementa un algoritmo híbrido que combina las
-metaheurísticas \emph{tabu search} y \emph{simulated annealing}. Los resultados
-experimentales muestran que la solución propuesta es capaz de escalar a un
-número grande de servicios y procesos abstractos. Cardellini et al. \cite{cardellini07}
-codifican una parte del problema de composición de servicios caracterizados por QoS
-como un problema de programación lineal \cite{cardellini07}. Por otra parte, Wada et al.
-\cite{Hiroshi2008} trata el problema como uno de optimización con múltiples objetivos donde
-los diferentes parámetros de QoS son considerados igualmente importantes en vez
-de ser agregados en una función. Luego se propone un algoritmo basado en
-algoritmos genéticos que identifica un conjunto de composiciones de servicios
-no dominadas que mejor satisfacen los requerimientos QoS.
-
-Alrifai y Risse \cite{alrifaiR09} proponen una solución de dos fases que usa un
-algoritmo de programación entera para obtener la descomposición de QoS
-globales en restricciones locales y luego selecciona los servicios que mejor
-cumplen las restricciones locales.
-
-Recientemente, dos soluciones basadas en planificación han sido propuestas.
-Kuter y Golbeck \cite{kuterG09} extienden el algoritmo de planificación SHOP2 para
-seleccionar la composición confiable de servicios que implementan un modelo de
-procesos OWL-S dado, mientras que Sohrabi y McIlraith \cite{sohrabiM09} proponen una
-solución basada en planificación HTN donde las métricas de preferencia del
-usario y regulaciones de dominio son utilizadas para guíar el planificador hacia
-el espacio de composiciones relevantes. Finalmente, Lécué \cite{lecue09} propone un
-algoritmo basado en genética para identificar la composición de servicios que
-mejor cumple con los criterios de calidad para un conjunto de parámetros QoS.
-
-Aunque estas soluciones son capaces de resolver el problema de optimización y
-escalar a un número de procesos abstractos, ninguno de ellos está ajustado para
-describir semánticamente los servicios en términos de procesos abstractos, ni
-para usar estas descripciones para identificar los servicios que implementan un
-flujo de trabajo dado o que mejor cumplen los criterios no funcionales del
-usuario.
-
-\subsubsection{Reescritura de Consultas}
-
-Un número de algoritmos han sido desarrollados para encontrar las reescrituras
-de una consulta dada; los más prominentes son el algoritmo \emph{bucket} \cite{levy:bucket}, el
-algoritmo de reglas inversas \cite{pods:DuschkaG97,Qian96}, el algoritmo
-\minicon \cite{pottinger:minicon}, y el
-algoritmo \mcdsat \cite{arvelo:aaai06}. Generalmente, la reescritura de consultas trabaja en
-dos fases. Durante la primera, el algoritmo identifica las vistas que reescriben
-al menos un subobjetivo de la consulta, y durante el segundo estas reescrituras
-parciales son combinadas en reescrituras completas. La diferencia principal
-entre los algoritmos es el conjunto de criterios usado para elegir las vistas
-relevantes para reducir el espacio de reescrituras no útiles.
-
-El algoritmo denominado \emph{bucket} reduce el número de posibilidades considerando sólo cada
-subobjetivo de la consulta aislado, y seleccionando las vistas que son capaces
-de producir al menos los atributos proyectados por la consulta. Como los
-atributos involucrados en los \emph{joins} de la consulta no son verificados, un número
-grande de reescrituras comprendido por productos cartesianos puede ser generado.
-
-El algoritmo de reglas inversas construye un conjunto de reglas que invierten
-las definiciones de vistas y establecen cómo computar tuplas para las relaciones
-de la base de datos a partir de las tuplas de las vistas. Al igual que el algoritmo
-bucket, puede producir un gran número de reescrituras no útiles.
-
-El algoritmo \minicon\ supera las limitaciones de los algoritmos previos
-identificando solamente vistas que reescriben un conjunto de los subobjetivos de
-la consulta, y que pueden ser combinados con el resto de los subobjetivos. La
-idea principal es identificar las asociaciones entre variables de cada
-subobjetivo a las variables en uno o más subojbetivos en las vistas, de manera
-que las variables de \emph{join} de la consulta son asociadas a las variables de
-\emph{join}
-del cuerpo de una vista o a variables distinguidas de la vista. Asociaciones
-entre variables y subojbetivos son representadas en las llamadas \emph{MiniCon
-Descriptions} (MCDs) \cite{pottinger:minicon}.
-
-Finalmente, el algoritmo \mcdsat\ es capaz de identificar las reescrituras de una
-consulta traduciendo el problema de reescritura al problema de enumeración de
-modelos de una teoría proposicional cuyos modelos están en correspondencia con
-las reescrituras de la consulta. El algoritmo explota las propiedades de los
-d-DNNF para enumerar eficientemente los modelos de la teoría.
-El algoritmo
-\mcdsat\ ha demostrado escalar mejor que el algoritmo \minicon\ sobre un número
-grande de experimentos donde frecuentemente muestra mejoras de rendimiento de
-varios órdenes de magnitud. Sin embargo, el algoritmo \mcdsat\ no fue diseñado
-para problemas de reescritura que involucren constantes explícitas, ni para
-computar las mejores reescrituras respecto a una función de utilidad o modelo de
-costo dado. En este trabajo se propone una codificación extendida que supera
-estas limitaciones y se aplica la codificación al problema de instanciación de
-flujos de trabajo.
-
-\subsubsection{Compilación del Conocimiento}
-
-Compilación del conocimiento es el área en inteligencia artificial que se ocupa
-del problema de traducir teorías lógicas en fragmentos apropiados que hagan
-tratables ciertas operaciones deseadas \cite{cadoli:compilation}. Diferentes lenguajes de
-compilación han sido definidos. Por ejemplo están los \emph{Ordered
-Binary Decision Diagrams} (OBDDs) \cite{bryant:obdd},
-\emph{Negation Normal Form} (NNF) \cite{barwise:handbook}, y
-\emph{Decomposable Negation Normal Form} (DNNF) \cite{darwiche:map}. En este trabajo se usan las
-propiedades de los DNNF determinísticos (d-DNNF) \cite{darwiche:d-dnnfs} para proveer una
-solución eficiente y escalable al problema de selección de servicios.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=.7\textwidth]{graphics/odd}
-\caption{Un NNF descomponible y determinístico.}
-\label{fig:dnnf}
-\end{figure}
-
-Una teoría en Negation Normal Form es construida desde literales usando sólo
-conjunciones y disjunciones \cite{barwise:handbook}, y puede ser representada como un grafo
-acíclico dirigido en que las hojas son etiquetadas con literales y los nodos
-internos con $\land$ y $\lor$; ver ejemplo en la figura~\ref{fig:dnnf}. Esta
-forma normal es universal, lo que significa que para cada fórmula lógica hay una
-equivalente en formato NNF. Se dice que un NNF es descomponible (DNNF)
-\cite{darwiche:d-dnnfs}
-si para cada conjunción $\bigwedge_i\phi_i$, sus variables son disjuntas por pares; es
-decir, $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ para $i\neq j$. Un DNNF soporta un
-número de operaciones en tiempo polinomial en el tamaño de su GAD. Por ejemplo,
-podemos probar si un DNNF es satisfactible haciendo un sólo recorrido 
-sobre su GAD en tiempo linear. Se dice que un DNNF es
-determinístico (d-DNNF) \cite{darwiche:d-dnnfs} si para cada disjunción $\bigvee_i\phi_i$, los
-disjuntos son lógicamente contradictorios por pares, es decir,
-$\phi_i\land\phi_j\equiv\textbf{false}$ para $i\neq j$.
-El NNF en la figura~\ref{fig:dnnf}, por ejemplo, es
-descomponible y determinístico. Un d-DNNF soporta conteo de modelos en
-tiempo polinomial en el tamaño de su GAD, y enumeración de modelos en tiempo
-polinomial en el tamaño de la salida. Además, dada una función de calidad de
-literales $r$, se puede computar la calidad del mejor modelo en tiempo
-polinomial para DNNFs \cite{darwiche:weighted}.
-
-Toda teoría puede representarse en las formas DNNF y d-DNNF pero traducir una teoría CNF a
-formato DNNF tiene costo exponencial en el peor caso. Esta traducción se
-llama compilación en este campo. Hay un compilador públicamente disponible,
-llamado c2d\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}, que realiza este proceso de compilación y que hace uso
-de técnicas de SAT modernas como backtracking dirigido por conflictos,
-aprendizaje de cláusulas y caching \cite{darwiche:compiler}. Este compilador incurre en el peor
-caso (peor gasto) en espacio exponencial en un parámetro llamado el ancho del árbol de
-descomposición que está relacionado a la ``conectividad'' de la teoría CNF.
-Sin embargo, en nuestros experimentos, las teorías CNF que son compiladas son de
-ancho pequeño.
-
-\subsection{Formalización}
-
-Consideramos catálogos de servicio de la forma $C=\tup{S,T}$ donde $S$ es un
-conjunto de predicados que representa servicios abstractos y $T=\{T_s\}_{s\in S}$ es una
-colección de tablas que representa el resultado de evaluar los servicios
-abstractos. En el contexto del WIP, un catálogo de servicio $C$ es una
-descripción virtual de la salida producida por flujos abstractos
-implementados por servicios concretos descritos como vistas. Un flujo de trabajo
-$W$ sobre $S$ es una consulta conjuntiva de la forma
-
-\[ W(\vvec{x})\ \qrule\ s_1(\vvec{x}_1),\,s_1(\vvec{x}_2),\,\ldots,\,s_m(\vvec{x}_m)\,, \]
-
-donde $s_i\in S$, $\vvec{x}$ es un vector de variables y cada $\vvec{x}_i$ es un vector de variables y
-constantes. El resultado de $W$ sobre $C$, denotado como $W(C)$, es la tabla con
- $|\vvec{x}|$ columnas que resulta de la proyección del \emph{join} relacional
-$\join\!\!\{T_{s_i}\}_{i=1}^n$ sobre $\vvec{x}$. Los átomos en el cuerpo de $W$ son llamados los
-(sub)objetivos de $W$, y las variables en la cabeza de $W$ son llamadas
-distinguidas.
-
-Un servicio concreto es descrito como una vista $V$ sobre $C$ que, siguiendo
-LAV, es una consulta sobre $S$. Dado un catálogo $C$, un flujo de trabajo $W$ y
-una colección de $n$ vistas $E=\tup{\{V_i\}_{i=1}^n,\{E_i\}_{i=1}^n}$, se pide
-conseguir todas las tuplas en
-$W(C)$ que puedan ser obtenidas de vistas en $E$. Es decir, se necesitan
-conseguir las \emph{instanciaciones}
-
-\[ R(\vvec{x})\ \qrule\ V_{i_1}(\vvec{x}_1),\,V_{i_2}(\vvec{x}_2),\,\ldots,\,V_{i_{m_i}}(\vvec{x}_{m_i}) \]
-
-tales que $R(E) \subseteq Q(C)$.
-Un problema de instanciación de flujo de trabajo es una
-tupla $\tup{S,W,\{V_i\}_i}$ donde $S$ es un conjunto de predicados
-que representa servicios
-abstractos, $W$ es un flujo de trabajo sobre $S$, y $\{V_i\}_i$ una colección de vistas
-que define servicios concretos. Suponemos trabajo con problemas \emph{seguros}
-en el sentido en que todas las variables mencionadas en la cabeza del flujo de
-trabajo (respectivamente en la cabeza de cada vista) aparecen en el cuerpo del
-flujo de trabajo (respectivamente en el cuerpo de cada vista). Además, sólo
-lidiamos con WIPs sin predicados aritméticos dentro del flujo de trabajo o
-vistas. Una instanciación $R$ es \emph{válida} si para todos los catálogos $C=\tup{S,T}$ y
-extensiones $\{E_i\}_i$, $R(E)\subseteq Q(C)$. Una colección $\R$ de instanciaciones válidas es una
-solución si para todos los catálogos de servicio $C=\tup{S,T}$ y extensiones $\{E_i\}_i$, no
-existe $\R'$ tal que $\R(E)\subset\R'(E)\subseteq Q(C)$.

### 2010-01-27 tesis/propuesta/secciones/portada.tex
casi todas las correcciones del segundo lote de ma esther

@@
 en un problema de reescritura de consultas, del área de sistemas de integración.
-Entonces, construyendo sobre trabajo relacionado, se idea una codificación del
+Entonces, en base a enfoques reportados en la literatura, se propone una codificación del
 problema de instanciación de flujos de trabajo como una teoría lógica cuyos
 modelos están en correspondencia con las instanciaciones del flujo de trabajo, y
-los modelos mejor puntuados en correspondencia con instanciaciones óptimas del
+los modelos mejor calificados en correspondencia con instanciaciones óptimas del
 flujo de trabajo. Así, explotando propiedades conocidas de teorías lógicas en
@@
 instancias grandes como muestran resultados experimentales ya realizados, sino
-que también es correcta y completa dado que, estando basada en la lógica, es
-amigable al análisis formal.
-
+que también podemos verificar que es correcta y completa dado que, estando basada en la lógica, es
+fácilmente sujeta al análisis formal.
 \end{abstract}

### 2010-01-27 tesis/propuesta/secciones/solucion.tex
casi todas las correcciones del segundo lote de ma esther

@@
 
-\subsection{Arquitectura del Sistema}
+\subsection{Formalización}
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=.9\textwidth]{graphics/architecture}
-\caption{Arquitectura del Sistema}
-\label{fig:architecture}
-\end{figure}
+Consideramos catálogos de servicio de la forma $C=\tup{S,T}$ donde $S$ es un
+conjunto de predicados que representa servicios abstractos y $T=\{T_s\}_{s\in S}$ es una
+colección de tablas que representa el resultado de evaluar los servicios
+abstractos. En el contexto del WIP, un catálogo de servicio $C$ es una
+descripción virtual de la salida producida por flujos abstractos
+implementados por servicios concretos descritos como vistas. Un flujo de trabajo
+$W$ sobre $S$ es una consulta conjuntiva de la forma
 
-Se propone una arquitectura comprendida por un catálogo de descripciones de
-servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
-el Decodificador. La figura~\ref{fig:architecture} muestra la arquitectura global del sistema.
-En esta infraestructura, una instancia del problema de instanciación de flujos
-de trabajo se define como un flujo abstracto representado por una consulta
-conjuntiva sobre servicios abstractos que es dada como entrada junto con un
-conjunto de servicios concretos definidos por vistas de servicios abstractos.
+\[ W(\vvec{x})\ \qrule\ s_1(\vvec{x}_1),\,s_1(\vvec{x}_2),\,\ldots,\,s_m(\vvec{x}_m)\,, \]
 
-El catálogo se carga con descripciones de servicios abstractos y concretos; cada
-servicio es descrito en términos de atributos de entrada y salida y anotado con
-un valor real que representa la utilidad QoS del servicio. La descripción de los
-servicios concretos, que son definidos como vistas de los servicios abstractos,
-puede ser generada de manera semiautomática o automática usando herramientas
-tales como el sistema DEIMOS \cite{AmbiteISWC09}.
+donde $s_i\in S$, $\vvec{x}$ es un vector de variables y cada $\vvec{x}_i$ es un vector de variables y
+constantes. El resultado de $W$ sobre $C$, denotado como $W(C)$, es la tabla con
+ $|\vvec{x}|$ columnas que resulta de la proyección del \emph{join} relacional
+$\join\!\!\{T_{s_i}\}_{i=1}^n$ sobre $\vvec{x}$. Los átomos en el cuerpo de $W$ son llamados los
+(sub)objetivos de $W$, y las variables en la cabeza de $W$ son llamadas
+distinguidas.
 
-Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
-corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
-compilador c2d, un componente \emph{off-the-shelf}, compila la fórmula CNF a
-d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
-convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
-parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
-importante remarcar que el proceso de compilación necesita ser realizado sólo
-una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
-compilación resulta ser costosa en términos de tiempo, este costo puede ser
-amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
-instanciaciones con respecto a múltiples valores de los parámetros QoS.
-Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscador
-a una instanciación de flujo de trabajo que resuelve el WIP.
+Un servicio concreto es descrito como una vista $V$ sobre $C$ que, siguiendo
+LAV, es una consulta sobre $S$. Dado un catálogo $C$, un flujo de trabajo $W$ y
+una colección de $n$ vistas $E=\tup{\{V_i\}_{i=1}^n,\{E_i\}_{i=1}^n}$, se pide
+conseguir todas las tuplas en
+$W(C)$ que puedan ser obtenidas de vistas en $E$. Es decir, se necesitan
+conseguir las \emph{instanciaciones}
 
-Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
-todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
-una manera libre de \emph{backtracking} todas las instanciaciones del flujo de
-trabajo. Si el usuario está interesado en una mejor instanciación dados
-parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamaño
-del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
-estas pueden ser computadas en tiempo lineal en su número. Finalmente, si el
-usuario está interesado en todas las instanciaciones, estas también pueden ser
-computadas en tiempo lineal en su número. En los últimos dos casos, si ese
-número es exponencial (en el tamaño de la entrada), la enumeración de las
-instanciaciones también lo es pero esta complejidad es intrínsica al problema y
-por lo tanto no puede ser evitada.
+\[ R(\vvec{x})\ \qrule\ V_{i_1}(\vvec{x}_1),\,V_{i_2}(\vvec{x}_2),\,\ldots,\,V_{i_{m_i}}(\vvec{x}_{m_i}) \]
+
+tales que $R(E) \subseteq Q(C)$.
+Un problema de instanciación de flujo de trabajo es una
+tupla $\tup{S,W,\{V_i\}_i}$ donde $S$ es un conjunto de predicados
+que representa servicios
+abstractos, $W$ es un flujo de trabajo sobre $S$, y $\{V_i\}_i$ una colección de vistas
+que define servicios concretos. Suponemos trabajo con problemas \emph{seguros}
+en el sentido en que todas las variables mencionadas en la cabeza del flujo de
+trabajo (respectivamente en la cabeza de cada vista) aparecen en el cuerpo del
+flujo de trabajo (respectivamente en el cuerpo de cada vista). Además, sólo
+lidiamos con WIPs sin predicados aritméticos dentro del flujo de trabajo o
+vistas. Una instanciación $R$ es \emph{válida} si para todos los catálogos $C=\tup{S,T}$ y
+extensiones $\{E_i\}_i$, $R(E)\subseteq Q(C)$. Una colección $\R$ de instanciaciones válidas es una
+solución si para todos los catálogos de servicio $C=\tup{S,T}$ y extensiones $\{E_i\}_i$, no
+existe $\R'$ tal que $\R(E)\subset\R'(E)\subseteq Q(C)$.
 
@@
 $\ell=v_i$, y $r(\ell)=0$ if $\ell\notin\{v_1,\ldots,v_n\}$.
+
+\subsection{Arquitectura del Sistema}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=.9\textwidth]{graphics/architecture}
+\caption{Arquitectura del Sistema}
+\label{fig:architecture}
+\end{figure}
+
+Se propone una arquitectura comprendida por un catálogo de descripciones de
+servicio, el Codificador, el compilador c2d, el Buscador de mejores modelos, y
+el Decodificador. La figura~\ref{fig:architecture} muestra la arquitectura global del sistema.
+En esta infraestructura, una instancia del problema de instanciación de flujos
+de trabajo se define como un flujo abstracto representado por una consulta
+conjuntiva sobre servicios abstractos que es dada como entrada junto con un
+conjunto de servicios concretos definidos por vistas de servicios abstractos.
+
+El catálogo se carga con descripciones de servicios abstractos y concretos; cada
+servicio es descrito en términos de atributos de entrada y salida y anotado con
+un valor real que representa la utilidad QoS del servicio. La descripción de los
+servicios concretos, que son definidos como vistas de los servicios abstractos,
+puede ser generada de manera semiautomática o automática usando herramientas
+tales como el sistema DEIMOS \cite{AmbiteISWC09}.
+
+Una instancia de entrada de WIP es codificada como una teoría CNF cuyos modelos
+corresponden a las intanciaciones del flujo de trabajo por el Codificador. El
+compilador c2d, un componente \emph{off-the-shelf}, compila la fórmula CNF a
+d-DNNF. El Codificador traduce las instancias WIP en teorías CNF que luego son
+convertidas en d-DNNF usando c2d. El Buscador computa un mejor modelo dados los
+parámetros QoS en tiempo lineal en el tamaño del d-DNNF resultante. Es
+importante remarcar que el proceso de compilación necesita ser realizado sólo
+una vez ya que no depende del valor de los parámetros QoS. Así, incluso si la
+compilación resulta ser costosa en términos de tiempo, este costo puede ser
+amortizado dado que el d-DNNF resultante puede ser usado para conseguir mejores
+instanciaciones con respecto a múltiples valores de los parámetros QoS.
+Finalmente, el Decodificador traduce el mejor modelo retornado por el Buscador
+a una instanciación de flujo de trabajo que resuelve el WIP.
+
+Dado un CNF que codifica un WIP, su d-DNNF es una representación compacta de
+todas las instanciaciones del flujo de trabajo. Es decir, uno puede generar de
+una manera libre de \emph{backtracking} todas las instanciaciones del flujo de
+trabajo. Si el usuario está interesado en una mejor instanciación dados
+parámetros de QoS, entonces esta se puede computar en tiempo lineal en el tamaño
+del d-DNNF. Si el usuario está interesado en todas las mejores instanciaciones,
+estas pueden ser computadas en tiempo lineal en su número. Finalmente, si el
+usuario está interesado en todas las instanciaciones, estas también pueden ser
+computadas en tiempo lineal en su número. En los últimos dos casos, si ese
+número es exponencial (en el tamaño de la entrada), la enumeración de las
+instanciaciones también lo es pero esta complejidad es intrínsica al problema y
+por lo tanto no puede ser evitada.
+
+\subsection{Resultados Iniciales}
+
+Se realizó un análisis empírico sobre tres experimentos. Todos los experimentos
+fueron realizados en una máquina de escritorio con un CPU Intel Core 2 Duo de
+2GHz y 4GB de memoria, y el tiempo fue medido con el comando de Unix
+\emph{time}.
+
+El objetivo de los experimentos es determinar el rendimiento de la propuesta con
+condiciones variantes. El principal beneficio de la solución propuesta es
+que se puede compilar la teoría lógica para una instancia del problema y luego
+calcular todas las instanciaciones, o las mejores, cualquier número de veces. El
+modelo de costos para conseguir mejores instanciaciones puede ser cambiado sin
+necesidad de recompilar la teoría lógica. Por lo tanto, la complejidad en tiempo
+de la solución es básicamente el tiempo para codificar el WIP como un
+CNF más el tiempo para compilar el CNF en un d-DNNF y el tiempo para decodificar
+los modelos. Los tiempos para codificar y decodificar son despreciables
+comparados con el tiempo para compilar el CNF. Por esto, el enfoque es en el tiempo
+necesario para compilar los problemas de los experimentos.
+
+El primer experimento consiste de problemas para consultas para viajes aéreos.
+Los servicios concretos son de la forma $V_i(x,y)\qrule\ \vuelo(x,y,\AL_i)$
+donde $\AL_i$ es una
+constante que denota el nombre de una aerolínea, y $\vuelo(x,y,\AL_i)$ relaciona las
+ciudades $x$ e $y$ tales que hay un vuelo entre $x$ e $y$ servido por $\AL_i$.
+Se supone que este servicio concreto retorna todos los vuelos entre dos ciudades
+con una aerolínea específica. El flujo de trabajo tiene la forma
+
+\[ W(x_1,\ldots,x_n)\ \qrule\ \vuelo(\PA,x_1,z),\,\vuelo(x_1,x_2,z),\,\ldots,\,\vuelo(x_n,\NY,z)\,. \]
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=1\textwidth]{graphics/plot1}
+\caption{Tiempos de compilación para los experimentos I y II para diferentes
+números de objetivos y de vistas. Los gráficos están en escala logarítmica, y el
+tiempo es en segundos.}
+\label{fig:plot1}
+\end{figure}
+
+El experimento incluye instancias para flujos de trabajo con 2 a 5 subobjetivos
+y conjuntos de 10 a 100 servicios concretos. Los resultados para la compilación
+son mostrados en el panel (a) de la figura ~\ref{fig:plot1}. Este es un gráfico en escala
+logarítmica que sugiere comportamiento sub-exponencial. En cualquier caso, los
+resultados muestran buen rendimiento dado que instancias realísticas del
+problema (conjuntos de 100 aerolíneas con vuelos de cinco paradas) pueden ser
+compiladas en 328 segundos. El tamaño en disco del d-DNNF para 100 aerolíneas y
+vuelos con cinco paradas es 3,4MB. En este d-DNNF, el mejor modelo puede ser
+computado en 0.29 segundos, y la enumeración de todos los modelos en 0.47
+segundos.
+
+En un intento de inducir crecimiento exponencial en el tamaño de compilación, en
+el segundo experimento se agrega un segundo servicio concreto para cada
+aerolínea. Esta modificación incrementa el número de instanciaciones válidas de
+lineal a exponencial dado que cada tramo de un vuelo puede ser ahora instanciado
+por dos servicios concretos y entonces un vuelo con $n$ tramos puede tener hasta
+$2^n$ instanciaciones. Se corrió el compilador para instancias que comprendían
+el mismo número de objetivos del flujo de trabajo y el número total de
+servicios concretos. Los resultados graficados en escala logarítmica se muestran
+en el panel (b) de la figura ~\ref{fig:plot1}.
+
+Estas pruebas muestran buen rendimiento para este tipo de problemas, pero no
+involucran servicios concretos con múltiples objetivos. Por lo tanto se
+diseñó un tercer experimento que consiste de instancias no estructuradas y
+generadas aleatoriamente. Cada instancia contiene tres variables por servicio
+abstracto, diez variables distintas y diez constantes distintas, seis
+objetivos en los flujos de trabajo, 2 a 5 objetivos en los servicios
+concretos, y número de servicios variante. La probabilidad de que un argumento
+de un servicio abstracto esté ligado a una constante es 50\%. Los resultados se
+muestran en la figura ~\ref{fig:plot3}. El tiempo de compilación de estas instancias no
+crece monótonamente dado que son generadas aleatoriamente. Lo mismo ocurre para
+el tamaño de las teorías y los números de modelos. Por ejemplo, el d-DNNF para
+un problema con 45 vistas cada una con 5 objetivos fue de tamaño 5,1Mb y tuvo
+$1.26\times 10^8$ modelos. El tiempo para buscar el mejor modelo para este
+d-DNNF es 0.46 segundos mientras que el tiempo para enumerar todos los modelos
+es alredededor de 17 horas.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=.8\textwidth]{graphics/plot3}
+\caption{Tiempos de compilación para el experimento III para diferentes números
+de objetivos y de vistas. Los gráficos están en escala logarítmica, y el
+tiempo es en segundos.}
+\label{fig:plot3}
+\end{figure}
+
+Estos son experimentos preliminares, pero los resultados muestran que la
+solución propuesta escala eficientemente para problemas con varios objetivos
+y vistas. Se cree que estos resultados son alentadores y motivan a continuar
+esta línea de investigación.

### 2010-01-27 tesis/propuesta/secciones/trabajorelacionado.tex
casi todas las correcciones del segundo lote de ma esther

\section{Trabajo Relacionado}

En esta sección se resumen los enfoques existentes que proveen soluciones
a los problemas de selección de servicios y reescritura de consultas y se
discute el trabajo relacionado en el área de inteligencia artificial llamado
compilación del conocimiento.

\subsection{Selección de Servicios}

El problema de seleccionar los servicios que implementan un flujo de trabajo
abstracto y cumplen mejor los criterios basados en QoS es conocido como el
problema de selección o composición de servicios consciente de QoS, que ha sido
demostrado ser \emph{NP-hard} \cite{Hiroshi2008}. Este es un problema de optimización
combinatoria y varias heurísticas han sido propuestas para conseguir una
solución relativamente buena en un período razonable de tiempo.

Rahmani et al. \cite{rahmani08} presentan una heurística basada en una métrica de distancia
que guía un algoritmo de búsqueda; esta métrica induce un orden de
los servicios en una manera en que los nodos vertedero (sin salida) tienen baja probabilidad
de ser visitados. En una serie de artículos, Berardi y otros \cite{berardi05,berardi08,berardi06}
describen
servicios y flujos de trabajo en términos de máquinas de estados finitos
determinísticas que son codificadas usando teorías de Lógicas Descriptivas
cuyos modelos corresponden a soluciones del problema. Aunque se podrían explotar
métodos para formalismos de Lógicas Descriptivas, no han sido reportados la
escalabilidad o el rendimiento de la solución propuesta.

Ko et al. \cite{myoung08} proponen un enfoque basado en restriccciones que codifica
los valores permisibles no funcionales como un conjunto de restricciones cuya
violación debe ser minimizada; para recorrer el espacio de soluciones
posiblemente óptima, se implementa un algoritmo híbrido que combina las
metaheurísticas \emph{tabu search} y \emph{simulated annealing}. Los resultados
experimentales muestran que la solución propuesta es capaz de escalar a un
número grande de servicios y procesos abstractos. Cardellini et al. \cite{cardellini07}
codifican una parte del problema de composición de servicios caracterizados por QoS
como un problema de programación lineal \cite{cardellini07}. Por otra parte, Wada et al.
\cite{Hiroshi2008} trata el problema como uno de optimización con múltiples objetivos donde
los diferentes parámetros de QoS son considerados igualmente importantes en vez
de ser agregados en una función. Luego se propone un algoritmo basado en
algoritmos genéticos que identifica un conjunto de composiciones de servicios
no dominadas que mejor satisfacen los requerimientos QoS.

Alrifai y Risse \cite{alrifaiR09} proponen una solución de dos fases que usa un
algoritmo de programación entera para obtener la descomposición de QoS
globales en restricciones locales y luego selecciona los servicios que mejor
cumplen las restricciones locales.

Recientemente, dos soluciones basadas en planificación han sido propuestas.
Kuter y Golbeck \cite{kuterG09} extienden el algoritmo de planificación SHOP2 para
seleccionar la composición confiable de servicios que implementan un modelo de
procesos OWL-S dado, mientras que Sohrabi y McIlraith \cite{sohrabiM09} proponen una
solución basada en planificación HTN donde las métricas de preferencia del
usario y regulaciones de dominio son utilizadas para guíar el planificador hacia
el espacio de composiciones relevantes. Finalmente, Lécué \cite{lecue09} propone un
algoritmo basado en genética para identificar la composición de servicios que
mejor cumple con los criterios de calidad para un conjunto de parámetros QoS.

Aunque estas soluciones son capaces de resolver el problema de optimización y
escalar a un número de procesos abstractos, ninguno de ellos está ajustado para
describir semánticamente los servicios en términos de procesos abstractos, ni
para usar estas descripciones para identificar los servicios que implementan un
flujo de trabajo dado o que mejor cumplen los criterios no funcionales del
usuario.

\subsection{Reescritura de Consultas}

Un número de algoritmos han sido desarrollados para encontrar las reescrituras
de una consulta dada; los más prominentes son el algoritmo \emph{bucket} \cite{levy:bucket}, el
algoritmo de reglas inversas \cite{pods:DuschkaG97,Qian96}, el algoritmo
\minicon \cite{pottinger:minicon}, y el
algoritmo \mcdsat \cite{arvelo:aaai06}. Generalmente, la reescritura de consultas trabaja en
dos fases. Durante la primera, el algoritmo identifica las vistas que reescriben
al menos un subobjetivo de la consulta, y durante el segundo estas reescrituras
parciales son combinadas en reescrituras completas. La diferencia principal
entre los algoritmos es el conjunto de criterios usado para elegir las vistas
relevantes para reducir el espacio de reescrituras no útiles.

El algoritmo denominado \emph{bucket} reduce el número de posibilidades considerando sólo cada
subobjetivo de la consulta aislado, y seleccionando las vistas que son capaces
de producir al menos los atributos proyectados por la consulta. Como los
atributos involucrados en los \emph{joins} de la consulta no son verificados, un número
grande de reescrituras comprendido por productos cartesianos puede ser generado.

El algoritmo de reglas inversas construye un conjunto de reglas que invierten
las definiciones de vistas y establecen cómo computar tuplas para las relaciones
de la base de datos a partir de las tuplas de las vistas. Al igual que el algoritmo
bucket, puede producir un gran número de reescrituras no útiles.

El algoritmo \minicon\ supera las limitaciones de los algoritmos previos
identificando solamente vistas que reescriben un conjunto de los subobjetivos de
la consulta, y que pueden ser combinados con el resto de los subobjetivos. La
idea principal es identificar las asociaciones entre variables de cada
subobjetivo a las variables en uno o más subojbetivos en las vistas, de manera
que las variables de \emph{join} de la consulta son asociadas a las variables de
\emph{join}
del cuerpo de una vista o a variables distinguidas de la vista. Asociaciones
entre variables y subojbetivos son representadas en las llamadas \emph{MiniCon
Descriptions} (MCDs) \cite{pottinger:minicon}.

Finalmente, el algoritmo \mcdsat\ es capaz de identificar las reescrituras de una
consulta traduciendo el problema de reescritura al problema de enumeración de
modelos de una teoría proposicional cuyos modelos están en correspondencia con
las reescrituras de la consulta. El algoritmo explota las propiedades de los
d-DNNF para enumerar eficientemente los modelos de la teoría.
El algoritmo
\mcdsat\ ha demostrado escalar mejor que el algoritmo \minicon\ sobre un número
grande de experimentos donde frecuentemente muestra mejoras de rendimiento de
varios órdenes de magnitud. Sin embargo, el algoritmo \mcdsat\ no fue diseñado
para problemas de reescritura que involucren constantes explícitas, ni para
computar las mejores reescrituras respecto a una función de utilidad o modelo de
costo dado. En este trabajo se propone una codificación extendida que supera
estas limitaciones y se aplica la codificación al problema de instanciación de
flujos de trabajo.

\subsection{Compilación del Conocimiento}

Compilación del conocimiento es el área en inteligencia artificial que se ocupa
del problema de traducir teorías lógicas en fragmentos apropiados que hagan
tratables ciertas operaciones deseadas \cite{cadoli:compilation}. Diferentes lenguajes de
compilación han sido definidos. Por ejemplo están los \emph{Ordered
Binary Decision Diagrams} (OBDDs) \cite{bryant:obdd},
\emph{Negation Normal Form} (NNF) \cite{barwise:handbook}, y
\emph{Decomposable Negation Normal Form} (DNNF) \cite{darwiche:map}. En este trabajo se usan las
propiedades de los DNNF determinísticos (d-DNNF) \cite{darwiche:d-dnnfs} para proveer una
solución eficiente y escalable al problema de selección de servicios.

\begin{figure}[t]
\centering
\includegraphics[width=.7\textwidth]{graphics/odd}
\caption{Un NNF descomponible y determinístico.}
\label{fig:dnnf}
\end{figure}

Una teoría en Negation Normal Form es construida desde literales usando sólo
conjunciones y disjunciones \cite{barwise:handbook}, y puede ser representada como un grafo
acíclico dirigido en que las hojas son etiquetadas con literales y los nodos
internos con $\land$ y $\lor$; ver ejemplo en la figura~\ref{fig:dnnf}. Esta
forma normal es universal, lo que significa que para cada fórmula lógica hay una
equivalente en formato NNF. Se dice que un NNF es descomponible (DNNF)
\cite{darwiche:d-dnnfs}
si para cada conjunción $\bigwedge_i\phi_i$, sus variables son disjuntas por pares; es
decir, $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ para $i\neq j$. Un DNNF soporta un
número de operaciones en tiempo polinomial en el tamaño de su GAD. Por ejemplo,
podemos probar si un DNNF es satisfactible haciendo un sólo recorrido 
sobre su GAD en tiempo linear. Se dice que un DNNF es
determinístico (d-DNNF) \cite{darwiche:d-dnnfs} si para cada disjunción $\bigvee_i\phi_i$, los
disjuntos son lógicamente contradictorios por pares, es decir,
$\phi_i\land\phi_j\equiv\textbf{false}$ para $i\neq j$.
El NNF en la figura~\ref{fig:dnnf}, por ejemplo, es
descomponible y determinístico. Un d-DNNF soporta conteo de modelos en
tiempo polinomial en el tamaño de su GAD, y enumeración de modelos en tiempo
polinomial en el tamaño de la salida. Además, dada una función de calidad de
literales $r$, se puede computar la calidad del mejor modelo en tiempo
polinomial para DNNFs \cite{darwiche:weighted}.

Toda teoría puede representarse en las formas DNNF y d-DNNF pero traducir una teoría CNF a
formato DNNF tiene costo exponencial en el peor caso. Esta traducción se
llama compilación en este campo. Hay un compilador públicamente disponible,
llamado c2d\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}, que realiza este proceso de compilación y que hace uso
de técnicas de SAT modernas como backtracking dirigido por conflictos,
aprendizaje de cláusulas y caching \cite{darwiche:compiler}. Este compilador incurre en el peor
caso (peor gasto) en espacio exponencial en un parámetro llamado el ancho del árbol de
descomposición que está relacionado a la ``conectividad'' de la teoría CNF.
Sin embargo, en nuestros experimentos, las teorías CNF que son compiladas son de
ancho pequeño.

### 2010-01-27 tesis/propuesta/secciones/cronograma.tex
cronograma corregido

@@
 \hline
-Experimentos sobre método \emph{Branch and Bound} & 3 semanas & 8 de marzo de 2010 \\
+Experimentos sobre método \emph{Branch and Bound} & 3 semanas & 5 de marzo de 2010 \\
+\hline
+Implementación de soporte para clases de constantes & 2 semanas & 12 de marzo de 2010 \\
+\hline
+Pruebas de soporte para clases de constantes & 1 semana & 19 de marzo de 2010 \\
 \hline

### 2010-01-27 tesis/propuesta/secciones/cronograma.tex
traducido a espanol graficos, justif. lista

@@
 \hline
-Implementación de sistema de eliminación de cláusulas para uso de \emph{Branch and Bound} & 5 semanas & 12 de febrero de 2010 \\
+Implementación de sistema para problemas de tamaño mayor & 5 semanas & 12 de febrero de 2010 \\
 \hline
-Experimentos sobre método \emph{Branch and Bound} & 3 semanas & 5 de marzo de 2010 \\
+Experimentos de sistema para problemas de tamaño mayor & 3 semanas & 5 de marzo de 2010 \\
 \hline

### 2010-01-27 tesis/propuesta/secciones/cronograma.tex
blargh

--- 2010-01-27 tesis/propuesta/secciones/cronograma.tex
@@
 \hline
-Experimentos con generación de cláusulas nuevas & 1 semana & 04 de diciembre de 2009 \\
+Experimentos con generación de cláusulas nuevas & 3 semanas & 18 de diciembre de 2009 \\
 \hline
-Implementación de búsqueda de mejor reescritura & 2 semanas & 11 de diciembre de 2009 \\
+Implementación de búsqueda de mejor reescritura & 2 semanas & 8 de enero de 2010 \\
 \hline
-Experimentos con búsqueda de mejor reescritura & 1 semana & 18 de diciembre de 2009 \\
+Experimentos con búsqueda de mejor reescritura & 2 semana & 22 de enero de 2010 \\
 \hline
-Implementación de sistema de eliminación de cláusulas para uso de \emph{Branch and Bound} & 5 semanas & 12 de febrero de 2010 \\
+Implementación de sistema para problemas de tamaño mayor & 5 semanas & 26 de febrero de 2010 \\
 \hline
-Experimentos sobre método \emph{Branch and Bound} & 3 semanas & 8 de marzo de 2010 \\
+Experimentos de sistema para problemas de tamaño mayor & 3 semanas & 19 de marzo de 2010 \\
+\hline
+Implementación de soporte para clases de constantes & 2 semanas & 2 de abril de 2010 \\
+\hline
+Pruebas de soporte para clases de constantes & 2 semana & 16 de abril de 2010 \\
 \hline

### 2010-02-18 tesis/src/models/filtermodels.cc
filtermodels

@@ -1,0 +1,193 @@
+/*
+ *  Copyright (C) 2006 Universidad Simon Bolivar
+ * 
+ *  Permission is hereby granted to distribute this software for
+ *  non-commercial research purposes, provided that this copyright
+ *  notice is included with any such distribution.
+ *  
+ *  THIS SOFTWARE IS PROVIDED "AS IS" WITHOUT WARRANTY OF ANY KIND,
+ *  EITHER EXPRESSED OR IMPLIED, INCLUDING, BUT NOT LIMITED TO, THE
+ *  IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
+ *  PURPOSE.  THE ENTIRE RISK AS TO THE QUALITY AND PERFORMANCE OF THE
+ *  SOFTWARE IS WITH YOU.  SHOULD THE PROGRAM PROVE DEFECTIVE, YOU
+ *  ASSUME THE COST OF ALL NECESSARY SERVICING, REPAIR OR CORRECTION.
+ *  
+ *  Blai Bonet, bonet@ldc.usb.ve
+ *
+ */
+
+#include <nnf.h>
+#include <satlib.h>
+#include <utils.h>
+
+#include <string.h>
+
+#include <iostream>
+#include <fstream>
+
+void
+banner( std::ostream &os )
+{
+  os << "Model enumerator for d-DNNF theories." << std::endl
+     << "Universidad Simon Bolivar, 2006." << std::endl;
+}
+
+void
+usage( std::ostream &os )
+{
+  os << "usage: models [--mp] [--literal-counts] [--num <n>] [--verbose] [--write-models] [--write-all-literals] <nnf>" << std::endl;
+}
+
+void
+read_simple_file( std::istream &is, std::vector<int> &contents )
+{
+  size_t n;
+  is >> n;
+  contents.clear();
+  contents.reserve(n);
+  for( size_t i = 0; i < n; ++i ) {
+    int entry;
+    is >> entry;
+    contents.push_back(entry);
+  }
+}
+
+int
+main( int argc, char **argv )
+{
+  const char *forget_file = 0;
+  float i_seconds, l_seconds, seconds;
+  bool use_mp = false;
+  bool literal_counts = false;
+  bool all = false;
+  bool output = false;
+  bool verbose = false;
+  size_t count = 0;
+  banner(std::cout);
+
+  ++argv;
+  --argc;
[…]

### 2010-02-18 tesis/src/imcdsat/mcdsat/CQ/CQ.py
cambios varios

@@ -24,6 +24,7 @@ class CQ:
         self.comparacion = comp
         self.vars,self.varsExist,self.varsDist = self.obtVariables();
         self.varsExistL = list(self.varsExist)
+        self.string = None
 
     def variables(self):
         return self.vars
@@ -57,6 +58,28 @@ class CQ:
 
     def __str__(self):
         return str(self.cabeza) + " :- " + ", ".join(str(x) for x in self.cuerpo)
+        if not self.string:
+            display_cabeza = self.cabeza
+            display_cuerpo = self.cuerpo
+            
+            display_cabeza.orden = []
+            display_cabeza.argumentos = {}
+
+            for v in self.cabeza.orden:
+                if es_const(v):
+                    nv = getConstantName(v)
+
+                    display_cabeza.orden.append(nv)
+                    display_cabeza.argumentos[nv] = '1'
+                else:
+                    nv = getVariableName(v)
+
+                    display_cabeza.orden.append(nv)
+                    display_cabeza.argumentos[nv] = '0'
+
+            self.string = str(display_cabeza) + " :- " + ", ".join(str(x) for x in display_cuerpo)
+
+        return str(self.cabeza) 
 
     __repr__ = __str__
 

### 2010-02-18 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py

@@ -121,7 +121,7 @@ def clausulasCombinarMCD(transf, n, q, vistas):
                                 except KeyError:
                                     continue
 
-                                #clausulas para constantes en teoria extendida
+                                #clausulas para constantes en teoria extendida (S6)
                                 clausulas2.append("%s %s 0\n" % (str(numVarT0), str(numVarT1)))
 
 
@@ -178,50 +178,51 @@ def generarTeoriaMCD(q, vistas):
     
     c5, c13 = clausulas513(q,vistas,ltaux)
     c15 = clausulas15(q, vistas)
-    #print "clausulas 1  \/ vi (por lo menos uno)"
+    #print "clausulas C1  \/ vi (por lo menos uno)"
     #pprint.pprint(c1) 
-    #print "clausulas 2  -vi \/ -vj (maximo uno)"
+    #print "clausulas C2  -vi \/ -vj (maximo uno)"
     #pprint.pprint(c2) 
-    #print "clausulas 3  \/ gk (por lo menos uno)"
+    #print "clausulas C3 y C4 \/ gk (por lo menos uno)"
     #pprint.pprint(c3) 
-    #print "clausulas 4  Vm /\ tij => -tik y property 1 C2 "
+    #print "clausulas C9 Vm /\ tij => -tik y property 1 C2 "
     #pprint.pprint(c4) 
-    #print "clausulas 5  vm => -tij (i Dist y j exist) "
+    #print "clausulas C10 vm => -tij (i Dist y j exist) "
     #pprint.pprint(c5)
-    #print "clausulas 6  gi /\ vm => \/ zir (r subob de Vm)"
+    #print "clausulas C7 (=>) gi /\ vm => \/ zir (r subob de Vm)"
     #pprint.pprint(c6) 
-    #print "clausulas 7  zir => tir"
+    #print "clausulas C12 zir => tir"
     #pprint.pprint(c7) 
-    #print "clausulas 8  gi /\ vm <= \/ zir (r subob de Vm)"
+    #print "clausulas C7 (<=) gi /\ vm <= \/ zir (r subob de Vm)"
     #pprint.pprint(c8)
-    #print "clausulas 9  maximo una z por vm, gi"
+    #print "clausulas C5 maximo una z por vm, gi"
     #pprint.pprint(c9)
-    #print "clausulas 10 t explicito"
[…]

### 2010-02-18 tesis/propuesta/secciones/planteamiento.tex

@@
 cumplidas, el Problema de Instanciación de Flujos de Trabajo (WIP).
-
-En este trabajo se considera una versión restringida de WIP que adopta el
-enfoque Local-As-View (LAV) \cite{levy:bucket}. En LAV, todos los elementos de un
-problema son especificados con un lenguaje común basado en servicios abstractos
-tal que los servicios concretos se describen como vistas de servicios
-abstractos, la calidad de una instanciación como una función de utilidad global
-que combina los diferentes parámetros QoS, y el flujo de trabajo abstracto como
-una consulta conjuntiva sobre los servicios abstractos; esta representación es
-similar a la que es generada de manera semiautomática para el sistema DEIMOS
-\cite{AmbiteISWC09}. En esta versión de WIP, las reglas que definen flujos de trabajo
-(consultas conjuntivas) y servicios concretos (vistas) son creados de manera que
-todas las restricciones funcionales sobre pre- y post-condiciones de los
-servicios y sus combinaciones sean satisfechas, y las medidas de QoS son
-representadas anotando cada descripción de servicio concreto con un número real
-que representa la utilidad QoS global del servicio.
 
@@
 que minimiza (o maximiza) la agregación de utilidades.
+
+\subsection{Propuesta}
+
+En este trabajo se considera una versión restringida de WIP que adopta el
+enfoque Local-As-View (LAV) \cite{levy:bucket}. En LAV, todos los elementos de un
+problema son especificados con un lenguaje común basado en servicios abstractos
+tal que los servicios concretos se describen como vistas de servicios
+abstractos, la calidad de una instanciación como una función de utilidad global
+que combina los diferentes parámetros QoS, y el flujo de trabajo abstracto como
+una consulta conjuntiva sobre los servicios abstractos; esta representación es
+similar a la que es generada de manera semiautomática para el sistema DEIMOS
+\cite{AmbiteISWC09}. En esta versión de WIP, las reglas que definen flujos de trabajo
+(consultas conjuntivas) y servicios concretos (vistas) son creados de manera que
+todas las restricciones funcionales sobre pre- y post-condiciones de los
+servicios y sus combinaciones sean satisfechas, y las medidas de QoS son
+representadas anotando cada descripción de servicio concreto con un número real
+que representa la utilidad QoS global del servicio.
+
+El resto de esta propuesta está distribuido como sigue. En la sección 2 se
+discute trabajo previo relacionado a la propuesta y conseguido en la literatura
+actual. En la sección 3 se justifica el trabajo sobre este problema. En la
+sección 4 se exponen los objetivos del trabajo. En la sección 5 se propone una
+metodología para alcanzar los objetivos. En la sección 6 se presenta un
+cronograma de actividades que permitan desarrollar el trabajo. Por último, en la
+sección 7 se dan a conocer los resultados del trabajo ya realizado.

### 2010-02-18 tesis/propuesta/secciones/portada.tex

@@
 \begin{flushleft}
-\includegraphics[scale=0.3]{graphics/logo_usb.png} \\
+\includegraphics[scale=0.3]{graphics/logo_usb} \\
 Universidad Simón Bolívar \\

### 2010-02-19 tesis/src/eswc_demo/main.py
ultima version de eswc_demo para screenshots del paper

@@ -11,7 +11,7 @@ class MyForm(QMainWindow):
         self.ui.setupUi(self)
 
         queryDisplay = self.ui.CentralWidget.findChild(QLineEdit, "queryDisplayBox")
-        queryDisplay.setText("q(X1,X2) :- r(X1,X3),r(X3,X4),r(X4,X5),r(X5,X2)")
+        queryDisplay.setText("W(x, w, y, z) :- uscity(x), flight(x, w), flight(w, y), flight(y, z), flight(z, x)")
 
         listView = self.ui.CentralWidget.findChild(QListWidget, "listView")
 
@@ -25,7 +25,8 @@ class MyForm(QMainWindow):
         #parte de archivo TODO
         viewsDisplayBox = self.ui.CentralWidget.findChild(QListWidget, "viewsDisplayBox")
 
-        f = open("views-60-5.txt")
+        #f = open("views-60-5.txt")
+        f = open("views-paper.txt")
         lines = [l for l in f]
         f.close()
 

### 2010-02-19 tesis/src/eswc_demo/ui/main.py

@@ -1,0 +1,66 @@
+# -*- coding: utf-8 -*-
+
+# Form implementation generated from reading ui file 'ui/main.ui'
+#
+# Created: Sat Jan 23 20:28:22 2010
+#      by: PyQt4 UI code generator 4.6
+#
+# WARNING! All changes made in this file will be lost!
+
+from PyQt4 import QtCore, QtGui
+
+class Ui_Main(object):
+    def setupUi(self, Main):
+        Main.setObjectName("Main")
+        Main.resize(602, 402)
+        self.CentralWid = QtGui.QWidget(Main)
+        self.CentralWid.setObjectName("CentralWid")
+        self.splitter_2 = QtGui.QSplitter(self.CentralWid)
+        self.splitter_2.setGeometry(QtCore.QRect(0, 1, 601, 401))
+        self.splitter_2.setOrientation(QtCore.Qt.Horizontal)
+        self.splitter_2.setObjectName("splitter_2")
+        self.listView = QtGui.QListView(self.splitter_2)
+        sizePolicy = QtGui.QSizePolicy(QtGui.QSizePolicy.Expanding, QtGui.QSizePolicy.Expanding)
+        sizePolicy.setHorizontalStretch(0)
+        sizePolicy.setVerticalStretch(0)
+        sizePolicy.setHeightForWidth(self.listView.sizePolicy().hasHeightForWidth())
+        self.listView.setSizePolicy(sizePolicy)
+        self.listView.setObjectName("listView")
+        self.splitter = QtGui.QSplitter(self.splitter_2)
+        self.splitter.setOrientation(QtCore.Qt.Vertical)
+        self.splitter.setObjectName("splitter")
+        self.horizontalLayoutWidget = QtGui.QWidget(self.splitter)
+        self.horizontalLayoutWidget.setObjectName("horizontalLayoutWidget")
+        self.horizontalLayout = QtGui.QHBoxLayout(self.horizontalLayoutWidget)
+        self.horizontalLayout.setObjectName("horizontalLayout")
+        self.label = QtGui.QLabel(self.horizontalLayoutWidget)
+        self.label.setObjectName("label")
+        self.horizontalLayout.addWidget(self.label)
+        self.queryDisplay = QtGui.QLineEdit(self.horizontalLayoutWidget)
+        self.queryDisplay.setReadOnly(True)
+        self.queryDisplay.setObjectName("queryDisplay")
[…]

### 2010-02-19 tesis/src/eswc_demo/ui/prueba.py

@@ -100,7 +100,7 @@ class Ui_Main(object):
 
         # top :: right side :: solutions :: rewritings
         self.rewritings_Widget = QGroupBox()
-        self.rewritings_Widget.setTitle("Rewritings")
+        self.rewritings_Widget.setTitle("Instantiations")
         self.solutions_Layout.addWidget(self.rewritings_Widget)
 
         self.rewritings_Layout = QVBoxLayout()
@@ -145,11 +145,11 @@ class Ui_Main(object):
         QtCore.QMetaObject.connectSlotsByName(Main)
 
     def retranslateUi(self, Main):
-        Main.setWindowTitle(QApplication.translate("Main", "SsdSat Demo", None, QApplication.UnicodeUTF8))
+        Main.setWindowTitle(QApplication.translate("Main", "SATWins", None, QApplication.UnicodeUTF8))
         self.queryDisplayLabel.setText(QApplication.translate("Main", "Workflow", None, QApplication.UnicodeUTF8))
         self.viewsDisplayLabel.setText(QApplication.translate("Main", "Concrete Services", None, QApplication.UnicodeUTF8))
 
-        self.dnnfLabel.setText(QApplication.translate("Main", "d-DNNF: loaded", None, QApplication.UnicodeUTF8))
+        self.dnnfLabel.setText(QApplication.translate("Main", "d-DNNF: loaded (28K)", None, QApplication.UnicodeUTF8))
         #TODO activate next label
         #self.dnnfStatusLabel.setText(QApplication.translate("Main", "loaded", None, QApplication.UnicodeUTF8))
 

### 2010-02-19 tesis/src/models/nnf/nnf.cc
filtermodels

@@ -279,6 +279,42 @@ Manager::enumerate_models( std::ostream &os, size_t count, bool all ) const
 }
 
 void
+Manager::enumerate_and_project_models( std::ostream &os, size_t count, bool all, const std::vector<int>& forgets ) const
+{
+  Model m;
+  bool next = true;
+  size_t i = 0;
+  while( next && ((count == 0) || (i < count)) ) {
+    // generate model
+    std::pair<bool,const Node*> rc = enumerate_models_recursively(root_,m,0);
+    if( rc.first ) {
+        for( int i = 0, isz = forgets.size(); i < isz; ++i ) {
+            int var = forgets[i]<<1;
+            m.erase(var);
+            m.erase(var+1);
+        }
+
+        m.print(os,all);
+        os << std::endl;
+        ++i;
+    }
+    m.clear();
+
+    // advance state
+    next = false;
+    for( const Node *n = rc.second; !next && (n != 0); n = (const Node*)n->cache_.second_ ) {
+      long i = 1+(long)n->cache_.first_;
+      if( n->children_[i] == 0 )
+        n->cache_.first_ = 0;
+      else {
+        n->cache_.first_ = (const void*)i;
+        next = true;
+      }
+    }
+  }
+}
+
+void
 Manager::enumerate_models( ModelList &models, size_t count ) const
 {
   Model m;
@@ -364,7 +400,7 @@ Manager::recursive_min_model( const Node *n, Model &model ) const
     recursive_min_model(*minchild,model);
   }
   else if( n->type_ == Value )
-    assert((long)n->children_);
+    0;//assert((long)n->children_);
   else if( n->type_ == Variable )
     model.insert((long)n->children_);
 }

### 2010-02-19 tesis/src/models/nnf/nnf.h

@@ -477,6 +477,7 @@ namespace nnf {
     mp::Int* mp_count_models( mp::Int **output = 0, const int *lit_map = 0 ) const;
     std::pair<bool,const Node*> enumerate_models_recursively( const Node *n, Model &m, const Node *last_or ) const;
     void enumerate_models( std::ostream &os, size_t count = 0, bool all = false ) const;
+    void enumerate_and_project_models( std::ostream &os, size_t count = 0, bool all = false, const std::vector<int>& forgets = std::vector<int>() ) const;
     void enumerate_models( ModelList &models, size_t count = 0 ) const;
     int min_cost( const int *costs = 0 ) const;
     void min_model( Model &model ) const;

### 2010-02-19 tesis/src/qrpgen/generators/structured.py
bug en el numero de sos generados en qrpgen

@@ -37,7 +37,7 @@ def randomviews(argv):
 
     for n in xrange(1, nviews+1):
         view_table = Table("v%s" % n, [Argument('var', i+1) for i in xrange(n_vars)])
-        view_sos = [Table(pred, [rand_constvar(i+1) for i in random.sample(range(n_vars)*n_vars, n_varsperpred)]) for pred in random.sample(predicates * n_preds, nquery_sos)]
+        view_sos = [Table(pred, [rand_constvar(i+1) for i in random.sample(range(n_vars)*n_vars, n_varsperpred)]) for pred in random.sample(predicates * n_preds, nview_sos)]
         view = Query(view_table, view_sos)
         views.append(view)
 

### 2010-02-19 tesis/src/qrpgen/qrpgen.py

@@ -17,7 +17,8 @@ def writelist(l, filename):
 
 def main():
     if len(sys.argv) < 4:
-        usage_and_quit()
+        print_usage()
+        sys.exit(1)
 
     viewsfilename = sys.argv[1]
     queryfilename = sys.argv[2]

### 2010-02-19 tesis/experimentos/5/runner.sh
experimentos 5-9

@@ -1,0 +1,30 @@
+#!/bin/bash
+
+INDIR=casos
+OUTFILE=experimento5-allrw.out
+
+echo '########' | tee -a ${OUTFILE}
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/mcdsat
+
+for nviews in `seq 55 5 80`; do
+    for nview_sos in `seq 2 5`; do
+        #casos largos
+        if [ $nviews = 40 -a $nview_sos = 5 ]; then
+            continue
+        fi
+        if [ $nviews = 45 -a $nview_sos = 4 ]; then
+            continue
+        fi
+
+        VIEWSFILE=${INDIR}/views-$nviews-$nview_sos-4-10-10-3-0.5.txt
+        QUERYFILE=${INDIR}/query-$nviews-$nview_sos-4-10-10-3-0.5.txt
+
+        echo ${nviews} ${nview_sos} 4 10 10 3 0.5 | tee -a ${OUTFILE}
+
+        for i in `seq 1`; do #n times to get average
+            ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+        done
+    done
+done

### 2010-02-19 tesis/experimentos/6/runner.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+INDIR=casos
+OUTFILE=experimento6.out
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
+
+echo '########' | tee -a ${OUTFILE}
+
+for vistas in `seq 5 5 40`; do
+    for pasos in `seq 2 5`; do
+        for copias in 2; do
+            VIEWSFILE=${INDIR}/views-${vistas}-${pasos}-${copias}.txt
+            QUERYFILE=${INDIR}/query-${vistas}-${pasos}-${copias}.txt
+
+            echo ${vistas} ${pasos} ${copias} | tee -a ${OUTFILE}
+
+            for i in `seq 1`; do #n times to get average
+                ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+            done
+        done
+    done
+done

### 2010-02-19 tesis/experimentos/7/runner.sh

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+INDIR=casos
+OUTFILE=experimento7.out
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
+
+echo '########' | tee -a ${OUTFILE}
+
+for vistas in `seq 100 10 100`; do
+    for pasos in `seq 2 4`; do
+        for copias in 2; do
+            VIEWSFILE=${INDIR}/views-${vistas}-${pasos}-${copias}.txt
+            QUERYFILE=${INDIR}/query-${vistas}-${pasos}-${copias}.txt
+
+            echo ${vistas} ${pasos} ${copias} | tee -a ${OUTFILE}
+
+            for i in `seq 1`; do #n times to get average
+                ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+            done
+        done
+    done
+done

### 2010-02-19 tesis/experimentos/8/runner.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+INDIR=casos
+OUTFILE=plot1.raw
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
+
+echo '################################################################################' | tee -a ${OUTFILE}
+
+for vistas in `seq 90 10 90`; do
+    for pasos in `seq 2 5`; do
+        VIEWSFILE=${INDIR}/views-${vistas}-${pasos}.txt
+        QUERYFILE=${INDIR}/query-${vistas}-${pasos}.txt
+
+        echo ${vistas} ${pasos} | tee -a ${OUTFILE}
+
+        for i in `seq 1`; do #n times to get average
+            ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+        done
+    done
+done

### 2010-02-19 tesis/experimentos/9/runner.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+INDIR=casos
+OUTFILE=experimento5-allrw.out
+
+echo '########' | tee -a ${OUTFILE}
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
+
+for nviews in `seq 10 5 80`; do
+    for nview_sos in `seq 2 5`; do
+        VIEWSFILE=${INDIR}/views-$nviews-$nview_sos-6-10-10-3-0.5.txt
+        QUERYFILE=${INDIR}/query-$nviews-$nview_sos-6-10-10-3-0.5.txt
+
+        echo ${nviews} ${nview_sos} 6 10 10 3 0.5 | tee -a ${OUTFILE}
+
+        for i in `seq 1`; do #n times to get average
+            ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+        done
+    done
+done

### 2010-03-26 coloquio/presentacion.tex
presentacion coloquio

\documentclass{beamer}

\usepackage[spanish]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{beamerthemesplit}

\title{Selección de Servicios Web de Máxima Utilidad Con Circuitos}
\author{Daniel Izquierdo}
\date{14 de diciembre de 2009}

\begin{document}

\frame{\titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section[Contenido]{}
\frame{\tableofcontents}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame
{
\section{Motivación}

\begin{itemize}
\item Explosión en número de servicios y fuentes de datos disponibles
\item Heterogeneidad de servicios concretos y fuentes
\item Diferencias en calidad de servicio existentes en el mundo real
    \item Necesidad de realizar consultas o \emph{workflows} bajo estas condiciones
\end{itemize}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame
{
\section{Objetivo}
Desarrollar un sistema eficiente y escalable que pueda resolver problemas de
reescritura de consultas o composición de servicios con constantes, tomando en
cuenta parámetros de calidad para obtener una mejor solución.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame
{
\section{Retos}
\begin{itemize}
\item Gran número de fuentes o servicios.
\item Manejo de transitividad en asociaciones de variables con constantes.
\end{itemize}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\frame
{
\section{Solución Propuesta}
\begin{enumerate}
\item Traducir instancias del problema a instancias de SAT (ya realizado).
\item Añadir cláusulas para tratar con constantes.
\item Compilar teorías a d-DNNF.
\item Calcular todas las reescrituras -- o la mejor.
\end{enumerate}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Resultados Estudio Experimental}

\subsection{Vuelos con escalas y aerolíneas}
\frame
{

\begin{center}
\includegraphics[width=0.5\textwidth]{plot1}
\end{center}

\begin{itemize}
\item Instancias de tamaños realistas
\item Buenos tiempos incluso para las instancias más grandes
\end{itemize}

Esto incluye cálculo de todas las reescrituras o de la mejor.
}

\subsection{Otros experimentos}
\frame
{
\begin{itemize}
\item Con dos vistas por aerolínea: compilación un minuto en caso más grande
\item Menos de seis minutos para problema aleatorio con
\begin{itemize}
\item 80 vistas
\item 5 objetivos de consulta
\item 4 predicados por vista
\end{itemize}

\end{itemize}

Tiempos dados para cálculo de todas las reescrituras.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusiones y Trabajo Futuro}

\frame
{
\begin{itemize}
\item Se puede escalar a tamaños razonables.
\item Es factible manejar constantes y rendimiento con la aproximación
propuesta.
\item Se podría pensar en aun más escalabilidad: posible combinación de este
sistema con algoritmo tipo branch and bound.
\end{itemize}
}

\end{document}

### 2010-03-26 coloquio/presentacion.tex
fin presentacion del coloquio

@@
 
-\title{Selección de Servicios Web de Máxima Utilidad Con Circuitos}
+\title{Instanciación Óptima de Flujos de Trabajo Abstractos Usando Lógica y Circuitos}
 \author{Daniel Izquierdo}
-\date{14 de diciembre de 2009}
+\date{26 de marzo de 2010}
+
+\newcommand{\qrule}{:\!\!-}
+\newcommand{\vuelo}{\text{\it vuelo}}
+\newcommand{\ciudadVE}{\text{\it ciudadVE}}
+\newcommand{\nacional}{\text{\it nacional}}
+\newcommand{\ida}{\text{\it ida}}
+\newcommand{\unaescala}{\text{\it una-escala}}
+\newcommand{\vueloCCS}{\text{\it vuelo-hacia-ccs}}
+\newcommand{\unaparadaCCS}{\text{\it unaescala-hacia-ccs}}
+\newcommand{\desdeVAL}{\text{\it desde-val}}
+\newcommand{\PA}{\text{CCS}}
+\newcommand{\NY}{\text{VAL}}
+\newcommand{\AL}{\text{AL}}
 
@@
 
+\section{Motivación}
 \frame
 {
-\section{Motivación}
 
@@
 
+\section{Objetivo}
 \frame
 {
-\section{Objetivo}
 Desarrollar un sistema eficiente y escalable que pueda resolver problemas de
@@
 
+\section{Ejemplo}
+
+\subsection{Servicios abstractos y consulta}
+
 \frame
 {
+    \begin{itemize}
+    \item Servicios abstractos:
+        \begin{itemize}
+        \item $\vuelo(x,y)$
+        \item $\ciudadVE(x)$
+        \end{itemize}
+    \item Consulta:\\ $W(x,w,y,z) \qrule$\\ $\ciudadVE(x),\,\vuelo(x,w),\,\vuelo(w,y),\,\vuelo(y,z),\,\vuelo(z,x)$
+    \end{itemize}
+}
+
+\subsection{Servicios concretos}
+
+\frame
+{
+\begin{itemize}
+\item {\tiny $\nacional(x,y)$ relaciona dos ciudades de Venezuela conectadas por un vuelo directo,} \\
+      {\tiny $\nacional(x,y)\   \qrule\ \vuelo(x,y),\,\ciudadVE(x),\,\ciudadVE(y)$}\\
+\item {\tiny $\ida(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,}\\
+      {\tiny $\ida(x,y)\     \qrule\ \vuelo(x,y)$}\\
+\item {\tiny $\unaescala(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,}\\
+      {\tiny $\unaescala(x,z)\    \qrule\ \vuelo(x,y),\,\vuelo(y,z)$}\\
+\item {\tiny $\vueloCCS(x)$ dice si hay un vuelo directo desde $x$ a Caracas,}\\
+      {\tiny $\vueloCCS(x)\     \qrule\ \vuelo(x,\PA)$}\\
+\item {\tiny $\unaparadaCCS(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a Caracas con una escala en $y$,}\\
+      {\tiny $\unaparadaCCS(x,y)\  \qrule\ \vuelo(x,y),\,\vuelo(y,\PA)$}\\
+\item {\tiny $\desdeVAL(x)$ dice si hay un vuelo de Valencia a $x$,}\\
+      {\tiny $\desdeVAL(x)\       \qrule\ \vuelo(\NY,x)$}\\
+\end{itemize}
+}
+
+\subsection{Composiciones}
+
+\frame
+{
+    Pueden existir composiciones válidas
+{\tiny $$I(x,w,y,z)\ \qrule\ \nacional(x,w),\,\vueloCCS(w),\,\ida(\PA,z),\,\nacional(z,x)$$}
+
+    y no válidas
+
+{\tiny $$I'(x,w,y,z)\  \qrule\ \nacional(x,y),\,\vueloCCS(y),\,\desdeVAL(z),\,\nacional(z,x)$$\\
+       $$I''(x,w,y,z)\ \qrule\ \unaescala(x,y),\,\ida(y,z),\,\nacional(z,x)$$\\
+}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Retos}
+\frame
+{
 \begin{itemize}
@@
 
+\section{Solución Propuesta}
 \frame
 {
-\section{Solución Propuesta}
 \begin{enumerate}
@@
 
-\section{Resultados Estudio Experimental}
+\section{Estudio Experimental}
 
 \subsection{Vuelos con escalas y aerolíneas}
+\frame
+{
+    Primer experimento: búsqueda de vuelos con $n$ escalas usando la misma
+    aerolínea para todas las escalas.
+}
+
 \frame
@@
 propuesta.
-\item Se podría pensar en aun más escalabilidad: posible combinación de este
-sistema con algoritmo tipo branch and bound.
+\item Queremos aun más escalabilidad: posible combinación de este
+sistema con algoritmo tipo branch and bound, o Max-Sat solvers.
 \end{itemize}

### 2010-03-31 tesis/propuesta/secciones/portada.tex
pequena correccion de resumen de leonid

@@
 eficientemente el problema de instanciación que está basada en
-la lógica proposicional. La infraestructura adopta el enfoque de definición denominado Local-As-View en la cual la
+la lógica proposicional. La infraestructura adopta el enfoque de definición denominado Local-As-View en el cual la
 funcionalidad de servicios concretos se describe usando vistas de servicios

### 2010-03-31 tesis/propuesta/secciones/justificacion.tex
muchas correcciones de la propuesta de tesis

@@
 
-Estamos tratando con un problema difícil \cite{lenzerini:dataintegration} en el
+Se está tratando con un problema difícil \cite{lenzerini:dataintegration} en el
 sentido de que es computacionalmente complejo. Más formalmente, el problema
@@
 
-Debido
+Como fue expuesto anteriormente, la explosión de datos y servicios web
+disponibles lleva a que los usuarios necesiten realizar consultas cada vez más
+complejas, y por lo tanto existe la necesidad de un sistema eficiente capaz de
+resolverlas.

### 2010-03-31 tesis/propuesta/secciones/planteamiento.tex
muchas correcciones de la propuesta de tesis

@@
 En el contexto de la Web Semántica, y con el soporte de las Arquitecturas Orientadas a
-Servicios (SOA), el número de fuentes de datos y servicios Web ha explotado en
+Servicios (SOA), el número de fuentes de datos y servicios Web ha crecido rápidamente en
 los últimos años. Por ejemplo, la colección de bases de datos de biología
@@
 esta variedad de datos, la tendencia de los usuarios es hacia depender cada vez
-más de métodos automáticos para manejarlos tales como recuperación de datos de
+más de métodos automáticos para manejarlos, tales como recuperación de datos de
 fuentes públicas y análisis utilizando herramientas o servicios Web compuestos
@@
 servicios concretos o la complejidad (número
-de objetivos, tamaño) de la estructura del flujo de trabajo  aumenta. Se llama al
+de objetivos, tamaño) de la estructura del flujo de trabajo  aumenta. Se denomina al
 problema de instanciar un flujo de trabajo dado con servicios concretos de un
 conjunto de servicios concretos dado, de manera que ciertas restricciones sobre QoS sean
-cumplidas, el Problema de Instanciación de Flujos de Trabajo (WIP).
+cumplidas, el Problema de Instanciación de Flujos de Trabajo (WIP, por sus siglas en inglés \emph{Workflow Instantiation Problem}).
 
 Bajo estas suposiciones, WIP puede ser convertido en el bien conocido Problema
-de Reescritura de Consultas (QRP) para LAV que es central a los sistemas de
+de Reescritura de Consultas (QRP, por sus siglas en inglés \emph{Query Rewriting Problem}) para Local-As-View (LAV) que es central a los sistemas de
 integración \cite{halevy:survey}. QRP consiste de una consulta conjuntiva que debe ser
@@
 de integración de datos \cite{Chen05,JaudoinPRST05}, y optimización de consultas y mantenimiento
-de datos \cite{AfratiLU07,levy:bucket}, y varias soluciones que escalan a un número grande de
+de datos \cite{AfratiLU07,levy:bucket}. Varias soluciones que escalan a un número grande de
 vistas han sido definidas \cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
@@
 y un conjunto de fuentes definidas como vistas, se construye una teoría
-lógica de manera que cada modelo de la teoría codifica una
-reescritura válida, y así todas las reescrituras son obtenidas enumerando los
+lógica de manera que cada modelo de la teoría está en correspondencia con una
+reescritura válida distinta, y así todas las reescrituras son obtenidas enumerando los
 modelos de la teoría. Esta enumeración puede ser realizada eficientemente si la
 teoría lógica está en cierta forma normal llamada la forma normal de negación
-determinística y descomponible (d-DNNF) \cite{darwiche:d-dnnfs}. Así,
-la solución consiste en transformar (llamado compilar en el campo de
-compilación del conocimiento) la teoría lógica en el formato d-DNNF para
-enumerar sus modelos eficientemente.
+determinística y descomponible (d-DNNF, por sus siglas en inglés \emph{Decomposable Deterministic Negation Normal Form}) \cite{darwiche:d-dnnfs}. Así,
+la solución consiste en construir la teoría lógica, transformarla al formato d-DNNF para
+enumerar sus modelos eficientemente, y recuperar las reescrituras a partir de los modelos.
 
@@
 los mejores modelos, en tiempo lineal en el tamaño del d-DNNF. Este cómputo
-transforma el GAD del d-DNNF en un circuito aritmético reemplazando los nodos
+transforma el grafo acíclico dirigido del d-DNNF en un circuito aritmético reemplazando los nodos
 AND con `+' y los nodos OR con `min'. La función de calidad de literales
@@
 
-Considere un sistema simple de información de vuelos que contiene información
+Considere un sistema sencillo de información de vuelos que contiene información
 sobre vuelos entre ciudades e información sobre qué ciudades están en los
-Estados Unidos. Un sistema así puede ser descrito usando LAV con los dos servicios
+Venezuela. Un sistema así puede ser descrito usando LAV con los dos servicios
 abstractos $\vuelo(x,y)$ y $\ciudadVE(x)$. El primero relaciona dos ciudades $x$ e $y$ si
@@
 \begin{enumerate}[--]
-\item $\nacional(x,y)$ relaciona dos ciudades de Venezuela conectadas por un vuelo directo,
-\item $\ida(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida,
-\item $\unaescala(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala,
-\item $\vueloCCS(x)$ dice si hay un vuelo directo desde $x$ a Caracas,
-\item $\unaparadaCCS(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a Caracas con una escala en $y$, y
-\item $\desdeVAL(x)$ dice si hay un vuelo de Valencia a $x$.
+\item $\nacional(x,y)$ relaciona dos ciudades de Venezuela conectadas por un vuelo directo
+\item $\ida(x,y)$ relaciona dos ciudades conectadas por un vuelo de ida
+\item $\unaescala(x,y)$ relaciona dos ciudades conectadas por un vuelo con una escala
+\item $\vueloCCS(x)$ dice si hay un vuelo directo desde $x$ a Caracas
+\item $\unaparadaCCS(x,y)$ relaciona $x$ y $y$ si hay un vuelo desde $x$ a Caracas con una escala en $y$
+\item $\desdeVAL(x)$ dice si hay un vuelo de Valencia a $x$
 \end{enumerate}

### 2010-03-31 tesis/propuesta/secciones/portada.tex
muchas correcciones de la propuesta de tesis

@@
 hecha en el momento, consiste de una búsqueda en un espacio de posibilidades
-combinatorio. En este trabajo se propone una infraestructura para resolver
-eficientemente el problema de instanciación que está basada en
-la lógica proposicional. La infraestructura adopta el enfoque de definición denominado Local-As-View en el cual la
+combinatorio. En este trabajo se propone una infraestructura basada en
+la lógica proposicional para resolver
+eficientemente el problema de instanciación. La infraestructura adopta el enfoque de definición denominado Local-As-View en el cual la
 funcionalidad de servicios concretos se describe usando vistas de servicios
@@
 instancias grandes como muestran resultados experimentales ya realizados, sino
-que también podemos verificar que es correcta y completa dado que, estando basada en la lógica, es
+que también se puede verificar que es correcta y completa dado que, estando basada en la lógica, es
 fácilmente sujeta al análisis formal.

### 2010-03-31 tesis/propuesta/secciones/solucion.tex
muchas correcciones de la propuesta de tesis

@@
 
-Consideramos catálogos de servicio de la forma $C=\tup{S,T}$ donde $S$ es un
+Se consideran catálogos de servicio de la forma $C=\tup{S,T}$ donde $S$ es un
 conjunto de predicados que representa servicios abstractos y $T=\{T_s\}_{s\in S}$ es una
@@
 
-El objetivo de los experimentos es determinar el rendimiento de la propuesta con
+El objetivo de los experimentos fue determinar el rendimiento de la propuesta con
 condiciones variantes. El principal beneficio de la solución propuesta es
@@
 y vistas. Se cree que estos resultados son alentadores y motivan a continuar
-esta línea de investigación.
+utilizando este enfoque para atacar el problema de WIP teniendo en mente la meta
+de obtener escalabilidad y eficiencia.

### 2010-03-31 tesis/propuesta/secciones/trabajorelacionado.tex
muchas correcciones de la propuesta de tesis

@@
 podemos probar si un DNNF es satisfactible haciendo un sólo recorrido 
-sobre su GAD en tiempo linear. Se dice que un DNNF es
+sobre su GAD en tiempo lineal. Se dice que un DNNF es
 determinístico (d-DNNF) \cite{darwiche:d-dnnfs} si para cada disjunción $\bigvee_i\phi_i$, los
@@
 descomposición que está relacionado a la ``conectividad'' de la teoría CNF.
-Sin embargo, en nuestros experimentos, las teorías CNF que son compiladas son de
-ancho pequeño.
+Sin embargo, en los experimentos realizados para esta propuesta las teorías CNF
+que fueron compiladas tuvieron ancho pequeño.

### 2010-06-25 tesis/src/omcdsat/mcdsat/Main.py
clausulas 6 con ontologia (inicio)

@@ -32,7 +32,13 @@ if __name__ == "__main__":
         else:
             archCostosSat = None
 
-        traducir(exp, archVistas, archCons, archVars, archTiempo, archSalida, archCostosSat)
+        if len(sys.argv) > 9:
+            archOntology = sys.argv[9]
+        else:
+            archOntology = None
+
+
+        traducir(exp, archVistas, archCons, archVars, archTiempo, archSalida, archCostosSat, archOntology)
     elif op == 'G':
         if len(sys.argv) > 7:
             archCostos = sys.argv[7]

### 2010-06-25 tesis/src/omcdsat/mcdsat/Traductor/Traductor3.py

@@ -4,6 +4,7 @@ import resource
 import pprint
 import random
 import cPickle
+import pickle
 
 from Parser import *
 from Parser.CQparser import *
@@ -24,10 +25,10 @@ varsT = {}
 varsZ = {}
 contribVT = {}
 
-def traducir(exp, archV, archC, archVars, archTiempo, archSalida, archCostosSat):
+def traducir(exp, archV, archC, archVars, archTiempo, archSalida, archCostosSat, archOntology=None):
     #tiempo = timeit.Timer('traducir1()', "from __main__ import traducir1; import psyco; psyco.full()").timeit(1)/1
     tiempoi = resource.getrusage(resource.RUSAGE_SELF)[0]
-    traducir1(exp, archV, archC, archVars, archSalida, archCostosSat)
+    traducir1(exp, archV, archC, archVars, archSalida, archCostosSat, archOntology)
     tiempof = resource.getrusage(resource.RUSAGE_SELF)[0]
     
     fileobject = open(archTiempo, 'w')
@@ -44,14 +45,19 @@ def traducirBig(exp, archV, archC, archVars, archTiempo, archSalida, archCostos)
     fileobject.write(str(tiempof-tiempoi))
     fileobject.close()
 
-def traducir1(exp, archVistas, archCons, archVars, archSalida, archCostosSat): #(archVistas, archCons):
+def traducir1(exp, archVistas, archCons, archVars, archSalida, archCostosSat, archOntology): #(archVistas, archCons):
     vistas = cargarCQ(archVistas)
     consultas = cargarCQ(archCons)
+
+    ao = open(archOntology, 'r')
+    ontology = pickle.load(ao)
+    ao.close()
+
     for q in consultas:
         if exp == 'SatRW' or exp == 'SatBestRW':
-            transf = traducirConsultaRW(q, vistas, archSalida)
+            transf = traducirConsultaRW(q, vistas, ontology, archSalida)
         elif exp == 'Sat':
-            transf = traducirConsultaMCD(q, vistas, archSalida)
+            transf = traducirConsultaMCD(q, vistas, ontology, archSalida)
         guardarVars(transf, archVars)
 
 def traducir1Big(exp, archVistas, archCons, archVars, archSalida, archCostos): #(archVistas, archCons):
@@ -66,19 +72,19 @@ def guardarVars(transf, archVars):
     fh.close()
 
 
-def traducirConsultaMCD(q, vistas, archSalida):
[…]

### 2010-06-26 tesis/src/omcdsat/mcdsat/Ontologies/ontologies.py
fix transitive closure infinito

@@ -112,7 +112,7 @@ def generate_specs(filename):
         tc.update(new)
         pos_len = len(tc)
 
-        if pre_len < pos_len:
+        if pre_len == pos_len:
             break
 
     tc_filename = "%s.ontology" % (filename)

### 2010-06-26 tesis/src/omcdsat/mcdsat/Traductor/Traductor3.py
usando can_specialize en Traductor3

@@ -72,19 +72,19 @@ def guardarVars(transf, archVars):
     fh.close()
 
 
-def traducirConsultaMCD(q, vistas, ontologia, archSalida):
-    variables, clausulas = generarTeoriaMCD(q, vistas, ontologia)
+def traducirConsultaMCD(q, vistas, ontology, archSalida):
+    variables, clausulas = generarTeoriaMCD(q, vistas, ontology)
     transf = TransformarFormula(variables)
     n = 1
     imprimirCopias(variables, clausulas, [], n, transf, archSalida)
     return transf
 
 
-def traducirConsultaRW(q, vistas, ontologia, archSalida):
+def traducirConsultaRW(q, vistas, ontology, archSalida):
     variables, clausulas = generarTeoriaMCD(q, vistas)
     transf = TransformarFormula(variables)
     n = len(q.cuerpo)
-    clausulas2 = clausulasCombinarMCD(transf, n, q, vistas, ontologia)
+    clausulas2 = clausulasCombinarMCD(transf, n, q, vistas, ontology)
     imprimirCopias(variables, clausulas, clausulas2, n, transf, archSalida)
     return transf
 
@@ -108,7 +108,7 @@ def traducirConsultaBigBestRW(q, vistas, archSalida, archCostos):
     imprimirCopiasPeso(variables, clausulas, clausulas2, n, costDict, transf, archSalida)
     return transf
 
-def clausulasCombinarMCD(transf, n, q, vistas, ontologia):
+def clausulasCombinarMCD(transf, n, q, vistas, ontology):
     clausulas2 = []
     for numG in xrange(n):
         varG = VariableSat(False, 'g', [numG])
@@ -200,16 +200,16 @@ def imprimirCopiasPeso(variables, clausulas, clausulas2, numCopias, pesos, trans
     #arch.write('%\n')
     arch.close()
 
-def generarTeoriaMCD(q, vistas, ontologia):
+def generarTeoriaMCD(q, vistas, ontology):
     global varsV
     global varsG
     global varsT
     global varsZ
     lv, c1, c2 = variablesV(q, vistas)
     lg, c3 = variablesG(q, lv)
-    lt, lz, c6, c7, c8, c9, c14, ltaux = clausulas678(q, vistas)
+    lt, lz, c6, c7, c8, c9, c14, ltaux = clausulas678(q, vistas, ontology)
 
-    print ontologia
+    print ontology
 
     # support for constants
 
[…]

### 2010-06-27 tesis/benchmarks/omcdsat/0/generator.sh
benchmark 0 de omcdsat

@@ -1,0 +1,12 @@
+#!/bin/bash
+
+QRPGEN=/home/daniel/proyectos/postgrado/tesis/src/qrpgen/qrpgen.py
+
+for nquery_sos in `seq 2 1 20`; do
+    for nviews in `seq 10 10 50`; do
+        VIEWSFILE=casos/views-$nviews-$nquery_sos.in
+        QUERYFILE=casos/query-$nviews-$nquery_sos.in
+
+        $QRPGEN $VIEWSFILE $QUERYFILE flights.sameairline_manyviews $nviews $nquery_sos 2
+    done
+done

### 2010-06-27 tesis/benchmarks/omcdsat/0/runner.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+INDIR=casos
+
+OUT_MCD=plot_mcd7.raw
+OUT_MMS=plot_mms7.raw
+
+TIME_MCD="timeout 120 /usr/bin/time -apo ${OUT_MCD}"
+TIME_MMS="timeout 120 /usr/bin/time -apo ${OUT_MMS}"
+
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
+
+MCDSAT=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/mcdsat
+MINIMAXSAT="/home/daniel/proyectos/postgrado/tesis/src/MiniMaxSat/minimaxsat1.0 -F=2"
+
+echo '################################################################################' | tee -a ${OUT_MCD}
+echo '################################################################################' | tee -a ${OUT_MMS}
+
+for nquery_sos in `seq 2 1 20`; do
+    for nviews in `seq 10 10 50`; do
+        VIEWSFILE=casos/views-$nviews-$nquery_sos.in
+        QUERYFILE=casos/query-$nviews-$nquery_sos.in
+        COSTFILE=../../../src/costs-pot2.txt
+
+        CNF=`basename $VIEWSFILE`.cnf
+
+        echo ${nviews} ${nquery_sos} | tee -a ${OUT_MCD}
+        echo ${nviews} ${nquery_sos} | tee -a ${OUT_MMS}
+
+        for i in `seq 1`; do #n times to get average
+            ${TIME_MCD} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+
+            ${TIME_MMS} ./mmsrunner.sh "$MCDSAT" "$VIEWSFILE" "$QUERYFILE" "$COSTFILE" "$MINIMAXSAT" "$CNF"
+        done
+    done
+done

### 2010-06-27 tesis/src/omcdsat/mcdsat/Ontologies/ontologies.py
fix ontologies creciente infinito

@@ -86,6 +86,23 @@ class Spec:
 
         return (new_subclass_args, new_class_args)
 
+    def __eq__(self, other):
+        return self.subclass_name == other.subclass_name and \
+               self.subclass_args == other.subclass_args and \
+               self.class_name == other.class_name       and \
+               self.class_args == other.class_args
+
+    def __ne__(self, other):
+        return not (self == other)
+
+    def __hash__(self):
+        th = "(%s):(%s):(%s):(%s)" % (self.subclass_name,
+                                      self.subclass_args,
+                                      self.class_name,
+                                      self.class_args)
+
+        return hash(th)
+
 def generate_specs(filename):
     o = load_ontology(filename)
 

### 2010-06-27 tesis/src/omcdsat/mcdsat/Traductor/Traductor3.py

@@ -209,7 +209,7 @@ def generarTeoriaMCD(q, vistas, ontology):
     lg, c3 = variablesG(q, lv)
     lt, lz, c6, c7, c8, c9, c14, ltaux = clausulas678(q, vistas, ontology)
 
-    print ontology
+    #print ontology
 
     # support for constants
 
@@ -547,9 +547,6 @@ def clausulas_d3(q, vistas):
 
 def can_specialize(preda, predb, ontology):
     for spec in ontology:
-        print "checking spec"
-        print "spec: %s [ %s" % (spec.subclass_name, spec.class_name)
-        print "pred: %s [ %s" % (preda, predb)
         if spec.subclass_name == preda and spec.class_name == predb:
             return True
 
@@ -583,15 +580,13 @@ def clausulas678(q, vistas, ontology):
             subObCubre = False
             for subObtemp in v.cuerpo:
                 predtemp = subObtemp.predicado
-                print "checking %s against %s" % (pred,predtemp)
                 if pred == predtemp or can_specialize(pred, predtemp, ontology):
-                    print "success"
                     varz = VariableSat(True, 'z', [i,j,m])
                     lz.append(varz)
                     varsZ[(i,j,m)]=varz
                     c6temp.append(varz)
                     c9temp.append(varz)
-                    lttemp = clausula78a(varz, varg, varm, subOb, subObtemp, m, ltaux, c7, c8, v)
+                    lttemp = clausula78a(varz, varg, varm, subOb, subObtemp, m, ltaux, c7, c8, v, pred != predtemp)
                     lt |= lttemp
                     subObCubre = True
                 j = j + 1
@@ -620,7 +615,7 @@ def clausula9(c9temp):
            c9.append([c9temp[x].negarVar(), c9temp[y].negarVar()])
     return c9
 
-def clausula78a(varz, varg, varm, subObQ, subObV, vis, ltaux, c7, c8, vista):
+def clausula78a(varz, varg, varm, subObQ, subObV, vis, ltaux, c7, c8, vista, is_spec):
     global varsT
     global contribVT
     c8temp1 = [varz, varm.negarVar()]
@@ -635,6 +630,9 @@ def clausula78a(varz, varg, varm, subObQ, subObV, vis, ltaux, c7, c8, vista):
         contribVT.setdefault(vista, set([])).add((int(x), int(y)))
         c7temp = [varm.negarVar(), varz.negarVar(), varT]
[…]

### 2010-06-27 tesis/benchmarks/omcdsat/0/generator.sh
generator de benchmark 0

@@ -1,12 +1,7 @@
 #!/bin/bash
 
-QRPGEN=/home/daniel/proyectos/postgrado/tesis/src/qrpgen/qrpgen.py
+GEN=./generator.py
 
-for nquery_sos in `seq 2 1 20`; do
-    for nviews in `seq 10 10 50`; do
-        VIEWSFILE=casos/views-$nviews-$nquery_sos.in
-        QUERYFILE=casos/query-$nviews-$nquery_sos.in
-
-        $QRPGEN $VIEWSFILE $QUERYFILE flights.sameairline_manyviews $nviews $nquery_sos 2
-    done
+for levels in `seq 7`; do
+        $GEN $levels > casos/${levels}levels.ontology
 done

### 2010-06-27 tesis/src/omcdsat/mcdsat/Traductor/Traductor3.py
fix para rewritings

@@ -81,7 +81,7 @@ def traducirConsultaMCD(q, vistas, ontology, archSalida):
 
 
 def traducirConsultaRW(q, vistas, ontology, archSalida):
-    variables, clausulas = generarTeoriaMCD(q, vistas)
+    variables, clausulas = generarTeoriaMCD(q, vistas, ontology)
     transf = TransformarFormula(variables)
     n = len(q.cuerpo)
     clausulas2 = clausulasCombinarMCD(transf, n, q, vistas, ontology)
@@ -101,7 +101,7 @@ def traducirConsultaBigBestRW(q, vistas, archSalida, archCostos):
 
     #impresion
 
-    variables, clausulas = generarTeoriaMCD(q, vistas)
+    variables, clausulas = generarTeoriaMCD(q, vistas, ontology)
     transf = TransformarFormula(variables)
     n = len(q.cuerpo)
     clausulas2 = clausulasCombinarMCD(transf, n, q, vistas)
@@ -547,7 +547,7 @@ def clausulas_d3(q, vistas):
 
 def can_specialize(preda, predb, ontology):
     for spec in ontology:
-        if spec.subclass_name == preda and spec.class_name == predb:
+        if spec.subclass_name == predb and spec.class_name == preda:
             return True
 
     return False

### 2010-06-27 tesis/benchmarks/omcdsat/0/orunner.sh
runners de benchmark0

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+GENONTOLOGIES="$1"
+OMCDSAT="$2"
+ONTOLOGY="$3"
+VIEWS="$4"
+QUERY="$5"
+
+${GENONTOLOGIES} $ONTOLOGY && \
+${OMCDSAT} RW ${VIEWS} ${QUERY} casos/costs.txt ${ONTOLOGY}.ontology > /dev/null
+true

### 2010-06-27 tesis/benchmarks/omcdsat/0/runner.sh

@@ -1,36 +1,21 @@
 #!/bin/bash
 
 INDIR=casos
+OUT=omcdsat_0.raw
+TIME="timeout 600 /usr/bin/time -apo ${OUT}"
+GENONTOLOGIES=/home/daniel/proyectos/postgrado/tesis/src/omcdsat/mcdsat/GenOntologies.py
+OMCDSAT=/home/daniel/proyectos/postgrado/tesis/src/omcdsat/mcdsat/mcdsat
 
-OUT_MCD=plot_mcd7.raw
-OUT_MMS=plot_mms7.raw
-
-TIME_MCD="timeout 120 /usr/bin/time -apo ${OUT_MCD}"
-TIME_MMS="timeout 120 /usr/bin/time -apo ${OUT_MMS}"
-
-PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
-
-MCDSAT=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/mcdsat
-MINIMAXSAT="/home/daniel/proyectos/postgrado/tesis/src/MiniMaxSat/minimaxsat1.0 -F=2"
-
-echo '################################################################################' | tee -a ${OUT_MCD}
 echo '################################################################################' | tee -a ${OUT_MMS}
 
-for nquery_sos in `seq 2 1 20`; do
-    for nviews in `seq 10 10 50`; do
-        VIEWSFILE=casos/views-$nviews-$nquery_sos.in
-        QUERYFILE=casos/query-$nviews-$nquery_sos.in
-        COSTFILE=../../../src/costs-pot2.txt
+for levels in `seq 2 3`; do
+    ONTOLOGY=casos/${levels}levels.ontology
+    VIEWS=casos/vistas.txt
+    QUERY=casos/query.txt
 
-        CNF=`basename $VIEWSFILE`.cnf
+    echo ${levels} | tee -a ${OUT}
 
-        echo ${nviews} ${nquery_sos} | tee -a ${OUT_MCD}
-        echo ${nviews} ${nquery_sos} | tee -a ${OUT_MMS}
-
-        for i in `seq 1`; do #n times to get average
-            ${TIME_MCD} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
-
[…]

### 2010-06-27 tesis/benchmarks/omcdsat/0/runner.sh
benchmark 0 corrido

@@ -8,7 +8,7 @@ OMCDSAT=/home/daniel/proyectos/postgrado/tesis/src/omcdsat/mcdsat/mcdsat
 
 echo '################################################################################' | tee -a ${OUT_MMS}
 
-for levels in `seq 2 3`; do
+for levels in `seq 2 7`; do
     ONTOLOGY=casos/${levels}levels.ontology
     VIEWS=casos/vistas.txt
     QUERY=casos/query.txt

### 2010-06-27 tesis/benchmarks/omcdsat/1/generator.py
benchmark 1 ontologias

@@ -1,0 +1,17 @@
+#!/usr/bin/env python
+
+import sys
+
+first = 3
+top = 4
+
+levels = int(sys.argv[1])
+
+while top <= 2**levels:
+    for i in xrange(first, top+1):
+        parent = (i+1)/2
+
+        print "r%d(X1, X2) :- r%d(X1,X2)" % (i, parent)
+
+    first = top+1
+    top *= 2

### 2010-06-27 tesis/benchmarks/omcdsat/1/generator.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+GEN=./generator.py
+
+for levels in `seq 7`; do
+        $GEN $levels > casos/${levels}levels.ontology
+done

### 2010-06-27 tesis/benchmarks/omcdsat/1/orunner.sh

@@ -1,0 +1,11 @@
+#!/bin/bash
+
+GENONTOLOGIES="$1"
+OMCDSAT="$2"
+ONTOLOGY="$3"
+VIEWS="$4"
+QUERY="$5"
+
+${GENONTOLOGIES} $ONTOLOGY && \
+${OMCDSAT} BESTRW ${VIEWS} ${QUERY} casos/costs.txt ${ONTOLOGY}.ontology > /dev/null
+true

### 2010-06-27 tesis/benchmarks/omcdsat/1/runner.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+
+INDIR=casos
+OUT=omcdsat_0.raw
+TIME="timeout 600 /usr/bin/time -apo ${OUT}"
+GENONTOLOGIES=/home/daniel/proyectos/postgrado/tesis/src/omcdsat/mcdsat/GenOntologies.py
+OMCDSAT=/home/daniel/proyectos/postgrado/tesis/src/omcdsat/mcdsat/mcdsat
+
+echo '################################################################################' | tee -a ${OUT_MMS}
+
+for levels in `seq 2 7`; do
+    ONTOLOGY=casos/${levels}levels.ontology
+    VIEWS=casos/vistas.txt
+    QUERY=casos/query.txt
+
+    echo ${levels} | tee -a ${OUT}
+
+    for i in `seq 1`; do #n times to get average
+        ${TIME} ./orunner.sh "$GENONTOLOGIES" "$OMCDSAT" "$ONTOLOGY" "$VIEWS" "$QUERY"
+    done
+done

### 2010-06-28 publicaciones/iswc10/paperISWC10v2.tex
paper v2

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{url}

\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\Omit}[1]{}
\newcommand{\vvec}[1]{\mathbf{#1}}
\newcommand{\join}{\bowtie}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\mcdsat}{\textsc{McdSat}}
\newcommand{\minicon}{{MiniCon}}

\newcommand{\qrule}{:\!\!-}
\newcommand{\orule}{\sqsubseteq}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\M}{\mathcal{M}}

% Ontology concepts
\newcommand{\trip}{\textit{trip}}
\newcommand{\planetrip}{\textit{plane-trip}}
\newcommand{\traintrip}{\textit{train-trip}}
\newcommand{\AAflight}{\textit{AA-flight}}
\newcommand{\UAflight}{\textit{UA-flight}}
\newcommand{\ATtrain}{\textit{AT-train}}
\newcommand{\UPtrain}{\textit{UP-train}}
\newcommand{\flight}{\textit{flight}}
\newcommand{\train}{\textit{train}}
\newcommand{\UScity}{\textit{uscity}}

% Constants
\renewcommand{\AA}{\text{AA}}
\newcommand{\UA}{\text{UA}}
\newcommand{\AT}{\text{AT}}
\newcommand{\UP}{\text{UP}}
\newcommand{\PA}{\text{Paris}}
\newcommand{\NY}{\text{NY}}
\newcommand{\LA}{\text{LA}}
\newcommand{\AL}{\text{AL}}

% Services
\newcommand{\nationalFlight}{\textit{national-flight}}
\newcommand{\onewayFlight}{\textit{one-way-flight}}
\newcommand{\nationalTrain}{\textit{national-train}}
\newcommand{\onewayTrain}{\textit{one-way-train}}
\newcommand{\onestop}{\textit{one-stop}}
\newcommand{\toPA}{\textit{to-pa}}
\newcommand{\onestopPA}{\textit{onestop-to-pa}}
\newcommand{\fromNY}{\textit{from-ny}}
\newcommand{\fromLA}{\textit{from-la}}

\begin{document}

\allowdisplaybreaks
\title{An Expressive and Efficient Solution to the Service Selection Problem}
\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
\institute{Departamento de Computaci\'on \\
           Universidad Sim\'on Bol\'{\i}var \\
           Caracas 89000, Venezuela \\
           \url{{idaniel,mvidal,bonet}@ldc.usb.ve}}
\maketitle

\begin{abstract}
Given the tremendous amount of potential Semantic Web 
Services that can be created from online sources by using existing 
annotation tools, expressive formalisms and efficient and scalable approaches to solve the Service Selection Problem (SSP) are required. In this paper, we propose a framework that 
is firmly grounded on logic. The framework adopts the Local-As-View 
approach~\cite{levy:bucket} that semantically describes services in terms of concepts in a domain 
ontology. A user request is comprised of a conjunctive query on concepts 
from the domain ontology and a set of preferences; the query expresses 
the abstract concepts that must be combined to solve the query, and 
the preferences are used to rank the solutions. Relations in the domain 
ontology are considered to augment the space of available services. Using 
this representation, the problem of service selection is cast as a problem 
of query rewriting, from the area of integration systems. Then, 
building on related work, we devise an encoding of SSP as a logical theory 
whose models are in correspondence with the service rewritings of the 
user query, and in presence of preferences, whose best models are in 
correspondence with the best-ranked valid solutions. Thus, by exploiting 
known properties of modern SAT solvers, we provide an efficient and 
scalable solution to SSP. The approach provides the basis to represent 
a large number of real-world situations; additionally, it does not 
only scale up to large instances, as the experimental results show, 
but is also sound and complete since, being founded on logic, is 
amenable to formal analysis.
\end{abstract}

\section{Introduction}
Existing Web infrastructures support the publication and access to a
tremendous amount of Web data sources, some of which can be labeled
and converted into  Semantic Web Services by using existing annotation
tools like the one proposed  by Ambite et al. \
\cite{AmbiteISWC09}. Once this huge dataset of Semantic Web  Services
become available, users more than ever will require techniques to
effectively select the services that meet their requirements. In order
to achieve  this goal, functional and non-functional properties of the
services should be  precisely described, as well as criteria to select
and rank the solutions that best meet user requests.   In this paper,
we extend an approach traditionally used in the area of data
integration to solve the Service Selection Problem. First, concepts
from a domain ontology provide a precise description of the semantics
of available sources. These  descriptions correspond to views on
concepts of the domain ontology and follow  the Local-As-View approach
(LAV)~\cite{levy:bucket}. LAV mappings are able to  scale to service
datasets that constantly change, in contrast to Global-As-View
definitions  (GAV), where generic concepts are defined in terms of the
available sources and  cannot be easily modified when sources
constantly change. A user request is  modeled as a conjunctive query
on the ontology concepts, and user preferences  are associated with a
cost and constraint the solutions of the query.  Based on this
integration framework, we define the problem of selecting the
services that best meet the user request, as the problem of
enumerating and  ranking the service rewritings of the user
query. Each rewriting is associated  with a cost equal to the cost of
the preferences violated by the rewriting, and the best ranked
rewriting is one that has minimum cost among all valid
rewritings. Knowledge encoded in the domain ontology is used to expand
the  space of rewritings. Thus, our proposed service selection
approach is cast into  the well-known Query Rewriting Problem (QRP)
for LAV which is central to  integration systems.  The QRP  has been shown to
be NP-complete~\cite{Ullman00}, and it  consists of a conjunctive
query that must be answered in terms of views in  which the query and
the views are described using LAV with abstract relations.  This
problem is important in the context of data
integration~\cite{Chen05,JaudoinPRST05},  and query optimization and
data maintenance~\cite{AfratiLU07,levy:bucket}, and  several
approaches have been defined that scale up to a large number of
views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.

We propose a solution that extends the recent approach of Arvelo et
al.\ \cite{arvelo:aaai06} which is based on the efficient enumeration
of models of a propositional logic  theory. Similarly, given an
instance of SSP, a logical theory is constructed such  that each model
of the theory encodes a valid rewriting; the theory can be con-
structed in polynomial time from the instance of the SSP. All valid
rewritings  are obtained by enumerating all models of the theory in
linear time by using  off-the-shelf model enumeration solvers such as
c2d~\cite{c2d} and Relsat~\cite{relsat}; off-the-  seft SAT solvers
such as Minisat~\cite{minisat} can be used to produce all the minimal
rewritings. Thus, we exploit known properties of modern SAT solvers
that  implement non-chronological backtacking via clause learning,
caching of common  subproblems and heuristics, to provide an efficient
and scalable solution to SSP.

To summarize the contributions of our work are as follows:
\begin{itemize}
\item Service descriptions as LAV mappings on ontology concepts and user preferences as logic constraints, to provide an expressive description framework.
\item Formalization of SSP as the well-known QRP, and the extension of efficient and scalable existing solutions to exploit the properties of modern SAT solvers to construct a logical theory that encodes an instance of SSP, in a way that the theory models encode the valid rewritings and best models 
correspond to the optimal (best) rewritings.
\item Reduction of the combinatorial search, by efficiently enumerating the solutions.
\end{itemize}

The paper contains five more sections. The next section defines our
approach  and illustrates the SSP with typical examples. Section 3
describes our solution  and system architecture; Section 4 reports our
empirical results over different benchmark problems. Sections 5
summarizes existing approaches; and the paper  concludes with a final
discussion in Section 6.

\section{Service Selection Problem}
We consider service selection problems comprised of the
description $IS$ of the integration framework and of a
user request $R$ over the $IS$.

Our setting is based on the assuption that services are
semantically described in terms of a domain ontology. 
Formally, the integration framework corresponds to a tuple
$IS=\tup{D,S,M}$ where $D$ is the domain ontology, $S$ describes
the available services, and $M$ is the collection of LAV mappings
that define the services in $S$ as views on the ontology $D$. 
On the other hand, the user input $R=\tup{Q,P}$ consists of a
request for composition of services, described as a conjunctive
query $Q$ over the ontology, and a set $P$ of preferences for
the requested composition.

In the following subsections, we describe in detail all the
ingredients making up a service selection problem, and illustrate
the framework with several examples.

\subsection{Domain Ontology}

The domain ontology $D$ is a tuple $\tup{\sigma,A}$ where $\sigma$ is a
\emph{signature} for a logical language and $A$ is a collection of axioms
describing the ontology.
A signature $\sigma$ is a tuple $\tup{R_1^{r_1},\ldots,R_n^{r_n},c_1,\ldots,c_m}$
where each $R_i$ is a relational symbol of arity $r_i$,\footnote{We use
the notation $R^r$ to say that $R$ is a relational symbol of arity $r$.}
and each $c_j$ is a constant symbol.
The set of axioms describe the ontology by defining the relationships
between the ontology relations. % Not sure if talk about Datalog here...
For the present work, we only assume subsumption relationships among
concepts that are expressed with rules of the form:
\begin{equation}
\label{eq:orule}
R(\bar x,\bar y, \bar a)\ \orule\ P(\bar x, \bar b)\,,
\end{equation}
where $R$ and $P$ are predicates in $\sigma$ of appropriate arity,
$\bar x$ and $\bar y$ are lists of variables (repetitions allowed),
and $\bar a$ and $\bar b$ are lists of constant symbols in $\sigma$
(repetitions allowed). All these lists except $\bar x$ can be empty.

Although limited in appearance, rules of this form are quite expressive
as they allow us to specify diverse relationships among predicates.
For example,
\begin{itemize}
\item \textbf{Hierarchy of classes and subclasses} (or types and subtypes):
classes are specified with unary predicates. A subclass relantionship
among two classes is directly specified with a simple rule; e.g., the
rule $\textit{penguin}(x) \orule \textit{bird}(x)$ declares that every
penguin is a bird.
\item \textbf{Subrelations via specialization:} A subrelation of $R^r$
can be specified by constraining another relation $P^s$ ($r<s$). 
For example, the rule:
\begin{equation}
\label{eq:QE2}
\textit{descendant}(\textit{Elizabeth II},x)\ \orule\ \textit{noble}(x)
\end{equation}
tells that any descendant of Queen Elizabeth II is a noble.
\item \textbf{Indirect subsumption:} it is even possible to specify
a subrelation via another seemengly unrelated predicate. For example,
the rule:
\[ \textit{citizen-of}(x,\textit{Montreal})\ \orule\ \textit{lives-in}(x,\textit{Canada}) \]
says that when the second argument of `CitizenOf' is fixed to the
constant `Montreal', the tuples in the relation `CitizenOf' are
contained in the set of tuples in the relation `LivesIn' whose
second component is `Canada'.
\end{itemize}
However, we require that the \emph{dependency graph} for the ontology
to be a \emph{forest of trees}. The dependency graph $G(D)$ is a
labeled directed graph that is constructed as follows: the nodes
are the relational symbols in the signature $\sigma$, and there
is an edge $(R,P)$ in the graph iff there is an rule of the
form: \eqref{eq:orule} that is labeled with the \emph{bindings}
induced by the rule. For example, if $\textit{descendant}(x,y)$
and $\textit{noble}(z)$ are two predicates in the signature
and there is the rule \eqref{eq:QE2}, there edge from 
Descendant to Noble is labeled with the bindings
$\{x=\textit{Elizabeth II},y=z\}$.

\subsection{Services and LAV Mappings}

The available services are represented by means of another signature
$S=\tau=\tup{S_1^{s_1},\ldots,S_k^{s_k}}$, called the \emph{services signature},
that consists of relational symbols $S_i$ of arity $s_i$: each such symbol
$S_i$ represents a concrete service in the Semantic Web that ``offers''
some information.

The semantic description of services is expressed following the
LAV paradigm in terms of LAV mappings. These mappings describe the
services in terms of concepts in the domain ontology \cite{Ullman00}:
for each service $S_i$, there is a LAV mapping that describes $S_i$
as a \emph{conjunctive query} on the concepts in the ontology; 
input and output attributes of the service are also distinguished.
For example, a service $S(x,y)$ that returns information about
flights originating at a given US city can be described with the view:
\[ S(\$x,y)\ \qrule\ \flight(x,y), \UScity(x)\,, \]
where $flight^2$ and $uscity^1$ are relational symbols in the
ontology; the symbol \$ denotes that $x$ is an input attribute. 
The semantic interpretation of a mapping like this one enforce the following:
\begin{itemize}
\item the service represented by $S$ provides information in the
      form of tuples $(x,y)$,
\item each tuple $(x,y)$ returned by the service satisfies the rhs
      of the view; i.e., $\flight(x,y)$ and $\UScity(x)$, and
\item the views are not necessarily \emph{complete}; i.e., there may be
      other tuples $(x,y)$ that satisfy the rhs of the view but which
      are not available through the service represented by $S$.
\end{itemize}

The LAV approach is commonly used in integration systems because it permits
the scalability of the system as new services become available \cite{Ullman00}.
Under LAV, the appearance of a new service only causes the addition of a new
mapping describing the service in terms of the concepts in the ontology.
Under GAV, on the other hand, the ontology concepts are semantically described
using views in terms of the sources of information.
Thus, the extension or modification of the ontology is an easy task in GAV as
it only involves the addition or local modification of few descriptions
\cite{Ullman00}.

Therefore, the LAV approach is best suited for applications with a stable
ontology but with changing data sources whereas the GAV approach is best
suited for applications with stable data sources and a changing ontology.
In any case, the chosen approach is guarantee to efficiently scale with
respect to the changing characteristic but not with respect to the stable one;
e.g., a LAV description can be efficiently updated as data sources
are added or deleted, yet a slight revision of the ontology implies a
revision of all LAV mappings.

For the Semantic Web, we assume that the ontology of concepts is the stable
component. We believe that this is a reasonable assumption since, once a
common language is agreed upon to describe Web resources, the only changing
characteristic is the number and nature of resources which constantly pop
up or disappear from the Web.

Up to here, we have described all elements in the integration framework
$IS=\tup{D,S,M}$ where $D=\tup{\sigma,A}$ is an ontology of concepts, 
$S=\tau$ represent the available services in the Web and $M$ is a
collection of LAV mappings that describe the services in terms of the
concepts in the ontology.
The integration framework can be thought as the ``knowledge base'' (KB)
in a system designed for answering requests about the selection and
composition of Web services.
Ideally, the KB should support the efficient processing of user requests.

\subsection{User Requests}

A user request is formally described by a tuple $R=\tup{Q,P}$
where $Q$ is a conjunctive query in terms of concepts from the
ontology that expresses how the abstract concepts must be 
combined to resolve a given task, and a set $P$ of preferences.
For example, the query:
\[ Q(x)\ \qrule\ \flight(\LA,x), \flight(x,\PA) \]
can be used to find all cities on which a two-leg flight
from Los Angeles to Paris stop. Continuing with the above example,
this query can be answered using the view represented by $S$
as $I(x) \qrule S(\LA,x), S(x,\PA)$. This rewriting is correct yet
not necessarily complete as there may be two-leg flights from
Los Angeles to Paris that stop at non-US cities, e.g.,\ London.

The preferences are used to rank the collection of all valid
rewritings of the query. Once this ranking is obtained, the solution
of request $R$ is any best-ranked valid rewriting.
In this work, we consider a simple yet expressive system of
preferences in which each preference is given as a 
`soft constraint' on the rewriting of the query.
A soft constraint is a tuple $\pi=\tup{\varphi,c}$ where $\varphi$
is a propositional formula and $c$ is the cost associated with $\varphi$.
The idea is that each valid rewriting of the query is associated with
a cost equal to the sum of the costs of the preferences violated by
the rewriting, and that the best ranked rewriting is one that has minimum
cost among all valid rewriting.

It only remains to say over which language the formulas $\varphi$
are allowed to be and when a preference is violated by a rewriting.
The set of propositions for constructing preferences is
$\L(IS)=\{ R : R\in\sigma \} \cup \{S : S\in\tau \}$ that corresponds
to the relational symbols either in the ontology signature $\sigma$
or in the services signature $\tau$. Elements of $\L(IS)$ are
propositional symbols that should not be confused with their
relational interpretation in $IS$; indeed, if the reader
is more comfortable, he may use a different symbols altogether
such as $P_R$, $[R]$, or other.

The validity of a preference is established with respect to the
propositional model $\M(I)$ (truth assignment for the symbols in $\L(IS)$)
constructed from an answer $I(\bar x)$.
The model $\M(I)$ is defined by $S=\textbf{true}$ iff the service
$S$ appears in $I(\bar x)$, for $S\in\tau$, or by $R=\textbf{true}$
iff the concept $R$ appears in the unique path from a concept 
$R'$ to the root in the dependency graph $G(D)$ where $R'$ appears
in a service $S(\bar y)$ used in $I(\bar x)$, for $R\in\sigma$.
That is, the model makes true the service symbols used in
$I(\bar x)$, or the ontology symbols used in services in $I(\bar x)$,
or the ontology symbols that can be reached from the latter in
the dependency graph $G(D)$.
For example, the answer $I(x)\qrule S(\LA,x),S(x,\PA)$ defines
the model $\M(I)=\{S=\textbf{true},\flight=\textbf{true},\UScity=\textbf{true}\}$.

A preference $\varphi$ holds in an answer $I(\bar x)$ iff
$\M(I)\vDash\varphi$. This simple system of preferences permits
us to represent interesting cases such as:
\begin{itemize}
\item \textbf{Hard constraints:} any constraint $\pi=\tup{\varphi,\infty}$
is treated as a hard constraint that must be satisfied by every rewriting,
\item \textbf{QoS preferences:} this type of preferences can be used to
assign absolute quantities of reward/cost to single services as the one
used for integrated QoS parameters. For example, if each service $S_i$
is associated with a QoS reward of $r_i$, then the collection of preferences
$\pi_i=\tup{\neg S_i,-r_i}$ selects a valid rewriting with services that
have the highest combined QoS as the best rewriting,
\item \textbf{Conditional preferences:} a user's preference of the type
`if service $S$ is used, then service $R$ should be used as well' can be
modeled with the hard constraint $S \Rightarrow R$,
\item \textbf{Preferences of the type at-least-one:} a user's preference
of the type that at least one of the services $S_1,\ldots,S_n$ should be
used in the rewriting, can be modeled with the hard constraint $S_1\lor\cdots\lor S_n$, and
\item \textbf{Preferences of the type at-most-one:} a user's preference
of the type that at most one of the services $S_1,\ldots,S_n$ should be
used in the rewriting, can be modeled with the collection
$\{\neg S_i\lor\neg S_j: i\neq j\}$ of hard constraints.
\end{itemize}

\Omit{
$S$ is a signature  defined as a tuple $<$$S_1^{j1}$,$\dots$,
$S_q^{jn}$, $d_1$,$\dots$,$d_k$$>$, where $S_i^{ji}$ corresponds
to a service predicate of arity $ji$ and $d_i$ is a constant.  
Mappings in $M$ correspond to semantic descriptions expressed
following the  Local as View paradigm (LAV). We assume that
services are not complete, i.e., every tuple produced by the
source in the head of the rule must satisfy the body, but not
viceversa. LAV rules are commonly used in integration systems
in order to scale up to new sources, in contrast to the Global
As View paradigm (GAV) where domain ontology predicates are
defined in terms of conjunctive formulas in the service predicates.
In addition, in LAV mappings, services are treated independently
and the integration system easily scales up to the evolution of
the available services. However, the problem of rewriting a query
on the domain ontology into a set of queries on the available
services, corresponds to the problem of rewriting queries using
views which is known to be NP-complete~\cite{RajaramanSU95,Ullman00}.  
}

\subsection{Examples}

Let us consider a simple travel-information system that contains
information about flight and train trips between cities and information
about which cities are in the US. The domain ontology is comprised
of the predicates $\trip^2$, $\AAflight^2$, $\UAflight^2$, $\ATtrain^2$,
$\UPtrain^2$, $\flight^3$, $\train^3$ and $\UScity^1$, and the constants
$\AA$, $\UA$, $\AT$, $\UP$, $\LA$, $\NY$, $\PA$.
The first predicate relates cities $(x,y)$ if there is a direct
trip either by plane or train between them.
The predicates $\AAflight$ and $\UAflight$ relate cities that are
connected by direct flights either with American (AA) or United (UA).
The predicates $\ATtrain$ and $\UPtrain$ relate cities that are
connected by direct train trips either with Amtrak (AT) or Union
Pacific Railway (UP). The flight predicate relates $(x,y,t)$ whenever
there is a direct flight from $x$ to $y$ operated by airline $t$,
and similarly for $\train$. Finally, $\UScity$ indicates when a
given city is a US city or not.
There are various subsumption relationships among these predicates
that are captured by the following axioms:
\begin{alignat*}{1}
\flight(x,y,t)\  &\orule\ \trip(x,y)\,, \\
\train(x,y,t)\   &\orule\ \trip(x,y)\,.
\end{alignat*}
For the services, assume that the available data sources on the
Internet contain the following information:
\begin{enumerate}[--]
\item $\nationalFlight(x,y)$ relates two US cities that are connected by a direct flight,
\item $\onewayFlight(x,y)$ relates two cities that are connected by a one-way flight,
\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
\item $\toPA(x)$ tells if there is a direct flight from $x$ to Paris,
\item $\onestopPA(x,y)$ if there is a flight from $x$ to Paris with a stop at $y$, 
\item $\fromLA(x)$ tells if there is a flight from Los Angeles to $x$,
\item $\nationalTrain(x,y)$ relates US cities that are connected by a direct train, and
\item $\onewayTrain(x,y)$ relates US cities that are connected by a one-way train.
\end{enumerate}
These services are semantically described using the concepts in the ontology by
the following LAV mappings:
\begin{alignat*}{1}
\nationalFlight(x,y)\ &\qrule\ \flight(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
\AAflight(x,y)\       &\qrule\ \flight(x,y,\AA)\,, \\
\UAflight(x,y)\       &\qrule\ \flight(x,y,\UA)\,, \\
\onestop(x,z)\        &\qrule\ \flight(x,y,t),\,\flight(y,z,t)\,, \\
\toPA(x)\             &\qrule\ \flight(x,\PA,\AA)\,, \\
%\onestopPA(x,y)\      &\qrule\ \flight(x,y),\,\flight(y,\PA)\,, \\
\fromLA(x)\           &\qrule\ \flight(\LA,x,\UA)\,, \\
\nationalTrain(x,y)\  &\qrule\ \train(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
\ATtrain(x,y)\        &\qrule\ \train(x,y,\AT)\,, \\
\UPtrain(x,y)\        &\qrule\ \train(x,y,\UP)\,.
\end{alignat*}
Observe that each tuple produced by each service satisfies the semantic description
expressed by the body of the rule. For example, the tuples that satisfy the predicate
$\nationalFlight(x,y)$ also meet the conjunctive formula:
\[ \exists t(\flight(x,y,t)\ \land\ \UScity(x)\ \land\ \UScity(y))\,. \]
However, there may be tuples that satisfy this formula that are not
produced by the service represented by $\nationalFlight(x,y)$.
These elements make up the integration framework $IS$. In the rest of this section.
we give examples of possible user requests and their solutions.

Suppose that a user is interested in identifying the services able to retrieve
one-stop round trips from a US city $x$ to any city $y$ in the world. Notice
that the trip from $x$ to $y$ stops at a city $u$, that the back trip from $y$
to $v$ stops at a city $v$, and that $u$ may not be equal to $v$.
This request can be modeled as the following connjunctive query:
\[ Q(x,u,y,v)\ \qrule\ \UScity(x),\,\trip(x,u),\,\trip(u,y),\,\trip(y,v),\,\trip(v,x)\,. \]
Any rewriting of the ontology predicates in terms of the services correspond
to a \emph{composition of services} that implements the user request.
For example, the following rewriting is a valid solution to the request:
\begin{alignat*}{1}
I(x,u,y,v)\ \qrule\ &\nationalFlight(x,u),\,\toPA(u),\, \\
                    &\onewayFlight(\PA,v),\,\nationalFlight(v,x)\,. 
\end{alignat*}
But, the following two rewritings are not valid solutions:
\begin{alignat*}{1}
 I'(x,u,y,v)\  &\qrule\ \nationalFlight(x,u),\,\toPA(u),\,\fromLA(v),\,\nationalFlight(v,x)\,, \\
I''(x,u,y,v)\  &\qrule\ \onestop(x,y),\,\onewayFlight(y,v),\,\nationalFlight(v,x)\,.
\end{alignat*}
The first is not valid because it maps the query variable $y$ into two different
constants \PA\ and \LA\ that denote different cities, and the second rewriting is
not valid because the service $\onestop(x,y)$ does not receive as input, or produce
as output, the middle city $u$ where the flight from $x$ to $y$ must stop.

As shown, one can use a system for rewriting queries in terms of views
for computing solutions to the Service Selection Problem, as the valid
solutions correspond to the valid rewritings of the query. However, in
the presence of user preferences, the solutions must be ranked according
to the preferences and the best solutions should be returned.

To illustrate the use of preferences, consider the following request:
\[ Q(x,y)\ \qrule\ \trip(\LA,x),\,\trip(x,\NY),\,\trip(\NY,y),\,\trip(y,\LA) \]
that looks for round-trips between Los Angeles and New York such that
each direction is a one-stop trip. Observe that the query is posed in 
a way that there are no restrictions whatsoever on the use of planes or
trains for any leg of the trip. However, as is typical, users have 
preferences about using planes or trains.
For this example, we study four different scenarios for user's preferences
and show how to model them in the proposed framework:
\begin{enumerate}[P1.]
\item The user prefers to flight rather than to travel by train. This can be
      modeled by assigning a high reward to the symbol $\flight$. Likewise, a
      preference of trains over airplanes can be modeled by assigning a high
      reward to the symbol $\train$.
\item The user is indiferent with respect to trains or airplanes, yet
      s/he does not want to mix both. This preference is an at-most-one preference
      over the set $\{\flight,\train\}$ that corresponds to the formula
      $\neg\flight \lor \neg\train$ and a cost for the violation of the
      preference.
\item If the user travels by airplane, s/he prefers to always use the same
      airline (independently of the airline). This preference can be modeled
      with the formula $\neg\AAflight \lor \neg\UAflight$ together with a
      cost. Additionally, the other means of air transportation should be
      `disabled' since they may return flights operated by any airline;
      e.g., add the constraint $\neg\nationalFlight$ with a high cost.
\item Finally, if the user travels by airplane, s/he prefers to use $\UA$.
      This is a complex preference that can be modeled with the formula
\[ (\flight \Rightarrow \UAflight) \land (\neg\UAflight \lor \neg\AAflight)\,. \]
      The first part says that if a leg of the trip is done by plane, then $\UA$
      must be used, while the second part says that whenever $\UA$ is used,
      $\AA$ should not be used. Also, the services that do not guarantee
      airline operators should be disabled as in the previous case.
\end{enumerate}
All these preferences correspond to formulas over the propositional language
$\L(IS)$. The formulas for the last three preferences can be treated as
hard constraints if the are associated with infinite cost, or as soft
constraints meaning that the user prefers (but is not limited to) solutions
that do not violate the preferences.

\Omit{

Additionally, we allow users to specify their preferences. For example,
a user could prefer a round trip by either train or plane but not a round
trip that combines them. User preferences are specified as conditional
rules on service predicates. Thus, this user preference is expressed
with the following two rules:

\begin{alignat*}{1}
\onewayFlight(x,y)\ &\Rightarrow \neg \onewayTrain(y,x)\,. \\
\onewayTrain(x,y)\ &\Rightarrow \neg \onewayFlight(y,x)\,. 
\end{alignat*}

Based on the integration system and the user preferences, the following
composition corresponds to one rewriting of the user request in the
available services; the aggregated value of this solution is 8.0.
We use aggregated utility values and the user preferences to rank the
solutions and select the best one.

\begin{alignat*}{1}
 I(x,w,y,z)\ \qrule\  & \nationalFlight(x,w),\,\toPA(w),\,\\
                                & \onewayFlight(\PA,z),\,\nationalFlight(z,x)\,. 
\end{alignat*}

Additionally, the integration system is used to avoid the enumeration of
no valid solutions. For example, the following two rewritings are not valid:

\begin{alignat*}{1}
I'(x,w,y,z)\  \qrule\ &\nationalFlight(x,y),\,\toPA(y),\,\fromNY(z),\,\nationalFlight(z,x)\,. \\
I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\onewayFlight(y,z),\,\nationalFlight(z,x)\,.
\end{alignat*}

The first rewriting is not valid because it maps the query variable $y$
into constants \PA\ and \NY\ that denote different cities.
On the other hand, the rewriting $I''$ does not implement the query because
the service $\onestop(x,y)$ does not receive as input, or produce as output,
the middle city where the flight stops and thus is not possible to ensure
that this city is bound to the city $w$ that must returned by selected services.

These examples show that a proper user request rewriting must handle
constants such that not two different constants are mapped to each
other either directly or indirectly via transitivity, and that all
attributes that appear in a join or in the output need to be produced
by the selected  services. In addition, user preferences and QoS parameters
should be considered to discriminate among the valid solutions, the ones
that do not satisfy the user preferences or have low values of the
aggregated utility function. For example, our approach assigns a low score
to the following rewriting:

\begin{alignat*}{1}
 I'''(x,w,y,z)\ \qrule\ & \nationalFlight(x,w),\,\nationalFlight(w,y), \,\\
& \onewayFlight(y,z), \,\onewayTrain(z,y), \,\nationalFlight(y,w)\,. 
\end{alignat*}

Formally, the Service Selection Problem (SSP) is defined as follows:
given an integration system {\it IS=$<$G,S,M$>$}, a user request $Q$ and
a set of user preferences $UP$, a valid solution $Q_S$ is a rewriting of
$Q$ in the services in $S$, such that the result of evaluating $Q_S$ is
contained in the result of evaluating $Q$ i.e., $Q_S$ is a maximally-contained
rewriting of $Q$ ~\cite{RajaramanSU95,Ullman00}. Additional, the best solution
is a rewriting  that minimizes (or maximizes) the aggregation of utilities
and the user preferences.

%We consider service catalogs of the form $C=\tup{P,T}$ where $P$ is a set of predicates in $\sigma_G$ and $T=\{T_p\}_{p\in P}$ is a collection of tables comprised of tuples that satisfy the ontology predicates. In the context of the SSP, a service catalog $C$ is an idealized description of the output produced by user request implemented by services or data sources described as views in $M$. A user request $Q$ over $S$ is a conjunctive query of the form  \[ Q(\vvec{x})\ \qrule\ s_1(\vvec{x}_1),\,s_1(\vvec{x}_2),\,\ldots,\,s_m(\vvec{x}_m)\,, \] where $s_i\in S$, $\vvec{x}$ is a vector of variables, and each $\vvec{x}_i$ is a vector of variables and constants. The result of $Q$ over $C$, denoted as $Q(C)$, is the table with $|\vvec{x}|$ columns that result of the projection of the relational join $\join\!\!\{T_{s_i}\}_{i=1}^n$ over $\vvec{x}$. The atoms in the body of $Q$ are called the (sub)goals of $Q$, and the variables in the head of $Q$ are called distinguished.

%A concrete service is described as a view $V$ over $C$ that, following LAV, is a query over $S$. Given a catalog $C$, a workflow $W$ and a collection of $n$ views $E=\tup{\{V_i\}_{i=1}^n,\{E_i\}_{i=1}^n}$, we are asked to find all the tuples in $W(C)$ that can be obtained from the views in $E$. That is, we need to find all the 
%\emph{instantiations}
%\[ R(\vvec{x})\ \qrule\ V_{i_1}(\vvec{x}_1),\,V_{i_2}(\vvec{x}_2),\,\ldots,\,V_{i_{m_i}}(\vvec{x}_{m_i}) \]
%such that $R(E) \subseteq Q(C)$.
%A Service Selection Problem is a tuple $\tup{S,Q,\{V_i\}_i}$ where $S$
%is a set of predicates, $Q$ is a user request  over
%$S$, and $\{V_i\}_i$ is a collection of views that define the services.
%We assume \emph{safe} problems in the sense that all variables mentioned in the
%head of the user request (resp.\ in the head of each view) appear in the body of the
%user request (resp.\ in the body of each view).
%Further, we only deal with SSPs with no arithmetic predicates inside the user request
%or views. An instantiation $R$ is \emph{valid} if for all catalogs $C=\tup{S,T}$
%and extensions $\{E_i\}_i$, $R(E)\subseteq Q(C)$. A collection $\R$ of valid
%instantiations is a solution if for all service catalogs $C=\tup{S,T}$ and
%extensions $\{E_i\}_i$, there is no $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(C)$.
}

\section{Solution and System Architecture}

We extend the \mcdsat\ system of Arvelo et al.\ \cite{arvelo:aaai06} for
Query Rewriting Problems (QRP). An instance of a QRP consists of a
collection of views and a query on abstract concepts. 
The problem consists is in ``re-writing'' the query in terms of the
views so that each tuple produced by the rewriting is a tuple of the
solution.
\mcdsat\ reduces the QRP into the problem of finding the models of
a propositional logic theory.
The theory constructed by \mcdsat\ satisfies the following properties:
(1)~there is a 1-1 correspondence between the valid rewritings of the
query and the models of the logical theory, (2)~given a model of the theory,
one can recover the corresponding rewriting in linear time, and
(3)~the theory can be constructed in polynomial time from the instance 
of the QRP.
Once the logical theory is constructed, one can be interested in finding
all minimal rewritings of the query as done in data integration systems
with incomplete sources \cite{Ullman00}, or just a model of the theory
\cite{refs}. For the former, off-the-shelf model enumeration solvers
such as c2d \cite{c2d} and Relsat \cite{relsat} can be used, while 
off-the-shelf SAT solvers such as Minisat \cite{minisat} can be used in the
latter case.

In this section, we have just enough space to explain how the logical
theory constructed by \mcdsat\ can be extended to capture the features
associated with SSPs that are not present in QRPs; namely, handling
constant symbols and input and output attributes, the ontology of
concepts with subsumption relationships, and user's preferences. The
result is an extended theory whose models are in correspondence with
the valid solutions of the SSP and, in the presence of preferences,
whose \emph{best models} are in correspondence with the best-ranked
valid solutions of the SSP.

\subsubsection{Constant Symbols and Input Output Attributes}

\mcdsat\ did not provide support for constant symbols in the language
or service  input and output attributes, yet incorporating these
functionality is straightforward.  First to consider constant symbols,
basically, one only has to track the unification of variables with
constants using new propositional symbols, and to propagate such
unifications transitively using propositional implications to avoiding
the unification of different constant symbols. Second, to respect the
input and output attributes of the available services, one has to
ensure that input attributes of a service unify with constant symbols
in the user query or unify with output attributes returned by other
services in the rewriting. These modifications involve the addition of
a relatively small number of new propositional symbols and clauses to
the CNF generated by \mcdsat.

\subsubsection{Ontology}

The subsumption relationships in the ontology make the rewriting 
process more complex as now one need to consider unification of
predicate symbols of different name and arity. Indeed, consider
the following four ontology concepts, where $a$, $b$, $c$ and $d$
are constant symbols, and $x$, $y$ and $z$ are variables:
\begin{alignat*}{1}
R(a,y)\   &\orule\ T(c,y)\,, \\
P(b,y,z)\ &\orule\ R(a,y)\,, \\
M(a,x)\   &\orule\ N(d,x)\,, \\
P(d,x,z)\ &\orule\ M(a,x)\,,
\end{alignat*}
and the following user request $Q(x,y)$ with the services 
$S1$, $S2$ and $S3$:
\begin{alignat*}{1}
Q(x,y)\   &\qrule\ T(z,y),\,N(z,x)\,, \\
S_1(y)\   &\qrule\ R(a,y)\,, \\
S_2(x,z)\ &\qrule\ N(z,x)\,, \\
S_3(x,z)\ &\qrule\ P(d,x,z)\,.
\end{alignat*}
Then, the system must be able to infer that the query can be 
rewritten as $I(x,y) \qrule S_1(y),S_2(x,c)$ since $R(a,y)$
unifies with $T(z,y)$ producing the \emph{binding} $\{z=c\}$,
and $S_2(x,z)$ unifies with $N(z,x)$ and becomes $S_2(x,c)$
once the binding is propagated. On the other hand, the query
cannot be rewritten as $I(x,y)\qrule S_1(y),S_3(x,z)$ because
$R(a,y)$ unifies with $T(z,y)$ with the binding $\{z=c\}$,
$P(d,x,z)$ unifies with $N(z,x)$ with the binding $\{z=d\}$,
and these bindings are non-unifiable since constants
are assumed to denote unique objects.

We incorporate the subsumtion relation into \mcdsat\ by means
of the dependency graph $G(D)$. Once the graph is built using
the subsumption rules, its transitive closure is computed along
with the bindings associated with each edge: edges generated by
the transitive closure have labels that correspond to the union
of the bindings along the edges that generate this edge (if the
set of bindings is inconsistent, then the label is assigned the
binding \textbf{false}).
These labels are unique and well defined as $G(D)$ is assumed
to be a forest of trees. Once the transitive closure $G(D)^*$
is computed, all edges with inconsistent labels can be dropped.
The transitive closure is then used to extend the rules in the
logical theory that permit the cover of relational symbols in
the query with symbols in the views: a predicate $P$ is allowed
to cover a predicate $R$ whenever there is an edge from $P$ to
$R$ in $G(D)^*$, yet if this covering becomes active, the 
bindings associated with it become active as well.

\subsubsection{Preferences}

In order to account for the preferences, we shall use the
concepts literal-ranking functions and best-ranked models for
propositonal logic. A literal ranking function $r$ is a function
that assign ranks (real numbers) to propositional literals.
Given a literal-ranking function $r$, the rank $r(\omega)$ of
a model $\omega$ is the aggregation of the ranks for each literal
made true by the model: $r(\omega)=\sum_{\omega\vDash\ell} r(\ell)$.
Thus, the models can be ordered by their rank and the best-ranked
models are the models with minimum rank.
Some model enumerators like c2d can be used to compute all the
best-ranked models of a propositional theory.
Likewise, Weighted-Max-SAT solvers such as MiniMaxSAT~\cite{HerasetalJAIR2008}
can be used to find a best ranked model of a propositional theory.

For SSPs, we accomodate the preferences by using a suitably defined
literal-ranking function $r^*$ and by computing best-ranked models
\cite{darwiche:weighted}.
First, a new propositional variable is created for each relational
symbol in the ontology and services signatures along with clauses
that turn this proposition true whenever the corresponding symbols
become active (true). Second, for each preference $\pi=\tup{\varphi,c}$,
a new propositional symbol $p_\pi$ is created along with the
formula $p_\pi \Leftrightarrow \varphi$. Thus, $p_\pi$ is true
iff $\varphi$ is satisfied in the model (rewriting).
Finally, the literal-ranking function $r^*$ is defined as 
$r^*(\neg p_\pi)=c$ for each such preference.
Clearly, the rank of a model correspond to the sum of the costs
associated with the preferences violated by the model, and thus
a best-ranked model correspond to a rewriting of minimum regret.

\subsection{System Architecture}

We define an architecture for solving SSPs that is comprised of a Catalog of
service descriptions, an ontology reasoner, the Encoder, the best model Finder,
and the Decoder.
Figure~\ref{fig:architecture} depicts the overall architecture of the system.
In this framework, an instance of SSP consists of an integration framework $IS$
and a user request $R$ made of a conjunctive query $Q$ and a collection of
preferences.
The Catalog of the system is populated with the components of the integration system,
i.e., the domain ontology including the subsumption rules, the services and the LAV
mappings between them.

The input instance of the SSP is then translated into a CNF theory and a
literal-ranking function $r^*$ by the Encoder module. The Encoder makes use
of the transitive closure $G(D)^*$ that is calculated by the ontology reasoner
together with the bindings associated with the edges. Once the theory is obtained,
it is fed to a best-model finder. Once the finder return a best model, it is
given to the Decoder that then re-construct the rewriting from the model.

\begin{figure}[t]
\centering
\includegraphics[width=.8\textwidth]{architecture}
\caption{System Architecture}
\label{fig:architecture}
\end{figure}

\Omit{
Once the CNF is generated, the compiler c2d, an off-the-shelf
component, compiles the CNF formula into d-DNNF.  A d-DNNF  theory is constructed from literals using only conjunctions  and disjunctions \cite{barwise:handbook}; for each conjunction its variables are pairwise disjoint, and for each disjunction, the disjuncts are pairwise logically contradictory; see Fig.~\ref{fig:dnnf} for an example.  In our approach, a d-DNNF theory is a compact representation of
all the rewritings. That is, the Finder generates in a backtrack-free
manner all the rewritings of the user request. If the user is interested in
a best rewriting, then the Finder computes it in
linear time in the size of the d-DNNF. If the user is interested in the 
all the best rewritings, these are computed in linear time in the
number of them. This computation transforms the DAG of the d-DNNF into a arithmetic
circuit by replacing the AND nodes with `+' and the OR nodes with `min'.
The literal ranking function assigns
values to the leaves of the circuit that are propagated to the
root in linear time. The value of the root is the rank of the
best model \cite{darwiche:weighted}.
Finally, if the user is interested in all rewritings,
these are also computed in linear time in the number of them.
In the latter two cases, if such number is exponential (in the size of the
input), the enumeration of the instantiations is also exponential but
this complexity is intrinsic to the problem and thus cannot be avoided.

It is important to remark
that the compilation process performed by the compiler c2d, needs to be performed only once as it does not
depend on the value of the QoS parameters or the user preferences. Thus, even if the compilation happens to be costly in terms of time, this cost can be amortized since the resulting
d-DNNF can be used to find best rewritings with respect to multiple values
of the QoS parameters or weights of user preferences.
Finally, the Decoder translates the best model returned by the Finder into
a rewriting that solves the SSP.

In order to make our results more accessible to a general audience, we decided
to present the experimental results in the next section, and leave the related work, and formal
and theoretical results for the following sections. In this way, the interested
reader may skip the formal details of the approach on a first reading.
}

\section{Experimental Results}

We conducted an empirical analysis on three benchmarks.
All the experiments were run on a desktop machine with an Intel
Core 2 Duo 2GHz CPU and 4Gb of memory, and the time was measured
with the Unix time command. We used the
compiler c2d, \footnote{\url{http://reasoning.cs.ucla.edu/c2d}} to compile 
the CNF formula into d-DNNF\cite{darwiche:weighted}; c2d makes use of
modern SAT techniques such as conflict-directed backtracking, clause learning
and caching\cite{darwiche:compiler} to enumerate the theory models from the d-DNNF.

The objective of the experiment is to assess the performance of the proposal
on varying conditions. The main benefit of our approach is that one can compile
the logical theory for a problem instance and then calculate all the instantiations,
or the best ones, any number of times. The cost model for finding best instantiations
can be changed with no need to recompile the logical theory. Therefore,
the time complexity of our approach is basically the time to encode the SSP
as a CNF plus the time to find the (best) models and the time to decode the
models. The times to encode and decode are negligible compared to the time to
enumerate the models. Because of this, we focus on the time to enumerate the
models of the benchmark problems. All the CNF theories are of low width.

The first benchmark consists of problems for air-travel queries.
Service views are of the form $V_i(x,y)\qrule\ \flight(x,y,\AL_i)$
where $\AL_i$ is a constant that denotes the name of an airline, 
and $\flight(x,y,\AL_i)$ relates the cities $x$ and $y$ such that
there is a flight between $x$ and $y$ served by $\AL_i$. This 
service is assumed to return all flights between two cities with an
specific airline. The user request has the form 
\[ Q(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,z),\,\flight(x_1,x_2,z),\,\ldots,\,\flight(x_n,\NY,z)\,. \]

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{plots/plot1}
\caption{Compilation times for experiments I and II for different
number of goals and different number of views. The plots are in
logarithmic scale, and the time is in seconds.}
\label{fig:plot1}
\end{figure}

The benchmark includes user requests with 2 to 5 sub-goals and
sets of 10 to 100 service views. The results for the compilation are
shown in panel (a) of Fig.~\ref{fig:plot1}. This is a plot in logarithmic
scale that suggests a sub-exponential behavior. In any case, the results
show good performance since realistic instances of the problem (sets of 100
airlines with 5-stop flights) can be compiled in 328 seconds.
The size in disk of the d-DNNF for 100 airlines and 5-stop flights is 3.4Mb.
On this d-DNNF, the best model can be computed in 0.29 seconds,
and the enumeration of all models in 0.47 seconds.

In an attempt to induce an exponential growth in the compilation time,
in the second benchmark we add a second service view for each airline.
This modification increases the number of valid rewritings from linear
to exponential since each leg of a flight can now be instantiated by two
services and thus a flight with $n$ legs may have up to $2^n$
rewritings.
We ran the compiler for instances comprising the same number of user request
sub-goals and total number of service views. The results plotted in
logarithmic scale are shown in panel (b) of Fig.~\ref{fig:plot1}.

These tests show good performance for this type of problems, but they do not
involve service views with multiple sub-goals. We therefore designed a
third experiment that consists unstructured, randomly generated instances.
Each instance contains three variables per domain ontology concept in the user request, ten distinct
variables and ten distinct constants, six sub-goals in the user request,
2 to 5 sub-goals in the service views, and a varying number of service views.
The chance that an argument of a domain ontology concept is bound to a constant is 50\%.
The results are depicted  in panel (a) of Fig.~\ref{fig:plot4}.
The compilation time for these instances does not grow monotonically
since they are randomly generated. The same happens for the size of
the theories and the number of models. For example, the d-DNNF for 
a problem with 45 views each with 5 sub-goals is of size 5.1Mb and has
$1.26\times 10^8$ models. The time to find the best model for
this d-DNNF is 0.46 seconds while the time to enumerate all models
is about 17 hours.

\begin{figure}[t]
\centering
\includegraphics[width=.9\textwidth]{plots/plot4}
\caption{Compilation times for experiment III for different number of goals
and different number of views. Time for enumerating rewritings for experiment IV. The plots are in logarithmic scale, and the time
is in seconds. }
\label{fig:plot4}
\end{figure}

Finally, to study the effects of the ontology size, we considered the first benchmark and built 
ontologies where the predicates $\flight(x,y,z)$ and $\UScity(x)$ were leaves; ontologies were binary trees comprised of 2, 3, 4, 5, 6 and 7 levels.  The user request has the form: 
\[ Q(x_1,x_2,x_3,x_4)\ \qrule\  \UScity(x_1),\,\trip(x_1,x_2),\,\trip(x_2,x_3),\,\trip(x_3,x_4),\,\trip(x_4,x_1).\]
where $trip(x,y)$ was the root of the ontologies.  The results are reported in panel 
(b) of Fig.~\ref{fig:plot4}. The time to enumerate all the rewritings suggests a sub-exponential behavior, while the time to find the best models does not grow monotonically. On the other hand, 
the size of the theories and the number of models grow exponentially.  For example, the d-DNNF for 
this user request and the ontology with 7 levels is of size 1.7Mb and has
$2.6\times 10^8$ models. ÊThe time to enumerate all models is about 147 hours.

These are preliminary experiments, yet the results show that the
proposed approach efficiently scales for problems with several goals
and views. We believe that these results are encouraging and motivate
us to continue this research. We plan to conduct additional experiments
with other types and sizes of user requests and, when possible, compare with
other approaches.

\section{Related Work}
The problem of selecting the services that satisfy a user request is a
combinatorial optimization problem and several heuristics have been
proposed to find a good solution in a reasonably short period of
time~\cite{alrifaiR09,berardi08,myoung08,kuterG09,lecue09,rahmani08,sohrabiM09,Hiroshi2008}.

In a series of papers, Berardi and
others~\cite{berardi05,berardi08,berardi06}  describe services and
user requests in terms of deterministic finite-state machines that are
encoded using Description Logic theories whose models correspond to
solutions of the problem; properties of Description Logics reasoning
methods are not exploited.

Ko et al.~\cite{myoung08} propose a
constraint-based approach that encodes the nonfunctional permissible
values as a set of constraints whose violation needs to
be minimized. Alrifai and Risse~\cite{alrifaiR09} develop a two-fold
solution that uses a hybrid integer programming algorithm to find the
decomposition of global QoS into local constraints, and then, selects
the services that best meet the local constraints.

Recently, two
planning-based approaches have been proposed. Kuter
and Golbeck~\cite{kuterG09} extend the SHOP2 planning algorithm to
select the trustworthy composition of services that implement a given
OWL-S process model, while Sohrabi and McIlraith~\cite{sohrabiM09}
propose a HTN planning-based solution where user preference metrics
and domain regulations are used to guide the planner into the space
of relevant compositions. Finally, L\'ecu\'e \cite{lecue09}  develops
a genetic-based algorithm to identify the composition of services
that best meet the quality criteria for a set of QoS parameters.

These existing solutions are able to scale up to relatively large
number of abstract concepts. In addition to scalability, our approach
provides a more expressive framework where services are semantically
described in terms of domain ontology concepts, user preferences
restrict the space of solutions, and ontology relationships augment
the space of possible solutions. Finally, our approach is sound and
complete.

\Omit{
\subsection{Query Rewriting}

A number of algorithms have been developed to find the rewritings of a given query;
the most prominent being the Bucket algorithm \cite{levy:bucket}, the Inverse Rules
algorithm \cite{pods:DuschkaG97,Qian96}, the \minicon\ algorithm \cite{pottinger:minicon},
and the \mcdsat\ algorithm \cite{arvelo:aaai06}.
Generally, query rewriting algorithms work in two phases. During the first phase,
the algorithms identify the views that rewrite at least one sub-goal of the query, and
during the second, these partial rewritings are combined into complete rewritings.
The main difference between the algorithms is the criteria used to choose the relevant
views to reduce the space of non-useful rewritings.

The Bucket algorithm reduces the number of possibilities by just considering each
sub-goal in the query in isolation, and selecting the views that are able to produce
at least the attributes projected by the query. Since the attributes involved in the
joins in the query are not verified, a large number of rewritings comprised of Cartesian
products may be generated.

The Inverse Rules algorithm constructs a set of rules that invert the view definitions
and establish how to compute tuples for the database relations from the tuples of the
views. Similarly to the Bucket algorithm, it can produce a large number of non-useful
rewritings.

The \minicon\ algorithm overcomes the limitations of the previous algorithms by identifying
only views that rewrite a set of the sub-goals of the query, and that can be combined with
the rest of the sub-goals. The key idea is to identify the mappings between the variables in
each sub-goal to the variables in one or more sub-goals in the views, in a way
that join variables in the query are mapped to join variables in the body of a view or to
the distinguished variables of the view. Mappings between variables and sub-goals are
represented in the so-called MiniCon Descriptions (MCDs) \cite{pottinger:minicon}.

Finally, the \mcdsat\ algorithm is able to identify the query rewritings of a query by
translating the problem of rewriting into the problem of enumerating the models of a
propositional theory whose models are in correspondence with the rewritings of the query.
The algorithm exploits the properties of d-DNNFs to efficiently enumerate the models of
the theory.
The \mcdsat\ algorithm has demonstrated to scale better than the \minicon\ algorithm over
a large number of benchmarks often showing performance improvements of several orders of
magnitude. However, the \mcdsat\ algorithm was not designed for rewriting problems
involving explicit constants, nor to compute the best rewritings with respect to a given
utility function or user preferences. We propose an extended encoding that 
overcomes these limitations and apply the encoding to SSP.

\subsection{Knowledge Compilation}

Knowledge compilation is the area in AI concerned with the problem of mapping
logical theories into suitable fragments that make certain desired operations
tractable \cite{cadoli:compilation}. Different compilation languages have been
defined, for instance, Ordered Binary Decision Diagrams (OBDDs) \cite{bryant:obdd},
Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation
Normal Form (DNNF) \cite{darwiche:map}. In this work, we make use of the properties
of the deterministic DNNFs (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable
and efficient solution to the Service Selection Problem.

A Negation Normal Form (NNF) theory is constructed from literals using only conjunctions
and disjunctions \cite{barwise:handbook}. An NNF is said to be decomposable (DNNF) \cite{darwiche:d-dnnfs} if for each conjunction, its variables are pairwise disjoint. A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each disjunction, the disjuncts are pairwise logically contradictory. A d-DNNF supports model counting in polytime in the size of its DAG, and model
enumeration in polytime in the size of the output.
Furthermore, given a literal ranking function $r$, one can compute the rank
of the best model in polytime for DNNFs \cite{darwiche:weighted}.

The fragments DNNF and d-DNNF are universal yet translating a CNF theory 
to DNNF format has an exponential cost in the worst case. This translation
is referred to as compilation in the field. There is a publicly available
compiler, called c2d,\footnote{\url{http://reasoning.cs.ucla.edu/c2d}}
that performs this compilation process and that makes
use of modern SAT techniques such as conflict-directed backtracking,
clause learning and caching \cite{darwiche:compiler}.
This compiler incurs in the worst case in exponential space in a parameter
called the width of the decomposition tree that is related to the ``connectivity''
of the CNF theory. However, in our experiments, the CNF theories that are
compiled are of low width.

\section{SSP Formalization into Logical Theories}
We use an approach similar to the one described in \cite{arvelo:aaai06} to
encode the SSP. We have limited space to make a comprehensive description of
the logical theory so the reader is referred there for details and formal results.

We identify service compositions  by enumerating the models of a logical
theory $\Delta=\Delta_{com}\cup\Delta_{id}^1\cup\cdots\Delta_{id}^N$
where $\Delta_{com}$ specifies how to combine $N$ independent copies theories
$\Delta_{id}$ that cover the goals in a user request $Q$.
Each $\Delta^i_{id}$ is a tagged copy $\Delta_{id}$ in which each literal
$\ell$ is tagged as $\ell^i$.
The Instantiation Description (ID) theory $\Delta_{id}$ consists of different
groups of clauses that guarantees that its models are in correspondence with
partial rewritings, while the theory $\Delta_{com}$ contains additional
clauses to guarantee a sound and complete combination of partial rewritings
into a complete rewriting.
The ID theory $\Delta_{id}$ consists of the following variables:
\begin{enumerate}[--]
\item $\{v_0,\ldots,v_n\}$ to indicate which source view $V_i$ is used, or $v_0$ to indicate the null ID.
\item $\{g_1,\ldots,g_m\}$ to indicate the goals covered by the view.
\item $\{z_{j,k,i}\}$ to indicate that the $j$th goal in $Q$ is covered by the $k$th goal in $V_i$.
\item $\{t_{x,y}\}$ to indicate that the variable/constant $x$ in $Q$ is mapped into the
      variable/constant $y$ in the view.
\end{enumerate}
The ranges of the indices for the $z$ and $t$ variables depend on the problem.

The following clauses encode the SSP problem in terms of the LAV mappings in $M$.
Rajaraman et al.\ \cite{RajaramanSU95} showed that for queries without negation or
arithmetic comparisons, but with constants, and $m$ goals and $k$ variables in the
head of the user request, it is enough to consider rewritings of length at most $N=m+k$.
\begin{enumerate}[C10.]
\item[C1.] (At least one view is used): $\bigvee_{i=0}^n v_i$.
\item[C2.] (At most one view is used): $\neg v_i\lor\neg v_j$ for $i\neq j$.
\item[C3.] (Null view equals null): $v_0 \Rightarrow \neg g_j$ for $1\leq j\leq m$.
\item[C4.] (Views are useful): $v_i \Rightarrow \bigvee_{j=1}^m g_j$ for $1\leq i\leq n$.
\item[C5.] (Subgoals covered at most once): $z_{j,k,i} \Rightarrow \neg z_{j,l,i}$ for appropriate $i,j,k,l$.
\item[C6.] (Scope of views): $v_i \Rightarrow \neg g_j$ for goals that cannot be covered by $V_i$.
\item[C7.] (Consistency): $v_i \land g_j \Leftrightarrow \bigvee z_{j,k,i}$ for appropriate $i,j,k$.
\item[C8.] (Dead variables): $v_i \Rightarrow \neg t_{x,y}$ for all $x,y$ with $y\notin V_i$.
\item[C9.] (1-1 for $\exists$ vars): $v_i \land t_{x,y} \Rightarrow \neg t_{x,y'}$ for all
           existential variables $y,y'\in V_i$.
\item[C10.] (Distinguished): $v_i \Rightarrow \neg t_{x,y}$ for distinguished $x$ and existential
            $y\in V_i$.
\item[C11.] (Existential): $v_i\land t_{x,y}\Rightarrow g_j$ for exist.\ $y\in V_i$ and
            goals $g_j$ that contain $x$.
\item[C12.] (Match): $v_i\land z_{j,k,i} \Rightarrow t_{x,y}$ for all $x,y$ that must match
            if $g_j$ is covered by goal $k$ in $V_i$.
\item[C13.] (If all vars in $V_i$ are distinguished, it covers only one goal):
            $v_i \land g_j \Rightarrow \neg g_k$ for appropriate views $v_i$.
\end{enumerate}
These clauses are the same clauses used by \mcdsat\ to encode QRPs \cite{arvelo:aaai06}.
In order to properly manage constant symbols, the clauses must be enhanced with:
\begin{enumerate}[C10.]
\item[C14.] (Direct inconsistency 1): $t_{x,A} \Rightarrow \neg t_{x,B}$.
\item[C15.] (Direct inconsistency 2): $t_{A,x} \Rightarrow \neg t_{B,x}$.
\item[C16.] (Direct inconsistency 3): $\neg t_{A,B}$.
\item[C17.] (Transitivity 1): $v_i\land t_{A,y}\land t_{x,y}\land t_{x,z}\Rightarrow t_{A,z}$.
\item[C18.] (Transitivity 2): $v_i\land t_{y,A}\land t_{y,x}\land t_{z,x}\Rightarrow t_{z,A}$.
\end{enumerate}
Recall that, as shown in Section~2, the main issue when handling
constants is to be sure that not two different constants are mapped
into each other either directly or indirectly.
Clauses C14--C16 prune direct inconsistent mappings, while the last
two clauses implement a restricted propagation of mappings that prune
indirect inconsistencies.
Likewise, the theory $\Delta_{com}$ that specifies complete
instantiations contains the following clauses:
\begin{enumerate}[C10.]
\item[C19.] (Cover all goals): $\bigvee_{j=1}^m g^i_j$ for $1\leq i\leq N$.
\item[C20.] (Disjunctive cover): $g^i_k \Rightarrow \neg g^j_k$ for $i\neq j$.
\item[C21.] (Prune symmetries): $g^i_j \Rightarrow \bigvee_{k=1}^{j-1} g^{i-1}_k$
            for $1\leq j\leq m$ and $1\leq i\leq N$.
\item[C22.] (Direct inconsistency 4): $t^i_{x,A} \Rightarrow \neg t^j_{x,B}$.
\end{enumerate}
These clauses provide a sound and complete characterization of SSPs in the
sense that their models are in correspondence with the rewritings of the user request.

For the QoS parameters, we assume a simple additive aggregation model in
which each view $V_i$ is associated with a cost $c(V_i)$ (negative if utility),
and a complete rewriting with the sum of the cost of its views. User preferences rules are transformed into propositional logic formulas whose cost is a high number.
An optimal or best solution is one with minimum cost, and the
optimal value of the SSP is the cost of an optimal rewriting.
A SSP always has a well-defined optimal value (if there are no rewritings,
its cost is $\infty$), but it may have multiple best rewritings.
The SSP with costs consists in finding all optimal rewritings or
one rewriting, this depends on the particular application.
In our formulation, this can be done from the d-DNNF that encodes $\Delta$
using the literal ranking function $r$ that assigns $r(\ell)=c(V_i)$ if
$\ell=v_i$, and $r(\ell)=0$ if $\ell\notin\{v_1,\ldots,v_n\}$.
}
\section{Conclusions and Future Work}
We have shown how the propositional theory used in \mcdsat\  for
computing rewritings of queries can be extended to efficiently encode
SSP; thus, by exploiting known properties of modern SAT solvers, we
provide an efficient and scalable solution to SSP. In addition, we
provide an expressive framework by defining LAV mappings to semantically
describe services, and by expressing user requests as a conjunctive
query over the domain ontology concepts and a set of preferences that
restrict the SSP solutions.

The experimental results show that the approach can be applied to
real-sized problems; however, if d-DNNF format is used, the approach
is only possible when the compilation of the CNF theory into d-DNNF
format succeeds. Thus, to overcome this limitation, in the future, we
plan to study the performance and scalability of our approach in
other off-the-self model enumeration and SAT solvers such as Relsat,
Minisat and MiniMaxSAT. We also are interested in using other target
compilation languages like Ordered Binary Decision Diagrams.

\bibliographystyle{abbrv}
\bibliography{ref}

\end{document}

### 2010-06-28 publicaciones/iswc10/paperISWC10v2.tex
spellchecking del paper v2

@@
 instance of SSP, a logical theory is constructed such  that each model
-of the theory encodes a valid rewriting; the theory can be con-
-structed in polynomial time from the instance of the SSP. All valid
+of the theory encodes a valid rewriting; the theory can be constructed
+in polynomial time from the instance of the SSP. All valid
 rewritings  are obtained by enumerating all models of the theory in
 linear time by using  off-the-shelf model enumeration solvers such as
-c2d~\cite{c2d} and Relsat~\cite{relsat}; off-the-  seft SAT solvers
+c2d~\cite{c2d} and Relsat~\cite{relsat}; off-the-shelf SAT solvers
 such as Minisat~\cite{minisat} can be used to produce all the minimal
 rewritings. Thus, we exploit known properties of modern SAT solvers
-that  implement non-chronological backtacking via clause learning,
+that  implement non-chronological backtracking via clause learning,
 caching of common  subproblems and heuristics, to provide an efficient
@@
 
-Our setting is based on the assuption that services are
+Our setting is based on the assumption that services are
 semantically described in terms of a domain ontology. 
@@
 \item \textbf{Hierarchy of classes and subclasses} (or types and subtypes):
-classes are specified with unary predicates. A subclass relantionship
+classes are specified with unary predicates. A subclass relationship
 among two classes is directly specified with a simple rule; e.g., the
@@
 \item \textbf{Indirect subsumption:} it is even possible to specify
-a subrelation via another seemengly unrelated predicate. For example,
+a subrelation via another seemingly unrelated predicate. For example,
 the rule:
@@
 source in the head of the rule must satisfy the body, but not
-viceversa. LAV rules are commonly used in integration systems
+vice versa. LAV rules are commonly used in integration systems
 in order to scale up to new sources, in contrast to the Global
@@
 These elements make up the integration framework $IS$. In the rest of this section.
-we give examples of possible user requests and their solutions.
+We give examples of possible user requests and their solutions.
 
@@
 to $v$ stops at a city $v$, and that $u$ may not be equal to $v$.
-This request can be modeled as the following connjunctive query:
+This request can be modeled as the following conjunctive query:
 \[ Q(x,u,y,v)\ \qrule\ \UScity(x),\,\trip(x,u),\,\trip(u,y),\,\trip(y,v),\,\trip(v,x)\,. \]
@@
       reward to the symbol $\train$.
-\item The user is indiferent with respect to trains or airplanes, yet
+\item The user is indifferent with respect to trains or airplanes, yet
       s/he does not want to mix both. This preference is an at-most-one preference
@@
 
-We incorporate the subsumtion relation into \mcdsat\ by means
+We incorporate the subsumption relation into \mcdsat\ by means
 of the dependency graph $G(D)$. Once the graph is built using
@@
 concepts literal-ranking functions and best-ranked models for
-propositonal logic. A literal ranking function $r$ is a function
+propositional logic. A literal ranking function $r$ is a function
 that assign ranks (real numbers) to propositional literals.
@@
 
-For SSPs, we accomodate the preferences by using a suitably defined
+For SSPs, we accommodate the preferences by using a suitably defined
 literal-ranking function $r^*$ and by computing best-ranked models
@@
 this user request and the ontology with 7 levels is of size 1.7Mb and has
-$2.6\times 10^8$ models. ÊThe time to enumerate all models is about 147 hours.
+$2.6\times 10^8$ models. The time to enumerate all models is about 147 hours.
 

### 2010-06-28 tesis/benchmarks/omcdsat/1/generator.sh
fix de generator y runner de omcdsat1

@@ -2,6 +2,6 @@
 
 GEN=./generator.py
 
-for levels in `seq 7`; do
+for levels in `seq 12`; do
         $GEN $levels > casos/${levels}levels.ontology
 done

### 2010-06-28 tesis/benchmarks/omcdsat/1/runner.sh

@@ -1,12 +1,12 @@
 #!/bin/bash
 
 INDIR=casos
-OUT=omcdsat_0.raw
+OUT=omcdsat_1.raw
 TIME="timeout 600 /usr/bin/time -apo ${OUT}"
 GENONTOLOGIES=/home/daniel/proyectos/postgrado/tesis/src/omcdsat/mcdsat/GenOntologies.py
 OMCDSAT=/home/daniel/proyectos/postgrado/tesis/src/omcdsat/mcdsat/mcdsat
 
-echo '################################################################################' | tee -a ${OUT_MMS}
+echo '################################################################################' | tee -a ${OUT}
 
 for levels in `seq 2 7`; do
     ONTOLOGY=casos/${levels}levels.ontology

### 2010-06-28 tesis/benchmarks/preferences/0/runner.sh
preference 0

@@ -1,0 +1,24 @@
+#!/bin/bash
+
+INDIR=casos
+OUTFILE=preferences0.out
+
+TIME="/usr/bin/time -apo ${OUTFILE}"
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
+
+echo '########' | tee -a ${OUTFILE}
+
+for vistas in `seq 5 5 40`; do
+    for pasos in `seq 2 5`; do
+        for copias in 2; do
+            VIEWSFILE=${INDIR}/views-${vistas}-${pasos}-${copias}.txt
+            QUERYFILE=${INDIR}/query-${vistas}-${pasos}-${copias}.txt
+
+            echo ${vistas} ${pasos} ${copias} | tee -a ${OUTFILE}
+
+            for i in `seq 1`; do #n times to get average
+                ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
+            done
+        done
+    done
+done

### 2010-06-28 tesis/benchmarks/preferences/0/runner.sh
preferences funcionando casi perfecto

@@ -4,21 +4,18 @@ INDIR=casos
 OUTFILE=preferences0.out
 
 TIME="/usr/bin/time -apo ${OUTFILE}"
-PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/compileonly
+PROGRAM=/home/daniel/proyectos/postgrado/tesis/src/imcdsat/mcdsat/mcdsat
 
-echo '########' | tee -a ${OUTFILE}
+#echo '########' | tee -a ${OUTFILE}
 
-for vistas in `seq 5 5 40`; do
-    for pasos in `seq 2 5`; do
-        for copias in 2; do
-            VIEWSFILE=${INDIR}/views-${vistas}-${pasos}-${copias}.txt
-            QUERYFILE=${INDIR}/query-${vistas}-${pasos}-${copias}.txt
+for vistas in `seq 5 5 5`; do
+    for pasos in `seq 2 2`; do
+            VIEWSFILE=${INDIR}/views-${vistas}-${pasos}.txt
+            QUERYFILE=${INDIR}/query-${vistas}-${pasos}.txt
 
-            echo ${vistas} ${pasos} ${copias} | tee -a ${OUTFILE}
+            #echo ${vistas} ${pasos} ${copias} | tee -a ${OUTFILE}
 
-            for i in `seq 1`; do #n times to get average
-                ${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE} > /dev/null
-            done
-        done
+            #${TIME} ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE}
+            ${PROGRAM} RW ${VIEWSFILE} ${QUERYFILE}
     done
 done

### 2010-06-28 tesis/src/imcdsat/mcdsat/Traductor/Traductor3.py

@@ -163,17 +163,61 @@ def clausulasCombinarMCD(transf, n, q, vistas):
 
 
 def imprimirCopias(variables, clausulas, clausulas2, numCopias, transf, archSalida):
-    arch = file(archSalida, 'w+')
-    numVars = len(variables)*numCopias
-    numCl = len(clausulas)*numCopias + len(clausulas2)
-    arch.write('p cnf '+ str(numVars) + ' ' + str(numCl) + '\n')
+    global varsV
 
+    arch = file(archSalida, 'w+')
+    cl_prefs = []
+
+    #INI BENCHMARK PREFERENCIAS
+    #esto supone el -1 al final
+
+    #variables P
+    P_ini = len(variables)*numCopias+1
+    P_idx = len(variables)*numCopias
+
+    for v in varsV:
+        if v < 0:
+            continue
+
+        vv = transf.variables[varsV[v]]
+        P_idx += 1
+
+        for numCopia in xrange(numCopias):
+            cl_prefs.append("%d %d 0" % (-(transf.n*numCopia + vv), P_idx))
+
+        cl_prefs.append("%d %s 0" % (-P_idx, " ".join([str(transf.n*numCopia + vv) for numCopia in xrange(numCopias)])))
+
+    Q_ini = P_idx+1
+    Q_idx = P_idx
+
+    for ii, vi in enumerate(varsV):
+        for ij, vj in enumerate(varsV):
+            if ii==ij or vi < 0 or vj < 0:
+                continue
+
+            Q_idx += 1
+
+            cl_prefs.append("%d %d %d 0" % (-Q_idx, -(P_ini+ii), -(P_ini+ij)))
+            cl_prefs.append("%d %d 0" % (-(P_ini+ii), Q_idx))
+            cl_prefs.append("%d %d 0" % (-(P_ini+ij), Q_idx))
+
+    for i in xrange(Q_ini, Q_idx+1):
+        cl_prefs.append("%d 0" % (i))
+    #FIN BENCHMARK PREFERENCIAS
+
+    #numVars = len(variables)*numCopias
+    numVars = Q_idx
+    numCl = len(clausulas)*numCopias + len(clausulas2) + len(cl_prefs)
+    arch.write('p cnf '+ str(numVars) + ' ' + str(numCl) + '\n')
 
     for cls in clausulas2:
         arch.write(cls)
 
     for numCopia in xrange(numCopias):
[…]

### 2010-06-29 publicaciones/iswc10/paperISWC10-v3.tex
paperISWC10-v3

--- 2010-06-28 publicaciones/iswc10/paperISWC10v2.tex
@@
 % Services
-\newcommand{\nationalFlight}{\textit{national-flight}}
-\newcommand{\onewayFlight}{\textit{one-way-flight}}
-\newcommand{\nationalTrain}{\textit{national-train}}
+\newcommand{\nationaltlight}{\textit{national-flight}}
+\newcommand{\onewayflight}{\textit{one-way-flight}}
+\newcommand{\nationaltrain}{\textit{national-train}}
 \newcommand{\onewayTrain}{\textit{one-way-train}}
@@
 \begin{abstract}
-Given the tremendous amount of potential Semantic Web 
-Services that can be created from online sources by using existing 
-annotation tools, expressive formalisms and efficient and scalable approaches to solve the Service Selection Problem (SSP) are required. In this paper, we propose a framework that 
-is firmly grounded on logic. The framework adopts the Local-As-View 
-approach~\cite{levy:bucket} that semantically describes services in terms of concepts in a domain 
-ontology. A user request is comprised of a conjunctive query on concepts 
-from the domain ontology and a set of preferences; the query expresses 
-the abstract concepts that must be combined to solve the query, and 
-the preferences are used to rank the solutions. Relations in the domain 
-ontology are considered to augment the space of available services. Using 
-this representation, the problem of service selection is cast as a problem 
-of query rewriting, from the area of integration systems. Then, 
-building on related work, we devise an encoding of SSP as a logical theory 
-whose models are in correspondence with the service rewritings of the 
-user query, and in presence of preferences, whose best models are in 
-correspondence with the best-ranked valid solutions. Thus, by exploiting 
-known properties of modern SAT solvers, we provide an efficient and 
-scalable solution to SSP. The approach provides the basis to represent 
-a large number of real-world situations; additionally, it does not 
-only scale up to large instances, as the experimental results show, 
-but is also sound and complete since, being founded on logic, is 
-amenable to formal analysis.
+Given the large number of Semantic Web Services that can be created from
+online sources by using existing annotation tools, expressive formalisms and
+efficient and scalable approaches to solve the service selection problem are
+required in order to make this services widely available to the users.
+In this paper, we propose such a framework that is grounded on logic and
+the Local-As-View approach for describing instances of the service selection problem.
+In our approach, Web services are semantically described using LAV mappings
+in terms of generic concepts from an ontology, while the user requests correspond
+to conjunctive queries on the generic concepts, that must be combined to solve
+a task, plus a set of preferences that are used to rank the possible solutions.
+The LAV formulation allows us to cast the service selection problem as a
+query rewriting problem that must consider the relationships among the concepts
+in the ontology and the ranks induced by the preferences of the user.
+Then, building on related work, we devise an encoding of the resulting
+query rewriting problem as a logical theory whose models are in correspondence
+with the solutions of the user request, and in presence of preferences, whose
+best models are in correspondence with the best-ranked solutions.
+Thus, by exploiting known properties of modern SAT solvers, we provide an
+efficient and scalable solution to the service selection problem.
+The approach provides the basis to represent a large number of real-world
+situations and interesting user requests.
 \end{abstract}
@@
 \section{Introduction}
+
 Existing Web infrastructures support the publication and access to a
-tremendous amount of Web data sources, some of which can be labeled
+tremendous amount of Web data sources, some of which that can be labeled
 and converted into Semantic Web Services by using existing annotation
-tools like the one proposed  by Ambite et al. \
-\cite{AmbiteISWC09}. Once this huge dataset of Semantic Web  Services
-become available, users more than ever will require techniques to
-effectively select the services that meet their requirements. In order
-to achieve  this goal, functional and non-functional properties of the
-services should be  precisely described, as well as criteria to select
-and rank the solutions that best meet user requests.   In this paper,
-we extend an approach traditionally used in the area of data
-integration to solve the Service Selection Problem. First, concepts
-from a domain ontology provide a precise description of the semantics
-of available sources. These  descriptions correspond to views on
-concepts of the domain ontology and follow  the Local-As-View approach
-(LAV)~\cite{levy:bucket}. LAV mappings are able to  scale to service
-datasets that constantly change, in contrast to Global-As-View
-definitions  (GAV), where generic concepts are defined in terms of the
-available sources and  cannot be easily modified when sources
-constantly change. A user request is  modeled as a conjunctive query
-on the ontology concepts, and user preferences  are associated with a
-cost and constraint the solutions of the query.  Based on this
-integration framework, we define the problem of selecting the
-services that best meet the user request, as the problem of
-enumerating and  ranking the service rewritings of the user
-query. Each rewriting is associated  with a cost equal to the cost of
-the preferences violated by the rewriting, and the best ranked
-rewriting is one that has minimum cost among all valid
-rewritings. Knowledge encoded in the domain ontology is used to expand
-the  space of rewritings. Thus, our proposed service selection
-approach is cast into  the well-known Query Rewriting Problem (QRP)
-for LAV which is central to  integration systems.  The QRP  has been shown to
-be NP-complete~\cite{Ullman00}, and it  consists of a conjunctive
-query that must be answered in terms of views in  which the query and
-the views are described using LAV with abstract relations.  This
-problem is important in the context of data
-integration~\cite{Chen05,JaudoinPRST05},  and query optimization and
-data maintenance~\cite{AfratiLU07,levy:bucket}, and  several
-approaches have been defined that scale up to a large number of
-views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
-
-We propose a solution that extends the recent approach of Arvelo et
-al.\ \cite{arvelo:aaai06} which is based on the efficient enumeration
-of models of a propositional logic  theory. Similarly, given an
-instance of SSP, a logical theory is constructed such  that each model
-of the theory encodes a valid rewriting; the theory can be constructed
-in polynomial time from the instance of the SSP. All valid
-rewritings  are obtained by enumerating all models of the theory in
-linear time by using  off-the-shelf model enumeration solvers such as
-c2d~\cite{c2d} and Relsat~\cite{relsat}; off-the-shelf SAT solvers
-such as Minisat~\cite{minisat} can be used to produce all the minimal
-rewritings. Thus, we exploit known properties of modern SAT solvers
-that  implement non-chronological backtracking via clause learning,
-caching of common  subproblems and heuristics, to provide an efficient
-and scalable solution to SSP.
-
-To summarize the contributions of our work are as follows:
-\begin{itemize}
-\item Service descriptions as LAV mappings on ontology concepts and user preferences as logic constraints, to provide an expressive description framework.
-\item Formalization of SSP as the well-known QRP, and the extension of efficient and scalable existing solutions to exploit the properties of modern SAT solvers to construct a logical theory that encodes an instance of SSP, in a way that the theory models encode the valid rewritings and best models
-correspond to the optimal (best) rewritings.
-\item Reduction of the combinatorial search, by efficiently enumerating the solutions.
-\end{itemize}
-
-The paper contains five more sections. The next section defines our
-approach  and illustrates the SSP with typical examples. Section 3
-describes our solution  and system architecture; Section 4 reports our
-empirical results over different benchmark problems. Sections 5
-summarizes existing approaches; and the paper  concludes with a final
-discussion in Section 6.
+tools like the one proposed by Ambite et al.\ \cite{AmbiteISWC09}.
+Once a large dataset of Semantic Web Services become available, users
+will require techniques to effectively select the services that meet
+their requirements.
+In order to achieve this goal, the services in the dataset must be
+tagged with their functional and non-functional properties, and
+the users' preferences and requirements must be formally described
+as well.
+In this paper, we extend an approach traditionally used in the area of
+data integration to solve the problem of selecting the best services
+that meet a user request, a problem that we call in this paper the
+Service Selection Problem (SSP).
+
+As in other approaches, we use domain ontologies for describing
+the services in the dataset, yet we differ in how the services are described.
+In this paper, we use the recent approach of Ambite et al.\ \cite{AmbiteISWC09}
+that describes services as views on the generic concepts of the ontology
+following the Local-As-View (LAV) approach that is widely used in integration
+systems~\cite{levy:bucket}, instead of the traditional Global-As-View (GAV)
+approach where the generic concepts are expressed in terms of the services.
+The adoption of the LAV approach instead of GAV is not accidental.
+LAV descriptions are tailored towards systems with constantly changing
+datasets and a relatively stable set of generic concepts, while GAV
+descriptions are tailored towards systems with a constantly changing
+set of generic concepts but a relatively stable dataset of services.
+
+As it is shown below, LAV descriptions are achieved with mappings
+that define the services as conjunctive queries involving the generic
+concepts in the ontology. Thus, every time that a service changes or
+a new one becomes available, only a tiny fraction of the mappings
+must be updated, usually just one mapping.
+Likewise, user requests can be modeled as conjunctive queries over the
+generic concepts in a way that the SSP can be cast as the problem
+of rewriting a query in terms of a set of views, the so-called Query
+Rewriting Problem (QRP) that is well-known in the area of data
+integration~\cite{Chen05,JaudoinPRST05}, query optimization and data
+maintenance~\cite{AfratiLU07,levy:bucket}, and for which several
+scalable approaches had been
+proposed~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
+Furthermore, users' preferences and constraints on the set of the
+selected services to serve a given request are supported by the
+adoption of a simple yet expressive language for preferences. 
+These preferences and constraints further refine and rank the
+the set of valid rewritings of the posed query, in a way that
+the best solution to the SSP correspond to the best-ranked 
+valid rewritings of the corresponding QRP.
+
+Our solution for the SSP extends the recent approach of Arvelo et
+al.~\cite{arvelo:aaai06} for QRPs that is based on the efficient 
+enumeration of models for a propositional logic theory.
+In our case, an input instance of the SSP is converted into an instance
+of a QRP with preferences and constraints that is then translated
+into a logical theory such that its best models are in correspondence
+with the solutions of the SSP. These translations, from SSP to QRP to
+logic, are performed efficiently, in (low) polynomial time, while the
+best models are found using off-the-shelf SAT tools.
+Thus, we are able to exploit the befits of modern SAT techniques such
+as conflict-directed backtracking via clause learning and caching and
+decomposition of common subproblems~\cite{beame:understanding} to
+perform the necessary search in the combinatorial space of solutions.
+
+In summary, we make the following crisp contributions to the problem of 
+service selection and composition: (1)~propose the use of LAV mappings
+as an scalable solution for describing the continuously changing set
+of available Web Services in terms of stable generic concepts,
+(2)~propose a simple yet powerful language for expressing preferences
+and constraints on the valid solutions of the SSP, 
+(3)~describe how to transform the SSP to the well-known QRP extended with
+preferences and constraints, and
+(4)~describe how to change an efficient and scalable solution to the QRP,
+based on propositional logic and SAT tools, to handle preferences and constraints.
+The rest of this paper is as follows. The next two sections describe the
+SSP and the language of preferences and constraints, and the proposed solution to the SSP.
+Then, we present preliminary experimental results, related work 
+and end with a brief discussion.
 
 \section{Service Selection Problem}
-We consider service selection problems comprised of the
-description $IS$ of the integration framework and of a
-user request $R$ over the $IS$.
-
-Our setting is based on the assumption that services are
-semantically described in terms of a domain ontology. 
-Formally, the integration framework corresponds to a tuple
-$IS=\tup{D,S,M}$ where $D$ is the domain ontology, $S$ describes
-the available services, and $M$ is the collection of LAV mappings
-that define the services in $S$ as views on the ontology $D$. 
-On the other hand, the user input $R=\tup{Q,P}$ consists of a
-request for composition of services, described as a conjunctive
-query $Q$ over the ontology, and a set $P$ of preferences for
-the requested composition.
-
-In the following subsections, we describe in detail all the
-ingredients making up a service selection problem, and illustrate
-the framework with several examples.
+
+An SSP consists of a description $IS$ of the integration framework
+and a user request $R$.
+Formally, the integration framework is a tuple $IS=\tup{D,S,M}$
+where $D$ is the ontology of generic concepts, $S$ is the set
+of available services, and $M$ is the collection of LAV mappings that
+semantically describe the services in terms of the ontology.
+On the other hand, a user request is a tuple $R=\tup{Q,P}$
+that is made of a query $Q$ expressed as a conjunctive query over
+the generic concepts and a set of preferences $P$.
+In the following, we describe all these elements in detail and
+illustrate the framework through a number of examples.
 
@@
 
-The domain ontology $D$ is a tuple $\tup{\sigma,A}$ where $\sigma$ is a
-\emph{signature} for a logical language and $A$ is a collection of axioms
-describing the ontology.
-A signature $\sigma$ is a tuple $\tup{R_1^{r_1},\ldots,R_n^{r_n},c_1,\ldots,c_m}$
-where each $R_i$ is a relational symbol of arity $r_i$,\footnote{We use
-the notation $R^r$ to say that $R$ is a relational symbol of arity $r$.}
-and each $c_j$ is a constant symbol.
-The set of axioms describe the ontology by defining the relationships
-between the ontology relations. % Not sure if talk about Datalog here...
-For the present work, we only assume subsumption relationships among
-concepts that are expressed with rules of the form:
+The domain ontology $D$ is a tuple $\tup{\sigma,A}$ where $\sigma$
+is a \emph{signature} for a logical language and $A$ is a collection
+of axioms describing the ontology.
+A signature $\sigma$ is a set of relational and constant symbols from
+which logical formulas can be constructed; it corresponds to a tuple
+$\tup{R_1^{r_1},\ldots,R_n^{r_n},c_1,\ldots,c_m}$ where each $R_i$
+is a relational symbol of arity $r_i$,\footnote{We use the notation
+$R^r$ to say that $R$ is a relational symbol of arity $r$.} and each
+$c_j$ is a constant symbol.
+The axioms describe the ontology by defining the relationships between
+the ontology concepts.
+For the present work, we only consider subsumption relationships
+between concepts that are expressed with rules of the form:
 \begin{equation}
@@
 \end{equation}
-where $R$ and $P$ are predicates in $\sigma$ of appropriate arity,
+where $R$ and $P$ are predicates in $\sigma$ (of appropriate arity),
 $\bar x$ and $\bar y$ are lists of variables (repetitions allowed),
-and $\bar a$ and $\bar b$ are lists of constant symbols in $\sigma$
-(repetitions allowed). All these lists except $\bar x$ can be empty.
-
-Although limited in appearance, rules of this form are quite expressive
-as they allow us to specify diverse relationships among predicates.
-For example,
+and $\bar a$ and $\bar b$ are lists of constant symbols (repetitions allowed).
+All these lists may be empty except $\bar x\cup \bar b$.
+
+Although limited in appearance, subsumption rules are quite expressive
+as they allow us to specify diverse relationships between concepts; e.g.,
 \begin{itemize}
 \item \textbf{Hierarchy of classes and subclasses} (or types and subtypes):
-classes are specified with unary predicates. A subclass relationship
-among two classes is directly specified with a simple rule; e.g., the
-rule $\textit{penguin}(x) \orule \textit{bird}(x)$ declares that every
-penguin is a bird.
+classes are specified with unary predicates. A subclass relationship can
+be specified with a simple rule; e.g.,
+$\textit{penguin}(x) \orule \textit{bird}(x)$ tells that penguins are birds.
 \item \textbf{Subrelations via specialization:} A subrelation of $R^r$
@@
 \end{equation}
-tells that any descendant of Queen Elizabeth II is a noble.
+tells that the descendants of Queen Elizabeth II are noble.
 \item \textbf{Indirect subsumption:} it is even possible to specify
@@
 \[ \textit{citizen-of}(x,\textit{Montreal})\ \orule\ \textit{lives-in}(x,\textit{Canada}) \]
-says that when the second argument of `CitizenOf' is fixed to the
-constant `Montreal', the tuples in the relation `CitizenOf' are
-contained in the set of tuples in the relation `LivesIn' whose
-second component is `Canada'.
+says that when the second argument of `\textit{citizen-of}' is fixed to the
+constant `\textit{Montreal}', the tuples in the relation `\textit{citizen-of}'
+are contained in the set of tuples in the relation `\textit{lives-in}' whose
+second component is `\textit{Canada}'.
 \end{itemize}
-However, we require that the \emph{dependency graph} for the ontology
-to be a \emph{forest of trees}. The dependency graph $G(D)$ is a
-labeled directed graph that is constructed as follows: the nodes
-are the relational symbols in the signature $\sigma$, and there
-is an edge $(R,P)$ in the graph iff there is an rule of the
-form: \eqref{eq:orule} that is labeled with the \emph{bindings}
-induced by the rule. For example, if $\textit{descendant}(x,y)$
-and $\textit{noble}(z)$ are two predicates in the signature
-and there is the rule \eqref{eq:QE2}, there edge from 
-Descendant to Noble is labeled with the bindings
-$\{x=\textit{Elizabeth II},y=z\}$.
-
-\subsection{Services and LAV Mappings}
+However, we require that the \emph{dependency graph} $G(D)$ of the ontology
+to be a \emph{forest of trees}. The dependency graph is a labeled directed
+graph that is constructed as follows: the nodes of the graph are the relational
+symbols in the signature, and there is an edge $(R,P)$ in the graph iff
+there is an rule of the form \eqref{eq:orule}. The edge is labeled with
+the \emph{bindings} induced by the rule; e.g., if $\textit{descendant}(x,y)$
+and $\textit{noble}(z)$ are two predicates in the signature and there
+is the rule \eqref{eq:QE2}, then there is an edge from \textit{descendant}
+to \textit{noble} labeled with the bindings $\{x=\textit{Elizabeth II},y=z\}$.
+
+\subsection{Services and Mappings}
 
 The available services are represented by means of another signature
-$S=\tau=\tup{S_1^{s_1},\ldots,S_k^{s_k}}$, called the \emph{services signature},
-that consists of relational symbols $S_i$ of arity $s_i$: each such symbol
-$S_i$ represents a concrete service in the Semantic Web that ``offers''
-some information.
+$S=\tau=\tup{S_1^{s_1},\ldots,S_k^{s_k}}$ called the \emph{services signature},
+where each symbol $S_i$ represents a concrete service in the Internet
+that ``offers'' some information.
 
 The semantic description of services is expressed following the
-LAV paradigm in terms of LAV mappings. These mappings describe the
+LAV paradigm in terms of mappings. These mappings describe the
 services in terms of concepts in the domain ontology \cite{Ullman00}:
-for each service $S_i$, there is a LAV mapping that describes $S_i$
-as a \emph{conjunctive query} on the concepts in the ontology; 
-input and output attributes of the service are also distinguished.
+for each service $S_i$, there is a mapping that describes $S_i$
+as a \emph{conjunctive query} on the concepts in the ontology that
+also distinguish input and output attributes of the service.
 For example, a service $S(x,y)$ that returns information about
@@
 where $flight^2$ and $uscity^1$ are relational symbols in the
-ontology; the symbol \$ denotes that $x$ is an input attribute. 
+ontology. The symbol `\$' denotes that $x$ is an input attribute. 
 The semantic interpretation of a mapping like this one enforce the following:
@@
       form of tuples $(x,y)$,
+\item the service is called with $x$ as input attribute and returns $(x,y)$,
 \item each tuple $(x,y)$ returned by the service satisfies the rhs
@@
       other tuples $(x,y)$ that satisfy the rhs of the view but which
-      are not available through the service represented by $S$.
+      are not available through $S$.
 \end{itemize}
@@
 \cite{Ullman00}.
-
 Therefore, the LAV approach is best suited for applications with a stable
@@
 suited for applications with stable data sources and a changing ontology.
-In any case, the chosen approach is guarantee to efficiently scale with
-respect to the changing characteristic but not with respect to the stable one;
-e.g., a LAV description can be efficiently updated as data sources
-are added or deleted, yet a slight revision of the ontology implies a
-revision of all LAV mappings.
-
 For the Semantic Web, we assume that the ontology of concepts is the stable
@@
 $S=\tau$ represent the available services in the Web and $M$ is a
-collection of LAV mappings that describe the services in terms of the
-concepts in the ontology.
+collection of LAV mappings describing the services in terms of the
+concepts in $D$.
 The integration framework can be thought as the ``knowledge base'' (KB)
@@
 
-A user request is formally described by a tuple $R=\tup{Q,P}$
-where $Q$ is a conjunctive query in terms of concepts from the
-ontology that expresses how the abstract concepts must be 
-combined to resolve a given task, and a set $P$ of preferences.
+A user request is a tuple $R=\tup{Q,P}$ where $Q$ is a conjunctive query
+in terms of concepts in the ontology that describes how these concepts
+must be combined to resolve a given task, and a set $P$ of preferences.
 For example, the query:
-\[ Q(x)\ \qrule\ \flight(\LA,x), \flight(x,\PA) \]
+\[ Q(x)\ \qrule\ \flight(\LA,x), \flight(x,\PA). \]
 can be used to find all cities on which a two-leg flight
-from Los Angeles to Paris stop. Continuing with the above example,
+from Los Angeles to Paris stop. Continuing with the previous example,
 this query can be answered using the view represented by $S$
@@
 
-The preferences are used to rank the collection of all valid
-rewritings of the query. Once this ranking is obtained, the solution
-of request $R$ is any best-ranked valid rewriting.
-In this work, we consider a simple yet expressive system of
-preferences in which each preference is given as a 
-`soft constraint' on the rewriting of the query.
+The preferences are used to rank the collection of valid rewritings.
+Once this ranking is obtained, the solution for the request $R$ is
+any best-ranked and valid rewriting.
+In this work, we consider a simple yet expressive language for 
+preferences in which each preference is treated as a ``soft constraint''
+on the rewriting of the query.
 A soft constraint is a tuple $\pi=\tup{\varphi,c}$ where $\varphi$
 is a propositional formula and $c$ is the cost associated with $\varphi$.
-The idea is that each valid rewriting of the query is associated with
-a cost equal to the sum of the costs of the preferences violated by
-the rewriting, and that the best ranked rewriting is one that has minimum
-cost among all valid rewriting.
-
-It only remains to say over which language the formulas $\varphi$
-are allowed to be and when a preference is violated by a rewriting.
+The idea is that each valid rewriting is associated with a cost equal to
+the sum of the costs of the preferences violated by the rewriting,
+and that the best ranked rewriting is one that has minimum cost among
+all valid ones.
+
+It only remains to say what type of propositional formulas $\varphi$
+are allowed and when a preference is violated by a rewriting.
 The set of propositions for constructing preferences is
 $\L(IS)=\{ R : R\in\sigma \} \cup \{S : S\in\tau \}$ that corresponds
-to the relational symbols either in the ontology signature $\sigma$
-or in the services signature $\tau$. Elements of $\L(IS)$ are
-propositional symbols that should not be confused with their
-relational interpretation in $IS$; indeed, if the reader
-is more comfortable, he may use a different symbols altogether
-such as $P_R$, $[R]$, or other.
-
-The validity of a preference is established with respect to the
-propositional model $\M(I)$ (truth assignment for the symbols in $\L(IS)$)
-constructed from an answer $I(\bar x)$.
-The model $\M(I)$ is defined by $S=\textbf{true}$ iff the service
-$S$ appears in $I(\bar x)$, for $S\in\tau$, or by $R=\textbf{true}$
-iff the concept $R$ appears in the unique path from a concept 
-$R'$ to the root in the dependency graph $G(D)$ where $R'$ appears
-in a service $S(\bar y)$ used in $I(\bar x)$, for $R\in\sigma$.
-That is, the model makes true the service symbols used in
-$I(\bar x)$, or the ontology symbols used in services in $I(\bar x)$,
-or the ontology symbols that can be reached from the latter in
-the dependency graph $G(D)$.
-For example, the answer $I(x)\qrule S(\LA,x),S(x,\PA)$ defines
+to the relational symbols either in the ontology signature or in the
+services signature. Elements of $\L(IS)$ are propositional symbols that
+should not be confused with their relational interpretation in $IS$;
+indeed, if the reader is more comfortable, he may think on a different
+symbol altogether such as $P_R$, $[R]$, or other.
+The validity of a preference is defined with respect to the propositional
+model $\M(I)$ (truth assignment for the symbols in $\L(IS)$) constructed
+from a valid rewriting $I(\bar x)$:
+$\M(I)$ is defined by $S=\textbf{true}$ iff the service $S$ appears in
+$I(\bar x)$, for $S\in\tau$, or by $R=\textbf{true}$ iff the concept $R$
+appears in the unique path from a concept $R'$ to the root in the dependency
+graph $G(D)$ where $R'$ appears in a service $S(\bar y)$ used in $I(\bar x)$,
+for $R\in\sigma$.
+That is, the model makes true the service symbols used in $I(\bar x)$, or
+the ontology symbols used in services in $I(\bar x)$, or the ontology symbols
+that can be reached from the latter in the dependency graph $G(D)$.
+For example, the rewriting $I(x)\qrule S(\LA,x),S(x,\PA)$ defines
 the model $\M(I)=\{S=\textbf{true},\flight=\textbf{true},\UScity=\textbf{true}\}$.
 
-A preference $\varphi$ holds in an answer $I(\bar x)$ iff
-$\M(I)\vDash\varphi$. This simple system of preferences permits
-us to represent interesting cases such as:
+A preference $\varphi$ holds in an answer $I(\bar x)$ iff $\M(I)\vDash\varphi$.
+This simple language permits us to express interesting preferences such as:
 \begin{itemize}
 \item \textbf{Hard constraints:} any constraint $\pi=\tup{\varphi,\infty}$
-is treated as a hard constraint that must be satisfied by every rewriting,
+is treated as a hard constraint that must be satisfied by every rewriting;
+i.e., if the best rewriting has infinite cost, then it violates a hard constraint,
 \item \textbf{QoS preferences:} this type of preferences can be used to
@@
 $\pi_i=\tup{\neg S_i,-r_i}$ selects a valid rewriting with services that
-have the highest combined QoS as the best rewriting,
+have the highest combined QoS,
 \item \textbf{Conditional preferences:} a user's preference of the type
@@
 
-Let us consider a simple travel-information system that contains
-information about flight and train trips between cities and information
-about which cities are in the US. The domain ontology is comprised
-of the predicates $\trip^2$, $\AAflight^2$, $\UAflight^2$, $\ATtrain^2$,
-$\UPtrain^2$, $\flight^3$, $\train^3$ and $\UScity^1$, and the constants
-$\AA$, $\UA$, $\AT$, $\UP$, $\LA$, $\NY$, $\PA$.
+Consider a travel-information system that contains information about
+flight and train trips between cities and information about which cities
+are in the US. The domain ontology is comprised of the predicates
+$\trip^2$, $\flight^3$, $\train^3$ and $\UScity^1$, and the constants
+%$\AAflight^2$, $\UAflight^2$, $\ATtrain^2$, $\UPtrain^2$,
+$\AA$, $\UA$, $\AT$, $\UP$, $\LA$, $\NY$, and $\PA$.
 The first predicate relates cities $(x,y)$ if there is a direct
 trip either by plane or train between them.
-The predicates $\AAflight$ and $\UAflight$ relate cities that are
-connected by direct flights either with American (AA) or United (UA).
-The predicates $\ATtrain$ and $\UPtrain$ relate cities that are
-connected by direct train trips either with Amtrak (AT) or Union
-Pacific Railway (UP). The flight predicate relates $(x,y,t)$ whenever
-there is a direct flight from $x$ to $y$ operated by airline $t$,
-and similarly for $\train$. Finally, $\UScity$ indicates when a
-given city is a US city or not.
-There are various subsumption relationships among these predicates
-that are captured by the following axioms:
+%The predicates $\AAflight$ and $\UAflight$ relate cities that are
+%connected by direct flights either with American (AA) or United (UA).
+%The predicates $\ATtrain$ and $\UPtrain$ relate cities that are
+%connected by direct train trips either with Amtrak (AT) or Union
+%Pacific Railway (UP).
+The flight predicate relates $(x,y,t)$ whenever there is a direct flight
+from $x$ to $y$ operated by airline $t$, and similarly for $\train$,
+and $\UScity$ indicates when a given city is a US city or not.
+The ontology axioms capture two subsumption relations:
 \begin{alignat*}{1}
@@
 \begin{enumerate}[--]
-\item $\nationalFlight(x,y)$ relates two US cities that are connected by a direct flight,
-\item $\onewayFlight(x,y)$ relates two cities that are connected by a one-way flight,
+\item $\nationaltlight(x,y)$ relates two US cities that are connected by a direct flight,
+\item $\AAflight(x,y)$ relates cities that are connected by American flights,
+\item $\UAflight(x,y)$ relates cities that are connected by United flights,
+\item $\onewayflight(x,y)$ relates two cities that are connected by a one-way flight,
 \item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
 \item $\toPA(x)$ tells if there is a direct flight from $x$ to Paris,
-\item $\onestopPA(x,y)$ if there is a flight from $x$ to Paris with a stop at $y$, 
 \item $\fromLA(x)$ tells if there is a flight from Los Angeles to $x$,
-\item $\nationalTrain(x,y)$ relates US cities that are connected by a direct train, and
-\item $\onewayTrain(x,y)$ relates US cities that are connected by a one-way train.
+\item $\nationaltrain(x,y)$ relates US cities that are connected by a direct train,
+\item $\ATtrain(x,y)$ relates cities that are connected by Amtrak trains, and
+\item $\UPtrain(x,y)$ relates cities that are connected by Union Pacific Railway trains.
 \end{enumerate}
@@
 \begin{alignat*}{1}
-\nationalFlight(x,y)\ &\qrule\ \flight(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
-\AAflight(x,y)\       &\qrule\ \flight(x,y,\AA)\,, \\
-\UAflight(x,y)\       &\qrule\ \flight(x,y,\UA)\,, \\
+\nationaltlight(\$x,y)\ &\qrule\ \flight(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
+\AAflight(\$x,y)\       &\qrule\ \flight(x,y,\AA)\,, \\
+\UAflight(\$x,y)\       &\qrule\ \flight(x,y,\UA)\,, \\
+\onewayflight(x,y)\     &\qrule\ \flight(x,y,t)\,, \\
 \onestop(x,z)\          &\qrule\ \flight(x,y,t),\,\flight(y,z,t)\,, \\
-\toPA(x)\             &\qrule\ \flight(x,\PA,\AA)\,, \\
-%\onestopPA(x,y)\      &\qrule\ \flight(x,y),\,\flight(y,\PA)\,, \\
-\fromLA(x)\           &\qrule\ \flight(\LA,x,\UA)\,, \\
-\nationalTrain(x,y)\  &\qrule\ \train(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
-\ATtrain(x,y)\        &\qrule\ \train(x,y,\AT)\,, \\
-\UPtrain(x,y)\        &\qrule\ \train(x,y,\UP)\,.
+\toPA(\$x)\             &\qrule\ \flight(x,\PA,\AA)\,, \\
+\fromLA(\$x)\           &\qrule\ \flight(\LA,x,\UA)\,, \\
+\nationaltrain(\$x,y)\  &\qrule\ \train(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
+\ATtrain(\$x,y)\        &\qrule\ \train(x,y,\AT)\,, \\
+\UPtrain(\$x,y)\        &\qrule\ \train(x,y,\UP)\,.
 \end{alignat*}
-Observe that each tuple produced by each service satisfies the semantic description
-expressed by the body of the rule. For example, the tuples that satisfy the predicate
-$\nationalFlight(x,y)$ also meet the conjunctive formula:
+Observe that each tuple produced by each service satisfies the semantic
+description given in the body of the rule; e.g., the tuples that satisfy
+$\nationaltlight(x,y)$ meet the conjunctive formula
 \[ \exists t(\flight(x,y,t)\ \land\ \UScity(x)\ \land\ \UScity(y))\,. \]
 However, there may be tuples that satisfy this formula that are not
-produced by the service represented by $\nationalFlight(x,y)$.
-These elements make up the integration framework $IS$. In the rest of this section.
-We give examples of possible user requests and their solutions.
-
-Suppose that a user is interested in identifying the services able to retrieve
-one-stop round trips from a US city $x$ to any city $y$ in the world. Notice
-that the trip from $x$ to $y$ stops at a city $u$, that the back trip from $y$
-to $v$ stops at a city $v$, and that $u$ may not be equal to $v$.
-This request can be modeled as the following conjunctive query:
+produced by $\nationaltlight(x,y)$.
+
+Consider now a user that is interested on identifying the services able to
+retrieve one-stop round trips from a US city $x$ to any city $y$ in the world.
+Notice that the trip from $x$ to $y$ stops at a city $u$, that the back trip
+from $y$ to $v$ stops at a city $v$, and that $u$ may not be equal to $v$.
+This request can be modeled with the conjunctive query:
 \[ Q(x,u,y,v)\ \qrule\ \UScity(x),\,\trip(x,u),\,\trip(u,y),\,\trip(y,v),\,\trip(v,x)\,. \]
-Any rewriting of the ontology predicates in terms of the services correspond
-to a \emph{composition of services} that implements the user request.
+Any rewriting of the ontology predicates in terms of the services that
+respect the input/output constraints on the parameters correspond to a
+\emph{composition of services} that implements the request.
 For example, the following rewriting is a valid solution to the request:
 \begin{alignat*}{1}
-I(x,u,y,v)\ \qrule\ &\nationalFlight(x,u),\,\toPA(u),\, \\
-                    &\onewayFlight(\PA,v),\,\nationalFlight(v,x)\,. 
+I(x,u,y,v)\ \qrule\ &\nationaltlight(x,u),\,\toPA(u),\, \\
+                    &\onewayflight(\PA,v),\,\nationaltlight(v,x)\,. 
 \end{alignat*}
@@
 \begin{alignat*}{1}
- I'(x,u,y,v)\  &\qrule\ \nationalFlight(x,u),\,\toPA(u),\,\fromLA(v),\,\nationalFlight(v,x)\,, \\
-I''(x,u,y,v)\  &\qrule\ \onestop(x,y),\,\onewayFlight(y,v),\,\nationalFlight(v,x)\,.
+ I'(x,u,y,v)\  &\qrule\ \nationaltlight(x,u),\,\toPA(u),\,\fromLA(v),\,\nationaltlight(v,x)\,, \\
+I''(x,u,y,v)\  &\qrule\ \onestop(x,y),\,\onewayflight(y,v),\,\nationaltlight(v,x)\,.
 \end{alignat*}
@@
 
-As shown, one can use a system for rewriting queries in terms of views
-for computing solutions to the Service Selection Problem, as the valid
-solutions correspond to the valid rewritings of the query. However, in
-the presence of user preferences, the solutions must be ranked according
-to the preferences and the best solutions should be returned.
-
+As shown, one can use a system for rewriting queries in terms of views for
+computing solutions to the SSP, since the valid solutions correspond to the
+valid rewritings of the query.
+However, in the presence of user preferences, the solutions must be ranked
+according to the preferences and the best solutions should be returned.
 To illustrate the use of preferences, consider the following request:
-\[ Q(x,y)\ \qrule\ \trip(\LA,x),\,\trip(x,\NY),\,\trip(\NY,y),\,\trip(y,\LA) \]
+\[ Q(x,y)\ \qrule\ \trip(\LA,x),\,\trip(x,\NY),\,\trip(\NY,y),\,\trip(y,\LA). \]
 that looks for round-trips between Los Angeles and New York such that
@@
 a way that there are no restrictions whatsoever on the use of planes or
-trains for any leg of the trip. However, as is typical, users have 
-preferences about using planes or trains.
+trains for any leg of the trip. However, users typically have preferences
+about using planes or trains.
 For this example, we study four different scenarios for user's preferences
@@
 \item The user is indifferent with respect to trains or airplanes, yet
-      s/he does not want to mix both. This preference is an at-most-one preference
+      she does not want to mix both. This preference is an at-most-one preference
       over the set $\{\flight,\train\}$ that corresponds to the formula
@@
       preference.
-\item If the user travels by airplane, s/he prefers to always use the same
+\item If the user travels by airplane, she prefers to always use the same
       airline (independently of the airline). This preference can be modeled
@@
       `disabled' since they may return flights operated by any airline;
-      e.g., add the constraint $\neg\nationalFlight$ with a high cost.
-\item Finally, if the user travels by airplane, s/he prefers to use $\UA$.
-      This is a complex preference that can be modeled with the formula
+      e.g., add the constraint $\neg\nationaltlight$ with a high cost.
+\item Finally, if the user travels by airplane, she prefers to use $\UA$.
+      This is a non-trivial preference that can be modeled with the formula:
 \[ (\flight \Rightarrow \UAflight) \land (\neg\UAflight \lor \neg\AAflight)\,. \]
@@
 All these preferences correspond to formulas over the propositional language
-$\L(IS)$. The formulas for the last three preferences can be treated as
-hard constraints if the are associated with infinite cost, or as soft
-constraints meaning that the user prefers (but is not limited to) solutions
+$\L(IS)$. The formulas for all but the first case involve preferences that can be
+treated as hard constraints if they are associated with infinite cost, or soft
+constraints meaning that the user prefers, but is not limited to, solutions
 that do not violate the preferences.
 
-\Omit{
-
-Additionally, we allow users to specify their preferences. For example,
-a user could prefer a round trip by either train or plane but not a round
-trip that combines them. User preferences are specified as conditional
-rules on service predicates. Thus, this user preference is expressed
-with the following two rules:
-
-\begin{alignat*}{1}
-\onewayFlight(x,y)\ &\Rightarrow \neg \onewayTrain(y,x)\,. \\
-\onewayTrain(x,y)\ &\Rightarrow \neg \onewayFlight(y,x)\,. 
-\end{alignat*}
-
-Based on the integration system and the user preferences, the following
-composition corresponds to one rewriting of the user request in the
-available services; the aggregated value of this solution is 8.0.
-We use aggregated utility values and the user preferences to rank the
-solutions and select the best one.
-
-\begin{alignat*}{1}
- I(x,w,y,z)\ \qrule\  & \nationalFlight(x,w),\,\toPA(w),\,\\
-                                & \onewayFlight(\PA,z),\,\nationalFlight(z,x)\,. 
-\end{alignat*}
-
-Additionally, the integration system is used to avoid the enumeration of
-no valid solutions. For example, the following two rewritings are not valid:
-
-\begin{alignat*}{1}
-I'(x,w,y,z)\  \qrule\ &\nationalFlight(x,y),\,\toPA(y),\,\fromNY(z),\,\nationalFlight(z,x)\,. \\
-I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\onewayFlight(y,z),\,\nationalFlight(z,x)\,.
-\end{alignat*}
-
-The first rewriting is not valid because it maps the query variable $y$
-into constants \PA\ and \NY\ that denote different cities.
-On the other hand, the rewriting $I''$ does not implement the query because
-the service $\onestop(x,y)$ does not receive as input, or produce as output,
-the middle city where the flight stops and thus is not possible to ensure
-that this city is bound to the city $w$ that must returned by selected services.
-
-These examples show that a proper user request rewriting must handle
-constants such that not two different constants are mapped to each
-other either directly or indirectly via transitivity, and that all
-attributes that appear in a join or in the output need to be produced
-by the selected  services. In addition, user preferences and QoS parameters
-should be considered to discriminate among the valid solutions, the ones
-that do not satisfy the user preferences or have low values of the
-aggregated utility function. For example, our approach assigns a low score
-to the following rewriting:
-
-\begin{alignat*}{1}
- I'''(x,w,y,z)\ \qrule\ & \nationalFlight(x,w),\,\nationalFlight(w,y), \,\\
-& \onewayFlight(y,z), \,\onewayTrain(z,y), \,\nationalFlight(y,w)\,. 
-\end{alignat*}
-
-Formally, the Service Selection Problem (SSP) is defined as follows:
-given an integration system {\it IS=$<$G,S,M$>$}, a user request $Q$ and
-a set of user preferences $UP$, a valid solution $Q_S$ is a rewriting of
-$Q$ in the services in $S$, such that the result of evaluating $Q_S$ is
-contained in the result of evaluating $Q$ i.e., $Q_S$ is a maximally-contained
-rewriting of $Q$ ~\cite{RajaramanSU95,Ullman00}. Additional, the best solution
-is a rewriting  that minimizes (or maximizes) the aggregation of utilities
-and the user preferences.
-
-%We consider service catalogs of the form $C=\tup{P,T}$ where $P$ is a set of predicates in $\sigma_G$ and $T=\{T_p\}_{p\in P}$ is a collection of tables comprised of tuples that satisfy the ontology predicates.
-In the context of the SSP, a service catalog $C$ is an idealized description of the output produced by user request implemented by services or data sources described as views in $M$.
-A user request $Q$ over $S$ is a conjunctive query of the form  \[ Q(\vvec{x})\ \qrule\ s_1(\vvec{x}_1),\,s_1(\vvec{x}_2),\,\ldots,\,s_m(\vvec{x}_m)\,, \] where $s_i\in S$, $\vvec{x}$ is a vector of variables, and each $\vvec{x}_i$ is a vector of variables and constants.
-The result of $Q$ over $C$, denoted as $Q(C)$, is the table with $|\vvec{x}|$ columns that result of the projection of the relational join $\join\!\!\{T_{s_i}\}_{i=1}^n$ over $\vvec{x}$.
-The atoms in the body of $Q$ are called the (sub)goals of $Q$, and the variables in the head of $Q$ are called distinguished.
-
-%A concrete service is described as a view $V$ over $C$ that, following LAV, is a query over $S$.
-Given a catalog $C$, a workflow $W$ and a collection of $n$ views $E=\tup{\{V_i\}_{i=1}^n,\{E_i\}_{i=1}^n}$, we are asked to find all the tuples in $W(C)$ that can be obtained from the views in $E$.
-That is, we need to find all the
-%\emph{instantiations}
-%\[ R(\vvec{x})\ \qrule\ V_{i_1}(\vvec{x}_1),\,V_{i_2}(\vvec{x}_2),\,\ldots,\,V_{i_{m_i}}(\vvec{x}_{m_i}) \]
-%such that $R(E) \subseteq Q(C)$.
-%A Service Selection Problem is a tuple $\tup{S,Q,\{V_i\}_i}$ where $S$
-%is a set of predicates, $Q$ is a user request  over
-%$S$, and $\{V_i\}_i$ is a collection of views that define the services.
-%We assume \emph{safe} problems in the sense that all variables mentioned in the
-%head of the user request (resp.\ in the head of each view) appear in the body of the
-%user request (resp.\ in the body of each view).
-%Further, we only deal with SSPs with no arithmetic predicates inside the user request
-%or views. An instantiation $R$ is \emph{valid} if for all catalogs $C=\tup{S,T}$
-%and extensions $\{E_i\}_i$, $R(E)\subseteq Q(C)$. A collection $\R$ of valid
-%instantiations is a solution if for all service catalogs $C=\tup{S,T}$ and
-%extensions $\{E_i\}_i$, there is no $\R'$ such that $\R(E)\subset\R'(E)\subseteq Q(C)$.
-}
-
 \section{Solution and System Architecture}
 
-We extend the \mcdsat\ system of Arvelo et al.\ \cite{arvelo:aaai06} for
-Query Rewriting Problems (QRP). An instance of a QRP consists of a
-collection of views and a query on abstract concepts. 
-The problem consists is in ``re-writing'' the query in terms of the
-views so that each tuple produced by the rewriting is a tuple of the
-solution.
-\mcdsat\ reduces the QRP into the problem of finding the models of
-a propositional logic theory.
-The theory constructed by \mcdsat\ satisfies the following properties:
+We extend the \mcdsat\ system of Arvelo et al.\ \cite{arvelo:aaai06} for QRP.
+An instance of QRP consists of a collection of views and a query on abstract
+concepts.  The problem consists in rewriting the query in terms of the
+views such that each tuple produced by the rewriting is a tuple of the
+solution \cite{Ullman00}.
+\mcdsat\ reduces QRP to the problem of finding the models of a propositional
+logic theory that satisfies the following properties:
 (1)~there is a 1-1 correspondence between the valid rewritings of the
-query and the models of the logical theory, (2)~given a model of the theory,
-one can recover the corresponding rewriting in linear time, and
-(3)~the theory can be constructed in polynomial time from the instance 
-of the QRP.
-Once the logical theory is constructed, one can be interested in finding
-all minimal rewritings of the query as done in data integration systems
-with incomplete sources \cite{Ullman00}, or just a model of the theory
-\cite{refs}. For the former, off-the-shelf model enumeration solvers
-such as c2d \cite{c2d} and Relsat \cite{relsat} can be used, while 
-off-the-shelf SAT solvers such as Minisat \cite{minisat} can be used in the
-latter case.
+query and the models of the logical theory,
+(2)~given a model of the theory, one can recover the corresponding rewriting
+in linear time, and
+(3)~the theory can be constructed in polynomial time from the QRP instance.
+Once the logical theory is constructed, one can be interested in finding all
+minimal rewritings of the query as done in data integration systems with incomplete
+sources, or just one rewriting as done when sources are complete \cite{Ullman00}.
+For the former, off-the-shelf model enumeration tools such as c2d \cite{c2d} and
+Relsat \cite{relsat} can be used, while off-the-shelf SAT solvers such as
+Minisat \cite{minisat} or Rsat \cite{knot07lightweight} can be used in the latter case.
 
@@
 associated with SSPs that are not present in QRPs; namely, handling
-constant symbols and input and output attributes, the ontology of
-concepts with subsumption relationships, and user's preferences. The
-result is an extended theory whose models are in correspondence with
-the valid solutions of the SSP and, in the presence of preferences,
-whose \emph{best models} are in correspondence with the best-ranked
-valid solutions of the SSP.
-
-\subsubsection{Constant Symbols and Input Output Attributes}
-
-\mcdsat\ did not provide support for constant symbols in the language
-or service  input and output attributes, yet incorporating these
-functionality is straightforward.  First to consider constant symbols,
-basically, one only has to track the unification of variables with
-constants using new propositional symbols, and to propagate such
-unifications transitively using propositional implications to avoiding
-the unification of different constant symbols. Second, to respect the
-input and output attributes of the available services, one has to
-ensure that input attributes of a service unify with constant symbols
-in the user query or unify with output attributes returned by other
-services in the rewriting. These modifications involve the addition of
-a relatively small number of new propositional symbols and clauses to
-the CNF generated by \mcdsat.
+constant symbols, input and output attributes, the ontology of concepts
+with subsumption relationships, and user's preferences. The result is an
+extended theory whose models are in correspondence with the valid solutions
+of the SSP and, in the presence of preferences, whose \emph{best models}
+are in correspondence with the best-ranked valid solutions of the SSP.
+
+\subsubsection{Constant Symbols}
+
+\mcdsat\ does not provide support for constant symbols, yet incorporating
+this functionality is straightforward.
+Basically, one only has to track the unification of variables with
+constants using new propositional symbols, and to propagate such unifications
+transitively using implications in order to avoiding the unification of different
+constant symbols.
+This modification involve the addition of a small number of propositional
+symbols and clauses to the CNF generated by \mcdsat.
+
+\subsubsection{Input and Output Attributes}
+
+The general principle for properly handling input and output attributes 
+is that every input attribute of a service must unify either with a
+constant symbol or with an output attribute of another service,
+while avoiding the cycles in the dependencies among the services that
+`produce' (output) and `consume' (input) attributes.
+
+This principle can be enforced by adding propositional symbols to
+the theory of the form `$In(z,R)$' and `$Out(z,R)$', where $z$ is
+a variable and $R$ is a service, and symbols $Prec(R,S)$ for each pair
+of services $R$ and $S$.
+The intended interpretation for these symbols is that $In(z,R)$
+holds iff attribute $z$ is an input attribute of $R$, that
+$Out(z,R)$ holds iff attribute $z$ is an output attribute of $R$, 
+and that $Prec(R,S)$ holds iff the service $R$ produces an attribute
+that is consumed by $S$. Accordingly, the theory is extended with
+clauses that enforce this interpretation plus rules of the form
+$Prec(R,S) \land Prec(S,T)\Rightarrow Prec(R,T)$ that propagate
+the precedence relation via transitivity, and $\neg Prec(S,S)$
+that prunes rewritings containing cycles.
 
@@
 
-The subsumption relationships in the ontology make the rewriting 
-process more complex as now one need to consider unification of
-predicate symbols of different name and arity. Indeed, consider
-the following four ontology concepts, where $a$, $b$, $c$ and $d$
-are constant symbols, and $x$, $y$ and $z$ are variables:
+The subsumption relationships make the rewriting process more complex
+as now one need to consider unification among predicate symbols of different
+name and arity. Indeed, consider the following four concepts, where $a$,
+$b$, $c$ and $d$ are constant symbols, and $x$, $y$ and $z$ are variables:
 \begin{alignat*}{1}
+P(b,y,z)\ &\orule\ R(a,y)\,, \\
 R(a,y)\   &\orule\ T(c,y)\,, \\
-P(b,y,z)\ &\orule\ R(a,y)\,, \\
-M(a,x)\   &\orule\ N(d,x)\,, \\
-P(d,x,z)\ &\orule\ M(a,x)\,,
+P(d,x,z)\ &\orule\ M(a,x)\,, \\
+M(a,x)\   &\orule\ N(d,x)\,,
 \end{alignat*}
-and the following user request $Q(x,y)$ with the services 
-$S1$, $S2$ and $S3$:
+and the user request $Q(x,y)$ with services $S1$, $S2$ and $S3$:
 \begin{alignat*}{1}
@@
 and $S_2(x,z)$ unifies with $N(z,x)$ and becomes $S_2(x,c)$
-once the binding is propagated. On the other hand, the query
+once the binding is propagated.
+On the other hand, the system must also infer that $Q(x,y)$
 cannot be rewritten as $I(x,y)\qrule S_1(y),S_3(x,z)$ because
-$R(a,y)$ unifies with $T(z,y)$ with the binding $\{z=c\}$,
-$P(d,x,z)$ unifies with $N(z,x)$ with the binding $\{z=d\}$,
-and these bindings are non-unifiable since constants
-are assumed to denote unique objects.
+$R(a,y)$ unifies with $T(z,y)$ with binding $\{z=c\}$,
+$P(d,x,z)$ unifies with $N(z,x)$ with binding $\{z=d\}$,
+and these two bindings are non-unifiable since constants
+denote unique objects.
 
@@
 set of bindings is inconsistent, then the label is assigned the
-binding \textbf{false}).
+binding $\{\textbf{false}\}$).
 These labels are unique and well defined as $G(D)$ is assumed
@@
 to cover a predicate $R$ whenever there is an edge from $P$ to
-$R$ in $G(D)^*$, yet if this covering becomes active, the 
+$R$ in $G(D)^*$, and when this covering becomes active, the 
 bindings associated with it become active as well.
@@
 
-In order to account for the preferences, we shall use the
-concepts literal-ranking functions and best-ranked models for
-propositional logic. A literal ranking function $r$ is a function
-that assign ranks (real numbers) to propositional literals.
-Given a literal-ranking function $r$, the rank $r(\omega)$ of
-a model $\omega$ is the aggregation of the ranks for each literal
-made true by the model: $r(\omega)=\sum_{\omega\vDash\ell} r(\ell)$.
+To incorporate preferences, we use the concepts of literal-ranking
+function and best-ranked models for propositional logic.
+A literal ranking function $r$ is a function that assign ranks
+(real numbers) to literals. Given a literal-ranking function $r$,
+the rank $r(\omega)$ of a model $\omega$ is the aggregation of the
+ranks for each literal made true by the model; i.e.,
+$r(\omega)=\sum_{\omega\vDash\ell} r(\ell)$ \cite{darwiche:weighted}.
 Thus, the models can be ordered by their rank and the best-ranked
@@
 Likewise, Weighted-Max-SAT solvers such as MiniMaxSAT~\cite{HerasetalJAIR2008}
-can be used to find a best ranked model of a propositional theory.
+can be used to find a best ranked model.
 
 For SSPs, we accommodate the preferences by using a suitably defined
-literal-ranking function $r^*$ and by computing best-ranked models
-\cite{darwiche:weighted}.
+literal-ranking function $r^*$ and by computing best-ranked models.
 First, a new propositional variable is created for each relational
@@
 $r^*(\neg p_\pi)=c$ for each such preference.
-Clearly, the rank of a model correspond to the sum of the costs
+Clearly, the rank of a model corresponds to the sum of the costs
 associated with the preferences violated by the model, and thus
-a best-ranked model correspond to a rewriting of minimum regret.
+a best-ranked model corresponds to a rewriting of minimum regret.
 
@@
 We define an architecture for solving SSPs that is comprised of a Catalog of
-service descriptions, an ontology reasoner, the Encoder, the best model Finder,
+service descriptions, an Ontology Reasoner, the Encoder, the best model Finder,
 and the Decoder.
@@
 In this framework, an instance of SSP consists of an integration framework $IS$
-and a user request $R$ made of a conjunctive query $Q$ and a collection of
-preferences.
+and a user request $R$.
 The Catalog of the system is populated with the components of the integration system,
@@
 
-The input instance of the SSP is then translated into a CNF theory and a
-literal-ranking function $r^*$ by the Encoder module. The Encoder makes use
-of the transitive closure $G(D)^*$ that is calculated by the ontology reasoner
-together with the bindings associated with the edges. Once the theory is obtained,
-it is fed to a best-model finder. Once the finder return a best model, it is
-given to the Decoder that then re-construct the rewriting from the model.
+The input instance is then translated into a CNF theory and a literal-ranking
+function $r^*$ by the Encoder module. The Encoder makes use of the transitive
+closure $G(D)^*$ that is calculated by the Ontology Reasoner together with the
+bindings associated with the edges. Once the theory is obtained, it is fed to the
+Finder that returns a best model. The model is given to the Decoder that 
+reconstructs the solution to the input instance.
 
@@
 
-\Omit{
-Once the CNF is generated, the compiler c2d, an off-the-shelf
-component, compiles the CNF formula into d-DNNF.
-A d-DNNF  theory is constructed from literals using only conjunctions  and disjunctions \cite{barwise:handbook}; for each conjunction its variables are pairwise disjoint, and for each disjunction, the disjuncts are pairwise logically contradictory; see Fig.~\ref{fig:dnnf} for an example.
-In our approach, a d-DNNF theory is a compact representation of
-all the rewritings. That is, the Finder generates in a backtrack-free
-manner all the rewritings of the user request. If the user is interested in
-a best rewriting, then the Finder computes it in
-linear time in the size of the d-DNNF. If the user is interested in the 
-all the best rewritings, these are computed in linear time in the
-number of them. This computation transforms the DAG of the d-DNNF into a arithmetic
-circuit by replacing the AND nodes with `+' and the OR nodes with `min'.
-The literal ranking function assigns
-values to the leaves of the circuit that are propagated to the
-root in linear time. The value of the root is the rank of the
-best model \cite{darwiche:weighted}.
-Finally, if the user is interested in all rewritings,
-these are also computed in linear time in the number of them.
-In the latter two cases, if such number is exponential (in the size of the
-input), the enumeration of the instantiations is also exponential but
-this complexity is intrinsic to the problem and thus cannot be avoided.
-
-It is important to remark
-that the compilation process performed by the compiler c2d, needs to be performed only once as it does not
-depend on the value of the QoS parameters or the user preferences. Thus, even if the compilation happens to be costly in terms of time, this cost can be amortized since the resulting
-d-DNNF can be used to find best rewritings with respect to multiple values
-of the QoS parameters or weights of user preferences.
-Finally, the Decoder translates the best model returned by the Finder into
-a rewriting that solves the SSP.
-
-In order to make our results more accessible to a general audience, we decided
-to present the experimental results in the next section, and leave the related work, and formal
-and theoretical results for the following sections. In this way, the interested
-reader may skip the formal details of the approach on a first reading.
-}
-
-\section{Experimental Results}
-
-We conducted an empirical analysis on three benchmarks.
-All the experiments were run on a desktop machine with an Intel
-Core 2 Duo 2GHz CPU and 4Gb of memory, and the time was measured
-with the Unix time command. We used the
-compiler c2d, \footnote{\url{http://reasoning.cs.ucla.edu/c2d}} to compile 
-the CNF formula into d-DNNF\cite{darwiche:weighted}; c2d makes use of
-modern SAT techniques such as conflict-directed backtracking, clause learning
-and caching\cite{darwiche:compiler} to enumerate the theory models from the d-DNNF.
-
-The objective of the experiment is to assess the performance of the proposal
-on varying conditions. The main benefit of our approach is that one can compile
-the logical theory for a problem instance and then calculate all the instantiations,
-or the best ones, any number of times. The cost model for finding best instantiations
-can be changed with no need to recompile the logical theory. Therefore,
-the time complexity of our approach is basically the time to encode the SSP
-as a CNF plus the time to find the (best) models and the time to decode the
-models. The times to encode and decode are negligible compared to the time to
-enumerate the models. Because of this, we focus on the time to enumerate the
-models of the benchmark problems. All the CNF theories are of low width.
+\section{Preliminary Experiments}
+
+We have developed a system prototype that implements the above ideas 
+except for the support to distinguish input and output attributes of
+services (all attributes are assumed to be output), and with 
+partial support for handling preferences; a complete implementation
+is ongoing work.
+With this prototype, we conducted experiments on two type of domains:
+airline domains of the type seen before and random domains.
+The Finder module is built using c2d (\url{http://reasoning.cs.ucla.edu/c2d})
+that compiles (transforms) the CNF formula for the propositional
+theory into deterministic and decomposable negation normal form (d-DNNF)
+from which all models or just the best models can be efficiently
+enumerated in linear time \cite{darwiche:compiler}.
+
+The objective of the experiments is to test several features of the 
+approach and to see the scalability of the approach.
+The main benefit of the approach is
+that one can compile the logical theory for a problem instance and then
+calculate all the rewritings, or the best ones, any number of times,
+and the cost/rewards associated with the preferences can also be changed
+without the need to recompile the theory.
+Therefore, the time complexity of our approach is basically the time to
+compile the CNF theory into d-DNNF since calculating the CNF from the SSP
+and decoding the models is negligible.
+Thus, we only report the time to compile the CNF into d-DNNF.
+
+\subsection{Airline Experiments}
 
 The first benchmark consists of problems for air-travel queries.
-Service views are of the form $V_i(x,y)\qrule\ \flight(x,y,\AL_i)$
-where $\AL_i$ is a constant that denotes the name of an airline, 
-and $\flight(x,y,\AL_i)$ relates the cities $x$ and $y$ such that
-there is a flight between $x$ and $y$ served by $\AL_i$. This 
-service is assumed to return all flights between two cities with an
-specific airline. The user request has the form 
-\[ Q(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,z),\,\flight(x_1,x_2,z),\,\ldots,\,\flight(x_n,\NY,z)\,. \]
+Service views are of the form $V_i(x,y)\qrule \flight(x,y,\AL_i)$
+where $\AL_i$ is a constant that denotes the name of an airline
+and the view is assumed to return flights between two cities served
+by the airline $\AL_i$. For the query, we consider a request to
+find trips between Paris and New York with a number of stops.
+The query returns the stops and has the form
+\[ Q(x_1,\ldots,x_n)\ \qrule\ \flight(\PA,x_1,t),\,\flight(x_1,x_2,t),\,\ldots,\,\flight(x_n,\NY,t)\,. \]
+Observe that the existentially quantified variable, $t$, is the same
+for each flight meaning that it can only be unified with the same
+constant; i.e., each leg of the flight is served by the same airline.
+We solved several instances for this type of query with a number of stops
+from 2 to 5 stops and a number of services from 10 to 100.
+Fig.~\ref{fig:plot1}(a) shows the results of the compilations:
+the vertical axis refers to the time in seconds in logarithmic
+scale and the horizontal axis to the number of views in the benchmark.
+The results show good performance since realistic instances of the
+problem (sets of 100 airlines with 5-stop flights) can be compiled
+in 328 seconds. The size in disk of the d-DNNF for 100 airlines and
+5-stop flights is 3.4Mb from which the best model can be computed
+in 0.29 seconds, and the enumeration of all models can be done
+in 0.47 seconds.
+
+In the second experiment, we test our system with user's preferences.
+The query is the same except that the existentially quantified variables
+are all distinct for each flight meaning that any combination of airlines
+can fulfill the user request.
+As user's preferences, we consider the set
+$\{\neg V_i \lor \neg V_j : 1\leq i\neq j\leq n\}$ of $n(n-1)/2$ constraints
+each with cost $c_{i,j}$, for a problem with $n$ services.
+Thus, a best model is one that violates the minimum number of preferences
+and this is equivalent to using the same airline
+for each leg of the flight. Fig.\ref{fig:plot1}(b) shows the result for
+the compilation also in logarithmic scale.
+As it can be seen,  the compilation times are very similar for the
+Experiment~I where there are no preferences.
+The largest instance is a complex problem involving 100 views, 5 subgoals
+in the query and 4,950 user preferences; the total number of rewritings
+is $100^5$, yet it can be compiled in 600 seconds.
 
@@
 \centering
-\includegraphics[width=1\textwidth]{plots/plot1}
+\includegraphics[width=1\textwidth]{plots-v2/plot5}
 \caption{Compilation times for experiments I and II for different
-number of goals and different number of views. The plots are in
-logarithmic scale, and the time is in seconds.}
+number of goals and different number of views.
+Experiment II involves $n(n-1)/2$ preference of a problem with $n$ views.
+The plots are in logarithmic scale, and the time is in seconds.}
 \label{fig:plot1}
@@
 
-The benchmark includes user requests with 2 to 5 sub-goals and
-sets of 10 to 100 service views. The results for the compilation are
-shown in panel (a) of Fig.~\ref{fig:plot1}. This is a plot in logarithmic
-scale that suggests a sub-exponential behavior. In any case, the results
-show good performance since realistic instances of the problem (sets of 100
-airlines with 5-stop flights) can be compiled in 328 seconds.
-The size in disk of the d-DNNF for 100 airlines and 5-stop flights is 3.4Mb.
-On this d-DNNF, the best model can be computed in 0.29 seconds,
-and the enumeration of all models in 0.47 seconds.
-
-In an attempt to induce an exponential growth in the compilation time,
-in the second benchmark we add a second service view for each airline.
-This modification increases the number of valid rewritings from linear
-to exponential since each leg of a flight can now be instantiated by two
-services and thus a flight with $n$ legs may have up to $2^n$
-rewritings.
-We ran the compiler for instances comprising the same number of user request
-sub-goals and total number of service views. The results plotted in
-logarithmic scale are shown in panel (b) of Fig.~\ref{fig:plot1}.
-
-These tests show good performance for this type of problems, but they do not
-involve service views with multiple sub-goals. We therefore designed a
-third experiment that consists unstructured, randomly generated instances.
-Each instance contains three variables per domain ontology concept in the user request, ten distinct
-variables and ten distinct constants, six sub-goals in the user request,
-2 to 5 sub-goals in the service views, and a varying number of service views.
-The chance that an argument of a domain ontology concept is bound to a constant is 50\%.
-The results are depicted  in panel (a) of Fig.~\ref{fig:plot4}.
-The compilation time for these instances does not grow monotonically
-since they are randomly generated. The same happens for the size of
-the theories and the number of models. For example, the d-DNNF for 
-a problem with 45 views each with 5 sub-goals is of size 5.1Mb and has
-$1.26\times 10^8$ models. The time to find the best model for
-this d-DNNF is 0.46 seconds while the time to enumerate all models
-is about 17 hours.
+In the third experiment, we test the system with ontologies of different sizes.
+Ontologies corresponding to full binary trees of depth 2 to 7 were generated with
+the predicate $\trip(x,y)$ at the root node. Then, for each node in the tree,
+there is a view that is described by the predicate at that node.
+The user request has the form
+\[ Q(x_1,\ldots,x_4)\ \qrule\  \UScity(x_1),\,\trip(x_1,x_2),\,\trip(x_2,x_3),\,\trip(x_3,x_4),\,\trip(x_4,x_1).\]
+In all cases, the compilation time was always less than 13 seconds.
+
+\subsection{Random Experiments}
+
+For the last experiments, we generated random unstructured instances of SSPs
+as follows: each user request contains 6 subgoals, 10 distinct variables and
+10 distinct constant symbols, while each service view contains between 2 and
+5 subgoals. The constants were randomly placed on the subgoals arguments with
+a 50\% probability.
+Fig,~\ref{fig:random} shows the compilation time for these instances. 
+The size of the compiled theories and number of models does not increase
+monotonically with the number of views given the random nature of the
+instances. As it can be seen, these are complex instances and the approach
+is able to solve them in reasonable time.
 
@@
 \centering
-\includegraphics[width=.9\textwidth]{plots/plot4}
-\caption{Compilation times for experiment III for different number of goals
-and different number of views. Time for enumerating rewritings for experiment IV. The plots are in logarithmic scale, and the time
-is in seconds. }
-\label{fig:plot4}
+\includegraphics[width=.8\textwidth]{plots-v2/plot3}
+\caption{Compilation times for random experiments with different number of views.
+The plots are in logarithmic scale and the time is in seconds. }
+\label{fig:random}
 \end{figure}
 
-Finally, to study the effects of the ontology size, we considered the first benchmark and built 
-ontologies where the predicates $\flight(x,y,z)$ and $\UScity(x)$ were leaves; ontologies were binary trees comprised of 2, 3, 4, 5, 6 and 7 levels.  The user request has the form: 
-\[ Q(x_1,x_2,x_3,x_4)\ \qrule\  \UScity(x_1),\,\trip(x_1,x_2),\,\trip(x_2,x_3),\,\trip(x_3,x_4),\,\trip(x_4,x_1).\]
-where $trip(x,y)$ was the root of the ontologies.  The results are reported in panel 
-(b) of Fig.~\ref{fig:plot4}. The time to enumerate all the rewritings suggests a sub-exponential behavior, while the time to find the best models does not grow monotonically. On the other hand, 
-the size of the theories and the number of models grow exponentially.  For example, the d-DNNF for 
-this user request and the ontology with 7 levels is of size 1.7Mb and has
-$2.6\times 10^8$ models. The time to enumerate all models is about 147 hours.
-
-These are preliminary experiments, yet the results show that the
-proposed approach efficiently scales for problems with several goals
-and views. We believe that these results are encouraging and motivate
-us to continue this research. We plan to conduct additional experiments
-with other types and sizes of user requests and, when possible, compare with
-other approaches.
-
 \section{Related Work}
+
 The problem of selecting the services that satisfy a user request is a
@@
 In a series of papers, Berardi and
-others~\cite{berardi05,berardi08,berardi06}  describe services and
-user requests in terms of deterministic finite-state machines that are
-encoded using Description Logic theories whose models correspond to
-solutions of the problem; properties of Description Logics reasoning
-methods are not exploited.
-
-Ko et al.~\cite{myoung08} propose a
-constraint-based approach that encodes the nonfunctional permissible
+others~\cite{berardi05,berardi08,berardi06} describe services and user
+requests in terms of deterministic finite-state machines that are
+encoded using Description Logics theories whose models correspond to
+solutions of the problem, yet there are no efficient methods to compute
+these models as in the case of SAT.
+
+Ko et al.~\cite{myoung08} propose a constraint-based approach that encodes
+the nonfunctional permissible
 values as a set of constraints whose violation needs to
@@
 
-Recently, two
-planning-based approaches have been proposed. Kuter
+Recently, two planning-based approaches have been proposed. Kuter
 and Golbeck~\cite{kuterG09} extend the SHOP2 planning algorithm to
@@
 
-These existing solutions are able to scale up to relatively large
-number of abstract concepts. In addition to scalability, our approach
-provides a more expressive framework where services are semantically
+These existing solutions are scale to a number of abstract concepts.
+In addition to scalability, our approach provides a more expressive
+framework where services are semantically
 described in terms of domain ontology concepts, user preferences
@@
 }
-\section{Conclusions and Future Work}
-We have shown how the propositional theory used in \mcdsat\  for
-computing rewritings of queries can be extended to efficiently encode
-SSP; thus, by exploiting known properties of modern SAT solvers, we
-provide an efficient and scalable solution to SSP. In addition, we
-provide an expressive framework by defining LAV mappings to semantically
-describe services, and by expressing user requests as a conjunctive
-query over the domain ontology concepts and a set of preferences that
-restrict the SSP solutions.
-
-The experimental results show that the approach can be applied to
-real-sized problems; however, if d-DNNF format is used, the approach
-is only possible when the compilation of the CNF theory into d-DNNF
-format succeeds. Thus, to overcome this limitation, in the future, we
-plan to study the performance and scalability of our approach in
-other off-the-self model enumeration and SAT solvers such as Relsat,
-Minisat and MiniMaxSAT. We also are interested in using other target
-compilation languages like Ordered Binary Decision Diagrams.
+
+\section{Discussion}
+
+We proposed a novel formalism for expressing Service Selection Problems
+involving an ontology of generic concepts, services described using views
+in terms on the concepts, following the LAV approach, and user's preferences.
+This is a general, well-defined and scalable framework since it is based
+on logic, the LAV approach, and permits the modeling of real-life scenarios
+and preferences.
+
+We also showed how the propositional theory used in \mcdsat\ for solving
+the QRP can be extended to handle SSPs.
+This formulation allows to exploit the properties of modern SAT solvers
+to provide an efficient and scalable solution to the SSP.
+
+The preliminary experiments show that the approach can be applied to
+real-sized problems. We are currently working on a complete implementation
+of the formalism in order to offer its full expressivity.
+In the future, we plan to use other off-the-shelf SAT tools such 
+as MiniMaxSat that is able to find a best model without the need 
+to compile the CNF into d-DNNF.
 

### 2010-06-29 publicaciones/iswc10/paperISWC10-v3.tex
typos corregidos en paperISWC10-v3

@@
 Existing Web infrastructures support the publication and access to a
-tremendous amount of Web data sources, some of which that can be labeled
+tremendous amount of Web data sources, some of which can be labeled
 and converted into Semantic Web Services by using existing annotation
@@
 the set of valid rewritings of the posed query, in a way that
-the best solution to the SSP correspond to the best-ranked 
+the best solution to the SSP corresponds to the best-ranked 
 valid rewritings of the corresponding QRP.
@@
 best models are found using off-the-shelf SAT tools.
-Thus, we are able to exploit the befits of modern SAT techniques such
+Thus, we are able to exploit the benefits of modern SAT techniques such
 as conflict-directed backtracking via clause learning and caching and
@@
 ontology. The symbol `\$' denotes that $x$ is an input attribute. 
-The semantic interpretation of a mapping like this one enforce the following:
+The semantic interpretation of a mapping like this one enforces the following:
 \begin{itemize}
@@
 constant symbols.
-This modification involve the addition of a small number of propositional
+This modification involves the addition of a small number of propositional
 symbols and clauses to the CNF generated by \mcdsat.
@@
 The subsumption relationships make the rewriting process more complex
-as now one need to consider unification among predicate symbols of different
+as now one needs to consider unification among predicate symbols of different
 name and arity. Indeed, consider the following four concepts, where $a$,
@@
 We solved several instances for this type of query with a number of stops
-from 2 to 5 stops and a number of services from 10 to 100.
+from 2 to 5 and a number of services from 10 to 100.
 Fig.~\ref{fig:plot1}(a) shows the results of the compilations:
@@
 number of goals and different number of views.
-Experiment II involves $n(n-1)/2$ preference of a problem with $n$ views.
+Experiment II involves $n(n-1)/2$ preferences of a problem with $n$ views.
 The plots are in logarithmic scale, and the time is in seconds.}
@@
 
-These existing solutions are scale to a number of abstract concepts.
+These existing solutions scale to a number of abstract concepts.
 In addition to scalability, our approach provides a more expressive

### 2010-06-29 publicaciones/iswc10/paperISWC10-v3.tex
dos typos mas en paperISWC10-v3

@@
 adoption of a simple yet expressive language for preferences. 
-These preferences and constraints further refine and rank the
+These preferences and constraints further refine and rank 
 the set of valid rewritings of the posed query, in a way that
@@
 real-sized problems. We are currently working on a complete implementation
-of the formalism in order to offer its full expressivity.
+of the formalism in order to offer its full expressiveness.
 In the future, we plan to use other off-the-shelf SAT tools such 

### 2010-06-29 publicaciones/iswc10/paperISWC10-v4.tex
paperISWC10-v4

--- 2010-06-29 publicaciones/iswc10/paperISWC10-v3.tex
@@
 efficient and scalable approaches to solve the service selection problem are
-required in order to make this services widely available to the users.
+required in order to make these services widely available to the users.
 In this paper, we propose such a framework that is grounded on logic and
@@
 flights originating at a given US city can be described with the view:
-\[ S(\$x,y)\ \qrule\ \flight(x,y), \UScity(x)\,, \]
+\[ S(\$x,y)\ \qrule\ \flight(x,y), \UScity(x)\,. \]
 where $flight^2$ and $uscity^1$ are relational symbols in the
@@
 there is a view that is described by the predicate at that node.
-The user request has the form
+The user request has the form:
 \[ Q(x_1,\ldots,x_4)\ \qrule\  \UScity(x_1),\,\trip(x_1,x_2),\,\trip(x_2,x_3),\,\trip(x_3,x_4),\,\trip(x_4,x_1).\]

### 2010-08-23 publicaciones/iswc10_demo/paperdemo.tex
demo de ISWC2010

 \documentclass{llncs}
 \usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{url}

\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\vvec}[1]{\mathbf{#1}}
\newcommand{\join}{\bowtie}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\qrule}{:\!\!-}
\newcommand{\orule}{\sqsubseteq}
\newcommand{\mcdsat}{\textsc{McdSat}}
\newcommand{\minicon}{{MiniCon}}
\newcommand{\Omit}[1]{}

% flights
\newcommand{\flight}{\text{\it flight}}
\newcommand{\UScity}{\text{\it uscity}}
\newcommand{\national}{\text{\it national}}

\newcommand{\trip}{\textit{trip}}
\newcommand{\planetrip}{\textit{plane-trip}}
\newcommand{\traintrip}{\textit{train-trip}}
\newcommand{\AAflight}{\textit{AA-flight}}
\newcommand{\UAflight}{\textit{UA-flight}}
\newcommand{\ATtrain}{\textit{AT-train}}
\newcommand{\UPtrain}{\textit{UP-train}}
\newcommand{\train}{\textit{train}}

% Constants
\renewcommand{\AA}{\text{AA}}
\newcommand{\UA}{\text{UA}}
\newcommand{\AT}{\text{AT}}
\newcommand{\UP}{\text{UP}}
\newcommand{\PA}{\text{Paris}}
\newcommand{\NY}{\text{NY}}
\newcommand{\LA}{\text{LA}}
\newcommand{\AL}{\text{AL}}

% Services
\newcommand{\nationaltlight}{\textit{nat-flight}}
\newcommand{\onewayflight}{\textit{one-way-flight}}
\newcommand{\nationaltrain}{\textit{nat-train}}
\newcommand{\onewayTrain}{\textit{one-way-train}}
\newcommand{\onestop}{\textit{one-stop}}
\newcommand{\toPA}{\textit{to-pa}}
\newcommand{\onestopPA}{\textit{onestop-to-pa}}
\newcommand{\fromNY}{\textit{from-ny}}
\newcommand{\fromLA}{\textit{from-la}}

\begin{document}
\allowdisplaybreaks
\title{STEREO: a SaT-based tool for an optimal solution of the sERvice selEctiOn problem}
\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
\institute{Departamento de Computaci\'on \\
           Universidad Sim\'on Bol\'{\i}var \\
           Caracas 89000, Venezuela \\
           \url{{idaniel,mvidal,bonet}@ldc.usb.ve}}
\maketitle

\begin{abstract}
We present STEREO, a system that offers an expressive formalism and implements techniques firmly grounded on logic, to solve the Service Selection Problem (SSP).
STEREO adopts the Local-As-View approach to represent service's functionality as 
views of ontology concepts, while user requests are expressed as conjunctive queries on these concepts. Additionally,  users can describe their preferences, which  are  used to rank the solutions. 
We discuss the LAV formulation  of SSP; then, we illustrate the encoding of SSP  as a logical theory  whose models are in correspondence with the problem solutions, and in presence of preferences, whose best models are in correspondence with the best-ranked solutions.  We demonstrate how by exploiting properties of  modern SAT solvers, STEREO provides an efficient and scalable solution  to  SSP. 
\end{abstract}

\section{Introduction}
Nowadays, several annotation tools like the one proposed by Ambite et al.\ \cite{AmbiteISWC09}, are able to labeled and convert existing Web data sources into Semantic Web Services.
Once these tremendous among of services become available, users
will require approaches  more than ever, to precisely express their selection needs and to effectively identify the services that best meet their requirements.
In order to achieve this goal, we developed  STEREO, a system that offers an expressive formalism for describing Web services and user requests,  and implements a logic-based solution to efficiently select the services that best meet the user requirements. STEREO formalism and the techniques have been reported in  \cite{IVB2010}.

 STEREO is tailored to constantly changing Web service datasets and a relatively stable set of ontology concepts. STEREO adopts  the recent approach of Ambite et al.\ \cite{AmbiteISWC09}
that describes services as views on ontology concepts following the Local-As-View (LAV) approach that is widely used in integration
systems~\cite{levy:bucket}. Thus, every time that a service changes or
a new one becomes available, only a tiny fraction of the mappings
must be updated. Additionally,   STEREO models user requests  as conjunctive queries on the ontology concepts and casts  SSP  as the problem
of rewriting a query in terms of a set of views, the so-called Query
Rewriting Problem (QRP). Thus, STEREO exploits existing scalable and efficient techniques that have been proposed in the data integration area~\cite{arvelo:aaai06,levy:bucket,pottinger:minicon}.
Furthermore, STEREO offers a simple yet expressive language for preferences and supports users' preferences and constraints on the set of the
selected services to serve a given request. 
These preferences and constraints further refine and rank 
the set of valid rewritings of the posed query, in a way that
the best solution to the SSP corresponds to the best-ranked 
valid rewritings of the corresponding QRP. STEREO  constructs a propositional logical theory  that captures 
the features associated with SSPs; each model of the theory encodes a valid combination of services, and these models are ordered by their rank and the best-ranked
models are the models with minimum rank. Knowledge encoded in the ontology is used to extend the rules in the logical theory to permit the cover of ontology symbols in
the query with symbols in the views according to subsumption relationships represented in the ontology. 
We will demonstrate the benefits of the propose approach and show the following key issues: (a) we illustrate the expressiveness of the approach; we discuss scalability  by demonstrating how STEREO is able to enumerate realistic instances of SSP in few seconds; and finally, we demonstrate the performance  by enumerating in the same logical theory, the optimal models based on different cost/rewards of the users' preferences.  The demo is published at \url{www.ldc.usb.ve/STEREO}.

\section{STEREO Architecture}
\begin{figure}[t]
\centering
\includegraphics[height=30mm,width=.6\textwidth]{architecture.pdf}
\caption{The STEREO Architecture}
\label{fig:architecture}
\end{figure}

 The STEREO architecture  is comprised of a Catalog of service descriptions,
 an Ontology Reasoner, the Encoder, the best model Finder,
and the Decoder.
Figure~\ref{fig:architecture} depicts the overall architecture. In STEREO, an instance of SSP consists of an integration framework $IS=\tup{D,S,M}$ where $D$ is an ontology, $S$ is the set
of services, and $M$ is the set LAV mappings. A user request is a tuple $R=\tup{Q,P}$ comprised of a query $Q$  and a set of preferences $P$. STEREO Catalog is populated with the components of the integration framework. The Ontology Reasoner computes the transitive
closure of the ontology subsumption relationships, and the Encoder module makes used of this information in conjunction with the LAV mappings,  constant symbols,   and user's preferences to encode  an input instance of SSP as a CNF theory.
STEREO uses  c2d (\url{http://reasoning.cs.ucla.edu/c2d}) to compute all the
best-ranked models of a propositional theory  in a certain 
normal form called deterministic and decomposable negation normal
form (d-DNNF) \cite{darwiche:d-dnnfs}; however, the formalization can be also used in Weighted-Max-SAT solvers.
 The Finder computes a best model by enumerating its models  and it can reuse the same d-DNNF for different cost/rewards associated with the preferences to calculate  the best service rewritings. 
   Finally, the Decoder translates the model(s)  into the input instance.

\section{Demonstration of Use Cases}
In this demonstration, we illustrate our approach on  a simple travel-information system that contains information about
flight and train trips between cities and information about which cities
are in the US. The ontology is comprised of the predicates
$\trip$, $\flight$, $\train$ and $\UScity$.
The first predicate relates cities $(x,y)$ if there is a direct
trip either by plane or train between them. The flight predicate relates $(x,y,t)$ whenever there is a direct flight from $x$ to $y$ operated by airline $t$, and similarly for $\train$,
and $\UScity$ indicates if a given city is or not a US city. 
The ontology axioms capture two subsumption relations:
\begin{alignat*}{1}
\flight(x,y,t)\  \orule\ \trip(x,y).\  & \train(x,y,t)\  \orule\ \trip(x,y).
\end{alignat*}
For the services, we assume the following type of data sources:
\begin{enumerate}[--]
\item $\nationaltlight(x,y)$ relates two US cities that are connected by a direct flight,
\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
\item $\nationaltrain(x,y)$ relates US cities that are connected by a direct train,
\item $\toPA(x)$ tells if there is a direct flight from $x$ to Paris,
\item $\fromLA(x)$ tells if there is a flight from Los Angeles to $x$.
\end{enumerate}

Based on the ontology concepts the services are described using LAV views:
\begin{alignat*}{1}
\nationaltlight(x,y)\ &\qrule\ \flight(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
\toPA(x)\             &\qrule\ \flight(x,\PA,\AA)\,, \\
\fromLA(x)\           &\qrule\ \flight(\LA,x,\UA)\,, \\
\onestop(x,z)\          &\qrule\ \flight(x,y,t),\,\flight(y,z,t)\,, \\
\nationaltrain(x,y)\  &\qrule\ \train(x,y,t),\,\UScity(x),\,\UScity(y)\,. 
\end{alignat*}

Consider a user that needs to select the services able to
retrieve one-stop round trips from a US city $x$ to any city $y$ in the world and the user prefers to fly rather than to travel by train. This request is represented as follows:
\[ Q(x,y)\ \qrule\ \UScity(x),\,\trip(x,u),\,\trip(u,y),\,\trip(y,v),\,\trip(v,x)\,. \]

 This preference is modeled by assigning a high reward to the symbol $\flight$. Any rewriting of the ontology predicates in terms of the services defined by these predicates, implements the request.
In addition, rewritings comprised of services defined in terms of the symbol $\flight$ will have a higher scores than those with services on the symbol $\train$.
Thus, this rewriting is  valid and  highly ranked:
\begin{alignat*}{1}
I(x,\PA)\ \qrule\ &\nationaltlight(x,u),\,\toPA(u),\, \onewayflight(\PA,v),\,\nationaltlight(v,x)\,. 
\end{alignat*}
But, the following rewritings is not a valid solution because it maps the query variable $y$ into two different constants \PA\ and \LA\ that denote different cities:
\begin{alignat*}{1}
 I'(x,y)\  &\qrule\ \nationaltlight(x,u),\,\toPA(u),\,\fromLA(v),\,\nationaltlight(v,x)\,.
\end{alignat*}

Using this travel domain, we  consider the following scenarios:
\begin{itemize}
\item We run STEREO in three different benchmarks.   We start with a benchmark of  queries with 2
to 5 sub-goals and sets of 10 to 100 services; we show that the sets of 100
airlines with 5-stop flights can be compiled in 328 secs, 
the best model can be computed in 0.29 secs, and the enumeration of all models in 0.47 secs. We add a second  service for each airline and show that STEREO behavior remains similar. Finally, we  try services with multiple sub-goals which are randomly generated and show that the compilation time  does not grow monotonically with the number of views.  The time to find the best model 
is 0.46 secs while the enumeration of all models is about 17 hours.
\item We test the system with ontologies of different sizes. Ontologies corresponding to full binary trees of depth 2 to 7  with the predicate $\trip(x,y)$ at the root node. We show that in all cases, the compilation time  is  around 13 secs.
\item To show the reusability, we assign different cost/reward to the  preferences,  
 and demonstrate that the same d-DNNF can be used to compute the best model, while the execution time remains almost the same. 
\end{itemize}

\section{Conclusions}
In this demonstration, we will present a logical-based approach that relies on the Local-As-View (LAV) formalization \cite{levy:bucket} to solve the SSP. We will show how by using simple conjunctive rules, the functionality of services  can be defined. In addition, we will demonstrate how a propositional logic-based approach can be used to represent instances of SSP, and different scenarios where the properties STEREO will be observed by  users.  We will demonstrate that the approach can be applied to
real-sized problems, while the whole approach is only possible
when the compilation of the CNF theory into d-DNNF  succeeds.
\bibliographystyle{abbrv}
\bibliography{ref}
 \end{document}

### 2010-08-24 publicaciones/iswc10_demo/paperdemo.tex
errores de ortografia del paper del demo

@@
 \begin{abstract}
-We present STEREO, a system that offers an expressive formalism and implements techniques firmly grounded on logic, to solve the Service Selection Problem (SSP).
-STEREO adopts the Local-As-View approach to represent service's functionality as 
-views of ontology concepts, while user requests are expressed as conjunctive queries on these concepts. Additionally,  users can describe their preferences, which  are  used to rank the solutions. 
+We present STEREO, a system that offers an expressive formalism and implements techniques firmly grounded on logic to solve the Service Selection Problem (SSP).
+STEREO adopts the Local-As-View approach to represent services' functionality as 
+views on ontology concepts, while user requests are expressed as conjunctive queries on these concepts. Additionally,  users can describe their preferences, which  are  used to rank the solutions. 
 We discuss the LAV formulation  of SSP; then, we illustrate the encoding of SSP  as a logical theory  whose models are in correspondence with the problem solutions, and in presence of preferences, whose best models are in correspondence with the best-ranked solutions.
@@
 \section{Introduction}
-Nowadays, several annotation tools like the one proposed by Ambite et al.\ \cite{AmbiteISWC09}, are able to labeled and convert existing Web data sources into Semantic Web Services.
-Once these tremendous among of services become available, users
-will require approaches  more than ever, to precisely express their selection needs and to effectively identify the services that best meet their requirements.
+Nowadays, several annotation tools like the one proposed by Ambite et al.\ \cite{AmbiteISWC09}, are able to label and convert existing Web data sources into Semantic Web Services.
+Once this tremendous amount of services becomes available, users
+will more than ever require approaches to precisely express their selection needs and to effectively identify the services that best meet their requirements.
 In order to achieve this goal, we developed  STEREO, a system that offers an expressive formalism for describing Web services and user requests,  and implements a logic-based solution to efficiently select the services that best meet the user requirements.
-STEREO formalism and the techniques have been reported in  \cite{IVB2010}.
+The STEREO formalism and the techniques have been reported in  \cite{IVB2010}.
 
@@
 the query with symbols in the views according to subsumption relationships represented in the ontology. 
-We will demonstrate the benefits of the propose approach and show the following key issues: (a) we illustrate the expressiveness of the approach; we discuss scalability  by demonstrating how STEREO is able to enumerate realistic instances of SSP in few seconds; and finally, we demonstrate the performance  by enumerating in the same logical theory, the optimal models based on different cost/rewards of the users' preferences.
+We will demonstrate the benefits of the proposed approach and show the following key issues: (a) we illustrate the expressiveness of the approach; we discuss scalability  by demonstrating how STEREO is able to enumerate realistic instances of SSP in a few seconds; and finally, we demonstrate the performance  by enumerating in the same logical theory, the optimal models based on different cost/rewards of the users' preferences.
 The demo is published at \url{www.ldc.usb.ve/STEREO}.
@@
 A user request is a tuple $R=\tup{Q,P}$ comprised of a query $Q$  and a set of preferences $P$.
-STEREO Catalog is populated with the components of the integration framework.
+The STEREO Catalog is populated with the components of the integration framework.
 The Ontology Reasoner computes the transitive

### 2010-08-24 publicaciones/iswc10_demo/paperdemo.tex
correcciones de ma esther

@@
 We present STEREO, a system that offers an expressive formalism and implements techniques firmly grounded on logic to solve the Service Selection Problem (SSP).
-STEREO adopts the Local-As-View approach to represent services' functionality as 
+STEREO adopts the Local-As-View approach (LAV) to represent services' functionality as 
 views on ontology concepts, while user requests are expressed as conjunctive queries on these concepts. Additionally,  users can describe their preferences, which  are  used to rank the solutions. 
-We discuss the LAV formulation  of SSP; then, we illustrate the encoding of SSP  as a logical theory  whose models are in correspondence with the problem solutions, and in presence of preferences, whose best models are in correspondence with the best-ranked solutions.
+We discuss the LAV formulation  of SSP; then, we illustrate the encoding of SSP  as a logical theory  whose models are in correspondence with the problem solutions, and in presence of preferences, the best models are in correspondence with the best-ranked solutions.
 We demonstrate how by exploiting properties of  modern SAT solvers, STEREO provides an efficient and scalable solution  to  SSP.
@@
 the query with symbols in the views according to subsumption relationships represented in the ontology. 
-We will demonstrate the benefits of the proposed approach and show the following key issues: (a) we illustrate the expressiveness of the approach; we discuss scalability  by demonstrating how STEREO is able to enumerate realistic instances of SSP in a few seconds; and finally, we demonstrate the performance  by enumerating in the same logical theory, the optimal models based on different cost/rewards of the users' preferences.
+We will demonstrate the benefits of the proposed approach and show the following key issues:   the expressiveness of the approach; the scalability  by demonstrating how STEREO is able to enumerate realistic instances of SSP in a few seconds; and finally, the performance  by enumerating in the same logical theory, the optimal models based on different cost/rewards of the users' preferences.
 The demo is published at \url{www.ldc.usb.ve/STEREO}.
@@
 Figure~\ref{fig:architecture} depicts the overall architecture. In STEREO, an instance of SSP consists of an integration framework $IS=\tup{D,S,M}$ where $D$ is an ontology, $S$ is the set
-of services, and $M$ is the set LAV mappings.
+of services, and $M$ is the set of LAV mappings.
 A user request is a tuple $R=\tup{Q,P}$ comprised of a query $Q$  and a set of preferences $P$.
@@
 The Ontology Reasoner computes the transitive
-closure of the ontology subsumption relationships, and the Encoder module makes used of this information in conjunction with the LAV mappings,  constant symbols,   and user's preferences to encode  an input instance of SSP as a CNF theory.
+closure of the ontology subsumption relationships, and the Encoder module makes used of this information in conjunction with the LAV mappings,  constant symbols,   and users' preferences to encode  an input instance of SSP as a CNF theory.
 STEREO uses  c2d (\url{http://reasoning.cs.ucla.edu/c2d}) to compute all the
@@
 form (d-DNNF) \cite{darwiche:d-dnnfs}; however, the formalization can be also used in Weighted-Max-SAT solvers.
- The Finder computes a best model by enumerating its models  and it can reuse the same d-DNNF for different cost/rewards associated with the preferences to calculate  the best service rewritings. 
+ The Finder computes a best model by enumerating the models  and it can reuse the same d-DNNF for different cost/rewards associated with the preferences to calculate  the best service rewritings. 
    Finally, the Decoder translates the model(s)  into the input instance.
@@
 \section{Demonstration of Use Cases}
-In this demonstration, we illustrate our approach on  a simple travel-information system that contains information about
+In this demonstration, we illustrate our approach on  a simple travel domain that contains information about
 flight and train trips between cities and information about which cities
@@
 The first predicate relates cities $(x,y)$ if there is a direct
-trip either by plane or train between them. The flight predicate relates $(x,y,t)$ whenever there is a direct flight from $x$ to $y$ operated by airline $t$, and similarly for $\train$,
+trip either by plane or train between them. The $\flight$ predicate relates $(x,y,t)$ whenever there is a direct flight from $x$ to $y$ operated by airline $t$, and similarly for $\train$,
 and $\UScity$ indicates if a given city is or not a US city. 
@@
 Consider a user that needs to select the services able to
-retrieve one-stop round trips from a US city $x$ to any city $y$ in the world and the user prefers to fly rather than to travel by train. This request is represented as follows:
+retrieve one-stop round trips from a US city $x$ to any city $y$ in the world, and the user prefers  flying than traveling by train. This request is represented as follows:
 \[ Q(x,y)\ \qrule\ \UScity(x),\,\trip(x,u),\,\trip(u,y),\,\trip(y,v),\,\trip(v,x)\,. \]
@@
  This preference is modeled by assigning a high reward to the symbol $\flight$. Any rewriting of the ontology predicates in terms of the services defined by these predicates, implements the request.
-In addition, rewritings comprised of services defined in terms of the symbol $\flight$ will have a higher scores than those with services on the symbol $\train$.
+In addition, rewritings comprised of services defined by the symbol $\flight$ will have higher scores than those with services defined by the symbol $\train$.
 Thus, this rewriting is  valid and  highly ranked:
@@
 \end{alignat*}
-But, the following rewritings is not a valid solution because it maps the query variable $y$ into two different constants \PA\ and \LA\ that denote different cities:
+But, the following rewriting is not a valid solution because it maps the query variable $y$ into two different constants \PA\ and \LA\ that denote different cities:
 \begin{alignat*}{1}
@@
 \item We test the system with ontologies of different sizes.
-Ontologies corresponding to full binary trees of depth 2 to 7  with the predicate $\trip(x,y)$ at the root node.
+The ontologies correspond to full binary trees of depth 2 to 7  with the predicate $\trip(x,y)$ at the root node.
 We show that in all cases, the compilation time  is  around 13 secs.
-\item To show the reusability, we assign different cost/reward to the  preferences,  
+\item We show reusability by assigning different cost/reward to the  preferences,  
  and demonstrate that the same d-DNNF can be used to compute the best model, while the execution time remains almost the same. 
@@
 \section{Conclusions}
-In this demonstration, we will present a logical-based approach that relies on the Local-As-View (LAV) formalization \cite{levy:bucket} to solve the SSP.
-We will show how by using simple conjunctive rules, the functionality of services  can be defined.
-In addition, we will demonstrate how a propositional logic-based approach can be used to represent instances of SSP, and different scenarios where the properties STEREO will be observed by  users.
-We will demonstrate that the approach can be applied to
+We present a logical-based approach that relies on the Local-As-View (LAV) formalization \cite{levy:bucket} to solve SSP.
+We show how by using simple conjunctive rules, the functionality of services  can be defined.
+In addition, we demonstrate how a propositional logic-based approach can be used to represent instances of SSP, and different scenarios where the properties STEREO can be observed by  users.
+We demonstrate that the approach can be applied to
 real-sized problems, while the whole approach is only possible
